### 2021-12-02 inst/MyArticle/MergePoster.Rmd
changes needed because prolfqua uses LazyLoad False.

---
title: prolfqua - R-package for Analysing Proteomics Label Free Quantification Experiments
author: 
  - name: Witold Wolski
    affil: 1; 2
    orcid: '0000-0002-6468-120X'
  - name: Paolo Nanni
    affil: 1
    orcid: '0000-0001-8429-3557'
  - name: Jonas Grossmann
    affil: 1; 2
    orcid: '0000-0002-6899-9020'
  - name: Maria D'Errico
    affil: 1; 2
    orcid: '0000-0001-9467-9058'
  - name: Ralph Schlapbach
    affil: 1
    orcid: '0000-0002-7488-4262'
  - name: Christian Panse
    affil: 1; 2
    orcid: '0000-0003-1975-3064'
affiliation:
  - num: 1
    address: Functional Genomics Center Zurich (https://www.fgcz.ch/)
  - num: 2
    address: Swiss Institute of Bioinformatics (https://www.sib.swiss/)
column_numbers: 5
poster_height: "30in"
poster_width: "58in"
title_textsize:	"90pt"
font_family: Monaco
primary_colour: "#000C66"
titletext_fontfamily: Arial
body_textsize: "35px"
logoleft_name: C:/Users/wolski/__checkout/prolfqua/inst/Figures/FGCZLogo.png
logoright_name: C:/Users/wolski/__checkout/prolfqua/inst/Figures/hexStickerProlfqua.png
output: 
  bookdown::html_document2:
    toc: true
  posterdown::posterdown_html:
bibliography: bibliography.bib
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(cowplot)
library(ggplot2)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 7, fig.height = 5)

```

See ACS guidelines [https://publish.acs.org/publish/author_guidelines?coden=ancham#manuscript_text_components]

# Abstract (80–250 words)

[Copied this paragraph from a different article. Needs to be adjusted for prolfqua and this article]
"Protein mass spectrometry with label-free quantification is widely used for quantitative proteomics studies. Nevertheless, well-principled statistical inference procedures are still lacking, and most practitioners adopt methods from transcriptomics. These, however, cannot properly treat the principal complication of label-free proteomics, namely many non-randomly missing values." (rewrite this sentence)

The _prolfqua_ package integrates essential steps of the mass spectrometry based label free quantification data analysis workflow: quality control, data normalization, protein aggregation, modeling, hypothesis testing and sample size estimation. We use _prolfqua_ to develop highly customizable, visually appealing, and interactive data analysis reports in pdf or HTML format.  We use _prolfqua_ to visualize and model simple experimental designs with a single explanatory variable and complex experiments with multiple factors. Furthermore, we use _prolfqua_ to benchmark data acquisition, data preprocessing or data modeling methods. We developed and improved the package by applying the "Eating your own dog food" principle, making it easy to use.

# Introduction

[Jonas/Paolo could you provide some references here [...] doi's will do.]
Proteins are carriers of function and structure in living cells. Hence, measuring changes in protein abundance is the subject of active research [....]. However, identifying and quantifying intact proteins in complex samples containing hundreds of proteins is not possible using existing methods at the moment. Hence, the bottom-up mass spectrometric method is employed [...]. Proteins that are extracted from biological samples are specifically and reproducible digested into protein fragments - peptides. The peptides are first separated by their chemical and physical properties using High-Pressure Liquid Chromatography (HPLC). Afterward, they are weighted, identified, and quantified using mass spectrometric techniques, e.g., Electro-Spray-Ionization Mass Spectrometry (ESI-MS). Finally, peptide identification is achieved by fragmenting and matching the measured fragment masses to theoretical masses computed from known peptide sequences[...]. For quantification, intact peptide ions [DDA ref - method and sofware] or products of peptide ion fragmentation [DIA ref method and software] are counted and aggregated. Finally, the identified and quantified peptides are assigned to proteins based on protein sequence information [...prot inference ref].

Label-Free Quantification (LFQ) Proteomics Experiments enable monitoring relative abundances of thousands of proteins in biological samples. For most studies, parallel designs with a single explanatory variable are used, e.g., treatment vs. control. More recently, more complex experimental designs are implemented with increasing numbers of samples, i.e., factorial designs and longitudinal studies (time series), sometimes with repeated measurements on the same subject [...]. The obtained data can be modeled using linear fixed, mixed, or random-effects models [@Bates2015FittingJSS]. Based on these models, tests can be applied to examine whether a factor or interactions are significant, or differences between factor levels can be estimated, and their significance tested.  Proteins can be ordered or grouped using the estimated fold-change, t-statistics, scaled p-value, or false discovery rate (FDR) and then subjected to gene set enrichment or over-representation analysis [...] or validation experiments.

Several packages exist to model mass spectrometry LFQ experiments, e.g. _limma_[@Ritchie2015], _MSStats_[@Choi2014], _ROPECA_[@Suomi2017bEnhanced],  _MSqRob2_ [@Goeminne2016] or _proDA_ [@bioxrvproDA2020], to mention some, all implemented in the R. Each of them has some unique features. For example, _MSStats_ infers the model from the sample annotation and generates the model formula. At the same time, _limma_ allows to specify a linear model formula and implements the empirical Bayes variance shrinkage method. The R-package _ROPECA_, performs a roll-up of peptide level fold change and p-value estimates, obtained from _limma_, to protein level estimates. Furthermore, _MSqRob2_ implements two models: robust linear models fitted to protein intensities and robust ridge regression fitted to peptide intensities and combines them with empirical Baye's variance shrinkage known from _limma_. The _proDA_ package implements a linear probabilistic dropout model to account for missing data without imputation.

When analyzing parallel designs using a single explanatory variable, we can use all packages for the statistical data analysis of the data; however, if we want to model factorial designs or repeated measurements, we can use only some of them. Table \ref@(tab:tableOverview) gives an overview of the models and features supported by these packages.

```{r tableOverview}
xx <- list(
  triqler = c("pd" = "Y"),
  MSstats = c("pd" = "Y", "rm" = "Y", "mem" = "y"),
  ROPECA = c("pd" = "Y", "rm" = "Y*", fd = "Y", int = "Y", eb = "Y"),
  limma  = c("pd" = "Y", "rm" = "Y*", fd = "Y", int = "Y",eb = "Y"),
  MSqRob2_rlm = c("pd" = "Y", "rm" = "Y*", fd = "Y", int = "Y",eb = "Y", md = "?"),
  proDA = c("pd" = "Y", "rm" = "Y*", fd = "Y", int = "Y",eb = "Y", md = "Y"),
  prolfqua = c("pd" = "Y", "rm" = "Y", fd = "Y", int = "Y", mem = "Y",eb = "Y", md = "Y"))

bb <- data.frame(bind_rows(xx))
rownames(bb) <- names(xx)
knitr::kable(data.frame(bb), caption = "Rows - R package, Columns - types of models supported: pd - parallel design, rm - repeated measurements, fd - factorial design, int - interactions among factors, mem - mixed effect models, eb - empirical Bayes, md - missing data modelling. Y - yes, y - yes but only for random effects, * - repeated measurements are only modeled using a fixed effects. ? - the hurdle model is not implemented in the msqrob2 package yet.")
```

We were inspired by the excellent R package caret, which enables to call various machine learning methods, making selecting the best ML algorithm for your problem easy. We wanted to have a similar tool for the analysis of LFQ data. However, after examining the R packages for modeling proteomics quantification data, we observed that model specification, input, and output formats differ widely and wildly. However, all of them fit some type of linear model either to peptide or protein intensities, determine contrasts among conditions, and afterward, some of them apply empirical Bayes variance shrinkage. Therefore, the revised goal was to provide a modular design, with R linear models as a core, where features such as p-value aggregation (ROPECA) or variance shrinkage could be easily enabled if required.

Models in _prolfqua_ are specified using R’s linear model and mixed model formula interface. Therefore, knowledge of the R regression model infrastructure is an advantage when using our package. Furthermore, this glass box approach should make it easy to reimplement an analysis performed with _prolfqua_ using base R or other programming languages by reading the analysis script and without looking at the package code. However, acknowledging the R formula interface’s complexity to non-statisticians and the popularity of MSstats, we provide the functionality to suggest a model formula from the sample annotation.

An important aspect of mass spectrometric data are missing values. Missing values lead to groups of varying size. Linear and mixed effect models can handle unbalanced designs, i.e. different group sizes. As long as there is at least one observation per group they do produce unbiased estimates. We use this feature instead of imputing missing observations. However, if there are not observation in one of the groups, then contrasts involving this group can not be estimated. To adres it, for the group without observations, we use a point prior equal to the average of $10\%$ smallest group mean intensities in the dataset and infer the variance from the variance in other groups. A way to adress this is to annotate the produced models with information about missigness which is done by for instance MSStats. How missigness can modelled using the Hurdle model is described here, however to our knowledge only the R package _proDA_ models missingness directly.

Furthermore, the functionality of _prolfqua_ also includes methods specific to proteomics data. For example, we implemented strategies to infer protein intensities from peptide intensities: top N, Tukey's median polish, robust linear models. Furthermore, peptide or protein intensities can be scaled using the robust scaling method or transformed using _quantile_ normalization or _vsn_ from Bioconductor to remove systematic differences among samples and heteroscedasticity. In this respect, _prolfqua_ can be compared with other R packages such _DEP_ [@DEP2018] or POMA [@POMA2021] which support the entire LFQ analysis pipeline.

We use the Ionstar [@shen2018ionstar] dataset to benchmark the analysis pipeline and modeling methods implemented within _prolfqua_ and to compare our results with those of _MSstats_ and _proDA_.

# Implementation & Methods (Experimental Section)

## Implementation

We store all the data needed for analysis in a single data frame in a tidy table, i.e., every column is a variable, every row is an observation, every cell is a single value[@tidydata2014]. Using an __R6__[@R6cite] configuration object, we specify what variable is in which column, making it easy to integrate new inputs in _prolfqua_ if provided in tidy tables. For example, to visualize tidy _Spectronaut_, or _Skyline_ outputs, or data in _MSStats_ format, only a few lines of code to update the _prolfqua_ configuration are needed. For popular software like _MaxQuant_ or _MSFragger_, which stores the same variable (e.g., intensity) in multiple columns, one for each sample, we implemented methods that transform the data into tidy tables. Relying on the tidy data table enabled us to easily interface with many data manipulation, visualization, and modeling methods implemented in base __R__[@RCoreTeam2021] and the tidyverse[@tidyverse2019]. We use __R6__ classes to structure the functionality of the package (see Figure \@ref(fig:LFQData) and Figure \@ref(fig:ContrastUML)). __R6__ classes are well supported (e.g. auto-completion) and easy to use (OOP).

```{r LFQData, echo=FALSE, fig.cap="Class Diagram for LFQData and related objects. The LFQData_Plotter class uses the LFQData class to implement methods for plotting.", out.width = '90%'}
fig_svg <- cowplot::ggdraw() + cowplot::draw_image("LFQDataUML.svg")
plot(fig_svg)

```

__R__ linear model and linear mixed effect models allow modeling parallel designs, repeated measurements, factorial designs, and many more. __R__'s formula interface for linear models is flexible, widely used, and well documented. This makes it easy to reproduce an analysis performed with _prolfqua_ in other statistical programming languages. We implemented features specific to high throughput experiments, such as the experimental Bayes variance and p-value moderation, which utilizes the parallel structure of the protein measurements and the analysis [@Ritchie2015]. We also compute probabilities of differential protein regulation based on peptide level models [@Suomi2017bEnhanced]. We use R6 to model the statistical modeling functionality in `proflqua` (see Figure \@ref(fig:ContrastUML)).

(ref:ContrastUML) UML diagram of modeling and contrast related classes. Different strategies e.g. _lm_, _lmer_, and _glm_ can be used to fit various types of models using a model builder (see code). All classes computing contrasts implement the _Contrast_ interface.

```{r ContrastUML, echo=FALSE, fig.cap="(ref:ContrastUML)", out.width = '90%'}
fig_svg <- cowplot::ggdraw() + cowplot::draw_image("ContrastClassesUML.svg")
plot(fig_svg)
```

## Benchmark dataset

Ionstar dataset, give Pride ids, describe MaxQuant settings and version. Give reference to Experiment hub and the _prolfquaData_ package.

## Modelling

### Estimating contrasts for linear models

Given a linear model contrasts can be computed by $\hat{\beta_{c}} = \sum l\beta_{m}$ and $\textrm{var} \hat\beta_c = l\sigma^2 (X^T X)^{-1} l^T$, with $X$ being the design matrix, $\beta_m$ the model parameters and $l$ an array of coefficients.  The degrees of freedom for the contrast are equal to the residual degrees of freedom of the linear model.
For estimating contrasts from mixed effects models we used the function `contest` implemented in the R package `lmerTest` [@Kuznetsova2017lmerTest] and used the Satterthwaite method to estimate the denominator degrees of freedom. These methods are implemented in the class `Contrast` (see Figure \@ref(fig:ContrastUML))

### Contrast estimation in presence of missing data.

If there are no observation in a group we assume that there are no observations because of the limit of detection (LOD). Furthermore, we assume that the variance of the protein is constant in all the conditions.
We impute the mean estimate of the group by using the average of $10\%$ smallest group mean intensities in the experiment. To estimate the variance in the condition, we use the pooled variance based on all conditions:

$$s_p^2=\frac{\sum_{i=1}^k (n_i - 1)s_i^2}{\sum_{i=1}^k (n_i - 1)},$$
with $n_i$ the number of observations, and $s_i$ the standard deviation in each condition.
The standard deviation for the t-statics is then given by:

$$
s = \sqrt{2}\sqrt{\frac{n_g s_p^2}{n}},
$$

Where, $n_G$ is the number of groups and $n$ is the number of observations.

This methods are implemented in the class `ContrastSimpleImpute` (see Figure \@ref(fig:ContrastUML)).

### P-value moderation

From the linear and the mixed effect models, we can obtain the residual standard deviation $\sigma$, and degrees of freedom.  [@Smyth2004linear] discuss how, using the empirical Bayes paradigm, to use the $\sigma$ and $df$ of all models to estimate a prior $\sigma$ and prior degrees of freedom, and posterior $\tilde \sigma$. These can be used to moderate the t-statistics by:
$\tilde{t}_{pj} = \frac{t_{pj} s_p}{\tilde{s}_p}$
and subsequently the p-values. We implemented this method in the class `ContrastModerated` (Figure \@ref(fig:ContrastUML)).

### Summarizing peptide level models

To summarize peptide level models to protein models we did applied the method suggested by [@Suomi2017bEnhanced], that is to use the median scaled p-value of the peptide models and cumulative distribution function of the beta distribution function to determine a regulation probability of the protein.

To obtain the median p-value of the protein, we first rescaled the peptide p-values by taking the direction of the fold-change $\hat \beta$ into account, i.e.:

\begin{equation}
p_{s} =
  \begin{cases}
1-p, & \textrm{if}~ \hat{\beta} > 0\\
p-1, & \textrm{otherwise}
\end{cases}
(\#eq:scalepvalue)
\end{equation}

Afterwards, the median scaled p-value $\tilde{p}_s$ is determined and using the transformation below, transformed back onto the original scale.

$$
\tilde{p} = 1 - |\tilde{p}_{s}|
$$
Because we used the median as the i-th order statistic  $i = n/2 +0.5$. Therefore, $\gamma = i = n/2 + 0.5$ and $\delta = n - i + 1 = n - (n/2 + 0.5) + 1 = n/2 + 0.5 = \gamma$ are used to parameterize the CDF of the Beta distribution. We implemented this method in the class `ContrastROPECA` (Figure \@ref(fig:ContrastUML)).

# Example Analysis

Peptide intensities are $\log_2$ transformed and robust z-score scaled but then rescaled to preserves the original range of the data using the method `robscale`. We show the distribution of the normalized intensities in Figure \@ref(fig:prepro) Panel A. Protein intensities are estimated using Tukey's median polish. Figure \@ref(fig:prepro) Panel B shows the peptide intensities and the estimated protein intensities in black. The boxplot (Panel D) shows the distribution of the protein intensity estimate in each condition. Finally, Panel C shows the distribution of the protein standard deviations for each condition.

```{r prepro,  echo = TRUE , fig.cap="(ref:scaling)", out.width = '90%', fig.width = 8, fig.height = 8}
library(prolfqua)
d <- prolfqua_data('data_ionstar')$filtered()
lfqd <- LFQData$new(d$data, d$config) 
t <- lfqd$get_Transformer()
lfqd <-  t$log2()$robscale()$lfq
pl <- lfqd$get_Plotter()
p_a <- pl$intensity_distribution_density() +
  labs(tag = "A") + theme(legend.position = "none")
agr <- lfqd$get_Aggregator()
lfqp <- agr$medpolish()
p_b <- agr$plot()$plots[[1]] + labs(tag = "B")
p_c <- lfqp$get_Stats()$violin() + labs(tag = "C")
pl <- lfqp$get_Plotter()
p_d <- pl$boxplots()$boxplot[[1]] + labs(tag = "D")
ggpubr::ggarrange(p_a, p_b, p_c, p_d)

```

(ref:scaling) Panel A - Peptide intensity distributions for 20 samples, Panel B - Peptide intensities for protein _5NTD_ and protein intensity estimate (black dashed line), Panel C - distribution of protein standard deviations for all dilutions, Panel D - Distribution of protein intensities for protein _5NTD_.

The following code example illustrates how we compute contrast comparing the protein intensities in the condition. First, the differences and the linear model are specified. Afterward, the model is fitted to the data using the `build_model` function. Next, we examine the importance of the factors in the model using the ANOVA function, which facilitates model selection for models with one explanatory variable. Finally, we estimate the contrasts either from the linear model using the `Contrasts` class or directly from the data using the `ContrastsSimpleImpute` class. Afterward, we apply t-statistic moderation using the `ContrastModerated` class. Finally, the `addContrastResults` function merges both sets of contrast estimates, preferring the one obtained from the linear model. Figure \@ref(fig:exampleContrasts) Panel A shows the distribution of the p-values, while figure Panel B shows the volcano-plot for each comparison.

(ref:exampleContrasts) Left panel - histogram of p-values. Right panel - volcano plots.

```{r exampleContrasts, fig.cap="(ref:exampleContrasts)", echo=TRUE, out.width = '90%'}
contrasts <- c(
  "dilution_(9/7.5)_1.2" =   "dilution.e - dilution.d",
  "dilution_(7.5/6)_1.25" =   "dilution.d - dilution.c"
  #,"dilution_(6/4.5)_1.3(3)" =   "dilution.c - dilution.b",
  #,"dilution_(4.5/3)_1.5" =   "dilution.b - dilution.a"
)
# fit model
lmmodel <- paste(lfqp$intensity_column()," ~ dilution.")
modelFunction <- strategy_lm( lmmodel, model_name = "lm")
models <- build_model(lfqp, modelFunction)
p0 <- models$anova_histogram()
# compute contrasts from linear model and with imputation
contr <- prolfqua::Contrasts$new(models, contrasts) %>% 
  prolfqua::ContrastsModerated$new()
conI <- prolfqua::ContrastsSimpleImpute$new( lfqp, contrasts) %>% 
  prolfqua::ContrastsModerated$new()
tmp <- prolfqua::addContrastResults(prefer = contr, add = conI)
# visualize results
pl <- tmp$merged$get_Plotter()
p1 <- pl$histogram()$p.value
p2 <- pl$volcano()$FDR +
   theme(legend.position = "bottom")
gridExtra::grid.arrange(p1, p2, ncol = 2)

```

# Benchmarking

The peptide intensities from the _peptide.txt_ file were $log_2$ transformed, and subsequent robust z-score transformation, where median and the mean average deviation (mad) were obtained from the human proteins only. We removed _one hit wonders_, i.e. proteins with a single peptide assignment. No further data filtering was used. We also did not use imputation. The protein level linear model was fitted to the protein intensities inferred from the peptide intensities using the Tukey's median polish. The mixed effect linear model, and the ROPECA model were fitted to the peptide intensities. The details of the _prolfqua_ benchmark implementation are available online  [BenchmarkingIonstarData](https://wolski.github.io/prolfqua/articles/BenchmarkingIonstarData.html).

For benchmarking of MSstats we used the defaults specified in the [Benchmarking_MSstats](https://wolski.github.io/prolfqua/articles/)

The Benchmark functionality of _prolfqua_ includes ROC curves and computes partial areas under those curves (pAUC) and other scores, e.g. the false discovery proportion FDP. We use it to study how well linear, mixed effect models or p-value moderation model quantitative mass spectrometric high throughput experiments and compare it with results produced by MSstats [@MSstats2014] and by proDA [@bioxrvproDA2020]. Figure \@ref(fig:benchmarkROC)

The IonStar  [@shen2018ionstar] benchmark dataset contains _H. sapiens_ proteins with constant concentrations and _E. coli_ proteins with varying concentrations. We know that for _H. sapiens_ proteins the difference $\beta$ between two dilutions should be $\beta = 0$ while for _E. coli_ proteins, we know that the difference between dilutions should be $\beta \ne 0$. To compare the performance of the various methods implemented in _prolfqua_ we use only the contrasts resulting in small fold-changes $\beta = 1.2,1.25,1.3(3),1.5$.

We can use various statistics to examine the alternative hypothesis $\beta \ne 0$: the contrast estimate, i.e. the $\log_2$ fold-change $\beta$, the t-statistic $\frac{\beta}{\sqrt{var(\beta)}}$, or the $p-value$ and moderated $p-value$. For each statistic and each value of the statistics we then compute a confusion matrix (see Table \@ref(tab:confusionMatrix)). From the confusion matrix we obtain measures such as true positive rate (TPR), false positive rate (FPR), or false discovery proportion (FDP) which are given by:

```{r confusionMatrix}
table <- data.frame( c("beta != 0", "beta == 0", "Total"), matrix(c("TP","FP","R","FN","TN","","P","N","m"), ncol = 3, byrow = T))
colnames(table) <- c("Prediction \\ Truth","E.coli", "H.sapiens", "Total")
knitr::kable(table, caption = "Confusion matrix, TP - true positive, FP - false positive, FN - false negative, P - all positive cases (all E. coli proteins), N - all negative cases (all H. sapiens proteins), m- all proteins.")

```

$$
\begin{aligned}
TPR &= \frac{TP}{TP+FN} = \frac{TP}{P}\\
FPR &= \frac{FP}{FP+TN} = \frac{FP}{N}\\
FDP &= \frac{FP}{TP + FP} = \frac{FP}{R}
\end{aligned}
$$

By plotting the $TPR$ versus the $FPR$ we obtain the receiver operator characteristic curve (ROC curve). The area under the curve (AUC) or partial areas under the curve (pAUC), at various values of the $FPR$, are measures of performance derived from the ROC curve. By using these measures we can compare the performances of the statistics produced by the various methods examined.

```{r}

allBenchmarks <- readRDS("../../inst/Benchresults/allBenchmarks.RDS")
benchmark_msstats <- readRDS("../../inst/Benchresults/benchmark_msstats.RDS")
benchmark_prodA <- readRDS("../../inst/Benchresults/benchmark_medpolish_proDA.RDS")
msFragger <- readRDS(file = "../../inst/Benchresults/MSFragger_medpol_benchmark.RDS")

allBenchmarks$benchmark_mssstats <- benchmark_msstats
allBenchmarks$benchmark_msFragger <- msFragger
allBenchmarks$benchmark_proDA <- benchmark_prodA
allBenchmarks <- allBenchmarks[c("benchmark_imputation","benchmark_ProtModerated",  "benchmark_mixedModerated", "benchmark_ropeca","benchmark_merged","benchmark_mssstats","benchmark_proDA"   )]

```

(ref:benchmarkROC) Left panel - Partial area under the ROC curve at $10\%$ FPR for all contrasts. Red line average area under the curve. Left panel - Difference to mean partial area under the ROC curve for various models, at $10\%$ FPR, by fold change.

```{r benchmarkROC, fig.cap="(ref:benchmarkROC)", out.width = '90%', fig.width=12, fig.height=10}

ttt <- sapply(allBenchmarks, function(x){x$complete(FALSE)})
res <- map_df(allBenchmarks, function(x){x$pAUC()})
res <- res %>% mutate(whatfix = case_when(what == "scaled.beta.based.significance" ~ "scaled.p.value", TRUE ~ what))

norm <- res %>% group_by(contrast,whatfix) %>% summarize(meanpAUC_10 = mean(pAUC_10))
res <- inner_join(res, norm)
res <- mutate(res , pAUC_10n = pAUC_10 - meanpAUC_10)

resAllB <- res %>% dplyr::filter(contrast == "all")

p1 <- ggplot(resAllB, aes(x = Name, y = pAUC_10)) +
  geom_bar(stat = "identity") +
  facet_wrap(~whatfix)  + 
  coord_cartesian(ylim = c(min(resAllB$pAUC_10),max(resAllB$pAUC_10))) + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = -90, vjust = 0.5)) +
  geom_hline(aes(yintercept = meanpAUC_10), color = "red") + xlab("")

p2 <- ggplot(res, aes(x = contrast, y = pAUC_10n, group = Name)) +
  geom_line(stat = "identity", aes(linetype = Name, color = Name), size = 1) + 
  facet_wrap(~ whatfix, scales = "free") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = -90, vjust = 0.5)) +
  geom_hline(aes(yintercept = 0), color = "red") + 
   theme(legend.position = "bottom", legend.title = element_blank())

gridExtra::grid.arrange(p1, p2, ncol = 2)

```

A relevant parameter is the number of contracts a given method estimates from the data(see Figure \@ref(fig:FDRfdp) left Panel). It indicates how robust the procedure works in the presence of missing data. For each protein, we can estimate four contrast which for 4046 proteins, at least two peptides resulted in $16184$ possible contrast estimates. Because the set of contrasts slightly differs for each method, it makes comparing them complicated. For instance, we observe that MSstas estimates $16058$ while the mixed effect models estimate the fewest with $15940$. Hence, to conclude that one method shows a better performance, the score, e.g., pAUC, needs to be greater, but also the number of estimated contrasts needs to be equal or larger.

Figure \@ref(fig:benchmarkROC) left panel, shows how various estimates obtained from the models, i.e. $\log_2(FC)$, $t-statistics$ and scaled p-values allow to identify true positives (TP) given a false positive rate. When testing for differential expression, the $\log_2(FC)$ is outperformed by than the t-statistic or p-value. 
If an accurate estimate of the $log_2(FC)$ is essential, the linear models fitted to protein intensities, estimated using Tukey's median polish, show the best performance. We speculate that this is because of the robustness of Tukey's median polish method. This hypothesis could be examined, by including other forms of protein intensity inferences implemented in `prolfqua`, e.g. `top-N` or `rlm`, into the benchmark.

There are only minor differences in the p-AUC when using the t-statistics or the scaled p-value  (see Figure \@ref(fig:benchmarkROC) right panel). In order to compute the confusion matrices based on the p-value we first need to rescale it (see Equation \@ref(eq:scalepvalue)). Thus, the p-value is a function of the t-statistics and the degrees of freedom. Interestingly, the pAUC increases when using p-values instead of the t-statistics for linear models, while it decreases for mixed effect models. The reason is erroneous denominator degrees of freedom estimation for many proteins in the mixed-effect models.

We also benchmark if the $FDR$ obtained from a model is an unbiased estimate of the false discovery proportion $FDP$. Figure \@ref(fig:FDRfdp) right panel draws on the horizontal axis the FDR determined from the model, and on the vertical axis, the FDP obtained from the confusion matrix.  We see that most of the lines are below the diagonal, indicating that the FDR estimates obtained with _prolfqua_ are modestly conservative for this benchmark dataset provide. In the case of _MSstats_, we observe a high proportion of false discoveries for small $FDR$ values. In the case of _ROPECA_, the FDR estimates, obtained when Benjamini-Hochberg correcting the regulation probabilities, strongly overestimate the $FDP$.

(ref:FDRfdp) Left panel - Number and percentage of estimated contrasts by modeling method. Right panel - Compare FDR estimate with false discovery proportion (FDP).

```{r FDRfdp, fig.cap = "(ref:FDRfdp)", out.width = '90%', fig.width=12, fig.height=10}

dd <- map_df(allBenchmarks, function(x){res <- x$smc$summary; res$name <- x$model_name;res})
dd
dd <- dd %>% mutate(nrcontrasts = protein_Id * (4 - nr_missing))
dd
dds <- dd %>% group_by(name) %>% summarize(nrcontrasts = sum(nrcontrasts))
dds
dds$percent <- dds$nrcontrasts/max(dds$nrcontrasts) * 100

nrgg <- dds %>% ggplot(aes(x = name, y = nrcontrasts )) + 
  geom_bar(stat = "identity", fill="white", colour = "black") + 
  coord_cartesian(ylim = c(min(dds$nrcontrasts) - 100, max(dds$nrcontrasts) + 10)) +
  theme(axis.text.x = element_text(angle = -90, vjust = 0.5)) +
  geom_text(aes(label = round(nrcontrasts, digits = 1)),
            vjust = 1, hjust = -0.2, angle = -90) + 
  geom_text(aes(label = paste0("(",round(percent, digits = 1),"%)")),
            vjust = -1, hjust = -0.2, angle = -90) 
p3 <- nrgg

dd <- map_df(allBenchmarks, function(x){res <- x$get_confusion_FDRvsFDP(); res$name <- x$model_name;res})

ddb <- filter(dd, contrast == "dilution_(4.5/3)_1.5")
ddb <- dd %>% dplyr::filter(contrast == "dilution_(7.5/6)_1.25")
ddb <- dd %>% dplyr::filter(contrast == "all")

p4 <- ddb %>% ggplot(aes(y = FDP_,  x  = scorecol )) + 
  geom_line(aes(color = model_name, linetype = model_name)) +
  facet_wrap(~contrast) + 
   geom_abline(intercept = 0, slope = 1, color = 2) + 
   theme(legend.position = "bottom")

gridExtra::grid.arrange(p3, p4, ncol = 2)

```

# Conclusions

We focused on comparing models to estimate fold changes, variances and to obtain false discovery rates. However, an essential parameter, not discussed further, is the normalization of the intensities within the samples to make them comparable. For this benchmark dataset, the fraction of regulated proteins (ECOLI) is relatively large. Therefore, to calibrate the measurement, we used human proteins only. In practice, it is unknown which proteins do and which don't change. Therefore, the assumption used when normalizing, scaling the data, and thus calibrating the measurement is that only a small subset of the proteins in the dataset is differentially regulated.

The method used to infer protein intensities from more peptide intensities is an additional important factor. The original _proDA_ publication uses MaxLFQ [@cox2014accurate] protein estimates. When using those, _proDA_ performs significantly worse than _proDA_ when using the protein estimates obtained from peptide intensity estimates using Tukey's median polish,  $pAUC_10(t-statistics) = 66%$ and $pAUC_10(t-statistics) = 73%$ respectively (see Articles on [@prolfquaGithubIO] for more details).

We assessed the performance of different modeling approaches implemented in prolfqua, MSstats and proDA, using a benchmark dataset whose ground truth was known. However, the dataset used has several limitations: first, the design is a parallel-group design, i.e. we could not benchmark how well interactions are modelled; secondly, essential sources of variation present in every experiment, such as the biochemical and biological, are not measured. The dataset captures only the variance from the chromatography, electro-spray, and mass spectrometric measurement method. Thus, while we can extrapolate some of the results obtained to more realistic datasets, we need to be careful not to over-interpret the results.

We can conclude that if we want to order the proteins according to the likelihood of being differentially regulated, for instance, to perform gene enrichment analysis, the t-statistic is better suited than the fold-change estimate. Modeling the degrees of freedom when computing the p-values might improve the inference further. However, the observed change is minuscule.

For the fixed effect linear model, the empirical Bayes variance shrinkage, as first suggested by [@Smyth2004linear], gives a consistent improvement compared with the unmoderated estimates (data not shown). However, there is no such effect for the mixed effect model, most likely because the degrees of freedom are erroneously estimated for a large proportion of models.

Computing the statistics at the peptide level, e.g., the t-statistics or p-value, then summarizing these statistics using their median produces the highest AUC scores among all the tested models. By using the beta distribution to model the number of peptides observed, we can further improve the pAUC scores (see Figure). However, the properties of beta-based probabilities are not well understood; for instance, they have no uniform distribution under the null hypothesis. The FDR estimates obtained when correcting multiple testing with the Benjamini-Hochberg method are strongly biased (see Figure \@ref(fig:FDRfdp) right panel). Therefore, we can not recommend this method if an unbiased estimate of FDR is essential, which is frequently the case. Furthermore, peptides are stronger affected by missing values, reducing the number of contrasts we could estimate for the dataset (see Figure \@ref(fig:FDRfdp) left panel). To adress the issue of missing observation we could either run proDA or the imputation of group averages method discussed here and afterwards apply ROPECA.

The probabilistic dropout analysis implemented in the _proDA_ produces inferences comparable to other methods (Figure \@ref(fig:benchmarkROC) for the largest number of proteins (Figure \@ref(fig:FDRfdp) left panel). Moreover, the estimated fold changes are unbiased and show high diagnostic accuracy. Furthermore, the scaled p-values and t-statics are on par with those obtained by the linear model with variance moderation. Therefore, we plan to integrate the _proDA_ package as an additional modeling option into _prolfqua_ (see Figure \@ref(fig:ContrastUML)).

Leveraging on these computer experiments, we can provide the following advice. First, aggregate the peptide intensities to protein intensities using a robust regression method. Second, use linear models as long as they suffice to model your experimental design. Use mixed effect models only if you have to (e.g., repeated measurements). If you use fixed-effect linear models, apply variance moderation since this improves the t-statistics and p-value estimates. If you want to sort your protein lists to perform gene set enrichment analysis, use the t-statistic instead of the fold change. You can select proteins to study them further based on the FDR, although expect a $bias$.

_prolfqua_ allows for considerable flexibility to model your experiments. Various types of models are available, and the contrast specification is consistent and intuitive for all models. The benchmarking results are comparable with other methods. The modular design of _prolfqua_, allows for customizing the analysis or adding new features, e.g., generalized linear models (_glm_'s) for missing data modeling or robust linear models (_rlm_'s), quickly.

Using the tidy table to model the data ensures interoperability with other proteomics-related packages that manage their data with tidy-tables, e.g., _protti_. Furthermore, the use of `R6` classes, which encapsulate the configuration and the data, allows writing very concise code (no function arguments needed), and auto-completion support in the Rstudio editor makes it easy for novices to explore _prolfquas_ functionality.

In summary _prolfqua_ is an easy-to-use and feature rich R package to analyse quantitative mass spectrometric data, report results, and benchmark MS software and statistical methods. We provide more details at the website https://wolski.github.io/prolfqua/. To install:

```{r echo=TRUE, eval=FALSE}
install.packages('remotes')
remotes::install_gitlab("wolski/prolfquaData",
                        host="gitlab.bfabric.org")
remotes::install_github('wolski/prolfqua',
                        build_vignettes = TRUE)

```

# Supplement

## Comparing benchmark results

Here we compare the benchmark results for the proDA model and the linear model with variance moderation, but using only the protein where inference could be performed for both models. We see that the fold change estimates obtained with the proda pacakge show better diagnostic accuracy.

```{r}

prodB <- allBenchmarks$benchmark_proDA$clone(deep=TRUE)
proda <- prodB$.data
modB <- allBenchmarks$benchmark_ProtModerated

mod <- modB$.data
mod <- mod[!is.na(mod$statistic),]
modSum <- modB$pAUC_summaries()

prodaMod <- semi_join(proda, mod, by = c("contrast", "protein_Id"))
prodB$.data <- prodaMod
d <- prodB$missing_contrasts()
prodSum <- prodB$pAUC_summaries()

tmp <- inner_join(prodSum$ftable$content, modSum$ftable$content, by=c("contrast","what"), suffix = c(".proDA",".lmMod"))
tmp <- tmp %>% unite("score_contr" , what, contrast, remove = FALSE)

ggplot(tmp, aes(x = contrast, y = pAUC_10.proDA - pAUC_10.lmMod)) + geom_bar(stat="identity") + facet_wrap(~what) +  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

```{r}
proda <- allBenchmarks$benchmark_proDA$data()
imp <- allBenchmarks$benchmark_imputation$data()

x <- inner_join(imp, proda, by = c("contrast","protein_Id", "TP", "species"), suffix = c(".imp",".proda"))

ggplot(x,aes( x = estimate.imp, y = estimate.proda) ) + geom_point() + facet_wrap(~contrast + species) + geom_hline(yintercept = 0, color="red") + geom_vline(xintercept = 0, color="red")

```

# Acknoledgements

The authors thank the technology platform fund of the University of Zurich.

# References

### 2021-12-02 inst/MyArticle/PosterBioconductor.Rmd
changes needed because prolfqua uses LazyLoad False.

---
title: |
  <center>The __prolfqua__ R-package</center> 
  <center>for Analysing Proteomics Label Free Quantification Experiments <center>
author: 
  - name: Witold Wolski
    affil: 1; 2
    orcid: '0000-0002-6468-120X'
  - name: Paolo Nanni
    affil: 1
    orcid: '0000-0001-8429-3557'
  - name: Jonas Grossmann
    affil: 1; 2
    orcid: '0000-0002-6899-9020'
  - name: Maria D'Errico
    affil: 1; 2
    orcid: '0000-0001-9467-9058'
  - name: Ralph Schlapbach
    affil: 1
    orcid: '0000-0002-7488-4262'
  - name: Christian Panse
    affil: 1; 2
    orcid: '0000-0003-1975-3064'

affiliation:
  - num: 1
    address: Functional Genomics Center Zurich - ETH Zurich/University of Zurich (https://www.fgcz.ch/)
  - num: 2
    address: Swiss Institute of Bioinformatics (https://www.sib.swiss/)
column_numbers: 5
poster_height: "38in"
poster_width: "65in"
title_textsize:	"100pt"
font_family: Monaco
primary_colour: "#6f9e73"
secondary_colour: "#7d5f60"
accent_colour: "#248ce0"
titletext_fontfamily: Arial
body_textsize: "44px"
logoleft_name: ../Figures/hexStickerProlfqua.png
logoright_name: FGCZLogo.png
output: 
  posterdown::posterdown_html
bibliography: bibliography.bib

---

# Introduction

```{r setup, include=FALSE}
library(tidyverse)

library(ggplot2)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 7, fig.height = 5)

```

We use `prolfqua` to develop highly customizable, visually appealing, and interactive data analysis reports in pdf or HTML format for mass spectrometry based protein quantification experiments. We use `prolfqua` to visualize and model simple experimental designs with a single explanatory variable and complex experiments with multiple factors. The `prolfqua` package integrates essential steps of the data analysis workflow: quality control, data normalization, protein intensity inference, modeling, hypothesis testing and sample size estimation. We further use `prolfqua` to benchmark data acquisition, data preprocessing or data modeling methods. We developed and improved the package by applying the "Eating your own dog food" principle, making it easy to use.

# Implementation & Methods

We store all the data needed for analysis in a single data frame in a tidy table, i.e., every column is a variable, every row is an observation, every cell is a single value. Using an __R6__ configuration object, we specify what variable is in which column, making it easy to integrate new inputs in `prolfqua` if provided in tidy tables. For example, to visualize tidy _Spectronaut_, or _Skyline_ outputs, or data in _MSStats_ format, only a few lines of code to update the `prolfqua` configuration are needed. For popular software like _MaxQuant_ or _MSFragger_, which stores the same variable (e.g., intensity) in multiple columns, one for each sample, we implemented methods that transform the data into tidy tables. Relying on the tidy data table enabled us to easily interface with many data manipulation, visualization, and modeling methods implemented in base __R__ and the tidyverse. We use __R6__ classes to structure the functionality of the package (see Figure \@ref(fig:LFQData) and Figure \@ref(fig:ContrastUML)). __R6__ classes are well supported (e.g. auto-completion) and easy to use (OOP).

```{r LFQData, echo=FALSE, fig.cap="Class Diagram for LFQData and related objects. The LFQData_Plotter class uses the LFQData class to implement methods for plotting.", out.width = '90%'}
knitr::include_graphics("LFQData.png")
```

__R__ linear model and linear mixed effect models allow modeling parallel designs, repeated measurements, factorial designs, and many more. __R__'s formula interface for linear models is flexible, widely used, and well documented. This makes it easy to reproduce an analysis performed with `prolfqua` in other statistical programming languages. We implemented features specific to high throughput experiments, such as the experimental Bayes variance and p-value moderation, which utilizes the parallel structure of the protein measurements and the analysis [@Ritchie2015]. We also compute probabilities of differential protein regulation based on peptide level models [@Suomi2017bEnhanced]. We also integrate the results of other packages, e.g. _SaintExpress_ or _proDA_. We use R6 to model the statistical modeling functionality in `proflqua` (see Figure \@ref(fig:ContrastUML)).

(ref:ContrastUML) UML diagram of modeling and contrast related classes. Different strategies e.g. _lm_, _lmer_, and _glm_ can be used to fit the models using a model builder (see code). All classes computing contrasts implement the _Contrast_ interface.

```{r ContrastUML, echo=FALSE, fig.cap="(ref:ContrastUML)", out.width = '90%'}
knitr::include_graphics("ContrastClassesUML.png")
```

# Example Analysis

The IonStar  [@shen2018ionstar] benchmark dataset contains _H. sapiens_ proteins with constant concentrations and _E. coli_ proteins with varying concentrations. We know that for _H. sapiens_ proteins the difference $\beta$ between two dilutions should be $\beta = 0$ while for _E. coli_ proteins, we know that the difference between dilutions should be $\beta \ne 0$. For benchmarking we used the contrasts resulting in small fold-changes $\beta = 1.2,1.25,1.3(3),1.5$. Peptide intensities are $\log_2$ transformed and robust z-score scaled but then rescaled to preserves the original range of the data. Protein intensities are estimated using Tukey's median polish (see code example and Figure \@ref(fig:prepro)).

```{r prepro,  echo = TRUE , fig.cap="(ref:scaling)", out.width = '90%'}
library(prolfqua)
d <- prolfqua_data('data_ionstar')$filtered()
lfqd <- LFQData$new(d$data, d$config) 
t <- lfqd$get_Transformer()
lfqd <-  t$log2()$robscale()$lfq
pl <- lfqd$get_Plotter()
p_1 <- pl$intensity_distribution_density() +
  labs(tag = "A") + theme(legend.position = "none")
agr <- lfqd$get_Aggregator()
lfqp <- agr$medpolish()
p_2 <- agr$plot()$plots[[1]] + labs(tag = "B")
p_3 <- lfqp$get_Stats()$violin() + labs(tag = "C")
pl <- lfqp$get_Plotter()
p_4 <- pl$boxplots()$boxplot[[1]] + labs(tag = "D")
ggpubr::ggarrange(p_1, p_2, p_3, p_4)
```

(ref:scaling) Panel A - Peptide intensity distributions for 20 samples, Panel B - Peptide intensities for protein _5NTD_ and protein intensity estimate (black dashed line), Panel C - distribution of protein standard deviations for all dilutions, Panel D - Distribution of protein intensities for protein _5NTD_.

(ref:exampleContrasts) Left panel - histogram of p-values. Right panel - volcano plots.

```{r exampleContrasts, fig.cap="(ref:exampleContrasts)", echo=TRUE, out.width = '90%'}
contrasts <- c(
  "dilution_(9/7.5)_1.2" =   "dilution.e - dilution.d",
  "dilution_(7.5/6)_1.25" =   "dilution.d - dilution.c",
  "dilution_(6/4.5)_1.3(3)" =   "dilution.c - dilution.b",
  "dilution_(4.5/3)_1.5" =   "dilution.b - dilution.a"
)
# fit model
lmmodel <- paste(lfqp$intensity_column()," ~ dilution.")
modelFunction <- strategy_lm( lmmodel, model_name = "lm")
models <- build_model(lfqp, modelFunction)
p1 <-  models$anova_histogram()
# compute contrasts from linear model and with imputation
contr <- Contrasts$new(models, contrasts) %>% 
  ContrastsModerated$new()
conI <- ContrastsSimpleImpute$new( lfqp, contrasts) %>% 
  ContrastsModerated$new()
tmp <- addContrastResults(contr, conI)
# visualize results
pl <- tmp$merged$get_Plotter()
p1 <- pl$histogram()$p.value
p2 <- pl$volcano()$FDR +
   theme(legend.position = "bottom")
gridExtra::grid.arrange(p1, p2, ncol = 2)

```

# Benchmarking

```{r}
allBenchmarks <- readRDS("../../inst/Benchresults/allBenchmarks.RDS")
benchmark_msstats <- readRDS("../../inst/Benchresults/benchmark_msstats.RDS")
msFragger <- readRDS(file = "../../inst/Benchresults/MSFragger_medpol_benchmark.RDS")

allBenchmarks$benchmark_mssstats <- benchmark_msstats
allBenchmarks$benchmark_msFragger <- msFragger
allBenchmarks <- allBenchmarks[c("benchmark_imputation","benchmark_ProtModerated",  "benchmark_mixedModerated", "benchmark_ropeca","benchmark_merged","benchmark_mssstats"   )]
```

(ref:benchmarkROC) Left panel - Partial area under the ROC curve at $10\%$ FPR for all contrasts computed using various modelling methods available in _prolfqua_ and for _MSstats_. Red line average area under the curve. Left panel - Difference to mean partial area under the ROC curve for various models (color coded line), at $10\%$ FPR, x - axis results for various fold changes.

```{r benchmarkROC, fig.cap="(ref:benchmarkROC)", out.width = '90%', fig.width=10, fig.height=10}

ttt <- sapply(allBenchmarks, function(x){x$complete(FALSE)})
res <- map_df(allBenchmarks, function(x){x$pAUC()})
res <- res %>% mutate(whatfix = case_when(what == "scaled.beta.based.significance" ~ "scaled.p.value", TRUE ~ what))

norm <- res %>% group_by(contrast,whatfix) %>% summarize(meanpAUC_10 = mean(pAUC_10))
res <- inner_join(res, norm)
res <- mutate(res , pAUC_10n = pAUC_10 - meanpAUC_10)

resAllB <- res %>% dplyr::filter(contrast == "all")

p1 <- ggplot2::ggplot(resAllB, aes(x = Name, y = pAUC_10)) +
  geom_bar(stat = "identity") +
  facet_wrap(~whatfix)  + 
  coord_cartesian(ylim = c(min(resAllB$pAUC_10),max(resAllB$pAUC_10))) + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = -90, vjust = 0.5)) +
  geom_hline(aes(yintercept = meanpAUC_10), color = "red") + xlab("")

p2 <- ggplot2::ggplot(res, aes(x = contrast, y = pAUC_10n, group = Name)) +
  geom_line(stat = "identity", aes(linetype = Name, color = Name), size = 1) + 
  facet_wrap(~ whatfix, scales = "free") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = -90, vjust = 0.5)) +
  geom_hline(aes(yintercept = 0), color = "red") + 
   theme(legend.position = "bottom", legend.title = element_blank())

gridExtra::grid.arrange(p1, p2, ncol = 2)

```

The Benchmark functionality of `prolfqua` includes ROC curves and computes partial areas under those curves (pAUC) and other scores (for more details see [BenchmarkingIonstarData](https://wolski.github.io/prolfqua/articles/BenchmarkingIonstarData.html)). We use it to study how well linear, mixed effect models or p-value moderation model quantitative mass spectrometric high throughput experiments and compare it with results produced by _MSstats_ [@MSstats2014]. Figure \@ref(fig:benchmarkROC)

A relevant parameter is the number of proteins for which we estimated the contrasts (see Figure \@ref(fig:FDRfdp)). It indicates how robust the models are in the presence of missing data. Also important is if the FDR produced is an unbiased estimate of the false discovery proportion (FDP) (see Figure \@ref(fig:FDRfdp) right panel).

(ref:FDRfdp) Left panel - Number and percentage of estimated contrasts by modeling method. Some modelling methods are more affected by missing data. Right panel - Compare FDR estimate with false discovery proportion (FDP).

```{r FDRfdp, fig.cap = "(ref:FDRfdp)", out.width = '90%', fig.width=10, fig.height=7}

dd <- map_df(allBenchmarks, function(x){res <- x$smc$summary; res$name <- x$model_name;res})
dd <- dd %>% mutate(nrcontrasts = protein_Id * (4 - nr_missing))
dds <- dd %>% group_by(name) %>% summarize(nrcontrasts = sum(nrcontrasts))
dds$percent <- dds$nrcontrasts/max(dds$nrcontrasts) * 100

nrgg <- dds %>% ggplot(aes(x = name, y = nrcontrasts )) + 
  geom_bar(stat = "identity", fill="white", colour = "black") + 
  coord_cartesian(ylim = c(min(dds$nrcontrasts) - 100, max(dds$nrcontrasts) + 10)) +
  theme(axis.text.x = element_text(angle = -90, vjust = 0.5)) +
  geom_text(aes(label = round(nrcontrasts, digits = 1)),
            vjust = 1, hjust = -0.2, angle = -90) + 
  geom_text(aes(label = paste0("(",round(percent, digits = 1),"%)")),
            vjust = -1, hjust = -0.2, angle = -90) 
p1 <- nrgg

dd <- map_df(allBenchmarks, function(x){res <- x$get_confusion_FDRvsFDP(); res$name <- x$model_name;res})

ddb <- filter(dd, contrast == "dilution_(4.5/3)_1.5")
ddb <- dd %>% dplyr::filter(contrast == "dilution_(7.5/6)_1.25")
ddb <- dd %>% dplyr::filter(contrast == "all")

p2 <- ddb %>% ggplot(aes(y = FDP_,  x  = scorecol )) + 
  geom_line(aes(color = model_name, linetype = model_name)) +
  facet_wrap(~contrast) + 
   geom_abline(intercept = 0, slope = 1, color = 2) + 
   theme(legend.position = "bottom")

gridExtra::grid.arrange(p1, p2, ncol = 2)

```

# Conclusions

`prolfqua` is an easy-to-use package to analyze quantitative mass spectrometric data, report results, and benchmark MS software and statistical methods. We provide more details at the website https://wolski.github.io/prolfqua/. To install:

```{r echo=TRUE, eval=FALSE}
install.packages('remotes')
remotes::install_gitlab("wolski/prolfquaData",
                        host="gitlab.bfabric.org")
remotes::install_github('wolski/prolfqua',
                        build_vignettes = TRUE)
```

The authors thank the technology platform fund of the University of Zurich.

# References

### 2021-12-02 inst/MyArticle/Prolfqua.Rmd
changes needed because prolfqua uses LazyLoad False.

---
journal: jacsat
type: article
author:
  - name: Witold E. Wolski
  - name: Christian Panse
  - name: Jonas Grossman
  - name: Paolo Nanni
  - name: Ralph Schlappbach
title:
  formatted: Prolfqua - \textsf{Pro}teomics label free quantification using linear models.
   class\footnote{A footnote for the title}
  short: prolfqua
abbr: IR,NMR,UV
keywords: American Chemical Society, \LaTeX
abstract: We introduce here the R package and benchmark linear modelling methods available in R for label free protein expression analysis.
bibliography: acstest.bib
csl: american-chemical-society.csl
output: rticles::acs_article
editor_options: 
  chunk_output_type: console
---

\begin{tocentry}
Some journals require a graphical entry for the Table of Contents.
This should be laid out ``print ready'' so that the sizing of the
text is correct.

Inside the \texttt{tocentry} environment, the font used is Helvetica
8\,pt, as required by \emph{Journal of the American Chemical
Society}.

The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum
permitted for  \emph{Journal of the American Chemical Society}
graphical table of content entries. The box will not resize if the
content is too big: instead it will overflow the edge of the box.

This box and the associated title will always be printed on a
separate page at the end of the document.
\end{tocentry}

# Introduction

```{r setup, include=FALSE}
library(tidyverse)
library(prolfqua)
library(ggplot2)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

  The R-package for **pro**teomics **l**abel **f**ree **qua**ntification `prolfqua` (read: prolewka) evolved from functions and code snippets used to visualize and analyze label-free quantification data. To compute protein fold changes among treatment conditions, we first used t-test, linear models, or functions implemented in the package limma. We evaluated [MSStats]( 10.18129/B9.bioc.MSstats), [ROPECA](10.1038/s41598-017-05949-y) or [MSqRob](https://github.com/statOmics/MSqRob) all implemented in R, with the idea to integrate the various approaches in our analysis pipeline. Although these packages are implemented in R, model specification, input, and output formats differ widely and wildly. Furthermore, to better understand the inferences produced, we decided to use the R linear and mixed effect models directly and transparently and to reimplement some features from the other packages when considered useful. The R-package prolfqua is the outcome of this venture.

The _prolfqua_ package integrates most of the steps of the LFQ data analysis workflow: quality control, data normalization, protein aggregation, sample size estimation, modeling, and hypothesis testing. For instance, the quality control module makes it easy to determine the coefficients of variations (CV) for peptides and proteins within conditions, summarize and visualize missing data or intensities.

When developing _prolfqua_, we draw inspiration from packages such as _sf_, which uses data in a long tidy table format. We represent the data needed for analysis using a single data-frame in long format and an R6 configuration object. Using the configuration object, we specify what data is in which column, making it easy to integrate new inputs if provided in long tidy tables. All that is needed to incorporate Spectronaut, Skyline text output, or the __MSStats__ data format is a few code lines of code to update the configuration object. For software like MaxQuant or MSFragger, which writes the data into a wide table, with several intensity columns, one for each sample, we implemented methods that transform the data into a long tidy table. Relying on the long tidy data table enabled us to easily interface with many data manipulation, visualization, and modeling methods implemented in base R and the _tidyverse_.

 R linear model and linear mixed effect models allow modeling parallel designs, repeated measurements, factorial designs, and many more.  R's formula interface for linear models is flexible, widely used, and well documented. We integrate R's linear model and mixed model formula interface in _prolfqua_'s API. This glass box approach should make it easy to reimplement an analysis performed with prolfqua using base R or other programming languages by reading the analysis script and without looking at the package code. Knowledge of the R regression model infrastructure is an advantage when using our package. Acknowledging the formula interface's complexity and the popularity of MSstats, we provided functionality to derive the model formula from an MSstats data format.

The prolfqua package fits the linear model to all the proteins in the dataset. Afterward, we can compute contrast to test the hypothesis, perform ANOVA, or model selection using the likelihood ratio test for thousands of proteins. By exploiting the parallel structure of the analysis, _prolfqua_ implements t-statistic moderation to improve inference for small sample sizes [_limma_]. It also computes probabilities of differential protein regulation based on peptide level models [_ropeca_].

To optimize a data analysis workflow, prolfqua implements and integrates benchmarking methods, e.g., computing ROC curves. We implemented these methods to assess differences between linear models, mixed effect models, p-value moderation, or Bayesian regression models using the IonStar dataset.

We developed and improved the package by applying the "Eating your own dog food" principle. We used it to analyze SRM, PRM, and LFQ data and to generate analysis reports. The current iteration of the packages API groups the functionality into R6 classes (e.g., LFQPlotter, LFQSummarizer, LFQStats, Model, Contrast, etc.). We choose R6 for its visual debugging and code completion support in RStudio.

We believe that the package makes it easy to perform proteomics quantitative data analysis, generate visualizations, and integrate them into reproducible reports using Rmarkdown. You can install the package from [www.github.com/wolski/prolfqua](www.github.com/wolski/prolfqua), and we are working on making it available on Bioconductor.

# Methods

## Benchmarking
The benchmark dataset contains _H. sapiens_ proteins with constant concentrations and _E. coli_ proteins with varying concentration [@shen2018ionstar]. We know that for _H. sapiens_ proteins the difference $\beta$ between two dilutions should be $\beta = 0$ while for _E. coli_ proteins we know that the difference between dilutions should be $\beta \ne 0$.

We can use various statistics to examine this hypothesis: the contrast estimate $\beta$, or because we first log2 transformed the data the log2 fold-change (log2FC), the t-statistic $\frac{\beta}{\sqrt{var(\beta)}}$, the $p-value$  and moderated $p-value$ for models where the p-value can be obtained.

For Bayesian models we can use the posterior probabilities that $\beta=0$, e.g. the probability that $\beta < 0$ when the mean estimate $\hat{\beta}$ is greater $0$ or vice versa, i.e.  $P(\beta < 0| \hat{\beta} > 0)$ or $P(\beta > 0 | \hat{\beta} < 0)$, which we can compute from the posterior sample of the $\beta$ parameters. We can also use the statistic $\frac{mean(\beta)}{\sqrt{var(\beta)}}$ with $var{\beta}$ estimated from the posterior sample of the $\beta$ parameter.

For each statistic and each value of the statistics we then compute a confusion matrix (see Table \@ref(tab:confusionMatrix)).
From the confusion matrix we can compute measures such as true positive rate (TPR), false positive rate (FPR), false discovery proportion (FDP) or accuracy (ACC) which are given by:

```{r confusionMatrix}
table <- data.frame( c("beta != 0", "beta == 0", "Total"), matrix(c("TP","FP","R","FN","TN","","P","N","m"), ncol = 3, byrow = T))
colnames(table) <- c("Prediction \\ Truth","E.coli", "H.sapiens", "Total")
knitr::kable(table, caption = "Confusion matrix, TP - true positive, FP - false positive, FN - false negative, P - all positive cases (all E. coli proteins), N - all negative cases (all H. sapiens proteins), m- all proteins.", format="latex")

```

$$
\begin{aligned}
TPR &= \frac{TP}{TP+FN} = \frac{TP}{P}\\
FPR &= \frac{FP}{FP+TN} = \frac{FP}{N}\\
FDP &= \frac{FP}{TP + FP} = \frac{FP}{R}\\
ACC &= \frac{TP + TN}{m}
\end{aligned}
$$

By plotting the $TPR$ versus the $FPR$ we obtain the receiver operator characteristic curve (ROC curve). The area under the curve (AUC) or partial areas under the curve (pAUC), at various values of the $FPR$, are further measures of performance. By using these measures we can compare the performances of the statistics produced by the various methods examined.

A further question we can examine using the benchmark data is, how well the false discovery estimate (FDR) obtained to from the statistical model matches the false discovery proportion (FDP). The FDR is the expected value of the false discovery proportion. Ideally, the FDR should be an unbiased estimate of the $FDP$. By plotting the FDR gainst the FDP we can asses visually if these assumptions are met.

For the Bayesian models we obtain an FDR estimate by first computing the p-value using the statistic $\frac{mean(\beta)}{\sqrt{var(\beta)}}$ and assuming that it is normally distributed. We then obtain the FDR by adjusting the  p-values using the Benjamini-Hochberg procedure.

For some of the proteins, quantification results are missing in some conditions. All we required when attempting to fit a model for a protein is that there are at least two peptides per protein quantified in any of the twenty samples. That means that for some proteins, not all model parameters are estimable. 
In case of the linear model, the contrast function we implemented can handle partial models and provides estimates of some of the contrasts (see Table \@ref(tab:tablesMedpolish)). For the mixed effect models, the used `lmerTest::contest` function does not support models with missing parameter estimates, and hence we have 4 (all) or 0 (no) contrast estimates for a protein (see Table \@ref(tab:tablesMM)). For the Bayesian models obtained with _brms_ or _rstan_ we also only computed contrasts for models with complete fixed effect estimates (see Tables \@ref(tab:countsBrms), \@ref(tab:countsBrmsStudent) and \@ref(tab:countsRstand)). The ROPECA approach starts by fitting a model to each peptide and the contrast function can handle missing parameter estimates (see Table \@ref(tab:tablesRopeca)).
These differences result in different numbers of fold-change estimates for each type of model, i.e. in a different set of ground truths, which makes comparing model performances involved.
To deal with these differences, we computed our benchmark estimates, including the peptides and proteins, without fold-change estimates, by accordingly updating the numbers in the confusion matrix. To compute the $FPR$ and $TPR$ including the missing estimates we set $P =$ #all _E. coli_ peptides/proteins detected, and $N =$ #all _H. sapiens_ proteins detected.

## Benchmark Data preprocessing

We fitted the models using all the $20$ samples of the IonStar dataset (see Figure \@ref(fig:ionstar)), i.e. five different dilutions, with four technical replicates each. This allows to estimate ten different contrasts shown in Table \@ref(tab:setupContrasts). For benchmarking we only used the contrasts resulting in small fold-changes $\beta = 1.2,1.25,1.3(3),1.5$ listed in Table \@ref(tab:usedContrasts). Only those small fold-changes allowed us to see differences among the methods. For the larger fold-changes no differences can be seen among the models or statistics.

Data was preprocessed by $log_2$ transforming the peptide intensities, and subsequent robust z-score transformation. No other data filtering was used than the removal of _one hit wonders_, i.e. proteins with a single peptide assignment (For details see section [Datasets and Data preprocessing]). We did not use imputation.  The only differences we will observe are because of using different modelling methods.

$$
z = \frac{x - \bar{x}}{S} \sim \frac{x - \tilde{x}}{\tilde{S}}
$$  t fold change on the original scale, we multiply the z-score by the average variance of all the $N$ samples in the experiment.

$$
z' = z \times 1/N\sum_{i=1}^N S_i
$$
We decided to apply this simple transformation because it only requires to estimate two parameters per sample and works for experiments with thousands of proteins as well as for experiments where only a few hundreds of proteins per sample are measured.

```{r setupContrasts}

outpath <- "results_modelling_all"
#modelName  <- "Model"

DEBUG <- FALSE

Contrasts <- c(
  "dilution_(9/3)_3" =   "dilution.e - dilution.a",
  "dilution_(9/4.5)_2" =   "dilution.e - dilution.b",
  "dilution_(9/6)_1.5" =   "dilution.e - dilution.c",
  "dilution_(9/7.5)_1.2" =   "dilution.e - dilution.d",

  "dilution_(7.5/3)_2.5" =   "dilution.d - dilution.a",
  "dilution_(7.5/4.5)_1.6(6)" =   "dilution.d - dilution.b",
  "dilution_(7.5/6)_1.25" =   "dilution.d - dilution.c",

  "dilution_(6/3)_2" =   "dilution.c - dilution.a",
  "dilution_(6/4.5)_1.3(3)" =   "dilution.c - dilution.b",

  "dilution_(4.5/3)_1.5" =   "dilution.b - dilution.a"
)

tt <- data.frame(contrastsName = names(Contrasts),  contrasts = Contrasts)
tt <- tt %>% separate(contrastsName, c("dilution", "ratio", "expected fold change"), sep =  "_", remove = FALSE)

tt
```

```{r usedContrasts}
relevantContrasts <- c("dilution_(9/7.5)_1.2","dilution_(7.5/6)_1.25", "dilution_(6/4.5)_1.3(3)", "dilution_(4.5/3)_1.5" )

tt <- tt %>% dplyr::filter(contrastsName  %in% relevantContrasts)
tt <- select(tt, all_of(c("contrasts","ratio", "expected fold change")))
knitr::kable(tt, caption = "Contrasts used for benchmark.", format = "latex")

```

Furthermore, we will only model protein fold-changes of proteins were at least two peptides were observed. The reason is that peptide identification has an associated error which is significantly reduced if two independent peptide identifications of a protein are observed. Many proteomics journals require to only report proteins identified by two peptides. Proteins identified by only a single peptide are sometimes called "one-hit wonders" [@mendoza2018flexible].

In order to remove systematic differences among samples, peptide intensities need to be transformed and scaled. The transformation aims to remove heteroscedasticity while the scaling aims to remove systematic differences among samples. Because the size of the error is proportional to the intensity, by $\log$ transforming the intensities this error can be modelled with $\epsilon \propto N(0, \sigma^2)$.

Valikangas [@valikangas2016systematic] and colleagues discuss and benchmark various methods of peptide or protein intensity normalization such as variance stabilizing normalization [@huber2002variance] or quantile normalization [@bolstad2003comparison]. In this work we will be using a robust version of the z-score, where instead of the mean we use the median and instead of the standard deviation the median absolute deviation (mad):

$$
z = \frac{x - \bar{x}}{S} \sim \frac{x - \tilde{x}}{\tilde{S}}
$$

Because we need to estimate the protein fold-changes on the original scale, we have to multiply the $z$-score by the average variance of all the $N$ samples in the experiment.

$$
z' = z \times 1/N\sum_{i=1}^N S_i
$$
We decided to apply this simple transformation because it only requires to estimate two parameters per sample and works for experiments with thousands of proteins as well as for experiments where only a few hundreds of proteins per sample are measured.
For the Ionstar dataset we used the intensities of _H. sapiens_ proteins, whose concentrations do not change, to determine $\bar{x}$ and $S$ and than applied it to all the intensities (including _E. coli_) in the sample.

Figure \@ref(fig:scaling) shows the distribution of the peptide intensities within the samples before and after the intensity scaling. Figure \@ref(fig:variancestable) shows the coefficient of variation before log transforming and scaling the data and the distribution of the standard deviations afterwards.

(ref:scaling) Density plot of peptide intensities before (panel A) and after (panel B) data transformation and scaling.

```{r scaling, fig.cap="(ref:scaling)", fig.width=7, fig.height=7}
is <- prolfqua_data('data_ionstar')
pepsum <- is$filtered()
summarised <- is$subset_normalized()

p1 <-
  prolfqua::plot_intensity_distribution_density(pepsum$data,
                                                  pepsum$config) +
  theme(legend.position = "none") +
  labs(tag = "A") + theme(legend.position = "none")
p2 <-
  prolfqua::plot_intensity_distribution_density(
    summarised$data,
    summarised$config) +
  labs(tag = "B") + theme(legend.position = "none")
rocp <-
      ggpubr::ggarrange(
        p1,
        p2,
        nrow = 1,
        common.legend = TRUE,
        legend = "none"
      )
rocp
```

## Linear model

Here we show how the two-stage modelling by first estimating the protein intensities using the Tukey median polish and then fitting a linear model performs.

# Results and discussion

In this section we compare the results for the various models implemented and discussed previously, namely:

* (`prot_med_lm`) : linear model fitted to protein estimates obtained using Tukey's median polish 
* (`prot_lme4`) : mixed effects model fitted using _lme4_ 
* (`prot_ROPECA`) : linear models fitted on peptide level and aggregation of moderated p-values using the beta distribution
* (`prot_rstan`) : mixed effects model fitted with _rstan_ and with prior over $\varepsilon$ 
* (`prot_brms`) : mixed effects model fitted with _brms_ and default priors 
* (`prot_brms_student_t`) : mixed effects model fitted with _brms_ and shrinkage prior over fixed effects coefficients .

A relevant parameter is the number of proteins for which we estimated the contrasts (see Table \@ref(tab:completeCasesTab)). It indicates how robust the models are in the presence of missing data. We observe, that when using the linear model on protein level intensities or the bayesian version of the mixed effects models we were able to estimate for 4024 (out of the 4099) proteins all the four examined contrasts (complete cases). When using the mixed effect model implementation in _lme4_ this number drops to $4013$, while when summarizing peptide level models it drops further to $3999$ proteins.

The different number of proteins for which we obtained statistics of contrasts makes comparing the scores for various models complicated because the set of proteins differs, e.g. $4024$ versus $3999$. 
Therefore, we incorporate the failure to generate statistics for contrasts into the benchmark by setting $P$ and $N$ equal to the number of all proteins (see Table \@ref(tab:confusionMatrix)) and then computing the $FPR$ and $TPR$.

## Outline

# Extra information when writing JACS Communications

\begin{acknowledgement}

The authors thank technology platform fund of the University of Zurich.

\end{acknowledgement}

\begin{suppinfo}

This will usually read something like: ``Experimental procedures and
characterization data for all new compounds. The class will
automatically add a sentence pointing to the information on-line:

\end{suppinfo}
## References

### 2021-12-02 inst/MyArticle/ProlfquaPlain.Rmd
changes needed because prolfqua uses LazyLoad False.

--- 2021-12-02 inst/MyArticle/Prolfqua.Rmd
@@
 ---
-journal: jacsat
-type: article
-author:
-  - name: Witold E. Wolski
-  - name: Christian Panse
-  - name: Jonas Grossman
-  - name: Paolo Nanni
-  - name: Ralph Schlappbach
-title:
-  formatted: Prolfqua - \textsf{Pro}teomics label free quantification using linear models.
-   class\footnote{A footnote for the title}
-  short: prolfqua
-abbr: IR,NMR,UV
-keywords: American Chemical Society, \LaTeX
-abstract: We introduce here the R package and benchmark linear modelling methods available in R for label free protein expression analysis.
-bibliography: acstest.bib
-csl: american-chemical-society.csl
-output: rticles::acs_article
-editor_options: 
-  chunk_output_type: console
+title: "Prolfqua - Proteomics label free quantification using linear models."
+author: 'Witold Wolski Std Nr:'
+date: "6/11/2021"
+output: 
+  bookdown::html_document
+bibliography: bibliography.bib
+
 ---
 
-\begin{tocentry}
-Some journals require a graphical entry for the Table of Contents.
-This should be laid out ``print ready'' so that the sizing of the
-text is correct.
-
-Inside the \texttt{tocentry} environment, the font used is Helvetica
-8\,pt, as required by \emph{Journal of the American Chemical
-Society}.
-
-The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum
-permitted for  \emph{Journal of the American Chemical Society}
-graphical table of content entries. The box will not resize if the
-content is too big: instead it will overflow the edge of the box.
-
-This box and the associated title will always be printed on a
-separate page at the end of the document.
-\end{tocentry}
+# Abstract
 
@@
 Afterward, we can compute contrast to test the hypothesis, perform ANOVA, or model selection using the likelihood ratio test for thousands of proteins.
-By exploiting the parallel structure of the analysis, _prolfqua_ implements t-statistic moderation to improve inference for small sample sizes [_limma_].
-It also computes probabilities of differential protein regulation based on peptide level models [_ropeca_].
+By exploiting the parallel structure of the analysis, _prolfqua_ implements t-statistic moderation to improve inference for small sample sizes [@Smyth2004linear].
+It also computes probabilities of differential protein regulation based on peptide level models [@Suomi2017bEnhanced].
 
@@
 ## Benchmarking
+
 The benchmark dataset contains _H.
@@
 For each statistic and each value of the statistics we then compute a confusion matrix (see Table \@ref(tab:confusionMatrix)).
-From the confusion matrix we can compute measures such as true positive rate (TPR), false positive rate (FPR), false discovery proportion (FDP) or accuracy (ACC) which are given by:
+From the confusion matrix we obtain measures such as true positive rate (TPR), false positive rate (FPR), false discovery proportion (FDP) or accuracy (ACC) which are given by:
 
@@
 coli proteins), N - all negative cases (all H.
-sapiens proteins), m- all proteins.", format="latex")
+sapiens proteins), m- all proteins.")
 
@@
 That means that for some proteins, not all model parameters are estimable.
-In case of the linear model, the contrast function we implemented can handle partial models and provides estimates of some of the contrasts (see Table \@ref(tab:tablesMedpolish)).
-For the mixed effect models, the used `lmerTest::contest` function does not support models with missing parameter estimates, and hence we have 4 (all) or 0 (no) contrast estimates for a protein (see Table \@ref(tab:tablesMM)).
-For the Bayesian models obtained with _brms_ or _rstan_ we also only computed contrasts for models with complete fixed effect estimates (see Tables \@ref(tab:countsBrms), \@ref(tab:countsBrmsStudent) and \@ref(tab:countsRstand)).
-The ROPECA approach starts by fitting a model to each peptide and the contrast function can handle missing parameter estimates (see Table \@ref(tab:tablesRopeca)).
+In case of the linear model, the contrast function we implemented can handle partial models and provides estimates of some of the contrasts.
+For the mixed effect models, the used `lmerTest::contest` function does not support models with missing parameter estimates, and hence we have either four or zero contrast estimates for a protein.
+The ROPECA approach starts by fitting a model to each peptide and the contrast function can handle missing parameter estimates.
 These differences result in different numbers of fold-change estimates for each type of model, i.e. in a different set of ground truths, which makes comparing model performances involved.
-To deal with these differences, we computed our benchmark estimates, including the peptides and proteins, without fold-change estimates, by accordingly updating the numbers in the confusion matrix.
-To compute the $FPR$ and $TPR$ including the missing estimates we set $P =$ #all _E.
-coli_ peptides/proteins detected, and $N =$ #all _H.
-sapiens_ proteins detected.
 
@@
 $$
-We decided to apply this simple transformation because it only requires to estimate two parameters per sample and works for experiments with thousands of proteins as well as for experiments where only a few hundreds of proteins per sample are measured.
+We decided to apply this simple transformation because it only requires to estimate two parameters per sample. For the ionstar dataset we estimated \bar{x} and $S$ using the human proteins only.
 
@@
 
-tt
 ```
@@
 tt <- tt %>% dplyr::filter(contrastsName  %in% relevantContrasts)
-tt <- select(tt, all_of(c("contrasts","ratio", "expected fold change")))
-knitr::kable(tt, caption = "Contrasts used for benchmark.", format = "latex")
+tt <- dplyr::select(tt, all_of(c("contrasts","ratio", "expected fold change")))
+knitr::kable(tt, caption = "Contrasts used for benchmark.")
 
@@
 
-## Linear model
+## Linear models
 
-Here we show how the two-stage modelling by first estimating the protein intensities using the Tukey median polish and then fitting a linear model performs.
+We fitted a linear model implemented by the R function `lm` to protein intensities inferred from peptide intensities using the Tukey's median polish.
+The second mixed-effects model, implemented in the R function `lmer`, fitted to peptide level intensities.
+We modeled the peptide measurements as repeated measurements of the protein.
+The third model is again a linear model but this time fitted to peptide intensities.
+By this we have for each protein several models, which we then summarize.
+
+## Contrast estimation with imputation
+
+To handle cases, where no observations were made in one of the conditions, we model the condition estimates and variances as follows.
+Either we use the mean of the protein intensity in the group, or if there are no observations we impute by using the average of $10\%$ smallest group mean intensities in the experiment.
+Here we assume that if there are no observations it is because of the limit of detection (LOD).
+To estimate the variance, we compute the pooled variance and standard deviation of all conditions $\hat{var}$.
+We assume that the variance of the protein is constant in all the conditions.
+The t-statistic is then given by $\frac{\bar x}{\hat{\sigma}}$.
+
+## Computing contrasts
+
+Given a linear model contrasts can be computed by $\hat{\beta_{c}} = \sum l\beta_{m}$ and $\textrm{var} \hat\beta_c = l\sigma^2 (X^T X)^{-1} l^T$, with $X$ being the design matrix, $\beta_m$ the model parameters and $l$ an array of coefficients.
+The degrees of freedom for the contrast are equal to the residual degrees of freedom of the linear model.
+For estimating contrasts from mixed effects models we used the function `contest` implemented in the R package `lmerTest` [@Kuznetsova2017lmerTest] and used the Satterthwaite method to estimate the denominator degrees of freedom.
+
+## P-value moderation
+
+From the linear and the mixed effect models, we can obtain the residual standard deviation $\sigma$, and degrees of freedom.
+[@Smyth2004linear] discuss how, using the empirical Bayes paradigm, to use the $\sigma$ and $df$ of all models to estimate a prior $\sigma$ and prior degrees of freedom, and posterior $\tilde \sigma$.
+These can be used to moderate the t-statistics by:
+$\tilde{t}_{pj} = \frac{t_{pj} s_p}{\tilde{s}_p}$
+and p-values.
+
+## Summarizing peptide level models
+
+To summarize peptide level models to protein models we did applied the method suggested by [@Suomi2017bEnhanced], that is to use the median scaled p-value of the peptide models and cumulative distribution function of the beta distribution function to determine a regulation probability of the protein.
+
+To obtain the median p-value of the protein, we first rescaled the peptide p-values by taking the direction of the fold-change $\hat \beta$ into account, i.e.:
+
+\begin{equation}
+p_{s} =
+  \begin{cases}
+1-p, & \textrm{if}~ \hat{\beta} > 0\\
+p-1, & \textrm{otherwise}
+\end{cases}
+\end{equation}
+
+Afterwards, the median scaled p-value $\tilde{p}_s$ is determined and using the transformation below, transformed back onto the original scale.
+
+$$
+\tilde{p} = 1 - |\tilde{p}_{s}|
+$$
+Because we used the median as the i-th order statistic  $i = n/2 +0.5$.
+Therefore, $\gamma = i = n/2 + 0.5$ and $\delta = n - i + 1 = n - (n/2 + 0.5) + 1 = n/2 + 0.5 = \gamma$ are used to parameterize the CDF of the Beta distribution.
 
 # Results and discussion
+
+## Method comparison
 
@@
 * (`prot_med_lm`) : linear model fitted to protein estimates obtained using Tukey's median polish 
-* (`prot_lme4`) : mixed effects model fitted using _lme4_ 
+* (`prot_lme4`) : mixed effects model fitted using _lme4_ to peptide intensities.
 * (`prot_ROPECA`) : linear models fitted on peptide level and aggregation of moderated p-values using the beta distribution
-* (`prot_rstan`) : mixed effects model fitted with _rstan_ and with prior over $\varepsilon$ 
-* (`prot_brms`) : mixed effects model fitted with _brms_ and default priors 
-* (`prot_brms_student_t`) : mixed effects model fitted with _brms_ and shrinkage prior over fixed effects coefficients .
+
+The table below summarizes the contrast estimates produced which will be benchmarked.
+
+|                    | Model                   | Contrast   | Moderation   | Aggregation  |
+|--------------------|:-----------------------:|------------|--------------|--------------|
+| Protein Intensity  |  lm                     | o          |  o           |              |
+| Protein Intensity Imputed | pooled variance  | o          |  o           |              |
+| Peptide Intensity  |  lmer                   | o          |  o           |              |
+| Peptide Intensity  |  lm                     |            |              |  o           |
 
@@
 When using the mixed effect model implementation in _lme4_ this number drops to $4013$, while when summarizing peptide level models it drops further to $3999$ proteins.
-
 The different number of proteins for which we obtained statistics of contrasts makes comparing the scores for various models complicated because the set of proteins differs, e.g.
@@
 
-## Outline
+```{r}
+allBenchmarks <- readRDS("../../inst/Benchresults/allBenchmarks.RDS")
+benchmark_msstats <- readRDS("../../inst/Benchresults/benchmark_msstats.RDS")
+msFragger <- readRDS(file = "../../inst/Benchresults/MSFragger_medpol_benchmark.RDS")
 
-# Extra information when writing JACS Communications
+allBenchmarks$benchmark_mssstats <- benchmark_msstats
+allBenchmarks$benchmark_msFragger <- msFragger
+
+names(allBenchmarks)
+allBenchmarks <- allBenchmarks[c("benchmark_imputation","benchmark_ProtModerated",  "benchmark_mixedModerated", "benchmark_ropeca","benchmark_merged","benchmark_mssstats", "benchmark_msFragger"   )]
+```
+
+```{r  fig.cap="Number and percentage of estimated contrasts by modelling method.", fig.width=5, fig.height=7}
+
+dd <- map_df(allBenchmarks, function(x){res <- x$smc$summary; res$name <- x$model_name;res})
+dd <- dd %>% mutate(nrcontrasts = protein_Id * (4 - nr_missing))
+dds <- dd %>% group_by(name) %>% summarize(nrcontrasts = sum(nrcontrasts))
+dds$percent <- dds$nrcontrasts/max(dds$nrcontrasts) * 100
+
+nrgg <- dds %>% ggplot(aes(x = name, y = nrcontrasts )) + 
+  geom_bar(stat = "identity", fill="white", colour = "black") + 
+  coord_cartesian(ylim = c(min(dds$nrcontrasts) - 100, max(dds$nrcontrasts) + 10)) +
+  theme(axis.text.x = element_text(angle = -90, vjust = 0.5)) +
+  geom_text(aes(label = round(nrcontrasts, digits = 1)),
+            vjust = 1, hjust = -0.2, angle = -90) + 
+  geom_text(aes(label = paste0("(",round(percent, digits = 1),"%)")),
+            vjust = -1, hjust = -0.2, angle = -90) 
+nrgg
+
+```
+
+```{r fig.cap="Partial area under the ROC curve at 10% FPR. Red line average area under the curve.", fig.width=10, fig.width=10}
+
+ttt <- sapply(allBenchmarks, function(x){x$complete(FALSE)})
+res <- map_df(allBenchmarks, function(x){x$pAUC()})
+res <- res %>% mutate(whatfix = case_when(what == "scaled.beta.based.significance" ~ "scaled.p.value", TRUE ~ what))
+
+res$contrast %>% unique()
+#resAllB <- res %>% dplyr::filter(contrast == "dilution_(9/7.5)_1.2")
+#resAllB <- res %>% dplyr::filter(contrast == "dilution_(4.5/3)_1.5")
+#resAllB <- res %>% dplyr::filter(contrast == "dilution_(7.5/6)_1.25")
+
+norm <- res %>% group_by(contrast,whatfix) %>% summarize(meanpAUC_10 = mean(pAUC_10))
+res <- inner_join(res, norm)
+res <- mutate(res , pAUC_10n = pAUC_10 - meanpAUC_10)
+
+resAllB <- res %>% dplyr::filter(contrast == "all")
+
+p1 <- ggplot(resAllB, aes(x = Name, y = pAUC_10)) +
+  geom_bar(stat = "identity") +
+  facet_wrap(~whatfix)  + 
+  coord_cartesian(ylim = c(min(resAllB$pAUC_10),max(resAllB$pAUC_10))) + 
+  theme_minimal() + 
+  theme(axis.text.x = element_text(angle = -90, vjust = 0.5)) +
+  geom_hline(aes(yintercept = meanpAUC_10), color="red")
+p1
+```
+
+```{r fig.cap="Difference to mean partial area under the ROC curve for various models, at 10% FPR, by fold change.", fig.width=10, fig.width=10}
+
+p2 <- ggplot(res, aes(x = contrast, y = pAUC_10n, group = Name)) +
+  geom_line(stat = "identity",aes(linetype = Name, color = Name)) + 
+  facet_wrap(~whatfix, scales = "free") +
+  theme_minimal() + 
+  theme(axis.text.x = element_text(angle = -90, vjust = 0.5)) +
+  geom_hline(aes(yintercept = 0), color = "red")
+
+p2
+```
+
+```{r FDRfdp, fig.cap = "Compare FDR estimate with false discovery proportion (FDP).", fig.width=8, fig.height=8}
+dd <- map_df(allBenchmarks, function(x){res <- x$get_confusion_FDRvsFDP(); res$name <- x$model_name;res})
+ddb <- filter(dd, contrast == "dilution_(4.5/3)_1.5")
+ddb <- dd %>% dplyr::filter(contrast == "dilution_(7.5/6)_1.25")
+ddb <- dd %>% dplyr::filter(contrast == "all")
+
+ddb %>% ggplot(aes(y = FDP_,  x  = scorecol )) + 
+  geom_line(aes(color = model_name, linetype = model_name)) +
+  facet_wrap(~contrast) + 
+   geom_abline(intercept = 0, slope = 1, color = 2)
+
+```
+
+```{r fig.cap="Distribution of the FDR values produced."}
+dd %>% ggplot(aes(x = scorecol)) + geom_histogram() + facet_wrap(~model_name)
+```
 
@@
 
-The authors thank technology platform fund of the University of Zurich.
+The authors thank the technology platform fund of the University of Zurich.
 
@@
 \end{suppinfo}
+
 ## References

### 2022-01-24 inst/MyArticle/MergePoster.Rmd
itration intro

@@
 
-When analyzing parallel designs using a single explanatory variable, we can use all packages for the statistical data analysis of the data; however, if we want to model factorial designs or repeated measurements, we can use only some of them.
+TODO link to table \@ref(tab:tableOverview)
+
+When analyzing parallel group designs using a single explanatory variable, we can use all packages for the statistical data analysis of the data; however, if we want to model factorial designs or repeated measurements, we can use only some of them.
 Table \ref@(tab:tableOverview) gives an overview of the models and features supported by these packages.
@@
 
-We were inspired by the excellent R package caret, which enables to call various machine learning methods, making selecting the best ML algorithm for your problem easy.
-We wanted to have a similar tool for the analysis of LFQ data.
+We were inspired by the R package caret [TODO cite], which enables to call various machine learning methods, making selecting the best ML algorithm for your problem easy.
+We aimed for a similar R package for the analysis of LFQ data.
 However, after examining the R packages for modeling proteomics quantification data, we observed that model specification, input, and output formats differ widely and wildly.
 However, all of them fit some type of linear model either to peptide or protein intensities, determine contrasts among conditions, and afterward, some of them apply empirical Bayes variance shrinkage.
-Therefore, the revised goal was to provide a modular design, with R linear models as a core, where features such as p-value aggregation (ROPECA) or variance shrinkage could be easily enabled if required.
+Therefore, the revised goal was to provide a modular design (TODO - clarity), with R linear models as a core, where features such as p-value aggregation (ROPECA) or variance shrinkage could be easily enabled if required.
 
 Models in _prolfqua_ are specified using R’s linear model and mixed model formula interface.
-Therefore, knowledge of the R regression model infrastructure is an advantage when using our package.
+Therefore, knowledge of the R regression model infrastructure (TODO cite) is an advantage when using our package.
 Furthermore, this glass box approach should make it easy to reimplement an analysis performed with _prolfqua_ using base R or other programming languages by reading the analysis script and without looking at the package code.
-However, acknowledging the R formula interface’s complexity to non-statisticians and the popularity of MSstats, we provide the functionality to suggest a model formula from the sample annotation.
+However, acknowledging the R formula interface’s complexity to non-statisticians and the popularity of _MSstats_, we provide the functionality to suggest a model formula from the sample annotation.
 
@@
 However, if there are not observation in one of the groups, then contrasts involving this group can not be estimated.
-To adres it, for the group without observations, we use a point prior equal to the average of $10\%$ smallest group mean intensities in the dataset and infer the variance from the variance in other groups.
-A way to adress this is to annotate the produced models with information about missigness which is done by for instance MSStats.
-How missigness can modelled using the Hurdle model is described here, however to our knowledge only the R package _proDA_ models missingness directly.
+To address this, for the group without observations, we use a point prior equal to the average of $1\%$ smallest group mean intensities in the dataset and infer the variance from the variance in other groups.
+(TODO use different start of sentence) A way to address this is to annotate the produced models with information about missingness which is done by for instance _MSStats_.
+How missingness can modelled using the Hurdle model is described here (TODO citations), however to our knowledge only the R package _proDA_ models missingness directly.
 
 Furthermore, the functionality of _prolfqua_ also includes methods specific to proteomics data.
-For example, we implemented strategies to infer protein intensities from peptide intensities: top N, Tukey's median polish, robust linear models.
+For example, we implemented strategies to estimate protein intensities from peptide intensities: top N, Tukey's median polish, robust linear models (cite?).
 Furthermore, peptide or protein intensities can be scaled using the robust scaling method or transformed using _quantile_ normalization or _vsn_ from Bioconductor to remove systematic differences among samples and heteroscedasticity.

### 2022-01-25 inst/MyArticle/MergePoster.Rmd
package names all lowercase, and remove library statements.

@@
 ```{r setup, include=FALSE}
-library(tidyverse)
-library(cowplot)
-library(ggplot2)
 
@@
 # compute contrasts from linear model and with imputation
-contr <- prolfqua::Contrasts$new(models, contrasts) %>% 
+contr <- prolfqua::Contrasts$new(models, contrasts) |> 
   prolfqua::ContrastsModerated$new()
-conI <- prolfqua::ContrastsSimpleImpute$new( lfqp, contrasts) %>% 
+conI <- prolfqua::ContrastsSimpleImpute$new( lfqp, contrasts) |> 
   prolfqua::ContrastsModerated$new()
@@
 res <- map_df(allBenchmarks, function(x){x$pAUC()})
-res <- res %>% mutate(whatfix = case_when(what == "scaled.beta.based.significance" ~ "scaled.p.value", TRUE ~ what))
+res <- res |> mutate(whatfix = case_when(what == "scaled.beta.based.significance" ~ "scaled.p.value", TRUE ~ what))
 
-norm <- res %>% group_by(contrast,whatfix) %>% summarize(meanpAUC_10 = mean(pAUC_10))
+norm <- res |> group_by(contrast,whatfix) |> summarize(meanpAUC_10 = mean(pAUC_10))
 res <- inner_join(res, norm)
@@
 
-resAllB <- res %>% dplyr::filter(contrast == "all")
+resAllB <- res |> dplyr::filter(contrast == "all")
 
@@
 dd
-dd <- dd %>% mutate(nrcontrasts = protein_Id * (4 - nr_missing))
+dd <- dd |> mutate(nrcontrasts = protein_Id * (4 - nr_missing))
 dd
-dds <- dd %>% group_by(name) %>% summarize(nrcontrasts = sum(nrcontrasts))
+dds <- dd |> group_by(name) |> summarize(nrcontrasts = sum(nrcontrasts))
 dds
@@
 
-nrgg <- dds %>% ggplot(aes(x = name, y = nrcontrasts )) + 
+nrgg <- dds |> ggplot(aes(x = name, y = nrcontrasts )) + 
   geom_bar(stat = "identity", fill="white", colour = "black") + 
@@
 ddb <- filter(dd, contrast == "dilution_(4.5/3)_1.5")
-ddb <- dd %>% dplyr::filter(contrast == "dilution_(7.5/6)_1.25")
-ddb <- dd %>% dplyr::filter(contrast == "all")
+ddb <- dd |> dplyr::filter(contrast == "dilution_(7.5/6)_1.25")
+ddb <- dd |> dplyr::filter(contrast == "all")
 
-p4 <- ddb %>% ggplot(aes(y = FDP_,  x  = scorecol )) + 
+p4 <- ddb |> ggplot(aes(y = FDP_,  x  = scorecol )) + 
   geom_line(aes(color = model_name, linetype = model_name)) +
@@
 install.packages('remotes')
-remotes::install_gitlab("wolski/prolfquaData",
+remotes::install_gitlab("wolski/prolfquadata",
                         host="gitlab.bfabric.org")
@@
 tmp <- inner_join(prodSum$ftable$content, modSum$ftable$content, by=c("contrast","what"), suffix = c(".proDA",".lmMod"))
-tmp <- tmp %>% unite("score_contr" , what, contrast, remove = FALSE)
+tmp <- tmp |> unite("score_contr" , what, contrast, remove = FALSE)
 

### 2022-01-25 inst/MyArticle/PosterBioconductor.Rmd
package names all lowercase, and remove library statements.

@@
 ```{r setup, include=FALSE}
-library(tidyverse)
-
-library(ggplot2)
 
@@
 # compute contrasts from linear model and with imputation
-contr <- Contrasts$new(models, contrasts) %>% 
+contr <- Contrasts$new(models, contrasts) |> 
   ContrastsModerated$new()
-conI <- ContrastsSimpleImpute$new( lfqp, contrasts) %>% 
+conI <- ContrastsSimpleImpute$new( lfqp, contrasts) |> 
   ContrastsModerated$new()
@@
 dd <- map_df(allBenchmarks, function(x){res <- x$smc$summary; res$name <- x$model_name;res})
-dd <- dd %>% mutate(nrcontrasts = protein_Id * (4 - nr_missing))
-dds <- dd %>% group_by(name) %>% summarize(nrcontrasts = sum(nrcontrasts))
+dd <- dd |> mutate(nrcontrasts = protein_Id * (4 - nr_missing))
+dds <- dd |> group_by(name) |> summarize(nrcontrasts = sum(nrcontrasts))
 dds$percent <- dds$nrcontrasts/max(dds$nrcontrasts) * 100
 
-nrgg <- dds %>% ggplot(aes(x = name, y = nrcontrasts )) + 
+nrgg <- dds |> ggplot(aes(x = name, y = nrcontrasts )) + 
   geom_bar(stat = "identity", fill="white", colour = "black") + 
@@
 
-ddb <- filter(dd, contrast == "dilution_(4.5/3)_1.5")
-ddb <- dd %>% dplyr::filter(contrast == "dilution_(7.5/6)_1.25")
-ddb <- dd %>% dplyr::filter(contrast == "all")
+ddb <- dplyr::filter(dd, contrast == "dilution_(4.5/3)_1.5")
+ddb <- dd |> dplyr::filter(contrast == "dilution_(7.5/6)_1.25")
+ddb <- dd |> dplyr::filter(contrast == "all")
 
-p2 <- ddb %>% ggplot(aes(y = FDP_,  x  = scorecol )) + 
+p2 <- ddb |> ggplot(aes(y = FDP_,  x  = scorecol )) + 
   geom_line(aes(color = model_name, linetype = model_name)) +

### 2022-01-25 inst/MyArticle/Prolfqua.Rmd
package names all lowercase, and remove library statements.

@@
 ```{r setup, include=FALSE}
-library(tidyverse)
-library(prolfqua)
-library(ggplot2)
 
@@
 tt <- data.frame(contrastsName = names(Contrasts),  contrasts = Contrasts)
-tt <- tt %>% separate(contrastsName, c("dilution", "ratio", "expected fold change"), sep =  "_", remove = FALSE)
+tt <- tt |> separate(contrastsName, c("dilution", "ratio", "expected fold change"), sep =  "_", remove = FALSE)
 
@@
 
-tt <- tt %>% dplyr::filter(contrastsName  %in% relevantContrasts)
+tt <- tt |> dplyr::filter(contrastsName  %in% relevantContrasts)
 tt <- select(tt, all_of(c("contrasts","ratio", "expected fold change")))

### 2022-01-25 inst/MyArticle/ProlfquaPlain.Rmd
package names all lowercase, and remove library statements.

@@
 
-tt <- tt %>% dplyr::filter(contrastsName  %in% relevantContrasts)
+tt <- tt |> dplyr::filter(contrastsName  %in% relevantContrasts)
 tt <- dplyr::select(tt, all_of(c("contrasts","ratio", "expected fold change")))
@@
 res <- map_df(allBenchmarks, function(x){x$pAUC()})
-res <- res %>% mutate(whatfix = case_when(what == "scaled.beta.based.significance" ~ "scaled.p.value", TRUE ~ what))
+res <- res |> mutate(whatfix = case_when(what == "scaled.beta.based.significance" ~ "scaled.p.value", TRUE ~ what))
 
-res$contrast %>% unique()
-#resAllB <- res %>% dplyr::filter(contrast == "dilution_(9/7.5)_1.2")
-#resAllB <- res %>% dplyr::filter(contrast == "dilution_(4.5/3)_1.5")
-#resAllB <- res %>% dplyr::filter(contrast == "dilution_(7.5/6)_1.25")
+res$contrast |> unique()
+#resAllB <- res |> dplyr::filter(contrast == "dilution_(9/7.5)_1.2")
+#resAllB <- res |> dplyr::filter(contrast == "dilution_(4.5/3)_1.5")
+#resAllB <- res |> dplyr::filter(contrast == "dilution_(7.5/6)_1.25")
 
-norm <- res %>% group_by(contrast,whatfix) %>% summarize(meanpAUC_10 = mean(pAUC_10))
+norm <- res |> group_by(contrast,whatfix) |> summarize(meanpAUC_10 = mean(pAUC_10))
 res <- inner_join(res, norm)
@@
 ddb <- filter(dd, contrast == "dilution_(4.5/3)_1.5")
-ddb <- dd %>% dplyr::filter(contrast == "dilution_(7.5/6)_1.25")
-ddb <- dd %>% dplyr::filter(contrast == "all")
+ddb <- dd |> dplyr::filter(contrast == "dilution_(7.5/6)_1.25")
+ddb <- dd |> dplyr::filter(contrast == "all")
 
-ddb %>% ggplot(aes(y = FDP_,  x  = scorecol )) + 
+ddb |> ggplot(aes(y = FDP_,  x  = scorecol )) + 
   geom_line(aes(color = model_name, linetype = model_name)) +
@@
 ```{r fig.cap="Distribution of the FDR values produced."}
-dd %>% ggplot(aes(x = scorecol)) + geom_histogram() + facet_wrap(~model_name)
+dd |> ggplot(aes(x = scorecol)) + geom_histogram() + facet_wrap(~model_name)
 ```

### 2022-01-25 inst/MyArticle/MergePoster.Rmd
explicit call to funcitons with <package>::<fun>

@@
 
-bb <- data.frame(bind_rows(xx))
+bb <- data.frame(dplyr::bind_rows(xx))
 rownames(bb) <- names(xx)
@@
 We also compute probabilities of differential protein regulation based on peptide level models [@Suomi2017bEnhanced].
-We use R6 to model the statistical modeling functionality in `proflqua` (see Figure \@ref(fig:ContrastUML)).
+We use R6 to model the statistical modeling functionality in `prolfqua` (see Figure \@ref(fig:ContrastUML)).
 
@@
 
+[TODO SAINTExpress integration, proDA]
+
 ## Benchmark dataset
@@
 The degrees of freedom for the contrast are equal to the residual degrees of freedom of the linear model.
+
 For estimating contrasts from mixed effects models we used the function `contest` implemented in the R package `lmerTest` [@Kuznetsova2017lmerTest] and used the Satterthwaite method to estimate the denominator degrees of freedom.
@@
 Furthermore, we assume that the variance of the protein is constant in all the conditions.
-We impute the mean estimate of the group by using the average of $10\%$ smallest group mean intensities in the experiment.
+We impute the mean estimate of the group by using the average of $1\%$ smallest group mean intensities in the experiment.
 To estimate the variance in the condition, we use the pooled variance based on all conditions:
@@
 $$s_p^2=\frac{\sum_{i=1}^k (n_i - 1)s_i^2}{\sum_{i=1}^k (n_i - 1)},$$
+
 with $n_i$ the number of observations, and $s_i$ the standard deviation in each condition.
-The standard deviation for the t-statics is then given by:
+The standard deviation for the $t$-statistics is then given by:
 
 $$
-s = \sqrt{2}\sqrt{\frac{n_g s_p^2}{n}},
+s = \sqrt{\frac{2n_g s_p^2}{n}},
 $$
 
-Where, $n_G$ is the number of groups and $n$ is the number of observations.
+Where, $n_g$ is the number of groups and $n$ is the number of observations.
 
@@
 
-From the linear and the mixed effect models, we can obtain the residual standard deviation $\sigma$, and degrees of freedom.
-[@Smyth2004linear] discuss how, using the empirical Bayes paradigm, to use the $\sigma$ and $df$ of all models to estimate a prior $\sigma$ and prior degrees of freedom, and posterior $\tilde \sigma$.
-These can be used to moderate the t-statistics by:
+From the linear and the mixed effect models, we can obtain the residual standard deviation $\sigma$, and degrees of freedom $df$.
+[@Smyth2004linear] discuss how, using the empirical Bayes paradigm, to use the $\sigma$ and $df$ of all models to estimate a prior $\sigma$ and prior $df$, and posterior $\tilde \sigma$.
+These can be used to moderate the $t$-statistics by:
+
 $\tilde{t}_{pj} = \frac{t_{pj} s_p}{\tilde{s}_p}$
-and subsequently the p-values. We implemented this method in the class `ContrastModerated` (Figure \@ref(fig:ContrastUML)).
+
+and subsequently the $p$-values. We implemented this method in the class `ContrastModerated` (Figure \@ref(fig:ContrastUML)).
 
@@
 
-Afterwards, the median scaled p-value $\tilde{p}_s$ is determined and using the transformation below, transformed back onto the original scale.
+Afterwards, the median scaled $p$-value $\tilde{p}_s$ is determined and using the transformation below, transformed back onto the original scale.
 
@@
 Finally, we estimate the contrasts either from the linear model using the `Contrasts` class or directly from the data using the `ContrastsSimpleImpute` class.
-Afterward, we apply t-statistic moderation using the `ContrastModerated` class.
+Afterward, we apply $t$-statistic moderation using the `ContrastModerated` class.
 Finally, the `addContrastResults` function merges both sets of contrast estimates, preferring the one obtained from the linear model.
-Figure \@ref(fig:exampleContrasts) Panel A shows the distribution of the p-values, while figure Panel B shows the volcano-plot for each comparison.
+Figure \@ref(fig:exampleContrasts) Panel A shows the distribution of the $p$-values, while figure Panel B shows the volcano-plot for each comparison.
 
@@
 We can use various statistics to examine the alternative hypothesis $\beta \ne 0$: the contrast estimate, i.e.
-the $\log_2$ fold-change $\beta$, the t-statistic $\frac{\beta}{\sqrt{var(\beta)}}$, or the $p-value$ and moderated $p-value$.
+the $\log_2$ fold-change $\beta$, the $t$-statistic $\frac{\beta}{\sqrt{var(\beta)}}$, or the $p-value$ and moderated $p-value$.
 For each statistic and each value of the statistics we then compute a confusion matrix (see Table \@ref(tab:confusionMatrix)).
@@
 
+(TODO(CP): benchmark load from package - first build package with vignettes )
 ```{r}
@@
 $\log_2(FC)$, $t-statistics$ and scaled p-values allow to identify true positives (TP) given a false positive rate.
-When testing for differential expression, the $\log_2(FC)$ is outperformed by than the t-statistic or p-value.
+When testing for differential expression, the $\log_2(FC)$ is outperformed by than the $t$-statistic or $p$-value.
 If an accurate estimate of the $log_2(FC)$ is essential, the linear models fitted to protein intensities, estimated using Tukey's median polish, show the best performance.
@@
 
-We can conclude that if we want to order the proteins according to the likelihood of being differentially regulated, for instance, to perform gene enrichment analysis, the t-statistic is better suited than the fold-change estimate.
-Modeling the degrees of freedom when computing the p-values might improve the inference further.
+We can conclude that if we want to order the proteins according to the likelihood of being differentially regulated, for instance, to perform gene enrichment analysis, the $t$-statistic is better suited than the fold-change estimate.
+Modeling the degrees of freedom when computing the $p$-values might improve the inference further.
 However, the observed change is minuscule.
@@
 
-Computing the statistics at the peptide level, e.g., the t-statistics or p-value, then summarizing these statistics using their median produces the highest AUC scores among all the tested models.
+Computing the statistics at the peptide level, e.g., the $t$-statistics or $p$-value, then summarizing these statistics using their median produces the highest AUC scores among all the tested models.
 By using the beta distribution to model the number of peptides observed, we can further improve the pAUC scores (see Figure).
@@
 If you use fixed-effect linear models, apply variance moderation since this improves the t-statistics and p-value estimates.
-If you want to sort your protein lists to perform gene set enrichment analysis, use the t-statistic instead of the fold change.
+If you want to sort your protein lists to perform gene set enrichment analysis, use the $t$-statistic instead of the fold change.
 You can select proteins to study them further based on the FDR, although expect a $bias$.

### 2022-01-25 inst/MyArticle/PosterBioconductor.Rmd
explicit call to funcitons with <package>::<fun>

@@
 _SaintExpress_ or _proDA_.
-We use R6 to model the statistical modeling functionality in `proflqua` (see Figure \@ref(fig:ContrastUML)).
+We use R6 to model the statistical modeling functionality in `prolfqua` (see Figure \@ref(fig:ContrastUML)).
 
@@
 d <- prolfqua_data('data_ionstar')$filtered()
-lfqd <- LFQData$new(d$data, d$config) 
+lfqd <- prolfqua::LFQData$new(d$data, d$config) 
 t <- lfqd$get_Transformer()

### 2022-02-05 inst/MyArticle/MergePoster.Rmd
adding references.

@@
 
-[Copied this paragraph from a different article. Needs to be adjusted for prolfqua and this article]
-"Protein mass spectrometry with label-free quantification is widely used for quantitative proteomics studies.
-Nevertheless, well-principled statistical inference procedures are still lacking, and most practitioners adopt methods from transcriptomics.
-These, however, cannot properly treat the principal complication of label-free proteomics, namely many non-randomly missing values." (rewrite this sentence)
-
-The _prolfqua_ package integrates essential steps of the mass spectrometry based label free quantification data analysis workflow: quality control, data normalization, protein aggregation, modeling, hypothesis testing and sample size estimation.
-We use _prolfqua_ to develop highly customizable, visually appealing, and interactive data analysis reports in pdf or HTML format.
+Protein mass spectrometry is widely used for quantitative proteomics studies and in particular for relative protein quantification.
+Nevertheless, there is still a need for a flexible and easy-to-use, transparently supporting a variety of well principled statistical inference procedures data analysis application programmer interface in R.
+The _prolfqua_ R package integrates essential steps of the mass spectrometry-based label-free quantification data analysis workflow: quality control, data normalization, protein aggregation, statistical modeling, hypothesis testing, and sample size estimation.
 We use _prolfqua_ to visualize and model simple experimental designs with a single explanatory variable and complex experiments with multiple factors.
 Furthermore, we use _prolfqua_ to benchmark data acquisition, data preprocessing or data modeling methods.
-We developed and improved the package by applying the "Eating your own dog food" principle, making it easy to use.
+Furthermore, we use _prolfqua_ to develop highly customizable, visually appealing, and interactive data analysis reports in pdf or HTML format.
+The R package is available from https://github.com/fgcz/prolfqua.
 
@@
 
-[Jonas/Paolo could you provide some references here [...] doi's will do.]
 Proteins are carriers of function and structure in living cells.
-Hence, measuring changes in protein abundance is the subject of active research [....].
+Hence, measuring changes in protein abundance is the subject of active research [@vidova2017review].
 However, identifying and quantifying intact proteins in complex samples containing hundreds of proteins is not possible using existing methods at the moment.
-Hence, the bottom-up mass spectrometric method is employed [...].
+Hence, the bottom-up mass spectrometric method is employed [@Bubis2017, @da2020philosopher].
 Proteins that are extracted from biological samples are specifically and reproducible digested into protein fragments - peptides.
@@
 Afterward, they are weighted, identified, and quantified using mass spectrometric techniques, e.g., Electro-Spray-Ionization Mass Spectrometry (ESI-MS).
-Finally, peptide identification is achieved by fragmenting and matching the measured fragment masses to theoretical masses computed from known peptide sequences[...].
-For quantification, intact peptide ions [DDA ref - method and sofware] or products of peptide ion fragmentation [DIA ref method and software] are counted and aggregated.
-Finally, the identified and quantified peptides are assigned to proteins based on protein sequence information [...prot inference ref].
+Finally, peptide identification is achieved by fragmenting and matching the measured fragment masses to theoretical masses computed from known peptide sequences[@eng2015deeper, @yu2016pipi, @kong2017msfragger].
+For quantification, intact peptide ions [@yu2020fast, @Cox2008MaxQuant] or products of peptide ion fragmentation [@rost2014openswath, @demichev2020dia] are counted and aggregated.
+Finally, the identified and quantified peptides are assigned to proteins based on protein sequence information.
 
@@
 p_a <- pl$intensity_distribution_density() +
-  labs(tag = "A") + theme(legend.position = "none")
+  ggplot2::labs(tag = "A") + ggplot2::theme(legend.position = "none")
 agr <- lfqd$get_Aggregator()
 lfqp <- agr$medpolish()
-p_b <- agr$plot()$plots[[1]] + labs(tag = "B")
-p_c <- lfqp$get_Stats()$violin() + labs(tag = "C")
+p_b <- agr$plot()$plots[[1]] + ggplot2::labs(tag = "B")
+p_c <- lfqp$get_Stats()$violin() + ggplot2::labs(tag = "C")
 pl <- lfqp$get_Plotter()
-p_d <- pl$boxplots()$boxplot[[1]] + labs(tag = "D")
+p_d <- pl$boxplots()$boxplot[[1]] + ggplot2::labs(tag = "D")
 ggpubr::ggarrange(p_a, p_b, p_c, p_d)
@@
 p2 <- pl$volcano()$FDR +
-   theme(legend.position = "bottom")
+   ggplot2::theme(legend.position = "bottom")
 gridExtra::grid.arrange(p1, p2, ncol = 2)
@@
 
-(TODO(CP): benchmark load from package - first build package with vignettes )
+(TODO(CP): benchmark load from package - first build package with vignettes)
 ```{r}
@@
 ttt <- sapply(allBenchmarks, function(x){x$complete(FALSE)})
-res <- map_df(allBenchmarks, function(x){x$pAUC()})
-res <- res |> mutate(whatfix = case_when(what == "scaled.beta.based.significance" ~ "scaled.p.value", TRUE ~ what))
+res <- purrr::map_df(allBenchmarks, function(x){x$pAUC()})
+res <- res |> dplyr::mutate(whatfix = case_when(what == "scaled.beta.based.significance" ~ "scaled.p.value", TRUE ~ what))
 
-norm <- res |> group_by(contrast,whatfix) |> summarize(meanpAUC_10 = mean(pAUC_10))
-res <- inner_join(res, norm)
-res <- mutate(res , pAUC_10n = pAUC_10 - meanpAUC_10)
+norm <- res |> dplyr::group_by(contrast,whatfix) |> dplyr::summarize(meanpAUC_10 = mean(pAUC_10))
+res <- dplyr::inner_join(res, norm)
+res <- dplyr::mutate(res , pAUC_10n = pAUC_10 - meanpAUC_10)
 
@@
 
-p1 <- ggplot(resAllB, aes(x = Name, y = pAUC_10)) +
-  geom_bar(stat = "identity") +
-  facet_wrap(~whatfix)  + 
-  coord_cartesian(ylim = c(min(resAllB$pAUC_10),max(resAllB$pAUC_10))) + 
-  theme_minimal() + 
-  theme(axis.text.x = element_text(angle = -90, vjust = 0.5)) +
-  geom_hline(aes(yintercept = meanpAUC_10), color = "red") + xlab("")
+p1 <- ggplot2::ggplot(resAllB, ggplot2::aes(x = Name, y = pAUC_10)) +
+  ggplot2::geom_bar(stat = "identity") +
+  ggplot2::facet_wrap(~whatfix)  + 
+  ggplot2::coord_cartesian(ylim = c(min(resAllB$pAUC_10),max(resAllB$pAUC_10))) + 
+  ggplot2::theme_minimal() + 
+  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = -90, vjust = 0.5)) +
+  ggplot2::geom_hline(ggplot2::aes(yintercept = meanpAUC_10), color = "red") + xlab("")
 
-p2 <- ggplot(res, aes(x = contrast, y = pAUC_10n, group = Name)) +
-  geom_line(stat = "identity", aes(linetype = Name, color = Name), size = 1) + 
-  facet_wrap(~ whatfix, scales = "free") +
-  theme_minimal() + 
-  theme(axis.text.x = element_text(angle = -90, vjust = 0.5)) +
-  geom_hline(aes(yintercept = 0), color = "red") + 
-   theme(legend.position = "bottom", legend.title = element_blank())
+p2 <- ggplot2::ggplot(res, ggplot2::aes(x = contrast, y = pAUC_10n, group = Name)) +
+  ggplot2::geom_line(stat = "identity", ggplot2::aes(linetype = Name, color = Name), size = 1) + 
+  ggplot2::facet_wrap(~ whatfix, scales = "free") +
+  ggplot2::theme_minimal() + 
+  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = -90, vjust = 0.5)) +
+  ggplot2::geom_hline(aes(yintercept = 0), color = "red") + 
+   ggplot2::theme(legend.position = "bottom", legend.title = element_blank())
 
@@
 
-# Conclusions
+# Results and Discussion.
 
 We focused on comparing models to estimate fold changes, variances and to obtain false discovery rates.
-However, an essential parameter, not discussed further, is the normalization of the intensities within the samples to make them comparable.
+
+[Goes in the discussion or introduction, and give references for normalization e.g.
+VSN paper.] However, an essential parameter, not discussed further, is the normalization of the intensities within the samples to make them comparable.
 For this benchmark dataset, the fraction of regulated proteins (ECOLI) is relatively large.
@@
 
-The method used to infer protein intensities from more peptide intensities is an additional important factor.
+[Goes to discussion] The method used to infer protein intensities from more peptide intensities is an additional important factor.
 The original _proDA_ publication uses MaxLFQ [@cox2014accurate] protein estimates.
@@
 
+# Conclusion (tbd)
+
 # Supplement

### 2022-02-07 inst/MyArticle/MergePoster.Rmd
adding references to introduction

@@
 Label-Free Quantification (LFQ) Proteomics Experiments enable monitoring relative abundances of thousands of proteins in biological samples.
-For most studies, parallel designs with a single explanatory variable are used, e.g., treatment vs.
-control.
-More recently, more complex experimental designs are implemented with increasing numbers of samples, i.e., factorial designs and longitudinal studies (time series), sometimes with repeated measurements on the same subject [...].
+For most studies, parallel group designs are used, e.g., treatment vs.
+control are used[cite].
+More recently, more complex experimental designs are implemented with increasing numbers of samples, i.e., factorial designs and longitudinal studies (time series), sometimes with repeated measurements on the same subject[cite].
 The obtained data can be modeled using linear fixed, mixed, or random-effects models [@Bates2015FittingJSS].

### 2022-02-10 inst/MyArticle/MergePoster.Rmd
fix vignette

@@
 ttt <- sapply(allBenchmarks, function(x){x$complete(FALSE)})
+names(allBenchmarks)
+
+b1 <- allBenchmarks$benchmark_imputation$pAUC
+b1$get_confusion_benchmark
+
+allBenchmarks[[1]]$pAUC
+
 res <- purrr::map_df(allBenchmarks, function(x){x$pAUC()})

### 2022-03-16 inst/MyArticle/MergePoster.Rmd
add caret reference

--- 2022-02-07 inst/MyArticle/MergePoster.Rmd
@@
 
-We were inspired by the R package caret [TODO cite], which enables to call various machine learning methods, making selecting the best ML algorithm for your problem easy.
+We were inspired by the R package caret [@caret2008], which enables to call various machine learning methods, making selecting the best ML algorithm for your problem easy.
 We aimed for a similar R package for the analysis of LFQ data.
@@
 We store all the data needed for analysis in a single data frame in a tidy table, i.e., every column is a variable, every row is an observation, every cell is a single value[@tidydata2014].
-Using an __R6__[@R6cite] configuration object, we specify what variable is in which column, making it easy to integrate new inputs in _prolfqua_ if provided in tidy tables.
+Using an __R6__[@R6cite] configuration object, we specify what variable is in which column, making it easy to integrate new inputs in _prolfqua_ if provided as a tidy table.
 For example, to visualize tidy _Spectronaut_, or _Skyline_ outputs, or data in _MSStats_ format, only a few lines of code to update the _prolfqua_ configuration are needed.
@@
 
-ttt <- sapply(allBenchmarks, function(x){x$complete(FALSE)})
+ttt <- sapply(allBenchmarks, function(x){x$complete(TRUE)})
+
 res <- purrr::map_df(allBenchmarks, function(x){x$pAUC()})
-res <- res |> dplyr::mutate(whatfix = case_when(what == "scaled.beta.based.significance" ~ "scaled.p.value", TRUE ~ what))
+res <- res |> dplyr::mutate(whatfix = dplyr::case_when(what == "scaled.beta.based.significance" ~ "scaled.p.value", TRUE ~ what))
 
@@
   ggplot2::theme(axis.text.x = ggplot2::element_text(angle = -90, vjust = 0.5)) +
-  ggplot2::geom_hline(ggplot2::aes(yintercept = meanpAUC_10), color = "red") + xlab("")
+  ggplot2::geom_hline(ggplot2::aes(yintercept = meanpAUC_10), color = "red") + ggplot2::xlab("")
 
@@
   ggplot2::theme(axis.text.x = ggplot2::element_text(angle = -90, vjust = 0.5)) +
-  ggplot2::geom_hline(aes(yintercept = 0), color = "red") + 
-   ggplot2::theme(legend.position = "bottom", legend.title = element_blank())
+  ggplot2::geom_hline(ggplot2::aes(yintercept = 0), color = "red") + 
+   ggplot2::theme(legend.position = "bottom", legend.title = ggplot2::element_blank())
 
@@
 
-dd <- map_df(allBenchmarks, function(x){res <- x$smc$summary; res$name <- x$model_name;res})
+dd <- purrr::map_df(allBenchmarks, function(x){res <- x$smc$summary; res$name <- x$model_name;res})
 dd
-dd <- dd |> mutate(nrcontrasts = protein_Id * (4 - nr_missing))
+dd <- dd |> dplyr::mutate(nrcontrasts = protein_Id * (4 - nr_missing))
 dd
-dds <- dd |> group_by(name) |> summarize(nrcontrasts = sum(nrcontrasts))
+dds <- dd |> dplyr::group_by(name) |> dplyr::summarize(nrcontrasts = sum(nrcontrasts))
 dds
@@
 
-nrgg <- dds |> ggplot(aes(x = name, y = nrcontrasts )) + 
-  geom_bar(stat = "identity", fill="white", colour = "black") + 
-  coord_cartesian(ylim = c(min(dds$nrcontrasts) - 100, max(dds$nrcontrasts) + 10)) +
-  theme(axis.text.x = element_text(angle = -90, vjust = 0.5)) +
-  geom_text(aes(label = round(nrcontrasts, digits = 1)),
+nrgg <- dds |> ggplot2::ggplot(ggplot2::aes(x = name, y = nrcontrasts )) + 
+  ggplot2::geom_bar(stat = "identity", fill="white", colour = "black") + 
+  ggplot2::coord_cartesian(ylim = c(min(dds$nrcontrasts) - 100, max(dds$nrcontrasts) + 10)) +
+  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = -90, vjust = 0.5)) +
+  ggplot2::geom_text(ggplot2::aes(label = round(nrcontrasts, digits = 1)),
             vjust = 1, hjust = -0.2, angle = -90) + 
-  geom_text(aes(label = paste0("(",round(percent, digits = 1),"%)")),
+  ggplot2::geom_text(ggplot2::aes(label = paste0("(",round(percent, digits = 1),"%)")),
             vjust = -1, hjust = -0.2, angle = -90) 
@@
 
-dd <- map_df(allBenchmarks, function(x){res <- x$get_confusion_FDRvsFDP(); res$name <- x$model_name;res})
+dd <- purrr::map_df(allBenchmarks, function(x){res <- x$get_confusion_FDRvsFDP(); res$name <- x$model_name;res})
 
-ddb <- filter(dd, contrast == "dilution_(4.5/3)_1.5")
+ddb <- dplyr::filter(dd, contrast == "dilution_(4.5/3)_1.5")
 ddb <- dd |> dplyr::filter(contrast == "dilution_(7.5/6)_1.25")
@@
 
-p4 <- ddb |> ggplot(aes(y = FDP_,  x  = scorecol )) + 
-  geom_line(aes(color = model_name, linetype = model_name)) +
-  facet_wrap(~contrast) + 
-   geom_abline(intercept = 0, slope = 1, color = 2) + 
-   theme(legend.position = "bottom")
+p4 <- ddb |> ggplot2::ggplot(ggplot2::aes(y = FDP_,  x  = scorecol )) + 
+  ggplot2::geom_line(ggplot2::aes(color = model_name, linetype = model_name)) +
+  ggplot2::facet_wrap(~contrast) + 
+   ggplot2::geom_abline(intercept = 0, slope = 1, color = 2) + 
+   ggplot2::theme(legend.position = "bottom")
 
@@
 
-prodaMod <- semi_join(proda, mod, by = c("contrast", "protein_Id"))
+prodaMod <- dplyr::semi_join(proda, mod, by = c("contrast", "protein_Id"))
 prodB$.data <- prodaMod
@@
 
-tmp <- inner_join(prodSum$ftable$content, modSum$ftable$content, by=c("contrast","what"), suffix = c(".proDA",".lmMod"))
-tmp <- tmp |> unite("score_contr" , what, contrast, remove = FALSE)
+tmp <- dplyr::inner_join(prodSum$ftable$content, modSum$ftable$content, by=c("contrast","what"), suffix = c(".proDA",".lmMod"))
+tmp <- tmp |> tidyr::unite("score_contr" , what, contrast, remove = FALSE)
 
-ggplot(tmp, aes(x = contrast, y = pAUC_10.proDA - pAUC_10.lmMod)) + geom_bar(stat="identity") + facet_wrap(~what) +  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+ggplot2::ggplot(tmp, ggplot2::aes(x = contrast, y = pAUC_10.proDA - pAUC_10.lmMod)) + ggplot2::geom_bar(stat="identity") + ggplot2::facet_wrap(~what) +  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust=1))
 
@@
 
-x <- inner_join(imp, proda, by = c("contrast","protein_Id", "TP", "species"), suffix = c(".imp",".proda"))
-
-ggplot(x,aes( x = estimate.imp, y = estimate.proda) ) + geom_point() + facet_wrap(~contrast + species) + geom_hline(yintercept = 0, color="red") + geom_vline(xintercept = 0, color="red")
+x <- dplyr::inner_join(imp, proda, by = c("contrast","protein_Id", "TP", "species"), suffix = c(".imp",".proda"))
+ggplot2::ggplot(x,ggplot2::aes( x = diff.imp, y = diff.proda) ) + ggplot2::geom_point() + ggplot2::facet_wrap(~contrast + species) + ggplot2::geom_hline(yintercept = 0, color="red") + ggplot2::geom_vline(xintercept = 0, color="red")
 

### 2022-03-17 inst/MyArticle/paper.Rmd
Merging benchmark figures into 1.

--- 2022-03-16 inst/MyArticle/MergePoster.Rmd
@@
     address: Swiss Institute of Bioinformatics (https://www.sib.swiss/)
-column_numbers: 5
-poster_height: "30in"
-poster_width: "58in"
-title_textsize:	"90pt"
-font_family: Monaco
-primary_colour: "#000C66"
-titletext_fontfamily: Arial
-body_textsize: "35px"
-logoleft_name: C:/Users/wolski/__checkout/prolfqua/inst/Figures/FGCZLogo.png
-logoright_name: C:/Users/wolski/__checkout/prolfqua/inst/Figures/hexStickerProlfqua.png
 output: 
-  bookdown::html_document2:
+  bookdown::pdf_document2:
     toc: true
-  posterdown::posterdown_html:
 bibliography: bibliography.bib
@@
 However, all of them fit some type of linear model either to peptide or protein intensities, determine contrasts among conditions, and afterward, some of them apply empirical Bayes variance shrinkage.
-Therefore, the revised goal was to provide a modular design (TODO - clarity), with R linear models as a core, where features such as p-value aggregation (ROPECA) or variance shrinkage could be easily enabled if required.
+Therefore, the revised goal was to provide a modular object oriented design, with R linear models as a core, where features such as p-value aggregation (ROPECA) or variance shrinkage could be easily enabled (see Figure \@ref(fig:ContrastUML)).
 
 Models in _prolfqua_ are specified using R’s linear model and mixed model formula interface.
-Therefore, knowledge of the R regression model infrastructure (TODO cite) is an advantage when using our package.
-Furthermore, this glass box approach should make it easy to reimplement an analysis performed with _prolfqua_ using base R or other programming languages by reading the analysis script and without looking at the package code.
-However, acknowledging the R formula interface’s complexity to non-statisticians and the popularity of _MSstats_, we provide the functionality to suggest a model formula from the sample annotation.
+Therefore, knowledge of the R regression model infrastructure [@faraway2016extending,@MASS2002] is an advantage when using our package.
+Furthermore, this glass box approach should make it easy to reimplement an analysis which was performed with _prolfqua_ using base R or other programming languages by reading the analysis script.
+However, acknowledging the R formula interface’s complexity to non-statisticians and the popularity of _MSstats_, we provide the functionality to suggest a model formula from the sample annotation provided in tabular form similarily to MSStats [@Choi2014].
 
@@
 We use this feature instead of imputing missing observations.
-However, if there are not observation in one of the groups, then contrasts involving this group can not be estimated.
-To address this, for the group without observations, we use a point prior equal to the average of $1\%$ smallest group mean intensities in the dataset and infer the variance from the variance in other groups.
-(TODO use different start of sentence) A way to address this is to annotate the produced models with information about missingness which is done by for instance _MSStats_.
-How missingness can modelled using the Hurdle model is described here (TODO citations), however to our knowledge only the R package _proDA_ models missingness directly.
+However, if there are no observation in one of the groups, then contrasts involving this group can not be estimated.
+To address this, for the group without observations, we use a point prior equal to the average of $3\%$ smallest group intensities means in the dataset and infer the variance using the pooled the variance of other groups.
+Proteins for which the fold change was imputed in such a way are explicitely annotated in the result table.
+
+Other means of modelling missing information in proteomics data using the Hurdle model is discussed by [@goeminne2020msqrob], while the R package _proDA_ models missingness using probabilistic dropout models [@bioxrvproDA2020].
 
@@
 
-```{r LFQData, echo=FALSE, fig.cap="Class Diagram for LFQData and related objects. The LFQData_Plotter class uses the LFQData class to implement methods for plotting.", out.width = '90%'}
-fig_svg <- cowplot::ggdraw() + cowplot::draw_image("LFQDataUML.svg")
-plot(fig_svg)
+(ref:LFQData) Class Diagram for LFQData and related classes.
+The LFQData_Plotter class uses the LFQData class to implement methods for plotting.
+Similarily the LFQData_Stats, LFQData_Summary, reference the LFQData class and group methods for variance and sample size estimation or for summarizing peptide and protein counts.
+
+```{r LFQData, echo=FALSE, fig.cap="(ref:LFQData)", out.width = '90%'}
+
+knitr::include_graphics("LFQDataUML.png")
 
@@
 ```{r ContrastUML, echo=FALSE, fig.cap="(ref:ContrastUML)", out.width = '90%'}
-fig_svg <- cowplot::ggdraw() + cowplot::draw_image("ContrastClassesUML.svg")
-plot(fig_svg)
+#fig_svg <- cowplot::ggdraw() + cowplot::draw_image("ContrastClassesUML.svn")
+#plot(fig_svg)
+
+knitr::include_graphics("ContrastClassesUML.png")
 ```
@@
 lfqp <- agr$medpolish()
+
 p_b <- agr$plot()$plots[[1]] + ggplot2::labs(tag = "B")
@@
   "dilution_(7.5/6)_1.25" =   "dilution.d - dilution.c"
-  #,"dilution_(6/4.5)_1.3(3)" =   "dilution.c - dilution.b",
-  #,"dilution_(4.5/3)_1.5" =   "dilution.b - dilution.a"
 )
@@
 We use it to study how well linear, mixed effect models or p-value moderation model quantitative mass spectrometric high throughput experiments and compare it with results produced by MSstats [@MSstats2014] and by proDA [@bioxrvproDA2020].
-Figure \@ref(fig:benchmarkROC)
+Figure \@ref(fig:FDRfdp)
 
@@
 
-(TODO(CP): benchmark load from package - first build package with vignettes)
 ```{r}
-
 allBenchmarks <- readRDS("../../inst/Benchresults/allBenchmarks.RDS")
@@
 
-(ref:benchmarkROC) Left panel - Partial area under the ROC curve at $10\%$ FPR for all contrasts.
-Red line average area under the curve.
-Left panel - Difference to mean partial area under the ROC curve for various models, at $10\%$ FPR, by fold change.
+(ref:benchmarkROC)  Left panel - Difference to mean partial area under the ROC curve for various models, at $10\%$ FPR, by fold change.
 
@@
 
-p1 <- ggplot2::ggplot(resAllB, ggplot2::aes(x = Name, y = pAUC_10)) +
+p1_pAUC <- ggplot2::ggplot(resAllB, ggplot2::aes(x = Name, y = pAUC_10)) +
   ggplot2::geom_bar(stat = "identity") +
@@
 
-p2 <- ggplot2::ggplot(res, ggplot2::aes(x = contrast, y = pAUC_10n, group = Name)) +
+p2_compare_variousLevels <- ggplot2::ggplot(res, ggplot2::aes(x = contrast, y = pAUC_10n, group = Name)) +
   ggplot2::geom_line(stat = "identity", ggplot2::aes(linetype = Name, color = Name), size = 1) + 
@@
   ggplot2::theme(legend.position = "bottom", legend.title = ggplot2::element_blank())
-
-gridExtra::grid.arrange(p1, p2, ncol = 2)
 
@@
 
-Figure \@ref(fig:benchmarkROC) left panel, shows how various estimates obtained from the models, i.e.
+Figure \@ref(fig:FDRfdp) left panel, shows how various estimates obtained from the models, i.e.
 $\log_2(FC)$, $t-statistics$ and scaled p-values allow to identify true positives (TP) given a false positive rate.
@@
 
-There are only minor differences in the p-AUC when using the t-statistics or the scaled p-value  (see Figure \@ref(fig:benchmarkROC) right panel).
+There are only minor differences in the p-AUC when using the t-statistics or the scaled p-value  (see Figure \@ref(fig:FDRfdp) right panel).
 In order to compute the confusion matrices based on the p-value we first need to rescale it (see Equation \@ref(eq:scalepvalue)).
@@
 
-(ref:FDRfdp) Left panel - Number and percentage of estimated contrasts by modeling method. Right panel - Compare FDR estimate with false discovery proportion (FDP).
+(ref:FDRfdp) Left panel - Partial area under the ROC curve at $10\%$ FPR for all contrasts and three different statistics: the difference among conditions, the scaled p-value (sign(diff) * p.value) and the t-statics (higher is better).
+The red line average area under the curve.
+Left panel - Number and percentage of estimated contrasts for each modeling method (higher is better).
+Right panel - Compare FDR estimate with false discovery proportion (FDP) closer to the diagnoal (red continous line is better).
 
@@
 dd <- purrr::map_df(allBenchmarks, function(x){res <- x$smc$summary; res$name <- x$model_name;res})
-dd
 dd <- dd |> dplyr::mutate(nrcontrasts = protein_Id * (4 - nr_missing))
-dd
 dds <- dd |> dplyr::group_by(name) |> dplyr::summarize(nrcontrasts = sum(nrcontrasts))
-dds
 dds$percent <- dds$nrcontrasts/max(dds$nrcontrasts) * 100
@@
 
-gridExtra::grid.arrange(p3, p4, ncol = 2)
+hlay = rbind(c(1,1),
+            c(2,3))
+
+gridExtra::grid.arrange(p1_pAUC, p3, p4, layout_matrix=hlay)
 
@@
 The original _proDA_ publication uses MaxLFQ [@cox2014accurate] protein estimates.
-When using those, _proDA_ performs significantly worse than _proDA_ when using the protein estimates obtained from peptide intensity estimates using Tukey's median polish,  $pAUC_10(t-statistics) = 66%$ and $pAUC_10(t-statistics) = 73%$ respectively (see Articles on [@prolfquaGithubIO] for more details).
+When using those, _proDA_ performs significantly worse than _proDA_ when using the protein estimates obtained from peptide intensity estimates using Tukey's median polish,  $pAUC_{10}(\textrm{t-statistics}) = 66%$ and $pAUC_{10}(\textrm{t-statistics}) = 73%$ respectively (see Articles on [@prolfquaGithubIO] for more details).
 
@@
 
-The probabilistic dropout analysis implemented in the _proDA_ produces inferences comparable to other methods (Figure \@ref(fig:benchmarkROC) for the largest number of proteins (Figure \@ref(fig:FDRfdp) left panel).
+The probabilistic dropout analysis implemented in the _proDA_ produces inferences comparable to other methods (Figure \@ref(fig:FDRfdp) for the largest number of proteins (Figure \@ref(fig:FDRfdp) left panel).
 Moreover, the estimated fold changes are unbiased and show high diagnostic accuracy.

### 2022-03-17 inst/MyArticle/paper.Rmd
removing latex warnings

@@
 
-TODO link to table \@ref(tab:tableOverview)
-
 When analyzing parallel group designs using a single explanatory variable, we can use all packages for the statistical data analysis of the data; however, if we want to model factorial designs or repeated measurements, we can use only some of them.
-Table \ref@(tab:tableOverview) gives an overview of the models and features supported by these packages.
+Table \@ref(tab:tableOverview) gives an overview of the models and features supported by these packages.
 
@@
 Furthermore, the functionality of _prolfqua_ also includes methods specific to proteomics data.
-For example, we implemented strategies to estimate protein intensities from peptide intensities: top N, Tukey's median polish, robust linear models (cite?).
-Furthermore, peptide or protein intensities can be scaled using the robust scaling method or transformed using _quantile_ normalization or _vsn_ from Bioconductor to remove systematic differences among samples and heteroscedasticity.
+For example, we implemented strategies to estimate protein intensities from peptide intensities: top N, Tukey's median polish, robust linear models [@Grossmann2010,@goeminne2020msqrob].
+Furthermore, peptide or protein intensities can be scaled using the robust scaling method or transformed using _quantile_ normalization or _vsn_ to remove systematic differences among samples and heteroscedasticity.
 In this respect, _prolfqua_ can be compared with other R packages such _DEP_ [@DEP2018] or POMA [@POMA2021] which support the entire LFQ analysis pipeline.
-
+However, we also implemented functionality to streamline the benchmarking of the implemented methods.
 We use the Ionstar [@shen2018ionstar] dataset to benchmark the analysis pipeline and modeling methods implemented within _prolfqua_ and to compare our results with those of _MSstats_ and _proDA_.
@@
 Protein intensities are estimated using Tukey's median polish.
-Figure \@ref(fig:prepro) Panel B shows the peptide intensities and the estimated protein intensities in black.
+Figure \@ref(fig:prepro) Panel B shows the peptide intensities and the estimated protein intensities.
 The boxplot (Panel D) shows the distribution of the protein intensity estimate in each condition.
@@
 
-(ref:scaling) Panel A - Peptide intensity distributions for 20 samples, Panel B - Peptide intensities for protein _5NTD_ and protein intensity estimate (black dashed line), Panel C - distribution of protein standard deviations for all dilutions, Panel D - Distribution of protein intensities for protein _5NTD_.
+(ref:scaling) Panel A - Peptide intensity distributions for 20 samples.
+For each sample a line with a different colour is shown.
+Panel B - Peptide intensities for protein _5NTD_ are shown using a line of different colour, and the protein intensity estimate is shown using a fat black line, Panel C - distribution of standard deviations of all proteins in each dilution group (a, b, c, d, e) and overall (all), Panel D - Distribution of protein intensities for protein _5NTD_.
 

### 2022-03-18 inst/MyArticle/paper.Rmd
remove line duplicate
Cosmetics Introduction
- added ``r BiocStyle::...` link when possible
- added LaTeX math annotation of $p$- or $t$- statistics
- typos

@@
 For most studies, parallel group designs are used, e.g., treatment vs.
-control are used[cite].
-More recently, more complex experimental designs are implemented with increasing numbers of samples, i.e., factorial designs and longitudinal studies (time series), sometimes with repeated measurements on the same subject[cite].
-The obtained data can be modeled using linear fixed, mixed, or random-effects models [@Bates2015FittingJSS].
+control are used [TODO: cite].
+More recently, more complex experimental designs are implemented with increasing numbers of samples, i.e., factorial designs and longitudinal studies (time series), sometimes with repeated measurements on the same subject [TODO: cite].
+The obtained data can be modeled using linear fixed-, mixed-, or random-effects models [@Bates2015FittingJSS].
 Based on these models, tests can be applied to examine whether a factor or interactions are significant, or differences between factor levels can be estimated, and their significance tested.
-Proteins can be ordered or grouped using the estimated fold-change, t-statistics, scaled p-value, or false discovery rate (FDR) and then subjected to gene set enrichment or over-representation analysis [...] or validation experiments.
+Proteins can be ordered or grouped using the estimated fold-change, $t$-statistics, scaled $p$-value, or false discovery rate (FDR) and then subjected to gene set enrichment or over-representation analysis [TODO ...] or validation experiments.
 
-Several packages exist to model mass spectrometry LFQ experiments, e.g.
-_limma_[@Ritchie2015], _MSStats_[@Choi2014], _ROPECA_[@Suomi2017bEnhanced],  _MSqRob2_ [@Goeminne2016] or _proDA_ [@bioxrvproDA2020], to mention some, all implemented in the R.
+Several packages exist to model mass spectrometry LFQ experiments, e.g., `r BiocStyle::Biocpkg("limma")` [@Ritchie2015], `r BiocStyle::Biocpkg("MSstats")` [@Choi2014], _ROPECA_[@Suomi2017bEnhanced],  `r BiocStyle::Biocpkg("msqrob2")` [@Goeminne2016] or `r BiocStyle::Biocpkg("proDA")` [@bioxrvproDA2020], to mention some, all implemented in the R.
 Each of them has some unique features.
-For example, _MSStats_ infers the model from the sample annotation and generates the model formula.
-At the same time, _limma_ allows to specify a linear model formula and implements the empirical Bayes variance shrinkage method.
-The R-package _ROPECA_, performs a roll-up of peptide level fold change and p-value estimates, obtained from _limma_, to protein level estimates.
-Furthermore, _MSqRob2_ implements two models: robust linear models fitted to protein intensities and robust ridge regression fitted to peptide intensities and combines them with empirical Baye's variance shrinkage known from _limma_.
-The _proDA_ package implements a linear probabilistic dropout model to account for missing data without imputation.
+For example, `r BiocStyle::Biocpkg("MSstats")` infers the model from the sample annotation and generates the model formula.
+At the same time, `r BiocStyle::Biocpkg("limma")` allows to specify a linear model formula and implements the empirical Bayes variance shrinkage method.
+The R-package _ROPECA_, performs a roll-up of peptide level fold change and $p$-value estimates, obtained from `r BiocStyle::Biocpkg("limma")`, to protein level estimates.
+Furthermore, `r BiocStyle::Biocpkg("msqrob2")` implements two models: robust linear models fitted to protein intensities and robust ridge regression fitted to peptide intensities and combines them with empirical Bayes variance shrinkage known from `r BiocStyle::Biocpkg("limma")`.
+The `r BiocStyle::Biocpkg("proDA")` package implements a linear probabilistic dropout model to account for missing data without imputation.
 
@@
 
-We were inspired by the R package caret [@caret2008], which enables to call various machine learning methods, making selecting the best ML algorithm for your problem easy.
+We were inspired by the R package `r BiocStyle::CRANpkg("caret")` [@caret2008], which enables to call various machine learning (ML) methods, making selecting the best ML algorithm for your problem easy.
 We aimed for a similar R package for the analysis of LFQ data.
@@
 However, all of them fit some type of linear model either to peptide or protein intensities, determine contrasts among conditions, and afterward, some of them apply empirical Bayes variance shrinkage.
-Therefore, the revised goal was to provide a modular object oriented design, with R linear models as a core, where features such as p-value aggregation (ROPECA) or variance shrinkage could be easily enabled (see Figure \@ref(fig:ContrastUML)).
+Therefore, the revised goal was to provide a modular object oriented (OO) design, with R linear models as a core, where features such as $p$-value aggregation (_ROPECA_) or variance shrinkage could be easily enabled (see Figure \@ref(fig:ContrastUML)).
 
-Models in _prolfqua_ are specified using R’s linear model and mixed model formula interface.
+Models in `r BiocStyle::Githubpkg("prolfqua")` are specified using R’s linear model and mixed model formula interface.
 Therefore, knowledge of the R regression model infrastructure [@faraway2016extending,@MASS2002] is an advantage when using our package.
-Furthermore, this glass box approach should make it easy to reimplement an analysis which was performed with _prolfqua_ using base R or other programming languages by reading the analysis script.
-However, acknowledging the R formula interface’s complexity to non-statisticians and the popularity of _MSstats_, we provide the functionality to suggest a model formula from the sample annotation provided in tabular form similarily to MSStats [@Choi2014].
+Furthermore, this glass box approach should make it easy to reimplement an analysis which was performed with `r BiocStyle::Githubpkg("prolfqua")` using base R or other programming languages by reading the analysis script.
+However, acknowledging the R formula interface’s complexity to non-statisticians and the popularity of `r BiocStyle::Biocpkg("MSstats")`, we provide the functionality to suggest a model formula from the sample annotation provided in tabular form similarily to `r BiocStyle::Biocpkg("MSstats")` [@Choi2014].
 
@@
 To address this, for the group without observations, we use a point prior equal to the average of $3\%$ smallest group intensities means in the dataset and infer the variance using the pooled the variance of other groups.
-Proteins for which the fold change was imputed in such a way are explicitely annotated in the result table.
+Proteins for which the fold change was imputed in such a way are explicitly annotated in the result table.
 
-Other means of modelling missing information in proteomics data using the Hurdle model is discussed by [@goeminne2020msqrob], while the R package _proDA_ models missingness using probabilistic dropout models [@bioxrvproDA2020].
+Other means of modelling missing information in proteomics data using the Hurdle model is discussed by [@goeminne2020msqrob], while the R package `r BiocStyle::Biocpkg("proDA")` models missingness using probabilistic dropout models [@bioxrvproDA2020].
 
-Furthermore, the functionality of _prolfqua_ also includes methods specific to proteomics data.
-For example, we implemented strategies to estimate protein intensities from peptide intensities: top N, Tukey's median polish, robust linear models [@Grossmann2010,@goeminne2020msqrob].
-Furthermore, peptide or protein intensities can be scaled using the robust scaling method or transformed using _quantile_ normalization or _vsn_ to remove systematic differences among samples and heteroscedasticity.
-In this respect, _prolfqua_ can be compared with other R packages such _DEP_ [@DEP2018] or POMA [@POMA2021] which support the entire LFQ analysis pipeline.
+Other means of modelling missing information in proteomics data using the Hurdle model is discussed by [@goeminne2020msqrob], while the R package `r BiocStyle::Biocpkg("proDA")` models missingness using probabilistic dropout models [@bioxrvproDA2020].
+
+Furthermore, the functionality of `r BiocStyle::Githubpkg("fgcz/prolfqua")` also includes methods specific to proteomics data.
+For example, we implemented strategies to estimate protein intensities from peptide intensities: top N, Tukey's median polish, robust linear models [@Grossmann2010, @goeminne2020msqrob].
+Furthermore, peptide or protein intensities can be scaled using the robust scaling method or transformed using _quantile_ normalization or `r BiocStyle::Biocpkg("vsa")` to remove systematic differences among samples and heteroscedasticity.
+In this respect, `r BiocStyle::Githubpkg("fgcz/prolfqua")` can be compared with other R packages such `r BiocStyle::Biocpkg("DEP")` [@DEP2018] or `r BiocStyle::Biocpkg("POMA")` [@POMA2021] which support the entire LFQ analysis pipeline.
 However, we also implemented functionality to streamline the benchmarking of the implemented methods.
-We use the Ionstar [@shen2018ionstar] dataset to benchmark the analysis pipeline and modeling methods implemented within _prolfqua_ and to compare our results with those of _MSstats_ and _proDA_.
+We use the Ionstar [@shen2018ionstar] dataset to benchmark the analysis pipeline and modeling methods implemented within `r BiocStyle::Githubpkg("fgcz/prolfqua")` and to compare our results with those of `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`.
 

### 2022-03-18 inst/MyArticle/paper.Rmd
Collect Abbreviations of Introduction

@@
 
-### P-value moderation
+### $p$-value moderation
 
@@
 
+# Acknoledgements
+
+The authors thank the technology platform fund (TPF) of the University of Zurich.
+
+# Abbreviations
+
+|       |                                                   |
+|:------|:--------------------------------------------------|
+|ESI-MS |Electro-Spray-Ionization Mass Spectrometry         |
+|FDR    |False Discovery Rate                               |
+|HPLC   |High-Pressure Liquid Chromatography                |
+|LC-MS  |Liquid Chromatography followed by Mass Spectrometry|
+|LFQ    |Label-Free Quantification                          |
+|OO     |Object-Oriented                                    |
+|MS     |mass spectrometry                                  |
+
+# References {-}
+
+<div id="refs"></div>
+
 # Supplement
@@
 
-# Acknoledgements
+## Session information
 
-The authors thank the technology platform fund of the University of Zurich.
-
-# References
+```{r sessioninfo, echo=FALSE}
+sessionInfo()
+```

### 2022-03-18 inst/MyArticle/paper.Rmd
Homogenize math notations
- abstract
- math notation,e.g., eqnarray t-statistics, p-statistics

--- 2022-03-18 inst/MyArticle/paper.Rmd
@@
 ---
-title: prolfqua - R-package for Analysing Proteomics Label Free Quantification Experiments
+title: prolfqua - An R-package for Analysing Proteomics Label Free Quantification Experiments
 author: 
@@
   - num: 1
-    address: Functional Genomics Center Zurich (https://www.fgcz.ch/)
+    address: Functional Genomics Center Zurich (FGCZ) - Swiss Federal Institute of Technology in Zurich and University of Zurich
   - num: 2
-    address: Swiss Institute of Bioinformatics (https://www.sib.swiss/)
+    address: Swiss Institute of Bioinformatics (SIB)
 output: 
@@
   chunk_output_type: console
+abstract: |
+  Protein mass spectrometry is widely used for quantitative proteomics
+  studies and in particular for relative protein quantification.
+  Nevertheless, there is still a need for a flexible and easy-to-use, 
+  transparently supporting a variety of well principled statistical
+  inference procedures data analysis application programmer interface
+  in R. The `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package integrates
+  essential steps of the mass spectrometry-based label-free quantification
+  data analysis workflow as: quality control, data normalization, protein
+  aggregation, statistical modeling, hypothesis testing, and sample size
+  estimation.
+  We use `r BiocStyle::Githubpkg("fgcz/prolfqua")` to visualize and model
+  simple experimental designs with a single explanatory variable and complex
+  experiments with multiple factors.
+  Furthermore, we use `r BiocStyle::Githubpkg("fgcz/prolfqua")` to benchmark
+  data acquisition, data preprocessing or data modeling methods. Furthermore,
+  we apply `r BiocStyle::Githubpkg("fgcz/prolfqua")` to develop highly
+  customizable, visually appealing, and interactive data analysis reports
+  in PDF or HTML format.
+  The R package is available through https://github.com/fgcz/prolfqua.
 ---
@@
 ```{r setup, include=FALSE}
-
+## See ACS guidelines [https://publish.acs.org/publish/author_guidelines?coden=ancham#manuscript_text_components]
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 7, fig.height = 5)
-
 ```
 
-See ACS guidelines [https://publish.acs.org/publish/author_guidelines?coden=ancham#manuscript_text_components]
-
-# Abstract (80–250 words)
-
-Protein mass spectrometry is widely used for quantitative proteomics studies and in particular for relative protein quantification.
-Nevertheless, there is still a need for a flexible and easy-to-use, transparently supporting a variety of well principled statistical inference procedures data analysis application programmer interface in R.
-The _prolfqua_ R package integrates essential steps of the mass spectrometry-based label-free quantification data analysis workflow: quality control, data normalization, protein aggregation, statistical modeling, hypothesis testing, and sample size estimation.
-We use _prolfqua_ to visualize and model simple experimental designs with a single explanatory variable and complex experiments with multiple factors.
-Furthermore, we use _prolfqua_ to benchmark data acquisition, data preprocessing or data modeling methods.
-Furthermore, we use _prolfqua_ to develop highly customizable, visually appealing, and interactive data analysis reports in pdf or HTML format.
-The R package is available from https://github.com/fgcz/prolfqua.
+\newpage
 
@@
 
-Label-Free Quantification (LFQ) Proteomics Experiments enable monitoring relative abundances of thousands of proteins in biological samples.
+Label-Free Quantification (LFQ) proteomics experiments enable monitoring relative abundances of thousands of proteins in biological samples.
 For most studies, parallel group designs are used, e.g., treatment vs.
@@
 
-Models in `r BiocStyle::Githubpkg("prolfqua")` are specified using R’s linear model and mixed model formula interface.
+Models in `r BiocStyle::Githubpkg("fgcz/prolfqua")` are specified using R’s linear model and mixed model formula interface.
 Therefore, knowledge of the R regression model infrastructure [@faraway2016extending,@MASS2002] is an advantage when using our package.
-Furthermore, this glass box approach should make it easy to reimplement an analysis which was performed with `r BiocStyle::Githubpkg("prolfqua")` using base R or other programming languages by reading the analysis script.
+Furthermore, this glass box approach should make it easy to reimplement an analysis which was performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` using base R or other programming languages by reading the analysis script.
 However, acknowledging the R formula interface’s complexity to non-statisticians and the popularity of `r BiocStyle::Biocpkg("MSstats")`, we provide the functionality to suggest a model formula from the sample annotation provided in tabular form similarily to `r BiocStyle::Biocpkg("MSstats")` [@Choi2014].
@@
 This makes it easy to reproduce an analysis performed with _prolfqua_ in other statistical programming languages.
-We implemented features specific to high throughput experiments, such as the experimental Bayes variance and p-value moderation, which utilizes the parallel structure of the protein measurements and the analysis [@Ritchie2015].
+We implemented features specific to high throughput experiments, such as the experimental Bayes variance and $p$-value moderation, which utilizes the parallel structure of the protein measurements and the analysis [@Ritchie2015].
 We also compute probabilities of differential protein regulation based on peptide level models [@Suomi2017bEnhanced].
@@
 (ref:ContrastUML) UML diagram of modeling and contrast related classes.
-Different strategies e.g.
-_lm_, _lmer_, and _glm_ can be used to fit various types of models using a model builder (see code).
+Different strategies, e.g., _lm_, _lmer_, and _glm_ can be used to fit various types of models using a model builder (see code).
 All classes computing contrasts implement the _Contrast_ interface.
@@
 
-$$s_p^2=\frac{\sum_{i=1}^k (n_i - 1)s_i^2}{\sum_{i=1}^k (n_i - 1)},$$
+\begin{equation} 
+s_p^2=\frac{\sum_{i=1}^k (n_i - 1)s_i^2}{\sum_{i=1}^k (n_i - 1)}
+\end{equation}
 
@@
 
-$$
+\begin{equation} 
 s = \sqrt{\frac{2n_g s_p^2}{n}},
-$$
+\end{equation}
 
@@
 
-$\tilde{t}_{pj} = \frac{t_{pj} s_p}{\tilde{s}_p}$
+\begin{equation} 
+\tilde{t}_{pj} = \frac{t_{pj} s_p}{\tilde{s}_p}
+\end{equation}
 
@@
 
-To summarize peptide level models to protein models we did applied the method suggested by [@Suomi2017bEnhanced], that is to use the median scaled p-value of the peptide models and cumulative distribution function of the beta distribution function to determine a regulation probability of the protein.
+To summarize peptide level models to protein models we did applied the method suggested by [@Suomi2017bEnhanced], that is to use the median scaled $p$-value of the peptide models and cumulative distribution function of the beta distribution function to determine a regulation probability of the protein.
 
-To obtain the median p-value of the protein, we first rescaled the peptide p-values by taking the direction of the fold-change $\hat \beta$ into account, i.e.:
+To obtain the median $p$-value of the protein, we first rescaled the peptide $p$-values by taking the direction of the fold-change $\hat \beta$ into account, i.e.:
 
@@
 
-$$
+\begin{equation}
 \tilde{p} = 1 - |\tilde{p}_{s}|
-$$
-Because we used the median as the i-th order statistic  $i = n/2 +0.5$.
-Therefore, $\gamma = i = n/2 + 0.5$ and $\delta = n - i + 1 = n - (n/2 + 0.5) + 1 = n/2 + 0.5 = \gamma$ are used to parameterize the CDF of the Beta distribution.
+\end{equation}
+
+Because we used the median as the i-th order statistic  $i = \frac{n}{2} + 0.5$.
+Therefore, $\gamma = i = \frac{n}{2} + 0.5$ and $\delta = n - i + 1 = n - (\frac{n}{2} + 0.5) + 1 = \frac{n}{2} + 0.5 = \gamma$ are used to parameterize the CDF of the Beta distribution.
 We implemented this method in the class `ContrastROPECA` (Figure \@ref(fig:ContrastUML)).
@@
 
-(ref:exampleContrasts) Left panel - histogram of p-values. Right panel - volcano plots.
+(ref:exampleContrasts) Left panel - histogram of $p$-values. Right panel - volcano plots.
 
@@
 the false discovery proportion FDP.
-We use it to study how well linear, mixed effect models or p-value moderation model quantitative mass spectrometric high throughput experiments and compare it with results produced by MSstats [@MSstats2014] and by proDA [@bioxrvproDA2020].
+We use it to study how well linear, mixed effect models or $p$-value moderation model quantitative mass spectrometric high throughput experiments and compare it with results produced by MSstats [@MSstats2014] and by proDA [@bioxrvproDA2020].
 Figure \@ref(fig:FDRfdp)
@@
 We can use various statistics to examine the alternative hypothesis $\beta \ne 0$: the contrast estimate, i.e.
-the $\log_2$ fold-change $\beta$, the $t$-statistic $\frac{\beta}{\sqrt{var(\beta)}}$, or the $p-value$ and moderated $p-value$.
+the $\log_2$ fold-change $\beta$, the $t$-statistic $\frac{\beta}{\sqrt{var(\beta)}}$, or the $p$-value and moderated $p$-value.
 For each statistic and each value of the statistics we then compute a confusion matrix (see Table \@ref(tab:confusionMatrix)).
@@
 
-$$
-\begin{aligned}
+\begin{eqnarray}
 TPR &= \frac{TP}{TP+FN} = \frac{TP}{P}\\
@@
 FDP &= \frac{FP}{TP + FP} = \frac{FP}{R}
-\end{aligned}
-$$
+\end{eqnarray}
 
@@
 Figure \@ref(fig:FDRfdp) left panel, shows how various estimates obtained from the models, i.e.
-$\log_2(FC)$, $t-statistics$ and scaled p-values allow to identify true positives (TP) given a false positive rate.
+$\log_2(FC)$, $t$-statistics and scaled $p$-values allow to identify true positives (TP) given a false positive rate.
 When testing for differential expression, the $\log_2(FC)$ is outperformed by than the $t$-statistic or $p$-value.
@@
 
-There are only minor differences in the p-AUC when using the t-statistics or the scaled p-value  (see Figure \@ref(fig:FDRfdp) right panel).
-In order to compute the confusion matrices based on the p-value we first need to rescale it (see Equation \@ref(eq:scalepvalue)).
-Thus, the p-value is a function of the t-statistics and the degrees of freedom.
-Interestingly, the pAUC increases when using p-values instead of the t-statistics for linear models, while it decreases for mixed effect models.
+There are only minor differences in the $p$-AUC when using the $t$-statistics or the scaled $p$-value  (see Figure \@ref(fig:FDRfdp) right panel).
+In order to compute the confusion matrices based on the $p$-value we first need to rescale it (see Equation \@ref(eq:scalepvalue)).
+Thus, the $p$-value is a function of the $t$-statistics and the degrees of freedom.
+Interestingly, the pAUC increases when using $p$-values instead of the $t$-statistics for linear models, while it decreases for mixed effect models.
 The reason is erroneous denominator degrees of freedom estimation for many proteins in the mixed-effect models.
@@
 
-(ref:FDRfdp) Left panel - Partial area under the ROC curve at $10\%$ FPR for all contrasts and three different statistics: the difference among conditions, the scaled p-value (sign(diff) * p.value) and the t-statics (higher is better).
+(ref:FDRfdp) Left panel - Partial area under the ROC curve at $10\%$ FPR for all contrasts and three different statistics: the difference among conditions, the scaled $p$-value (sign(diff) * p.value) and the $t$-statistics (higher is better).
 The red line average area under the curve.
@@
 Moreover, the estimated fold changes are unbiased and show high diagnostic accuracy.
-Furthermore, the scaled p-values and t-statics are on par with those obtained by the linear model with variance moderation.
+Furthermore, the scaled $p$-values and $t$-statics are on par with those obtained by the linear model with variance moderation.
 Therefore, we plan to integrate the _proDA_ package as an additional modeling option into _prolfqua_ (see Figure \@ref(fig:ContrastUML)).
@@
 Use mixed effect models only if you have to (e.g., repeated measurements).
-If you use fixed-effect linear models, apply variance moderation since this improves the t-statistics and p-value estimates.
+If you use fixed-effect linear models, apply variance moderation since this improves the $t$-statistics and $p$-value estimates.
 If you want to sort your protein lists to perform gene set enrichment analysis, use the $t$-statistic instead of the fold change.
@@
 # Conclusion (tbd)
+
+# Acknoledgements
+
+The authors thank the technology platform fund (TPF) of the University of Zurich.
+
+# Abbreviations
+
+|       |                                                   |
+|:------|:--------------------------------------------------|
+|AUC    |Area Under the Curve                               |
+|ESI-MS |Electro-Spray-Ionization Mass Spectrometry         |
+|FDR    |False Discovery Rate                               |
+|HPLC   |High-Pressure Liquid Chromatography                |
+|LC-MS  |Liquid Chromatography followed by Mass Spectrometry|
+|LFQ    |Label-Free Quantification                          |
+|LOD    |Limit Of Detection                                 |
+|MS     |mass spectrometry                                  |
+|OO     |Object-Oriented                                    |
+
+# References {-}
+
+<div id="refs"></div>
 

### 2022-03-18 inst/MyArticle/paper.Rmd
added citations.

--- 2022-03-18 inst/MyArticle/paper.Rmd
@@
 For most studies, parallel group designs are used, e.g., treatment vs.
-control are used [TODO: cite].
-More recently, more complex experimental designs are implemented with increasing numbers of samples, i.e., factorial designs and longitudinal studies (time series), sometimes with repeated measurements on the same subject [TODO: cite].
+control are used [@de2022apoe2,@laubscher2021baf].
+More recently, more complex experimental designs are implemented with increasing numbers of samples, i.e., factorial designs and longitudinal studies (time series), sometimes with repeated measurements on the same subject [@tan2022proteomic,@meier2021reduced].
 The obtained data can be modeled using linear fixed-, mixed-, or random-effects models [@Bates2015FittingJSS].
@@
    ggplot2::theme(legend.position = "bottom")
+p1
 gridExtra::grid.arrange(p1, p2, ncol = 2)

### 2022-03-18 inst/MyArticle/paper.Rmd

--- 2022-03-18 inst/MyArticle/paper.Rmd
@@
 
-Label-Free Quantification (LFQ) proteomics experiments enable monitoring relative abundances of thousands of proteins in biological samples.
+Label-Free Quantification (LFQ) Proteomics Experiments enable monitoring relative abundances of thousands of proteins in biological samples.
 For most studies, parallel group designs are used, e.g., treatment vs.
-control are used [TODO: cite].
-More recently, more complex experimental designs are implemented with increasing numbers of samples, i.e., factorial designs and longitudinal studies (time series), sometimes with repeated measurements on the same subject [TODO: cite].
+control are used [@de2022apoe2,@laubscher2021baf].
+More recently, more complex experimental designs are implemented with increasing numbers of samples, i.e., factorial designs and longitudinal studies (time series), sometimes with repeated measurements on the same subject [@tan2022proteomic,@meier2021reduced].
 The obtained data can be modeled using linear fixed-, mixed-, or random-effects models [@Bates2015FittingJSS].

### 2022-03-18 inst/MyArticle/paper.Rmd
Add sticker

@@
 ggplot2::ggplot(x,ggplot2::aes( x = diff.imp, y = diff.proda) ) + ggplot2::geom_point() + ggplot2::facet_wrap(~contrast + species) + ggplot2::geom_hline(yintercept = 0, color="red") + ggplot2::geom_vline(xintercept = 0, color="red")
-
 ```
@@
 ```
+
+## Package sticker
+
+```{r proLFQuaSticker, echo=TRUE, out.height="5cm", eval=TRUE, fig.cap="Sticker maintainer: Witold E.
+Wolski; License: Creative Commons Attribution CC-BY.
+Feel free to share and adapt, but don't forget to credit the author."}
+
+file.path(system.file(package = "prolfqua"), 
+          "Figures/hexStickerProlfqua.png") |>
+  knitr::include_graphics()
+```

### 2022-03-18 inst/MyArticle/paper.Rmd
Replace ROOPECA with PECA link

@@
 
-Several packages exist to model mass spectrometry LFQ experiments, e.g., `r BiocStyle::Biocpkg("limma")` [@Ritchie2015], `r BiocStyle::Biocpkg("MSstats")` [@Choi2014], _ROPECA_[@Suomi2017bEnhanced],  `r BiocStyle::Biocpkg("msqrob2")` [@Goeminne2016] or `r BiocStyle::Biocpkg("proDA")` [@bioxrvproDA2020], to mention some, all implemented in the R.
+Several packages exist to model mass spectrometry LFQ experiments, e.g., `r BiocStyle::Biocpkg("limma")` [@Ritchie2015], `r BiocStyle::Biocpkg("MSstats")` [@Choi2014], `r BiocStyle::Biocpkg("PECA")` `r BiocStyle::Biocpkg("PECA")` [@Suomi2017bEnhanced],  `r BiocStyle::Biocpkg("msqrob2")` [@Goeminne2016] or `r BiocStyle::Biocpkg("proDA")` [@bioxrvproDA2020], to mention some, all implemented in the R.
 Each of them has some unique features.
@@
 At the same time, `r BiocStyle::Biocpkg("limma")` allows to specify a linear model formula and implements the empirical Bayes variance shrinkage method.
-The R-package _ROPECA_, performs a roll-up of peptide level fold change and $p$-value estimates, obtained from `r BiocStyle::Biocpkg("limma")`, to protein level estimates.
+The R-package `r BiocStyle::Biocpkg("PECA")`, performs a roll-up of peptide level fold change and $p$-value estimates, obtained from `r BiocStyle::Biocpkg("limma")`, to protein level estimates.
 Furthermore, `r BiocStyle::Biocpkg("msqrob2")` implements two models: robust linear models fitted to protein intensities and robust ridge regression fitted to peptide intensities and combines them with empirical Bayes variance shrinkage known from `r BiocStyle::Biocpkg("limma")`.
@@
 However, all of them fit some type of linear model either to peptide or protein intensities, determine contrasts among conditions, and afterward, some of them apply empirical Bayes variance shrinkage.
-Therefore, the revised goal was to provide a modular object oriented (OO) design, with R linear models as a core, where features such as $p$-value aggregation (_ROPECA_) or variance shrinkage could be easily enabled (see Figure \@ref(fig:ContrastUML)).
+Therefore, the revised goal was to provide a modular object oriented (OO) design, with R linear models as a core, where features such as $p$-value aggregation, e.g., `r BiocStyle::Biocpkg("PECA")` or variance shrinkage could be easily enabled (see Figure \@ref(fig:ContrastUML)).
 
@@
 The protein level linear model was fitted to the protein intensities inferred from the peptide intensities using the Tukey's median polish.
-The mixed effect linear model, and the ROPECA model were fitted to the peptide intensities.
+The mixed effect linear model, and the `r BiocStyle::Biocpkg("PECA")` model were fitted to the peptide intensities.
 The details of the _prolfqua_ benchmark implementation are available online  [BenchmarkingIonstarData](https://wolski.github.io/prolfqua/articles/BenchmarkingIonstarData.html).
@@
 In the case of _MSstats_, we observe a high proportion of false discoveries for small $FDR$ values.
-In the case of _ROPECA_, the FDR estimates, obtained when Benjamini-Hochberg correcting the regulation probabilities, strongly overestimate the $FDP$.
+In the case of `r BiocStyle::Biocpkg("PECA")`, the FDR estimates, obtained when Benjamini-Hochberg correcting the regulation probabilities, strongly overestimate the $FDP$.
 
@@
 Furthermore, peptides are stronger affected by missing values, reducing the number of contrasts we could estimate for the dataset (see Figure \@ref(fig:FDRfdp) left panel).
-To adress the issue of missing observation we could either run proDA or the imputation of group averages method discussed here and afterwards apply ROPECA.
+To adress the issue of missing observation we could either run proDA or the imputation of group averages method discussed here and afterwards apply `r BiocStyle::Biocpkg("PECA")`.
 

### 2022-03-18 inst/MyArticle/paper.Rmd
Cosmetic text edits

@@
 For most studies, parallel group designs are used, e.g., treatment vs.
-control are used [@de2022apoe2,@laubscher2021baf].
-More recently, more complex experimental designs are implemented with increasing numbers of samples, i.e., factorial designs and longitudinal studies (time series), sometimes with repeated measurements on the same subject [@tan2022proteomic,@meier2021reduced].
+control are used [@de2022apoe2, @laubscher2021baf].
+More recently, more complex experimental designs are implemented with increasing numbers of samples, i.e., factorial designs and longitudinal studies (time series), sometimes with repeated measurements on the same subject [@tan2022proteomic, @meier2021reduced].
 The obtained data can be modeled using linear fixed-, mixed-, or random-effects models [@Bates2015FittingJSS].
@@
 We store all the data needed for analysis in a single data frame in a tidy table, i.e., every column is a variable, every row is an observation, every cell is a single value[@tidydata2014].
-Using an __R6__[@R6cite] configuration object, we specify what variable is in which column, making it easy to integrate new inputs in _prolfqua_ if provided as a tidy table.
-For example, to visualize tidy _Spectronaut_, or _Skyline_ outputs, or data in _MSStats_ format, only a few lines of code to update the _prolfqua_ configuration are needed.
+Using an __R6__[@R6cite] configuration object, we specify what variable is in which column, making it easy to integrate new inputs in `r BiocStyle::Githubpkg("fgcz/prolfqua")` if provided as a tidy table.
+For example, to visualize tidy _Spectronaut_, or _Skyline_ outputs, or data in _MSStats_ format, only a few lines of code to update the `r BiocStyle::Githubpkg("fgcz/prolfqua")` configuration are needed.
 For popular software like _MaxQuant_ or _MSFragger_, which stores the same variable (e.g., intensity) in multiple columns, one for each sample, we implemented methods that transform the data into tidy tables.
@@
 __R__'s formula interface for linear models is flexible, widely used, and well documented.
-This makes it easy to reproduce an analysis performed with _prolfqua_ in other statistical programming languages.
+This makes it easy to reproduce an analysis performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` in other statistical programming languages.
 We implemented features specific to high throughput experiments, such as the experimental Bayes variance and $p$-value moderation, which utilizes the parallel structure of the protein measurements and the analysis [@Ritchie2015].
@@
 The mixed effect linear model, and the `r BiocStyle::Biocpkg("PECA")` model were fitted to the peptide intensities.
-The details of the _prolfqua_ benchmark implementation are available online  [BenchmarkingIonstarData](https://wolski.github.io/prolfqua/articles/BenchmarkingIonstarData.html).
+The details of the `r BiocStyle::Githubpkg("fgcz/prolfqua")` benchmark implementation are available online  [BenchmarkingIonstarData](https://wolski.github.io/prolfqua/articles/BenchmarkingIonstarData.html).
 
@@
 
-The Benchmark functionality of _prolfqua_ includes ROC curves and computes partial areas under those curves (pAUC) and other scores, e.g.
+The Benchmark functionality of `r BiocStyle::Githubpkg("fgcz/prolfqua")` includes ROC curves and computes partial areas under those curves (pAUC) and other scores, e.g.
 the false discovery proportion FDP.
@@
 coli_ proteins, we know that the difference between dilutions should be $\beta \ne 0$.
-To compare the performance of the various methods implemented in _prolfqua_ we use only the contrasts resulting in small fold-changes $\beta = 1.2,1.25,1.3(3),1.5$.
+To compare the performance of the various methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` we use only the contrasts resulting in small fold-changes $\beta = 1.2,1.25,1.3(3),1.5$.
 
@@
 Figure \@ref(fig:FDRfdp) right panel draws on the horizontal axis the FDR determined from the model, and on the vertical axis, the FDP obtained from the confusion matrix.
-We see that most of the lines are below the diagonal, indicating that the FDR estimates obtained with _prolfqua_ are modestly conservative for this benchmark dataset provide.
+We see that most of the lines are below the diagonal, indicating that the FDR estimates obtained with `r BiocStyle::Githubpkg("fgcz/prolfqua")` are modestly conservative for this benchmark dataset provide.
 In the case of _MSstats_, we observe a high proportion of false discoveries for small $FDR$ values.
@@
 Furthermore, the scaled $p$-values and $t$-statics are on par with those obtained by the linear model with variance moderation.
-Therefore, we plan to integrate the _proDA_ package as an additional modeling option into _prolfqua_ (see Figure \@ref(fig:ContrastUML)).
+Therefore, we plan to integrate the _proDA_ package as an additional modeling option into `r BiocStyle::Githubpkg("fgcz/prolfqua")` (see Figure \@ref(fig:ContrastUML)).
 
@@
 
-_prolfqua_ allows for considerable flexibility to model your experiments.
+`r BiocStyle::Githubpkg("fgcz/prolfqua")` allows for considerable flexibility to model your experiments.
 Various types of models are available, and the contrast specification is consistent and intuitive for all models.
 The benchmarking results are comparable with other methods.
-The modular design of _prolfqua_, allows for customizing the analysis or adding new features, e.g., generalized linear models (_glm_'s) for missing data modeling or robust linear models (_rlm_'s), quickly.
+The modular design of `r BiocStyle::Githubpkg("fgcz/prolfqua")`, allows for customizing the analysis or adding new features, e.g., generalized linear models (_glm_'s) for missing data modeling or robust linear models (_rlm_'s), quickly.
 
-Using the tidy table to model the data ensures interoperability with other proteomics-related packages that manage their data with tidy-tables, e.g., _protti_.
-Furthermore, the use of `R6` classes, which encapsulate the configuration and the data, allows writing very concise code (no function arguments needed), and auto-completion support in the Rstudio editor makes it easy for novices to explore _prolfquas_ functionality.
+Using the tidy table to model the data ensures interoperability with other proteomics-related packages that manage their data with tidy-tables, e.g., `r BiocStyle::CRANpkg("protti")`.
+Furthermore, the use of `R6` classes, which encapsulate the configuration and the data, allows writing very concise code (no function arguments needed), and auto-completion support in the Rstudio editor makes it easy for novices to explore `r BiocStyle::Githubpkg("fgcz/prolfqua")`s functionality.
 
-In summary _prolfqua_ is an easy-to-use and feature rich R package to analyse quantitative mass spectrometric data, report results, and benchmark MS software and statistical methods.
+In summary `r BiocStyle::Githubpkg("fgcz/prolfqua")` is an easy-to-use and feature rich R package to analyse quantitative mass spectrometric data, report results, and benchmark MS software and statistical methods.
 We provide more details at the website https://wolski.github.io/prolfqua/.
@@
 |OO     |Object-Oriented                                    |
+|UML    |Unified Modeling Language                          |
 

### 2022-03-18 inst/MyArticle/paper.Rmd
Edits on code snippets

@@
 
+The code snippets in this section demonstrate the basic usage of the `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package.
 Peptide intensities are $\log_2$ transformed and robust z-score scaled but then rescaled to preserves the original range of the data using the method `robscale`.
@@
 ```{r prepro,  echo = TRUE , fig.cap="(ref:scaling)", out.width = '90%', fig.width = 8, fig.height = 8}
-library(prolfqua)
-d <- prolfqua_data('data_ionstar')$filtered()
-lfqd <- LFQData$new(d$data, d$config) 
+## load peptide level data
+d <- prolfqua::prolfqua_data('data_ionstar')$filtered()
+
+## create R6 obejct
+lfqd <- prolfqua::LFQData$new(d$data, d$config)
+
+##  transform intensities
 t <- lfqd$get_Transformer()
+
 lfqd <-  t$log2()$robscale()$lfq
+
+## plotting
 pl <- lfqd$get_Plotter()
@@
   ggplot2::labs(tag = "A") + ggplot2::theme(legend.position = "none")
+
+## infer protein intensities from peptide intens
 agr <- lfqd$get_Aggregator()
@@
 
+## plotting
 p_b <- agr$plot()$plots[[1]] + ggplot2::labs(tag = "B")
@@
 ggpubr::ggarrange(p_a, p_b, p_c, p_d)
-
 ```
@@
 )
-# fit model
+## fit model
 lmmodel <- paste(lfqp$intensity_column()," ~ dilution.")
-modelFunction <- strategy_lm( lmmodel, model_name = "lm")
-models <- build_model(lfqp, modelFunction)
+modelFunction <- prolfqua::strategy_lm( lmmodel, model_name = "lm")
+models <- prolfqua::build_model(lfqp, modelFunction)
 p0 <- models$anova_histogram()
-# compute contrasts from linear model and with imputation
+
+## compute contrasts from linear model and with imputation
 contr <- prolfqua::Contrasts$new(models, contrasts) |> 
   prolfqua::ContrastsModerated$new()
+
 conI <- prolfqua::ContrastsSimpleImpute$new( lfqp, contrasts) |> 
   prolfqua::ContrastsModerated$new()
+
+# TODO(wew): give tmp a meaningful name
 tmp <- prolfqua::addContrastResults(prefer = contr, add = conI)
-# visualize results
+
+## visualize results
 pl <- tmp$merged$get_Plotter()
@@
    ggplot2::theme(legend.position = "bottom")
-p1
+
 gridExtra::grid.arrange(p1, p2, ncol = 2)
-
 ```
@@
 
-The peptide intensities from the _peptide.txt_ file were $log_2$ transformed, and subsequent robust z-score transformation, where median and the mean average deviation (mad) were obtained from the human proteins only.
+The peptide intensities from the MaxQuant _peptide.txt_ file were $log_2$ transformed, and subsequent robust z-score transformation, where median and the mean average deviation (mad) were obtained from the human proteins only.
 We removed _one hit wonders_, i.e.
@@
 
-# Supplement
+# Supplements
+
+## Creating a prolfqa configuration
+
+The following code snipet demomnstarte the usage from MSFragger ouput.
 

### 2022-03-18 inst/MyArticle/paper.Rmd
Edit authors

@@
 author: 
-  - name: Witold Wolski
-    affil: 1; 2
-    orcid: '0000-0002-6468-120X'
-  - name: Paolo Nanni
-    affil: 1
-    orcid: '0000-0001-8429-3557'
-  - name: Jonas Grossmann
-    affil: 1; 2
-    orcid: '0000-0002-6899-9020'
-  - name: Maria D'Errico
-    affil: 1; 2
-    orcid: '0000-0001-9467-9058'
-  - name: Ralph Schlapbach
-    affil: 1
-    orcid: '0000-0002-7488-4262'
-  - name: Christian Panse
-    affil: 1; 2
-    orcid: '0000-0003-1975-3064'
-affiliation:
-  - num: 1
-    address: Functional Genomics Center Zurich (FGCZ) - Swiss Federal Institute of Technology in Zurich and University of Zurich
-  - num: 2
-    address: Swiss Institute of Bioinformatics (SIB)
+  - Witold E. Wolski^[Functional Genomics Center Zurich (FGCZ) - Swiss Federal Institute of Technology in Zurich \| University of Zurich]
+  - Paolo Nanni$^*$
+  - Jonas Grossmann$^*$
+  - Maria d'Errico$^*$
+  - Ralph Schlapbach$^*$
+  - Christian Panse$^*$
 output: 

### 2022-03-19 inst/MyArticle/paper.Rmd
Edits in Method and Discussion
- code snippets
- text
- add links to packages
- add TODOs

@@
 author: 
-  - Witold E. Wolski^[Functional Genomics Center Zurich (FGCZ) - Swiss Federal Institute of Technology in Zurich \| University of Zurich]
+  - Witold E. Wolski^[[Functional Genomics Center Zurich (FGCZ)](https://fgcz.ch) - Swiss Federal Institute of Technology in Zurich \| University of Zurich]
   - Paolo Nanni$^*$
@@
 Label-Free Quantification (LFQ) Proteomics Experiments enable monitoring relative abundances of thousands of proteins in biological samples.
-For most studies, parallel group designs are used, e.g., treatment vs.
-control are used [@de2022apoe2, @laubscher2021baf].
+For most studies, parallel group designs are used, e.g., treatment versus control are used [@de2022apoe2, @laubscher2021baf].
 More recently, more complex experimental designs are implemented with increasing numbers of samples, i.e., factorial designs and longitudinal studies (time series), sometimes with repeated measurements on the same subject [@tan2022proteomic, @meier2021reduced].
@@
 
-Several packages exist to model mass spectrometry LFQ experiments, e.g., `r BiocStyle::Biocpkg("limma")` [@Ritchie2015], `r BiocStyle::Biocpkg("MSstats")` [@Choi2014], `r BiocStyle::Biocpkg("PECA")` `r BiocStyle::Biocpkg("PECA")` [@Suomi2017bEnhanced],  `r BiocStyle::Biocpkg("msqrob2")` [@Goeminne2016] or `r BiocStyle::Biocpkg("proDA")` [@bioxrvproDA2020], to mention some, all implemented in the R.
+Several packages exist to model mass spectrometry LFQ experiments, e.g., `r BiocStyle::Biocpkg("limma")` [@Ritchie2015], `r BiocStyle::Biocpkg("MSstats")` [@Choi2014], `r BiocStyle::Biocpkg("PECA")` [@Suomi2017bEnhanced],  `r BiocStyle::Biocpkg("msqrob2")` [@Goeminne2016] or `r BiocStyle::Biocpkg("proDA")` [@bioxrvproDA2020], to mention some, all implemented in the R.
 Each of them has some unique features.
@@
 
-# Implementation & Methods (Experimental Section)
+# Methods
 
@@
 Using an __R6__[@R6cite] configuration object, we specify what variable is in which column, making it easy to integrate new inputs in `r BiocStyle::Githubpkg("fgcz/prolfqua")` if provided as a tidy table.
-For example, to visualize tidy _Spectronaut_, or _Skyline_ outputs, or data in _MSStats_ format, only a few lines of code to update the `r BiocStyle::Githubpkg("fgcz/prolfqua")` configuration are needed.
-For popular software like _MaxQuant_ or _MSFragger_, which stores the same variable (e.g., intensity) in multiple columns, one for each sample, we implemented methods that transform the data into tidy tables.
-Relying on the tidy data table enabled us to easily interface with many data manipulation, visualization, and modeling methods implemented in base __R__[@RCoreTeam2021] and the tidyverse[@tidyverse2019].
+For example, to visualize tidy _Spectronaut_, or _Skyline_ outputs, or data in `r BiocStyle::Biocpkg("MSstats")` format, only a few lines of code are needed to update the `r BiocStyle::Githubpkg("fgcz/prolfqua")` configuration.
+For popular software like _MaxQuant_ or _MSFragger_, which stores the same variable, e.g., intensity, in multiple columns, one for each sample, we implemented methods that transform the data into tidy tables.
+Relying on the tidy data table enable us to easily interface with many data manipulation, visualization, and modeling methods implemented in base __R__[@RCoreTeam2021] and the tidyverse[@tidyverse2019].
 We use __R6__ classes to structure the functionality of the package (see Figure \@ref(fig:LFQData) and Figure \@ref(fig:ContrastUML)).
-__R6__ classes are well supported (e.g.
-auto-completion) and easy to use (OOP).
+__R6__ classes are well supported, e.g.
+by auto-completion features, and easy to use (OOP).
 
@@
 ```{r LFQData, echo=FALSE, fig.cap="(ref:LFQData)", out.width = '90%'}
-
 knitr::include_graphics("LFQDataUML.png")
-
 ```
@@
 __R__ linear model and linear mixed effect models allow modeling parallel designs, repeated measurements, factorial designs, and many more.
-__R__'s formula interface for linear models is flexible, widely used, and well documented.
+__R__'s formula interface for linear models is flexible, widely used, and well documented (TODO(wew): cite the docu).
 This makes it easy to reproduce an analysis performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` in other statistical programming languages.
@@
 We also compute probabilities of differential protein regulation based on peptide level models [@Suomi2017bEnhanced].
-We use R6 to model the statistical modeling functionality in `prolfqua` (see Figure \@ref(fig:ContrastUML)).
+We use R6 to model the statistical modeling functionality in `r BiocStyle::Biocpkg("prolfqua")` (see Figure \@ref(fig:ContrastUML)).
 
@@
 
-## Benchmark dataset
+## Dataset
 
-Ionstar dataset, give Pride ids, describe MaxQuant settings and version. Give reference to Experiment hub and the _prolfquaData_ package.
+TODO(wew): Ionstar dataset, give Pride ids, describe MaxQuant settings and version. Give reference to Experiment hub and the _prolfquaData_ package.
 
@@
 
-For estimating contrasts from mixed effects models we used the function `contest` implemented in the R package `lmerTest` [@Kuznetsova2017lmerTest] and used the Satterthwaite method to estimate the denominator degrees of freedom.
+For estimating contrasts from mixed effects models we used the function `contest` implemented in the R package `r BiocStyle::CRANpkg("lmerTest")` [@Kuznetsova2017lmerTest] and used the Satterthwaite method to estimate the denominator degrees of freedom.
 These methods are implemented in the class `Contrast` (see Figure \@ref(fig:ContrastUML))
@@
 
-If there are no observation in a group we assume that there are no observations because of the limit of detection (LOD).
+If there are no observation in a group, we assume that there are no observations because of the limit of detection (LOD).
 Furthermore, we assume that the variance of the protein is constant in all the conditions.
@@
 
-To summarize peptide level models to protein models we did applied the method suggested by [@Suomi2017bEnhanced], that is to use the median scaled $p$-value of the peptide models and cumulative distribution function of the beta distribution function to determine a regulation probability of the protein.
+To summarize peptide level models to protein models, we applied the method suggested by [@Suomi2017bEnhanced]  to use the median scaled $p$-value of the peptide models and cumulative distribution function of the beta distribution function to determine a regulation probability of the protein.
 
-To obtain the median $p$-value of the protein, we first rescaled the peptide $p$-values by taking the direction of the fold-change $\hat \beta$ into account, i.e.:
+To obtain the median $p$-value of a protein, we first rescaled the peptide $p$-values by taking the direction of the fold-change $\hat \beta$ into account, i.e.:
 
@@
 Because we used the median as the i-th order statistic $i = \frac{n}{2} + 0.5$.
-Therefore, $\gamma = i = \frac{n}{2} + 0.5$ and $\delta = n - i + 1 = n - (\frac{n}{2} + 0.5) + 1 = \frac{n}{2} + 0.5 = \gamma$ are used to parameterize the CDF of the Beta distribution.
+Therefore, $\gamma = i = \frac{n}{2} + 0.5$ and $\delta = n - i + 1 = n - (\frac{n}{2} + 0.5) + 1 = \frac{n}{2} + 0.5 = \gamma$ are used to parameterize the (TODO(wew): what is CDF) CDF of the Beta distribution.
 We implemented this method in the class `ContrastROPECA` (Figure \@ref(fig:ContrastUML)).
 
-# Example Analysis
+## Usage
 
-The code snippets in this section demonstrate the basic usage of the `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package.
+The code snippets in this section demonstrate the basic usage of the `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package. As input data we use Ionstar dataset.
 Peptide intensities are $\log_2$ transformed and robust z-score scaled but then rescaled to preserves the original range of the data using the method `robscale`.
@@
 
-# Benchmarking
+## Benchmarking
 
-The peptide intensities from the MaxQuant _peptide.txt_ file were $log_2$ transformed, and subsequent robust z-score transformation, where median and the mean average deviation (mad) were obtained from the human proteins only.
+The peptide intensities from the _MaxQuant_ _peptide.txt_ file were $log_2$ transformed, and subsequent robust z-score transformation, where median and the mean average deviation (mad) were obtained from the human proteins only.
 We removed _one hit wonders_, i.e.
@@
 
-For benchmarking of MSstats we used the defaults specified in the [Benchmarking_MSstats](https://wolski.github.io/prolfqua/articles/)
+For benchmarking of `r BiocStyle::Biocpkg("MSstats")` we used the defaults specified in the [Benchmarking_MSstats](https://wolski.github.io/prolfqua/articles/)
 
-The Benchmark functionality of `r BiocStyle::Githubpkg("fgcz/prolfqua")` includes ROC curves and computes partial areas under those curves (pAUC) and other scores, e.g.
-the false discovery proportion FDP.
-We use it to study how well linear, mixed effect models or $p$-value moderation model quantitative mass spectrometric high throughput experiments and compare it with results produced by MSstats [@MSstats2014] and by proDA [@bioxrvproDA2020].
-Figure \@ref(fig:FDRfdp)
+The Benchmark functionality of `r BiocStyle::Githubpkg("fgcz/prolfqua")` includes ROC curves and computes partial areas under those curves (pAUC) and other scores, e.g., the false discovery proportion FDP.
+We use those plots to study how well linear, mixed effect models or $p$-value moderation models quantitative mass spectrometric high throughput experiments and compare them with results produced by `r BiocStyle::Biocpkg("MSstats")` [@MSstats2014] and `r BiocStyle::Biocpkg("proDA")` [@bioxrvproDA2020] (see also Figure \@ref(fig:FDRfdp))
 
-The IonStar  [@shen2018ionstar] benchmark dataset contains _H.
+The IonStar [@shen2018ionstar] dataset contains _H.
 sapiens_ proteins with constant concentrations and _E.
@@
 We know that for _H.
-sapiens_ proteins the difference $\beta$ between two dilutions should be $\beta = 0$ while for _E.
+sapiens_ proteins, the difference $\beta$ between two dilutions should be $\beta = 0$.
+While for _E.
 coli_ proteins, we know that the difference between dilutions should be $\beta \ne 0$.
-To compare the performance of the various methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` we use only the contrasts resulting in small fold-changes $\beta = 1.2,1.25,1.3(3),1.5$.
+To compare the performance of the various methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` we use only the contrasts resulting in small fold-changes $\beta = (1.20, 1.25, 1.30, 1.50)$.
 
@@
 For each statistic and each value of the statistics we then compute a confusion matrix (see Table \@ref(tab:confusionMatrix)).
-From the confusion matrix we obtain measures such as true positive rate (TPR), false positive rate (FPR), or false discovery proportion (FDP) which are given by:
+From the confusion matrix we obtain measures such as true positive rate ($TPR$), false positive rate ($FPR$), or false discovery proportion ($FDP$) which are given by:
 
-```{r confusionMatrix}
-table <- data.frame( c("beta != 0", "beta == 0", "Total"), matrix(c("TP","FP","R","FN","TN","","P","N","m"), ncol = 3, byrow = T))
+```{r knitrConfusionMatrix}
+table <- data.frame( c("beta != 0", "beta == 0", "Total"),
+                     matrix(c("TP", "FP", "R", "FN", "TN", "", "P", "N", "m"),
+                            ncol = 3, byrow = T))
 colnames(table) <- c("Prediction \\ Truth","E.coli", "H.sapiens", "Total")
+
 knitr::kable(table, caption = "Confusion matrix, TP - true positive, FP - false positive, FN - false negative, P - all positive cases (all E.
@@
 sapiens proteins), m- all proteins.")
+```
 
-```
+with
 
@@
 
-```{r}
+```{r readBenchmarkData}
 allBenchmarks <- readRDS("../../inst/Benchresults/allBenchmarks.RDS")
@@
 allBenchmarks <- allBenchmarks[c("benchmark_imputation","benchmark_ProtModerated",  "benchmark_mixedModerated", "benchmark_ropeca","benchmark_merged","benchmark_mssstats","benchmark_proDA"   )]
-
 ```
@@
 ```{r benchmarkROC, fig.cap="(ref:benchmarkROC)", out.width = '90%', fig.width=12, fig.height=10}
-
 ttt <- sapply(allBenchmarks, function(x){x$complete(TRUE)})
@@
 Because the set of contrasts slightly differs for each method, it makes comparing them complicated.
-For instance, we observe that MSstas estimates $16058$ while the mixed effect models estimate the fewest with $15940$.
+For instance, we observe that `r BiocStyle::Biocpkg("MSstats")` estimates $16058$ while the mixed effect models estimate the fewest with $15940$.
 Hence, to conclude that one method shows a better performance, the score, e.g., pAUC, needs to be greater, but also the number of estimated contrasts needs to be equal or larger.
@@
 
-We also benchmark if the $FDR$ obtained from a model is an unbiased estimate of the false discovery proportion $FDP$.
+We also benchmark, if the $FDR$ obtained from a model is an unbiased estimate of the false discovery proportion $FDP$.
 Figure \@ref(fig:FDRfdp) right panel draws on the horizontal axis the FDR determined from the model, and on the vertical axis, the FDP obtained from the confusion matrix.
 We see that most of the lines are below the diagonal, indicating that the FDR estimates obtained with `r BiocStyle::Githubpkg("fgcz/prolfqua")` are modestly conservative for this benchmark dataset provide.
-In the case of _MSstats_, we observe a high proportion of false discoveries for small $FDR$ values.
+In the case of `r BiocStyle::Biocpkg("MSstats")`, we observe a high proportion of false discoveries for small $FDR$ values.
 In the case of `r BiocStyle::Biocpkg("PECA")`, the FDR estimates, obtained when Benjamini-Hochberg correcting the regulation probabilities, strongly overestimate the $FDP$.
@@
 ```{r FDRfdp, fig.cap = "(ref:FDRfdp)", out.width = '90%', fig.width=12, fig.height=10}
-
 dd <- purrr::map_df(allBenchmarks, function(x){res <- x$smc$summary; res$name <- x$model_name;res})
@@
 gridExtra::grid.arrange(p1_pAUC, p3, p4, layout_matrix=hlay)
-
 ```
@@
 
-[Goes in the discussion or introduction, and give references for normalization e.g.
+[TODO(wew): Goes in the discussion or introduction, and give references for normalization e.g.
 VSN paper.] However, an essential parameter, not discussed further, is the normalization of the intensities within the samples to make them comparable.
-For this benchmark dataset, the fraction of regulated proteins (ECOLI) is relatively large.
+For this benchmark dataset, the fraction of regulated proteins (_E.
+coli_) is relatively large.
 Therefore, to calibrate the measurement, we used human proteins only.
@@
 
-[Goes to discussion] The method used to infer protein intensities from more peptide intensities is an additional important factor.
+[TODO(wew): Goes to discussion] The method used to infer protein intensities from more peptide intensities is an additional important factor.
 The original _proDA_ publication uses MaxLFQ [@cox2014accurate] protein estimates.
@@
 
-We assessed the performance of different modeling approaches implemented in prolfqua, MSstats and proDA, using a benchmark dataset whose ground truth was known.
-However, the dataset used has several limitations: first, the design is a parallel-group design, i.e.
-we could not benchmark how well interactions are modelled; secondly, essential sources of variation present in every experiment, such as the biochemical and biological, are not measured.
+We assessed the performance of different modeling approaches implemented in prolfqua, `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`, using a benchmark dataset whose ground truth is known.
+However, the dataset used has several limitations: first, the design is a parallel-group design, i.e., we could not benchmark how well interactions are modelled; secondly, essential sources of variation present in every experiment, such as the biochemical and biological, are not measured.
 The dataset captures only the variance from the chromatography, electro-spray, and mass spectrometric measurement method.
@@
 
-For the fixed effect linear model, the empirical Bayes variance shrinkage, as first suggested by [@Smyth2004linear], gives a consistent improvement compared with the unmoderated estimates (data not shown).
+For the fixed effect linear model, the empirical Bayes variance shrinkage, as suggested by [@Smyth2004linear], consistently improves compared with the unmoderated estimates (data not shown).
 However, there is no such effect for the mixed effect model, most likely because the degrees of freedom are erroneously estimated for a large proportion of models.
@@
 Computing the statistics at the peptide level, e.g., the $t$-statistics or $p$-value, then summarizing these statistics using their median produces the highest AUC scores among all the tested models.
-By using the beta distribution to model the number of peptides observed, we can further improve the pAUC scores (see Figure).
+By using the beta distribution to model the number of peptides observed, we can further improve the pAUC scores (see Figure TODO(wew)).
 However, the properties of beta-based probabilities are not well understood; for instance, they have no uniform distribution under the null hypothesis.
@@
 Furthermore, peptides are stronger affected by missing values, reducing the number of contrasts we could estimate for the dataset (see Figure \@ref(fig:FDRfdp) left panel).
-To adress the issue of missing observation we could either run proDA or the imputation of group averages method discussed here and afterwards apply `r BiocStyle::Biocpkg("PECA")`.
+To address the issue of missing observation, we could either run `r BiocStyle::Biocpkg("proDA")` or the imputation of group averages method discussed here and afterwards apply `r BiocStyle::Biocpkg("PECA")`.
 
-The probabilistic dropout analysis implemented in the _proDA_ produces inferences comparable to other methods (Figure \@ref(fig:FDRfdp) for the largest number of proteins (Figure \@ref(fig:FDRfdp) left panel).
+The probabilistic dropout analysis implemented in the `r BiocStyle::Biocpkg("proDA")` produces inferences comparable to other methods (Figure \@ref(fig:FDRfdp) for the largest number of proteins (Figure \@ref(fig:FDRfdp) left panel).
 Moreover, the estimated fold changes are unbiased and show high diagnostic accuracy.
-Furthermore, the scaled $p$-values and $t$-statics are on par with those obtained by the linear model with variance moderation.
-Therefore, we plan to integrate the _proDA_ package as an additional modeling option into `r BiocStyle::Githubpkg("fgcz/prolfqua")` (see Figure \@ref(fig:ContrastUML)).
+Furthermore, the scaled $p$-values and $t$-statistics are on par with those obtained by the linear model with variance moderation.
+Therefore, we plan to integrate the `r BiocStyle::Biocpkg("proDA")` package as an additional modeling option into `r BiocStyle::Githubpkg("fgcz/prolfqua")` (see Figure \@ref(fig:ContrastUML)).
 
@@
 Second, use linear models as long as they suffice to model your experimental design.
-Use mixed effect models only if you have to (e.g., repeated measurements).
+Use mixed effect models only if you have to, e.g., repeated measurements.
 If you use fixed-effect linear models, apply variance moderation since this improves the $t$-statistics and $p$-value estimates.
@@
 `r BiocStyle::Githubpkg("fgcz/prolfqua")` allows for considerable flexibility to model your experiments.
-Various types of models are available, and the contrast specification is consistent and intuitive for all models.
-The benchmarking results are comparable with other methods.
+Various types of models are available (for an overview see table TODO(wew)), and the contrast specification is consistent and intuitive for all models.
+The benchmarking results are comparable with existing methods.
 The modular design of `r BiocStyle::Githubpkg("fgcz/prolfqua")`, allows for customizing the analysis or adding new features, e.g., generalized linear models (_glm_'s) for missing data modeling or robust linear models (_rlm_'s), quickly.
@@
 In summary `r BiocStyle::Githubpkg("fgcz/prolfqua")` is an easy-to-use and feature rich R package to analyse quantitative mass spectrometric data, report results, and benchmark MS software and statistical methods.
-We provide more details at the website https://wolski.github.io/prolfqua/.
-To install:
-
-```{r echo=TRUE, eval=FALSE}
-install.packages('remotes')
-remotes::install_gitlab("wolski/prolfquadata",
-                        host="gitlab.bfabric.org")
-remotes::install_github('wolski/prolfqua',
-                        build_vignettes = TRUE)
-
-```
+We provide more details at the website https://github.com/fgcz/prolfqua/.
 
@@
 |ESI-MS |Electro-Spray-Ionization Mass Spectrometry         |
-|FDR    |False Discovery Rate                               |
+|FC     |Fold Change                                        |
+|$FDP$  |False Discovery Proportion                         |
+|$FDR$  |False Discovery Rate                               |
 |HPLC   |High-Pressure Liquid Chromatography                |
@@
 
+TODO(wew)
+
 ## Comparing benchmark results
 
-Here we compare the benchmark results for the proDA model and the linear model with variance moderation, but using only the protein where inference could be performed for both models.
-We see that the fold change estimates obtained with the proda pacakge show better diagnostic accuracy.
+Here we compare the benchmark results for the `r BiocStyle::Biocpkg("proDA")` model and the linear model with variance moderation, but using only the protein where inference could be performed for both models.
+We see that the fold change estimates obtained with the `r BiocStyle::Biocpkg("proDA")` package show better diagnostic accuracy.
 
-```{r}
-
+```{r proDABenchmarks}
 prodB <- allBenchmarks$benchmark_proDA$clone(deep=TRUE)
@@
 
-tmp <- dplyr::inner_join(prodSum$ftable$content, modSum$ftable$content, by=c("contrast","what"), suffix = c(".proDA",".lmMod"))
-tmp <- tmp |> tidyr::unite("score_contr" , what, contrast, remove = FALSE)
+tmp <- dplyr::inner_join(prodSum$ftable$content,
+                         modSum$ftable$content,
+                         by=c("contrast","what"), suffix = c(".proDA",".lmMod")) |>
+  tidyr::unite("score_contr" , what, contrast, remove = FALSE)
 
 ggplot2::ggplot(tmp, ggplot2::aes(x = contrast, y = pAUC_10.proDA - pAUC_10.lmMod)) + ggplot2::geom_bar(stat="identity") + ggplot2::facet_wrap(~what) +  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust=1))
-
 ```
 
-```{r}
+```{r proDAAllBenchmarks}
 proda <- allBenchmarks$benchmark_proDA$data()

### 2022-03-28 inst/MyArticle/paper.Rmd
diff

@@
 __R__ linear model and linear mixed effect models allow modeling parallel designs, repeated measurements, factorial designs, and many more.
-__R__'s formula interface for linear models is flexible, widely used, and well documented (TODO(wew): cite the docu).
+__R__'s formula interface for linear models is flexible, widely used, and well documented ([@faraway2016extending], page : ).
 This makes it easy to reproduce an analysis performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` in other statistical programming languages.
@@
 ```
-
-[TODO SAINTExpress integration, proDA]
 
@@
 Because we used the median as the i-th order statistic $i = \frac{n}{2} + 0.5$.
-Therefore, $\gamma = i = \frac{n}{2} + 0.5$ and $\delta = n - i + 1 = n - (\frac{n}{2} + 0.5) + 1 = \frac{n}{2} + 0.5 = \gamma$ are used to parameterize the (TODO(wew): what is CDF) CDF of the Beta distribution.
+Therefore, $\gamma = i = \frac{n}{2} + 0.5$ and $\delta = n - i + 1 = n - (\frac{n}{2} + 0.5) + 1 = \frac{n}{2} + 0.5 = \gamma$ are used to parameterize the cumulative distribution function (CDF) of the Beta distribution.
 We implemented this method in the class `ContrastROPECA` (Figure \@ref(fig:ContrastUML)).
@@
 By using the beta distribution to model the number of peptides observed, we can further improve the pAUC scores (see Figure TODO(wew)).
-However, the properties of beta-based probabilities are not well understood; for instance, they have no uniform distribution under the null hypothesis.
+However, the properties of beta-based probabilities are not well understood; for instance, they have no uniform distribution under the null hypothesis (TODO remove they).
 The FDR estimates obtained when correcting multiple testing with the Benjamini-Hochberg method are strongly biased (see Figure \@ref(fig:FDRfdp) right panel).
@@
 `r BiocStyle::Githubpkg("fgcz/prolfqua")` allows for considerable flexibility to model your experiments.
-Various types of models are available (for an overview see table TODO(wew)), and the contrast specification is consistent and intuitive for all models.
+Various types of models are available (for an overview see table TODO(wew) create table with methods implemented in prolfqua), and the contrast specification is consistent and intuitive for all models.
 The benchmarking results are comparable with existing methods.
@@
 |AUC    |Area Under the Curve                               |
+|CDF    |Cumulative Distribution Function                   |
 |ESI-MS |Electro-Spray-Ionization Mass Spectrometry         |

### 2022-04-01 inst/MyArticle/paper.Rmd
added GSEA reference

@@
 Based on these models, tests can be applied to examine whether a factor or interactions are significant, or differences between factor levels can be estimated, and their significance tested.
-Proteins can be ordered or grouped using the estimated fold-change, $t$-statistics, scaled $p$-value, or false discovery rate (FDR) and then subjected to gene set enrichment or over-representation analysis [TODO ...] or validation experiments.
+Proteins can be ordered or grouped using the estimated fold-change, $t$-statistics, scaled $p$-value, or false discovery rate (FDR) and then subjected to gene set enrichment or over-representation analysis [@subramanian2005gene] or validation experiments.
 

### 2022-04-01 inst/MyArticle/paper.Rmd
update bib

@@
 
-Label-Free Quantification (LFQ) Proteomics Experiments enable monitoring relative abundances of thousands of proteins in biological samples.
+Protein quantification proteomics Experiments enable monitoring relative abundances of thousands of proteins in biological samples.
 For most studies, parallel group designs are used, e.g., treatment versus control are used [@de2022apoe2, @laubscher2021baf].
-More recently, more complex experimental designs are implemented with increasing numbers of samples, i.e., factorial designs and longitudinal studies (time series), sometimes with repeated measurements on the same subject [@tan2022proteomic, @meier2021reduced].
+More recently, more complex experimental designs with increasing numbers of samples are studid, i.e., factorial designs and longitudinal studies (time series), sometimes with repeated measurements on the same subject [@tan2022proteomic, @meier2021reduced].
 The obtained data can be modeled using linear fixed-, mixed-, or random-effects models [@Bates2015FittingJSS].
-Based on these models, tests can be applied to examine whether a factor or interactions are significant, or differences between factor levels can be estimated, and their significance tested.
+Based on these models, tests can be applied to examine whether a specific factors and their interactions are significant.
+Furthermore it can be tested if differences between factor levels can be estimated, and their significance tested.
 Proteins can be ordered or grouped using the estimated fold-change, $t$-statistics, scaled $p$-value, or false discovery rate (FDR) and then subjected to gene set enrichment or over-representation analysis [@subramanian2005gene] or validation experiments.
@@
 
-When analyzing parallel group designs using a single explanatory variable, we can use all packages for the statistical data analysis of the data; however, if we want to model factorial designs or repeated measurements, we can use only some of them.
+When analyzing parallel-group designs using a single explanatory variable, we can use all R packages; however, we can use only some of them if we want to model factorial designs or repeated measurements.
 Table \@ref(tab:tableOverview) gives an overview of the models and features supported by these packages.
+We see that packages such as $limma$ and $proDA$ allow us to a comprehensive variety of hypotheses; however, in-depth knowledge of design matrix specification is required [@law2020guide].
+
+Models in `r BiocStyle::Githubpkg("fgcz/prolfqua")` are specified using R’s linear model and mixed model formula interface.
+Therefore, knowledge of the R regression model infrastructure [@faraway2016extending,@MASS2002] is an advantage when using our package.
+Furthermore, this glass box approach should make it easy to reimplement an analysis which was performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` using base R or other programming languages by reading the analysis script.
+However, acknowledging the R formula interface’s complexity to non-statisticians and the popularity of `r BiocStyle::Biocpkg("MSstats")`, we provide the functionality to suggest a model formula from the sample annotation provided in tabular form similarily to `r BiocStyle::Biocpkg("MSstats")` [@Choi2014].
 
@@
 However, after examining the R packages for modeling proteomics quantification data, we observed that model specification, input, and output formats differ widely and wildly.
-However, all of them fit some type of linear model either to peptide or protein intensities, determine contrasts among conditions, and afterward, some of them apply empirical Bayes variance shrinkage.
+However, many of these packages fit linear model either to peptide or protein intensities, determine contrasts among conditions, and afterward, apply empirical Bayes variance shrinkage.
 Therefore, the revised goal was to provide a modular object oriented (OO) design, with R linear models as a core, where features such as $p$-value aggregation, e.g., `r BiocStyle::Biocpkg("PECA")` or variance shrinkage could be easily enabled (see Figure \@ref(fig:ContrastUML)).
-
-Models in `r BiocStyle::Githubpkg("fgcz/prolfqua")` are specified using R’s linear model and mixed model formula interface.
-Therefore, knowledge of the R regression model infrastructure [@faraway2016extending,@MASS2002] is an advantage when using our package.
-Furthermore, this glass box approach should make it easy to reimplement an analysis which was performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` using base R or other programming languages by reading the analysis script.
-However, acknowledging the R formula interface’s complexity to non-statisticians and the popularity of `r BiocStyle::Biocpkg("MSstats")`, we provide the functionality to suggest a model formula from the sample annotation provided in tabular form similarily to `r BiocStyle::Biocpkg("MSstats")` [@Choi2014].
 
@@
 Proteins for which the fold change was imputed in such a way are explicitly annotated in the result table.
-
 Other means of modelling missing information in proteomics data using the Hurdle model is discussed by [@goeminne2020msqrob], while the R package `r BiocStyle::Biocpkg("proDA")` models missingness using probabilistic dropout models [@bioxrvproDA2020].
-
-Other means of modelling missing information in proteomics data using the Hurdle model is discussed by [@goeminne2020msqrob], while the R package `r BiocStyle::Biocpkg("proDA")` models missingness using probabilistic dropout models [@bioxrvproDA2020].
+We show using a benchmark dataset that this, although trivial approach, can outperform methods based on data imputation or dropout models.
 
@@
 We use the Ionstar [@shen2018ionstar] dataset to benchmark the analysis pipeline and modeling methods implemented within `r BiocStyle::Githubpkg("fgcz/prolfqua")` and to compare our results with those of `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`.
+
+An essential parameter of any protein quantification pipeline, not discussed further, is the normalization of the intensities within the samples to make them comparable [@pursiheimo2015optimization].
+All benchmarks we run with the same data normalization method.
+The method used to infer protein intensities from peptide intensities is an additional important factor.
+For instance, the original _proDA_ publication uses MaxLFQ [@cox2014accurate] protein estimates.
+However, when using those protein abundance estimates, _proDA_ performs significantly worse when using the protein estimates obtained from peptide intensity estimates using Tukey's median polish,  $pAUC_{10}(\textrm{t-statistics}) = 66%$ and $pAUC_{10}(\textrm{t-statistics}) = 73%$ respectively (see Articles on [@prolfquaGithubIO] for more details).
+Last but not least, the software, identifying proteins, and generating the quantification values can significantly contribute to the performance of the entire pipeline [@Cox2008MaxQuant,@yu2020fast].
+However, here we focus on the comparison of the various modeling approaches.
 
@@
 TODO(wew): Ionstar dataset, give Pride ids, describe MaxQuant settings and version. Give reference to Experiment hub and the _prolfquaData_ package.
+
+For this benchmark dataset, the fraction of regulated proteins (_E.
+coli_) is relatively large.
+Therefore, to calibrate the measurement, we used human proteins only.
+In practice, it is unknown which proteins do and which don't change.
+Therefore, the assumption used when normalizing, scaling the data, and thus calibrating the measurement is that only a small subset of the proteins in the dataset is differentially regulated.
 
@@
 
-[TODO(wew): Goes in the discussion or introduction, and give references for normalization e.g.
-VSN paper.] However, an essential parameter, not discussed further, is the normalization of the intensities within the samples to make them comparable.
-For this benchmark dataset, the fraction of regulated proteins (_E.
-coli_) is relatively large.
-Therefore, to calibrate the measurement, we used human proteins only.
-In practice, it is unknown which proteins do and which don't change.
-Therefore, the assumption used when normalizing, scaling the data, and thus calibrating the measurement is that only a small subset of the proteins in the dataset is differentially regulated.
-
-[TODO(wew): Goes to discussion] The method used to infer protein intensities from more peptide intensities is an additional important factor.
-The original _proDA_ publication uses MaxLFQ [@cox2014accurate] protein estimates.
-When using those, _proDA_ performs significantly worse than _proDA_ when using the protein estimates obtained from peptide intensity estimates using Tukey's median polish,  $pAUC_{10}(\textrm{t-statistics}) = 66%$ and $pAUC_{10}(\textrm{t-statistics}) = 73%$ respectively (see Articles on [@prolfquaGithubIO] for more details).
-
 We assessed the performance of different modeling approaches implemented in prolfqua, `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`, using a benchmark dataset whose ground truth is known.
@@
 
-We can conclude that if we want to order the proteins according to the likelihood of being differentially regulated, for instance, to perform gene enrichment analysis, the $t$-statistic is better suited than the fold-change estimate.
+We can conclude that if we want to order the proteins according to the likelihood of being differentially regulated, for instance, to perform gene set enrichment analysis, the $t$-statistic is better suited than the fold-change estimate.
 Modeling the degrees of freedom when computing the $p$-values might improve the inference further.

### 2022-04-05 inst/MyArticle/paper.Rmd
added dataset description.

@@
 
-TODO(wew): Ionstar dataset, give Pride ids, describe MaxQuant settings and version. Give reference to Experiment hub and the _prolfquaData_ package.
-
-For this benchmark dataset, the fraction of regulated proteins (_E.
-coli_) is relatively large.
-Therefore, to calibrate the measurement, we used human proteins only.
-In practice, it is unknown which proteins do and which don't change.
-Therefore, the assumption used when normalizing, scaling the data, and thus calibrating the measurement is that only a small subset of the proteins in the dataset is differentially regulated.
+To evaluate the performance of differential expression analysis, we use the _IonStar_ benchmark dataset[@shen2018ionstar], available from the Proteomics Identifications Database (PRIDE) identifier PXD003881.
+$DH5\alpha$ _E.
+coli_ lysate was spiked in human pancreatic cancer cells(Panc-1) lysate at five different levels: $3\%$ _E.
+coli_, $4.5\%$ _E.
+coli_, $6\%$ _E.
+coli_, $7.5\%$ _E.
+coli_ and $9\%$ _E.
+coli_.
+We annotated these dilutions from smallest to largest with the letters a, b, c, d, e.
+We obtain different fold changes by comparing the various dilutions, e.g., when comparing dilution _e_ (9%) against dilution _d_ (7.5%), the expected fold change is 1.2 for E.
+coli proteins and 1 for human proteins.
+We processed the raw data using MaxQuant [@Cox2008MaxQuant] version Version 1.6.10.43, with MaxQuant default settings.
+The text files generated by Maxquant are available in the _prolfquadata_ R package [@prolfquadata].
 
@@
 
+The Benchmark functionality of `r BiocStyle::Githubpkg("fgcz/prolfqua")` includes ROC curves and computes partial areas under those curves (pAUC) and other scores, e.g., the false discovery proportion FDP.
+We use those plots to study how well linear, mixed effect models or $p$-value moderation models quantitative mass spectrometric high throughput experiments and compare them with results produced by `r BiocStyle::Biocpkg("MSstats")` [@MSstats2014] and `r BiocStyle::Biocpkg("proDA")` [@bioxrvproDA2020] (see also Figure \@ref(fig:FDRfdp))
+
 The peptide intensities from the _MaxQuant_ _peptide.txt_ file were $log_2$ transformed, and subsequent robust z-score transformation, where median and the mean average deviation (mad) were obtained from the human proteins only.
@@
 The mixed effect linear model, and the `r BiocStyle::Biocpkg("PECA")` model were fitted to the peptide intensities.
-The details of the `r BiocStyle::Githubpkg("fgcz/prolfqua")` benchmark implementation are available online  [BenchmarkingIonstarData](https://wolski.github.io/prolfqua/articles/BenchmarkingIonstarData.html).
+The details of the `r BiocStyle::Githubpkg("fgcz/prolfqua")` benchmark implementation are available online  [BenchmarkingIonstarData](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkingIonstarData.html).
 
-For benchmarking of `r BiocStyle::Biocpkg("MSstats")` we used the defaults specified in the [Benchmarking_MSstats](https://wolski.github.io/prolfqua/articles/)
-
-The Benchmark functionality of `r BiocStyle::Githubpkg("fgcz/prolfqua")` includes ROC curves and computes partial areas under those curves (pAUC) and other scores, e.g., the false discovery proportion FDP.
-We use those plots to study how well linear, mixed effect models or $p$-value moderation models quantitative mass spectrometric high throughput experiments and compare them with results produced by `r BiocStyle::Biocpkg("MSstats")` [@MSstats2014] and `r BiocStyle::Biocpkg("proDA")` [@bioxrvproDA2020] (see also Figure \@ref(fig:FDRfdp))
+For benchmarking of `r BiocStyle::Biocpkg("MSstats")` we used the MSstats default parameters (for details see: [Benchmarking_MSstats](https://wolski.github.io/prolfquabenchmark/articles/))
 

### 2022-04-05 inst/MyArticle/paper.Rmd
added table, adressed TODO's

@@
 ---
-title: prolfqua - An R-package for Analysing Proteomics Label Free Quantification Experiments
+title: prolfqua (prodea) - An R-package for Proteomics Differential Expression Analysis.
 author: 
@@
 We annotated these dilutions from smallest to largest with the letters a, b, c, d, e.
-We obtain different fold changes by comparing the various dilutions, e.g., when comparing dilution _e_ (9%) against dilution _d_ (7.5%), the expected fold change is 1.2 for E.
+As a result, we obtain different fold changes by comparing the various dilutions, e.g., when comparing dilution _e_ (9%) against dilution _d_ (7.5%), the expected fold change is 1.2 for E.
 coli proteins and 1 for human proteins.
-We processed the raw data using MaxQuant [@Cox2008MaxQuant] version Version 1.6.10.43, with MaxQuant default settings.
+We processed the raw data using MaxQuant [cite] version Version 1.6.10.43, with MaxQuant default settings.
 The text files generated by Maxquant are available in the _prolfquadata_ R package [@prolfquadata].
@@
 We implemented this method in the class `ContrastROPECA` (Figure \@ref(fig:ContrastUML)).
+
+```{r prolfquaModels}
+xM <- data.frame( 
+    lm = c("lm", "linear model on peptide or protein abundances", "strategy_lm, build_models, Contrasts"),
+    lmer = c("lmer" , "mixed effect models on peptide and protein abundances", "strategy_lmer, build,models, Contrasts"),
+    gmi = c("group mean imputation", "impute missing group means", "ContrastsSimpleImpute"),
+    ROPECA = c("ROPECA","linear models on peptide abundances", "strategy_lm, build_models, Contrast, ContrastsROPECA"),
+    limma = c("limma" , "empirical Bayes variance shrinkage", "ContrastsModerated"),
+    pic = c("Protein interaction scoring", "by wrapping SAINT express binaries", "runSaint, ContrastsSaintExpress"),
+    prodA = c("dropout model", "by warpping proDA functionality", "strategy_proDA, build_models, ContrastsProDa"), check.names =  FALSE)
+xM <- data.frame(t(xM))
+colnames(xM) <- c("m", "model", "prolfqua functions")
+knitr::kable(xM[,2:3], caption = "Models available in prolfqua.")
+
+```
 
@@
 
-# TODO(wew): give tmp a meaningful name
-tmp <- prolfqua::addContrastResults(prefer = contr, add = conI)
+contrasts <- prolfqua::addContrastResults(prefer = contr, add = conI)
 
 ## visualize results
-pl <- tmp$merged$get_Plotter()
+pl <- contrasts$merged$get_Plotter()
 p1 <- pl$histogram()$p.value
@@
 the $\log_2$ fold-change $\beta$, the $t$-statistic $\frac{\beta}{\sqrt{var(\beta)}}$, or the $p$-value and moderated $p$-value.
-For each statistic and each value of the statistics we then compute a confusion matrix (see Table \@ref(tab:confusionMatrix)).
+For each statistic and each value of the statistics we then compute a confusion matrix (see Table \@ref(tab:knitrConfusionMatrix)).
 From the confusion matrix we obtain measures such as true positive rate ($TPR$), false positive rate ($FPR$), or false discovery proportion ($FDP$) which are given by:
@@
 Computing the statistics at the peptide level, e.g., the $t$-statistics or $p$-value, then summarizing these statistics using their median produces the highest AUC scores among all the tested models.
-By using the beta distribution to model the number of peptides observed, we can further improve the pAUC scores (see Figure TODO(wew)).
-However, the properties of beta-based probabilities are not well understood; for instance, they have no uniform distribution under the null hypothesis (TODO remove they).
-The FDR estimates obtained when correcting multiple testing with the Benjamini-Hochberg method are strongly biased (see Figure \@ref(fig:FDRfdp) right panel).
+By using the beta distribution to model the number of peptides observed, we can further improve the pAUC scores (see Figure \@ref(fig:FDRfdp) top).
+However, the properties of beta-based probabilities are not well understood; for instance, the p-values are not uniformely distributed under the null hypothesis (not shown).
+Furthermore, the FDR estimates obtained when correcting for multiple testing with the Benjamini-Hochberg method are biased overestimating the false discovery proprotion (see Figure \@ref(fig:FDRfdp) right panel).
 Therefore, we can not recommend this method if an unbiased estimate of FDR is essential, which is frequently the case.
-Furthermore, peptides are stronger affected by missing values, reducing the number of contrasts we could estimate for the dataset (see Figure \@ref(fig:FDRfdp) left panel).
+In addition, peptides are stronger affected by missing values, reducing the number of contrasts we could estimate for the dataset (see Figure \@ref(fig:FDRfdp) bottom left).
 To address the issue of missing observation, we could either run `r BiocStyle::Biocpkg("proDA")` or the imputation of group averages method discussed here and afterwards apply `r BiocStyle::Biocpkg("PECA")`.
@@
 `r BiocStyle::Githubpkg("fgcz/prolfqua")` allows for considerable flexibility to model your experiments.
-Various types of models are available (for an overview see table TODO(wew) create table with methods implemented in prolfqua), and the contrast specification is consistent and intuitive for all models.
+Various types of models are available (see Figure \@ref(fig:ContrastUML) and Table \@ref(tab:prolfquaModels)), and the contrast specification is consistent and intuitive for all models.
 The benchmarking results are comparable with existing methods.
@@
 
-TODO(wew)
+```{r}
+
+```
 

### 2022-04-06 inst/MyArticle/paper.Rmd
Add correspondence

@@
 author: 
-  - Witold E. Wolski^[[Functional Genomics Center Zurich (FGCZ)](https://fgcz.ch) - Swiss Federal Institute of Technology in Zurich \| University of Zurich]
+  - Witold E.
+Wolski $^{\star,}$^[[Functional Genomics Center Zurich (FGCZ)](https://fgcz.ch) - Swiss Federal Institute of Technology in Zurich \| University of Zurich; Winterthurerstrasse 190, CH-8057 Zurich, Switzerland.
+Correspondence$^\star$\texttt{wew@fgcz.ethz.ch}]
   - Paolo Nanni$^*$
@@
   inference procedures data analysis application programmer interface
-  in R. The `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package integrates
   essential steps of the mass spectrometry-based label-free quantification
-  data analysis workflow as: quality control, data normalization, protein
   aggregation, statistical modeling, hypothesis testing, and sample size
   estimation.
-  We use `r BiocStyle::Githubpkg("fgcz/prolfqua")` to visualize and model
   simple experimental designs with a single explanatory variable and complex
   experiments with multiple factors.
-  Furthermore, we use `r BiocStyle::Githubpkg("fgcz/prolfqua")` to benchmark
   data acquisition, data preprocessing or data modeling methods. Furthermore,
-  we apply `r BiocStyle::Githubpkg("fgcz/prolfqua")` to develop highly
   customizable, visually appealing, and interactive data analysis reports
   in PDF or HTML format.
+  in R. The `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package integrates
+  data analysis workflow as: quality control, data normalization, protein
+  We use `r BiocStyle::Githubpkg("fgcz/prolfqua")` to visualize and model
+  Furthermore, we use `r BiocStyle::Githubpkg("fgcz/prolfqua")` to benchmark
+  we apply `r BiocStyle::Githubpkg("fgcz/prolfqua")` to develop highly
   The R package is available through https://github.com/fgcz/prolfqua.

### 2022-04-06 inst/MyArticle/paper.Rmd
Format tables

@@
 colnames(xM) <- c("m", "model", "prolfqua functions")
-knitr::kable(xM[,2:3], caption = "Models available in prolfqua.")
-
+knitr::kable(xM[,2:3], caption = "Models available in prolfqua.",
+             format = "latex", booktabs=TRUE) |>
+    kableExtra::kable_styling(latex_options="scale_down")
 ```
@@
 
-|       |                                                   |
-|:------|:--------------------------------------------------|
-|AUC    |Area Under the Curve                               |
-|CDF    |Cumulative Distribution Function                   |
-|ESI-MS |Electro-Spray-Ionization Mass Spectrometry         |
-|FC     |Fold Change                                        |
-|$FDP$  |False Discovery Proportion                         |
-|$FDR$  |False Discovery Rate                               |
-|HPLC   |High-Pressure Liquid Chromatography                |
-|LC-MS  |Liquid Chromatography followed by Mass Spectrometry|
-|LFQ    |Label-Free Quantification                          |
-|LOD    |Limit Of Detection                                 |
-|MS     |mass spectrometry                                  |
-|OO     |Object-Oriented                                    |
-|UML    |Unified Modeling Language                          |
+\begin{table}[H]
+\centering
+\begin{tabular}[H]{ll}
+\toprule
+Abbreviations & Explaination\\
+\midrule
+AUC    &Area Under the Curve                               \\
+CDF    &Cumulative Distribution Function                   \\
+ESI-MS &Electro-Spray-Ionization Mass Spectrometry         \\
+FC     &Fold Change                                        \\
+$FDP$  &False Discovery Proportion                         \\
+$FDR$  &False Discovery Rate                               \\
+HPLC   &High-Pressure Liquid Chromatography                \\
+LC-MS  &Liquid Chromatography followed by Mass Spectrometry\\
+LFQ    &Label-Free Quantification                          \\
+LOD    &Limit Of Detection                                 \\
+MS     &mass spectrometry                                  \\
+OO     &Object-Oriented                                    \\
+UML     &       Unified Modeling Language                  \\
+\bottomrule
+\end{tabular}
+\end{table}
 

### 2022-04-06 inst/MyArticle/paper.Rmd
my latest

--- 2022-04-05 inst/MyArticle/paper.Rmd
@@
 xM <- data.frame( 
-    lm = c("lm", "linear model on peptide or protein abundances", "strategy_lm, build_models, Contrasts"),
-    lmer = c("lmer" , "mixed effect models on peptide and protein abundances", "strategy_lmer, build,models, Contrasts"),
+    lm = c("lm", "linear model for peptides or proteins", "strategy_lm, Contrasts"),
+    lmer = c("lmer" , "mixed effect models for peptides or proteins", "strategy_lmer, Contrasts"),
     gmi = c("group mean imputation", "impute missing group means", "ContrastsSimpleImpute"),
-    ROPECA = c("ROPECA","linear models on peptide abundances", "strategy_lm, build_models, Contrast, ContrastsROPECA"),
+    ROPECA = c("ROPECA"," ", "ContrastsROPECA"),
     limma = c("limma" , "empirical Bayes variance shrinkage", "ContrastsModerated"),
-    pic = c("Protein interaction scoring", "by wrapping SAINT express binaries", "runSaint, ContrastsSaintExpress"),
-    prodA = c("dropout model", "by warpping proDA functionality", "strategy_proDA, build_models, ContrastsProDa"), check.names =  FALSE)
+    pic = c("SAINTexpress", "protein interaction scoring", "runSaint, ContrastsSaintExpress"),
+    prodA = c("proDA", "dropout models", "strategy_proDA, Contrasts_proDa"), check.names =  FALSE)
 xM <- data.frame(t(xM))
-colnames(xM) <- c("m", "model", "prolfqua functions")
-knitr::kable(xM[,2:3], caption = "Models available in prolfqua.")
+colnames(xM) <- c("id", "model", "prolfqua functions")
+knitr::kable(tibble::as_tibble(xM[,1:3]), caption = "Models available in prolfqua.")
 

### 2022-04-06 inst/MyArticle/paper.Rmd

--- 2022-04-06 inst/MyArticle/paper.Rmd
@@
 xM <- data.frame( 
-    lm = c("lm", "linear model on peptide or protein abundances", "strategy_lm, build_models, Contrasts"),
-    lmer = c("lmer" , "mixed effect models on peptide and protein abundances", "strategy_lmer, build,models, Contrasts"),
+    lm = c("lm", "linear model for peptides or proteins", "strategy_lm, Contrasts"),
+    lmer = c("lmer" , "mixed effect models for peptides or proteins", "strategy_lmer, Contrasts"),
     gmi = c("group mean imputation", "impute missing group means", "ContrastsSimpleImpute"),
-    ROPECA = c("ROPECA","linear models on peptide abundances", "strategy_lm, build_models, Contrast, ContrastsROPECA"),
+    ROPECA = c("ROPECA"," ", "ContrastsROPECA"),
     limma = c("limma" , "empirical Bayes variance shrinkage", "ContrastsModerated"),
-    pic = c("Protein interaction scoring", "by wrapping SAINT express binaries", "runSaint, ContrastsSaintExpress"),
-    prodA = c("dropout model", "by warpping proDA functionality", "strategy_proDA, build_models, ContrastsProDa"), check.names =  FALSE)
+    pic = c("SAINTexpress", "protein interaction scoring", "runSaint, ContrastsSaintExpress"),
+    prodA = c("proDA", "dropout models", "strategy_proDA, Contrasts_proDa"), check.names =  FALSE)
 xM <- data.frame(t(xM))
-colnames(xM) <- c("m", "model", "prolfqua functions")
-knitr::kable(xM[,2:3], caption = "Models available in prolfqua.",
+
+colnames(xM) <- c("id", "model", "prolfqua functions")
+knitr::kable(tibble::as_tibble(xM[,1:3]), caption = "Models available in prolfqua.",
              format = "latex", booktabs=TRUE) |>
     kableExtra::kable_styling(latex_options="scale_down")
+
 ```

### 2022-04-15 inst/MyArticle/paper.Rmd
title.
last

@@
 ---
-title: prolfqua (prodea) - An R-package for Proteomics Differential Expression Analysis.
+title: prolfqua - An R-package for Proteomics Differential Expression Analysis.
 author: 
   - Witold E.
-Wolski $^{\star,}$^[[Functional Genomics Center Zurich (FGCZ)](https://fgcz.ch) - Swiss Federal Institute of Technology in Zurich \| University of Zurich; Winterthurerstrasse 190, CH-8057 Zurich, Switzerland.
+Wolski $^{\star}$^[[Functional Genomics Center Zurich (FGCZ)](https://fgcz.ch) - Swiss Federal Institute of Technology in Zurich \| University of Zurich; Winterthurerstrasse 190, CH-8057 Zurich, Switzerland.
 Correspondence$^\star$\texttt{wew@fgcz.ethz.ch}]
-  - Paolo Nanni$^*$
-  - Jonas Grossmann$^*$
-  - Maria d'Errico$^*$
-  - Ralph Schlapbach$^*$
-  - Christian Panse$^*$
+  - Paolo Nanni$^{\star}$
+  - Jonas Grossmann$^{\star}$
+  - Maria d'Errico$^{\star}$
+  - Ralph Schlapbach$^{\star}$
+  - Christian Panse$^{\star}$
 output: 

### 2022-05-03 inst/MyArticle/paper.Rmd
updated abstract

@@
 author: 
-  - Witold E.
-Wolski $^{\star}$^[[Functional Genomics Center Zurich (FGCZ)](https://fgcz.ch) - Swiss Federal Institute of Technology in Zurich \| University of Zurich; Winterthurerstrasse 190, CH-8057 Zurich, Switzerland.
-Correspondence$^\star$\texttt{wew@fgcz.ethz.ch}]
   - Paolo Nanni$^{\star}$
@@
   - Christian Panse$^{\star}$
+  - Witold E.
+Wolski $^{\star}$^[[Functional Genomics Center Zurich (FGCZ)](https://fgcz.ch) - Swiss Federal Institute of Technology in Zurich \| University of Zurich; Winterthurerstrasse 190, CH-8057 Zurich, Switzerland.
+Correspondence$^\star$\texttt{wew@fgcz.ethz.ch}]
 output: 
@@
 abstract: |
-  Protein mass spectrometry is widely used for quantitative proteomics
-  studies and in particular for relative protein quantification.
-  Nevertheless, there is still a need for a flexible and easy-to-use, 
-  transparently supporting a variety of well principled statistical
-  inference procedures data analysis application programmer interface
-  essential steps of the mass spectrometry-based label-free quantification
-  aggregation, statistical modeling, hypothesis testing, and sample size
-  estimation.
-  simple experimental designs with a single explanatory variable and complex
-  experiments with multiple factors.
-  data acquisition, data preprocessing or data modeling methods. Furthermore,
-  customizable, visually appealing, and interactive data analysis reports
-  in PDF or HTML format.
-  in R. The `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package integrates
-  data analysis workflow as: quality control, data normalization, protein
-  We use `r BiocStyle::Githubpkg("fgcz/prolfqua")` to visualize and model
-  Furthermore, we use `r BiocStyle::Githubpkg("fgcz/prolfqua")` to benchmark
-  we apply `r BiocStyle::Githubpkg("fgcz/prolfqua")` to develop highly
+    Protein mass spectrometry is widely used for quantitative proteomics studies, relative protein quantification, and differential expression analysis of proteins.
+Nevertheless, there is still a need for a flexible and easy-to-use application programmer interface in R that transparently supports a variety of well principled statistical inference procedures.
+    The `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package integrates essential steps of the mass spectrometry-based label-free quantification data analysis workflow: quality control, data normalization, protein aggregation, statistical modeling, hypothesis testing, and sample size estimation.
+The package can model simple experimental designs with a single explanatory variable and complex experiments with multiple factors using various statistical models.
+It also can be used to develop applications that produce interactive data analysis reports in PDF or HTML format.
+     Furthermore, the package can be used to benchmark data acquisition, data preprocessing, or data modeling methods using gold standard datasets.
 The R package is available through https://github.com/fgcz/prolfqua.
+
 ---

### 2022-05-04 inst/MyArticle/paper.Rmd
rename intensity to abudance. move to knitr tables.

@@
      Furthermore, the package can be used to benchmark data acquisition, data preprocessing, or data modeling methods using gold standard datasets.
+Using the benchmark functionality, we show that the methods implemented allow performing sensitive and specific differential expression analysis.
      The R package is available through https://github.com/fgcz/prolfqua.
@@
 
-We focused on comparing models to estimate fold changes, variances and to obtain false discovery rates.
-
-We assessed the performance of different modeling approaches implemented in prolfqua, `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`, using a benchmark dataset whose ground truth is known.
-However, the dataset used has several limitations: first, the design is a parallel-group design, i.e., we could not benchmark how well interactions are modelled; secondly, essential sources of variation present in every experiment, such as the biochemical and biological, are not measured.
+We assessed the performance of different modeling approaches implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")`, `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`, using a benchmark dataset whose ground truth is known.
+However, since the for each dilution only technical replicates are available, essential sources of variation present in every experiment, such as the biochemical and biological, are not measured.
 The dataset captures only the variance from the chromatography, electro-spray, and mass spectrometric measurement method.
 Thus, while we can extrapolate some of the results obtained to more realistic datasets, we need to be careful not to over-interpret the results.
+Specifically, in more realistic experiments the observed variances will be higher and sensitivities and specificity, will be worse given the same sample sizes.
+Furthermore, the proportion of missing observations in real world datasets might be higher.
 
-We can conclude that if we want to order the proteins according to the likelihood of being differentially regulated, for instance, to perform gene set enrichment analysis, the $t$-statistic is better suited than the fold-change estimate.
-Modeling the degrees of freedom when computing the $p$-values might improve the inference further.
-However, the observed change is minuscule.
-
-For the fixed effect linear model, the empirical Bayes variance shrinkage, as suggested by [@Smyth2004linear], consistently improves compared with the unmoderated estimates (data not shown).
+However, we can conclude that if we want to order the proteins according to the likelihood of being differentially regulated, to perform gene set enrichment analysis [@subramanian2005gene], the $t$-statistic is better suited than the fold-change estimate.
+Modeling the degrees of freedom when computing the $p$-values might further improve the inference, However, the observed change is minuscule (see Figure \@ref(fig:)).
+Furthermore, for the fixed effect linear model, the empirical Bayes variance shrinkage, as suggested by [@Smyth2004linear], consistently improves the ranking of proteins compared with the unmoderated estimates (data not shown).
 However, there is no such effect for the mixed effect model, most likely because the degrees of freedom are erroneously estimated for a large proportion of models.
@@
 Moreover, the estimated fold changes are unbiased and show high diagnostic accuracy.
-Furthermore, the scaled $p$-values and $t$-statistics are on par with those obtained by the linear model with variance moderation.
+Furthermore, the scaled $p$-values or the $t$-statistics are on par with those obtained by the linear model with variance moderation.
 Therefore, we plan to integrate the `r BiocStyle::Biocpkg("proDA")` package as an additional modeling option into `r BiocStyle::Githubpkg("fgcz/prolfqua")` (see Figure \@ref(fig:ContrastUML)).
+
+Using the tidy table to model the data ensures interoperability with other proteomics-related packages that manage their data with tidy-tables, e.g., `r BiocStyle::CRANpkg("protti")`.
+Furthermore, the use of `R6` classes, which encapsulate the configuration and the data, allows writing very concise code (no function arguments needed), and auto-completion support in the Rstudio editor makes it easy for novices to explore `r BiocStyle::Githubpkg("fgcz/prolfqua")`s functionality.
+
+# Conclusion
+
+`r BiocStyle::Githubpkg("fgcz/prolfqua")` allows for considerable flexibility to model your experiments.
+Various types of models are available (see Figure \@ref(fig:ContrastUML) and Table \@ref(tab:prolfquaModels)), and the contrast specification is consistent and intuitive for all models.
+The modular design of `r BiocStyle::Githubpkg("fgcz/prolfqua")`, allows for adding new features, e.g., generalized linear models (_glm_'s) for missing data modeling or robust linear models (_rlm_'s), in the future.
 
@@
 Second, use linear models as long as they suffice to model your experimental design.
-Use mixed effect models only if you have to, e.g., repeated measurements.
+Use mixed effect models only if you have to, e.g., repeated measurements and if the sample sizes are large.
 If you use fixed-effect linear models, apply variance moderation since this improves the $t$-statistics and $p$-value estimates.
@@
 
-`r BiocStyle::Githubpkg("fgcz/prolfqua")` allows for considerable flexibility to model your experiments.
-Various types of models are available (see Figure \@ref(fig:ContrastUML) and Table \@ref(tab:prolfquaModels)), and the contrast specification is consistent and intuitive for all models.
-The benchmarking results are comparable with existing methods.
-The modular design of `r BiocStyle::Githubpkg("fgcz/prolfqua")`, allows for customizing the analysis or adding new features, e.g., generalized linear models (_glm_'s) for missing data modeling or robust linear models (_rlm_'s), quickly.
-
-Using the tidy table to model the data ensures interoperability with other proteomics-related packages that manage their data with tidy-tables, e.g., `r BiocStyle::CRANpkg("protti")`.
-Furthermore, the use of `R6` classes, which encapsulate the configuration and the data, allows writing very concise code (no function arguments needed), and auto-completion support in the Rstudio editor makes it easy for novices to explore `r BiocStyle::Githubpkg("fgcz/prolfqua")`s functionality.
-
 In summary `r BiocStyle::Githubpkg("fgcz/prolfqua")` is an easy-to-use and feature rich R package to analyse quantitative mass spectrometric data, report results, and benchmark MS software and statistical methods.
 We provide more details at the website https://github.com/fgcz/prolfqua/.
-
-# Conclusion (tbd)
 
@@
 ggplot2::ggplot(x,ggplot2::aes( x = diff.imp, y = diff.proda) ) + ggplot2::geom_point() + ggplot2::facet_wrap(~contrast + species) + ggplot2::geom_hline(yintercept = 0, color="red") + ggplot2::geom_vline(xintercept = 0, color="red")
+
 ```

### 2022-05-04 inst/MyArticle/paper.Rmd
updates to abstract and to the introduction.

@@
     Protein mass spectrometry is widely used for quantitative proteomics studies, relative protein quantification, and differential expression analysis of proteins.
-Nevertheless, there is still a need for a flexible and easy-to-use application programmer interface in R that transparently supports a variety of well principled statistical inference procedures.
-    The `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package integrates essential steps of the mass spectrometry-based label-free quantification data analysis workflow: quality control, data normalization, protein aggregation, statistical modeling, hypothesis testing, and sample size estimation.
-The package can model simple experimental designs with a single explanatory variable and complex experiments with multiple factors using various statistical models.
-It also can be used to develop applications that produce interactive data analysis reports in PDF or HTML format.
-     Furthermore, the package can be used to benchmark data acquisition, data preprocessing, or data modeling methods using gold standard datasets.
-Using the benchmark functionality, we show that the methods implemented allow performing sensitive and specific differential expression analysis.
+Nevertheless, there is still a need for a flexible and easy-to-use application programmer interface in R that transparently supports a variety of well principled statistical procedures.
+The package can model simple experimental designs with a single explanatory variable and complex experiments with multiple factors and hypothesis testing.
+    The `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package integrates, using an R6 application programmer interface, essential steps of the mass spectrometry-based differential expression analysis workflow: quality control, data normalization, protein aggregation, statistical modeling, hypothesis testing, and sample size estimation.
+     Furthermore, the package implements benchmark functionality that can help to compare data acquisition, data preprocessing, or data modeling methods using a gold standard dataset.
+Finally, we show that the implemented methods allow sensitive and specific differential expression analysis.
      The R package is available through https://github.com/fgcz/prolfqua.
@@
 ## See ACS guidelines [https://publish.acs.org/publish/author_guidelines?coden=ancham#manuscript_text_components]
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 7, fig.height = 5)
+knitr::opts_chunk$set(echo = FALSE,
+                      warning = FALSE,
+                      message = FALSE,
+                      fig.width = 7,
+                      fig.height = 5)
 ```
@@
 Finally, peptide identification is achieved by fragmenting and matching the measured fragment masses to theoretical masses computed from known peptide sequences[@eng2015deeper, @yu2016pipi, @kong2017msfragger].
-For quantification, intact peptide ions [@yu2020fast, @Cox2008MaxQuant] or products of peptide ion fragmentation [@rost2014openswath, @demichev2020dia] are counted and aggregated.
+For quantification, intact peptide ions [@yu2020fast, @Cox2008MaxQuant] or products of peptide ion fragmentation [@rost2014openswath, @demichev2020dia] are counted and aggregated to obtain precursor abundances.
 Finally, the identified and quantified peptides are assigned to proteins based on protein sequence information.
 
-Protein quantification proteomics Experiments enable monitoring relative abundances of thousands of proteins in biological samples.
+Protein quantification proteomics experiments enable monitoring relative abundances of thousands of proteins in biological samples.
 For most studies, parallel group designs are used, e.g., treatment versus control are used [@de2022apoe2, @laubscher2021baf].
-More recently, more complex experimental designs with increasing numbers of samples are studid, i.e., factorial designs and longitudinal studies (time series), sometimes with repeated measurements on the same subject [@tan2022proteomic, @meier2021reduced].
+More recently, more complex experimental designs with increasing numbers of samples are studied, i.e., factorial designs and longitudinal studies (time series), sometimes with repeated measurements on the same subject [@tan2022proteomic, @meier2021reduced].
 The obtained data can be modeled using linear fixed-, mixed-, or random-effects models [@Bates2015FittingJSS].
 Based on these models, tests can be applied to examine whether a specific factors and their interactions are significant.
-Furthermore it can be tested if differences between factor levels can be estimated, and their significance tested.
-Proteins can be ordered or grouped using the estimated fold-change, $t$-statistics, scaled $p$-value, or false discovery rate (FDR) and then subjected to gene set enrichment or over-representation analysis [@subramanian2005gene] or validation experiments.
+Furthermore it can be tested if differences of protein abundance between groups are significant.
+For this application it is important that the model provides an unbiased estimate of the false discovery rate (FDR) so that protein lists can be selected for follow-up experiments [TODO reference].
+Furthermore, the quantified proteins can be ranked using the estimated fold-change, $t$-statistics or scaled $p$-value, and then subjected to gene set enrichment (GSEA) or over-representation analysis [@subramanian2005gene] to determine up or down-regulated groups of proteins.
 
-Several packages exist to model mass spectrometry LFQ experiments, e.g., `r BiocStyle::Biocpkg("limma")` [@Ritchie2015], `r BiocStyle::Biocpkg("MSstats")` [@Choi2014], `r BiocStyle::Biocpkg("PECA")` [@Suomi2017bEnhanced],  `r BiocStyle::Biocpkg("msqrob2")` [@Goeminne2016] or `r BiocStyle::Biocpkg("proDA")` [@bioxrvproDA2020], to mention some, all implemented in the R.
-Each of them has some unique features.
-For example, `r BiocStyle::Biocpkg("MSstats")` infers the model from the sample annotation and generates the model formula.
+Several packages exist to model mass spectrometry protein quantification experiments, e.g., `r BiocStyle::Biocpkg("limma")` [@Ritchie2015], `r BiocStyle::Biocpkg("MSstats")` [@Choi2014], `r BiocStyle::Biocpkg("PECA")` [@Suomi2017bEnhanced],  `r BiocStyle::Biocpkg("msqrob2")` [@Goeminne2016] or `r BiocStyle::Biocpkg("proDA")` [@bioxrvproDA2020], to mention some, all implemented in the R.
+Each of them has some unique features, for example, `r BiocStyle::Biocpkg("MSstats")` infers the model and generates the model formula from the sample annotation, allowing users with limited statistical knowledge to perform differential expression analisis (DEA).
 At the same time, `r BiocStyle::Biocpkg("limma")` allows to specify a linear model formula and implements the empirical Bayes variance shrinkage method.
-The R-package `r BiocStyle::Biocpkg("PECA")`, performs a roll-up of peptide level fold change and $p$-value estimates, obtained from `r BiocStyle::Biocpkg("limma")`, to protein level estimates.
+The R-package `r BiocStyle::Biocpkg("PECA")`, performs a roll-up of peptide level fold change and peptide level $p$-value estimates, obtained from `r BiocStyle::Biocpkg("limma")`, to protein level estimates.
 Furthermore, `r BiocStyle::Biocpkg("msqrob2")` implements two models: robust linear models fitted to protein intensities and robust ridge regression fitted to peptide intensities and combines them with empirical Bayes variance shrinkage known from `r BiocStyle::Biocpkg("limma")`.
@@
 
-When analyzing parallel-group designs using a single explanatory variable, we can use all R packages; however, we can use only some of them if we want to model factorial designs or repeated measurements.
+When analyzing parallel-group designs using a single explanatory variable, and contrasting the groups, we can use all R packages; however, we can use only some of them if we want to model factorial designs or repeated measurements.
 Table \@ref(tab:tableOverview) gives an overview of the models and features supported by these packages.
-We see that packages such as $limma$ and $proDA$ allow us to a comprehensive variety of hypotheses; however, in-depth knowledge of design matrix specification is required [@law2020guide].
+We see that packages such as $limma$ and $proDA$ allow us to fit a comprehensive variety of models and test various hypotheses; however, in-depth knowledge of design matrix specification is required [@law2020guide].
 

### 2022-05-05 inst/MyArticle/paper.Rmd
Improve description of UML diagrams.

@@
 author: 
+  - Witold E.
+Wolski $^{\star}$^[[Functional Genomics Center Zurich (FGCZ)](https://fgcz.ch) - Swiss Federal Institute of Technology in Zurich \| University of Zurich; Winterthurerstrasse 190, CH-8057 Zurich, Switzerland.
+Correspondence$^\star$\texttt{wew@fgcz.ethz.ch}]
   - Paolo Nanni$^{\star}$
@@
   - Christian Panse$^{\star}$
-  - Witold E.
-Wolski $^{\star}$^[[Functional Genomics Center Zurich (FGCZ)](https://fgcz.ch) - Swiss Federal Institute of Technology in Zurich \| University of Zurich; Winterthurerstrasse 190, CH-8057 Zurich, Switzerland.
-Correspondence$^\star$\texttt{wew@fgcz.ethz.ch}]
 output: 
@@
 
-(ref:LFQData) Class Diagram for LFQData and related classes.
-The LFQData_Plotter class uses the LFQData class to implement methods for plotting.
-Similarily the LFQData_Stats, LFQData_Summary, reference the LFQData class and group methods for variance and sample size estimation or for summarizing peptide and protein counts.
+(ref:LFQData) Class Diagram of classes representing the proteomics data.
+The LFQData class encapsulates the data.
+Other classes are decorating the LFQData class with additional functionality.
+The LFQData_Plotter class uses the LFQData class to implement methods for data visualization.
+Similarly, the LFQData_Stats, LFQData_Summary, reference the LFQData class, group methods for variance and sample size estimation, or summarizing peptide and protein counts.
 
@@
 __R__ linear model and linear mixed effect models allow modeling parallel designs, repeated measurements, factorial designs, and many more.
-__R__'s formula interface for linear models is flexible, widely used, and well documented ([@faraway2016extending], page : ).
-This makes it easy to reproduce an analysis performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` in other statistical programming languages.
-We implemented features specific to high throughput experiments, such as the experimental Bayes variance and $p$-value moderation, which utilizes the parallel structure of the protein measurements and the analysis [@Ritchie2015].
+__R__'s formula interface for linear models is flexible, widely used, and well documented [@faraway2016extending].
+We use the formula interface to specify the models which should make it easy to reproduce an analysis performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` in other statistical programming languages.
+In addition we implemented features specific to high throughput experiments, such as the experimental Bayes variance and $p$-value moderation, which utilizes the parallel structure of the protein measurements and the analysis [@Ritchie2015].
 We also compute probabilities of differential protein regulation based on peptide level models [@Suomi2017bEnhanced].
-We use R6 to model the statistical modeling functionality in `r BiocStyle::Biocpkg("prolfqua")` (see Figure \@ref(fig:ContrastUML)).
+We used R6 classes encapsulate the statistical modelling functionality in `r BiocStyle::Biocpkg("prolfqua")` (see Figure \@ref(fig:ContrastUML)).
+We did specify a contrast interface (ContrastInterface), and there are various implementations allowing to perform differential expression analysis given linear or mixed effect models (Contrast), variance shrinkage (ContrastModerated), or impute contrasts in case of observations missing for an entire group of samples (ContrastSimpleImpute) can also be used to encapsulate and integrate differential expression analysis results of external tools such as SAINTExpress or proDA.
 
 (ref:ContrastUML) UML diagram of modeling and contrast related classes.
-Different strategies, e.g., _lm_, _lmer_, and _glm_ can be used to fit various types of models using a model builder (see code).
-All classes computing contrasts implement the _Contrast_ interface.
+Different strategies, e.g., _lm_, _lmer_, and _glm_ can be used to fit various types of models.
+The model builder method fit's the statistical model given the data and a strategy.
+The obtained model can then be used for analysis of variance or to estimate contrasts.
+All classes estimating the contrasts implement the _Contrast_ interface.
+The variance of the contrasts can be moderated using the `ContrastModerated` class.
 

### 2022-05-05 inst/MyArticle/paper.Rmd
Add dagger for wew

@@
   - Witold E.
-Wolski $^{\star}$^[[Functional Genomics Center Zurich (FGCZ)](https://fgcz.ch) - Swiss Federal Institute of Technology in Zurich \| University of Zurich; Winterthurerstrasse 190, CH-8057 Zurich, Switzerland.
-Correspondence$^\star$\texttt{wew@fgcz.ethz.ch}]
-  - Paolo Nanni$^{\star}$
-  - Jonas Grossmann$^{\star}$
-  - Maria d'Errico$^{\star}$
-  - Ralph Schlapbach$^{\star}$
-  - Christian Panse$^{\star}$
+Wolski $^{\dagger}$^[[Functional Genomics Center Zurich (FGCZ)](https://fgcz.ch) - Swiss Federal Institute of Technology in Zurich \| University of Zurich; Winterthurerstrasse 190, CH-8057 Zurich, Switzerland.
+$^\dagger$Correspondence \texttt{wew@fgcz.ethz.ch}]
+  - Paolo Nanni$^{\ast}$
+  - Jonas Grossmann$^{\ast}$
+  - Maria d'Errico$^{\ast}$
+  - Ralph Schlapbach$^{\ast}$
+  - Christian Panse$^{\ast}$
 output: 

### 2022-05-05 inst/MyArticle/paper.Rmd
Add urlcolor: blue to the yaml.
code snippet panel naming

@@
      The R package is available through https://github.com/fgcz/prolfqua.
-
+urlcolor: blue
 ---
@@
 
-## plotting
+## plotting panels
 pl <- lfqd$get_Plotter()
-p_a <- pl$intensity_distribution_density() +
+panelA <- pl$intensity_distribution_density() +
   ggplot2::labs(tag = "A") + ggplot2::theme(legend.position = "none")
@@
 ## plotting
-p_b <- agr$plot()$plots[[1]] + ggplot2::labs(tag = "B")
-p_c <- lfqp$get_Stats()$violin() + ggplot2::labs(tag = "C")
+panelB <- agr$plot()$plots[[1]] + ggplot2::labs(tag = "B")
+panelC <- lfqp$get_Stats()$violin() + ggplot2::labs(tag = "C")
 pl <- lfqp$get_Plotter()
-p_d <- pl$boxplots()$boxplot[[1]] + ggplot2::labs(tag = "D")
-ggpubr::ggarrange(p_a, p_b, p_c, p_d)
+panelD <- pl$boxplots()$boxplot[[1]] + ggplot2::labs(tag = "D")
+ggpubr::ggarrange(panelA, panelB, panelC, panelD)
 ```

### 2022-05-05 inst/MyArticle/paper.Rmd
Code snippet introdice panel A,B,...

@@
   - Witold E.
-Wolski $^{\dagger}$^[[Functional Genomics Center Zurich (FGCZ)](https://fgcz.ch) - Swiss Federal Institute of Technology in Zurich \| University of Zurich; Winterthurerstrasse 190, CH-8057 Zurich, Switzerland.
+Wolski $^{\dagger}$^[[Functional Genomics Center Zurich (FGCZ)](https://fgcz.ch) - [Swiss Federal Institute of Technology in Zurich](https://ethz.ch) \| [University of Zurich](https://uzh.ch); Winterthurerstrasse 190, CH-8057 Zurich, Switzerland.
 $^\dagger$Correspondence \texttt{wew@fgcz.ethz.ch}]
@@
 
-## visualize results
+## visualize results in panels
 pl <- contrasts$merged$get_Plotter() 
-p1 <- pl$histogram()$p.value
-p2 <- pl$volcano()$FDR +
-   ggplot2::theme(legend.position = "bottom")
+panelA <- pl$histogram()$p.value + ggplot2::labs(tag = "A")                                                        
+panelB <- pl$volcano()$FDR +
+   ggplot2::theme(legend.position = "bottom") +
+   ggplot2::labs(tag = "B")
 
-gridExtra::grid.arrange(p1, p2, ncol = 2)
+gridExtra::grid.arrange(panelA, panelB, ncol = 2)
 ```
@@
 
-norm <- res |> dplyr::group_by(contrast,whatfix) |> dplyr::summarize(meanpAUC_10 = mean(pAUC_10))
+norm <- res |> dplyr::group_by(contrast,whatfix) |>
+    dplyr::summarize(meanpAUC_10 = mean(pAUC_10))
 res <- dplyr::inner_join(res, norm)
@@
   ggplot2::theme(axis.text.x = ggplot2::element_text(angle = -90, vjust = 0.5)) +
-  ggplot2::geom_hline(ggplot2::aes(yintercept = meanpAUC_10), color = "red") + ggplot2::xlab("")
+  ggplot2::geom_hline(ggplot2::aes(yintercept = meanpAUC_10), color = "red") +
+  ggplot2::xlab("") +
+  ggplot2::labs(tag = "A")
 
@@
             vjust = -1, hjust = -0.2, angle = -90) 
-p3 <- nrgg
+p3 <- nrgg + ggplot2::labs(tag = "B")
 
@@
    ggplot2::geom_abline(intercept = 0, slope = 1, color = 2) + 
-   ggplot2::theme(legend.position = "bottom")
+   ggplot2::theme(legend.position = "bottom") +
+  ggplot2::labs(tag = "C")
 

### 2022-05-05 inst/MyArticle/paper.Rmd
change order of code

@@
 
-## plotting panels
+## infer protein intensities from peptide intensity
+agr <- lfqd$get_Aggregator()
+lfqp <- agr$medpolish()
+
+## plotting panels A-D
 pl <- lfqd$get_Plotter()
@@
   ggplot2::labs(tag = "A") + ggplot2::theme(legend.position = "none")
-
-## infer protein intensities from peptide intens
-agr <- lfqd$get_Aggregator()
-lfqp <- agr$medpolish()
-
-## plotting
 panelB <- agr$plot()$plots[[1]] + ggplot2::labs(tag = "B")

### 2022-05-05 inst/MyArticle/paper.Rmd
Text cosmetics R6

@@
 Using the tidy table to model the data ensures interoperability with other proteomics-related packages that manage their data with tidy-tables, e.g., `r BiocStyle::CRANpkg("protti")`.
-Furthermore, the use of `R6` classes, which encapsulate the configuration and the data, allows writing very concise code (no function arguments needed), and auto-completion support in the Rstudio editor makes it easy for novices to explore `r BiocStyle::Githubpkg("fgcz/prolfqua")`s functionality.
+Furthermore, the consistent implementation of `R6` classes, which encapsulate the configuration and the data, allows writing very concise code (no function arguments needed), and auto-completion support in the Rstudio editor makes it easy for novices to explore `r BiocStyle::Githubpkg("fgcz/prolfqua")`s functionality.
 

### 2022-05-05 inst/MyArticle/paper.Rmd
Add bfabric paragraph in Discussion

@@
 
-## plotting panels A-D
+## plot panels A-D
 pl <- lfqd$get_Plotter()
@@
 
+To ease the usage barriers of the R package to users not being so familiar with statistics and R programming, we integrated some workflows into our Local Laboratory Information Management System (LIMS) [b-fabric](https://fgcz-bfabric.uzh.ch/) @bfabric.
+This integration enables our users to select the input and basic settings in a graphical user interface.
+As an output, the user receives a report including the R markdown file and all necessary input files to reprocess the analysis on their devices.
+In this way, we also support the reproducible science efforts of the community.
+
 # Conclusion
@@
 <div id="refs"></div>
+
+\newpage
 

### 2022-05-05 inst/MyArticle/paper.Rmd
Update conclusions

--- 2022-05-05 inst/MyArticle/paper.Rmd
@@
   - Witold E.
-Wolski $^{\dagger}$^[[Functional Genomics Center Zurich (FGCZ)](https://fgcz.ch) - [Swiss Federal Institute of Technology in Zurich](https://ethz.ch) \| [University of Zurich](https://uzh.ch); Winterthurerstrasse 190, CH-8057 Zurich, Switzerland.
+Wolski $^{\dagger}$^[[Functional Genomics Center Zurich (FGCZ)](https://fgcz.ch) - Swiss Federal Institute of Technology in Zurich \| University of Zurich; Winterthurerstrasse 190, CH-8057 Zurich, Switzerland.
 $^\dagger$Correspondence \texttt{wew@fgcz.ethz.ch}]
@@
 For example, we implemented strategies to estimate protein intensities from peptide intensities: top N, Tukey's median polish, robust linear models [@Grossmann2010, @goeminne2020msqrob].
-Furthermore, peptide or protein intensities can be scaled using the robust scaling method or transformed using _quantile_ normalization or `r BiocStyle::Biocpkg("vsa")` to remove systematic differences among samples and heteroscedasticity.
+Furthermore, peptide or protein abundances can be scaled and transformed, using robust scaling, transformed using _quantile_ normalization or `r BiocStyle::Biocpkg("vsa")` to remove systematic differences among samples and heteroscedasticity.
 In this respect, `r BiocStyle::Githubpkg("fgcz/prolfqua")` can be compared with other R packages such `r BiocStyle::Biocpkg("DEP")` [@DEP2018] or `r BiocStyle::Biocpkg("POMA")` [@POMA2021] which support the entire LFQ analysis pipeline.
-However, we also implemented functionality to streamline the benchmarking of the implemented methods.
+
+We also implemented functionality to streamline the benchmarking of the implemented methods.
+We focus on the comparison of the various modeling approaches.
 We use the Ionstar [@shen2018ionstar] dataset to benchmark the analysis pipeline and modeling methods implemented within `r BiocStyle::Githubpkg("fgcz/prolfqua")` and to compare our results with those of `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`.
-
-An essential parameter of any protein quantification pipeline, not discussed further, is the normalization of the intensities within the samples to make them comparable [@pursiheimo2015optimization].
-All benchmarks we run with the same data normalization method.
-The method used to infer protein intensities from peptide intensities is an additional important factor.
-For instance, the original _proDA_ publication uses MaxLFQ [@cox2014accurate] protein estimates.
-However, when using those protein abundance estimates, _proDA_ performs significantly worse when using the protein estimates obtained from peptide intensity estimates using Tukey's median polish,  $pAUC_{10}(\textrm{t-statistics}) = 66%$ and $pAUC_{10}(\textrm{t-statistics}) = 73%$ respectively (see Articles on [@prolfquaGithubIO] for more details).
-Last but not least, the software, identifying proteins, and generating the quantification values can significantly contribute to the performance of the entire pipeline [@Cox2008MaxQuant,@yu2020fast].
-However, here we focus on the comparison of the various modeling approaches.
 
@@
 
-## Usage
+## Implementing an analysis workflow
 
@@
 
-## infer protein intensities from peptide intensity
+## plotting panels
+pl <- lfqd$get_Plotter()
+panelA <- pl$intensity_distribution_density() +
+  ggplot2::labs(tag = "A") + ggplot2::theme(legend.position = "none")
+
+## infer protein intensities from peptide intens
 agr <- lfqd$get_Aggregator()
@@
 
-## plotting panels A-D
-pl <- lfqd$get_Plotter()
-panelA <- pl$intensity_distribution_density() +
-  ggplot2::labs(tag = "A") + ggplot2::theme(legend.position = "none")
+## plotting
 panelB <- agr$plot()$plots[[1]] + ggplot2::labs(tag = "B")
@@
 
-## visualize results in panels
+## visualize results
 pl <- contrasts$merged$get_Plotter()
-panelA <- pl$histogram()$p.value + ggplot2::labs(tag = "A")                                                        
-panelB <- pl$volcano()$FDR +
-   ggplot2::theme(legend.position = "bottom") +
-   ggplot2::labs(tag = "B")
+p1 <- pl$histogram()$p.value
+p2 <- pl$volcano()$FDR +
+   ggplot2::theme(legend.position = "bottom")
 
-gridExtra::grid.arrange(panelA, panelB, ncol = 2)
+gridExtra::grid.arrange(p1, p2, ncol = 2)
 ```
@@
 
-(ref:benchmarkROC)  Left panel - Difference to mean partial area under the ROC curve for various models, at $10\%$ FPR, by fold change.
-
-```{r benchmarkROC, fig.cap="(ref:benchmarkROC)", out.width = '90%', fig.width=12, fig.height=10}
-ttt <- sapply(allBenchmarks, function(x){x$complete(TRUE)})
+```{r benchmarkROC}
+ttt <- sapply(allBenchmarks, function(x){x$complete(FALSE)})
 
@@
 
-norm <- res |> dplyr::group_by(contrast,whatfix) |>
-    dplyr::summarize(meanpAUC_10 = mean(pAUC_10))
+norm <- res |> dplyr::group_by(contrast,whatfix) |> dplyr::summarize(meanpAUC_10 = mean(pAUC_10))
 res <- dplyr::inner_join(res, norm)
@@
   ggplot2::theme(axis.text.x = ggplot2::element_text(angle = -90, vjust = 0.5)) +
-  ggplot2::geom_hline(ggplot2::aes(yintercept = meanpAUC_10), color = "red") +
-  ggplot2::xlab("") +
-  ggplot2::labs(tag = "A")
+  ggplot2::geom_hline(ggplot2::aes(yintercept = meanpAUC_10), color = "red") + ggplot2::xlab("")
 
@@
 We also benchmark, if the $FDR$ obtained from a model is an unbiased estimate of the false discovery proportion $FDP$.
-Figure \@ref(fig:FDRfdp) right panel draws on the horizontal axis the FDR determined from the model, and on the vertical axis, the FDP obtained from the confusion matrix.
+Figure \@ref(fig:FDRfdp) panel C draws on the horizontal axis the FDR determined from the model, and on the vertical axis, the FDP obtained from the confusion matrix.
 We see that most of the lines are below the diagonal, indicating that the FDR estimates obtained with `r BiocStyle::Githubpkg("fgcz/prolfqua")` are modestly conservative for this benchmark dataset provide.
@@
 
-(ref:FDRfdp) Left panel - Partial area under the ROC curve at $10\%$ FPR for all contrasts and three different statistics: the difference among conditions, the scaled $p$-value (sign(diff) * p.value) and the $t$-statistics (higher is better).
-The red line average area under the curve.
-Left panel - Number and percentage of estimated contrasts for each modeling method (higher is better).
-Right panel - Compare FDR estimate with false discovery proportion (FDP) closer to the diagnoal (red continous line is better).
+(ref:FDRfdp) Panel A - Partial area under the ROC curve at $10\%$ FPR (pAUC_10) for all contrasts and three different statistics: the difference among conditions, the scaled $p$-value (sign(diff) $\cdot$ p.value) and the $t$-statistics (higher is better).
+The red line indicates the average area under the curve of all methods.
+Panel B - Number of estimated contrasts for each modeling method (higher is better).
+Panel C - Plots the false discovery proportion (FDP) as a function of the FDR.
+Ideally the FDR should be equal to the FDP.
+Therefore larger distances from the diagonal are worse.
 
@@
    ggplot2::geom_abline(intercept = 0, slope = 1, color = 2) + 
-   ggplot2::theme(legend.position = "bottom") +
-  ggplot2::labs(tag = "C")
+   ggplot2::theme(legend.position = "bottom") +  ggplot2::labs(tag = "C")
 
@@
 
+We focused in our benchmark on the comparison of the statistical modelling methods and fixing the pre-processing steps.
+However, there are other equally or even more important parameters of any protein quantification pipeline.
+One of them is the normalization of the intensities within the samples to remove systematic differences and to make them comparable [@pursiheimo2015optimization].
+The method used to infer protein intensities from peptide intensities is an additional important factor [cite TopN].
+For instance, the original _proDA_ publication uses MaxLFQ [@cox2014accurate] protein estimates.
+However, when using MaxLFQ inensities reported by MaxQuant, _proDA_ performs significantly worse ($pAUC_{10}(\textrm{t-statistics}) = 66%$) compared with results obtained when protein abundances are estimated from peptide abundances using Tukey's median polish ($pAUC_{10}(\textrm{t-statistics}) = 73%$).
+Last but not least, the software [@Cox2008MaxQuant,@yu2020fast], identifying proteins, and generating the quantification values can also significantly contribute to the performance of the entire pipeline, altering the number of identified proteins and the sensitivity and specificity of the differential expression analysis.
+
 We assessed the performance of different modeling approaches implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")`, `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`, using a benchmark dataset whose ground truth is known.
@@
 Specifically, in more realistic experiments the observed variances will be higher and sensitivities and specificity, will be worse given the same sample sizes.
-Furthermore, the proportion of missing observations in real world datasets might be higher.
+Furthermore, the proportion of missing observations in real world datasets might also be higher.
 
-However, we can conclude that if we want to order the proteins according to the likelihood of being differentially regulated, to perform gene set enrichment analysis [@subramanian2005gene], the $t$-statistic is better suited than the fold-change estimate.
-Modeling the degrees of freedom when computing the $p$-values might further improve the inference, However, the observed change is minuscule (see Figure \@ref(fig:)).
-Furthermore, for the fixed effect linear model, the empirical Bayes variance shrinkage, as suggested by [@Smyth2004linear], consistently improves the ranking of proteins compared with the unmoderated estimates (data not shown).
+We can conclude that if we want to order the proteins according to the likelihood of being differentially regulated, to perform gene set enrichment analysis [@subramanian2005gene], the $t$-statistic is better suited than the fold-change estimate.
+Modeling the degrees of freedom when computing the $p$-values might further improve the inference, however, the observed this improvement is minuscule (see Figure \@ref(fig:FDRfdp) panel A).
 However, there is no such effect for the mixed effect model, most likely because the degrees of freedom are erroneously estimated for a large proportion of models.
+Therefore, for the fixed effect linear model, the empirical Bayes variance shrinkage, as suggested by [@Smyth2004linear], consistently improves the ranking of proteins compared with the unmoderated estimates (not shown), however fails to do so for mixed effect models.
 
-Computing the statistics at the peptide level, e.g., the $t$-statistics or $p$-value, then summarizing these statistics using their median produces the highest AUC scores among all the tested models.
-By using the beta distribution to model the number of peptides observed, we can further improve the pAUC scores (see Figure \@ref(fig:FDRfdp) top).
+Computing the statistics at the peptide level, e.g., the $t$-statistics or $p$-value, then summarizing these statistics using their median produces the highest AUC scores among all the tested models (see Figure \@ref(fig:FDRfdp) Panel A left).
+By using the beta distribution to model the number of peptides observed, we can further improve the pAUC scores (see Figure \@ref(fig:FDRfdp) Panel A center).
 However, the properties of beta-based probabilities are not well understood; for instance, the p-values are not uniformely distributed under the null hypothesis (not shown).
-Furthermore, the FDR estimates obtained when correcting for multiple testing with the Benjamini-Hochberg method are biased overestimating the false discovery proprotion (see Figure \@ref(fig:FDRfdp) right panel).
+Furthermore, the FDR estimates obtained when correcting for multiple testing with the Benjamini-Hochberg method are biased overestimating the false discovery proportion (see Figure \@ref(fig:FDRfdp) Panel C).
 Therefore, we can not recommend this method if an unbiased estimate of FDR is essential, which is frequently the case.
-In addition, peptides are stronger affected by missing values, reducing the number of contrasts we could estimate for the dataset (see Figure \@ref(fig:FDRfdp) bottom left).
-To address the issue of missing observation, we could either run `r BiocStyle::Biocpkg("proDA")` or the imputation of group averages method discussed here and afterwards apply `r BiocStyle::Biocpkg("PECA")`.
+In addition, peptides are stronger affected by missing values, reducing the number of contrasts we could estimate for the dataset using this method (see Figure \@ref(fig:FDRfdp) Panel C).
 
-The probabilistic dropout analysis implemented in the `r BiocStyle::Biocpkg("proDA")` produces inferences comparable to other methods (Figure \@ref(fig:FDRfdp) for the largest number of proteins (Figure \@ref(fig:FDRfdp) left panel).
-Moreover, the estimated fold changes are unbiased and show high diagnostic accuracy.
+The probabilistic dropout analysis implemented in the `r BiocStyle::Biocpkg("proDA")` produces inferences comparable to those of other methods (Figure \@ref(fig:FDRfdp) Panel A) for the largest number of proteins (Figure \@ref(fig:FDRfdp) Panel B).
+Moreover, the estimated fold changes are unbiased (not shown) and show high diagnostic accuracy (\@ref(fig:FDRfdp) Panel A).
 Furthermore, the scaled $p$-values or the $t$-statistics are on par with those obtained by the linear model with variance moderation.
@@
 Furthermore, the use of `R6` classes, which encapsulate the configuration and the data, allows writing very concise code (no function arguments needed), and auto-completion support in the Rstudio editor makes it easy for novices to explore `r BiocStyle::Githubpkg("fgcz/prolfqua")`s functionality.
+To simplify integration of prolfqua with Bioconductor based workflows there is an method to convert the LFQData class into a `r BiocStyle::Biocpkg("SummarizedExperiment")`.
 
@@
 
-Leveraging on these computer experiments, we can provide the following advice.
-First, aggregate the peptide intensities to protein intensities using a robust regression method.
-Second, use linear models as long as they suffice to model your experimental design.
-Use mixed effect models only if you have to, e.g., repeated measurements and if the sample sizes are large.
-If you use fixed-effect linear models, apply variance moderation since this improves the $t$-statistics and $p$-value estimates.
+When comparing statistical modeling methods for the differential expression analysis, we assessed performance measures such as the number of estimated contrasts, the pAUC, and if the FDR is an unbiased estimate of the FDR.
+It is relevant that an analysis pipeline shows good performance in all these categories.
+Leveraging these computer experiments, we can provide the following advice.
+First, estimate protein abundances from peptide abundances using a robust or nonparametric regression method.
+Second, use linear models as long as they suffice to model your experimental design because they show good performance in all categories.
+Third, use mixed effect models only if you have to, e.g., repeated measurements and if the sample sizes are large.
+If you use fixed-effect linear models, apply variance moderation to improve the $t$-statistics and $p$-value estimates.
 If you want to sort your protein lists to perform gene set enrichment analysis, use the $t$-statistic instead of the fold change.
-You can select proteins to study them further based on the FDR, although expect a $bias$.
 
-In summary `r BiocStyle::Githubpkg("fgcz/prolfqua")` is an easy-to-use and feature rich R package to analyse quantitative mass spectrometric data, report results, and benchmark MS software and statistical methods.
+The r-package proDA and prolfqua model the missing data directly, performing similarly, while MSstats imputes the data using an accelerated failure model.
+MSstats estimates group differences for fewer proteins and does not achieve a higher pAUC score despite imputation.
+Furthermore, \@ref(fig:FDRfdp) Panel C shows that when using MSstats the false proportion of false discoveries might be very high even when filtering for a low FDR.
+
+In summary, `r BiocStyle::Githubpkg("fgcz/prolfqua")` is an easy-to-use and feature-rich R package to analyze quantitative mass spectrometric data, report results, and benchmark MS software and statistical methods.
 We provide more details at the website https://github.com/fgcz/prolfqua/.

### 2022-05-05 inst/MyArticle/paper.Rmd
code snippet for MSFragger

@@
 
-The following code snipet demomnstarte the usage from MSFragger ouput.
+The following code demonstrates how an analysis of MSFragger ouput can be performed using prolfqua.
 
 ```{r}
+datadir <- file.path(find.package("prolfquadata") , "quantdata")
+inputFragfile <-  file.path(datadir, "MSFragger_IonStar2018_PXD003881.zip")
+inputAnnotation <- file.path(datadir, "annotation_Ionstar2018_PXD003881.xlsx")
+# read input annotation
+annotation <- readxl::read_xlsx(inputAnnotation)
+
+protein <- tibble::as_tibble(read.csv(unz(inputFragfile,"IonstarWithMSFragger/combined_protein.tsv"),
+                             header = TRUE, sep = "\t", stringsAsFactors = FALSE))
+
+# read combined_protein.tsv 
+protein <- prolfqua::tidy_MSFragger_combined_protein(protein)
+# remove proteins identified by a single peptide
+protein <- protein |> dplyr::filter(unique.stripped.peptides > 1)
+
+# annotate the data
+merged <- dplyr::inner_join(annotation, protein)
+
+atable <- prolfqua::AnalysisTableAnnotation$new()
+atable$fileName = "raw.file"
+# specify column containing protein identifiers
+atable$hierarchy[["protein_Id"]] <- c("protein")
+
+# column with protein abundances
+atable$setWorkIntensity("total.intensity")
+
+# the factors of the analysis
+atable$factors[["dilution."]] = "sample"
+atable$factors[["run"]] = "run_ID"
+
+config <- prolfqua::AnalysisConfiguration$new(atable)
+
+adata <- prolfqua::setup_analysis(merged, config)
+lfqdata <- prolfqua::LFQData$new(adata, config)
+# remove small intensities
+lfqdata$remove_small_intensities()
 

### 2022-05-05 inst/MyArticle/paper.Rmd

@@
 
-## plotting panels
+## infer protein intensities from peptide intensity
+agr <- lfqd$get_Aggregator()
+lfqp <- agr$medpolish()
+
+## plot panels A-D
 pl <- lfqd$get_Plotter()
@@
    ggplot2::geom_abline(intercept = 0, slope = 1, color = 2) + 
-   ggplot2::theme(legend.position = "bottom") +  ggplot2::labs(tag = "C")
+   ggplot2::theme(legend.position = "bottom") +
+  ggplot2::labs(tag = "C")
 
@@
 To simplify integration of prolfqua with Bioconductor based workflows there is an method to convert the LFQData class into a `r BiocStyle::Biocpkg("SummarizedExperiment")`.
+
+To ease the usage barriers of the R package to users not being so familiar with statistics and R programming, we integrated some workflows into our Local Laboratory Information Management System (LIMS) [b-fabric](https://fgcz-bfabric.uzh.ch/) @bfabric.
+This integration enables our users to select the input and basic settings in a graphical user interface.
+As an output, the user receives a report including the R markdown file and all necessary input files to reprocess the analysis on their devices.
+In this way, we also support the reproducible science efforts of the community.
 

### 2022-05-05 inst/MyArticle/paper.Rmd
some additions to the supplment

@@
 
-If there are no observation in a group, we assume that there are no observations because of the limit of detection (LOD).
-Furthermore, we assume that the variance of the protein is constant in all the conditions.
-We impute the mean estimate of the group by using the average of $1\%$ smallest group mean intensities in the experiment.
-To estimate the variance in the condition, we use the pooled variance based on all conditions:
+If there is no observation in a group, we assume that the protein was unobserved because the protein abundance is below the limit of detection (LOD) of the MS instrument.
+We estimate the detection limit as follows: if a protein is quantified just in a single sample of a treatment group, then it is at the instrument's detection limit.
+We compute the mean of those abundances and use it as the group mean if a protein is absent in a treatment group.
+
+When computing differences among two groups $a$ and $b$, we will use the group mean $\bar{a}$ estimated from the data or the detection limit $a_d$ if no observations are present in the group.
+If for instance there are no observations in group $b$, then, $\Delta_{a,b} = \bar{a} - b_{dl}$ if $a > b_{dl}$ or $\Delta = 0$ if $a < b_{dl}$.
+
+Furthermore, we assume that the variance of the protein is constant in all the conditions. To estimate the variance in the condition, we use the pooled variance based on all conditions:
 
@@
 
-```{r exampleContrasts, fig.cap="(ref:exampleContrasts)", echo=TRUE, out.width = '90%'}
+```{r exampleContrasts, fig.cap="(ref:exampleContrasts)", echo=TRUE, out.width = '90%', fig.width = 8, fig.height = 6}
 contrasts <- c(
@@
 
+The r-package proDA and prolfqua model the missing data directly, while MSstats imputes the data using an accelerated failure model.
+Despite imputation, MSstats do not estimate group differences for more proteins and do not achieve a higher pAUC score than prolfqua.
+Furthermore, \@ref(fig:FDRfdp) Panel C shows that when using MSstats the proportion of false discoveries might be very high even when filtering for a low FDR.
+
 Using the tidy table to model the data ensures interoperability with other proteomics-related packages that manage their data with tidy-tables, e.g., `r BiocStyle::CRANpkg("protti")`.
@@
 This integration enables our users to select the input and basic settings in a graphical user interface.
-As an output, the user receives a report including the R markdown file and all necessary input files to reprocess the analysis on their devices.
+As an output, the user receives a report including the R markdown file and all necessary input files to reprocess the analysis using R on their devices.
 In this way, we also support the reproducible science efforts of the community.
@@
 If you want to sort your protein lists to perform gene set enrichment analysis, use the $t$-statistic instead of the fold change.
-
-The r-package proDA and prolfqua model the missing data directly, performing similarly, while MSstats imputes the data using an accelerated failure model.
-MSstats estimates group differences for fewer proteins and does not achieve a higher pAUC score despite imputation.
-Furthermore, \@ref(fig:FDRfdp) Panel C shows that when using MSstats the false proportion of false discoveries might be very high even when filtering for a low FDR.
+Last but not least, do not impute missing data but model it instead.
 
@@
 We provide more details at the website https://github.com/fgcz/prolfqua/.
+This document was created using Rmarkdown, and all the code need to replicate the document and the presented benchmarks is available at : https://github.com/wolski/prolfquabenchmark/.
 
@@
 
-## Creating a prolfqa configuration
+## Creating a prolfqua configuration
 
-The following code demonstrates how an analysis of MSFragger ouput can be performed using prolfqua.
+The following code demonstrates how an analysis of MSFragger output can be performed using prolfqua.
+First, a tidy table with protein abundances is generated by reading protein abundances reported by MSFragger in the combined_protein.tsv file.
+The samples are annotated using the data stored in the annotation.xlsx file. Next, an _AnalysisTableAnnotation_ R6 object is created. 
+Since bottom-up proteomics data is hierarchical a protein has peptides, which might be modified and create precursors.
+To describe this hierarchy, AnalysisTableAnnotation has a hierarchy field that stores a list with an entry for each hierarchy level.
+Since combined_portein.tsv only holds protein level data, the hierarchy list has only one element, and we need to specify which column does contain the protein identifiers.
+We also need to specify which column contains the protein abundances we want to use for the data analysis.
+Finally, we have to specify which columns contain the explanatory variables of the analysis.
+The AnalysisTableAnnotation has the field factors, a list with as many entries as explanatory variables.
+Here we have only one explanatory variable, the dilution, specified in the column 'sample.'
 
@@
 
-## Comparing benchmark results
+## Comparing benchmark results for proDA and prolfqua
 
@@
 We see that the fold change estimates obtained with the `r BiocStyle::Biocpkg("proDA")` package show better diagnostic accuracy.
+Figure \@ref(fig:proDAAllBenchmarks), compares the fold change estimates obtained using both methods.
+However, the $t-statistics$ or $p-values$ show slightly worse diagnostic accuracy.
 
@@
 
-```{r proDAAllBenchmarks}
+```{r proDAAllBenchmarks, fig.cap="Comparison of fold change estimates produced by proDA and using the missing data modelling method implemented in prolfqua."}
 proda <- allBenchmarks$benchmark_proDA$data()
 imp <- allBenchmarks$benchmark_imputation$data()
+imp <- allBenchmarks$benchmark_ProtModerated$data()
 

### 2022-05-05 inst/MyArticle/paper.Rmd
fix methods formula

@@
 When computing differences among two groups $a$ and $b$, we will use the group mean $\bar{a}$ estimated from the data or the detection limit $a_d$ if no observations are present in the group.
-If for instance there are no observations in group $b$, then, $\Delta_{a,b} = \bar{a} - b_{dl}$ if $a > b_{dl}$ or $\Delta = 0$ if $a < b_{dl}$.
+If for instance there are no observations in group $b$, then, $\Delta_{a,b} = \bar{a} - b_{dl}$ if $\bar{a} > b_{dl}$ or $\Delta_{a,b} = 0$ if $\bar{a} < b_{dl}$.
 

### 2022-05-06 inst/MyArticle/paper.Rmd
Add command-line completion graphic in supplement

@@
 
-```{r}
+```{r error=TRUE}
 datadir <- file.path(find.package("prolfquadata") , "quantdata")
@@
 
+(ref:tabCompletion) The screenshot displays the command-line completion (tab completion) of RStudio on the `prolfqua::LFQData` R6 object. In the example, it shows the getter methods of the object.
+
+```{r tabCompletion, echo=FALSE, fig.cap="(ref:tabCompletion)", out.width = '66%'}
+knitr::include_graphics("codeSnippet1TabCompletion.png")
+```
+
 ## Session information

### 2022-05-06 inst/MyArticle/paper.Rmd
Add formular 1 and 2 as eqmarray

@@
 
+\newpage
 # Methods
@@
 
-Given a linear model contrasts can be computed by $\hat{\beta_{c}} = \sum l\beta_{m}$ and $\textrm{var} \hat\beta_c = l\sigma^2 (X^T X)^{-1} l^T$, with $X$ being the design matrix, $\beta_m$ the model parameters and $l$ an array of coefficients.
+Given a linear model contrasts can be computed by 
+\begin{eqnarray} 
+\hat{\beta_{c}} &=& \sum l\beta_{m} 
+\end{eqnarray}
+
+and
+
+\begin{eqnarray} 
+\textrm{var} \hat\beta_c &=& l\sigma^2 (X^T X)^{-1} l^T
+\end{eqnarray}
+
+with $X$ being the design matrix, $\beta_m$ the model parameters and $l$ an array of coefficients.
 The degrees of freedom for the contrast are equal to the residual degrees of freedom of the linear model.

### 2022-05-06 inst/MyArticle/paper.Rmd
supp

--- 2022-05-06 inst/MyArticle/paper.Rmd
@@
 
-The code snippets in this section demonstrate the basic usage of the `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package. As input data we use Ionstar dataset.
+The code snippets in this section demonstrate the basic usage of the `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package.
+As input data we use subset of the Ionstar dataset containing $163$ proteins and $1258$ peptides.
 Peptide intensities are $\log_2$ transformed and robust z-score scaled but then rescaled to preserves the original range of the data using the method `robscale`.
@@
 
-(ref:exampleContrasts) Left panel - histogram of $p$-values. Right panel - volcano plots.
+(ref:exampleContrasts) Panel A - Histogram showing the distribution of p-values for $163$ proteins.
+Panel B -  Volcano plot showing $-\log_{10}$ transformed FDR as function of the difference between groups for $163$ proteins.
 
-```{r exampleContrasts, fig.cap="(ref:exampleContrasts)", echo=TRUE, out.width = '90%', fig.width = 8, fig.height = 6}
+```{r exampleContrasts, fig.cap="(ref:exampleContrasts)", echo=TRUE, out.width = '90%', fig.width = 8, fig.height = 5}
+# specify differences among conditions
 contrasts <- c(
@@
 This integration enables our users to select the input and basic settings in a graphical user interface.
-As an output, the user receives a report including the R markdown file and all necessary input files to reprocess the analysis using R on their devices.
+Then, as an output, the user receives a report including the R markdown file and all necessary input files to reprocess the analysis using R on their devices.
 In this way, we also support the reproducible science efforts of the community.
@@
 Finally, we have to specify which columns contain the explanatory variables of the analysis.
-The AnalysisTableAnnotation has the field factors, a list with as many entries as explanatory variables.
-Here we have only one explanatory variable, the dilution, specified in the column 'sample.'
+The AnalysisTableAnnotation has the field factors, a list with as many entries as there are explanatory variables.
+Here we have included two explanatory variable, the dilution, specified in the column 'sample', and 'run' in the column 'run_ID' representing the order of the measurement.
 
@@
 
-Here we compare the benchmark results for the `r BiocStyle::Biocpkg("proDA")` model and the linear model with variance moderation, but using only the protein where inference could be performed for both models.
+Here we compare the benchmark results for the `r BiocStyle::Biocpkg("proDA")` model and the linear model with variance moderation.
 We see that the fold change estimates obtained with the `r BiocStyle::Biocpkg("proDA")` package show better diagnostic accuracy.
+However, the $t-statistics$ or $p-values$ show slightly worse diagnostic accuracy (Figure \@ref(fig:proDABenchmarks).
 Figure \@ref(fig:proDAAllBenchmarks), compares the fold change estimates obtained using both methods.
-However, the $t-statistics$ or $p-values$ show slightly worse diagnostic accuracy.
 
-```{r proDABenchmarks}
+(ref:proDABenchmarks) The barplots show the difference of the $pAUC_10$ obtained when using the proDA package or when using prolfqua.
+Postitive values indicate a better diagnostic performance of proDA while negative values a better diagnostic performance of prolfqua.
+
+```{r proDABenchmarks, fig.cap="(ref:proDABenchmarks)"}
 prodB <- allBenchmarks$benchmark_proDA$clone(deep=TRUE)
@@
 modSum <- modB$pAUC_summaries()
-
 prodaMod <- dplyr::semi_join(proda, mod, by = c("contrast", "protein_Id"))
 prodB$.data <- prodaMod
+
 d <- prodB$missing_contrasts()
@@
 
-```{r proDAAllBenchmarks, fig.cap="Comparison of fold change estimates produced by proDA and using the missing data modelling method implemented in prolfqua."}
+```{r proDAAllBenchmarks, fig.cap="Comparison of fold change estimates produced by proDA and prolfqua simple impute method."}
 proda <- allBenchmarks$benchmark_proDA$data()
@@
 x <- dplyr::inner_join(imp, proda, by = c("contrast","protein_Id", "TP", "species"), suffix = c(".imp",".proda"))
-ggplot2::ggplot(x,ggplot2::aes( x = diff.imp, y = diff.proda) ) + ggplot2::geom_point() + ggplot2::facet_wrap(~contrast + species) + ggplot2::geom_hline(yintercept = 0, color="red") + ggplot2::geom_vline(xintercept = 0, color="red")
+ggplot2::ggplot(x,ggplot2::aes( x = diff.imp, y = diff.proda) ) + ggplot2::geom_point() + ggplot2::facet_wrap(~contrast + species) + ggplot2::geom_hline(yintercept = 0, color="lightgray") + ggplot2::geom_vline(xintercept = 0, color="lightgray") + ggplot2::geom_abline(slope = 1) + ggplot2::theme_minimal()
 
-```
-
-(ref:tabCompletion) The screenshot displays the command-line completion (tab completion) of RStudio on the `prolfqua::LFQData` R6 object. In the example, it shows the getter methods of the object.
-
-```{r tabCompletion, echo=FALSE, fig.cap="(ref:tabCompletion)", out.width = '66%'}
-knitr::include_graphics("codeSnippet1TabCompletion.png")
 ```

### 2022-05-06 inst/MyArticle/paper.Rmd

@@
 We use __R6__ classes to structure the functionality of the package (see Figure \@ref(fig:LFQData) and Figure \@ref(fig:ContrastUML)).
-__R6__ classes are well supported, e.g.
-by auto-completion features, and easy to use (OOP).
+__R6__ classes are well supported, e.g., by command-line completion features, see Figure \@ref{fig:tabCompletion}, and easy to use (OOP).
 
@@
 
-```{r proDAAllBenchmarks, fig.cap="Comparison of fold change estimates produced by proDA and prolfqua simple impute method."}
+(ref:proDAAllBenchmarks) Comparison of fold change estimates produced by proDA and using the missing data modelling method implemented in prolfqua.
+
+```{r proDAAllBenchmarks, fig.cap="(ref:proDAAllBenchmarks)"}
 proda <- allBenchmarks$benchmark_proDA$data()
@@
 
-x <- dplyr::inner_join(imp, proda, by = c("contrast","protein_Id", "TP", "species"), suffix = c(".imp",".proda"))
-ggplot2::ggplot(x,ggplot2::aes( x = diff.imp, y = diff.proda) ) + ggplot2::geom_point() + ggplot2::facet_wrap(~contrast + species) + ggplot2::geom_hline(yintercept = 0, color="lightgray") + ggplot2::geom_vline(xintercept = 0, color="lightgray") + ggplot2::geom_abline(slope = 1) + ggplot2::theme_minimal()
+x <- dplyr::inner_join(imp, proda, by = c("contrast","protein_Id", "TP", "species"), suffix = c(".lm",".proda"))
+ggplot2::ggplot(x,ggplot2::aes( x = diff.lm, y = diff.proda) ) + ggplot2::geom_point() + ggplot2::facet_wrap(~contrast + species) + ggplot2::geom_hline(yintercept = 0, color="lightgray") + ggplot2::geom_vline(xintercept = 0, color="lightgray") + ggplot2::geom_abline(slope = 1) + ggplot2::theme_minimal()
 
@@
 
-## Package sticker
+\newpage
+## Miscellaneous
+
+(ref:tabCompletion) The screenshot displays the command-line completion (tab completion) of RStudio on the `prolfqua::LFQData` R6 object. In the example, it shows the getter methods of the object.
+
+```{r tabCompletion, echo=FALSE, fig.cap="(ref:tabCompletion)", out.width = '66%'}
+knitr::include_graphics("codeSnippet1TabCompletion.png")
+```
 

### 2022-05-06 inst/MyArticle/paper.Rmd
Add SIB affiliation

--- 2022-05-06 inst/MyArticle/paper.Rmd
@@
   - Witold E.
-Wolski $^{\dagger}$^[[Functional Genomics Center Zurich (FGCZ)](https://fgcz.ch) - [Swiss Federal Institute of Technology in Zurich](https://ethz.ch) \| [University of Zurich](https://uzh.ch); Winterthurerstrasse 190, CH-8057 Zurich, Switzerland.
-$^\dagger$Correspondence \texttt{wew@fgcz.ethz.ch}]
+Wolski $^{\dagger\ddagger}$^[[Functional Genomics Center Zurich (FGCZ)](https://fgcz.ch) - [Swiss Federal Institute of Technology in Zurich](https://ethz.ch) \| [University of Zurich](https://uzh.ch); Winterthurerstrasse 190, CH-8057 Zurich, Switzerland.
+$^\ddagger$[Swiss Institute of Bioinformatics (SIB)](https://www.sib.swiss/) $^\dagger$Correspondence \texttt{wew@fgcz.ethz.ch}]
   - Paolo Nanni$^{\ast}$
-  - Jonas Grossmann$^{\ast}$
-  - Maria d'Errico$^{\ast}$
+  - Jonas Grossmann$^{\ast\ddagger}$
+  - Maria d'Errico$^{\ast\ddagger}$
   - Ralph Schlapbach$^{\ast}$
-  - Christian Panse$^{\ast}$
+  - Christian Panse$^{\ast\ddagger}$
 output: 
@@
 
+\newpage
 # References {-}
@@
 
-(ref:tabCompletion) The screenshot displays the command-line completion (tab completion) of RStudio on the `prolfqua::LFQData` R6 object. In the example, it shows the getter methods of the object.
-
-```{r tabCompletion, echo=FALSE, fig.cap="(ref:tabCompletion)", out.width = '66%'}
-knitr::include_graphics("codeSnippet1TabCompletion.png")
-```
-
 ## Session information

### 2022-05-06 inst/MyArticle/paper.Rmd
cosmetics in referencing R packages

@@
 
-The r-package proDA and prolfqua model the missing data directly, while MSstats imputes the data using an accelerated failure model.
+The R-package `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Biocpkg("prolfqua")` model the missing data directly, while MSstats imputes the data using an accelerated failure model.
 Despite imputation, MSstats do not estimate group differences for more proteins and do not achieve a higher pAUC score than prolfqua.

### 2022-05-09 inst/MyArticle/paper.Rmd
Adding DEqMS for further references. Otherwise integrate CP's comments.

--- 2022-05-06 inst/MyArticle/paper.Rmd
@@
 Furthermore it can be tested if differences of protein abundance between groups are significant.
-For this application it is important that the model provides an unbiased estimate of the false discovery rate (FDR) so that protein lists can be selected for follow-up experiments [TODO reference].
+For this application it is important that the model provides an unbiased estimate of the false discovery rate (FDR) so that protein lists can be selected for follow-up experiments.
 Furthermore, the quantified proteins can be ranked using the estimated fold-change, $t$-statistics or scaled $p$-value, and then subjected to gene set enrichment (GSEA) or over-representation analysis [@subramanian2005gene] to determine up or down-regulated groups of proteins.
@@
 knitr::kable(data.frame(bb), caption = "Rows - R package, Columns - types of models supported: pd - parallel design, rm - repeated measurements, fd - factorial design, int - interactions among factors, mem - mixed effect models, eb - empirical Bayes, md - missing data modelling.
-Y - yes, y - yes but only for random effects, * - repeated measurements are only modeled using a fixed effects.
+Y - yes, * - repeated measurements are modeled using a fixed effects.
 ?
-- the hurdle model is not implemented in the msqrob2 package yet.")
+- the hurdle model was published but unclear if available in the msqrob2 package.")
 ```
@@
 We use __R6__ classes to structure the functionality of the package (see Figure \@ref(fig:LFQData) and Figure \@ref(fig:ContrastUML)).
-__R6__ classes are well supported, e.g., by command-line completion features, see Figure \@ref{fig:tabCompletion}, and easy to use (OOP).
+__R6__ classes are well supported, e.g., by command-line completion features, see Figure \@ref(fig:tabCompletion), and easy to use (OOP).
 
@@
 coli proteins and 1 for human proteins.
-We processed the raw data using MaxQuant [cite] version Version 1.6.10.43, with MaxQuant default settings.
-The text files generated by Maxquant are available in the _prolfquadata_ R package [@prolfquadata].
+We processed the raw data using MaxQuant [@Cox2008MaxQuant] version Version 1.6.10.43, with MaxQuant default settings.
+The text files generated by MaxQuant are available in the _prolfquadata_ R package [@prolfquadata].
 
@@
 
+### Robust scaling of the data
+
+[TODO : Describe robust scaling]
+
 ### Estimating contrasts for linear models
 
-Given a linear model contrasts can be computed by 
+To allow for the comparison of different treatments, given a linear model $y = \beta X$, a contrast is computed by a linear combination $l$ of linear model parameters $\beta_m$.
+The difference estimate $\beta_c$, is given by:
+
 \begin{eqnarray} 
@@
 
-and
+and the variance of $\beta_c$ by:
 
@@
 
-with $X$ being the design matrix, $\beta_m$ the model parameters and $l$ an array of coefficients.
+with $X$ being the design matrix, $\beta_m$ the model parameters and $l$ an array of weights, defining the contrast $c$.
 The degrees of freedom for the contrast are equal to the residual degrees of freedom of the linear model.
-
 For estimating contrasts from mixed effects models we used the function `contest` implemented in the R package `r BiocStyle::CRANpkg("lmerTest")` [@Kuznetsova2017lmerTest] and used the Satterthwaite method to estimate the denominator degrees of freedom.
 These methods are implemented in the class `Contrast` (see Figure \@ref(fig:ContrastUML))
+
+### Determining linear parameter combinations for treatment comparison
+
+[TODO : Describe how we produce linear combinations from model specification]
 
@@
 
-A relevant parameter is the number of contracts a given method estimates from the data(see Figure \@ref(fig:FDRfdp) left Panel).
+A relevant parameter is the number of contracts a given method estimates from the data (see Figure \@ref(fig:FDRfdp) Panel B).
 It indicates how robust the procedure works in the presence of missing data.
@@
 
-Figure \@ref(fig:FDRfdp) left panel, shows how various estimates obtained from the models, i.e.
-$\log_2(FC)$, $t$-statistics and scaled $p$-values allow to identify true positives (TP) given a false positive rate.
-When testing for differential expression, the $\log_2(FC)$ is outperformed by than the $t$-statistic or $p$-value.
-If an accurate estimate of the $log_2(FC)$ is essential, the linear models fitted to protein intensities, estimated using Tukey's median polish, show the best performance.
+Figure \@ref(fig:FDRfdp) Panel A, shows how various estimates obtained from the models, i.e.
+the difference between groups, $t$-statistics and scaled $p$-values allow to identify true positives (TP) given a false positive rate.
+When ordering the proteins by the probablity of differential expression the difference is outperformed by than the $t$-statistic or $p$-value.
+If an accurate estimate of the difference is essential, the linear models fitted to protein intensities, estimated using Tukey's median polish, show the best performance.
 We speculate that this is because of the robustness of Tukey's median polish method.
@@
 
-There are only minor differences in the $p$-AUC when using the $t$-statistics or the scaled $p$-value  (see Figure \@ref(fig:FDRfdp) right panel).
+There are only minor differences in the $p$-AUC when using the $t$-statistics or the scaled $p$-value  (see Figure \@ref(fig:FDRfdp) Panel A).
 In order to compute the confusion matrices based on the $p$-value we first need to rescale it (see Equation \@ref(eq:scalepvalue)).
@@
 We also benchmark, if the $FDR$ obtained from a model is an unbiased estimate of the false discovery proportion $FDP$.
-Figure \@ref(fig:FDRfdp) panel C draws on the horizontal axis the FDR determined from the model, and on the vertical axis, the FDP obtained from the confusion matrix.
+Figure \@ref(fig:FDRfdp) Panel C draws on the horizontal axis the FDR determined from the model, and on the vertical axis, the FDP obtained from the confusion matrix.
 We see that most of the lines are below the diagonal, indicating that the FDR estimates obtained with `r BiocStyle::Githubpkg("fgcz/prolfqua")` are modestly conservative for this benchmark dataset provide.
@@
 One of them is the normalization of the intensities within the samples to remove systematic differences and to make them comparable [@pursiheimo2015optimization].
-The method used to infer protein intensities from peptide intensities is an additional important factor [cite TopN].
+The method used to infer protein intensities from peptide intensities is an additional important factor [@Grossmann2010].
 For instance, the original _proDA_ publication uses MaxLFQ [@cox2014accurate] protein estimates.
-However, when using MaxLFQ inensities reported by MaxQuant, _proDA_ performs significantly worse ($pAUC_{10}(\textrm{t-statistics}) = 66%$) compared with results obtained when protein abundances are estimated from peptide abundances using Tukey's median polish ($pAUC_{10}(\textrm{t-statistics}) = 73%$).
+However, when using MaxLFQ intensities reported by MaxQuant, _proDA_ performs significantly worse ($pAUC_{10}(\textrm{t-statistics}) = 66%$) compared with results obtained when protein abundances are estimated from peptide abundances using Tukey's median polish ($pAUC_{10}(\textrm{t-statistics}) = 73%$).
 Last but not least, the software [@Cox2008MaxQuant,@yu2020fast], identifying proteins, and generating the quantification values can also significantly contribute to the performance of the entire pipeline, altering the number of identified proteins and the sensitivity and specificity of the differential expression analysis.
@@
 Moreover, the estimated fold changes are unbiased (not shown) and show high diagnostic accuracy (\@ref(fig:FDRfdp) Panel A).
-Furthermore, the scaled $p$-values or the $t$-statistics are on par with those obtained by the linear model with variance moderation.
-Therefore, we plan to integrate the `r BiocStyle::Biocpkg("proDA")` package as an additional modeling option into `r BiocStyle::Githubpkg("fgcz/prolfqua")` (see Figure \@ref(fig:ContrastUML)).
+Furthermore, the performance of the scaled $p$-values or the $t$-statistics is comparable with that of the linear model with variance moderation.
+Because of the robustness of the dropout model to missing observations, we intend to integrate the `r BiocStyle::Biocpkg("proDA")` package as an additional modeling option into `r BiocStyle::Githubpkg("fgcz/prolfqua")` (see Figure \@ref(fig:ContrastUML)).
 
@@
 
-(ref:proDAAllBenchmarks) Comparison of fold change estimates produced by proDA and using the missing data modelling method implemented in prolfqua.
+(ref:proDAAllBenchmarks) Comparison of fold change estimates produced by proDA and produced by linear models implemented in prolfqua.
 

### 2022-05-09 inst/MyArticle/paper.Rmd
merge from cp. Remove fold changes.

@@
 At the same time, `r BiocStyle::Biocpkg("limma")` allows to specify a linear model formula and implements the empirical Bayes variance shrinkage method.
-The R-package `r BiocStyle::Biocpkg("PECA")`, performs a roll-up of peptide level fold change and peptide level $p$-value estimates, obtained from `r BiocStyle::Biocpkg("limma")`, to protein level estimates.
+The R-package `r BiocStyle::Biocpkg("PECA")`, performs a roll-up of peptide level differences and peptide level $p$-value estimates, obtained from `r BiocStyle::Biocpkg("limma")`, to protein level estimates.
 Furthermore, `r BiocStyle::Biocpkg("msqrob2")` implements two models: robust linear models fitted to protein intensities and robust ridge regression fitted to peptide intensities and combines them with empirical Bayes variance shrinkage known from `r BiocStyle::Biocpkg("limma")`.
@@
 To address this, for the group without observations, we use a point prior equal to the average of $3\%$ smallest group intensities means in the dataset and infer the variance using the pooled the variance of other groups.
-Proteins for which the fold change was imputed in such a way are explicitly annotated in the result table.
+Proteins for which the difference was imputed in such a way are explicitly annotated in the result table.
 Other means of modelling missing information in proteomics data using the Hurdle model is discussed by [@goeminne2020msqrob], while the R package `r BiocStyle::Biocpkg("proDA")` models missingness using probabilistic dropout models [@bioxrvproDA2020].
@@
 We annotated these dilutions from smallest to largest with the letters a, b, c, d, e.
-As a result, we obtain different fold changes by comparing the various dilutions, e.g., when comparing dilution _e_ (9%) against dilution _d_ (7.5%), the expected fold change is 1.2 for E.
+As a result, we obtain various effect sizes by comparing the various dilutions, e.g., when comparing dilution _e_ (9%) against dilution _d_ (7.5%), the expected difference is 1.2 for E.
 coli proteins and 1 for human proteins.
@@
 The probabilistic dropout analysis implemented in the `r BiocStyle::Biocpkg("proDA")` produces inferences comparable to those of other methods (Figure \@ref(fig:FDRfdp) Panel A) for the largest number of proteins (Figure \@ref(fig:FDRfdp) Panel B).
-Moreover, the estimated fold changes are unbiased (not shown) and show high diagnostic accuracy (\@ref(fig:FDRfdp) Panel A).
+Moreover, the estimated differences are unbiased (not shown) and show high diagnostic accuracy (\@ref(fig:FDRfdp) Panel A).
 Furthermore, the performance of the scaled $p$-values or the $t$-statistics is comparable with that of the linear model with variance moderation.
@@
 If you use fixed-effect linear models, apply variance moderation to improve the $t$-statistics and $p$-value estimates.
-If you want to sort your protein lists to perform gene set enrichment analysis, use the $t$-statistic instead of the fold change.
+If you want to sort your protein lists to perform gene set enrichment analysis, use the $t$-statistic instead of the difference.
 Last but not least, do not impute missing data but model it instead.
@@
 ESI-MS &Electro-Spray-Ionization Mass Spectrometry         \\
-FC     &Fold Change                                        \\
 $FDP$  &False Discovery Proportion                         \\
@@
 Here we compare the benchmark results for the `r BiocStyle::Biocpkg("proDA")` model and the linear model with variance moderation.
-We see that the fold change estimates obtained with the `r BiocStyle::Biocpkg("proDA")` package show better diagnostic accuracy.
+We see that the difference estimates obtained with the `r BiocStyle::Biocpkg("proDA")` package show better diagnostic accuracy.
 However, the $t-statistics$ or $p-values$ show slightly worse diagnostic accuracy (Figure \@ref(fig:proDABenchmarks).
-Figure \@ref(fig:proDAAllBenchmarks), compares the fold change estimates obtained using both methods.
+Figure \@ref(fig:proDAAllBenchmarks), compares the difference estimates obtained using both methods.
 
@@
 
-(ref:proDAAllBenchmarks) Comparison of fold change estimates produced by proDA and produced by linear models implemented in prolfqua.
+(ref:proDAAllBenchmarks) Comparison of difference estimates produced by proDA and produced by linear models implemented in prolfqua.
 

### 2022-05-19 inst/MyArticle/paper.Rmd
incorporating suggestions from PN and CP.

@@
 Hence, measuring changes in protein abundance is the subject of active research [@vidova2017review].
-However, identifying and quantifying intact proteins in complex samples containing hundreds of proteins is not possible using existing methods at the moment.
-Hence, the bottom-up mass spectrometric method is employed [@Bubis2017, @da2020philosopher].
-Proteins that are extracted from biological samples are specifically and reproducible digested into protein fragments - peptides.
+Bottom-up mass spectrometric methods, where proteins are specifically and reproducible digested into protein fragments - peptides, are employed to identify and quantify proteins in complex samples containing hundreds to thousand of proteins, [@Bubis2017, @da2020philosopher].
 The peptides are first separated by their chemical and physical properties using High-Pressure Liquid Chromatography (HPLC).
@@
 
-Protein quantification proteomics experiments enable monitoring relative abundances of thousands of proteins in biological samples.
-For most studies, parallel group designs are used, e.g., treatment versus control are used [@de2022apoe2, @laubscher2021baf].
-More recently, more complex experimental designs with increasing numbers of samples are studied, i.e., factorial designs and longitudinal studies (time series), sometimes with repeated measurements on the same subject [@tan2022proteomic, @meier2021reduced].
-The obtained data can be modeled using linear fixed-, mixed-, or random-effects models [@Bates2015FittingJSS].
-Based on these models, tests can be applied to examine whether a specific factors and their interactions are significant.
-Furthermore it can be tested if differences of protein abundance between groups are significant.
-For this application it is important that the model provides an unbiased estimate of the false discovery rate (FDR) so that protein lists can be selected for follow-up experiments.
-Furthermore, the quantified proteins can be ranked using the estimated fold-change, $t$-statistics or scaled $p$-value, and then subjected to gene set enrichment (GSEA) or over-representation analysis [@subramanian2005gene] to determine up or down-regulated groups of proteins.
+Proteomics quantification experiments enable monitoring relative abundances of thousands of proteins in biological samples.
+Most studies use parallel-group designs, where one or many treatment groups are compared to the control group [@de2022apoe2, @laubscher2021baf].
+More recently, more complex experimental designs with an increasing number of samples are studied, e.g., factorial designs and longitudinal studies (time series), sometimes with repeated measurements on the same subject [@tan2022proteomic, @meier2021reduced].
+The data can be modeled using linear fixed-, mixed-, or random-effects models [@Bates2015FittingJSS].
+Based on these models, tests can be applied to examine whether specific factors and factor interactions are significant, e.g., it can be tested if differences in protein abundance between groups are statistically significant.
+
+An important aspect of mass spectrometric data are missing peptide and protein quantifications.
+[@rubin1976inference] classified missing data problems into three categories: missing completely at random (MCAR), missing at random (MAR), and missing not at random (MNAR).
+For instance, in data-dependent acquisition (DDA) MS, only a limited number of MS1 signals are selected for fragmentation and identified.
+Peptide quantification algorithms transfer identification information between MS1 features in different samples by aligning the data using retention time and mass information, thus reducing the amount of missing data.
+Nevertheless, highly abundant proteins can suppress the detection of other proteins in a sample, an MAR mechanism.
+Furthermore, a weak correlation between the number of missing measurements in a condition and the intensities of the observed features can be observed, which is caused by the limit of detection (LOD), an MNAR mechanism [@mcgurk2020use].
 
 Several packages exist to model mass spectrometry protein quantification experiments, e.g., `r BiocStyle::Biocpkg("limma")` [@Ritchie2015], `r BiocStyle::Biocpkg("MSstats")` [@Choi2014], `r BiocStyle::Biocpkg("PECA")` [@Suomi2017bEnhanced],  `r BiocStyle::Biocpkg("msqrob2")` [@Goeminne2016] or `r BiocStyle::Biocpkg("proDA")` [@bioxrvproDA2020], to mention some, all implemented in the R.
-Each of them has some unique features, for example, `r BiocStyle::Biocpkg("MSstats")` infers the model and generates the model formula from the sample annotation, allowing users with limited statistical knowledge to perform differential expression analisis (DEA).
-At the same time, `r BiocStyle::Biocpkg("limma")` allows to specify a linear model formula and implements the empirical Bayes variance shrinkage method.
-The R-package `r BiocStyle::Biocpkg("PECA")`, performs a roll-up of peptide level differences and peptide level $p$-value estimates, obtained from `r BiocStyle::Biocpkg("limma")`, to protein level estimates.
-Furthermore, `r BiocStyle::Biocpkg("msqrob2")` implements two models: robust linear models fitted to protein intensities and robust ridge regression fitted to peptide intensities and combines them with empirical Bayes variance shrinkage known from `r BiocStyle::Biocpkg("limma")`.
+Each of them has some unique features, for example, `r BiocStyle::Biocpkg("MSstats")` infers the model and generates the model formula from the sample annotation, allowing users with limited statistical knowledge to perform differential expression analysis (DEA).
+At the same time, `r BiocStyle::Biocpkg("limma")` allows to specify a design matrix using a linear model formula and implements the empirical Bayes variance shrinkage method.
+The R-package `r BiocStyle::Biocpkg("PECA")`, performs a roll-up of peptide level differences and peptide level $p$-value estimates, obtained from `r BiocStyle::Biocpkg("limma")` or `r BiocStyle::Biocpkg("PECA")`, to protein level estimates.
+Furthermore, `r BiocStyle::Biocpkg("msqrob2")` implements two models: robust linear models fitted to protein intensities and robust ridge regression fitted to peptide intensities and combines them with empirical Bayes variance shrinkage.
 The `r BiocStyle::Biocpkg("proDA")` package implements a linear probabilistic dropout model to account for missing data without imputation.
+`r BiocStyle::Biocpkg("MSstats")` handles missing data by feature filtering and using imputation.
+Other means of modelling missing observations are the Hurdle model discussed by [@goeminne2020msqrob], while the R package `r BiocStyle::Biocpkg("proDA")` models missingness using probabilistic dropout models [@bioxrvproDA2020].
 
-When analyzing parallel-group designs using a single explanatory variable, and contrasting the groups, we can use all R packages; however, we can use only some of them if we want to model factorial designs or repeated measurements.
+When analyzing parallel-group designs using a single explanatory variable, and contrasting groups, we can use all R packages; however, we can use only some of them if we want to model factorial designs or repeated measurements.
 Table \@ref(tab:tableOverview) gives an overview of the models and features supported by these packages.
-We see that packages such as $limma$ and $proDA$ allow us to fit a comprehensive variety of models and test various hypotheses; however, in-depth knowledge of design matrix specification is required [@law2020guide].
-
-Models in `r BiocStyle::Githubpkg("fgcz/prolfqua")` are specified using R’s linear model and mixed model formula interface.
-Therefore, knowledge of the R regression model infrastructure [@faraway2016extending,@MASS2002] is an advantage when using our package.
-Furthermore, this glass box approach should make it easy to reimplement an analysis which was performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` using base R or other programming languages by reading the analysis script.
-However, acknowledging the R formula interface’s complexity to non-statisticians and the popularity of `r BiocStyle::Biocpkg("MSstats")`, we provide the functionality to suggest a model formula from the sample annotation provided in tabular form similarily to `r BiocStyle::Biocpkg("MSstats")` [@Choi2014].
+We see that packages such as $limma$ and $proDA$ allow us to fit a comprehensive variety of models and test various hypotheses; however, in-depth knowledge of design matrix specification using the R formula interface is required [@law2020guide].
 
@@
 xx <- list(
-  triqler = c("pd" = "Y"),
   MSstats = c("pd" = "Y", "rm" = "Y", "mem" = "y"),
@@
 ?
-- the hurdle model was published but unclear if available in the msqrob2 package.")
+- the hurdle model was published but is not available in the msqrob2 package.")
 ```
 
-We were inspired by the R package `r BiocStyle::CRANpkg("caret")` [@caret2008], which enables to call various machine learning (ML) methods, making selecting the best ML algorithm for your problem easy.
+When developing the R package `r BiocStyle::Githubpkg("fgcz/prolfqua")` we were inspired by the R package `r BiocStyle::CRANpkg("caret")` [@caret2008], which enables to call various machine learning (ML) methods, making selecting the best ML algorithm for your problem easy.
 We aimed for a similar R package for the analysis of LFQ data.
 However, after examining the R packages for modeling proteomics quantification data, we observed that model specification, input, and output formats differ widely and wildly.
-However, many of these packages fit linear model either to peptide or protein intensities, determine contrasts among conditions, and afterward, apply empirical Bayes variance shrinkage.
-Therefore, the revised goal was to provide a modular object oriented (OO) design, with R linear models as a core, where features such as $p$-value aggregation, e.g., `r BiocStyle::Biocpkg("PECA")` or variance shrinkage could be easily enabled (see Figure \@ref(fig:ContrastUML)).
-
-An important aspect of mass spectrometric data are missing values.
-Missing values lead to groups of varying size.
-Linear and mixed effect models can handle unbalanced designs, i.e.
-different group sizes.
-As long as there is at least one observation per group they do produce unbiased estimates.
-We use this feature instead of imputing missing observations.
-However, if there are no observation in one of the groups, then contrasts involving this group can not be estimated.
-To address this, for the group without observations, we use a point prior equal to the average of $3\%$ smallest group intensities means in the dataset and infer the variance using the pooled the variance of other groups.
-Proteins for which the difference was imputed in such a way are explicitly annotated in the result table.
-Other means of modelling missing information in proteomics data using the Hurdle model is discussed by [@goeminne2020msqrob], while the R package `r BiocStyle::Biocpkg("proDA")` models missingness using probabilistic dropout models [@bioxrvproDA2020].
-We show using a benchmark dataset that this, although trivial approach, can outperform methods based on data imputation or dropout models.
+However, what these packages have in common is that they fit linear model either to peptide or protein intensities, determine contrasts among conditions, and afterward, apply empirical Bayes variance shrinkage.
+Therefore, the revised goal was to provide a modular object oriented (OO) design, with R linear models as a core, where features such as $p$-value aggregation, e.g., `r BiocStyle::Biocpkg("PECA")`, variance shrinkage, or modelling of missing observations can be added.
 
@@
 We also implemented functionality to streamline the benchmarking of the implemented methods.
-We focus on the comparison of the various modeling approaches.
-We use the Ionstar [@shen2018ionstar] dataset to benchmark the analysis pipeline and modeling methods implemented within `r BiocStyle::Githubpkg("fgcz/prolfqua")` and to compare our results with those of `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`.
+We use the Ionstar [@shen2018ionstar] dataset to benchmark the modeling methods implemented within `r BiocStyle::Githubpkg("fgcz/prolfqua")` and to compare our results with those of `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`.
+Since group sizes are relatively small, typically with four or five subjects per group, a high power of the tests is a relevant criterion to asses the modelling method.
+Furthermore, the quantified proteins can be ranked using the estimated fold-change, $t$-statistics, or scaled $p$-value, and then subjected to gene set enrichment (GSEA) or over-representation analysis [@subramanian2005gene] to determine up or down-regulated groups of proteins.
+However, the model must also provide an unbiased estimate of the false discovery rate (FDR) to manage expectations when selecting protein lists for follow-up experiments.
 
@@
 For popular software like _MaxQuant_ or _MSFragger_, which stores the same variable, e.g., intensity, in multiple columns, one for each sample, we implemented methods that transform the data into tidy tables.
-Relying on the tidy data table enable us to easily interface with many data manipulation, visualization, and modeling methods implemented in base __R__[@RCoreTeam2021] and the tidyverse[@tidyverse2019].
+Relying on the tidy data table enable us to easily interface with many data manipulation, visualization, and modeling methods implemented in base R[@RCoreTeam2021] and the tidyverse[@tidyverse2019].
 We use __R6__ classes to structure the functionality of the package (see Figure \@ref(fig:LFQData) and Figure \@ref(fig:ContrastUML)).
@@
 
-__R__ linear model and linear mixed effect models allow modeling parallel designs, repeated measurements, factorial designs, and many more.
-__R__'s formula interface for linear models is flexible, widely used, and well documented [@faraway2016extending].
-We use the formula interface to specify the models which should make it easy to reproduce an analysis performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` in other statistical programming languages.
 In addition we implemented features specific to high throughput experiments, such as the experimental Bayes variance and $p$-value moderation, which utilizes the parallel structure of the protein measurements and the analysis [@Ritchie2015].
@@
 We used R6 classes encapsulate the statistical modelling functionality in `r BiocStyle::Biocpkg("prolfqua")` (see Figure \@ref(fig:ContrastUML)).
-We did specify a contrast interface (ContrastInterface), and there are various implementations allowing to perform differential expression analysis given linear or mixed effect models (Contrast), variance shrinkage (ContrastModerated), or impute contrasts in case of observations missing for an entire group of samples (ContrastSimpleImpute) can also be used to encapsulate and integrate differential expression analysis results of external tools such as SAINTExpress or proDA.
+We did specify a contrast interface (ContrastInterface), and there are various implementations allowing to perform differential expression analysis given linear or mixed effect models (Contrast), variance shrinkage (ContrastModerated), or impute contrasts in case of observations missing for an entire group of samples (ContrastSimpleImpute) can also be used to encapsulate and integrate differential expression analysis results of external tools such as proDA or to analyse data from protein interaction studies using SAINTexpress [@teo2014saintexpress].
 
@@
 
+```{r prolfquaModels}
+xM <- data.frame( 
+    lm = c("lm", "linear model for peptides or proteins", "strategy_lm, Contrasts"),
+    lmer = c("lmer" , "mixed effect models for peptides or proteins", "strategy_lmer, Contrasts"),
+    gmi = c("group mean imputation", "impute missing group means", "ContrastsSimpleImpute"),
+    ROPECA = c("ROPECA"," ", "ContrastsROPECA"),
+    limma = c("limma" , "empirical Bayes variance shrinkage", "ContrastsModerated"),
+    pic = c("SAINTexpress", "protein interaction scoring", "runSaint, ContrastsSaintExpress"),
+    prodA = c("proDA", "dropout models", "strategy_proDA*, Contrasts_proDA*"), check.names =  FALSE)
+xM <- data.frame(t(xM))
+
+colnames(xM) <- c("id", "model", "prolfqua functions")
+knitr::kable(tibble::as_tibble(xM[,1:3]), caption = "Models available in prolfqua. * in development",
+             format = "latex", booktabs=TRUE) |>
+    kableExtra::kable_styling(latex_options="scale_down")
+
+```
+
 ## Dataset
@@
 We annotated these dilutions from smallest to largest with the letters a, b, c, d, e.
-As a result, we obtain various effect sizes by comparing the various dilutions, e.g., when comparing dilution _e_ (9%) against dilution _d_ (7.5%), the expected difference is 1.2 for E.
-coli proteins and 1 for human proteins.
+By comparing the various dilutions, we obtain various effect sizes , e.g., when comparing dilution _e_ (9%) against dilution _d_ (7.5%), the expected difference is $1.2$ for E.
+coli proteins and $1$ for human proteins.
+There are four technical replicates for each dilution, and hence 20 raw files in total.
 We processed the raw data using MaxQuant [@Cox2008MaxQuant] version Version 1.6.10.43, with MaxQuant default settings.
@@
 
-## Modelling
+## Data preprocessing for model comparison
 
-### Robust scaling of the data
+The peptide abundances (from the _MaxQuant_ _peptide.txt_ file), were $log_2$ transformed, and subsequent robust z-score transformation, where median and the mean average deviation (mad) were obtained from the human proteins only.
+We removed _one hit wonders_, i.e., proteins with a single peptide assignment.
+We also did not use imputation.
+Protein abundances are inferred from the peptide intensities using Tukey's median polish.
+We fitted the fixed effect linear models using protein abundances, the mixed effect linear model, and the `r BiocStyle::Biocpkg("PECA")` model we fitted to peptide intensities.
 
-[TODO : Describe robust scaling]
-
-### Estimating contrasts for linear models
-
-To allow for the comparison of different treatments, given a linear model $y = \beta X$, a contrast is computed by a linear combination $l$ of linear model parameters $\beta_m$.
-The difference estimate $\beta_c$, is given by:
-
-\begin{eqnarray} 
-\hat{\beta_{c}} &=& \sum l\beta_{m} 
-\end{eqnarray}
-
-and the variance of $\beta_c$ by:
-
-\begin{eqnarray} 
-\textrm{var} \hat\beta_c &=& l\sigma^2 (X^T X)^{-1} l^T
-\end{eqnarray}
-
-with $X$ being the design matrix, $\beta_m$ the model parameters and $l$ an array of weights, defining the contrast $c$.
-The degrees of freedom for the contrast are equal to the residual degrees of freedom of the linear model.
-For estimating contrasts from mixed effects models we used the function `contest` implemented in the R package `r BiocStyle::CRANpkg("lmerTest")` [@Kuznetsova2017lmerTest] and used the Satterthwaite method to estimate the denominator degrees of freedom.
-These methods are implemented in the class `Contrast` (see Figure \@ref(fig:ContrastUML))
-
-### Determining linear parameter combinations for treatment comparison
-
-[TODO : Describe how we produce linear combinations from model specification]
-
-### Contrast estimation in presence of missing data.
-
-If there is no observation in a group, we assume that the protein was unobserved because the protein abundance is below the limit of detection (LOD) of the MS instrument.
-We estimate the detection limit as follows: if a protein is quantified just in a single sample of a treatment group, then it is at the instrument's detection limit.
-We compute the mean of those abundances and use it as the group mean if a protein is absent in a treatment group.
-
-When computing differences among two groups $a$ and $b$, we will use the group mean $\bar{a}$ estimated from the data or the detection limit $a_d$ if no observations are present in the group.
-If for instance there are no observations in group $b$, then, $\Delta_{a,b} = \bar{a} - b_{dl}$ if $\bar{a} > b_{dl}$ or $\Delta_{a,b} = 0$ if $\bar{a} < b_{dl}$.
-
-Furthermore, we assume that the variance of the protein is constant in all the conditions. To estimate the variance in the condition, we use the pooled variance based on all conditions:
-
-\begin{equation} 
-s_p^2=\frac{\sum_{i=1}^k (n_i - 1)s_i^2}{\sum_{i=1}^k (n_i - 1)}
-\end{equation}
-
-with $n_i$ the number of observations, and $s_i$ the standard deviation in each condition.
-The standard deviation for the $t$-statistics is then given by:
-
-\begin{equation} 
-s = \sqrt{\frac{2n_g s_p^2}{n}},
-\end{equation}
-
-Where, $n_g$ is the number of groups and $n$ is the number of observations.
-
-This methods are implemented in the class `ContrastSimpleImpute` (see Figure \@ref(fig:ContrastUML)).
-
-### $p$-value moderation
-
-From the linear and the mixed effect models, we can obtain the residual standard deviation $\sigma$, and degrees of freedom $df$.
-[@Smyth2004linear] discuss how, using the empirical Bayes paradigm, to use the $\sigma$ and $df$ of all models to estimate a prior $\sigma$ and prior $df$, and posterior $\tilde \sigma$.
-These can be used to moderate the $t$-statistics by:
-
-\begin{equation} 
-\tilde{t}_{pj} = \frac{t_{pj} s_p}{\tilde{s}_p}
-\end{equation}
-
-and subsequently the $p$-values. We implemented this method in the class `ContrastModerated` (Figure \@ref(fig:ContrastUML)).
-
-### Summarizing peptide level models
-
-To summarize peptide level models to protein models, we applied the method suggested by [@Suomi2017bEnhanced]  to use the median scaled $p$-value of the peptide models and cumulative distribution function of the beta distribution function to determine a regulation probability of the protein.
-
-To obtain the median $p$-value of a protein, we first rescaled the peptide $p$-values by taking the direction of the fold-change $\hat \beta$ into account, i.e.:
-
-\begin{equation}
-p_{s} =
-  \begin{cases}
-1-p, & \textrm{if}~ \hat{\beta} > 0\\
-p-1, & \textrm{otherwise}
-\end{cases}
-(\#eq:scalepvalue)
-\end{equation}
-
-Afterwards, the median scaled $p$-value $\tilde{p}_s$ is determined and using the transformation below, transformed back onto the original scale.
-
-\begin{equation}
-\tilde{p} = 1 - |\tilde{p}_{s}|
-\end{equation}
-
-Because we used the median as the i-th order statistic $i = \frac{n}{2} + 0.5$.
-Therefore, $\gamma = i = \frac{n}{2} + 0.5$ and $\delta = n - i + 1 = n - (\frac{n}{2} + 0.5) + 1 = \frac{n}{2} + 0.5 = \gamma$ are used to parameterize the cumulative distribution function (CDF) of the Beta distribution.
-We implemented this method in the class `ContrastROPECA` (Figure \@ref(fig:ContrastUML)).
-
-```{r prolfquaModels}
-xM <- data.frame( 
-    lm = c("lm", "linear model for peptides or proteins", "strategy_lm, Contrasts"),
-    lmer = c("lmer" , "mixed effect models for peptides or proteins", "strategy_lmer, Contrasts"),
-    gmi = c("group mean imputation", "impute missing group means", "ContrastsSimpleImpute"),
-    ROPECA = c("ROPECA"," ", "ContrastsROPECA"),
-    limma = c("limma" , "empirical Bayes variance shrinkage", "ContrastsModerated"),
-    pic = c("SAINTexpress", "protein interaction scoring", "runSaint, ContrastsSaintExpress"),
-    prodA = c("proDA", "dropout models", "strategy_proDA, Contrasts_proDa"), check.names =  FALSE)
-xM <- data.frame(t(xM))
-
-colnames(xM) <- c("id", "model", "prolfqua functions")
-knitr::kable(tibble::as_tibble(xM[,1:3]), caption = "Models available in prolfqua.",
-             format = "latex", booktabs=TRUE) |>
-    kableExtra::kable_styling(latex_options="scale_down")
-
-```
-
-## Implementing an analysis workflow
-
-The code snippets in this section demonstrate the basic usage of the `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package.
-As input data we use subset of the Ionstar dataset containing $163$ proteins and $1258$ peptides.
-Peptide intensities are $\log_2$ transformed and robust z-score scaled but then rescaled to preserves the original range of the data using the method `robscale`.
-We show the distribution of the normalized intensities in Figure \@ref(fig:prepro) Panel A.
-Protein intensities are estimated using Tukey's median polish.
-Figure \@ref(fig:prepro) Panel B shows the peptide intensities and the estimated protein intensities.
-The boxplot (Panel D) shows the distribution of the protein intensity estimate in each condition.
-Finally, Panel C shows the distribution of the protein standard deviations for each condition.
-
-```{r prepro,  echo = TRUE , fig.cap="(ref:scaling)", out.width = '90%', fig.width = 8, fig.height = 8}
-## load peptide level data
-d <- prolfqua::prolfqua_data('data_ionstar')$filtered()
-
-## create R6 obejct
-lfqd <- prolfqua::LFQData$new(d$data, d$config)
-
-##  transform intensities
-t <- lfqd$get_Transformer()
-
-lfqd <-  t$log2()$robscale()$lfq
-
-## infer protein intensities from peptide intensity
-agr <- lfqd$get_Aggregator()
-lfqp <- agr$medpolish()
-
-## plot panels A-D
-pl <- lfqd$get_Plotter()
-panelA <- pl$intensity_distribution_density() +
-  ggplot2::labs(tag = "A") + ggplot2::theme(legend.position = "none")
-panelB <- agr$plot()$plots[[1]] + ggplot2::labs(tag = "B")
-panelC <- lfqp$get_Stats()$violin() + ggplot2::labs(tag = "C")
-pl <- lfqp$get_Plotter()
-panelD <- pl$boxplots()$boxplot[[1]] + ggplot2::labs(tag = "D")
-ggpubr::ggarrange(panelA, panelB, panelC, panelD)
-```
-
-(ref:scaling) Panel A - Peptide intensity distributions for 20 samples.
-For each sample a line with a different colour is shown.
-Panel B - Peptide intensities for protein _5NTD_ are shown using a line of different colour, and the protein intensity estimate is shown using a fat black line, Panel C - distribution of standard deviations of all proteins in each dilution group (a, b, c, d, e) and overall (all), Panel D - Distribution of protein intensities for protein _5NTD_.
-
-The following code example illustrates how we compute contrast comparing the protein intensities in the condition.
-First, the differences and the linear model are specified.
-Afterward, the model is fitted to the data using the `build_model` function.
-Next, we examine the importance of the factors in the model using the ANOVA function, which facilitates model selection for models with one explanatory variable.
-Finally, we estimate the contrasts either from the linear model using the `Contrasts` class or directly from the data using the `ContrastsSimpleImpute` class.
-Afterward, we apply $t$-statistic moderation using the `ContrastModerated` class.
-Finally, the `addContrastResults` function merges both sets of contrast estimates, preferring the one obtained from the linear model.
-Figure \@ref(fig:exampleContrasts) Panel A shows the distribution of the $p$-values, while figure Panel B shows the volcano-plot for each comparison.
-
-(ref:exampleContrasts) Panel A - Histogram showing the distribution of p-values for $163$ proteins.
-Panel B -  Volcano plot showing $-\log_{10}$ transformed FDR as function of the difference between groups for $163$ proteins.
-
-```{r exampleContrasts, fig.cap="(ref:exampleContrasts)", echo=TRUE, out.width = '90%', fig.width = 8, fig.height = 5}
-# specify differences among conditions
-contrasts <- c(
-  "dilution_(9/7.5)_1.2" =   "dilution.e - dilution.d",
-  "dilution_(7.5/6)_1.25" =   "dilution.d - dilution.c"
-)
-## fit model
-lmmodel <- paste(lfqp$intensity_column()," ~ dilution.")
-modelFunction <- prolfqua::strategy_lm( lmmodel, model_name = "lm")
-models <- prolfqua::build_model(lfqp, modelFunction)
-p0 <- models$anova_histogram()
-
-## compute contrasts from linear model and with imputation
-contr <- prolfqua::Contrasts$new(models, contrasts) |> 
-  prolfqua::ContrastsModerated$new()
-
-conI <- prolfqua::ContrastsSimpleImpute$new( lfqp, contrasts) |> 
-  prolfqua::ContrastsModerated$new()
-
-contrasts <- prolfqua::addContrastResults(prefer = contr, add = conI)
-
-## visualize results in panels
-pl <- contrasts$merged$get_Plotter() 
-panelA <- pl$histogram()$p.value + ggplot2::labs(tag = "A")                                                        
-panelB <- pl$volcano()$FDR +
-   ggplot2::theme(legend.position = "bottom") +
-   ggplot2::labs(tag = "B")
-
-gridExtra::grid.arrange(panelA, panelB, ncol = 2)
-```
-
-## Benchmarking
-
-The Benchmark functionality of `r BiocStyle::Githubpkg("fgcz/prolfqua")` includes ROC curves and computes partial areas under those curves (pAUC) and other scores, e.g., the false discovery proportion FDP.
-We use those plots to study how well linear, mixed effect models or $p$-value moderation models quantitative mass spectrometric high throughput experiments and compare them with results produced by `r BiocStyle::Biocpkg("MSstats")` [@MSstats2014] and `r BiocStyle::Biocpkg("proDA")` [@bioxrvproDA2020] (see also Figure \@ref(fig:FDRfdp))
-
-The peptide intensities from the _MaxQuant_ _peptide.txt_ file were $log_2$ transformed, and subsequent robust z-score transformation, where median and the mean average deviation (mad) were obtained from the human proteins only.
-We removed _one hit wonders_, i.e.
-proteins with a single peptide assignment.
-No further data filtering was used.
-We also did not use imputation.
-The protein level linear model was fitted to the protein intensities inferred from the peptide intensities using the Tukey's median polish.
-The mixed effect linear model, and the `r BiocStyle::Biocpkg("PECA")` model were fitted to the peptide intensities.
-The details of the `r BiocStyle::Githubpkg("fgcz/prolfqua")` benchmark implementation are available online  [BenchmarkingIonstarData](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkingIonstarData.html).
-
-For benchmarking of `r BiocStyle::Biocpkg("MSstats")` we used the MSstats default parameters (for details see: [Benchmarking_MSstats](https://wolski.github.io/prolfquabenchmark/articles/))
+## Benchmark metrics
 
@@
 By using these measures we can compare the performances of the statistics produced by the various methods examined.
+
+In order to compute the confusion matrices based on the $p$-value we first need to rescale it (see Equation \@ref(eq:scalepvalue)).
+Thus, the $p$-value is a function of the $t$-statistics and the degrees of freedom.
+
+## Modelling
+
+### Robust scaling of the data
+
+Valikangas [@valikangas2016systematic] and colleagues discuss and benchmark various methods of peptide or protein intensity normalization such as variance stabilizing normalization [@huber2002variance] or quantile normalization [@bolstad2003comparison].
+In this work we will be using a robust version of the z-score, where instead of the mean $\bar{x}$ we use the median and instead of the standard deviation $\tilde{S}$ the median absolute deviation (mad):
+
+$$
+z = \frac{x - \bar{x}}{S} \sim \frac{x - \tilde{x}}{\tilde{S}}
+$$
+
+Because we need to estimate the protein fold-changes on the original scale, we have to multiply the $z$-score by the average variance of all the $N$ samples in the experiment.
+
+$$
+z' = z \times 1/N\sum_{i=1}^N S_i
+$$
+
+To apply this transformation, we need to estimate two parameters per sample.
+It works for experiments with thousands of proteins and experiments where only a few hundred proteins per sample are measured.
+For the Ionstar dataset, we used the intensities of _H.
+sapiens_ proteins, whose concentrations do not change, to determine $\bar{x}$ and $S$ and then applied it to all the intensities (including _E.
+coli_) in the sample.
+
+### Estimating differences between groups
+
+Given a linear model $y = \beta X$, we can compute the difference between two groups $\beta_{c}$ by a linear combination $c$ of linear model parameters $\beta$.
+$c$ is vector with as many elements as there are coefficients in the linear model.
+If $c$ has $0$ for one or more of its rows, then the corresponding coefficient in $\beta$ is not involved in determining the contrast [@ph525xseries].
+
+The difference estimate $\beta_c$, is given by:
+
+\begin{eqnarray} 
+\hat{\beta_{c}} &=& \sum l\beta 
+\end{eqnarray}
+
+and the variance of $\beta_c$ by:
+
+\begin{eqnarray} 
+\textrm{var} \hat\beta_c &=& l\sigma^2 (X^T X)^{-1} l^T
+\end{eqnarray}
+
+with $X$ being the design matrix, $\beta$ the model parameters and $c$ an array of weights, defining the contrast (difference among conditions).
+The degrees of freedom for the contrast are equal to the residual degrees of freedom of the linear model.
+For estimating contrasts from mixed effects models we used the function `contest` implemented in the R package `r BiocStyle::CRANpkg("lmerTest")` [@Kuznetsova2017lmerTest] and used the Satterthwaite method to estimate the denominator degrees of freedom.
+These methods are available in the class `Contrast` (see Figure \@ref(fig:ContrastUML))
+
+### Determining linear parameter combinations for treatment comparison
+
+The package prolfqua provides function to determine the vector of $parameter$ weights $c$, from a contrast specification:
+
+- `<contrast name>` = `<factor><level_1> - <factor><level_2>`
+
+Furthermore, contrasts for subgroups also can be specified using the syntax:
+
+- `<contrast name>` = `<factor_1><subgroup>:<factor_2><level_1>` - `<factor_1><subgroup>:<factor_2><level_2>`
+
+where `<factor_x>` is the name of the explanatory variable, and `<level_x>` are group labels.
+
+The following code shows an example where we specify two contrast: 
+a) where we compare Cells of type CMP/MEP with cells of type HSC, and b) where we contrast therapy NO with therapy NU for the celltype CMP/MEP [@meier2021reduced].
+
+```{r echo=TRUE}
+Contrasts <- 
+    c("CMP/MEPvsHSC" = "`CelltypeCMP/MEP` - `CelltypeHSC`",
+   "NOvsHU" = "`class_therapyc.NO:CelltypeCMP/MEP` - `class_therapyp.HU:CelltypeCMP/MEP`")
+
+m <- prolfqua::prolfqua_data('data_basicModel_p1807')
+linfct <- prolfqua::linfct_from_model(m,as_list = FALSE)
+
+prolfqua::linfct_matrix_contrasts(linfct, Contrasts )
+```
+
+### Contrast estimation in presence of missing data.
+
+Missing observations lead to different group sizes, that is, unbalanced designs.
+Linear and mixed effect models implemented in R can handle unbalanced designs.
+Therefore, as long as one observation per group is available, they will produce unbiased estimates, and no imputation is needed.
+If there is no observation in a group, we assume that the protein is unobserved because the protein abundance is below the limit of detection (LOD) of the MS instrument.
+In this case, we will impute the mean using the protein abundance at the detection limit $A_{LOD}$.
+We estimate the abundance at the detection limit using the abundances of the proteins observed only in one sample of a group of samples.
+Then, we compute the median of the distribution and use it as the group mean if a protein is absent in a treatment group.
+
+When computing differences among two groups $a$ and $b$, we will use the group mean $\bar{a}$ or $\bar{b}$ estimated from the data, or if no observations are present we use $A_{LOD}$.
+If there are no observations in group $b$ then
+
+$$
+\Delta_{a,b} = 
+\begin{cases}
+\bar{a} - A_{LOD} & \text{if} ~~ \bar{a} > A_{LOD}\\
+0  &\text{if} ~~ \bar{a} < A_{LOD}.
+\end{cases}
+$$
+
+To estimate the variance, we assume that the variance of the protein is constant in all the groups and use the pooled variance based on all data:
+
+\begin{equation} 
+s_p^2=\frac{\sum_{i=1}^k (n_i - 1)s_i^2}{\sum_{i=1}^k (n_i - 1)}
+\end{equation}
+
+with $n_i$ the number of observations, and $s_i$ the standard deviation in each condition.
+The standard deviation for the $t$-statistics is then given by:
+
+\begin{equation} 
+s = \sqrt{\frac{2n_g s_p^2}{n}},
+\end{equation}
+
+Where, $n_g$ is the number of groups and $n$ is the number of observations.
+If variance can not be estimated for a protein, we will use the median pooled variance of all the proteins.
+
+This methods are implemented in the class `ContrastSimpleImpute` (see Figure \@ref(fig:ContrastUML)).
+
+### $p$-value moderation
+
+From the linear and the mixed effect models, we can obtain the residual standard deviation $\sigma$, and degrees of freedom $df$.
+[@Smyth2004linear] discuss how, to use the $\sigma$ and $df$ of all models to estimate a prior $\sigma$ and prior $df$, and posterior $\tilde \sigma$.
+These can be used to moderate the $t$-statistics by:
+
+\begin{equation} 
+\tilde{t}_{pj} = \frac{t_{pj} s_p}{\tilde{s}_p}.
+\end{equation}
+
+We implemented this method in the class `ContrastModerated` (Figure \@ref(fig:ContrastUML)).
+
+### Summarizing peptide level differences and p-values on protein level
+
+To summarize peptide level models to protein models, we applied the method suggested by [@Suomi2017bEnhanced]  to use the median scaled $p$-value of the peptide models and cumulative distribution function of the beta distribution function to determine a regulation probability of the protein.
+
+To obtain the median $p$-value of a protein, we first rescaled the peptide $p$-values by taking the sign of the fold-change $\hat \beta$ into account, i.e.:
+
+\begin{equation}
+p_{s} =
+  \begin{cases}
+1-p, & \textrm{if}~ \hat{\beta} > 0\\
+p-1, & \textrm{otherwise}
+\end{cases}
+(\#eq:scalepvalue)
+\end{equation}
+
+Afterwards, the median scaled $p$-value $\tilde{p}_s$ is determined and using the transformation below, transformed back onto the original scale:
+
+\begin{equation}
+\tilde{p} = 1 - |\tilde{p}_{s}|
+\end{equation}
+
+Because we used the median as the i-th order statistic $i = \frac{n}{2} + 0.5$.
+Therefore, $\gamma = i = \frac{n}{2} + 0.5$ and $\delta = n - i + 1 = n - (\frac{n}{2} + 0.5) + 1 = \frac{n}{2} + 0.5 = \gamma$ are used to parameterize the cumulative distribution function (CDF) of the Beta distribution.
+We implemented this method in the class `ContrastROPECA` (Figure \@ref(fig:ContrastUML)).
+
+# Results and Discussion.
+
+## Example analysis workflow
+
+The code snippets in this section demonstrate how a differential expression analysis workflow can be implemented using the `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package.
+As input data, we use a subset of the Ionstar dataset containing $163$ proteins and $1258$ peptides.
+Peptide abundances are $\log_2$ transformed and robust z-score scaled using the method `robscale`.
+Using the LFQDataPlotter class, we show the distribution of the normalized peptide abundances in Figure \@ref(fig:prepro) Panel A.
+Afterwards, protein intensities are estimated from peptide intensities using Tukey's median polish.
+Figure \@ref(fig:prepro) Panel B shows the peptide intensities and the estimated protein intensities.
+Next, we compute the standard deviation of all the proteins in each group and display their distribution using violin plots (Panel C).
+Finally, we create a boxplot (Panel D) showing the abundance of one protein.
+
+```{r prepro,  echo = TRUE , fig.cap="(ref:scaling)", out.width = '90%', fig.width = 8, fig.height = 8}
+## load peptide level data
+d <- prolfqua::prolfqua_data('data_ionstar')$filtered()
+
+## create R6 obejct
+lfqd <- prolfqua::LFQData$new(d$data, d$config)
+
+##  transform intensities
+t <- lfqd$get_Transformer()
+
+lfqd <-  t$log2()$robscale()$lfq
+
+## infer protein intensities from peptide intensity
+agr <- lfqd$get_Aggregator()
+lfqp <- agr$medpolish()
+
+## plot panels A-D
+pl <- lfqd$get_Plotter()
+panelA <- pl$intensity_distribution_density() +
+  ggplot2::labs(tag = "A") + ggplot2::theme(legend.position = "none")
+panelB <- agr$plot()$plots[[1]] + ggplot2::labs(tag = "B")
+panelC <- lfqp$get_Stats()$violin() + ggplot2::labs(tag = "C")
+pl <- lfqp$get_Plotter()
+panelD <- pl$boxplots()$boxplot[[1]] + ggplot2::labs(tag = "D")
+ggpubr::ggarrange(panelA, panelB, panelC, panelD)
+```
+
+(ref:scaling) Panel A - Peptide intensity distributions for $20$ samples.
+For each sample a line with a different colour is shown.
+Panel B - Peptide intensities for protein _5NTD_ are shown using a line of different colour, and the protein intensity estimate is shown using a fat black line, Panel C - distribution of standard deviations of all proteins in each dilution group (a, b, c, d, e) and overall (all), Panel D - Distribution of protein intensities for protein _5NTD_.
+
+The following code example illustrates how we compute contrast by comparing the protein intensities among groups.
+First, the linear model and the differences are specified.
+Afterward, the model is fitted to the data using the `build_model` function.
+Next, we estimate the contrasts either from the linear model using the `Contrasts` class or directly from the data using the `ContrastsSimpleImpute` class.
+Afterward, we apply $t$-statistic moderation using the `ContrastModerated` class.
+Finally, the `addContrastResults` function merges both sets of contrast estimates, preferring the one obtained from the linear model if both are available.
+Then we create the plots shown in Figure \@ref(fig:exampleContrasts).
+Panel A shows the distribution of the $p$-values, while figure Panel B shows the volcano plot for each comparison.
+
+(ref:exampleContrasts) Panel A - Histogram showing the distribution of p-values for $163$ proteins.
+Panel B -  Volcano plot showing $-\log_{10}$ transformed FDR as function of the difference between groups for $163$ proteins.
+
+```{r exampleContrasts, fig.cap="(ref:exampleContrasts)", echo=TRUE, out.width = '90%', fig.width = 8, fig.height = 5}
+# specify differences among conditions
+contrasts <- c(
+  "dilution_(9/7.5)_1.2" =   "dilution.e - dilution.d",
+  "dilution_(7.5/6)_1.25" =   "dilution.d - dilution.c"
+)
+## fit model
+lmmodel <- paste(lfqp$intensity_column()," ~ dilution.")
+modelFunction <- prolfqua::strategy_lm( lmmodel, model_name = "lm")
+models <- prolfqua::build_model(lfqp, modelFunction)
+
+## compute contrasts from linear model and with imputation
+contr <- prolfqua::Contrasts$new(models, contrasts) |> 
+  prolfqua::ContrastsModerated$new()
+
+conI <- prolfqua::ContrastsSimpleImpute$new( lfqp, contrasts) |> 
+  prolfqua::ContrastsModerated$new()
+
+contrasts <- prolfqua::addContrastResults(prefer = contr, add = conI)
+
+## visualize results in panels
+pl <- contrasts$merged$get_Plotter() 
+panelA <- pl$histogram()$p.value + ggplot2::labs(tag = "A")                                                        
+panelB <- pl$volcano()$FDR +
+   ggplot2::theme(legend.position = "bottom") +
+   ggplot2::labs(tag = "B")
+
+gridExtra::grid.arrange(panelA, panelB, ncol = 2)
+```
+
+R linear model and linear mixed effect models allow modeling parallel designs, repeated measurements, factorial designs, and many more. 
+Models in `r BiocStyle::Githubpkg("fgcz/prolfqua")` are specified using R’s linear model and mixed model formula interface.
+Therefore, knowledge of the R regression model infrastructure [@faraway2016extending,@MASS2002] is an advantage when using our package.
+Furthermore, this glass box approach should make it easy to reimplement an analysis which was performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` using base R or other programming languages by reading the analysis script.
+However, acknowledging the R formula interface’s complexity to non-statisticians and the popularity of `r BiocStyle::Biocpkg("MSstats")`, we provide the functionality to suggest a model formula from the sample annotation provided in tabular form similarly to `r BiocStyle::Biocpkg("MSstats")` [@Choi2014].
+
+Using the tidy table to model the data ensures interoperability with other proteomics-related packages that manage their data with tidy-tables, e.g., `r BiocStyle::CRANpkg("protti")` [@quast2022protti].
+Furthermore, the use of `R6` classes, which encapsulate the configuration and the data, allows writing very concise code (no function arguments needed), and auto-completion support in the Rstudio editor makes it easy for novices to explore `r BiocStyle::Githubpkg("fgcz/prolfqua")`s functionality \@ref(fig:tabCompletion).
+To simplify integration of `r BiocStyle::Githubpkg("fgcz/prolfqua")` with Bioconductor based workflows there is an method to convert the LFQData class into a `r BiocStyle::Biocpkg("SummarizedExperiment")`.
+
+To ease the usage barriers of the R package to users not being familiar with statistics and R programming, we integrated some workflows into our Local Laboratory Information Management System (LIMS) [b-fabric](https://fgcz-bfabric.uzh.ch/) @bfabric.
+This integration enables our users to select the input and basic settings in a graphical user interface.
+Then, as an output, the user receives a report including input files, the R markdown file and R scripts necessary to replicate the analysis using their in-house R installation.
+In this way, we also support the reproducible science efforts of the community.
+
+## Benchmarking modelling approaches
+
+The Benchmark functionality of `r BiocStyle::Githubpkg("fgcz/prolfqua")` includes receiver operator curves and computes partial areas under those curves (pAUC) and other scores, e.g., the false discovery proportion FDP.
+We use those scores, i.e.
+the $pAUC$ at $10%$ FDR and the FDP, to examine how well the methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` model quantitative mass spectrometric high throughput data and compare them with results produced by `r BiocStyle::Biocpkg("MSstats")` [@MSstats2014] and `r BiocStyle::Biocpkg("proDA")` [@bioxrvproDA2020] (see also Figure \@ref(fig:FDRfdp))
 
@@
 For each protein, we can estimate four contrast which for 4046 proteins, at least two peptides resulted in $16184$ possible contrast estimates.
-Because the set of contrasts slightly differs for each method, it makes comparing them complicated.
+Because the set of proteins for which differences are estimated differs slightly for each method, it makes comparing them complicated.
 For instance, we observe that `r BiocStyle::Biocpkg("MSstats")` estimates $16058$ while the mixed effect models estimate the fewest with $15940$.
-Hence, to conclude that one method shows a better performance, the score, e.g., pAUC, needs to be greater, but also the number of estimated contrasts needs to be equal or larger.
+Hence, to conclude that one method shows a better performance, the score, e.g., pAUC, needs to be greater, but also the number of proteins with differential expression results needs to be equal or larger.
 
@@
 the difference between groups, $t$-statistics and scaled $p$-values allow to identify true positives (TP) given a false positive rate.
-When ordering the proteins by the probablity of differential expression the difference is outperformed by than the $t$-statistic or $p$-value.
-If an accurate estimate of the difference is essential, the linear models fitted to protein intensities, estimated using Tukey's median polish, show the best performance.
-We speculate that this is because of the robustness of Tukey's median polish method.
+Furthermore, ordering the proteins by the $t$-statistic or $p$-value leads to a higher $pAUC_{10}$ then when ordering by the estimated difference among groups.
+Suppose an accurate estimate of the difference among groups is essential.
+In that case, the linear models fitted to protein intensities, estimated using Tukey's median polish, show a better performance than the methods directly modeling peptide intensities.
+We speculate that outliers at the peptide level do not affect the protein estimates when using Tukey's median polish method.
 This hypothesis could be examined, by including other forms of protein intensity inferences implemented in `prolfqua`, e.g.
@@
 
-There are only minor differences in the $p$-AUC when using the $t$-statistics or the scaled $p$-value  (see Figure \@ref(fig:FDRfdp) Panel A).
-In order to compute the confusion matrices based on the $p$-value we first need to rescale it (see Equation \@ref(eq:scalepvalue)).
-Thus, the $p$-value is a function of the $t$-statistics and the degrees of freedom.
-Interestingly, the pAUC increases when using $p$-values instead of the $t$-statistics for linear models, while it decreases for mixed effect models.
+There are only minor differences in the $pAUC_{10}$ when comparing the $t$-statistics or the scaled $p$-value  (see Figure \@ref(fig:FDRfdp) Panel A).
+Interestingly, the $pAUC$ increases when using $p$-values instead of the $t$-statistics for linear models, while it decreases for mixed effect models.
 The reason is erroneous denominator degrees of freedom estimation for many proteins in the mixed-effect models.
@@
 
-(ref:FDRfdp) Panel A - Partial area under the ROC curve at $10\%$ FPR (pAUC_10) for all contrasts and three different statistics: the difference among conditions, the scaled $p$-value (sign(diff) $\cdot$ p.value) and the $t$-statistics (higher is better).
+(ref:FDRfdp) Panel A - Partial area under the ROC curve at $10\%$ FPR ($pAUC_10$) for all contrasts and three different statistics: the difference among conditions, the scaled $p$-value (sign(diff) $\cdot$ p.value) and the $t$-statistics (higher is better).
 The red line indicates the average area under the curve of all methods.
@@
             vjust = -1, hjust = -0.2, angle = -90) 
-p3 <- nrgg + ggplot2::labs(tag = "B")
+pB <- nrgg + ggplot2::labs(tag = "B")
 
@@
 
-p4 <- ddb |> ggplot2::ggplot(ggplot2::aes(y = FDP_,  x  = scorecol )) + 
+pC <- ddb |> ggplot2::ggplot(ggplot2::aes(y = FDP_,  x  = scorecol )) + 
   ggplot2::geom_line(ggplot2::aes(color = model_name, linetype = model_name)) +
@@
 
-gridExtra::grid.arrange(p1_pAUC, p3, p4, layout_matrix=hlay)
+gridExtra::grid.arrange(p1_pAUC, pB, pC, layout_matrix=hlay)
 ```
 
-# Results and Discussion.
-
 We focused in our benchmark on the comparison of the statistical modelling methods and fixing the pre-processing steps.
-However, there are other equally or even more important parameters of any protein quantification pipeline.
+However, there are other equally or even more important parameters of a protein quantification pipeline [some paper with workflow evaluation].
 One of them is the normalization of the intensities within the samples to remove systematic differences and to make them comparable [@pursiheimo2015optimization].
@@
 We assessed the performance of different modeling approaches implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")`, `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`, using a benchmark dataset whose ground truth is known.
-However, since the for each dilution only technical replicates are available, essential sources of variation present in every experiment, such as the biochemical and biological, are not measured.
+However, since for each dilution only technical replicates are available, essential sources of variation present in every experiment, such as the biochemical and biological, are not measured.
 The dataset captures only the variance from the chromatography, electro-spray, and mass spectrometric measurement method.
@@
 
-Computing the statistics at the peptide level, e.g., the $t$-statistics or $p$-value, then summarizing these statistics using their median produces the highest AUC scores among all the tested models (see Figure \@ref(fig:FDRfdp) Panel A left).
-By using the beta distribution to model the number of peptides observed, we can further improve the pAUC scores (see Figure \@ref(fig:FDRfdp) Panel A center).
+Computing the statistics at the peptide level, e.g., the $t$-statistics or $p$-value, then summarizing these statistics using their median produces the highest $AUC$ scores among all the tested models (see Figure \@ref(fig:FDRfdp) Panel A left).
+By using the beta distribution to model the number of peptides observed, we can further improve the $pAUC$ scores (see Figure \@ref(fig:FDRfdp) Panel A center).
 However, the properties of beta-based probabilities are not well understood; for instance, the p-values are not uniformely distributed under the null hypothesis (not shown).
@@
 The R-package `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Biocpkg("prolfqua")` model the missing data directly, while MSstats imputes the data using an accelerated failure model.
-Despite imputation, MSstats do not estimate group differences for more proteins and do not achieve a higher pAUC score than prolfqua.
-Furthermore, \@ref(fig:FDRfdp) Panel C shows that when using MSstats the proportion of false discoveries might be very high even when filtering for a low FDR.
+Despite imputation, `r BiocStyle::Biocpkg("MSstats")` do not estimate group differences for more proteins and do not achieve a higher $pAUC$ score than `r BiocStyle::Githubpkg("fgcz/prolfqua")`.
+Furthermore, \@ref(fig:FDRfdp) Panel C shows that when using `r BiocStyle::Biocpkg("MSstats")` the proportion of false discoveries might be very high even when filtering for a low FDR.
 
-Using the tidy table to model the data ensures interoperability with other proteomics-related packages that manage their data with tidy-tables, e.g., `r BiocStyle::CRANpkg("protti")`.
-Furthermore, the use of `R6` classes, which encapsulate the configuration and the data, allows writing very concise code (no function arguments needed), and auto-completion support in the Rstudio editor makes it easy for novices to explore `r BiocStyle::Githubpkg("fgcz/prolfqua")`s functionality.
-To simplify integration of prolfqua with Bioconductor based workflows there is an method to convert the LFQData class into a `r BiocStyle::Biocpkg("SummarizedExperiment")`.
-
-To ease the usage barriers of the R package to users not being so familiar with statistics and R programming, we integrated some workflows into our Local Laboratory Information Management System (LIMS) [b-fabric](https://fgcz-bfabric.uzh.ch/) @bfabric.
-This integration enables our users to select the input and basic settings in a graphical user interface.
-Then, as an output, the user receives a report including the R markdown file and all necessary input files to reprocess the analysis using R on their devices.
-In this way, we also support the reproducible science efforts of the community.
+We make the R-markdown files to replicate the benchmarking available at `r BiocStyle::Githubpkg("wolski/prolfquabenchamrk")` and at [BenchmarkingIonstarData](https://wolski.github.io/prolfquabenchmark).
 
@@
 The modular design of `r BiocStyle::Githubpkg("fgcz/prolfqua")`, allows for adding new features, e.g., generalized linear models (_glm_'s) for missing data modeling or robust linear models (_rlm_'s), in the future.
+R's formula interface for linear models is flexible, widely used, and well documented [@faraway2016extending].
+We use the formula interface to specify the models which should make it easy to reproduce an analysis performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` in other statistical programming languages.
 
-When comparing statistical modeling methods for the differential expression analysis, we assessed performance measures such as the number of estimated contrasts, the pAUC, and if the FDR is an unbiased estimate of the FDR.
+When comparing statistical modeling methods for the differential expression analysis, we assessed performance measures such as the number of estimated contrasts, the $pAUC$, and if the FDR is an unbiased estimate of the FDR.
 It is relevant that an analysis pipeline shows good performance in all these categories.
@@
 If you want to sort your protein lists to perform gene set enrichment analysis, use the $t$-statistic instead of the difference.
-Last but not least, do not impute missing data but model it instead.
+Last but not least, do not impute missing data directly but model it when estimating parameters.
 
@@
 We provide more details at the website https://github.com/fgcz/prolfqua/.
-This document was created using Rmarkdown, and all the code need to replicate the document and the presented benchmarks is available at : https://github.com/wolski/prolfquabenchmark/.
+This document was created using Rmarkdown, and all the code need to replicate the document and the presented benchmarks is available at : https://github.com/wolski/prolfquabenchmark.
 
@@
 
-The following code demonstrates how an analysis of MSFragger output can be performed using prolfqua.
-First, a tidy table with protein abundances is generated by reading protein abundances reported by MSFragger in the combined_protein.tsv file.
-The samples are annotated using the data stored in the annotation.xlsx file. Next, an _AnalysisTableAnnotation_ R6 object is created. 
-Since bottom-up proteomics data is hierarchical a protein has peptides, which might be modified and create precursors.
-To describe this hierarchy, AnalysisTableAnnotation has a hierarchy field that stores a list with an entry for each hierarchy level.
-Since combined_portein.tsv only holds protein level data, the hierarchy list has only one element, and we need to specify which column does contain the protein identifiers.
+The following code demonstrates how we use prolfqua to analyze protein intensities reported in the MSFragger `combined_protein.tsv` file.
+First, we create a tidy table containing the protein abundances by reading the `combined_protein.tsv` file using  `tidy_MSFragger_combined_protein.` Then, we read the sample annotation from the file `annotation.xlsx` file.
+Next, we create an `AnalysisTableAnnotation` R6 object.
+
+Bottom-up proteomics data is hierarchical, i.e., a protein has peptides, peptides might be modified, etc.
+Therefore, the `AnalysisTableAnnotation` has a `hierarchy` field storing a list with an entry for each hierarchy level.
+Since combined_portein.tsv only holds protein level data, the hierarchy list has one element, and we use it to specify which column contains the protein identifiers.
 We also need to specify which column contains the protein abundances we want to use for the data analysis.
 Finally, we have to specify which columns contain the explanatory variables of the analysis.
-The AnalysisTableAnnotation has the field factors, a list with as many entries as there are explanatory variables.
-Here we have included two explanatory variable, the dilution, specified in the column 'sample', and 'run' in the column 'run_ID' representing the order of the measurement.
+The `AnalysisTableAnnotation` has the field `factors,` a list with as many entries as explanatory variables.
+Here we include two explanatory variables, the dilution, specified in the column 'sample', and 'run' stored in the column 'run_ID', representing the order of the measurement.
 
-```{r error=TRUE}
+```{r echo=TRUE}
 datadir <- file.path(find.package("prolfquadata") , "quantdata")
@@
 
-protein <- tibble::as_tibble(read.csv(unz(inputFragfile,"IonstarWithMSFragger/combined_protein.tsv"),
+protein <- tibble::as_tibble(
+    read.csv(unz(inputFragfile,"IonstarWithMSFragger/combined_protein.tsv"),
              header = TRUE, sep = "\t", stringsAsFactors = FALSE))
@@
 merged <- dplyr::inner_join(annotation, protein)
-
 atable <- prolfqua::AnalysisTableAnnotation$new()
@@
 # specify column containing protein identifiers
-atable$hierarchy[["protein_Id"]] <- c("protein")
+atable$hierarchy[["protein_Id"]] = "protein"
 
@@
 
-## Comparing benchmark results for proDA and prolfqua
+## Comparing benchmark results for proDA and prolfqua [REMOVE?]
 
@@
 We see that the difference estimates obtained with the `r BiocStyle::Biocpkg("proDA")` package show better diagnostic accuracy.
-However, the $t-statistics$ or $p-values$ show slightly worse diagnostic accuracy (Figure \@ref(fig:proDABenchmarks).
+However, the $t-statistics$ or $p-values$ show slightly worse diagnostic accuracy (Figure \@ref(fig:proDABenchmarks)).
 Figure \@ref(fig:proDAAllBenchmarks), compares the difference estimates obtained using both methods.

### 2022-05-19 inst/MyArticle/paper.Rmd
fix formulas with CP.

@@
 
-$$
-z = \frac{x - \bar{x}}{S} \sim \frac{x - \tilde{x}}{\tilde{S}}
-$$
+\begin{eqnarray}
+z = \frac{x - \tilde{x}}{\tilde{S}}
+\end{eqnarray}
 
-Because we need to estimate the protein fold-changes on the original scale, we have to multiply the $z$-score by the average variance of all the $N$ samples in the experiment.
+Because we need to estimate the protein fold-changes on the original scale, we have to multiply the $z$-score by the average variance of all the $n$ samples in the experiment.
 
-$$
-z' = z \times 1/N\sum_{i=1}^N S_i
-$$
+\begin{eqnarray} 
+z' = z \cdot \frac{1}{n}\sum_{i=1}^n S_i
+\end{eqnarray}
 
@@
 Given a linear model $y = \beta X$, we can compute the difference between two groups $\beta_{c}$ by a linear combination $c$ of linear model parameters $\beta$.
-$c$ is vector with as many elements as there are coefficients in the linear model.
+$c$ is a column vector with as many elements as there are coefficients in the linear model.
 If $c$ has $0$ for one or more of its rows, then the corresponding coefficient in $\beta$ is not involved in determining the contrast [@ph525xseries].
 
-The difference estimate $\beta_c$, is given by:
+The difference estimate $\beta_c$, is given by the dot product:
 
 \begin{eqnarray} 
-\hat{\beta_{c}} &=& \sum l\beta 
+\hat{\beta_{c}} &=& c^T \beta 
 \end{eqnarray}
@@
 \begin{eqnarray} 
-\textrm{var} \hat\beta_c &=& l\sigma^2 (X^T X)^{-1} l^T
+\textrm{var} \hat\beta_c &=& c\sigma^2 (X^T X)^{-1} c^T
 \end{eqnarray}
@@
 
-The package prolfqua provides function to determine the vector of $parameter$ weights $c$, from a contrast specification:
+The package prolfqua provides function to determine the vector of $parameter$ weights $c$, from a contrast specification string:
 
-- `<contrast name>` = `<factor><level_1> - <factor><level_2>`
+```{r eval=FALSE, echo=TRUE}
+c("contrast_name" = "factorlevel_1 - factorlevel_2")
+```
 
@@
 
-- `<contrast name>` = `<factor_1><subgroup>:<factor_2><level_1>` - `<factor_1><subgroup>:<factor_2><level_2>`
+```{r eval=FALSE, echo=TRUE}
+c("contrast_name" = "`factor_1level_a:factor_2level_1` - `factor_1level_a:factor_2level_2`")
 
-where `<factor_x>` is the name of the explanatory variable, and `<level_x>` are group labels.
+```
+where `factor_x` is the name of the explanatory variable, and `level_x`  are group labels.
 
@@
 
-When computing differences among two groups $a$ and $b$, we will use the group mean $\bar{a}$ or $\bar{b}$ estimated from the data, or if no observations are present we use $A_{LOD}$.
-If there are no observations in group $b$ then
+When computing differences $\Delta$ among two groups $a$ and $b$, we will use the group mean $\bar{a}$ or $\bar{b}$ estimated from the data, or if no observations are present in a group we use $A_{LOD}$, e.g., If there are no observations in group $b$ then
 
 $$
-\Delta_{a,b} = 
+\Delta = 
 \begin{cases}

### 2022-05-19 inst/MyArticle/paper.Rmd
delta

@@
 
-When computing differences $\Delta$ among two groups $a$ and $b$, we will use the group mean $\bar{a}$ or $\bar{b}$ estimated from the data, or if no observations are present in a group we use $A_{LOD}$, e.g., If there are no observations in group $b$ then
+When computing differences $\Delta$ among two groups $a$ and $b$, we will use the group mean $\bar{a}$ or $\bar{b}$ estimated from the data, or if no observations are present in a group we use $A_{LOD}$, e.g., If there are no observations in group $b$ then :
 

### 2022-05-19 inst/MyArticle/paper.Rmd
cosmetics in sup

--- 2022-05-19 inst/MyArticle/paper.Rmd
@@
 Next, we create an `AnalysisTableAnnotation` R6 object.
-
 Bottom-up proteomics data is hierarchical, i.e., a protein has peptides, peptides might be modified, etc.
 Therefore, the `AnalysisTableAnnotation` has a `hierarchy` field storing a list with an entry for each hierarchy level.
-Since combined_portein.tsv only holds protein level data, the hierarchy list has one element, and we use it to specify which column contains the protein identifiers.
-We also need to specify which column contains the protein abundances we want to use for the data analysis.
+Since `combined_portein.tsv` only holds protein level data, the hierarchy list has one element, and we use it to specify which column contains the protein identifiers.
+We also need to define which column contains the protein abundances we want to use for the data analysis.
 Finally, we have to specify which columns contain the explanatory variables of the analysis.

### 2022-05-19 inst/MyArticle/paper.Rmd
update formula
Added bioc links

@@
       toc: true
+      toc_appendix: false
+      toc_bib: false
 bibliography: bibliography.bib
@@
 Table \@ref(tab:tableOverview) gives an overview of the models and features supported by these packages.
-We see that packages such as $limma$ and $proDA$ allow us to fit a comprehensive variety of models and test various hypotheses; however, in-depth knowledge of design matrix specification using the R formula interface is required [@law2020guide].
+We see that packages such as `r BiocStyle::Biocpkg("limma")` and `r BiocStyle::Biocpkg("proDA")` allow us to fit a comprehensive variety of models and test various hypotheses; however, in-depth knowledge of design matrix specification using the R formula interface is required [@law2020guide].
 
@@
 ```{r eval=FALSE, echo=TRUE}
-c("contrast_name" = "`factor_1level_a:factor_2level_1` - `factor_1level_a:factor_2level_2`")
+c("contrast_name" =
+    "`factor_1level_a:factor_2level_1` - `factor_1level_a:factor_2level_2`")
 
@@
 LOD    &Limit Of Detection                                 \\
+MAR    & Missing At Random                                 \\ 
+MCAR   & Missing Completely At Random                      \\
 MS     &mass spectrometry                                  \\
@@
 
-# Supplements
+# Appendix
 

### 2022-05-30 inst/MyArticle/paper.Rmd
adding changes from Maria

@@
     Protein mass spectrometry is widely used for quantitative proteomics studies, relative protein quantification, and differential expression analysis of proteins.
-Nevertheless, there is still a need for a flexible and easy-to-use application programmer interface in R that transparently supports a variety of well principled statistical procedures.
-The package can model simple experimental designs with a single explanatory variable and complex experiments with multiple factors and hypothesis testing.
-    The `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package integrates, using an R6 application programmer interface, essential steps of the mass spectrometry-based differential expression analysis workflow: quality control, data normalization, protein aggregation, statistical modeling, hypothesis testing, and sample size estimation.
+Nevertheless, there is still a need for a flexible and easy-to-use application programming interface in R that transparently supports a variety of well principled statistical procedures.
+The prolfqua package can model simple experimental designs with a single explanatory variable and complex experiments with multiple factors and hypothesis testing.
+    It is an package that integrates essential steps of the mass spectrometry-based differential expression analysis workflow: quality control, data normalization, protein aggregation, statistical modeling, hypothesis testing, and sample size estimation.
      Furthermore, the package implements benchmark functionality that can help to compare data acquisition, data preprocessing, or data modeling methods using a gold standard dataset.
 Finally, we show that the implemented methods allow sensitive and specific differential expression analysis.
-     The R package is available through https://github.com/fgcz/prolfqua.
+     The R package is available on github https://github.com/fgcz/prolfqua.
 urlcolor: blue
@@
 Hence, measuring changes in protein abundance is the subject of active research [@vidova2017review].
-Bottom-up mass spectrometric methods, where proteins are specifically and reproducible digested into protein fragments - peptides, are employed to identify and quantify proteins in complex samples containing hundreds to thousand of proteins, [@Bubis2017, @da2020philosopher].
+Bottom-up mass spectrometric methods, where proteins are specifically and reproducible digested into protein fragments - peptides, are employed to identify and quantify proteins in complex samples containing hundreds to thousand of proteins, [@Bubis2017; @da2020philosopher].
 The peptides are first separated by their chemical and physical properties using High-Pressure Liquid Chromatography (HPLC).
 Afterward, they are weighted, identified, and quantified using mass spectrometric techniques, e.g., Electro-Spray-Ionization Mass Spectrometry (ESI-MS).
-Finally, peptide identification is achieved by fragmenting and matching the measured fragment masses to theoretical masses computed from known peptide sequences[@eng2015deeper, @yu2016pipi, @kong2017msfragger].
-For quantification, intact peptide ions [@yu2020fast, @Cox2008MaxQuant] or products of peptide ion fragmentation [@rost2014openswath, @demichev2020dia] are counted and aggregated to obtain precursor abundances.
+Finally, peptide identification is achieved by fragmenting and matching the measured fragment masses to theoretical masses computed from known peptide sequences [@eng2015deeper; @yu2016pipi; @kong2017msfragger].
+For quantification, intact peptide ions [@yu2020fast; @Cox2008MaxQuant] or products of peptide ion fragmentation [@rost2014openswath; @demichev2020dia] are counted and aggregated to obtain precursor abundances.
 Finally, the identified and quantified peptides are assigned to proteins based on protein sequence information.
@@
 Proteomics quantification experiments enable monitoring relative abundances of thousands of proteins in biological samples.
-Most studies use parallel-group designs, where one or many treatment groups are compared to the control group [@de2022apoe2, @laubscher2021baf].
-More recently, more complex experimental designs with an increasing number of samples are studied, e.g., factorial designs and longitudinal studies (time series), sometimes with repeated measurements on the same subject [@tan2022proteomic, @meier2021reduced].
+Most studies use parallel-group designs, where one or many treatment groups are compared to the control group [@de2022apoe2; @laubscher2021baf].
+More recently, more complex experimental designs with an increasing number of samples are studied, e.g., factorial designs and longitudinal studies (time series), sometimes with repeated measurements on the same subject [@tan2022proteomic; @meier2021reduced].
 The data can be modeled using linear fixed-, mixed-, or random-effects models [@Bates2015FittingJSS].
@@
 Nevertheless, highly abundant proteins can suppress the detection of other proteins in a sample, an MAR mechanism.
-Furthermore, a weak correlation between the number of missing measurements in a condition and the intensities of the observed features can be observed, which is caused by the limit of detection (LOD), an MNAR mechanism [@mcgurk2020use].
+Furthermore, a weak correlation between the number of missing measurements in a condition and the abundance of the quantified protein can be observed, which is caused by the limit of detection (LOD), an MNAR mechanism [@mcgurk2020use].
 
@@
 At the same time, `r BiocStyle::Biocpkg("limma")` allows to specify a design matrix using a linear model formula and implements the empirical Bayes variance shrinkage method.
-The R-package `r BiocStyle::Biocpkg("PECA")`, performs a roll-up of peptide level differences and peptide level $p$-value estimates, obtained from `r BiocStyle::Biocpkg("limma")` or `r BiocStyle::Biocpkg("PECA")`, to protein level estimates.
+The package `r BiocStyle::Biocpkg("PECA")`, performs a roll-up of peptide level differences and peptide level $p$-value estimates, obtained from `r BiocStyle::Biocpkg("limma")` or `r BiocStyle::Biocpkg("PECA")`, to protein level estimates.
 Furthermore, `r BiocStyle::Biocpkg("msqrob2")` implements two models: robust linear models fitted to protein intensities and robust ridge regression fitted to peptide intensities and combines them with empirical Bayes variance shrinkage.
@@
   MSqRob2_rlm = c("pd" = "Y", "rm" = "Y*", fd = "Y", int = "Y",eb = "Y", md = "?"),
-  proDA = c("pd" = "Y", "rm" = "Y*", fd = "Y", int = "Y",eb = "Y", md = "Y"),
+  proDA = c("pd" = "Y", "rm" = "Y*", fd = "Y", int = "Y", eb = "Y", md = "Y"),
   prolfqua = c("pd" = "Y", "rm" = "Y", fd = "Y", int = "Y", mem = "Y",eb = "Y", md = "Y"))
@@
 
-When developing the R package `r BiocStyle::Githubpkg("fgcz/prolfqua")` we were inspired by the R package `r BiocStyle::CRANpkg("caret")` [@caret2008], which enables to call various machine learning (ML) methods, making selecting the best ML algorithm for your problem easy.
-We aimed for a similar R package for the analysis of LFQ data.
+When developing the R package `r BiocStyle::Githubpkg("fgcz/prolfqua")` we were inspired by the R package `r BiocStyle::CRANpkg("caret")` [@caret2008], which enables us to call various machine learning (ML) methods, which makes selecting the best ML algorithm for your problem easy.
+We aimed for a similar R package for the differential expression analysis of proteomics data.
 However, after examining the R packages for modeling proteomics quantification data, we observed that model specification, input, and output formats differ widely and wildly.
-However, what these packages have in common is that they fit linear model either to peptide or protein intensities, determine contrasts among conditions, and afterward, apply empirical Bayes variance shrinkage.
-Therefore, the revised goal was to provide a modular object oriented (OO) design, with R linear models as a core, where features such as $p$-value aggregation, e.g., `r BiocStyle::Biocpkg("PECA")`, variance shrinkage, or modelling of missing observations can be added.
+However, these packages have in common: they fit linear models either to peptide or protein intensities, determine contrasts among conditions, and afterward, apply empirical Bayes variance shrinkage.
+Therefore, the revised goal was to provide a modular object-oriented (OO) design, with R linear models as a core, where we can add features such as $p$-value aggregation, e.g., `r BiocStyle::Biocpkg("PECA")`, variance shrinkage, or modeling of missing observations.
 
 Furthermore, the functionality of `r BiocStyle::Githubpkg("fgcz/prolfqua")` also includes methods specific to proteomics data.
-For example, we implemented strategies to estimate protein intensities from peptide intensities: top N, Tukey's median polish, robust linear models [@Grossmann2010, @goeminne2020msqrob].
+For example, we implemented strategies to estimate protein intensities from peptide intensities: top N [@Grossmann2010], Tukey's median polish, robust linear models [@goeminne2020msqrob].
 Furthermore, peptide or protein abundances can be scaled and transformed, using robust scaling, transformed using _quantile_ normalization or `r BiocStyle::Biocpkg("vsa")` to remove systematic differences among samples and heteroscedasticity.
-In this respect, `r BiocStyle::Githubpkg("fgcz/prolfqua")` can be compared with other R packages such `r BiocStyle::Biocpkg("DEP")` [@DEP2018] or `r BiocStyle::Biocpkg("POMA")` [@POMA2021] which support the entire LFQ analysis pipeline.
+In this respect, `r BiocStyle::Githubpkg("fgcz/prolfqua")` can be compared with other R packages such as `r BiocStyle::Biocpkg("DEP")` [@DEP2018] or `r BiocStyle::Biocpkg("POMA")` [@POMA2021] which support the entire differential expression analysis pipeline.
 
@@
 We use the Ionstar [@shen2018ionstar] dataset to benchmark the modeling methods implemented within `r BiocStyle::Githubpkg("fgcz/prolfqua")` and to compare our results with those of `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`.
-Since group sizes are relatively small, typically with four or five subjects per group, a high power of the tests is a relevant criterion to asses the modelling method.
+Since group sizes are relatively small, typically with four or five subjects per group, the high power of the tests is a relevant criterion to assess the modeling method.
 Furthermore, the quantified proteins can be ranked using the estimated fold-change, $t$-statistics, or scaled $p$-value, and then subjected to gene set enrichment (GSEA) or over-representation analysis [@subramanian2005gene] to determine up or down-regulated groups of proteins.
-However, the model must also provide an unbiased estimate of the false discovery rate (FDR) to manage expectations when selecting protein lists for follow-up experiments.
+Furthermore, the statistical model must provide an unbiased estimate of the false discovery rate (FDR) to manage expectations when selecting protein lists for follow-up experiments.
+We will use the partial area under the receiver operator curve (ROC) to assess the power of the tests and compare the FDR with the false discovery proportion (FDP).
 
@@
 
-We store all the data needed for analysis in a single data frame in a tidy table, i.e., every column is a variable, every row is an observation, every cell is a single value[@tidydata2014].
+We store all the data needed for analysis in a single data frame in a tidy table, i.e., every column is a variable, every row is an observation, every cell is a single value [@tidydata2014].
 Using an __R6__[@R6cite] configuration object, we specify what variable is in which column, making it easy to integrate new inputs in `r BiocStyle::Githubpkg("fgcz/prolfqua")` if provided as a tidy table.
@@
 For popular software like _MaxQuant_ or _MSFragger_, which stores the same variable, e.g., intensity, in multiple columns, one for each sample, we implemented methods that transform the data into tidy tables.
-Relying on the tidy data table enable us to easily interface with many data manipulation, visualization, and modeling methods implemented in base R[@RCoreTeam2021] and the tidyverse[@tidyverse2019].
+Relying on the tidy data table enable us to easily interface with many data manipulation, visualization, and modeling methods implemented in base R [@RCoreTeam2021] and the tidyverse [@tidyverse2019].
 We use __R6__ classes to structure the functionality of the package (see Figure \@ref(fig:LFQData) and Figure \@ref(fig:ContrastUML)).
@@
 \begin{eqnarray} 
-\textrm{var} \hat\beta_c &=& c\sigma^2 (X^T X)^{-1} c^T
+\textrm{var} (\hat\beta_c) &=& \sqrt{ c^T \sigma^2 (X^T X)^{-1} c }
 \end{eqnarray}
@@
 
-The package prolfqua provides function to determine the vector of $parameter$ weights $c$, from a contrast specification string:
+[TODO WEW : provide model formula]
+
+The package `r BiocStyle::Githubpkg("fgcz/prolfqua")` provides function to determine the vector of $parameter$ weights $c$, from a linear models and a contrast specification string.
+
+If the linear model was specified using the following R formula:
+
+```{r}
+Intensity ~ factor_1 + factor_2 + factor_1:factor_2
+```
+
+then the contrasts ca
 
 ```{r eval=FALSE, echo=TRUE}
-c("contrast_name" = "factorlevel_1 - factorlevel_2")
+c("contrast_name" = "factor_1level_1 - factor_1level_2")
 ```
@@
 Models in `r BiocStyle::Githubpkg("fgcz/prolfqua")` are specified using R’s linear model and mixed model formula interface.
-Therefore, knowledge of the R regression model infrastructure [@faraway2016extending,@MASS2002] is an advantage when using our package.
+Therefore, knowledge of the R regression model infrastructure [@faraway2016extending; @MASS2002] is an advantage when using our package.
 Furthermore, this glass box approach should make it easy to reimplement an analysis which was performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` using base R or other programming languages by reading the analysis script.
@@
 However, when using MaxLFQ intensities reported by MaxQuant, _proDA_ performs significantly worse ($pAUC_{10}(\textrm{t-statistics}) = 66%$) compared with results obtained when protein abundances are estimated from peptide abundances using Tukey's median polish ($pAUC_{10}(\textrm{t-statistics}) = 73%$).
-Last but not least, the software [@Cox2008MaxQuant,@yu2020fast], identifying proteins, and generating the quantification values can also significantly contribute to the performance of the entire pipeline, altering the number of identified proteins and the sensitivity and specificity of the differential expression analysis.
+Last but not least, the software [@Cox2008MaxQuant; @yu2020fast], identifying proteins, and generating the quantification values can also significantly contribute to the performance of the entire pipeline, altering the number of identified proteins and the sensitivity and specificity of the differential expression analysis.
 
@@
 
-\begin{table}[H]
+\begin{table}
 \centering
-\begin{tabular}[H]{ll}
+\begin{tabular}{ll}
 \toprule
@@
 LC-MS  &Liquid Chromatography followed by Mass Spectrometry\\
-LFQ    &Label-Free Quantification                          \\
 LOD    &Limit Of Detection                                 \\

### 2022-05-30 inst/MyArticle/paper.Rmd
Omit appendix subsections in the table of contents
Link prolfqua in abstract

@@
 Nevertheless, there is still a need for a flexible and easy-to-use application programming interface in R that transparently supports a variety of well principled statistical procedures.
-The prolfqua package can model simple experimental designs with a single explanatory variable and complex experiments with multiple factors and hypothesis testing.
+The `r BiocStyle::Githubpkg("fgcz/prolfqua")` package can model simple experimental designs with a single explanatory variable and complex experiments with multiple factors and hypothesis testing.
     It is an package that integrates essential steps of the mass spectrometry-based differential expression analysis workflow: quality control, data normalization, protein aggregation, statistical modeling, hypothesis testing, and sample size estimation.
@@
 
-## Creating a prolfqua configuration
+## Creating a prolfqua configuration {.unlisted .unnumbered}
 
-The following code demonstrates how we use prolfqua to analyze protein intensities reported in the MSFragger `combined_protein.tsv` file.
+The following code demonstrates how we use `r BiocStyle::Githubpkg("fgcz/prolfqua")` to analyze protein intensities reported in the MSFragger `combined_protein.tsv` file.
 First, we create a tidy table containing the protein abundances by reading the `combined_protein.tsv` file using  `tidy_MSFragger_combined_protein.` Then, we read the sample annotation from the file `annotation.xlsx` file.
@@
 
-## Comparing benchmark results for proDA and prolfqua [REMOVE?]
+## Comparing benchmark results for proDA and prolfqua {.unlisted .unnumbered}
+
+TODO(WEW): remove?
 
@@
 
-## Session information
+## Session information {.unlisted .unnumbered}
 
@@
 \newpage
-## Miscellaneous
+## Miscellaneous {.unlisted .unnumbered}
 

### 2022-05-31 inst/MyArticle/paper.Rmd
Print abbreviation table in section
unlist appendix
adding JG comments.

@@
 ---
-title: prolfqua - An R-package for Proteomics Differential Expression Analysis.
+title: prolfqua - A Hany and Comprehensive R-package for Proteomics Differential Expression Analysis.
 author: 
@@
     Protein mass spectrometry is widely used for quantitative proteomics studies, relative protein quantification, and differential expression analysis of proteins.
-Nevertheless, there is still a need for a flexible and easy-to-use application programming interface in R that transparently supports a variety of well principled statistical procedures.
+Nevertheless, there is a need for a flexible and easy-to-use application programming interface in R that transparently supports a variety of well principled statistical procedures.
 The `r BiocStyle::Githubpkg("fgcz/prolfqua")` package can model simple experimental designs with a single explanatory variable and complex experiments with multiple factors and hypothesis testing.
-    It is an package that integrates essential steps of the mass spectrometry-based differential expression analysis workflow: quality control, data normalization, protein aggregation, statistical modeling, hypothesis testing, and sample size estimation.
+It integrates essential steps of the mass spectrometry-based differential expression analysis workflow: quality control, data normalization, protein aggregation, statistical modeling, hypothesis testing, and sample size estimation.
+The application programmer interface strives to be clear,  predictable, discoverable, and consistent to make proteomics data analysis easy and exciting.
      Furthermore, the package implements benchmark functionality that can help to compare data acquisition, data preprocessing, or data modeling methods using a gold standard dataset.
@@
 
-Proteins are carriers of function and structure in living cells.
+Proteins carry out most crucial functions and give structure to living cells.
 Hence, measuring changes in protein abundance is the subject of active research [@vidova2017review].
@@
 The peptides are first separated by their chemical and physical properties using High-Pressure Liquid Chromatography (HPLC).
-Afterward, they are weighted, identified, and quantified using mass spectrometric techniques, e.g., Electro-Spray-Ionization Mass Spectrometry (ESI-MS).
+Afterward, they are ionised, weighted, identified, and quantified using mass spectrometric techniques, e.g., Electro-Spray-Ionization Mass Spectrometry (ESI-MS).
 Finally, peptide identification is achieved by fragmenting and matching the measured fragment masses to theoretical masses computed from known peptide sequences [@eng2015deeper; @yu2016pipi; @kong2017msfragger].
-For quantification, intact peptide ions [@yu2020fast; @Cox2008MaxQuant] or products of peptide ion fragmentation [@rost2014openswath; @demichev2020dia] are counted and aggregated to obtain precursor abundances.
+For quantification, intact peptide ions [@yu2020fast; @Cox2008MaxQuant] or products of peptide ion fragmentation [@rost2014openswath; @demichev2020dia] are counted and aggregated to obtain peptide abundances.
 Finally, the identified and quantified peptides are assigned to proteins based on protein sequence information.
@@
 xx <- list(
-  MSstats = c("pd" = "Y", "rm" = "Y", "mem" = "y"),
+  MSstats = c("pd" = "Y", "rm" = "Y", "mem" = "Y"),
   ROPECA = c("pd" = "Y", "rm" = "Y*", fd = "Y", int = "Y", eb = "Y"),
@@
 knitr::kable(data.frame(bb), caption = "Rows - R package, Columns - types of models supported: pd - parallel design, rm - repeated measurements, fd - factorial design, int - interactions among factors, mem - mixed effect models, eb - empirical Bayes, md - missing data modelling.
-Y - yes, * - repeated measurements are modeled using a fixed effects.
+Y - yes, * - repeated measurements are modeled using a fixed effect.
 ?
@@
 We aimed for a similar R package for the differential expression analysis of proteomics data.
-However, after examining the R packages for modeling proteomics quantification data, we observed that model specification, input, and output formats differ widely and wildly.
+However, after examining the R packages for modeling proteomics quantification data, we observed that model specification, input, and output formats differ widely (and wildly).
 However, these packages have in common: they fit linear models either to peptide or protein intensities, determine contrasts among conditions, and afterward, apply empirical Bayes variance shrinkage.
@@
     gmi = c("group mean imputation", "impute missing group means", "ContrastsSimpleImpute"),
-    ROPECA = c("ROPECA"," ", "ContrastsROPECA"),
+    ROPECA = c("ROPECA","summarize peptide level models ", "ContrastsROPECA"),
     limma = c("limma" , "empirical Bayes variance shrinkage", "ContrastsModerated"),
@@
 colnames(xM) <- c("id", "model", "prolfqua functions")
-knitr::kable(tibble::as_tibble(xM[,1:3]), caption = "Models available in prolfqua. * in development",
+knitr::kable(tibble::as_tibble(xM[,1:3]), caption = "Models available in prolfqua.",
              format = "latex", booktabs=TRUE) |>
-    kableExtra::kable_styling(latex_options="scale_down")
+    kableExtra::kable_styling(latex_options="scale_down") |> 
+    kableExtra::add_footnote(label = "* - in development")
 
@@
 
-## Dataset
+## Dataset for Benchmarking
 
@@
 We annotated these dilutions from smallest to largest with the letters a, b, c, d, e.
-By comparing the various dilutions, we obtain various effect sizes , e.g., when comparing dilution _e_ (9%) against dilution _d_ (7.5%), the expected difference is $1.2$ for E.
+By comparing the various dilutions, we obtain different effect sizes, e.g., when comparing dilution _e_ (9%) against dilution _d_ (7.5%), the expected difference is $1.2$ for E.
 coli proteins and $1$ for human proteins.
@@
 The text files generated by MaxQuant are available in the _prolfquadata_ R package [@prolfquadata].
+To compare the performance of the various methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` we use only the contrasts resulting in minor differences $\Delta = (1.20, 1.25, 1.30, 1.50)$, because for bigger differences all models perform similarly.
 
@@
 
-The peptide abundances (from the _MaxQuant_ _peptide.txt_ file), were $log_2$ transformed, and subsequent robust z-score transformation, where median and the mean average deviation (mad) were obtained from the human proteins only.
+The peptide abundances (from the _MaxQuant_ _peptide.txt_ file), were $log_2$ transformed and subsequently scaled, where median and the mean average deviation (mad) were obtained from the human proteins only.
 We removed _one hit wonders_, i.e., proteins with a single peptide assignment.
-We also did not use imputation.
-Protein abundances are inferred from the peptide intensities using Tukey's median polish.
-We fitted the fixed effect linear models using protein abundances, the mixed effect linear model, and the `r BiocStyle::Biocpkg("PECA")` model we fitted to peptide intensities.
+Protein abundances are inferred from the peptide intensities using Tukey's median polish [@tukey1977exploratory].
+Finally, we fitted the fixed effect linear models, the dropout model `r BiocStyle::Biocpkg("proDA")` to protein abundances, the mixed effect linear model, and the `r BiocStyle::Biocpkg("PECA")` model to peptide intensities.
 
@@
 coli_ proteins, we know that the difference between dilutions should be $\beta \ne 0$.
-To compare the performance of the various methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` we use only the contrasts resulting in small fold-changes $\beta = (1.20, 1.25, 1.30, 1.50)$.
 
@@
 
-Valikangas [@valikangas2016systematic] and colleagues discuss and benchmark various methods of peptide or protein intensity normalization such as variance stabilizing normalization [@huber2002variance] or quantile normalization [@bolstad2003comparison].
-In this work we will be using a robust version of the z-score, where instead of the mean $\bar{x}$ we use the median and instead of the standard deviation $\tilde{S}$ the median absolute deviation (mad):
+Valikangas [@valikangas2016systematic] and colleagues discuss and benchmark various methods of peptide or protein intensity normalization, such as variance stabilizing normalization [@huber2002variance] or quantile normalization [@bolstad2003comparison].
+In this work, we will be using a robust version of the z-score, where instead of the mean $\bar{x}$ we use the median, and instead of the standard deviation $\tilde{S}$ the median absolute deviation (mad):
 
@@
 
-Because we need to estimate the protein fold-changes on the original scale, we have to multiply the $z$-score by the average variance of all the $n$ samples in the experiment.
+Because we need to estimate the differences among groups on the original scale, we must multiply the $z$-score by the average variance of all the $n$ samples in the experiment.
 
@@
 
-[TODO WEW : provide model formula]
-
 The package `r BiocStyle::Githubpkg("fgcz/prolfqua")` provides function to determine the vector of $parameter$ weights $c$, from a linear models and a contrast specification string.
 
-If the linear model was specified using the following R formula:
+The linear model below explain the observed protein abundances using the explanatory variables `factor_1` and `factor_2` and the interaction among them `factor_1:factor_2`,
 
@@
 
-then the contrasts ca
+then the contrasts among `group_1` and `group_2` defined by `factor_1` can be specified using the string below.
 
 ```{r eval=FALSE, echo=TRUE}
-c("contrast_name" = "factor_1level_1 - factor_1level_2")
+c("contrast_name" = "factor_1group_1 - factor_1group_2")
 ```
 
-Furthermore, contrasts for subgroups also can be specified using the syntax:
+Furthermore, contrasts for subgroups can be specified using the code below,
 
@@
 c("contrast_name" =
-    "`factor_1level_a:factor_2level_1` - `factor_1level_a:factor_2level_2`")
+    "`factor_1group_1:factor_2group_a` - `factor_1group_1:factor_2group_b`")
 
 ```
-where `factor_x` is the name of the explanatory variable, and `level_x`  are group labels.
 
-The following code shows an example where we specify two contrast: 
-a) where we compare Cells of type CMP/MEP with cells of type HSC, and b) where we contrast therapy NO with therapy NU for the celltype CMP/MEP [@meier2021reduced].
+where `factor_x` is the name of the explanatory variable, and `group_x`  are group labels.
+
+The following code shows an example where we specified two contrast, the first to compare Cells of type CMP/MEP with cells of type HSC, and the second where we contrast therapy NO with therapy NU for the celltype CMP/MEP [@meier2021reduced].
+Finally, we compute the array of weights defining the contrast, used to multiply the model coefficient $\beta$.
 
@@
 
-Missing observations lead to different group sizes, that is, unbalanced designs.
+Missing observations lead to different group sizes, which results in unbalanced designs.
 Linear and mixed effect models implemented in R can handle unbalanced designs.
 Therefore, as long as one observation per group is available, they will produce unbiased estimates, and no imputation is needed.
+
 If there is no observation in a group, we assume that the protein is unobserved because the protein abundance is below the limit of detection (LOD) of the MS instrument.
@@
 
-When computing differences $\Delta$ among two groups $a$ and $b$, we will use the group mean $\bar{a}$ or $\bar{b}$ estimated from the data, or if no observations are present in a group we use $A_{LOD}$, e.g., If there are no observations in group $b$ then :
+When computing differences $\Delta$ among two groups $a$ and $b$, we will use the group mean $\bar{a}$ or $\bar{b}$ estimated from the data, or if no observations are present in a group, we use $A_{LOD}$, e.g., If there are no observations in group $b$ then :
 
@@
 Because we used the median as the i-th order statistic $i = \frac{n}{2} + 0.5$.
-Therefore, $\gamma = i = \frac{n}{2} + 0.5$ and $\delta = n - i + 1 = n - (\frac{n}{2} + 0.5) + 1 = \frac{n}{2} + 0.5 = \gamma$ are used to parameterize the cumulative distribution function (CDF) of the Beta distribution.
+Therefore, to parameterize the cumulative distribution function (CDF) of the Beta distribution, $\gamma = i = \frac{n}{2} + 0.5$ and $\delta = n - i + 1 = n - (\frac{n}{2} + 0.5) + 1 = \frac{n}{2} + 0.5 = \gamma$.
 We implemented this method in the class `ContrastROPECA` (Figure \@ref(fig:ContrastUML)).
@@
 The code snippets in this section demonstrate how a differential expression analysis workflow can be implemented using the `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package.
-As input data, we use a subset of the Ionstar dataset containing $163$ proteins and $1258$ peptides.
+To speed up the computation of these examples, we use a subset of the Ionstar dataset containing $163$ proteins and $1258$ peptides.
 Peptide abundances are $\log_2$ transformed and robust z-score scaled using the method `robscale`.
-Using the LFQDataPlotter class, we show the distribution of the normalized peptide abundances in Figure \@ref(fig:prepro) Panel A.
-Afterwards, protein intensities are estimated from peptide intensities using Tukey's median polish.
+Using the `LFQDataPlotter` class, we show the distribution of the normalized peptide abundances in Figure \@ref(fig:prepro) Panel A.
+Afterwards, protein intensities are estimated from peptide intensities using Tukey's median polish [@tukey1977exploratory].
 Figure \@ref(fig:prepro) Panel B shows the peptide intensities and the estimated protein intensities.
@@
 lfqd <-  t$log2()$robscale()$lfq
+lfqd$rename_response("peptide_abundance")
 
@@
 lfqp <- agr$medpolish()
+lfqp$rename_response("protein_abundance")
 
@@
 
-## compute contrasts from linear model and with imputation
+## compute contrasts from linear model, moderate t-statistics and p-values
 contr <- prolfqua::Contrasts$new(models, contrasts) |> 
@@
 
+# compute contrasts using imputation imputation and moderate
 conI <- prolfqua::ContrastsSimpleImpute$new( lfqp, contrasts) |> 
@@
 
+# merge contrasts, to obtain differences for all proteins
 contrasts <- prolfqua::addContrastResults(prefer = contr, add = conI)
@@
 Models in `r BiocStyle::Githubpkg("fgcz/prolfqua")` are specified using R’s linear model and mixed model formula interface.
-Therefore, knowledge of the R regression model infrastructure [@faraway2016extending; @MASS2002] is an advantage when using our package.
-Furthermore, this glass box approach should make it easy to reimplement an analysis which was performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` using base R or other programming languages by reading the analysis script.
+Therefore, knowledge of the R regression model infrastructure [@faraway2016extending; @MASS2002] is advantageous when using our package.
+Furthermore, this glass box approach should make it easy to reimplement an analysis performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` using base R or other programming languages by reading the analysis script.
 However, acknowledging the R formula interface’s complexity to non-statisticians and the popularity of `r BiocStyle::Biocpkg("MSstats")`, we provide the functionality to suggest a model formula from the sample annotation provided in tabular form similarly to `r BiocStyle::Biocpkg("MSstats")` [@Choi2014].
@@
 Using the tidy table to model the data ensures interoperability with other proteomics-related packages that manage their data with tidy-tables, e.g., `r BiocStyle::CRANpkg("protti")` [@quast2022protti].
-Furthermore, the use of `R6` classes, which encapsulate the configuration and the data, allows writing very concise code (no function arguments needed), and auto-completion support in the Rstudio editor makes it easy for novices to explore `r BiocStyle::Githubpkg("fgcz/prolfqua")`s functionality \@ref(fig:tabCompletion).
+Furthermore, the use of `R6` classes, which encapsulate the configuration and the data, allows for writing very concise code (no function arguments needed).
+Auto-completion support in the Rstudio editor makes it easy for novices to explore `r BiocStyle::Githubpkg("fgcz/prolfqua")`s functionality (Figure \@ref(fig:tabCompletion)).
 To simplify integration of `r BiocStyle::Githubpkg("fgcz/prolfqua")` with Bioconductor based workflows there is an method to convert the LFQData class into a `r BiocStyle::Biocpkg("SummarizedExperiment")`.
@@
 This integration enables our users to select the input and basic settings in a graphical user interface.
-Then, as an output, the user receives a report including input files, the R markdown file and R scripts necessary to replicate the analysis using their in-house R installation.
+Then, as an output, the user receives a report including input files, the R markdown file, and R scripts necessary to replicate the analysis using their in-house R installation.
 In this way, we also support the reproducible science efforts of the community.
@@
 
-A relevant parameter is the number of contracts a given method estimates from the data (see Figure \@ref(fig:FDRfdp) Panel B).
-It indicates how robust the procedure works in the presence of missing data.
-For each protein, we can estimate four contrast which for 4046 proteins, at least two peptides resulted in $16184$ possible contrast estimates.
-Because the set of proteins for which differences are estimated differs slightly for each method, it makes comparing them complicated.
+A relevant parameter is the number of proteins for which a method estimated differences (see Figure \@ref(fig:FDRfdp) Panel B).
+It indicates how robust the procedure works in the presence of missing observations.
+For each protein, we tried to determine four differences ($\Delta = (1.20, 1.25, 1.30, 1.50)$).
+Given 4046 proteins with at least two peptides, there are in total $16184$ possible differences.
+However, some methods can not estimate all of them.
+The set of proteins with effect size estimates might differ slightly for each method, biasing direct comparison of scores such as pAUC.
 For instance, we observe that `r BiocStyle::Biocpkg("MSstats")` estimates $16058$ while the mixed effect models estimate the fewest with $15940$.
-Hence, to conclude that one method shows a better performance, the score, e.g., pAUC, needs to be greater, but also the number of proteins with differential expression results needs to be equal or larger.
+Hence, to conclude that one method shows a better performance, the pAUC needs to be greater, and the number of proteins with differential expression results needs to be equal or larger.
 
-Figure \@ref(fig:FDRfdp) Panel A, shows how various estimates obtained from the models, i.e.
-the difference between groups, $t$-statistics and scaled $p$-values allow to identify true positives (TP) given a false positive rate.
-Furthermore, ordering the proteins by the $t$-statistic or $p$-value leads to a higher $pAUC_{10}$ then when ordering by the estimated difference among groups.
+Figure \@ref(fig:FDRfdp) Panel A shows how various estimates obtained from the models, i.e., the difference between groups, $t$-statistics, and scaled $p$-values allow identifying true positives (TP) given a false positive rate.
+Furthermore, ordering proteins by the $t$-statistic or $p$-value leads to a higher $pAUC_{10}$ than when ordering by the estimated difference among groups.
 Suppose an accurate estimate of the difference among groups is essential.
-In that case, the linear models fitted to protein intensities, estimated using Tukey's median polish, show a better performance than the methods directly modeling peptide intensities.
+In that case, the linear models fitted to protein intensities, calculated using Tukey's median polish, show a better performance than the methods directly modeling peptide intensities.
 We speculate that outliers at the peptide level do not affect the protein estimates when using Tukey's median polish method.
@@
 Interestingly, the $pAUC$ increases when using $p$-values instead of the $t$-statistics for linear models, while it decreases for mixed effect models.
-The reason is erroneous denominator degrees of freedom estimation for many proteins in the mixed-effect models.
+The reason is erroneous denominator degrees of freedom estimation for many proteins in the case of the mixed-effect models.
 
-We also benchmark, if the $FDR$ obtained from a model is an unbiased estimate of the false discovery proportion $FDP$.
+We also benchmark if the $FDR$ obtained from a model is an unbiased estimate of the false discovery proportion $FDP$.
 Figure \@ref(fig:FDRfdp) Panel C draws on the horizontal axis the FDR determined from the model, and on the vertical axis, the FDP obtained from the confusion matrix.
-We see that most of the lines are below the diagonal, indicating that the FDR estimates obtained with `r BiocStyle::Githubpkg("fgcz/prolfqua")` are modestly conservative for this benchmark dataset provide.
+Most lines are below the diagonal, indicating that the FDR estimates are modestly conservative for this benchmark dataset.
 In the case of `r BiocStyle::Biocpkg("MSstats")`, we observe a high proportion of false discoveries for small $FDR$ values.
-In the case of `r BiocStyle::Biocpkg("PECA")`, the FDR estimates, obtained when Benjamini-Hochberg correcting the regulation probabilities, strongly overestimate the $FDP$.
+In the case of `r BiocStyle::Biocpkg("PECA")`, the FDR estimates, obtained by Benjamini-Hochberg correcting the regulation probabilities, strongly overestimate the $FDP$.
 
@@
 Panel C - Plots the false discovery proportion (FDP) as a function of the FDR.
-Ideally the FDR should be equal to the FDP.
+Ideally, the FDR should be equal to the FDP.
 Therefore larger distances from the diagonal are worse.
@@
 
-We focused in our benchmark on the comparison of the statistical modelling methods and fixing the pre-processing steps.
-However, there are other equally or even more important parameters of a protein quantification pipeline [some paper with workflow evaluation].
-One of them is the normalization of the intensities within the samples to remove systematic differences and to make them comparable [@pursiheimo2015optimization].
+We focused in our benchmark on comparing the statistical modeling methods and fixing the pre-processing steps.
+However, there are other equally or even more important parameters of a protein quantification pipeline [@frohlich2022benchmarking].
+One of them is the normalization of the intensities within the samples to remove systematic differences [@pursiheimo2015optimization].
 The method used to infer protein intensities from peptide intensities is an additional important factor [@Grossmann2010].
-For instance, the original _proDA_ publication uses MaxLFQ [@cox2014accurate] protein estimates.
-However, when using MaxLFQ intensities reported by MaxQuant, _proDA_ performs significantly worse ($pAUC_{10}(\textrm{t-statistics}) = 66%$) compared with results obtained when protein abundances are estimated from peptide abundances using Tukey's median polish ($pAUC_{10}(\textrm{t-statistics}) = 73%$).
-Last but not least, the software [@Cox2008MaxQuant; @yu2020fast], identifying proteins, and generating the quantification values can also significantly contribute to the performance of the entire pipeline, altering the number of identified proteins and the sensitivity and specificity of the differential expression analysis.
+For instance, the original `r BiocStyle::Biocpkg("proDA") publication uses MaxLFQ [@cox2014accurate] protein estimates.
+However, when using MaxLFQ intensities reported by MaxQuant, the $pAUC_{10}$ is significantly lower ($pAUC_{10}(\textrm{t-statistics}) = 66%$) compared with results obtained when protein abundances are estimated from peptide abundances using Tukey's median polish ($pAUC_{10}(\textrm{t-statistics}) = 73%$).
+Last but not least, the software [@Cox2008MaxQuant; @yu2020fast], identifying proteins and generating the quantification values can also significantly contribute to the performance of the entire pipeline, altering the number of identified proteins and the sensitivity and specificity of the differential expression analysis.
 
 We assessed the performance of different modeling approaches implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")`, `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`, using a benchmark dataset whose ground truth is known.
-However, since for each dilution only technical replicates are available, essential sources of variation present in every experiment, such as the biochemical and biological, are not measured.
+However, since only technical replicates are available for each dilution, essential sources of variation present in every experiment, such as the biochemical and biological, are not measured.
 The dataset captures only the variance from the chromatography, electro-spray, and mass spectrometric measurement method.
 Thus, while we can extrapolate some of the results obtained to more realistic datasets, we need to be careful not to over-interpret the results.
-Specifically, in more realistic experiments the observed variances will be higher and sensitivities and specificity, will be worse given the same sample sizes.
-Furthermore, the proportion of missing observations in real world datasets might also be higher.
+Specifically, the observed variances will be higher in more realistic experiments, and sensitivities and specificity will be worse given the same sample sizes.
+Furthermore, the proportion of missing observations in real-life datasets might also be higher.
 
@@
 
-\begin{table}
+\begin{table}[H]
 \centering
@@
 \newpage
+
 # References {-}
@@
 
-# Appendix
+# Appendix  {.unnumbered}
 
@@
 
-## Comparing benchmark results for proDA and prolfqua {.unlisted .unnumbered}
-
-TODO(WEW): remove?
-
-Here we compare the benchmark results for the `r BiocStyle::Biocpkg("proDA")` model and the linear model with variance moderation.
-We see that the difference estimates obtained with the `r BiocStyle::Biocpkg("proDA")` package show better diagnostic accuracy.
-However, the $t-statistics$ or $p-values$ show slightly worse diagnostic accuracy (Figure \@ref(fig:proDABenchmarks)).
-Figure \@ref(fig:proDAAllBenchmarks), compares the difference estimates obtained using both methods.
-
-(ref:proDABenchmarks) The barplots show the difference of the $pAUC_10$ obtained when using the proDA package or when using prolfqua.
-Postitive values indicate a better diagnostic performance of proDA while negative values a better diagnostic performance of prolfqua.
-
-```{r proDABenchmarks, fig.cap="(ref:proDABenchmarks)"}
-prodB <- allBenchmarks$benchmark_proDA$clone(deep=TRUE)
-proda <- prodB$.data
-modB <- allBenchmarks$benchmark_ProtModerated
-
-mod <- modB$.data
-mod <- mod[!is.na(mod$statistic),]
-modSum <- modB$pAUC_summaries()
-prodaMod <- dplyr::semi_join(proda, mod, by = c("contrast", "protein_Id"))
-prodB$.data <- prodaMod
-
-d <- prodB$missing_contrasts()
-prodSum <- prodB$pAUC_summaries()
-
-tmp <- dplyr::inner_join(prodSum$ftable$content,
-                         modSum$ftable$content,
-                         by=c("contrast","what"), suffix = c(".proDA",".lmMod")) |>
-  tidyr::unite("score_contr" , what, contrast, remove = FALSE)
-
-ggplot2::ggplot(tmp, ggplot2::aes(x = contrast, y = pAUC_10.proDA - pAUC_10.lmMod)) + ggplot2::geom_bar(stat="identity") + ggplot2::facet_wrap(~what) +  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust=1))
-```
-
-(ref:proDAAllBenchmarks) Comparison of difference estimates produced by proDA and produced by linear models implemented in prolfqua.
-
-```{r proDAAllBenchmarks, fig.cap="(ref:proDAAllBenchmarks)"}
-proda <- allBenchmarks$benchmark_proDA$data()
-imp <- allBenchmarks$benchmark_imputation$data()
-imp <- allBenchmarks$benchmark_ProtModerated$data()
-
-x <- dplyr::inner_join(imp, proda, by = c("contrast","protein_Id", "TP", "species"), suffix = c(".lm",".proda"))
-ggplot2::ggplot(x,ggplot2::aes( x = diff.lm, y = diff.proda) ) + ggplot2::geom_point() + ggplot2::facet_wrap(~contrast + species) + ggplot2::geom_hline(yintercept = 0, color="lightgray") + ggplot2::geom_vline(xintercept = 0, color="lightgray") + ggplot2::geom_abline(slope = 1) + ggplot2::theme_minimal()
-
-```
-
 ## Session information {.unlisted .unnumbered}

### 2022-05-31 inst/MyArticle/paper.Rmd
fix

@@
 Thus, while we can extrapolate some of the results obtained to more realistic datasets, we need to be careful not to over-interpret the results.
-Specifically, the observed variances will be higher in more realistic experiments, and sensitivities and specificity will be worse given the same sample sizes.
+Specifically, the observed variances will be higher in more realistic experiments, and sensitivity and specificity will worsen, given the same sample sizes.
 Furthermore, the proportion of missing observations in real-life datasets might also be higher.
 
-We can conclude that if we want to order the proteins according to the likelihood of being differentially regulated, to perform gene set enrichment analysis [@subramanian2005gene], the $t$-statistic is better suited than the fold-change estimate.
-Modeling the degrees of freedom when computing the $p$-values might further improve the inference, however, the observed this improvement is minuscule (see Figure \@ref(fig:FDRfdp) panel A).
-However, there is no such effect for the mixed effect model, most likely because the degrees of freedom are erroneously estimated for a large proportion of models.
-Therefore, for the fixed effect linear model, the empirical Bayes variance shrinkage, as suggested by [@Smyth2004linear], consistently improves the ranking of proteins compared with the unmoderated estimates (not shown), however fails to do so for mixed effect models.
+We can conclude that if we want to order the proteins according to the likelihood of being differentially regulated to perform gene set enrichment analysis [@subramanian2005gene], the $t$-statistic is better suited than the fold-change estimate.
+Modeling the degrees of freedom when computing the $p$-values might improve the inference.
+However, this improvement is minuscule (see Figure \@ref(fig:FDRfdp) panel A).
+There is no such effect for the mixed effect model, most likely because the degrees of freedom are erroneously estimated for many models.
+Therefore, for the fixed effect linear model, the empirical Bayes variance shrinkage, as suggested by [@Smyth2004linear], consistently improves the ranking of proteins compared with the unmoderated estimates (not shown) and fails to do so for mixed effect models.
 
 Computing the statistics at the peptide level, e.g., the $t$-statistics or $p$-value, then summarizing these statistics using their median produces the highest $AUC$ scores among all the tested models (see Figure \@ref(fig:FDRfdp) Panel A left).
-By using the beta distribution to model the number of peptides observed, we can further improve the $pAUC$ scores (see Figure \@ref(fig:FDRfdp) Panel A center).
-However, the properties of beta-based probabilities are not well understood; for instance, the p-values are not uniformely distributed under the null hypothesis (not shown).
-Furthermore, the FDR estimates obtained when correcting for multiple testing with the Benjamini-Hochberg method are biased overestimating the false discovery proportion (see Figure \@ref(fig:FDRfdp) Panel C).
+Furthermore, by using the beta distribution to model the number of peptides observed, we can further improve the $pAUC$ scores (see Figure \@ref(fig:FDRfdp) Panel A center).
+However, the properties of beta-based probabilities are not well understood; for instance, the p-values are not uniformly distributed under the null hypothesis (not shown).
+Furthermore, the FDR estimates obtained when correcting for multiple testing with the Benjamini-Hochberg method are biased and overestimate the false discovery proportion (see Figure \@ref(fig:FDRfdp) Panel C).
 Therefore, we can not recommend this method if an unbiased estimate of FDR is essential, which is frequently the case.

### 2022-05-31 inst/MyArticle/paper.Rmd
remove handy from title

@@
 ---
-title: prolfqua - A Hany and Comprehensive R-package for Proteomics Differential Expression Analysis.
+title: prolfqua - A Comprehensive R-package for Proteomics Differential Expression Analysis.
 author: 
@@
 The method used to infer protein intensities from peptide intensities is an additional important factor [@Grossmann2010].
-For instance, the original `r BiocStyle::Biocpkg("proDA") publication uses MaxLFQ [@cox2014accurate] protein estimates.
+For instance, the original `r BiocStyle::Biocpkg("proDA")` publication uses MaxLFQ [@cox2014accurate] protein estimates.
 However, when using MaxLFQ intensities reported by MaxQuant, the $pAUC_{10}$ is significantly lower ($pAUC_{10}(\textrm{t-statistics}) = 66%$) compared with results obtained when protein abundances are estimated from peptide abundances using Tukey's median polish ($pAUC_{10}(\textrm{t-statistics}) = 73%$).

### 2022-05-31 inst/MyArticle/paper.Rmd
Add FAIR
b-fabric and FAIR text cosmetics

@@
 
-To ease the usage barriers of the R package to users not being familiar with statistics and R programming, we integrated some workflows into our Local Laboratory Information Management System (LIMS) [b-fabric](https://fgcz-bfabric.uzh.ch/) @bfabric.
+To ease the usage barriers of the R package to users not being familiar with statistics and R programming, we integrated some workflows into our data management platform [b-fabric](https://fgcz-bfabric.uzh.ch/) @bfabric.
 This integration enables our users to select the input and basic settings in a graphical user interface.
 Then, as an output, the user receives a report including input files, the R markdown file, and R scripts necessary to replicate the analysis using their in-house R installation.
-In this way, we also support the reproducible science efforts of the community.
+In this way,  [b-fabric](https://fgcz-bfabric.uzh.ch/) helps scientists meet requirements from funding agencies, journals, and academic institutions to publish data according to the FAIR (Findable, Accessible, Interoperable and Reusable) [@FAIR] data principles.
 

### 2022-06-01 inst/MyArticle/paper.Rmd
Add quote
Remove dot in title

@@
 ---
-title: prolfqua - A Comprehensive R-package for Proteomics Differential Expression Analysis.
+title: prolfqua - A Comprehensive R-package for Proteomics Differential Expression Analysis
 author: 
@@
 # Introduction
+
+ \begin{quote}
+``Es ist nicht das Wissen, sondern das Lernen, nicht das Besitzen, sondern das Erwerben, nicht das Dasein, sondern das Hinkommen, was den größten Genuß gewährt.''
+
+Carl Friedrich Gauß
+ \end{quote}
 

### 2022-06-01 inst/MyArticle/paper.Rmd

@@
 \begin{quote}
-``Es ist nicht das Wissen, sondern das Lernen, nicht das Besitzen, sondern das Erwerben, nicht das Dasein, sondern das Hinkommen, was den größten Genuß gewährt.''
+To paraphrase provocatively, `machine learning is statistics minus any checking of models and assumptions'.
 
-Carl Friedrich Gauß
+-- Brian D. Ripley (about the difference between machine learning and statistics) useR! 2004, Vienna (May 2004).
 \end{quote}

### 2022-06-01 inst/MyArticle/paper.Rmd
Add license information and system requirements in abstract

@@
 Finally, we show that the implemented methods allow sensitive and specific differential expression analysis.
-     The R package is available on github https://github.com/fgcz/prolfqua.
+     The `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package is available on GitHub https://github.com/fgcz/prolfqua, distributed under the MIT licence, and runs on all plattforms supported by the [R free software environment for statistical computing and graphics](https://www.r-project.org/).
 urlcolor: blue
@@
 
--- Brian D. Ripley (about the difference between machine learning and statistics) useR! 2004, Vienna (May 2004).
+\begin{flushright}
+-- Brian D. Ripley, useR! 2004, Vienna.
+\end{flushright}
 \end{quote}

### 2022-06-02 inst/MyArticle/paper.Rmd
Mostly updates to the implementation section.

--- 2022-06-01 inst/MyArticle/paper.Rmd
@@
 
-We store all the data needed for analysis in a single data frame in a tidy table, i.e., every column is a variable, every row is an observation, every cell is a single value [@tidydata2014].
-Using an __R6__[@R6cite] configuration object, we specify what variable is in which column, making it easy to integrate new inputs in `r BiocStyle::Githubpkg("fgcz/prolfqua")` if provided as a tidy table.
-For example, to visualize tidy _Spectronaut_, or _Skyline_ outputs, or data in `r BiocStyle::Biocpkg("MSstats")` format, only a few lines of code are needed to update the `r BiocStyle::Githubpkg("fgcz/prolfqua")` configuration.
-For popular software like _MaxQuant_ or _MSFragger_, which stores the same variable, e.g., intensity, in multiple columns, one for each sample, we implemented methods that transform the data into tidy tables.
-Relying on the tidy data table enable us to easily interface with many data manipulation, visualization, and modeling methods implemented in base R [@RCoreTeam2021] and the tidyverse [@tidyverse2019].
+We store all the data needed for analysis in a single data frame in a tidy table, i.e., every column is a variable, every row is an observation, and every cell is a single value [@tidydata2014].
+Using an __R6__[@R6cite] configuration object \@ref(fig:LFQData), we specify what variable is in which column, making it easy to integrate new inputs in `r BiocStyle::Githubpkg("fgcz/prolfqua")` if provided as a tidy table.
+For example, to visualize tidy _Spectronaut_, _DiaNN_, or _Skyline_ outputs, or data in `r BiocStyle::Biocpkg("MSstats")` format, only a few lines of code are needed to update the `r BiocStyle::Githubpkg("fgcz/prolfqua")` configuration.
+The configuration encapsulates the differences of column names among the various input formats and enables the methods of prolfqua to be transparent, i.e., it preserves the differences of the processing tools in the figures and analysis.
+We show an example code for creating an MSFragger configuration in the Supplement.
+For popular software like _MaxQuant_, or_MSFragger_, which stores the same variable, e.g., intensity, in multiple columns, one for each sample, we implemented methods that transform the data into tidy tables.
+Furthermore, relying on the tidy data table enables us to interface with many data manipulation, visualization, and modeling methods, implemented in base R [@RCoreTeam2021] and the tidyverse [@tidyverse2019], easily.
 We use __R6__ classes to structure the functionality of the package (see Figure \@ref(fig:LFQData) and Figure \@ref(fig:ContrastUML)).
@@
 (ref:LFQData) Class Diagram of classes representing the proteomics data.
-The LFQData class encapsulates the data.
-Other classes are decorating the LFQData class with additional functionality.
-The LFQData_Plotter class uses the LFQData class to implement methods for data visualization.
-Similarly, the LFQData_Stats, LFQData_Summary, reference the LFQData class, group methods for variance and sample size estimation, or summarizing peptide and protein counts.
+The LFQData class encapsulates the quantitative proteomics data stored in a tidy table.
+An instance of the AnalysisTableConfiguration class specifies a mapping of columns in the tidy table.
+The LFQDataPlotter class and other classes decorate the LFQData class with additional functionality.
+Similarly, the `LFQDataStats` and`LFQDataSummary`, reference the LFQData class and group methods for variance and sample size estimation or summarizing peptide and protein counts.
 
 ```{r LFQData, echo=FALSE, fig.cap="(ref:LFQData)", out.width = '90%'}
-knitr::include_graphics("LFQDataUML.png")
+knitr::include_graphics("LFQData_UML_V2.png")
 ```
@@
 knitr::include_graphics("ContrastClassesUML.png")
-```
-
-```{r prolfquaModels}
-xM <- data.frame( 
-    lm = c("lm", "linear model for peptides or proteins", "strategy_lm, Contrasts"),
-    lmer = c("lmer" , "mixed effect models for peptides or proteins", "strategy_lmer, Contrasts"),
-    gmi = c("group mean imputation", "impute missing group means", "ContrastsSimpleImpute"),
-    ROPECA = c("ROPECA","summarize peptide level models ", "ContrastsROPECA"),
-    limma = c("limma" , "empirical Bayes variance shrinkage", "ContrastsModerated"),
-    pic = c("SAINTexpress", "protein interaction scoring", "runSaint, ContrastsSaintExpress"),
-    prodA = c("proDA", "dropout models", "strategy_proDA*, Contrasts_proDA*"), check.names =  FALSE)
-xM <- data.frame(t(xM))
-
-colnames(xM) <- c("id", "model", "prolfqua functions")
-knitr::kable(tibble::as_tibble(xM[,1:3]), caption = "Models available in prolfqua.",
-             format = "latex", booktabs=TRUE) |>
-    kableExtra::kable_styling(latex_options="scale_down") |> 
-    kableExtra::add_footnote(label = "* - in development")
-
 ```
@@
 We use those scores, i.e.
-the $pAUC$ at $10%$ FDR and the FDP, to examine how well the methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` model quantitative mass spectrometric high throughput data and compare them with results produced by `r BiocStyle::Biocpkg("MSstats")` [@MSstats2014] and `r BiocStyle::Biocpkg("proDA")` [@bioxrvproDA2020] (see also Figure \@ref(fig:FDRfdp))
+the $pAUC$ at $10%$ FDR and the FDP, to examine how well the methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` model quantitative mass spectrometric high throughput data and compare them with results produced by `r BiocStyle::Biocpkg("MSstats")` [@MSstats2014] and `r BiocStyle::Biocpkg("proDA")` [@bioxrvproDA2020].
+Table \@ref(tab:describeBenchmarkedModels) summarizes all modelling methods we evaluated.
 
@@
 allBenchmarks <- allBenchmarks[c("benchmark_imputation","benchmark_ProtModerated",  "benchmark_mixedModerated", "benchmark_ropeca","benchmark_merged","benchmark_mssstats","benchmark_proDA"   )]
+
 ```
@@
 res <- purrr::map_df(allBenchmarks, function(x){x$pAUC()})
+
 res <- res |> dplyr::mutate(whatfix = dplyr::case_when(what == "scaled.beta.based.significance" ~ "scaled.p.value", TRUE ~ what))
@@
 
-A relevant parameter is the number of proteins for which a method estimated differences (see Figure \@ref(fig:FDRfdp) Panel B).
+When comparing differential expression analysis performance, a relevant parameter is the number of proteins for which a method estimated differences (see Figure \@ref(fig:FDRfdp) Panel A).
 It indicates how robust the procedure works in the presence of missing observations.
@@
 However, some methods can not estimate all of them.
-The set of proteins with effect size estimates might differ slightly for each method, biasing direct comparison of scores such as pAUC.
-For instance, we observe that `r BiocStyle::Biocpkg("MSstats")` estimates $16058$ while the mixed effect models estimate the fewest with $15940$.
-Hence, to conclude that one method shows a better performance, the pAUC needs to be greater, and the number of proteins with differential expression results needs to be equal or larger.
+The set of proteins with effect size estimates might differ for each method, biasing direct comparison of scores such as pAUC.
+For instance, we observe that `r BiocStyle::Biocpkg("MSstats")` estimates $16058$ group differences while the mixed effect models estimates the fewest with $15940$.
+Hence, to conclude that one method shows a better performance, it does not suffice that pAUC is greater, but the number of proteins with differential expression results needs to be equal or larger.
 
-Figure \@ref(fig:FDRfdp) Panel A shows how various estimates obtained from the models, i.e., the difference between groups, $t$-statistics, and scaled $p$-values allow identifying true positives (TP) given a false positive rate.
-Furthermore, ordering proteins by the $t$-statistic or $p$-value leads to a higher $pAUC_{10}$ than when ordering by the estimated difference among groups.
+Figure \@ref(fig:FDRfdp) Panel B shows how various estimates obtained from the models, i.e., the difference between groups, $t$-statistics, and scaled $p$-values allow identifying true positives (TP) given a false positive rate (FPR) of $10%$, by displaying the partial area under the ROC.
+Ordering proteins by the $t$-statistic or $p$-value leads to a higher $pAUC_{10}$ than when ordering by the estimated difference among groups.
 Suppose an accurate estimate of the difference among groups is essential.
-In that case, the linear models fitted to protein intensities, calculated using Tukey's median polish, show a better performance than the methods directly modeling peptide intensities.
-We speculate that outliers at the peptide level do not affect the protein estimates when using Tukey's median polish method.
+In that case, the linear models fitted to protein intensities, calculated using Tukey's median polish, perform better than those directly modeling peptide intensities, e.g., `ropeca` or `prot_mixed_effect_moderated` (see Table \@ref(tab:describeBenchmarkedModels) Abundance column).
+We speculate that outliers at the peptide level do not affect the protein estimates when using Tukey's median polish method, a non-parametric method to infer protein abundances.
 This hypothesis could be examined, by including other forms of protein intensity inferences implemented in `prolfqua`, e.g.
@@
 
-There are only minor differences in the $pAUC_{10}$ when comparing the $t$-statistics or the scaled $p$-value  (see Figure \@ref(fig:FDRfdp) Panel A).
+There are only minor differences in the $pAUC_{10}$ between the $t$-statistics or the scaled $p$-value  (see Figure \@ref(fig:FDRfdp) Panel A).
 Interestingly, the $pAUC$ increases when using $p$-values instead of the $t$-statistics for linear models, while it decreases for mixed effect models.
@@
 
-(ref:FDRfdp) Panel A - Partial area under the ROC curve at $10\%$ FPR ($pAUC_10$) for all contrasts and three different statistics: the difference among conditions, the scaled $p$-value (sign(diff) $\cdot$ p.value) and the $t$-statistics (higher is better).
+(ref:FDRfdp) Panel A - Number of estimated contrasts for each modeling method (higher is better).
+Panel B - Partial area under the ROC curve at $10\%$ FPR ($pAUC_10$) for all contrasts and three different statistics: the difference among conditions, the scaled $p$-value (sign(diff) $\cdot$ p.value) and the $t$-statistics (higher is better).
 The red line indicates the average area under the curve of all methods.
-Panel B - Number of estimated contrasts for each modeling method (higher is better).
 Panel C - Plots the false discovery proportion (FDP) as a function of the FDR.
@@
 
-```{r FDRfdp, fig.cap = "(ref:FDRfdp)", out.width = '90%', fig.width=12, fig.height=10}
+```{r FDRfdp, fig.cap = "(ref:FDRfdp)", out.width = '90%', fig.width=10, fig.height=10}
 dd <- purrr::map_df(allBenchmarks, function(x){res <- x$smc$summary; res$name <- x$model_name;res})
@@
   ggplot2::geom_bar(stat = "identity", fill="white", colour = "black") + 
-  ggplot2::coord_cartesian(ylim = c(min(dds$nrcontrasts) - 100, max(dds$nrcontrasts) + 10)) +
+  ggplot2::coord_cartesian(ylim = c(min(dds$nrcontrasts) - 200, max(dds$nrcontrasts) + 10)) +
   ggplot2::theme(axis.text.x = ggplot2::element_text(angle = -90, vjust = 0.5)) +
-  ggplot2::geom_text(ggplot2::aes(label = round(nrcontrasts, digits = 1)),
-            vjust = 1, hjust = -0.2, angle = -90) + 
-  ggplot2::geom_text(ggplot2::aes(label = paste0("(",round(percent, digits = 1),"%)")),
-            vjust = -1, hjust = -0.2, angle = -90) 
+  ggplot2::geom_text(ggplot2::aes(label = paste0(round(nrcontrasts, digits = 1), paste0("  (",round(percent, digits = 1),"%)"))),
+            vjust = 0, hjust = -0.2, angle = -90) #+ 
+  #ggplot2::geom_text(ggplot2::aes(label = paste0("(",round(percent, digits = 1),"%)")),
+            #vjust = -1, hjust = -0.2, angle = -90) 
 pB <- nrgg + ggplot2::labs(tag = "B")
@@
 
-hlay = rbind(c(1,1),
-            c(2,3))
+hlay = rbind(c(2,1,1),
+            c(2,3,3))
 
@@
 
-We focused in our benchmark on comparing the statistical modeling methods and fixing the pre-processing steps.
-However, there are other equally or even more important parameters of a protein quantification pipeline [@frohlich2022benchmarking].
-One of them is the normalization of the intensities within the samples to remove systematic differences [@pursiheimo2015optimization].
-The method used to infer protein intensities from peptide intensities is an additional important factor [@Grossmann2010].
-For instance, the original `r BiocStyle::Biocpkg("proDA")` publication uses MaxLFQ [@cox2014accurate] protein estimates.
-However, when using MaxLFQ intensities reported by MaxQuant, the $pAUC_{10}$ is significantly lower ($pAUC_{10}(\textrm{t-statistics}) = 66%$) compared with results obtained when protein abundances are estimated from peptide abundances using Tukey's median polish ($pAUC_{10}(\textrm{t-statistics}) = 73%$).
-Last but not least, the software [@Cox2008MaxQuant; @yu2020fast], identifying proteins and generating the quantification values can also significantly contribute to the performance of the entire pipeline, altering the number of identified proteins and the sensitivity and specificity of the differential expression analysis.
+```{r prolfquaModels}
+xM <- data.frame( 
+    lm = c("strategy_lm, Contrasts", "linear modelling of peptide or protein abundances and group difference estimation"),
+    lmer = c("strategy_lmer, Contrasts", "mixed effect modelling of peptide or protein abundances and group differences estimation" ),
+    gmi = c("ContrastsSimpleImpute" , "group difference estimation when no observations in one of the groups" ),
+    ROPECA = c("ContrastsROPECA", "estimating group differences for proteins by summarizing peptide differences"),
+    limma = c( "ContrastsModerated", "empirical Bayes variance shrinkage for group difference estimates (limma)"),
+    pic = c("runSaint, ContrastsSaintExpress", "protein interaction scoring (SAINTExpress)"),
+    prodA = c( "strategy_proDA*, Contrasts_proDA*", "adapter to the probabilistic dropout model implemented in proDA"), check.names =  FALSE)
+xM <- data.frame(t(xM))
 
-We assessed the performance of different modeling approaches implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")`, `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`, using a benchmark dataset whose ground truth is known.
-However, since only technical replicates are available for each dilution, essential sources of variation present in every experiment, such as the biochemical and biological, are not measured.
-The dataset captures only the variance from the chromatography, electro-spray, and mass spectrometric measurement method.
+colnames(xM) <- c("prolfqua functions", "model")
+knitr::kable(tibble::as_tibble(xM), caption = "Prolfqua functions to fit various models.",
+             format = "latex", booktabs=TRUE) |>
+    kableExtra::kable_styling(latex_options="scale_down") |> 
+    kableExtra::add_footnote(label = "* in development")
+
+```
+
+Using a benchmark dataset whose ground truth is known (see Methods), we assessed the performance of different modeling approaches implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` (Tables \@ref(tab:prolfquaModels) and \@ref(tab:describeBenchmarkedModels)), `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`.
+Table \@ref(tab:describeBenchmarkedModels) summarizes which methods we have evaluated, which MaxQuant results we used, and if the models are fitted to peptide or protein intensities.
+We make the R-markdown files to replicate the benchmarking available at `r BiocStyle::Githubpkg("wolski/prolfquabenchamrk")` and at [BenchmarkingIonstarData](https://wolski.github.io/prolfquabenchmark).
+
+```{r describeBenchmarkedModels}
+xMM <- data.frame(
+    msstats = c("MSstats", "preprocess with default parameters",  "", "evidence.txt"),
+    proDA = c("proDA", "probabilistic dropout model",  "protein", "peptide.txt"),
+    imputation = c("prot_imputation", "ContrastsSimpleImpute, ContrastsModerated", "protein",  "peptide.txt"),
+    lmmed = c("prot_med_lm_moderated", "strategy_lm, Contrasts, ContrastsModerated",  "protein", "peptide.txt"),
+    prot_merged = c("prot_merged", "addContrastResults( prefer = prot_med_lm_moderated , add = prot_imputation) *",  "protein", "peptide.txt"),
+    mixed = c("prot_mixed_effect_moderated", "strategy_lmer, Contrasts, ContrastsModerated", "peptide", "peptide.txt" ),
+    ropeca = c("ropeca", "strategy_lm, Contrasts, ContrastsModerated, ContrastsROPECA", "peptide", "peptide.txt" )
+)
+
+xMM <- data.frame(t(xMM))
+rownames(xMM) <- NULL
+colnames(xMM) <-  c("Label","Description","Abundance","Input File")
+kableExtra::kable(xMM,
+             caption = paste0("All benchmarked models.",
+             "Description - prolfqua function names",
+             "Abudances - indicates if model is fitted to peptide or protein abundances",
+             "Input File - MaxQuant file used as input to the analysis."),
+             format = "latex", booktabs=TRUE) |>
+    kableExtra::kable_styling(latex_options="scale_down") |> 
+    kableExtra::add_footnote(label = "* Merges results of the prot_med_lm_moderated and prot_imputation modeling pipeline, preferring those of prot_med_lm_moderated if available.")
+
+```
+
+Since only technical replicates are available for each dilution, essential sources of variation present in every experiment, such as the biochemical and biological, are not measured.
+Furthermore, the dataset captures only the variance from the chromatography, electro-spray, and mass spectrometric measurement method.
 Thus, while we can extrapolate some of the results obtained to more realistic datasets, we need to be careful not to over-interpret the results.
-Specifically, the observed variances will be higher in more realistic experiments, and sensitivity and specificity will worsen, given the same sample sizes.
+Specifically, the observed variances will be higher in more realistic experiments, and the power will be lower for the same sample sizes.
 Furthermore, the proportion of missing observations in real-life datasets might also be higher.
@@
 Modeling the degrees of freedom when computing the $p$-values might improve the inference.
-However, this improvement is minuscule (see Figure \@ref(fig:FDRfdp) panel A).
-There is no such effect for the mixed effect model, most likely because the degrees of freedom are erroneously estimated for many models.
-Therefore, for the fixed effect linear model, the empirical Bayes variance shrinkage, as suggested by [@Smyth2004linear], consistently improves the ranking of proteins compared with the unmoderated estimates (not shown) and fails to do so for mixed effect models.
+However, this improvement is minuscule (see Figure \@ref(fig:FDRfdp) panel B).
+There is no such improvement for the mixed effect model, most likely because the degrees of freedom are erroneously estimated for many models.
+Furthermore, for the fixed effect linear model, the empirical Bayes variance shrinkage, as suggested by [@Smyth2004linear], consistently improves the ranking of proteins compared with the unmoderated estimates (not shown) and fails to do so for mixed effect models.
 
@@
 
-We make the R-markdown files to replicate the benchmarking available at `r BiocStyle::Githubpkg("wolski/prolfquabenchamrk")` and at [BenchmarkingIonstarData](https://wolski.github.io/prolfquabenchmark).
+We focused our benchmark on comparing the statistical modeling methods using three different scores, while we fixed the pre-processing steps.
+However, there are other equally or even more important parameters of a protein quantification pipeline [@frohlich2022benchmarking].
+One of them is the normalization of the intensities within the samples to remove systematic differences [@pursiheimo2015optimization].
+The method used to infer protein intensities from peptide intensities is an additional important factor [@Grossmann2010].
+For instance, the original `r BiocStyle::Biocpkg("proDA")` publication uses MaxLFQ [@cox2014accurate] protein estimates.
+However, when using MaxLFQ intensities reported by MaxQuant, the $pAUC_{10}$ is significantly lower ($pAUC_{10}(\textrm{t-statistics}) = 66%$) compared with results obtained when protein abundances are estimated from peptide abundances using Tukey's median polish ($pAUC_{10}(\textrm{t-statistics}) = 73%$).
+Last but not least, the software [@Cox2008MaxQuant; @yu2020fast], identifying proteins and generating the quantification values can also significantly contribute to the performance of the entire pipeline, altering the number of identified proteins and the sensitivity and specificity of the differential expression analysis.
 
@@
 First, estimate protein abundances from peptide abundances using a robust or nonparametric regression method.
-Second, use linear models as long as they suffice to model your experimental design because they show good performance in all categories.
-Third, use mixed effect models only if you have to, e.g., repeated measurements and if the sample sizes are large.
-If you use fixed-effect linear models, apply variance moderation to improve the $t$-statistics and $p$-value estimates.
+Second, use linear models because they show good performance in all categories.
+If the measurements are correlated, as for technical replicates, mixed effect models might work if the sample sizes are large; if not, aggregate the replicates and fit a linear model instead.
+Third, if you use fixed-effect linear models, apply variance moderation to improve the $t$-statistics and $p$-value estimates.
 If you want to sort your protein lists to perform gene set enrichment analysis, use the $t$-statistic instead of the difference.
-Last but not least, do not impute missing data directly but model it when estimating parameters.
+Last but not least, do not impute missing observation but rather statistically model missingness to estimate parameters, i.e., group differences.
 
 In summary, `r BiocStyle::Githubpkg("fgcz/prolfqua")` is an easy-to-use and feature-rich R package to analyze quantitative mass spectrometric data, report results, and benchmark MS software and statistical methods.
-We provide more details at the website https://github.com/fgcz/prolfqua/.
-This document was created using Rmarkdown, and all the code need to replicate the document and the presented benchmarks is available at : https://github.com/wolski/prolfquabenchmark.
+We provide documentation, in vignette format, at the website https://github.com/fgcz/prolfqua/.
+This document was created using Rmarkdown.
+All the code needed to replicate the document or the benchmark results is available at : https://github.com/wolski/prolfquabenchmark.
 

### 2022-06-03 inst/MyArticle/paper.Rmd
Set keep_tex: true

--- 2022-06-01 inst/MyArticle/paper.Rmd
@@
       toc_bib: false
+      keep_tex: true
 bibliography: bibliography.bib

### 2022-06-03 inst/MyArticle/paper.Rmd

--- 2022-06-02 inst/MyArticle/paper.Rmd
@@
 Then, as an output, the user receives a report including input files, the R markdown file, and R scripts necessary to replicate the analysis using their in-house R installation.
+Under the hood of the b-fabric system runs a computing infrastructure controlled by a local resource management system that also supports cloud-bursting [@VMMAD].
 In this way, [b-fabric](https://fgcz-bfabric.uzh.ch/) helps scientists meet requirements from funding agencies, journals, and academic institutions to publish data according to the FAIR (Findable, Accessible, Interoperable and Reusable) [@FAIR] data principles.
@@
 
-# Acknoledgements
+# Acknoledgements {.unlisted .unnumbered}
 
@@
 
-# Abbreviations
+# Abbreviations {.unlisted .unnumbered}
 

### 2022-06-03 inst/MyArticle/paper.Rmd
more edits to implementation.

--- 2022-06-02 inst/MyArticle/paper.Rmd
@@
 
-We store all the data needed for analysis in a single data frame in a tidy table, i.e., every column is a variable, every row is an observation, and every cell is a single value [@tidydata2014].
-Using an __R6__[@R6cite] configuration object \@ref(fig:LFQData), we specify what variable is in which column, making it easy to integrate new inputs in `r BiocStyle::Githubpkg("fgcz/prolfqua")` if provided as a tidy table.
-For example, to visualize tidy _Spectronaut_, _DiaNN_, or _Skyline_ outputs, or data in `r BiocStyle::Biocpkg("MSstats")` format, only a few lines of code are needed to update the `r BiocStyle::Githubpkg("fgcz/prolfqua")` configuration.
-The configuration encapsulates the differences of column names among the various input formats and enables the methods of prolfqua to be transparent, i.e., it preserves the differences of the processing tools in the figures and analysis.
+We store all the data needed for analysis in a tidy table, i.e., every column is a variable, every row is an observation, and every cell is a single value [@tidydata2014].
+Using an __R6__[@R6cite] configuration object (Figure \@ref(fig:LFQData)), we specify what variable is in which column, making it easy to integrate new inputs in `r BiocStyle::Githubpkg("fgcz/prolfqua")` if provided as a tidy table.
+For example, to visualize tidy _Spectronaut_, _DiaNN_, or _Skyline_ outputs, or data in `r BiocStyle::Biocpkg("MSstats")` format, only a few lines of code are needed to update the `r BiocStyle::Githubpkg("fgcz/prolfqua")` `AnalysisTableConfiguration` configuration.
+The configuration encapsulates the differences in column names among the various input formats, enables the use of the methods without additional arguments, and preserves the processing tools' differences in the figures and analysis.
 We show an example code for creating an MSFragger configuration in the Supplement.
@@
 
-In addition we implemented features specific to high throughput experiments, such as the experimental Bayes variance and $p$-value moderation, which utilizes the parallel structure of the protein measurements and the analysis [@Ritchie2015].
+In addition, we implemented features specific to high throughput experiments, such as the experimental Bayes variance and $p$-value moderation, which utilizes the parallel structure of the protein measurements and the analysis [@Ritchie2015].
 We also compute probabilities of differential protein regulation based on peptide level models [@Suomi2017bEnhanced].
-We used R6 classes encapsulate the statistical modelling functionality in `r BiocStyle::Biocpkg("prolfqua")` (see Figure \@ref(fig:ContrastUML)).
-We did specify a contrast interface (ContrastInterface), and there are various implementations allowing to perform differential expression analysis given linear or mixed effect models (Contrast), variance shrinkage (ContrastModerated), or impute contrasts in case of observations missing for an entire group of samples (ContrastSimpleImpute) can also be used to encapsulate and integrate differential expression analysis results of external tools such as proDA or to analyse data from protein interaction studies using SAINTexpress [@teo2014saintexpress].
+We used R6 classes to encapsulate the statistical modelling functionality in `r BiocStyle::Githubpkg("fgcz/prolfqua")` (see Figure \@ref(fig:ContrastUML)).
+We did specify a contrast interface (`ContrastsInterface`).
+Several implementations allow to perform differential expression analysis given linear or mixed effect models (`Contrasts`), variance shrinkage (`ContrastsModerated`), or impute contrasts in cases when observations are missing for an entire group of samples (`ContrastsSimpleImpute`).
+Further implementations of the interface encapsulate and integrate differential expression analysis results of external tools such as `r `BiocStyle::Biocpkg("proDA")` or of SAINTexpress [@teo2014saintexpress] used to analyze data from protein interaction studies using.
 
 (ref:ContrastUML)  UML diagram of modeling and contrast related classes.
-Different strategies, e.g., _lm_, _lmer_, and _glm_ can be used to fit various types of models.
+Different strategies, e.g., _lm_, _lmer_, and _glm_, can fit various types of models.
 The model builder method fit's the statistical model given the data and a strategy.
-The obtained model can then be used for analysis of variance or to estimate contrasts.
-All classes estimating the contrasts implement the _Contrast_ interface.
-The variance of the contrasts can be moderated using the `ContrastModerated` class.
+The obtained model can then be used for the analysis of variance or to estimate contrasts.
+All classes estimating contrasts implement the _Contrasts_ interface.
+Results of external tools, e.g., SAINTExpress, can be adapted by implementing the Contrasts Interface.
 
@@
     kableExtra::kable_styling(latex_options="scale_down") |> 
-    kableExtra::add_footnote(label = "* in development")
+    kableExtra::add_footnote(label = "in development")
 
@@
     kableExtra::kable_styling(latex_options="scale_down") |> 
-    kableExtra::add_footnote(label = "* Merges results of the prot_med_lm_moderated and prot_imputation modeling pipeline, preferring those of prot_med_lm_moderated if available.")
+    kableExtra::add_footnote(label = "Merges results of the prot_med_lm_moderated` and prot_imputation modeling pipeline, preferring those of prot_med_lm_moderated if available.")
 
@@
 
-The probabilistic dropout analysis implemented in the `r BiocStyle::Biocpkg("proDA")` produces inferences comparable to those of other methods (Figure \@ref(fig:FDRfdp) Panel A) for the largest number of proteins (Figure \@ref(fig:FDRfdp) Panel B).
-Moreover, the estimated differences are unbiased (not shown) and show high diagnostic accuracy (\@ref(fig:FDRfdp) Panel A).
-Furthermore, the performance of the scaled $p$-values or the $t$-statistics is comparable with that of the linear model with variance moderation.
-Because of the robustness of the dropout model to missing observations, we intend to integrate the `r BiocStyle::Biocpkg("proDA")` package as an additional modeling option into `r BiocStyle::Githubpkg("fgcz/prolfqua")` (see Figure \@ref(fig:ContrastUML)).
+The probabilistic dropout analysis implemented in the `r BiocStyle::Biocpkg("proDA")` produces inferences comparable to those of other methods (Figure \@ref(fig:FDRfdp) Panel A).
+Because of the robustness of the dropout model to missing observations we obtain difference estimates for all proteins and contrasts (Figure \@ref(fig:FDRfdp) Panel B).
+Moreover, the estimated differences are unbiased and show high diagnostic accuracy (\@ref(fig:FDRfdp) Panel A).
+Furthermore, the performance of the scaled $p$-values or the $t$-statistics is comparable with that of the linear model with variance moderation (`prot_med_lm_moderated`).
+Therefore, we started to integrate the `r BiocStyle::Biocpkg("proDA")` package as an additional modeling option into `r BiocStyle::Githubpkg("fgcz/prolfqua")` (see Figure \@ref(fig:ContrastUML)).
 
@@
 `r BiocStyle::Githubpkg("fgcz/prolfqua")` allows for considerable flexibility to model your experiments.
-Various types of models are available (see Figure \@ref(fig:ContrastUML) and Table \@ref(tab:prolfquaModels)), and the contrast specification is consistent and intuitive for all models.
-The modular design of `r BiocStyle::Githubpkg("fgcz/prolfqua")`, allows for adding new features, e.g., generalized linear models (_glm_'s) for missing data modeling or robust linear models (_rlm_'s), in the future.
+Various types of models are available (see Figure \@ref(fig:ContrastUML) and Table \@ref(tab:prolfquaModels)), and the contrast specification is explicit and consistent for all models.
+The modular design of `r BiocStyle::Githubpkg("fgcz/prolfqua")`, allows for adding new features, e.g., generalized linear models (_glm_'s) to model the presence or absence information of a protein, or robust linear models (_rlm_'s), in the future.
 R's formula interface for linear models is flexible, widely used, and well documented [@faraway2016extending].
-We use the formula interface to specify the models which should make it easy to reproduce an analysis performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` in other statistical programming languages.
+We use the formula interface to specify the models, making it easy to reproduce an analysis performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` in other statistical methods programming languages.
+The developed framework is flexible to integrate other modeling methods, e.g., the probabilistic dropout model [@bioxrvproDA2020] or accurate variance estimation [@zhu2020deqms].
 

### 2022-06-03 inst/MyArticle/paper.Rmd

@@
 Then, as an output, the user receives a report including input files, the R markdown file, and R scripts necessary to replicate the analysis using their in-house R installation.
+The b-fabric system runs a computing infrastructure controlled by a local resource management system that supports cloud-bursting [@VMMAD].
 In this way, [b-fabric](https://fgcz-bfabric.uzh.ch/) helps scientists meet requirements from funding agencies, journals, and academic institutions to publish data according to the FAIR (Findable, Accessible, Interoperable and Reusable) [@FAIR] data principles.

### 2022-06-03 inst/MyArticle/paper.Rmd
Add date

@@
       keep_tex: true
+date: "`r format(Sys.time())`"
 bibliography: bibliography.bib

### 2022-06-03 inst/MyArticle/paper.Rmd
fix references by PN.

--- 2022-06-03 inst/MyArticle/paper.Rmd
@@
       toc_bib: false
-      keep_tex: true
 bibliography: bibliography.bib
@@
 abstract: |
-    Protein mass spectrometry is widely used for quantitative proteomics studies, relative protein quantification, and differential expression analysis of proteins.
+    Mass spectrometry is widely used for quantitative proteomics studies, relative protein quantification, and differential expression analysis of proteins.
 Nevertheless, there is a need for a flexible and easy-to-use application programming interface in R that transparently supports a variety of well principled statistical procedures.
@@
 An important aspect of mass spectrometric data are missing peptide and protein quantifications.
-[@rubin1976inference] classified missing data problems into three categories: missing completely at random (MCAR), missing at random (MAR), and missing not at random (MNAR).
+@rubin1976inference classified missing data problems into three categories: missing completely at random (MCAR), missing at random (MAR), and missing not at random (MNAR).
 For instance, in data-dependent acquisition (DDA) MS, only a limited number of MS1 signals are selected for fragmentation and identified.
@@
 `r BiocStyle::Biocpkg("MSstats")` handles missing data by feature filtering and using imputation.
-Other means of modelling missing observations are the Hurdle model discussed by [@goeminne2020msqrob], while the R package `r BiocStyle::Biocpkg("proDA")` models missingness using probabilistic dropout models [@bioxrvproDA2020].
+Other means of modelling missing observations are the Hurdle model discussed by @goeminne2020msqrob, while the R package `r BiocStyle::Biocpkg("proDA")` models missingness using probabilistic dropout models [@bioxrvproDA2020].
 
@@
 
-Valikangas [@valikangas2016systematic] and colleagues discuss and benchmark various methods of peptide or protein intensity normalization, such as variance stabilizing normalization [@huber2002variance] or quantile normalization [@bolstad2003comparison].
+@valikangas2016systematic discuss and benchmark various methods of peptide or protein intensity normalization, such as variance stabilizing normalization [@huber2002variance] or quantile normalization [@bolstad2003comparison].
 In this work, we will be using a robust version of the z-score, where instead of the mean $\bar{x}$ we use the median, and instead of the standard deviation $\tilde{S}$ the median absolute deviation (mad):
@@
 From the linear and the mixed effect models, we can obtain the residual standard deviation $\sigma$, and degrees of freedom $df$.
-[@Smyth2004linear] discuss how, to use the $\sigma$ and $df$ of all models to estimate a prior $\sigma$ and prior $df$, and posterior $\tilde \sigma$.
+@Smyth2004linear discuss how, to use the $\sigma$ and $df$ of all models to estimate a prior $\sigma$ and prior $df$, and posterior $\tilde \sigma$.
 These can be used to moderate the $t$-statistics by:

### 2022-06-03 inst/MyArticle/paper.Rmd
NOP

@@
             vjust = 0, hjust = -0.2, angle = -90) #+ 
-  #ggplot2::geom_text(ggplot2::aes(label = paste0("(",round(percent, digits = 1),"%)")),
-            #vjust = -1, hjust = -0.2, angle = -90) 
 pB <- nrgg + ggplot2::labs(tag = "B")

### 2022-06-07 inst/MyArticle/paper.Rmd

@@
       toc_bib: false
+      keep_tex: true
+date: "`r format(Sys.time())`"
 bibliography: bibliography.bib

### 2022-06-07 inst/MyArticle/paper.Rmd
set R code snippet width=70

@@
                       fig.height = 5)
+
+options(width=70)
 ```
@@
 
-\newpage
 ## Miscellaneous {.unlisted .unnumbered}

### 2022-06-07 inst/MyArticle/paper.Rmd
more references. Changes from Paolo.

--- 2022-06-03 inst/MyArticle/paper.Rmd
@@
 Finally, we show that the implemented methods allow sensitive and specific differential expression analysis.
-     The `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package is available on GitHub https://github.com/fgcz/prolfqua, distributed under the MIT licence, and runs on all plattforms supported by the [R free software environment for statistical computing and graphics](https://www.r-project.org/).
+     The `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package is available on GitHub https://github.com/fgcz/prolfqua, distributed under the MIT licence, and runs on all platforms supported by the [R free software environment for statistical computing and graphics](https://www.r-project.org/).
 urlcolor: blue
@@
 Hence, measuring changes in protein abundance is the subject of active research [@vidova2017review].
-Bottom-up mass spectrometric methods, where proteins are specifically and reproducible digested into protein fragments - peptides, are employed to identify and quantify proteins in complex samples containing hundreds to thousand of proteins, [@Bubis2017; @da2020philosopher].
-The peptides are first separated by their chemical and physical properties using High-Pressure Liquid Chromatography (HPLC).
-Afterward, they are ionised, weighted, identified, and quantified using mass spectrometric techniques, e.g., Electro-Spray-Ionization Mass Spectrometry (ESI-MS).
+Bottom-up mass spectrometric methods, where proteins are specifically and reproducibly digested into protein fragments - peptides, are employed to identify and quantify proteins in complex samples containing hundreds to thousand of proteins [@Bubis2017; @da2020philosopher].
+The peptides are first separated by their chemical and physical properties using liquid chromatography (LC).
+Afterward, they are ionised, weighed, identified, and quantified using mass spectrometric techniques, e.g., electro-spray-ionization mass spectrometry (ESI-MS).
 Finally, peptide identification is achieved by fragmenting and matching the measured fragment masses to theoretical masses computed from known peptide sequences [@eng2015deeper; @yu2016pipi; @kong2017msfragger].
@@
 Peptide quantification algorithms transfer identification information between MS1 features in different samples by aligning the data using retention time and mass information, thus reducing the amount of missing data.
-Nevertheless, highly abundant proteins can suppress the detection of other proteins in a sample, an MAR mechanism.
-Furthermore, a weak correlation between the number of missing measurements in a condition and the abundance of the quantified protein can be observed, which is caused by the limit of detection (LOD), an MNAR mechanism [@mcgurk2020use].
+Nevertheless, highly abundant proteins can suppress the detection of other proteins in a sample, a MAR mechanism.
+Furthermore, a weak correlation between the number of missing measurements in a condition and the abundance of the quantified protein can be observed, which is caused by the limit of detection (LOD), a MNAR mechanism [@mcgurk2020use].
 
-Several packages exist to model mass spectrometry protein quantification experiments, e.g., `r BiocStyle::Biocpkg("limma")` [@Ritchie2015], `r BiocStyle::Biocpkg("MSstats")` [@Choi2014], `r BiocStyle::Biocpkg("PECA")` [@Suomi2017bEnhanced],  `r BiocStyle::Biocpkg("msqrob2")` [@Goeminne2016] or `r BiocStyle::Biocpkg("proDA")` [@bioxrvproDA2020], to mention some, all implemented in the R.
+Several data analysis packages exist to model mass spectrometry protein quantification experiments, e.g., `r BiocStyle::Biocpkg("limma")` [@Ritchie2015], `r BiocStyle::Biocpkg("MSstats")` [@Choi2014], `r BiocStyle::Biocpkg("PECA")` [@Suomi2017bEnhanced],  `r BiocStyle::Biocpkg("msqrob2")` [@Goeminne2016] or `r BiocStyle::Biocpkg("proDA")` [@bioxrvproDA2020], to mention some, all implemented in R.
 Each of them has some unique features, for example, `r BiocStyle::Biocpkg("MSstats")` infers the model and generates the model formula from the sample annotation, allowing users with limited statistical knowledge to perform differential expression analysis (DEA).
@@
 The package `r BiocStyle::Biocpkg("PECA")`, performs a roll-up of peptide level differences and peptide level $p$-value estimates, obtained from `r BiocStyle::Biocpkg("limma")` or `r BiocStyle::Biocpkg("PECA")`, to protein level estimates.
-Furthermore, `r BiocStyle::Biocpkg("msqrob2")` implements two models: robust linear models fitted to protein intensities and robust ridge regression fitted to peptide intensities and combines them with empirical Bayes variance shrinkage.
+Furthermore, `r BiocStyle::Biocpkg("msqrob2")` implements two models: robust linear models fitted to protein intensities and robust ridge regression fitted to peptide intensities, and combines them with empirical Bayes variance shrinkage.
 The `r BiocStyle::Biocpkg("proDA")` package implements a linear probabilistic dropout model to account for missing data without imputation.
@@
 
-When analyzing parallel-group designs using a single explanatory variable, and contrasting groups, we can use all R packages; however, we can use only some of them if we want to model factorial designs or repeated measurements.
+When analyzing parallel-group designs using a single explanatory variable, and contrasting groups, we can use all R packages; but, we can use only some of them if we want to model factorial designs or repeated measurements.
 Table \@ref(tab:tableOverview) gives an overview of the models and features supported by these packages.
@@
 
-When developing the R package `r BiocStyle::Githubpkg("fgcz/prolfqua")` we were inspired by the R package `r BiocStyle::CRANpkg("caret")` [@caret2008], which enables us to call various machine learning (ML) methods, which makes selecting the best ML algorithm for your problem easy.
+When developing the R package `r BiocStyle::Githubpkg("fgcz/prolfqua")` we were inspired by the R package `r BiocStyle::CRANpkg("caret")` [@caret2008], which enables us to call various machine learning (ML) methods, and makes selecting the best ML algorithm for your problem easy.
 We aimed for a similar R package for the differential expression analysis of proteomics data.
-However, after examining the R packages for modeling proteomics quantification data, we observed that model specification, input, and output formats differ widely (and wildly).
-However, these packages have in common: they fit linear models either to peptide or protein intensities, determine contrasts among conditions, and afterward, apply empirical Bayes variance shrinkage.
+However, after examining the R packages for modeling proteomics quantification data, we observed that model specification, input, and output formats differ widely.
+At the same time, they have in common that: they fit linear models either to peptide or protein intensities, determine contrasts among conditions, and afterward apply empirical Bayes variance shrinkage.
 Therefore, the revised goal was to provide a modular object-oriented (OO) design, with R linear models as a core, where we can add features such as $p$-value aggregation, e.g., `r BiocStyle::Biocpkg("PECA")`, variance shrinkage, or modeling of missing observations.
@@
 Furthermore, the functionality of `r BiocStyle::Githubpkg("fgcz/prolfqua")` also includes methods specific to proteomics data.
-For example, we implemented strategies to estimate protein intensities from peptide intensities: top N [@Grossmann2010], Tukey's median polish, robust linear models [@goeminne2020msqrob].
-Furthermore, peptide or protein abundances can be scaled and transformed, using robust scaling, transformed using _quantile_ normalization or `r BiocStyle::Biocpkg("vsa")` to remove systematic differences among samples and heteroscedasticity.
+For example, we implemented strategies to estimate protein intensities from peptide intensities: top N [@Grossmann2010], Tukey's median polish  [@tukey1977exploratory], robust linear models [@goeminne2020msqrob].
+Furthermore, peptide or protein abundances can be scaled and transformed, using robust scaling, _quantile_ normalization or `r BiocStyle::Biocpkg("vsn")` to remove systematic differences among samples and heteroscedasticity.
 In this respect, `r BiocStyle::Githubpkg("fgcz/prolfqua")` can be compared with other R packages such as `r BiocStyle::Biocpkg("DEP")` [@DEP2018] or `r BiocStyle::Biocpkg("POMA")` [@POMA2021] which support the entire differential expression analysis pipeline.
 
-We also implemented functionality to streamline the benchmarking of the implemented methods.
-We use the Ionstar [@shen2018ionstar] dataset to benchmark the modeling methods implemented within `r BiocStyle::Githubpkg("fgcz/prolfqua")` and to compare our results with those of `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`.
+We also implemented functionality and use the Ionstar [@shen2018ionstar] dataset to benchmark the modeling methods implemented within `r BiocStyle::Githubpkg("fgcz/prolfqua")` and to compare our results with those of `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`.
 Since group sizes are relatively small, typically with four or five subjects per group, the high power of the tests is a relevant criterion to assess the modeling method.
-Furthermore, the quantified proteins can be ranked using the estimated fold-change, $t$-statistics, or scaled $p$-value, and then subjected to gene set enrichment (GSEA) or over-representation analysis [@subramanian2005gene] to determine up or down-regulated groups of proteins.
+The quantified proteins can be ranked using the estimated fold-change, $t$-statistics, or scaled $p$-value, and afterward subjected to gene set enrichment (GSEA) or over-representation analysis [@subramanian2005gene] to determine up or down-regulated groups of proteins.
 Furthermore, the statistical model must provide an unbiased estimate of the false discovery rate (FDR) to manage expectations when selecting protein lists for follow-up experiments.
@@
 We store all the data needed for analysis in a tidy table, i.e., every column is a variable, every row is an observation, and every cell is a single value [@tidydata2014].
-Using an __R6__[@R6cite] configuration object (Figure \@ref(fig:LFQData)), we specify what variable is in which column, making it easy to integrate new inputs in `r BiocStyle::Githubpkg("fgcz/prolfqua")` if provided as a tidy table.
-For example, to visualize tidy _Spectronaut_, _DiaNN_, or _Skyline_ outputs, or data in `r BiocStyle::Biocpkg("MSstats")` format, only a few lines of code are needed to update the `r BiocStyle::Githubpkg("fgcz/prolfqua")` `AnalysisTableConfiguration` configuration.
-The configuration encapsulates the differences in column names among the various input formats, enables the use of the methods without additional arguments, and preserves the processing tools' differences in the figures and analysis.
-We show an example code for creating an MSFragger configuration in the Supplement.
-For popular software like _MaxQuant_, or_MSFragger_, which stores the same variable, e.g., intensity, in multiple columns, one for each sample, we implemented methods that transform the data into tidy tables.
-Furthermore, relying on the tidy data table enables us to interface with many data manipulation, visualization, and modeling methods, implemented in base R [@RCoreTeam2021] and the tidyverse [@tidyverse2019], easily.
-We use __R6__ classes to structure the functionality of the package (see Figure \@ref(fig:LFQData) and Figure \@ref(fig:ContrastUML)).
-__R6__ classes are well supported, e.g., by command-line completion features, see Figure \@ref(fig:tabCompletion), and easy to use (OOP).
+Using an R6 [@R6cite] configuration object (Figure \@ref(fig:LFQData)), we specify what variable is in which column, making it easy to integrate new inputs in `r BiocStyle::Githubpkg("fgcz/prolfqua")` if provided as a tidy table.
+For example, to visualize tidy Spectronaut[@bruderer2015extending], DiaNN[@demichev2020dia], or Skyline[@maclean2010skyline] outputs, or data in `r BiocStyle::Biocpkg("MSstats")`[@Choi2014] format, only a few lines of code are needed to update the `r BiocStyle::Githubpkg("fgcz/prolfqua")` `AnalysisTableConfiguration` configuration.
+The configuration encapsulates the differences in column names among the various input formats, enables the use of the methods without additional arguments, and enables to preserve the differences in the analysis and figures.
+We show an example code for creating an MSFragger[@yu2020fast] configuration in the Appendix.
+For popular software like MaxQuant[@Cox2008MaxQuant], or MSFragger, which stores the same variable, e.g., intensity, in multiple columns, one for each sample, we implemented methods that transform the data into tidy tables.
+Relying on the tidy data table enables us to interface with many data manipulation, visualization, and modeling methods, implemented in base R [@RCoreTeam2021] and the tidyverse [@tidyverse2019], easily.
+We use R6 classes to structure the functionality of the package (see Figure \@ref(fig:LFQData) and Figure \@ref(fig:ContrastUML)).
+R6 classes are well supported by command-line completion features in RStudio [@RStudio2022], and help to implement argument free functions (Appendix Figure \@ref(fig:tabCompletion).
 
@@
 The LFQDataPlotter class and other classes decorate the LFQData class with additional functionality.
-Similarly, the `LFQDataStats` and`LFQDataSummary`, reference the LFQData class and group methods for variance and sample size estimation or summarizing peptide and protein counts.
+Similarly, the `LFQDataStats` and `LFQDataSummary`, reference the `LFQData` class and group methods for variance and sample size estimation or summarizing peptide and protein counts.
 
@@
 
-In addition, we implemented features specific to high throughput experiments, such as the experimental Bayes variance and $p$-value moderation, which utilizes the parallel structure of the protein measurements and the analysis [@Ritchie2015].
+In addition, we implement features specific to high throughput experiments, such as the experimental Bayes variance and $p$-value moderation, which utilizes the parallel structure of the protein measurements and the analysis [@Ritchie2015].
 We also compute probabilities of differential protein regulation based on peptide level models [@Suomi2017bEnhanced].
 We used R6 classes to encapsulate the statistical modelling functionality in `r BiocStyle::Githubpkg("fgcz/prolfqua")` (see Figure \@ref(fig:ContrastUML)).
-We did specify a contrast interface (`ContrastsInterface`).
+We specify a contrast interface (`ContrastsInterface`).
 Several implementations allow to perform differential expression analysis given linear or mixed effect models (`Contrasts`), variance shrinkage (`ContrastsModerated`), or impute contrasts in cases when observations are missing for an entire group of samples (`ContrastsSimpleImpute`).
-Further implementations of the interface encapsulate and integrate differential expression analysis results of external tools such as `r `BiocStyle::Biocpkg("proDA")` or of SAINTexpress [@teo2014saintexpress] used to analyze data from protein interaction studies using.
+Further implementations of the interface encapsulate and integrate differential expression analysis results of external tools such as `r BiocStyle::Biocpkg("proDA")` or of SAINTexpress [@teo2014saintexpress] used to analyze data from protein interaction studies.
 
-(ref:ContrastUML)  UML diagram of modeling and contrast related classes.
-Different strategies, e.g., _lm_, _lmer_, and _glm_, can fit various types of models.
-The model builder method fit's the statistical model given the data and a strategy.
+(ref:ContrastUML)  The UML (unified modeling language) diagram of modeling and contrast related classes.
+Different strategies, e.g., _lm_, _lmer_, and _glm_ (Table \@ref(tab:prolfquaModels)), can fit various types of models.
+The model builder method fits the statistical model given the data and a strategy.
 The obtained model can then be used for the analysis of variance or to estimate contrasts.
 All classes estimating contrasts implement the _Contrasts_ interface.
-Results of external tools, e.g., SAINTExpress, can be adapted by implementing the Contrasts Interface.
+Results of external tools, e.g., SAINTexpress, can be adapted by implementing the Contrasts Interface.
 
@@
 
-## Dataset for Benchmarking
+## Dataset for benchmarking
 
-To evaluate the performance of differential expression analysis, we use the _IonStar_ benchmark dataset[@shen2018ionstar], available from the Proteomics Identifications Database (PRIDE) identifier PXD003881.
+To evaluate the performance of differential expression analysis, we use the IonStar benchmark dataset[@shen2018ionstar], available from the Proteomics Identifications Database (PRIDE) identifier PXD003881.
 $DH5\alpha$ _E.
-coli_ lysate was spiked in human pancreatic cancer cells(Panc-1) lysate at five different levels: $3\%$ _E.
+coli_ lysate was spiked in human pancreatic cancer cells (Panc-1) lysate at five different levels: $3\%$ _E.
 coli_, $4.5\%$ _E.
@@
 coli_.
-We annotated these dilutions from smallest to largest with the letters a, b, c, d, e.
+We annotated these dilutions from smallest to largest with the letters _a_, _b_, _c_, _d_, _e_.
 By comparing the various dilutions, we obtain different effect sizes, e.g., when comparing dilution _e_ (9%) against dilution _d_ (7.5%), the expected difference is $1.2$ for E.
@@
 
-The peptide abundances (from the _MaxQuant_ _peptide.txt_ file), were $log_2$ transformed and subsequently scaled, where median and the mean average deviation (mad) were obtained from the human proteins only.
+The peptide abundances (from the MaxQuant _peptide.txt_ file), were $log_2$ transformed and subsequently scaled, where median and the mean average deviation (mad) were obtained from the human proteins only.
 We removed _one hit wonders_, i.e., proteins with a single peptide assignment.
-Protein abundances are inferred from the peptide intensities using Tukey's median polish [@tukey1977exploratory].
+Protein abundances are inferred from the peptide intensities using Tukey's median polish.
 Finally, we fitted the fixed effect linear models, the dropout model `r BiocStyle::Biocpkg("proDA")` to protein abundances, the mixed effect linear model, and the `r BiocStyle::Biocpkg("PECA")` model to peptide intensities.
@@
 
-The IonStar [@shen2018ionstar] dataset contains _H.
+The IonStar dataset contains _H.
 sapiens_ proteins with constant concentrations and _E.
@@
 
-knitr::kable(table, caption = "Confusion matrix, TP - true positive, FP - false positive, FN - false negative, P - all positive cases (all E.
-coli proteins), N - all negative cases (all H.
-sapiens proteins), m- all proteins.")
+knitr::kable(table, caption = "Confusion matrix, TP - true positive, FP - false positive, FN - false negative, TN - true negatives, P - all positive cases (all $E.~coli$ proteins), N - all negative cases (all $H.~sapiens$ proteins), m - all proteins.")
 ```
@@
 @valikangas2016systematic discuss and benchmark various methods of peptide or protein intensity normalization, such as variance stabilizing normalization [@huber2002variance] or quantile normalization [@bolstad2003comparison].
-In this work, we will be using a robust version of the z-score, where instead of the mean $\bar{x}$ we use the median, and instead of the standard deviation $\tilde{S}$ the median absolute deviation (mad):
+In this work, we use a robust version of the z-score, where instead of the mean $\bar{x}$ we use the median, and instead of the standard deviation $\tilde{S}$ the median absolute deviation (mad):
 
@@
 
-Given a linear model $y = \beta X$, we can compute the difference between two groups $\beta_{c}$ by a linear combination $c$ of linear model parameters $\beta$.
-$c$ is a column vector with as many elements as there are coefficients in the linear model.
+Given a linear model $y = \beta X$, we can compute the difference between two groups $\beta_{c}$ by a linear combination $c$ of linear model parameters $\beta$, where $c$ is a column vector with as many elements as there are coefficients $\beta$ in the linear model.
 If $c$ has $0$ for one or more of its rows, then the corresponding coefficient in $\beta$ is not involved in determining the contrast [@ph525xseries].
@@
 The degrees of freedom for the contrast are equal to the residual degrees of freedom of the linear model.
-For estimating contrasts from mixed effects models we used the function `contest` implemented in the R package `r BiocStyle::CRANpkg("lmerTest")` [@Kuznetsova2017lmerTest] and used the Satterthwaite method to estimate the denominator degrees of freedom.
+For estimating contrasts from mixed effects models we used the function `contest` implemented in the R package `r BiocStyle::CRANpkg("lmerTest")`  and used the Satterthwaite [@Kuznetsova2017lmerTest] method to estimate the denominator degrees of freedom.
 These methods are available in the class `Contrast` (see Figure \@ref(fig:ContrastUML))
@@
 
-To summarize peptide level models to protein models, we applied the method suggested by [@Suomi2017bEnhanced] to use the median scaled $p$-value of the peptide models and cumulative distribution function of the beta distribution function to determine a regulation probability of the protein.
+To summarize peptide level models to protein models, we applied the method suggested by @Suomi2017bEnhanced to use the median scaled $p$-value of the peptide models and the cumulative distribution function of the beta distribution function to determine a regulation probability of the protein.
 
@@
 Using the `LFQDataPlotter` class, we show the distribution of the normalized peptide abundances in Figure \@ref(fig:prepro) Panel A.
-Afterwards, protein intensities are estimated from peptide intensities using Tukey's median polish [@tukey1977exploratory].
+Afterwards, protein intensities are estimated from peptide intensities using Tukey's median polish.
 Figure \@ref(fig:prepro) Panel B shows the peptide intensities and the estimated protein intensities.
@@
 For each sample a line with a different colour is shown.
-Panel B - Peptide intensities for protein _5NTD_ are shown using a line of different colour, and the protein intensity estimate is shown using a fat black line, Panel C - distribution of standard deviations of all proteins in each dilution group (a, b, c, d, e) and overall (all), Panel D - Distribution of protein intensities for protein _5NTD_.
+Panel B - Peptide intensities for protein _5NTD_ are shown using a line of different colour, and the protein intensity estimate is shown using a fat black line, Panel C - distribution of standard deviations of all proteins in each dilution group ($a$, $b$, $c$, $d$, $e$) and overall (all), Panel D - Distribution of protein intensities for protein _5NTD_.
 
@@
 Then we create the plots shown in Figure \@ref(fig:exampleContrasts).
-Panel A shows the distribution of the $p$-values, while figure Panel B shows the volcano plot for each comparison.
+Panel A shows the distribution of the $p$-values, while Panel B shows the volcano plot for each comparison.
 
@@
 Using the tidy table to model the data ensures interoperability with other proteomics-related packages that manage their data with tidy-tables, e.g., `r BiocStyle::CRANpkg("protti")` [@quast2022protti].
-Furthermore, the use of `R6` classes, which encapsulate the configuration and the data, allows for writing very concise code (no function arguments needed).
+The use of R6 classes, which encapsulate the configuration and the data, allows for writing very concise code (no function arguments needed).
 Auto-completion support in the Rstudio editor makes it easy for novices to explore `r BiocStyle::Githubpkg("fgcz/prolfqua")`s functionality (Figure \@ref(fig:tabCompletion)).
-To simplify integration of `r BiocStyle::Githubpkg("fgcz/prolfqua")` with Bioconductor based workflows there is an method to convert the LFQData class into a `r BiocStyle::Biocpkg("SummarizedExperiment")`.
+To simplify integration of `r BiocStyle::Githubpkg("fgcz/prolfqua")` with Bioconductor based workflows there is a method to convert the LFQData class into a `r BiocStyle::Biocpkg("SummarizedExperiment")`.
 
-To ease the usage barriers of the R package to users not being familiar with statistics and R programming, we integrated some workflows into our data management platform [b-fabric](https://fgcz-bfabric.uzh.ch/) @bfabric.
+To ease the usage barriers of the R package to users not being familiar with statistics and R programming, we integrated some workflows into our data management platform [b-fabric](https://fgcz-bfabric.uzh.ch/) [@bfabric].
 This integration enables our users to select the input and basic settings in a graphical user interface.
@@
 
-# Acknoledgements {.unlisted .unnumbered}
+# Acknowledgements {.unlisted .unnumbered}
 
@@
 $FDR$  &False Discovery Rate                               \\
-HPLC   &High-Pressure Liquid Chromatography                \\
+LC     &Liquid Chromatography                              \\
 LC-MS  &Liquid Chromatography followed by Mass Spectrometry\\

### 2022-06-07 inst/MyArticle/paper.Rmd
temp

@@
 For example, to visualize tidy Spectronaut[@bruderer2015extending], DiaNN[@demichev2020dia], or Skyline[@maclean2010skyline] outputs, or data in `r BiocStyle::Biocpkg("MSstats")`[@Choi2014] format, only a few lines of code are needed to update the `r BiocStyle::Githubpkg("fgcz/prolfqua")` `AnalysisTableConfiguration` configuration.
-The configuration encapsulates the differences in column names among the various input formats, enables the use of the methods without additional arguments, and enables to preserve the differences in the analysis and figures.
+The configuration encapsulates the differences in column names among the various input formats, and enables the use of the methods without additional arguments.
 We show an example code for creating an MSFragger[@yu2020fast] configuration in the Appendix.

### 2022-06-07 inst/MyArticle/paper.Rmd
Update paper.Rmd

@@
 Finally, we show that the implemented methods allow sensitive and specific differential expression analysis.
-     The `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package is available on GitHub https://github.com/fgcz/prolfqua, distributed under the MIT licence, and runs on all platforms supported by the [R free software environment for statistical computing and graphics](https://www.r-project.org/).
+     The `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package is available on GitHub https://github.com/fgcz/prolfqua, distributed under the MIT licence, and runs on all plattforms supported by the [R free software environment for statistical computing and graphics](https://www.r-project.org/).
 urlcolor: blue
@@
 Hence, measuring changes in protein abundance is the subject of active research [@vidova2017review].
-Bottom-up mass spectrometric methods, where proteins are specifically and reproducibly digested into protein fragments - peptides, are employed to identify and quantify proteins in complex samples containing hundreds to thousand of proteins [@Bubis2017; @da2020philosopher].
-The peptides are first separated by their chemical and physical properties using liquid chromatography (LC).
-Afterward, they are ionised, weighed, identified, and quantified using mass spectrometric techniques, e.g., electro-spray-ionization mass spectrometry (ESI-MS).
+Bottom-up mass spectrometric methods, where proteins are specifically and reproducible digested into protein fragments - peptides, are employed to identify and quantify proteins in complex samples containing hundreds to thousand of proteins, [@Bubis2017; @da2020philosopher].
+The peptides are first separated by their chemical and physical properties using High-Pressure Liquid Chromatography (HPLC).
+Afterward, they are ionised, weighted, identified, and quantified using mass spectrometric techniques, e.g., Electro-Spray-Ionization Mass Spectrometry (ESI-MS).
 Finally, peptide identification is achieved by fragmenting and matching the measured fragment masses to theoretical masses computed from known peptide sequences [@eng2015deeper; @yu2016pipi; @kong2017msfragger].
@@
 Peptide quantification algorithms transfer identification information between MS1 features in different samples by aligning the data using retention time and mass information, thus reducing the amount of missing data.
-Nevertheless, highly abundant proteins can suppress the detection of other proteins in a sample, a MAR mechanism.
-Furthermore, a weak correlation between the number of missing measurements in a condition and the abundance of the quantified protein can be observed, which is caused by the limit of detection (LOD), a MNAR mechanism [@mcgurk2020use].
+Nevertheless, highly abundant proteins can suppress the detection of other proteins in a sample, an MAR mechanism.
+Furthermore, a weak correlation between the number of missing measurements in a condition and the abundance of the quantified protein can be observed, which is caused by the limit of detection (LOD), an MNAR mechanism [@mcgurk2020use].
 
-Several data analysis packages exist to model mass spectrometry protein quantification experiments, e.g., `r BiocStyle::Biocpkg("limma")` [@Ritchie2015], `r BiocStyle::Biocpkg("MSstats")` [@Choi2014], `r BiocStyle::Biocpkg("PECA")` [@Suomi2017bEnhanced],  `r BiocStyle::Biocpkg("msqrob2")` [@Goeminne2016] or `r BiocStyle::Biocpkg("proDA")` [@bioxrvproDA2020], to mention some, all implemented in R.
+Several packages exist to model mass spectrometry protein quantification experiments, e.g., `r BiocStyle::Biocpkg("limma")` [@Ritchie2015], `r BiocStyle::Biocpkg("MSstats")` [@Choi2014], `r BiocStyle::Biocpkg("PECA")` [@Suomi2017bEnhanced],  `r BiocStyle::Biocpkg("msqrob2")` [@Goeminne2016] or `r BiocStyle::Biocpkg("proDA")` [@bioxrvproDA2020], to mention some, all implemented in the R.
 Each of them has some unique features, for example, `r BiocStyle::Biocpkg("MSstats")` infers the model and generates the model formula from the sample annotation, allowing users with limited statistical knowledge to perform differential expression analysis (DEA).
@@
 The package `r BiocStyle::Biocpkg("PECA")`, performs a roll-up of peptide level differences and peptide level $p$-value estimates, obtained from `r BiocStyle::Biocpkg("limma")` or `r BiocStyle::Biocpkg("PECA")`, to protein level estimates.
-Furthermore, `r BiocStyle::Biocpkg("msqrob2")` implements two models: robust linear models fitted to protein intensities and robust ridge regression fitted to peptide intensities, and combines them with empirical Bayes variance shrinkage.
+Furthermore, `r BiocStyle::Biocpkg("msqrob2")` implements two models: robust linear models fitted to protein intensities and robust ridge regression fitted to peptide intensities and combines them with empirical Bayes variance shrinkage.
 The `r BiocStyle::Biocpkg("proDA")` package implements a linear probabilistic dropout model to account for missing data without imputation.
@@
 
-When analyzing parallel-group designs using a single explanatory variable, and contrasting groups, we can use all R packages; but, we can use only some of them if we want to model factorial designs or repeated measurements.
+When analyzing parallel-group designs using a single explanatory variable, and contrasting groups, we can use all R packages; however, we can use only some of them if we want to model factorial designs or repeated measurements.
 Table \@ref(tab:tableOverview) gives an overview of the models and features supported by these packages.
@@
 
-When developing the R package `r BiocStyle::Githubpkg("fgcz/prolfqua")` we were inspired by the R package `r BiocStyle::CRANpkg("caret")` [@caret2008], which enables us to call various machine learning (ML) methods, and makes selecting the best ML algorithm for your problem easy.
+When developing the R package `r BiocStyle::Githubpkg("fgcz/prolfqua")` we were inspired by the R package `r BiocStyle::CRANpkg("caret")` [@caret2008], which enables us to call various machine learning (ML) methods, which makes selecting the best ML algorithm for your problem easy.
 We aimed for a similar R package for the differential expression analysis of proteomics data.
-However, after examining the R packages for modeling proteomics quantification data, we observed that model specification, input, and output formats differ widely.
-At the same time, they have in common that: they fit linear models either to peptide or protein intensities, determine contrasts among conditions, and afterward apply empirical Bayes variance shrinkage.
+However, after examining the R packages for modeling proteomics quantification data, we observed that model specification, input, and output formats differ widely (and wildly).
+However, these packages have in common: they fit linear models either to peptide or protein intensities, determine contrasts among conditions, and afterward, apply empirical Bayes variance shrinkage.
 Therefore, the revised goal was to provide a modular object-oriented (OO) design, with R linear models as a core, where we can add features such as $p$-value aggregation, e.g., `r BiocStyle::Biocpkg("PECA")`, variance shrinkage, or modeling of missing observations.
@@
 Furthermore, the functionality of `r BiocStyle::Githubpkg("fgcz/prolfqua")` also includes methods specific to proteomics data.
-For example, we implemented strategies to estimate protein intensities from peptide intensities: top N [@Grossmann2010], Tukey's median polish  [@tukey1977exploratory], robust linear models [@goeminne2020msqrob].
-Furthermore, peptide or protein abundances can be scaled and transformed, using robust scaling, _quantile_ normalization or `r BiocStyle::Biocpkg("vsn")` to remove systematic differences among samples and heteroscedasticity.
+For example, we implemented strategies to estimate protein intensities from peptide intensities: top N [@Grossmann2010], Tukey's median polish, robust linear models [@goeminne2020msqrob].
+Furthermore, peptide or protein abundances can be scaled and transformed, using robust scaling, transformed using _quantile_ normalization or `r BiocStyle::Biocpkg("vsa")` to remove systematic differences among samples and heteroscedasticity.
 In this respect, `r BiocStyle::Githubpkg("fgcz/prolfqua")` can be compared with other R packages such as `r BiocStyle::Biocpkg("DEP")` [@DEP2018] or `r BiocStyle::Biocpkg("POMA")` [@POMA2021] which support the entire differential expression analysis pipeline.
 
-We also implemented functionality and use the Ionstar [@shen2018ionstar] dataset to benchmark the modeling methods implemented within `r BiocStyle::Githubpkg("fgcz/prolfqua")` and to compare our results with those of `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`.
+We also implemented functionality to streamline the benchmarking of the implemented methods.
+We use the Ionstar [@shen2018ionstar] dataset to benchmark the modeling methods implemented within `r BiocStyle::Githubpkg("fgcz/prolfqua")` and to compare our results with those of `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`.
 Since group sizes are relatively small, typically with four or five subjects per group, the high power of the tests is a relevant criterion to assess the modeling method.
-The quantified proteins can be ranked using the estimated fold-change, $t$-statistics, or scaled $p$-value, and afterward subjected to gene set enrichment (GSEA) or over-representation analysis [@subramanian2005gene] to determine up or down-regulated groups of proteins.
+Furthermore, the quantified proteins can be ranked using the estimated fold-change, $t$-statistics, or scaled $p$-value, and then subjected to gene set enrichment (GSEA) or over-representation analysis [@subramanian2005gene] to determine up or down-regulated groups of proteins.
 Furthermore, the statistical model must provide an unbiased estimate of the false discovery rate (FDR) to manage expectations when selecting protein lists for follow-up experiments.
@@
 We store all the data needed for analysis in a tidy table, i.e., every column is a variable, every row is an observation, and every cell is a single value [@tidydata2014].
-Using an R6 [@R6cite] configuration object (Figure \@ref(fig:LFQData)), we specify what variable is in which column, making it easy to integrate new inputs in `r BiocStyle::Githubpkg("fgcz/prolfqua")` if provided as a tidy table.
-For example, to visualize tidy Spectronaut[@bruderer2015extending], DiaNN[@demichev2020dia], or Skyline[@maclean2010skyline] outputs, or data in `r BiocStyle::Biocpkg("MSstats")`[@Choi2014] format, only a few lines of code are needed to update the `r BiocStyle::Githubpkg("fgcz/prolfqua")` `AnalysisTableConfiguration` configuration.
-The configuration encapsulates the differences in column names among the various input formats, and enables the use of the methods without additional arguments.
-We show an example code for creating an MSFragger[@yu2020fast] configuration in the Appendix.
-For popular software like MaxQuant[@Cox2008MaxQuant], or MSFragger, which stores the same variable, e.g., intensity, in multiple columns, one for each sample, we implemented methods that transform the data into tidy tables.
-Relying on the tidy data table enables us to interface with many data manipulation, visualization, and modeling methods, implemented in base R [@RCoreTeam2021] and the tidyverse [@tidyverse2019], easily.
-We use R6 classes to structure the functionality of the package (see Figure \@ref(fig:LFQData) and Figure \@ref(fig:ContrastUML)).
-R6 classes are well supported by command-line completion features in RStudio [@RStudio2022], and help to implement argument free functions (Appendix Figure \@ref(fig:tabCompletion).
+Using an __R6__[@R6cite] configuration object (Figure \@ref(fig:LFQData)), we specify what variable is in which column, making it easy to integrate new inputs in `r BiocStyle::Githubpkg("fgcz/prolfqua")` if provided as a tidy table.
+For example, to visualize tidy _Spectronaut_, _DiaNN_, or _Skyline_ outputs, or data in `r BiocStyle::Biocpkg("MSstats")` format, only a few lines of code are needed to update the `r BiocStyle::Githubpkg("fgcz/prolfqua")` `AnalysisTableConfiguration` configuration.
+The configuration encapsulates the differences in column names among the various input formats, enables the use of the methods without additional arguments, and preserves the processing tools' differences in the figures and analysis.
+We show an example code for creating an MSFragger configuration in the Supplement.
+For popular software like _MaxQuant_, or_MSFragger_, which stores the same variable, e.g., intensity, in multiple columns, one for each sample, we implemented methods that transform the data into tidy tables.
+Furthermore, relying on the tidy data table enables us to interface with many data manipulation, visualization, and modeling methods, implemented in base R [@RCoreTeam2021] and the tidyverse [@tidyverse2019], easily.
+We use __R6__ classes to structure the functionality of the package (see Figure \@ref(fig:LFQData) and Figure \@ref(fig:ContrastUML)).
+__R6__ classes are well supported, e.g., by command-line completion features, see Figure \@ref(fig:tabCompletion), and easy to use (OOP).
 
@@
 The LFQDataPlotter class and other classes decorate the LFQData class with additional functionality.
-Similarly, the `LFQDataStats` and `LFQDataSummary`, reference the `LFQData` class and group methods for variance and sample size estimation or summarizing peptide and protein counts.
+Similarly, the `LFQDataStats` and`LFQDataSummary`, reference the LFQData class and group methods for variance and sample size estimation or summarizing peptide and protein counts.
 
@@
 
-In addition, we implement features specific to high throughput experiments, such as the experimental Bayes variance and $p$-value moderation, which utilizes the parallel structure of the protein measurements and the analysis [@Ritchie2015].
+In addition, we implemented features specific to high throughput experiments, such as the experimental Bayes variance and $p$-value moderation, which utilizes the parallel structure of the protein measurements and the analysis [@Ritchie2015].
 We also compute probabilities of differential protein regulation based on peptide level models [@Suomi2017bEnhanced].
 We used R6 classes to encapsulate the statistical modelling functionality in `r BiocStyle::Githubpkg("fgcz/prolfqua")` (see Figure \@ref(fig:ContrastUML)).
-We specify a contrast interface (`ContrastsInterface`).
+We did specify a contrast interface (`ContrastsInterface`).
 Several implementations allow to perform differential expression analysis given linear or mixed effect models (`Contrasts`), variance shrinkage (`ContrastsModerated`), or impute contrasts in cases when observations are missing for an entire group of samples (`ContrastsSimpleImpute`).
-Further implementations of the interface encapsulate and integrate differential expression analysis results of external tools such as `r BiocStyle::Biocpkg("proDA")` or of SAINTexpress [@teo2014saintexpress] used to analyze data from protein interaction studies.
+Further implementations of the interface encapsulate and integrate differential expression analysis results of external tools such as `r `BiocStyle::Biocpkg("proDA")` or of SAINTexpress [@teo2014saintexpress] used to analyze data from protein interaction studies using.
 
-(ref:ContrastUML)  The UML (unified modeling language) diagram of modeling and contrast related classes.
-Different strategies, e.g., _lm_, _lmer_, and _glm_ (Table \@ref(tab:prolfquaModels)), can fit various types of models.
-The model builder method fits the statistical model given the data and a strategy.
+(ref:ContrastUML)  UML diagram of modeling and contrast related classes.
+Different strategies, e.g., _lm_, _lmer_, and _glm_, can fit various types of models.
+The model builder method fit's the statistical model given the data and a strategy.
 The obtained model can then be used for the analysis of variance or to estimate contrasts.
 All classes estimating contrasts implement the _Contrasts_ interface.
-Results of external tools, e.g., SAINTexpress, can be adapted by implementing the Contrasts Interface.
+Results of external tools, e.g., SAINTExpress, can be adapted by implementing the Contrasts Interface.
 
@@
 
-## Dataset for benchmarking
+## Dataset for Benchmarking
 
-To evaluate the performance of differential expression analysis, we use the IonStar benchmark dataset[@shen2018ionstar], available from the Proteomics Identifications Database (PRIDE) identifier PXD003881.
+To evaluate the performance of differential expression analysis, we use the _IonStar_ benchmark dataset[@shen2018ionstar], available from the Proteomics Identifications Database (PRIDE) identifier PXD003881.
 $DH5\alpha$ _E.
-coli_ lysate was spiked in human pancreatic cancer cells (Panc-1) lysate at five different levels: $3\%$ _E.
+coli_ lysate was spiked in human pancreatic cancer cells(Panc-1) lysate at five different levels: $3\%$ _E.
 coli_, $4.5\%$ _E.
@@
 coli_.
-We annotated these dilutions from smallest to largest with the letters _a_, _b_, _c_, _d_, _e_.
+We annotated these dilutions from smallest to largest with the letters a, b, c, d, e.
 By comparing the various dilutions, we obtain different effect sizes, e.g., when comparing dilution _e_ (9%) against dilution _d_ (7.5%), the expected difference is $1.2$ for E.
@@
 
-The peptide abundances (from the MaxQuant _peptide.txt_ file), were $log_2$ transformed and subsequently scaled, where median and the mean average deviation (mad) were obtained from the human proteins only.
+The peptide abundances (from the _MaxQuant_ _peptide.txt_ file), were $log_2$ transformed and subsequently scaled, where median and the mean average deviation (mad) were obtained from the human proteins only.
 We removed _one hit wonders_, i.e., proteins with a single peptide assignment.
-Protein abundances are inferred from the peptide intensities using Tukey's median polish.
+Protein abundances are inferred from the peptide intensities using Tukey's median polish [@tukey1977exploratory].
 Finally, we fitted the fixed effect linear models, the dropout model `r BiocStyle::Biocpkg("proDA")` to protein abundances, the mixed effect linear model, and the `r BiocStyle::Biocpkg("PECA")` model to peptide intensities.
@@
 
-The IonStar dataset contains _H.
+The IonStar [@shen2018ionstar] dataset contains _H.
 sapiens_ proteins with constant concentrations and _E.
@@
 
-knitr::kable(table, caption = "Confusion matrix, TP - true positive, FP - false positive, FN - false negative, TN - true negatives, P - all positive cases (all $E.~coli$ proteins), N - all negative cases (all $H.~sapiens$ proteins), m - all proteins.")
+knitr::kable(table, caption = "Confusion matrix, TP - true positive, FP - false positive, FN - false negative, P - all positive cases (all E.
+coli proteins), N - all negative cases (all H.
+sapiens proteins), m- all proteins.")
 ```
@@
 @valikangas2016systematic discuss and benchmark various methods of peptide or protein intensity normalization, such as variance stabilizing normalization [@huber2002variance] or quantile normalization [@bolstad2003comparison].
-In this work, we use a robust version of the z-score, where instead of the mean $\bar{x}$ we use the median, and instead of the standard deviation $\tilde{S}$ the median absolute deviation (mad):
+In this work, we will be using a robust version of the z-score, where instead of the mean $\bar{x}$ we use the median, and instead of the standard deviation $\tilde{S}$ the median absolute deviation (mad):
 
@@
 
-Given a linear model $y = \beta X$, we can compute the difference between two groups $\beta_{c}$ by a linear combination $c$ of linear model parameters $\beta$, where $c$ is a column vector with as many elements as there are coefficients $\beta$ in the linear model.
+Given a linear model $y = \beta X$, we can compute the difference between two groups $\beta_{c}$ by a linear combination $c$ of linear model parameters $\beta$.
+$c$ is a column vector with as many elements as there are coefficients in the linear model.
 If $c$ has $0$ for one or more of its rows, then the corresponding coefficient in $\beta$ is not involved in determining the contrast [@ph525xseries].
@@
 The degrees of freedom for the contrast are equal to the residual degrees of freedom of the linear model.
-For estimating contrasts from mixed effects models we used the function `contest` implemented in the R package `r BiocStyle::CRANpkg("lmerTest")`  and used the Satterthwaite [@Kuznetsova2017lmerTest] method to estimate the denominator degrees of freedom.
+For estimating contrasts from mixed effects models we used the function `contest` implemented in the R package `r BiocStyle::CRANpkg("lmerTest")` [@Kuznetsova2017lmerTest] and used the Satterthwaite method to estimate the denominator degrees of freedom.
 These methods are available in the class `Contrast` (see Figure \@ref(fig:ContrastUML))
@@
 
-To summarize peptide level models to protein models, we applied the method suggested by @Suomi2017bEnhanced to use the median scaled $p$-value of the peptide models and the cumulative distribution function of the beta distribution function to determine a regulation probability of the protein.
+To summarize peptide level models to protein models, we applied the method suggested by [@Suomi2017bEnhanced] to use the median scaled $p$-value of the peptide models and cumulative distribution function of the beta distribution function to determine a regulation probability of the protein.
 
@@
 Using the `LFQDataPlotter` class, we show the distribution of the normalized peptide abundances in Figure \@ref(fig:prepro) Panel A.
-Afterwards, protein intensities are estimated from peptide intensities using Tukey's median polish.
+Afterwards, protein intensities are estimated from peptide intensities using Tukey's median polish [@tukey1977exploratory].
 Figure \@ref(fig:prepro) Panel B shows the peptide intensities and the estimated protein intensities.
@@
 For each sample a line with a different colour is shown.
-Panel B - Peptide intensities for protein _5NTD_ are shown using a line of different colour, and the protein intensity estimate is shown using a fat black line, Panel C - distribution of standard deviations of all proteins in each dilution group ($a$, $b$, $c$, $d$, $e$) and overall (all), Panel D - Distribution of protein intensities for protein _5NTD_.
+Panel B - Peptide intensities for protein _5NTD_ are shown using a line of different colour, and the protein intensity estimate is shown using a fat black line, Panel C - distribution of standard deviations of all proteins in each dilution group (a, b, c, d, e) and overall (all), Panel D - Distribution of protein intensities for protein _5NTD_.
 
@@
 Then we create the plots shown in Figure \@ref(fig:exampleContrasts).
-Panel A shows the distribution of the $p$-values, while Panel B shows the volcano plot for each comparison.
+Panel A shows the distribution of the $p$-values, while figure Panel B shows the volcano plot for each comparison.
 
@@
 Using the tidy table to model the data ensures interoperability with other proteomics-related packages that manage their data with tidy-tables, e.g., `r BiocStyle::CRANpkg("protti")` [@quast2022protti].
-The use of R6 classes, which encapsulate the configuration and the data, allows for writing very concise code (no function arguments needed).
+Furthermore, the use of `R6` classes, which encapsulate the configuration and the data, allows for writing very concise code (no function arguments needed).
 Auto-completion support in the Rstudio editor makes it easy for novices to explore `r BiocStyle::Githubpkg("fgcz/prolfqua")`s functionality (Figure \@ref(fig:tabCompletion)).
-To simplify integration of `r BiocStyle::Githubpkg("fgcz/prolfqua")` with Bioconductor based workflows there is a method to convert the LFQData class into a `r BiocStyle::Biocpkg("SummarizedExperiment")`.
+To simplify integration of `r BiocStyle::Githubpkg("fgcz/prolfqua")` with Bioconductor based workflows there is an method to convert the LFQData class into a `r BiocStyle::Biocpkg("SummarizedExperiment")`.
 
-To ease the usage barriers of the R package to users not being familiar with statistics and R programming, we integrated some workflows into our data management platform [b-fabric](https://fgcz-bfabric.uzh.ch/) [@bfabric].
+To ease the usage barriers of the R package to users not being familiar with statistics and R programming, we integrated some workflows into our data management platform [b-fabric](https://fgcz-bfabric.uzh.ch/) @bfabric.
 This integration enables our users to select the input and basic settings in a graphical user interface.
@@
 
-`r BiocStyle::Githubpkg("fgcz/prolfqua")` allows for considerable flexibility to model your experiments.
+`r BiocStyle::Githubpkg("fgcz/prolfqua")` allows for considerable flexibility to model quantitative proteomics experiments.
 Various types of models are available (see Figure \@ref(fig:ContrastUML) and Table \@ref(tab:prolfquaModels)), and the contrast specification is explicit and consistent for all models.
@@
 
-# Acknowledgements {.unlisted .unnumbered}
+# Acknoledgements {.unlisted .unnumbered}
 
@@
 $FDR$  &False Discovery Rate                               \\
-LC     &Liquid Chromatography                              \\
+HPLC   &High-Pressure Liquid Chromatography                \\
 LC-MS  &Liquid Chromatography followed by Mass Spectrometry\\

### 2022-06-07 inst/MyArticle/paper.Rmd
Update paper.Rmd
Incooperate Marias changes 3.2 Benchmarking modelling approaches

--- 2022-06-07 inst/MyArticle/paper.Rmd
@@
 
-When comparing differential expression analysis performance, a relevant parameter is the number of proteins for which a method estimated differences (see Figure \@ref(fig:FDRfdp) Panel A).
-It indicates how robust the procedure works in the presence of missing observations.
+When comparing differential expression analysis performance, a relevant parameter is the number of proteins for which a method estimates differences (see Figure \@ref(fig:FDRfdp) Panel A).
+The number of proteins indicates how robust the procedure works in the presence of missing observations.
 For each protein, we tried to determine four differences ($\Delta = (1.20, 1.25, 1.30, 1.50)$).
@@
 Interestingly, the $pAUC$ increases when using $p$-values instead of the $t$-statistics for linear models, while it decreases for mixed effect models.
-The reason is erroneous denominator degrees of freedom estimation for many proteins in the case of the mixed-effect models.
+The reason is an erroneous denominator degrees of freedom estimation for many proteins in the case of the mixed-effect models.
 

### 2022-06-07 inst/MyArticle/paper.Rmd
Update paper.Rmd
more on conclusion

--- 2022-06-07 inst/MyArticle/paper.Rmd
@@
 ---
-title: prolfqua - A Comprehensive R-package for Proteomics Differential Expression Analysis
+title: prolfqua: A Comprehensive R-package for Proteomics Differential Expression Analysis
 author: 
@@
 
-In summary, `r BiocStyle::Githubpkg("fgcz/prolfqua")` is an easy-to-use and feature-rich R package to analyze quantitative mass spectrometric data, report results, and benchmark MS software and statistical methods.
+In summary, `r BiocStyle::Githubpkg("fgcz/prolfqua")` is an easy-to-use and feature-rich R package to analyze quantitative mass spectrometric data with simple and complex experimental designs.
+It also can generate conclusive reports, and benchmark MS software and statistical methods.
+Furthermore, with very little adaptations this R package can also be adapted to analyze different quantitative proteomics data (e.g.
+labeling based TMT-, PRM- DIA-data or even other omics scale datasets such as microarray data).
 We provide documentation, in vignette format, at the website https://github.com/fgcz/prolfqua/.

### 2022-06-07 inst/MyArticle/paper.Rmd
tmp

--- 2022-06-07 inst/MyArticle/paper.Rmd
@@
 We use R6 classes to structure the functionality of the package (see Figure \@ref(fig:LFQData) and Figure \@ref(fig:ContrastUML)).
-R6 classes are well supported by command-line completion features in RStudio [@RStudio2022], and help to implement argument free functions (Appendix Figure \@ref(fig:tabCompletion).
+R6 classes are well supported by command-line completion features in RStudio [@RStudio2022], and help to implement argument free functions (Figure \@ref(fig:tabCompletion).
 
@@
 The LFQDataPlotter class and other classes decorate the LFQData class with additional functionality.
-Similarly, the `LFQDataStats` and `LFQDataSummary`, reference the `LFQData` class and group methods for variance and sample size estimation or summarizing peptide and protein counts.
+Similarly, the `LFQDataStats` and `LFQDataSummary` reference the `LFQData` class and group methods for variance and sample size estimation or summarizing peptide and protein counts.
 
@@
 We specify a contrast interface (`ContrastsInterface`).
-Several implementations allow to perform differential expression analysis given linear or mixed effect models (`Contrasts`), variance shrinkage (`ContrastsModerated`), or impute contrasts in cases when observations are missing for an entire group of samples (`ContrastsSimpleImpute`).
+Several implementations allow to perform differential expression analysis given linear or mixed effect models (`Contrasts`), variance shrinkage (`ContrastsModerated`), or to impute contrasts in cases when observations are missing for an entire group of samples (`ContrastsSimpleImpute`).
 Further implementations of the interface encapsulate and integrate differential expression analysis results of external tools such as `r BiocStyle::Biocpkg("proDA")` or of SAINTexpress [@teo2014saintexpress] used to analyze data from protein interaction studies.
@@
 
-The peptide abundances (from the MaxQuant _peptide.txt_ file), were $log_2$ transformed and subsequently scaled, where median and the mean average deviation (mad) were obtained from the human proteins only.
+The peptide abundances (from the MaxQuant _peptide.txt_ file) were $log_2$ transformed and subsequently scaled, where median and the mean average deviation (mad) were obtained from the human proteins only.
 We removed _one hit wonders_, i.e., proteins with a single peptide assignment.
@@
 We know that for _H.
-sapiens_ proteins, the difference $\beta$ between two dilutions should be $\beta = 0$.
-While for _E.
+sapiens_ proteins, the difference $\beta$ between two dilutions should be $\beta = 0$, while for _E.
 coli_ proteins, we know that the difference between dilutions should be $\beta \ne 0$.
@@
 @valikangas2016systematic discuss and benchmark various methods of peptide or protein intensity normalization, such as variance stabilizing normalization [@huber2002variance] or quantile normalization [@bolstad2003comparison].
-In this work, we use a robust version of the z-score, where instead of the mean $\bar{x}$ we use the median, and instead of the standard deviation $\tilde{S}$ the median absolute deviation (mad):
+In this work, we use a robust version of the z-score, where instead of the mean $\bar{x}$ we use the median, and instead of the standard deviation $\tilde{S}$ we use the median absolute deviation (mad):
 
@@
 
-To apply this transformation, we need to estimate two parameters per sample.
-It works for experiments with thousands of proteins and experiments where only a few hundred proteins per sample are measured.
+To apply this transformation, we need to estimate two parameters per sample, therefore it works for experiments with thousands of proteins and experiments where only a few hundred proteins per sample are measured.
 For the Ionstar dataset, we used the intensities of _H.
@@
 
-Given a linear model $y = \beta X$, we can compute the difference between two groups $\beta_{c}$ by a linear combination $c$ of linear model parameters $\beta$, where $c$ is a column vector with as many elements as there are coefficients $\beta$ in the linear model.
+Given a linear model $y = \beta X$, we can compute the difference $\beta_{c}$ between two groups by a linear combination $c$ of linear model parameters $\beta$, where $c$ is a column vector with as many elements as there are coefficients $\beta$ in the linear model.
 If $c$ has $0$ for one or more of its rows, then the corresponding coefficient in $\beta$ is not involved in determining the contrast [@ph525xseries].
@@
 
-with $X$ being the design matrix, $\beta$ the model parameters and $c$ an array of weights, defining the contrast (difference among conditions).
+with $X$ being the design matrix.
 The degrees of freedom for the contrast are equal to the residual degrees of freedom of the linear model.
@@
 
-The package `r BiocStyle::Githubpkg("fgcz/prolfqua")` provides function to determine the vector of $parameter$ weights $c$, from a linear models and a contrast specification string.
+The package `r BiocStyle::Githubpkg("fgcz/prolfqua")` provides a function to determine the vector of $parameter$ weights $c$, from a linear models and a contrast specification string.
 
@@
 
-The following code shows an example where we specified two contrast, the first to compare Cells of type CMP/MEP with cells of type HSC, and the second where we contrast therapy NO with therapy NU for the celltype CMP/MEP [@meier2021reduced].
+The following code shows an example where we specify two contrast: the first to compare Cells of type CMP/MEP with cells of type HSC, and the second to compare therapy NO with therapy NU for the celltype CMP/MEP [@meier2021reduced].
 Finally, we compute the array of weights defining the contrast, used to multiply the model coefficient $\beta$.
@@
 Missing observations lead to different group sizes, which results in unbalanced designs.
-Linear and mixed effect models implemented in R can handle unbalanced designs.
-Therefore, as long as one observation per group is available, they will produce unbiased estimates, and no imputation is needed.
+Linear and mixed effect models can handle unbalanced designs, as long as one observation per group is available, they will produce unbiased estimates, and no imputation is needed.
 
@@
 
-with $n_i$ the number of observations, and $s_i$ the standard deviation in each condition.
+with $n_i$ the number of observations, and $s_i$ the standard deviation in each group.
 The standard deviation for the $t$-statistics is then given by:
@@
 
-To summarize peptide level models to protein models, we applied the method suggested by @Suomi2017bEnhanced to use the median scaled $p$-value of the peptide models and the cumulative distribution function of the beta distribution function to determine a regulation probability of the protein.
+To summarize peptide level models to protein models, we apply the method suggested by @Suomi2017bEnhanced that use the median scaled $p$-value of the peptide models and the cumulative distribution function of the Beta distribution (CDF) to determine a regulation probability of the protein.
 
-To obtain the median $p$-value of a protein, we first rescaled the peptide $p$-values by taking the sign of the fold-change $\hat \beta$ into account, i.e.:
+To obtain the $\tilde{p}$ of a protein, we first rescaled the peptide $p$-values by taking the sign of the fold-change $\hat \beta$ into account, i.e.:
 
@@
 
-Because we used the median as the i-th order statistic $i = \frac{n}{2} + 0.5$.
-Therefore, to parameterize the cumulative distribution function (CDF) of the Beta distribution, $\gamma = i = \frac{n}{2} + 0.5$ and $\delta = n - i + 1 = n - (\frac{n}{2} + 0.5) + 1 = \frac{n}{2} + 0.5 = \gamma$.
+Because we use the median, with the i-th order statistic $i = \frac{n}{2} + 0.5$, we parametrize the CDF of the Beta distribution with $\gamma = i = \frac{n}{2} + 0.5$ and $\delta = n - i + 1 = n - (\frac{n}{2} + 0.5) + 1 = \frac{n}{2} + 0.5 = \gamma$.
 We implemented this method in the class `ContrastROPECA` (Figure \@ref(fig:ContrastUML)).
@@
 The code snippets in this section demonstrate how a differential expression analysis workflow can be implemented using the `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package.
-To speed up the computation of these examples, we use a subset of the Ionstar dataset containing $163$ proteins and $1258$ peptides.
+To speed up the computation of these examples, we use a random subset of the Ionstar dataset containing $163$ proteins and $1258$ peptides.
 Peptide abundances are $\log_2$ transformed and robust z-score scaled using the method `robscale`.
@@
 
-The following code example illustrates how we compute contrast by comparing the protein intensities among groups.
+The following code example illustrates how we compute differences among groups.
 First, the linear model and the differences are specified.
@@
 The b-fabric system runs a computing infrastructure controlled by a local resource management system that supports cloud-bursting [@VMMAD].
-In this way, [b-fabric](https://fgcz-bfabric.uzh.ch/) helps scientists meet requirements from funding agencies, journals, and academic institutions to publish data according to the FAIR (Findable, Accessible, Interoperable and Reusable) [@FAIR] data principles.
+In this way, [b-fabric](https://fgcz-bfabric.uzh.ch/) helps scientists to meet requirements from funding agencies, journals, and academic institutions to publish data according to the FAIR (Findable, Accessible, Interoperable and Reusable) [@FAIR] data principles.
 
@@
 
-When comparing differential expression analysis performance, a relevant parameter is the number of proteins for which a method estimates differences (see Figure \@ref(fig:FDRfdp) Panel A).
-The number of proteins indicates how robust the procedure works in the presence of missing observations.
+When comparing differential expression analysis performance, a relevant parameter is the number of proteins for which a method estimated differences (see Figure \@ref(fig:FDRfdp) Panel A).
+It indicates how robust the procedure works in the presence of missing observations.
 For each protein, we tried to determine four differences ($\Delta = (1.20, 1.25, 1.30, 1.50)$).
@@
 Interestingly, the $pAUC$ increases when using $p$-values instead of the $t$-statistics for linear models, while it decreases for mixed effect models.
-The reason is an erroneous denominator degrees of freedom estimation for many proteins in the case of the mixed-effect models.
+The reason is erroneous denominator degrees of freedom estimation for many proteins in the case of the mixed-effect models.
 

### 2022-06-07 inst/MyArticle/paper.Rmd
Affiliation

@@
 ---
-title: prolfqua: A Comprehensive R-package for Proteomics Differential Expression Analysis
+title: prolfqua - A Comprehensive R-package for Proteomics Differential Expression Analysis
 author: 
   - Witold E.
-Wolski $^{\dagger\ddagger}$^[[Functional Genomics Center Zurich (FGCZ)](https://fgcz.ch) - [Swiss Federal Institute of Technology in Zurich](https://ethz.ch) \| [University of Zurich](https://uzh.ch); Winterthurerstrasse 190, CH-8057 Zurich, Switzerland.
-$^\ddagger$[Swiss Institute of Bioinformatics (SIB)](https://www.sib.swiss/) $^\dagger$Correspondence \texttt{wew@fgcz.ethz.ch}]
+Wolski $^{\dagger\ddagger}$^[[Functional Genomics Center Zurich (FGCZ)](https://fgcz.ch) - [University of Zurich](https://uzh.ch)/[ETH Zurich](https://ethz.ch), Winterthurerstrasse 190, CH-8057 Zurich, Switzerland.
+$^\ddagger$[Swiss Institute of Bioinformatics (SIB), Quartier Sorge - Batiment Amphipole, 1015 Lausanne, Switzerland](https://www.sib.swiss/) $^\dagger$Correspondence \texttt{wew@fgcz.ethz.ch}]
   - Paolo Nanni$^{\ast}$
@@
 
-`r BiocStyle::Githubpkg("fgcz/prolfqua")` allows for considerable flexibility to model quantitative proteomics experiments.
+`r BiocStyle::Githubpkg("fgcz/prolfqua")` allows for considerable flexibility to model your experiments.
 Various types of models are available (see Figure \@ref(fig:ContrastUML) and Table \@ref(tab:prolfquaModels)), and the contrast specification is explicit and consistent for all models.
@@
 
-In summary, `r BiocStyle::Githubpkg("fgcz/prolfqua")` is an easy-to-use and feature-rich R package to analyze quantitative mass spectrometric data with simple and complex experimental designs.
-It also can generate conclusive reports, and benchmark MS software and statistical methods.
-Furthermore, with very little adaptations this R package can also be adapted to analyze different quantitative proteomics data (e.g.
-labeling based TMT-, PRM- DIA-data or even other omics scale datasets such as microarray data).
+In summary, `r BiocStyle::Githubpkg("fgcz/prolfqua")` is an easy-to-use and feature-rich R package to analyze quantitative mass spectrometric data, report results, and benchmark MS software and statistical methods.
 We provide documentation, in vignette format, at the website https://github.com/fgcz/prolfqua/.

### 2022-06-07 inst/MyArticle/paper.Rmd
more suggestions from PN.

--- 2022-06-07 inst/MyArticle/paper.Rmd
@@
 ---
-title: prolfqua: A Comprehensive R-package for Proteomics Differential Expression Analysis
+title: prolfqua - A Comprehensive R-package for Proteomics Differential Expression Analysis
 author: 
@@
 Furthermore, this glass box approach should make it easy to reimplement an analysis performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` using base R or other programming languages by reading the analysis script.
-However, acknowledging the R formula interface’s complexity to non-statisticians and the popularity of `r BiocStyle::Biocpkg("MSstats")`, we provide the functionality to suggest a model formula from the sample annotation provided in tabular form similarly to `r BiocStyle::Biocpkg("MSstats")` [@Choi2014].
+However, acknowledging the R formula interface’s complexity to non-statisticians and the popularity of `r BiocStyle::Biocpkg("MSstats")`, we provide the functionality to suggest a model formula from the sample annotation provided in tabular form similarly to `r BiocStyle::Biocpkg("MSstats")`.
 
@@
 
-The Benchmark functionality of `r BiocStyle::Githubpkg("fgcz/prolfqua")` includes receiver operator curves and computes partial areas under those curves (pAUC) and other scores, e.g., the false discovery proportion FDP.
+The Benchmark functionality of `r BiocStyle::Githubpkg("fgcz/prolfqua")` includes receiver operator curves (ROC) and computes partial areas under those curves (pAUC) and other scores, e.g., the false discovery proportion FDP.
 We use those scores, i.e.
-the $pAUC$ at $10%$ FDR and the FDP, to examine how well the methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` model quantitative mass spectrometric high throughput data and compare them with results produced by `r BiocStyle::Biocpkg("MSstats")` [@MSstats2014] and `r BiocStyle::Biocpkg("proDA")` [@bioxrvproDA2020].
+the $pAUC$ at $10%$ FDR and the FDP, to examine how well the methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` model quantitative mass spectrometric high throughput data and compare them with results produced by `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`.
 Table \@ref(tab:describeBenchmarkedModels) summarizes all modelling methods we evaluated.
@@
 For instance, we observe that `r BiocStyle::Biocpkg("MSstats")` estimates $16058$ group differences while the mixed effect models estimates the fewest with $15940$.
-Hence, to conclude that one method shows a better performance, it does not suffice that pAUC is greater, but the number of proteins with differential expression results needs to be equal or larger.
+Hence, to conclude that one method shows a better performance, it does not suffice if the pAUC is greater, but the number of proteins with differential expression results needs to be equal or larger.
 
@@
 Interestingly, the $pAUC$ increases when using $p$-values instead of the $t$-statistics for linear models, while it decreases for mixed effect models.
-The reason is an erroneous denominator degrees of freedom estimation for many proteins in the case of the mixed-effect models.
+The reason is erroneous denominator degrees of freedom estimation for many proteins in the case of the mixed effect models.
 
@@
 kableExtra::kable(xMM,
-             caption = paste0("All benchmarked models.",
-             "Description - prolfqua function names",
-             "Abudances - indicates if model is fitted to peptide or protein abundances",
-             "Input File - MaxQuant file used as input to the analysis."),
+             caption = paste0("All benchmarked models. ",
+             "Description - prolfqua function names, ",
+             "Abudances - indicates if model is fitted to peptide or protein abundances, ",
+             "Input File - name of MaxQuant file used as input."),
              format = "latex", booktabs=TRUE) |>
     kableExtra::kable_styling(latex_options="scale_down") |> 
-    kableExtra::add_footnote(label = "Merges results of the prot_med_lm_moderated` and prot_imputation modeling pipeline, preferring those of prot_med_lm_moderated if available.")
+    kableExtra::add_footnote(label = "Merges results of the 'prot_med_lm_moderated' and 'prot_imputation' modeling pipeline, preferring those of 'prot_med_lm_moderated' if available.")
 
@@
 Since only technical replicates are available for each dilution, essential sources of variation present in every experiment, such as the biochemical and biological, are not measured.
-Furthermore, the dataset captures only the variance from the chromatography, electro-spray, and mass spectrometric measurement method.
+Therefore, the dataset captures only the variance from the chromatography, electro-spray, and mass spectrometric measurement method.
 Thus, while we can extrapolate some of the results obtained to more realistic datasets, we need to be careful not to over-interpret the results.
@@
 
-We can conclude that if we want to order the proteins according to the likelihood of being differentially regulated to perform gene set enrichment analysis [@subramanian2005gene], the $t$-statistic is better suited than the fold-change estimate.
+We can conclude that if we want to sort the proteins according to the likelihood of being differentially regulated to perform gene set enrichment analysis [@subramanian2005gene], the $t$-statistic is better suited than the fold-change estimate.
 Modeling the degrees of freedom when computing the $p$-values might improve the inference.
@@
 There is no such improvement for the mixed effect model, most likely because the degrees of freedom are erroneously estimated for many models.
-Furthermore, for the fixed effect linear model, the empirical Bayes variance shrinkage, as suggested by [@Smyth2004linear], consistently improves the ranking of proteins compared with the unmoderated estimates (not shown) and fails to do so for mixed effect models.
+Furthermore, for the fixed effect linear model, the empirical Bayes variance shrinkage, as suggested by @Smyth2004linear, consistently improves the ranking of proteins compared with the unmoderated estimates (not shown) and fails to do so for mixed effect models.
 
 Computing the statistics at the peptide level, e.g., the $t$-statistics or $p$-value, then summarizing these statistics using their median produces the highest $AUC$ scores among all the tested models (see Figure \@ref(fig:FDRfdp) Panel A left).
-Furthermore, by using the beta distribution to model the number of peptides observed, we can further improve the $pAUC$ scores (see Figure \@ref(fig:FDRfdp) Panel A center).
-However, the properties of beta-based probabilities are not well understood; for instance, the p-values are not uniformly distributed under the null hypothesis (not shown).
+Furthermore, by using the Beta distribution to model the number of peptides observed, we can further improve the $pAUC$ scores (see Figure \@ref(fig:FDRfdp) Panel A center).
+However, the properties of Beta-based probabilities are not well understood; for instance, the p-values are not uniformly distributed under the null hypothesis (not shown).
 Furthermore, the FDR estimates obtained when correcting for multiple testing with the Benjamini-Hochberg method are biased and overestimate the false discovery proportion (see Figure \@ref(fig:FDRfdp) Panel C).
@@
 Because of the robustness of the dropout model to missing observations we obtain difference estimates for all proteins and contrasts (Figure \@ref(fig:FDRfdp) Panel B).
-Moreover, the estimated differences are unbiased and show high diagnostic accuracy (\@ref(fig:FDRfdp) Panel A).
+Moreover, the estimated differences are unbiased and show high diagnostic accuracy (Figure \@ref(fig:FDRfdp) Panel A).
 Furthermore, the performance of the scaled $p$-values or the $t$-statistics is comparable with that of the linear model with variance moderation (`prot_med_lm_moderated`).
-Therefore, we started to integrate the `r BiocStyle::Biocpkg("proDA")` package as an additional modeling option into `r BiocStyle::Githubpkg("fgcz/prolfqua")` (see Figure \@ref(fig:ContrastUML)).
+Therefore, we are planning to integrate the `r BiocStyle::Biocpkg("proDA")` package as an additional modeling option into `r BiocStyle::Githubpkg("fgcz/prolfqua")` (see Figure \@ref(fig:ContrastUML)).
 
 The R-package `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Biocpkg("prolfqua")` model the missing data directly, while MSstats imputes the data using an accelerated failure model.
-Despite imputation, `r BiocStyle::Biocpkg("MSstats")` do not estimate group differences for more proteins and do not achieve a higher $pAUC$ score than `r BiocStyle::Githubpkg("fgcz/prolfqua")`.
-Furthermore, \@ref(fig:FDRfdp) Panel C shows that when using `r BiocStyle::Biocpkg("MSstats")` the proportion of false discoveries might be very high even when filtering for a low FDR.
+Despite imputation, `r BiocStyle::Biocpkg("MSstats")` does not estimate group differences for more proteins and does not achieve a higher $pAUC$ score than `r BiocStyle::Githubpkg("fgcz/prolfqua")`.
+Furthermore, Figure \@ref(fig:FDRfdp) Panel C shows that when using `r BiocStyle::Biocpkg("MSstats")` the proportion of false discoveries might be very high even when filtering for a low FDR.
 

### 2022-06-07 inst/MyArticle/paper.Rmd

@@
 ---
-title: prolfqua - A Comprehensive R-package for Proteomics Differential Expression Analysis
+title: prolfqua: A Comprehensive R-package for Proteomics Differential Expression Analysis
 author: 
   - Witold E.
-Wolski $^{\dagger\ddagger}$^[[Functional Genomics Center Zurich (FGCZ)](https://fgcz.ch) - [Swiss Federal Institute of Technology in Zurich](https://ethz.ch) \| [University of Zurich](https://uzh.ch); Winterthurerstrasse 190, CH-8057 Zurich, Switzerland.
-$^\ddagger$[Swiss Institute of Bioinformatics (SIB)](https://www.sib.swiss/) $^\dagger$Correspondence \texttt{wew@fgcz.ethz.ch}]
+Wolski $^{\dagger\ddagger}$^[[Functional Genomics Center Zurich (FGCZ)](https://fgcz.ch) - [University of Zurich](https://uzh.ch)/[ETH Zurich](https://ethz.ch), Winterthurerstrasse 190, CH-8057 Zurich, Switzerland.
+$^\ddagger$[Swiss Institute of Bioinformatics (SIB), Quartier Sorge - Batiment Amphipole, 1015 Lausanne, Switzerland](https://www.sib.swiss/) $^\dagger$Correspondence \texttt{wew@fgcz.ethz.ch}]
   - Paolo Nanni$^{\ast}$
@@
 For instance, we observe that `r BiocStyle::Biocpkg("MSstats")` estimates $16058$ group differences while the mixed effect models estimates the fewest with $15940$.
-Hence, to conclude that one method shows a better performance, it does not suffice if the pAUC is greater, but the number of proteins with differential expression results needs to be equal or larger.
+Hence, to conclude that one method shows a better performance, it does not suffice that pAUC is greater, but the number of proteins with differential expression results needs to be equal or larger.
 
@@
 Interestingly, the $pAUC$ increases when using $p$-values instead of the $t$-statistics for linear models, while it decreases for mixed effect models.
-The reason is erroneous denominator degrees of freedom estimation for many proteins in the case of the mixed effect models.
+The reason is an erroneous denominator degrees of freedom estimation for many proteins in the case of the mixed-effect models.
 

### 2022-06-07 inst/MyArticle/paper.Rmd
make it shorter

--- 2022-06-07 inst/MyArticle/paper.Rmd
@@
 
-When comparing differential expression analysis performance, a relevant parameter is the number of proteins for which a method estimated differences (see Figure \@ref(fig:FDRfdp) Panel A).
-It indicates how robust the procedure works in the presence of missing observations.
+When comparing differential expression analysis performance, a relevant parameter is the number of proteins for which a method estimated differences (see Figure \@ref(fig:FDRfdp) Panel A), which indicates how robust the procedure works in the presence of missing observations.
 For each protein, we tried to determine four differences ($\Delta = (1.20, 1.25, 1.30, 1.50)$).
@@
 Interestingly, the $pAUC$ increases when using $p$-values instead of the $t$-statistics for linear models, while it decreases for mixed effect models.
-The reason is erroneous denominator degrees of freedom estimation for many proteins in the case of the mixed effect models.
+The reason is an erroneous denominator degrees of freedom estimation for many proteins in the case of the mixed effect models.
 

### 2022-06-07 inst/MyArticle/paper.Rmd

--- 2022-06-07 inst/MyArticle/paper.Rmd
@@
 
-When comparing differential expression analysis performance, a relevant parameter is the number of proteins for which a method estimates differences (see Figure \@ref(fig:FDRfdp) Panel A).
-The number of proteins indicates how robust the procedure works in the presence of missing observations.
+When comparing differential expression analysis performance, a relevant parameter is the number of proteins for which a method estimated differences (see Figure \@ref(fig:FDRfdp) Panel A), which indicates how robust the procedure works in the presence of missing observations.
 For each protein, we tried to determine four differences ($\Delta = (1.20, 1.25, 1.30, 1.50)$).
@@
 Interestingly, the $pAUC$ increases when using $p$-values instead of the $t$-statistics for linear models, while it decreases for mixed effect models.
-The reason is an erroneous denominator degrees of freedom estimation for many proteins in the case of the mixed-effect models.
+The reason is an erroneous denominator degrees of freedom estimation for many proteins in the case of the mixed effect models.
 

### 2022-06-07 inst/MyArticle/paper.Rmd
replace condition with group

@@
 Nevertheless, highly abundant proteins can suppress the detection of other proteins in a sample, a MAR mechanism.
-Furthermore, a weak correlation between the number of missing measurements in a condition and the abundance of the quantified protein can be observed, which is caused by the limit of detection (LOD), a MNAR mechanism [@mcgurk2020use].
+Furthermore, a weak correlation between the number of missing measurements in a group and the abundance of the quantified protein can be observed, which is caused by the limit of detection (LOD), a MNAR mechanism [@mcgurk2020use].
 
@@
 However, after examining the R packages for modeling proteomics quantification data, we observed that model specification, input, and output formats differ widely.
-At the same time, they have in common that: they fit linear models either to peptide or protein intensities, determine contrasts among conditions, and afterward apply empirical Bayes variance shrinkage.
+At the same time, they have in common that: they fit linear models either to peptide or protein intensities, determine differences among groups, and afterward apply empirical Bayes variance shrinkage.
 Therefore, the revised goal was to provide a modular object-oriented (OO) design, with R linear models as a core, where we can add features such as $p$-value aggregation, e.g., `r BiocStyle::Biocpkg("PECA")`, variance shrinkage, or modeling of missing observations.
@@
 ```{r exampleContrasts, fig.cap="(ref:exampleContrasts)", echo=TRUE, out.width = '90%', fig.width = 8, fig.height = 5}
-# specify differences among conditions
+# specify differences among groups
 contrasts <- c(
@@
 (ref:FDRfdp) Panel A - Number of estimated contrasts for each modeling method (higher is better).
-Panel B - Partial area under the ROC curve at $10\%$ FPR ($pAUC_10$) for all contrasts and three different statistics: the difference among conditions, the scaled $p$-value (sign(diff) $\cdot$ p.value) and the $t$-statistics (higher is better).
+Panel B - Partial area under the ROC curve at $10\%$ FPR ($pAUC_10$) for all contrasts and three different statistics: the difference among groups, the scaled $p$-value (sign(diff) $\cdot$ p.value) and the $t$-statistics (higher is better).
 The red line indicates the average area under the curve of all methods.

### 2022-06-07 inst/MyArticle/paper.Rmd
Replace colon in yaml header

--- 2022-06-07 inst/MyArticle/paper.Rmd
@@
 ---
-title: prolfqua: A Comprehensive R-package for Proteomics Differential Expression Analysis
+title: prolfqua&colon; A Comprehensive R-package for Proteomics Differential Expression Analysis
 author: 
@@
 Wolski $^{\dagger\ddagger}$^[[Functional Genomics Center Zurich (FGCZ)](https://fgcz.ch) - [University of Zurich](https://uzh.ch)/[ETH Zurich](https://ethz.ch), Winterthurerstrasse 190, CH-8057 Zurich, Switzerland.
-$^\ddagger$[Swiss Institute of Bioinformatics (SIB), Quartier Sorge - Batiment Amphipole, 1015 Lausanne, Switzerland](https://www.sib.swiss/) $^\dagger$Correspondence \texttt{wew@fgcz.ethz.ch}]
+$^\ddagger$[Swiss Institute of Bioinformatics (SIB), Quartier Sorge - Batiment Amphipole, 1015 Lausanne, Switzerland](https://www.sib.swiss/) $^\dagger$Correspondence&colon; \texttt{wew@fgcz.ethz.ch}]
   - Paolo Nanni$^{\ast}$
@@
 
-When comparing differential expression analysis performance, a relevant parameter is the number of proteins for which a method estimated differences (see Figure \@ref(fig:FDRfdp) Panel A), which indicates how robust the procedure works in the presence of missing observations.
+When comparing differential expression analysis performance, a relevant parameter is the number of proteins for which a method estimates differences (see Figure \@ref(fig:FDRfdp) Panel A).
+The number of proteins indicates how robust the procedure works in the presence of missing observations.
 For each protein, we tried to determine four differences ($\Delta = (1.20, 1.25, 1.30, 1.50)$).
@@
 Interestingly, the $pAUC$ increases when using $p$-values instead of the $t$-statistics for linear models, while it decreases for mixed effect models.
-The reason is an erroneous denominator degrees of freedom estimation for many proteins in the case of the mixed effect models.
+The reason is an erroneous denominator degrees of freedom estimation for many proteins in the case of the mixed-effect models.
 

### 2022-06-07 inst/MyArticle/paper.Rmd

--- 2022-06-07 inst/MyArticle/paper.Rmd
@@
 ---
-title: prolfqua - A Comprehensive R-package for Proteomics Differential Expression Analysis
+title: prolfqua&colon; A Comprehensive R-package for Proteomics Differential Expression Analysis
 author: 

### 2022-06-07 inst/MyArticle/paper.Rmd
Acknowledgements update
conclusions

@@
 We use the formula interface to specify the models, making it easy to reproduce an analysis performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` in other statistical methods programming languages.
-The developed framework is flexible to integrate other modeling methods, e.g., the probabilistic dropout model [@bioxrvproDA2020] or accurate variance estimation [@zhu2020deqms].
+However, the developed framework is flexible enough to in the future integrate other modeling methods, e.g., the probabilistic dropout model [@bioxrvproDA2020] or accurate variance estimation [@zhu2020deqms].
+Hence, `r BiocStyle::Githubpkg("fgcz/prolfqua")` enables you to call various methods and makes selecting the best differential expression analysis algorithm for your problem easy.
 
@@
 It is relevant that an analysis pipeline shows good performance in all these categories.
-Leveraging these computer experiments, we can provide the following advice.
-First, estimate protein abundances from peptide abundances using a robust or nonparametric regression method.
-Second, use linear models because they show good performance in all categories.
-If the measurements are correlated, as for technical replicates, mixed effect models might work if the sample sizes are large; if not, aggregate the replicates and fit a linear model instead.
-Third, if you use fixed-effect linear models, apply variance moderation to improve the $t$-statistics and $p$-value estimates.
-If you want to sort your protein lists to perform gene set enrichment analysis, use the $t$-statistic instead of the difference.
-Last but not least, do not impute missing observation but rather statistically model missingness to estimate parameters, i.e., group differences.
+Leveraging these computer experiments, we can provide the following advice: i) estimate protein abundances from peptide abundances using a robust or nonparametric regression method; ii) use linear models because they show good performance in all categories; iii) if the measurements are correlated, as for technical replicates, mixed effect models might work if the sample sizes are large; if not, aggregate the replicates and fit a linear model instead; iv) if you use fixed-effect linear models, apply variance moderation to improve the $t$-statistics and $p$-value estimates; v) If you want to sort your protein lists to perform gene set enrichment analysis, use the $t$-statistic instead of the difference; vi) do not impute missing observation but statistically model missingness to estimate parameters, i.e., group differences.
+Finally, the differential expression analysis result obtained with prolfqua are comparable to or better than when using other differential expression analysis tools.
 
-In summary, `r BiocStyle::Githubpkg("fgcz/prolfqua")` is an easy-to-use and feature-rich R package to analyze quantitative mass spectrometric data with simple and complex experimental designs.
-It also can generate conclusive reports, and benchmark MS software and statistical methods.
-Furthermore, with very little adaptations this R package can also be adapted to analyze different quantitative proteomics data (e.g.
-labeling based TMT-, PRM- DIA-data or even other omics scale datasets such as microarray data).
+In summary, `r BiocStyle::Githubpkg("fgcz/prolfqua")` is an easy-to-use and feature-rich R package to analyze quantitative mass spectrometric data with simple or complex experimental designs.
+It also can generate conclusive reports and to benchmark MS software and statistical methods.
+Furthermore, with minimal adaptations, this R package can analyze different quantitative proteomics data (e.g., labeling-based TMT-, PRM- DIA-data).
 We provide documentation, in vignette format, at the website https://github.com/fgcz/prolfqua/.
 This document was created using Rmarkdown.
-All the code needed to replicate the document or the benchmark results is available at : https://github.com/wolski/prolfquabenchmark.
+All the code needed to replicate the document or the benchmark results is available at: https://github.com/wolski/prolfquabenchmark.
 
@@
 
-The authors thank the technology platform fund (TPF) of the University of Zurich.
+The authors thank the technology platform fund (TPF) of the University of Zurich and all FGCZ proteomics colleagues for fruitful discussions.
 

### 2022-06-07 inst/MyArticle/paper.Rmd
temp

@@
 
-```{r FDRfdp, fig.cap = "(ref:FDRfdp)", out.width = '90%', fig.width=10, fig.height=10}
+```{r FDRfdp, fig.cap = "(ref:FDRfdp)", out.width = '90%', fig.width=12, fig.height=10}
 dd <- purrr::map_df(allBenchmarks, function(x){res <- x$smc$summary; res$name <- x$model_name;res})

### 2022-06-09 inst/MyArticle/paper.Rmd
automatic path

@@
 bibliography: bibliography.bib
+vignette: >
+  %\VignetteIndexEntry{prolfqua&colon; A Comprehensive R-package for Proteomics Differential Expression Analysis} 
+  %\VignetteEncoding{UTF-8}
+  %\VignetteEngine{knitr::rmarkdown}
 editor_options: 
@@
 $DH5\alpha$ _E.
-coli_ lysate was spiked in human pancreatic cancer cells (Panc-1) lysate at five different levels: $3\%$ _E.
-coli_, $4.5\%$ _E.
-coli_, $6\%$ _E.
-coli_, $7.5\%$ _E.
-coli_ and $9\%$ _E.
+coli_ lysate was spiked in human pancreatic cancer cells (Panc-1) lysate at five different levels: $3\%$, $4.5\%$, $6\%$, $7.5\%$ and $9\%$ _E.
 coli_.
@@
 ```{r readBenchmarkData}
-allBenchmarks <- readRDS("../../inst/Benchresults/allBenchmarks.RDS")
-benchmark_msstats <- readRDS("../../inst/Benchresults/benchmark_msstats.RDS")
-benchmark_prodA <- readRDS("../../inst/Benchresults/benchmark_medpolish_proDA.RDS")
-msFragger <- readRDS(file = "../../inst/Benchresults/MSFragger_medpol_benchmark.RDS")
+
+getpath <- function(filN){
+    f1 <- system.file("Benchresults",filN, package = "prolfquabenchmark")
+    if(f1 == ""){
+        f1 <- paste0("../../inst/Benchresults/",filN)
+    }
+    return(f1)
+}
+allBenchmarks <- readRDS(getpath("allBenchmarks.RDS"))
+benchmark_msstats <- readRDS(getpath("benchmark_msstats.RDS"))
+benchmark_prodA <- readRDS(getpath("benchmark_medpolish_proDA.RDS"))
+msFragger <- readRDS(getpath("MSFragger_medpol_benchmark.RDS"))
 
@@
 
-\begin{table}[H]
+\begin{table}
 \centering

### 2022-06-14 inst/MyArticle/PosterBioconductor.Rmd
New hexsticker

@@
     orcid: '0000-0003-1975-3064'
-
 affiliation:
@@
 body_textsize: "44px"
-logoleft_name: ../Figures/hexStickerProlfqua.png
+logoleft_name: hexStickerProlfqua.png
 logoright_name: FGCZLogo.png
@@
 bibliography: bibliography.bib
-
+editor_options: 
+  chunk_output_type: console
 ---
@@
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 7, fig.height = 5)
+library(ggplot2)
+library(prolfqua)
+library(tidyverse)
 
@@
 ```{r LFQData, echo=FALSE, fig.cap="Class Diagram for LFQData and related objects. The LFQData_Plotter class uses the LFQData class to implement methods for plotting.", out.width = '90%'}
-knitr::include_graphics("LFQData.png")
+knitr::include_graphics("LFQData_UML_V2.png")
 ```
@@
 # compute contrasts from linear model and with imputation
-contr <- Contrasts$new(models, contrasts) |> 
-  ContrastsModerated$new()
-conI <- ContrastsSimpleImpute$new( lfqp, contrasts) |> 
-  ContrastsModerated$new()
-tmp <- addContrastResults(contr, conI)
+contr <- prolfqua::Contrasts$new(models, contrasts) |> 
+  prolfqua::ContrastsModerated$new()
+conI <- prolfqua::ContrastsSimpleImpute$new( lfqp, contrasts) |> 
+  prolfqua::ContrastsModerated$new()
+tmp <- prolfqua::addContrastResults(contr, conI)
 # visualize results

### 2022-06-14 inst/MyArticle/paper.Rmd
New hexsticker

@@
                       fig.height = 5)
+ggplot2::theme_set(ggplot2::theme_classic())
 
@@
 getpath <- function(filN){
+    f1 <- paste0("../../inst/Benchresults/",filN)
+    if(f1 == ""){
         f1 <- system.file("Benchresults",filN, package = "prolfquabenchmark")
-    if(f1 == ""){
-        f1 <- paste0("../../inst/Benchresults/",filN)
     }
+    message(f1)
     return(f1)
@@
 gridExtra::grid.arrange(p1_pAUC, pB, pC, layout_matrix=hlay)
+
 ```
@@
 
-file.path(system.file(package = "prolfqua"), 
-          "Figures/hexStickerProlfqua.png") |>
+file.path("hexStickerProlfqua.png") |>
   knitr::include_graphics()

### 2022-06-16 inst/MyArticle/paper.Rmd
rename table.

@@
     proDA = c("proDA", "probabilistic dropout model",  "protein", "peptide.txt"),
-    imputation = c("prot_imputation", "ContrastsSimpleImpute, ContrastsModerated", "protein",  "peptide.txt"),
-    lmmed = c("prot_med_lm_moderated", "strategy_lm, Contrasts, ContrastsModerated",  "protein", "peptide.txt"),
-    prot_merged = c("prot_merged", "addContrastResults( prefer = prot_med_lm_moderated , add = prot_imputation) *",  "protein", "peptide.txt"),
-    mixed = c("prot_mixed_effect_moderated", "strategy_lmer, Contrasts, ContrastsModerated", "peptide", "peptide.txt" ),
-    ropeca = c("ropeca", "strategy_lm, Contrasts, ContrastsModerated, ContrastsROPECA", "peptide", "peptide.txt" )
+    imputation = c("prolfqua_impute", "ContrastsSimpleImpute, ContrastsModerated", "protein",  "peptide.txt"),
+    lmmed = c("prolfqua_lm_mod", "strategy_lm, Contrasts, ContrastsModerated",  "protein", "peptide.txt"),
+    prot_merged = c("prolfqua_merged", "addContrastResults( prefer = prot_med_lm_moderated , add = prot_imputation) *",  "protein", "peptide.txt"),
+    mixed = c("prolfqua_mix_eff_mod", "strategy_lmer, Contrasts, ContrastsModerated", "peptide", "peptide.txt" ),
+    ropeca = c("prolfqua_ropeca", "strategy_lm, Contrasts, ContrastsModerated, ContrastsROPECA", "peptide", "peptide.txt" )
 )

### 2022-07-06 inst/MyArticle/paper.Rmd
simplify

@@
 Table \@ref(tab:describeBenchmarkedModels) summarizes which methods we have evaluated, which MaxQuant results we used, and if the models are fitted to peptide or protein intensities.
-We make the R-markdown files to replicate the benchmarking available at `r BiocStyle::Githubpkg("wolski/prolfquabenchamrk")` and at [BenchmarkingIonstarData](https://wolski.github.io/prolfquabenchmark).
+We make the R-markdown files to replicate the benchmarking available at `r BiocStyle::Githubpkg("wolski/prolfquabenchmark")`.
 
@@
 # read combined_protein.tsv 
-protein <- prolfqua::tidy_MSFragger_combined_protein(protein)
+protein <- prolfqua::tidy_MSFragger_combined_protein_deprec(protein)
 # remove proteins identified by a single peptide

### 2022-07-06 inst/MyArticle/paper.Rmd
fix

@@
 # read combined_protein.tsv 
-protein <- prolfqua::tidy_MSFragger_combined_protein_deprec(protein)
+protein <- prolfqua::tidy_FragPipe_combined_protein_deprec(protein)
 # remove proteins identified by a single peptide

### 2022-07-15 inst/MyArticle/paper.Rmd
update abstract.

@@
     Mass spectrometry is widely used for quantitative proteomics studies, relative protein quantification, and differential expression analysis of proteins.
-Nevertheless, there is a need for a flexible and easy-to-use application programming interface in R that transparently supports a variety of well principled statistical procedures.
-The `r BiocStyle::Githubpkg("fgcz/prolfqua")` package can model simple experimental designs with a single explanatory variable and complex experiments with multiple factors and hypothesis testing.
-It integrates essential steps of the mass spectrometry-based differential expression analysis workflow: quality control, data normalization, protein aggregation, statistical modeling, hypothesis testing, and sample size estimation.
-The application programmer interface strives to be clear,  predictable, discoverable, and consistent to make proteomics data analysis easy and exciting.
+There is a large variety of signal processing software and analysis tools.
+Nevertheless, there is a need for a modular, easy-to-use application programming interface in R that transparently supports a variety of well principled statistical procedures to make applying them to proteomics data, comparing and understanding their differences easy.
+   The `r Githubpkg("fgcz/prolfqua")` package integrates essential steps of the mass spectrometry-based differential expression analysis workflow: quality control, data normalization, protein aggregation, statistical modeling, hypothesis testing, and sample size estimation.
+The package makes integrating new data formats easy.
+It can be used to model simple experimental designs with a single explanatory variable and complex experiments with multiple factors and hypothesis testing.
+The implemented methods allow sensitive and specific differential expression analysis.
 Furthermore, the package implements benchmark functionality that can help to compare data acquisition, data preprocessing, or data modeling methods using a gold standard dataset.
-Finally, we show that the implemented methods allow sensitive and specific differential expression analysis.
-The `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package is available on GitHub https://github.com/fgcz/prolfqua, distributed under the MIT licence, and runs on all platforms supported by the [R free software environment for statistical computing and graphics](https://www.r-project.org/).
+The application programmer interface of `r Githubpkg("fgcz/prolfqua")`  strives to be clear,  predictable, discoverable, and consistent to make proteomics data analysis easy and exciting.
+      Finally,  The `r Githubpkg("fgcz/prolfqua")` R package is available on GitHub https://github.com/fgcz/prolfqua, distributed under the MIT license.
+It runs on all platforms supported by the [R free software environment for statistical computing and graphics](https://www.r-project.org/).
+
 urlcolor: blue
@@
 The modular design of `r BiocStyle::Githubpkg("fgcz/prolfqua")`, allows for adding new features, e.g., generalized linear models (_glm_'s) to model the presence or absence information of a protein, or robust linear models (_rlm_'s), in the future.
-R's formula interface for linear models is flexible, widely used, and well documented [@faraway2016extending].
+The developed framework is flexible enough to in the future integrate other modeling methods, e.g., the probabilistic dropout model [@bioxrvproDA2020] or accurate variance estimation [@zhu2020deqms].
+(While other R packages typically only implement one modelling approach.) Hence, `r BiocStyle::Githubpkg("fgcz/prolfqua")` enables you to call various methods and makes selecting the best differential expression analysis algorithm for your problem easy.
+
+# R's formula interface for linear models is flexible, widely used, and well documented [@faraway2016extending].
 We use the formula interface to specify the models, making it easy to reproduce an analysis performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` in other statistical methods programming languages.
-However, the developed framework is flexible enough to in the future integrate other modeling methods, e.g., the probabilistic dropout model [@bioxrvproDA2020] or accurate variance estimation [@zhu2020deqms].
-Hence, `r BiocStyle::Githubpkg("fgcz/prolfqua")` enables you to call various methods and makes selecting the best differential expression analysis algorithm for your problem easy.
 
-When comparing statistical modeling methods for the differential expression analysis, we assessed performance measures such as the number of estimated contrasts, the $pAUC$, and if the FDR is an unbiased estimate of the FDR.
+When comparing statistical modeling methods for the differential expression analysis, we assessed performance measures such as the number of estimated contrasts, the $pAUC$, and if the FDR is an unbiased estimate of the FDP.
 It is relevant that an analysis pipeline shows good performance in all these categories.
@@
 Finally, the differential expression analysis result obtained with prolfqua are comparable to or better than when using other differential expression analysis tools.
+All the code needed to replicate the benchmark results and this document is available at: https://github.com/wolski/prolfquabenchmark.
 
-In summary, `r BiocStyle::Githubpkg("fgcz/prolfqua")` is an easy-to-use and feature-rich R package to analyze quantitative mass spectrometric data with simple or complex experimental designs.
+(merge with first paragraph of the Conclusions)
+`r BiocStyle::Githubpkg("fgcz/prolfqua")` is an easy-to-use and feature-rich R package to analyze quantitative mass spectrometric data with simple or complex experimental designs.
 It also can generate conclusive reports and to benchmark MS software and statistical methods.
 Furthermore, with minimal adaptations, this R package can analyze different quantitative proteomics data (e.g., labeling-based TMT-, PRM- DIA-data).
-We provide documentation, in vignette format, at the website https://github.com/fgcz/prolfqua/.
-This document was created using Rmarkdown.
-All the code needed to replicate the document or the benchmark results is available at: https://github.com/wolski/prolfquabenchmark.
 

### 2022-07-20 inst/MyArticle/paper.Rmd
updated paper

@@
 Nevertheless, there is a need for a modular, easy-to-use application programming interface in R that transparently supports a variety of well principled statistical procedures to make applying them to proteomics data, comparing and understanding their differences easy.
-   The `r Githubpkg("fgcz/prolfqua")` package integrates essential steps of the mass spectrometry-based differential expression analysis workflow: quality control, data normalization, protein aggregation, statistical modeling, hypothesis testing, and sample size estimation.
+The `r BiocStyle::Githubpkg("fgcz/prolfqua")` package integrates essential steps of the mass spectrometry-based differential expression analysis workflow: quality control, data normalization, protein aggregation, statistical modeling, hypothesis testing, and sample size estimation.
 The package makes integrating new data formats easy.
@@
 Furthermore, the package implements benchmark functionality that can help to compare data acquisition, data preprocessing, or data modeling methods using a gold standard dataset.
-The application programmer interface of `r Githubpkg("fgcz/prolfqua")`  strives to be clear,  predictable, discoverable, and consistent to make proteomics data analysis easy and exciting.
-      Finally,  The `r Githubpkg("fgcz/prolfqua")` R package is available on GitHub https://github.com/fgcz/prolfqua, distributed under the MIT license.
+The application programmer interface of `r BiocStyle::Githubpkg("fgcz/prolfqua")`  strives to be clear,  predictable, discoverable, and consistent to make proteomics data analysis easy and exciting.
+Finally, the `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package is available on GitHub https://github.com/fgcz/prolfqua, distributed under the MIT license.
 It runs on all platforms supported by the [R free software environment for statistical computing and graphics](https://www.r-project.org/).
@@
 ggplot2::theme_set(ggplot2::theme_classic())
-
+library(BiocStyle)
 options(width=70)
@@
 rownames(bb) <- names(xx)
-knitr::kable(data.frame(bb), caption = "Rows - R package, Columns - types of models supported: pd - parallel design, rm - repeated measurements, fd - factorial design, int - interactions among factors, mem - mixed effect models, eb - empirical Bayes, md - missing data modelling.
+knitr::kable(data.frame(bb), caption = "Rows - R packages, Columns - types of models supported: pd - parallel design, rm - repeated measurements, fd - factorial design, int - interactions among factors, mem - mixed effect models, eb - empirical Bayes, md - missing data modelling.
 Y - yes, * - repeated measurements are modeled using a fixed effect.
@@
 Furthermore, the statistical model must provide an unbiased estimate of the false discovery rate (FDR) to manage expectations when selecting protein lists for follow-up experiments.
-We will use the partial area under the receiver operator curve (ROC) to assess the power of the tests and compare the FDR with the false discovery proportion (FDP).
+We will use the partial area under the receiver operator curve (ROC) to assess the power of the tests and compare the FDR with the false discovery proportion ($FDP$).
 
@@
 (ref:LFQData) Class Diagram of classes representing the proteomics data.
-The LFQData class encapsulates the quantitative proteomics data stored in a tidy table.
+The `LFQData` class encapsulates the quantitative proteomics data stored in a tidy table.
 An instance of the AnalysisTableConfiguration class specifies a mapping of columns in the tidy table.
-The LFQDataPlotter class and other classes decorate the LFQData class with additional functionality.
+The `LFQDataPlotter` class and other classes decorate the `LFQData` class with additional functionality.
 Similarly, the `LFQDataStats` and `LFQDataSummary` reference the `LFQData` class and group methods for variance and sample size estimation or summarizing peptide and protein counts.
@@
 ```{r LFQData, echo=FALSE, fig.cap="(ref:LFQData)", out.width = '90%'}
-knitr::include_graphics("LFQData_UML_V2.png")
+knitr::include_graphics("LFQData_UML_LARGE.png")
 ```
 
+R's formula interface for linear models is flexible, widely used, and well documented [@faraway2016extending].
+We use the formula interface to specify the models, making it easy to reproduce an analysis performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` in other statistical methods programming languages.
 In addition, we implement features specific to high throughput experiments, such as the experimental Bayes variance and $p$-value moderation, which utilizes the parallel structure of the protein measurements and the analysis [@Ritchie2015].
@@
 
-knitr::include_graphics("ContrastClassesUML.png")
+knitr::include_graphics("ContrastClassesUML_LARGE.png")
 ```
@@
 By plotting the $TPR$ versus the $FPR$ we obtain the receiver operator characteristic curve (ROC curve).
-The area under the curve (AUC) or partial areas under the curve (pAUC), at various values of the $FPR$, are measures of performance derived from the ROC curve.
+The area under the curve ($AUC$) or partial areas under the curve ($pAUC$), at various values of the $FPR$, are measures of performance derived from the ROC curve.
 By using these measures we can compare the performances of the statistics produced by the various methods examined.
@@
 
-(ref:scaling) Panel A - Peptide intensity distributions for $20$ samples.
+(ref:scaling) Panel A - Density plot of peptide intensity distributions for $20$ samples.
 For each sample a line with a different colour is shown.
-Panel B - Peptide intensities for protein _5NTD_ are shown using a line of different colour, and the protein intensity estimate is shown using a fat black line, Panel C - distribution of standard deviations of all proteins in each dilution group ($a$, $b$, $c$, $d$, $e$) and overall (all), Panel D - Distribution of protein intensities for protein _5NTD_.
+Panel B - Peptide intensities for protein _5NTD_ are shown using a line of different colour, and the protein intensity estimate is shown using a fat black line, Panel C - distribution of standard deviations of all proteins in each dilution group ($a$, $b$, $c$, $d$, $e$) and overall (all), Panel D - Distribution of protein intensities of Protein _5NTD_ in each dilution group.
 
@@
 (ref:exampleContrasts) Panel A - Histogram showing the distribution of p-values for $163$ proteins.
-Panel B -  Volcano plot showing $-\log_{10}$ transformed FDR as function of the difference between groups for $163$ proteins.
+Panel B -  Volcano plot showing $-\log_{10}$ transformed $FDR$ as function of the difference between groups for $163$ proteins.
 
@@
 Auto-completion support in the Rstudio editor makes it easy for novices to explore `r BiocStyle::Githubpkg("fgcz/prolfqua")`s functionality (Figure \@ref(fig:tabCompletion)).
-To simplify integration of `r BiocStyle::Githubpkg("fgcz/prolfqua")` with Bioconductor based workflows there is a method to convert the LFQData class into a `r BiocStyle::Biocpkg("SummarizedExperiment")`.
+To simplify integration of `r BiocStyle::Githubpkg("fgcz/prolfqua")` with Bioconductor based workflows there is a method to convert the `LFQData` class into a `r BiocStyle::Biocpkg("SummarizedExperiment")`.
 
@@
 
-The Benchmark functionality of `r BiocStyle::Githubpkg("fgcz/prolfqua")` includes receiver operator curves (ROC) and computes partial areas under those curves (pAUC) and other scores, e.g., the false discovery proportion FDP.
+The Benchmark functionality of `r BiocStyle::Githubpkg("fgcz/prolfqua")` includes receiver operator curves (ROC) and computes partial areas under those curves ($pAUC$) and other scores, e.g., the false discovery proportion $FDP$.
 We use those scores, i.e.
-the $pAUC$ at $10%$ FDR and the FDP, to examine how well the methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` model quantitative mass spectrometric high throughput data and compare them with results produced by `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`.
+the $pAUC$ at $10%$ $FDR$ and the $FDP$, to examine how well the methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` model quantitative mass spectrometric high throughput data and compare them with results produced by `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`.
 Table \@ref(tab:describeBenchmarkedModels) summarizes all modelling methods we evaluated.
@@
 However, some methods can not estimate all of them.
-The set of proteins with effect size estimates might differ for each method, biasing direct comparison of scores such as pAUC.
+The set of proteins with effect size estimates might differ for each method, biasing direct comparison of scores such as $pAUC$.
 For instance, we observe that `r BiocStyle::Biocpkg("MSstats")` estimates $16058$ group differences while the mixed effect models estimates the fewest with $15940$.
-Hence, to conclude that one method shows a better performance, it does not suffice if the pAUC is greater, but the number of proteins with differential expression results needs to be equal or larger.
+Hence, to conclude that one method shows a better performance, it does not suffice if the $pAUC$ is greater, but the number of proteins with differential expression results needs to be equal or larger.
 
@@
 We also benchmark if the $FDR$ obtained from a model is an unbiased estimate of the false discovery proportion $FDP$.
-Figure \@ref(fig:FDRfdp) Panel C draws on the horizontal axis the FDR determined from the model, and on the vertical axis, the FDP obtained from the confusion matrix.
-Most lines are below the diagonal, indicating that the FDR estimates are modestly conservative for this benchmark dataset.
+Figure \@ref(fig:FDRfdp) Panel C draws on the horizontal axis the $FDR$ determined from the model, and on the vertical axis, the $FDP$ obtained from the confusion matrix.
+Most lines are below the diagonal, indicating that the $FDR$ estimates are modestly conservative for this benchmark dataset.
 In the case of `r BiocStyle::Biocpkg("MSstats")`, we observe a high proportion of false discoveries for small $FDR$ values.
-In the case of `r BiocStyle::Biocpkg("PECA")`, the FDR estimates, obtained by Benjamini-Hochberg correcting the regulation probabilities, strongly overestimate the $FDP$.
+In the case of `r BiocStyle::Biocpkg("PECA")`, the $FDR$ estimates, obtained by Benjamini-Hochberg correcting the regulation probabilities, strongly overestimate the $FDP$.
 
@@
 The red line indicates the average area under the curve of all methods.
-Panel C - Plots the false discovery proportion (FDP) as a function of the FDR.
-Ideally, the FDR should be equal to the FDP.
+Panel C - Plots the false discovery proportion ($FDP$) as a function of the $FDR$.
+Ideally, the $FDR$ should be equal to the $FDP$.
 Therefore larger distances from the diagonal are worse.
@@
 colnames(xM) <- c("prolfqua functions", "model")
-knitr::kable(tibble::as_tibble(xM), caption = "Prolfqua functions to fit various models.",
+knitr::kable(tibble::as_tibble(xM), caption = "Prolfqua functions that can be used to fit various models.",
              format = "latex", booktabs=TRUE) |>
@@
 
-```{r describeBenchmarkedModels}
+```{r describeBenchmarkedModels, results='asis' }
 xMM <- data.frame(
@@
     lmmed = c("prolfqua_lm_mod", "strategy_lm, Contrasts, ContrastsModerated",  "protein", "peptide.txt"),
-    prot_merged = c("prolfqua_merged", "addContrastResults( prefer = prot_med_lm_moderated , add = prot_imputation) *",  "protein", "peptide.txt"),
+    prot_merged = c("prolfqua_merged", "addContrastResults( prefer = prot_med_lm_moderated , add = prot_imputation)HACKIT ",  "protein", "peptide.txt"),
     mixed = c("prolfqua_mix_eff_mod", "strategy_lmer, Contrasts, ContrastsModerated", "peptide", "peptide.txt" ),
@@
 colnames(xMM) <-  c("Label","Description","Abundance","Input File")
-kableExtra::kable(xMM,
+tmp <- kableExtra::kable(xMM,
              caption = paste0("All benchmarked models. ",
@@
     kableExtra::kable_styling(latex_options="scale_down") |> 
-    kableExtra::add_footnote(label = "Merges results of the 'prot_med_lm_moderated' and 'prot_imputation' modeling pipeline, preferring those of 'prot_med_lm_moderated' if available.")
+    kableExtra::add_footnote(label = "prolfqua_merged - Merges results of the prot_med_lm_moderated and prot_imputation modeling pipeline, preferring those of prot_med_lm_moderated if available.")
 
+knitr::asis_output(stringr::str_replace(tmp, "HACKIT", "$^{a}$"))
 ```
@@
 However, the properties of Beta-based probabilities are not well understood; for instance, the p-values are not uniformly distributed under the null hypothesis (not shown).
-Furthermore, the FDR estimates obtained when correcting for multiple testing with the Benjamini-Hochberg method are biased and overestimate the false discovery proportion (see Figure \@ref(fig:FDRfdp) Panel C).
-Therefore, we can not recommend this method if an unbiased estimate of FDR is essential, which is frequently the case.
+Furthermore, the $FDR$ estimates obtained when correcting for multiple testing with the Benjamini-Hochberg method are biased and overestimate the false discovery proportion (see Figure \@ref(fig:FDRfdp) Panel C).
+Therefore, we can not recommend this method if an unbiased estimate of $FDR$ is essential, which is frequently the case.
 In addition, peptides are stronger affected by missing values, reducing the number of contrasts we could estimate for the dataset using this method (see Figure \@ref(fig:FDRfdp) Panel C).
@@
 Despite imputation, `r BiocStyle::Biocpkg("MSstats")` does not estimate group differences for more proteins and does not achieve a higher $pAUC$ score than `r BiocStyle::Githubpkg("fgcz/prolfqua")`.
-Furthermore, Figure \@ref(fig:FDRfdp) Panel C shows that when using `r BiocStyle::Biocpkg("MSstats")` the proportion of false discoveries might be very high even when filtering for a low FDR.
+Furthermore, Figure \@ref(fig:FDRfdp) Panel C shows that when using `r BiocStyle::Biocpkg("MSstats")` the proportion of false discoveries might be very high even when filtering for a low $FDR$.
 
@@
 
-`r BiocStyle::Githubpkg("fgcz/prolfqua")` allows for considerable flexibility to model quantitative proteomics experiments.
-Various types of models are available (see Figure \@ref(fig:ContrastUML) and Table \@ref(tab:prolfquaModels)), and the contrast specification is explicit and consistent for all models.
-The modular design of `r BiocStyle::Githubpkg("fgcz/prolfqua")`, allows for adding new features, e.g., generalized linear models (_glm_'s) to model the presence or absence information of a protein, or robust linear models (_rlm_'s), in the future.
-The developed framework is flexible enough to in the future integrate other modeling methods, e.g., the probabilistic dropout model [@bioxrvproDA2020] or accurate variance estimation [@zhu2020deqms].
-(While other R packages typically only implement one modelling approach.) Hence, `r BiocStyle::Githubpkg("fgcz/prolfqua")` enables you to call various methods and makes selecting the best differential expression analysis algorithm for your problem easy.
+`r BiocStyle::Githubpkg("fgcz/prolfqua")` is an easy-to-use and feature-rich R package to analyze quantitative mass spectrometric data with simple or complex experimental designs.
+While other R packages for differential expression analysis of proteins typically only implement one modeling approach, `r BiocStyle::Githubpkg("fgcz/prolfqua")` supports various types of models (see Figure 2 and Table 3).
+The contrast specification is explicit and consistent for all models.
+The modular design of `r BiocStyle::Githubpkg("fgcz/prolfqua")` allows for adding new features, e.g., generalized linear models to model the presence or absence information of a protein, or robust linear models, in the future.
+The developed framework is flexible enough to integrate other modeling methods, e.g., the probabilistic dropout model [@bioxrvproDA2020] or accurate variance estimation [@zhu2020deqms].
+Hence, `r BiocStyle::Githubpkg("fgcz/prolfqua")` enables you to call various methods and quickly select the best differential expression analysis algorithm for your problem.
+It also can generate conclusive reports and can be used to benchmark MS software and statistical methods.
+Furthermore, with minimal adaptations, this R package can analyze different quantitative proteomics data (e.g., labeling-based TMT-, PRM- DIA-data).
 
-# R's formula interface for linear models is flexible, widely used, and well documented [@faraway2016extending].
-We use the formula interface to specify the models, making it easy to reproduce an analysis performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` in other statistical methods programming languages.
-
-When comparing statistical modeling methods for the differential expression analysis, we assessed performance measures such as the number of estimated contrasts, the $pAUC$, and if the FDR is an unbiased estimate of the FDP.
+When comparing statistical modeling methods for the differential expression analysis, we assessed performance measures such as the number of estimated contrasts, the $pAUC$, and if the $FDR$ is an unbiased estimate of the $FDP$.
 It is relevant that an analysis pipeline shows good performance in all these categories.
 Leveraging these computer experiments, we can provide the following advice: i) estimate protein abundances from peptide abundances using a robust or nonparametric regression method; ii) use linear models because they show good performance in all categories; iii) if the measurements are correlated, as for technical replicates, mixed effect models might work if the sample sizes are large; if not, aggregate the replicates and fit a linear model instead; iv) if you use fixed-effect linear models, apply variance moderation to improve the $t$-statistics and $p$-value estimates; v) If you want to sort your protein lists to perform gene set enrichment analysis, use the $t$-statistic instead of the difference; vi) do not impute missing observation but statistically model missingness to estimate parameters, i.e., group differences.
-Finally, the differential expression analysis result obtained with prolfqua are comparable to or better than when using other differential expression analysis tools.
-All the code needed to replicate the benchmark results and this document is available at: https://github.com/wolski/prolfquabenchmark.
-
-(merge with first paragraph of the Conclusions)
-`r BiocStyle::Githubpkg("fgcz/prolfqua")` is an easy-to-use and feature-rich R package to analyze quantitative mass spectrometric data with simple or complex experimental designs.
-It also can generate conclusive reports and to benchmark MS software and statistical methods.
-Furthermore, with minimal adaptations, this R package can analyze different quantitative proteomics data (e.g., labeling-based TMT-, PRM- DIA-data).
+Finally, the differential expression analysis results obtained with prolfqua are comparable to or better than those of other differential expression analysis tools.
 
@@
 # Appendix  {.unnumbered}
+
+## Benchmark
+
+All the code needed to replicate the benchmark results and this document is available at:
+
+https://github.com/wolski/prolfquabenchmark.
+
+- [Benchmarking of methods implemented in prolfqua](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkingIonstarData.html)
+- [Benchmarking of MSstats](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_MSStats.html)
+- [Benchmarking of proDA](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_proDA_medpolish.html)
 

### 2022-07-20 inst/MyArticle/SIB.Rmd
add figs

---
title: prolfqua&colon; A Comprehensive R-package for Proteomics Differential Expression Analysis
author: 
  - name: Witold Wolski
    affil: 1; 2
    orcid: '0000-0002-6468-120X'
  - name: Paolo Nanni
    affil: 1
    orcid: '0000-0001-8429-3557'
  - name: Jonas Grossmann
    affil: 1; 2
    orcid: '0000-0002-6899-9020'
  - name: Maria D'Errico
    affil: 1; 2
    orcid: '0000-0001-9467-9058'
  - name: Ralph Schlapbach
    affil: 1
    orcid: '0000-0002-7488-4262'
  - name: Christian Panse
    affil: 1; 2
    orcid: '0000-0003-1975-3064'
affiliation:
  - num: 1
    address: Functional Genomics Center Zurich - ETH Zurich/University of Zurich (https://www.fgcz.ch/)
  - num: 2
    address: Swiss Institute of Bioinformatics (https://www.sib.swiss/)
column_numbers: 4
poster_height: "1009mm"
poster_width: "1426mm"
title_textsize:	"75pt"
author_textsize: "35pt"
font_family: Monaco
primary_colour: "cornflowerblue"
secondary_colour: "darkblue"
accent_colour: "slateblue"
titletext_fontfamily: Arial
logoleft_name: hexstrickerPRL2.png
logoright_name: FGCZLogo.png
output: 
  posterdown::posterdown_html
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
options(width = 999)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 7, fig.height = 5)
#library(ggplot2)
#library(prolfqua)
#library(tidyverse)
ggplot2::theme_set(ggplot2::theme_classic())
#ggplot2::theme_set(ggplot2::theme_gray())

```

# Implementation

(ref:LFQData) Class Diagram of classes representing the proteomics data. The LFQData class encapsulates the quantitative proteomics data stored in a tidy table. An instance of the AnalysisTableConfiguration class specifies a mapping of columns in the tidy table. The LFQDataPlotter class and other classes decorate the LFQData class with additional functionality. Similarly, the `LFQDataStats` and `LFQDataSummary` reference the `LFQData` class and group methods for variance and sample size estimation or summarizing peptide and protein counts.

```{r LFQData, echo=FALSE, fig.cap="(ref:LFQData)", out.width = '100%'}
knitr::include_graphics("LFQData_UML_V2.png")
```

(ref:ContrastUML) The UML (unified modeling language) diagram of modeling and contrast related classes. Different strategies, e.g., _lm_, _lmer_, and _glm_, can fit various types of models. The model builder method fits the statistical model given the data and a strategy. The obtained model can then be used for the analysis of variance or to estimate contrasts. All classes estimating contrasts implement the _Contrasts_ interface. Results of external tools, e.g., SAINTexpress, can be adapted by implementing the Contrasts Interface.

```{r ContrastUML, echo=FALSE, fig.cap="(ref:ContrastUML)", out.width = '100%'}
knitr::include_graphics("ContrastClassesUML.png")
```

<br/>
<br/>

# Example Analysis

```{r prepro,  echo = TRUE , fig.cap="(ref:scaling)", out.width = '90%', fig.height=10}
## load peptide abundances and configuration
d <- prolfqua::prolfqua_data('data_ionstar')$filtered()
## create R6 obejct
lfqd <- prolfqua::LFQData$new(d$data, d$config) 
##  transform intensities
t <- lfqd$get_Transformer()
lfqd <-  t$log2()$robscale()$lfq
lfqd$rename_response("peptide_abundance")
## infer protein intensities from peptide intensity
agr <- lfqd$get_Aggregator()
lfqp <- agr$medpolish()
lfqp$rename_response("protein_abundance")
## plot panels A-D
pl <- lfqd$get_Plotter()
panelA <- pl$intensity_distribution_density() + 
    ggplot2::labs(tag = "A") + 
    ggplot2::theme(legend.position = "none")

panelB <- agr$plot()$plots[[1]] + ggplot2::labs(tag = "B")
panelC <- lfqp$get_Stats()$violin() + 
    ggplot2::labs(tag = "C")
pl <- lfqp$get_Plotter()
panelD <- pl$boxplots()$boxplot[[1]] + 
    ggplot2::labs(tag = "D")

## specify differences among groups
contrasts <- c(
    "dilution_(9/7.5)_1.2" = "dilution.e - dilution.d",
    "dilution_(7.5/6)_1.25" = "dilution.d - dilution.c" )
## fit model
lmmodel <- paste(lfqp$intensity_column()," ~ dilution.")
modelFunction <- 
    prolfqua::strategy_lm(lmmodel, model_name = "lm")
models <- prolfqua::build_model(lfqp, modelFunction)
## compute differences from linear model
## and moderate variance
contr <- prolfqua::Contrasts$new(models, contrasts) |> 
    prolfqua::ContrastsModerated$new()
## compute differences using imputation
conI <- 
    prolfqua::ContrastsSimpleImpute$new(lfqp, contrasts) |> 
    prolfqua::ContrastsModerated$new()
## merge contrasts, to obtain differences for all proteins
contrasts <- prolfqua::addContrastResults(
    prefer = contr, add = conI)
## plot panel E and F
pl <- contrasts$merged$get_Plotter() 
panelE <- pl$histogram()$p.value + ggplot2::labs(tag = "E")                                                        
panelF <- pl$volcano()$FDR +
    ggplot2::theme(legend.position = "bottom") +
    ggplot2::labs(tag = "F")
gridExtra::grid.arrange(panelA, panelB,
                        panelC, panelD,
                        panelE, panelF, ncol = 2)
```

(ref:scaling) Panel A - Peptide intensity distributions for $20$ samples. For each sample a line with a different colour is shown. Panel B - Peptide intensities for protein _5NTD_ are shown using a line of different colour, and the protein intensity estimate is shown using a fat black line, Panel C - distribution of standard deviations of all proteins in each dilution group ($a$, $b$, $c$, $d$, $e$) and overall (all), Panel D - Distribution of protein intensities for protein _5NTD_. Panel E - Histogram showing the distribution of p-values for $163$ proteins. Panel F -  Volcano plot showing $-\log_{10}$ transformed FDR as function of the difference between groups for $163$ proteins.

# Benchmarking

```{r describeBenchmarkedModels}
xMM <- data.frame(
    msstats = c("MSstats", "preprocess with default parameters",  "", "evidence.txt"),
    proDA = c("proDA", "probabilistic dropout model",  "protein", "peptide.txt"),
    imputation = c("prot_imputation", "ContrastsSimpleImpute, ContrastsModerated", "protein",  "peptide.txt"),
    lmmed = c("prot_med_lm_moderated", "strategy_lm, Contrasts, ContrastsModerated",  "protein", "peptide.txt"),
    prot_merged = c("prot_merged", "addContrastResults(prefer = prot_med_lm_moderated , add = prot_imputation) *",  "protein", "peptide.txt"),
    mixed = c("prot_mixed_effect_moderated", "strategy_lmer, Contrasts, ContrastsModerated", "peptide", "peptide.txt" ),
    ropeca = c("ropeca", "strategy_lm, Contrasts, ContrastsModerated, ContrastsROPECA", "peptide", "peptide.txt" )
)

xMM <- data.frame(t(xMM))
rownames(xMM) <- NULL
colnames(xMM) <-  c("Label","Description","Abundance","Input File")
kableExtra::kable(xMM,
                  caption = paste0("All benchmarked models. ",
                                   "Description - prolfqua function names, ",
                                   "Abudances - indicates if model is fitted to peptide or protein abundances, ",
                                   "Input File - name of MaxQuant file used as input."),
                  booktabs=TRUE) |>
    kableExtra::kable_styling(font_size=20) |> 
    kableExtra::add_footnote(label = "Merges results of the 'prot_med_lm_moderated' and 'prot_imputation' modeling pipeline, preferring those of 'prot_med_lm_moderated' if available.")

```

```{r readBenchmarkData}
getpath <- function(filN){
    f1 <- paste0("../../inst/Benchresults/",filN)
    if(f1 == ""){
       f1 <- system.file("Benchresults",filN,
                         package = "prolfquabenchmark")
    }
    return(f1)
}
allBenchmarks <- readRDS(getpath("allBenchmarks.RDS"))
benchmark_msstats <- readRDS(getpath("benchmark_msstats.RDS"))
benchmark_prodA <- readRDS(getpath("benchmark_medpolish_proDA.RDS"))
msFragger <- readRDS(getpath("MSFragger_medpol_benchmark.RDS"))

allBenchmarks$benchmark_mssstats <- benchmark_msstats
allBenchmarks$benchmark_msFragger <- msFragger
allBenchmarks$benchmark_proDA <- benchmark_prodA
allBenchmarks <- allBenchmarks[c("benchmark_imputation","benchmark_ProtModerated",  "benchmark_mixedModerated", "benchmark_ropeca","benchmark_merged","benchmark_mssstats","benchmark_proDA"   )]

```

```{r benchmarkROC}
ttt <- sapply(allBenchmarks, function(x){x$complete(FALSE)})

res <- purrr::map_df(allBenchmarks, function(x){x$pAUC()})

res <- res |> dplyr::mutate(whatfix = dplyr::case_when(what == "scaled.beta.based.significance" ~ "scaled.p.value", TRUE ~ what))

norm <- res |> dplyr::group_by(contrast,whatfix) |>
    dplyr::summarize(meanpAUC_10 = mean(pAUC_10))
res <- dplyr::inner_join(res, norm)
res <- dplyr::mutate(res , pAUC_10n = pAUC_10 - meanpAUC_10)

resAllB <- res |> dplyr::filter(contrast == "all")

p1_pAUC <- ggplot2::ggplot(resAllB, ggplot2::aes(x = Name, y = pAUC_10)) +
    ggplot2::geom_bar(stat = "identity") +
    ggplot2::facet_wrap(~whatfix)  + 
    ggplot2::coord_cartesian(ylim = c(min(resAllB$pAUC_10),max(resAllB$pAUC_10))) + 
    ggplot2::theme(axis.text.x = ggplot2::element_text(angle = -90, vjust = 0.5)) +
    ggplot2::geom_hline(ggplot2::aes(yintercept = meanpAUC_10), color = "red") +
    ggplot2::xlab("") +
    ggplot2::labs(tag = "A")

p2_compare_variousLevels <- ggplot2::ggplot(res, ggplot2::aes(x = contrast, y = pAUC_10n, group = Name)) +
    ggplot2::geom_line(stat = "identity", ggplot2::aes(linetype = Name, color = Name), size = 1) + 
    ggplot2::facet_wrap(~ whatfix, scales = "free") +
    ggplot2::theme(axis.text.x = ggplot2::element_text(angle = -90, vjust = 0.5)) +
    ggplot2::geom_hline(ggplot2::aes(yintercept = 0), color = "red") + 
    ggplot2::theme(legend.position = "bottom", legend.title = ggplot2::element_blank())

```

(ref:FDRfdp) Panel A - Number of estimated contrasts for each modeling method (higher is better). Panel B - Partial area under the ROC curve at $10\%$ FPR ($pAUC_10$) for all contrasts and three different statistics: the difference among groups, the scaled $p$-value (sign(diff) $\cdot$ p.value) and the $t$-statistics (higher is better). The red line indicates the average area under the curve of all methods.  Panel C - Plots the false discovery proportion (FDP) as a function of the FDR. Ideally, the FDR should be equal to the FDP. Therefore larger distances from the diagonal are worse.

```{r FDRfdp, fig.cap = "(ref:FDRfdp)", out.width = '90%', fig.width=11, fig.height=11}
dd <- purrr::map_df(allBenchmarks, function(x){res <- x$smc$summary; res$name <- x$model_name;res})
dd <- dd |> dplyr::mutate(nrcontrasts = protein_Id * (4 - nr_missing))
dds <- dd |> dplyr::group_by(name) |> dplyr::summarize(nrcontrasts = sum(nrcontrasts))
dds$percent <- dds$nrcontrasts/max(dds$nrcontrasts) * 100

nrgg <- dds |> ggplot2::ggplot(ggplot2::aes(x = name, y = nrcontrasts )) + 
    ggplot2::geom_bar(stat = "identity", fill="white", colour = "black") + 
    ggplot2::coord_cartesian(ylim = c(min(dds$nrcontrasts) - 200, max(dds$nrcontrasts) + 10)) +
    ggplot2::theme(axis.text.x = ggplot2::element_text(angle = -90, vjust = 0.5)) +
    ggplot2::geom_text(ggplot2::aes(label = paste0(round(nrcontrasts, digits = 1), paste0("  (",round(percent, digits = 1),"%)"))),
                       vjust = 0, hjust = -0.2, angle = -90) #+ 
pB <- nrgg + ggplot2::labs(tag = "B")

dd <- purrr::map_df(allBenchmarks, function(x){res <- x$get_confusion_FDRvsFDP(); res$name <- x$model_name;res})

ddb <- dplyr::filter(dd, contrast == "dilution_(4.5/3)_1.5")
ddb <- dd |> dplyr::filter(contrast == "dilution_(7.5/6)_1.25")
ddb <- dd |> dplyr::filter(contrast == "all")

pC <- ddb |> ggplot2::ggplot(ggplot2::aes(y = FDP_,  x  = scorecol )) + 
    ggplot2::geom_line(ggplot2::aes(color = model_name, linetype = model_name)) +
    ggplot2::facet_wrap(~contrast) + 
    ggplot2::geom_abline(intercept = 0, slope = 1, color = 2) + 
    ggplot2::theme(legend.position = "bottom") +
    ggplot2::labs(tag = "C")

hlay = rbind(c(2,1,1),
             c(2,3,3))

gridExtra::grid.arrange(p1_pAUC, pB, pC, layout_matrix=hlay)
```

# Availability

https://github.com/fgcz/prolfqua

```{r echo=TRUE, eval=FALSE}
install.packages('remotes')
remotes::install_github('fgcz/prolfqua', build_vignettes = TRUE)
```

doi: 
https://doi.org/10.1101/2022.06.07.494524

```{r url, out.width="20%", fig.align="center"}
knitr::include_graphics("prolfqua-qr.png")
```

The authors thank the technology platform fund of the University of Zurich.

```{r sib, out.width="20%", fig.align="center"}
knitr::include_graphics("SIB_LogoQ_GBv.svg")
```

### 2022-07-20 inst/MyArticle/paper.Rmd
Add citation_package: natbib

@@
       keep_tex: true
+      citation_package: natbib
 date: "`r format(Sys.time())`"

### 2022-09-13 inst/MyArticle/prolfqua_supplement.Rmd
prolfqua supplement

---
title: "Supporting Information:\n prolfqua - A Comprehensive R-package for Proteomics Differential Expression Analysis"
author: 
  - Witold E. Wolski^[Functional Genomics Center Zurich (FGCZ), ETH Zurich / University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland] ^[Swiss Institute of Bioinformatics (SIB), Quartier Sorge - Batiment Amphipole, 1015 Lausanne, Switzerland]
  - Paolo Nanni$^*$
  - Jonas Grossmann$^*$$^\dag$
  - Maria d'Errico$^*$$^\dag$
  - Ralph Schlapbach$^*$
  - Christian Panse$^*$$^\dag$
output: 
  pdf_document:
    keep_tex: yes
    latex_engine: pdflatex
abstract: |
  This document contains the supplements for the prolfqua manuscript available through \url{https://doi.org/10.1101/2022.06.07.494524 }.
urlcolor: blue
toc: true
header-includes:
  - \usepackage{pdfpages}
---
\renewcommand{\thepage}{S--\arabic{page}}
\renewcommand{\tablename}{Supp. Table}
\renewcommand{\figurename}{Supp. Figure}

\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = TRUE)

```

# This
# And that
# And whatsoever

# R Session Information

```{r sessioninfo, echo=FALSE}
sessionInfo()
```

### 2022-09-22 inst/MyArticle/paper.Rmd
adding to bibliography,
fix caption

@@
 (ref:FDRfdp) Panel A - Number of estimated contrasts for each modeling method (higher is better).
-Panel B - Partial area under the ROC curve at $10\%$ FPR ($pAUC_10$) for all contrasts and three different statistics: the difference among groups, the scaled $p$-value (sign(diff) $\cdot$ p.value) and the $t$-statistics (higher is better).
+Panel B - Partial area under the ROC curve at $10\%$ FPR ($pAUC_{10}$) for all contrasts and three different statistics: the difference among groups (diff - Panel B left), the scaled $p$-value (sign(diff) $\cdot$ p.value) (scaled.p.value - Panel B center) and the $t$-statistics (statistic - Panel B right), where a higher pAUC is better.
 The red line indicates the average area under the curve of all methods.
@@
 
-## Benchmark
+## Benchmark Code
 

### 2022-10-06 inst/MyArticle/paper.Rmd
updating comparison table.

@@
 ```{r tableOverview}
+caption = "Rows - R packages, Columns - types of models supported: pd - parallel design, rm - repeated measurements, fd - factorial design, int - interactions among factors, mem - mixed effect models, eb - empirical Bayes, md - missing data modelling.
+Y - yes, * - repeated measurements are modeled using a fixed effect."
+
 xx <- list(
   MSstats = c("pd" = "Y", "rm" = "Y", "mem" = "Y"),
-  ROPECA = c("pd" = "Y", "rm" = "Y*", fd = "Y", int = "Y", eb = "Y"),
-  limma  = c("pd" = "Y", "rm" = "Y*", fd = "Y", int = "Y",eb = "Y"),
-  MSqRob2_rlm = c("pd" = "Y", "rm" = "Y*", fd = "Y", int = "Y",eb = "Y", md = "?"),
-  proDA = c("pd" = "Y", "rm" = "Y*", fd = "Y", int = "Y", eb = "Y", md = "Y"),
+  ROPECA = c("pd" = "Y", "rm" = "Y", fd = "Y", int = "Y", eb = "Y"),
+  limma  = c("pd" = "Y", "rm" = "Y", fd = "Y", int = "Y",eb = "Y"),
+  msqrob2 = c("pd" = "Y", "rm" = "Y", fd = "Y", int = "Y", mem = "Y", eb = "Y", md = "Y"),
+  proDA = c("pd" = "Y", "rm" = "Y", fd = "Y", int = "Y", eb = "Y", md = "Y"),
   prolfqua = c("pd" = "Y", "rm" = "Y", fd = "Y", int = "Y", mem = "Y",eb = "Y", md = "Y"))
@@
 rownames(bb) <- names(xx)
-knitr::kable(data.frame(bb), caption = "Rows - R packages, Columns - types of models supported: pd - parallel design, rm - repeated measurements, fd - factorial design, int - interactions among factors, mem - mixed effect models, eb - empirical Bayes, md - missing data modelling.
-Y - yes, * - repeated measurements are modeled using a fixed effect.
-?
-- the hurdle model was published but is not available in the msqrob2 package.")
+knitr::kable(data.frame(bb), format = "latex", caption = caption)
 ```
@@
 
-knitr::kable(table, caption = "Confusion matrix, TP - true positive, FP - false positive, FN - false negative, TN - true negatives, P - all positive cases (all $E.~coli$ proteins), N - all negative cases (all $H.~sapiens$ proteins), m - all proteins.")
+knitr::kable(table, caption = "Confusion matrix, TP - true positive, FP - false positive, FN - false negative, TN - true negatives, P - all positive cases (all $E.~coli$ proteins), N - all negative cases (all $H.~sapiens$ proteins), m - all proteins.", format = "latex" )
 ```
@@
 ## load peptide level data
-d <- prolfqua::prolfqua_data('data_ionstar')$filtered()
-
+d <- prolfqua::old2new(prolfqua::prolfqua_data('data_ionstar')$filtered())
+d <- prolfqua::LFQData$new(d$data, d$config)
 ## create R6 obejct
@@
 ## fit model
-lmmodel <- paste(lfqp$intensity_column()," ~ dilution.")
+lmmodel <- paste(lfqp$response()," ~ dilution.")
 modelFunction <- prolfqua::strategy_lm( lmmodel, model_name = "lm")
@@
 # merge contrasts, to obtain differences for all proteins
-contrasts <- prolfqua::addContrastResults(prefer = contr, add = conI)
+contrasts <- prolfqua::merge_contrasts_results(prefer = contr, add = conI)
 
@@
     f1 <- paste0("../../inst/Benchresults/",filN)
-    if(f1 == ""){
+    if (f1 == "") {
         f1 <- system.file("Benchresults",filN, package = "prolfquabenchmark")
@@
 msFragger <- readRDS(getpath("MSFragger_medpol_benchmark.RDS"))
+benchmark_msqrob <- readRDS(getpath("benchmark_msqrob.RDS"))
 
@@
 allBenchmarks$benchmark_proDA <- benchmark_prodA
-allBenchmarks <- allBenchmarks[c("benchmark_imputation","benchmark_ProtModerated",  "benchmark_mixedModerated", "benchmark_ropeca","benchmark_merged","benchmark_mssstats","benchmark_proDA"   )]
+allBenchmarks$benchmark_msqrob <- benchmark_msqrob
+
+allBenchmarks <- allBenchmarks[c("benchmark_imputation","benchmark_ProtModerated",  "benchmark_mixedModerated", "benchmark_ropeca","benchmark_merged","benchmark_mssstats","benchmark_proDA", "benchmark_msqrob")]
 
@@
 ttt <- sapply(allBenchmarks, function(x){x$complete(FALSE)})
-
 res <- purrr::map_df(allBenchmarks, function(x){x$pAUC()})
-
 res <- res |> dplyr::mutate(whatfix = dplyr::case_when(what == "scaled.beta.based.significance" ~ "scaled.p.value", TRUE ~ what))
 
-norm <- res |> dplyr::group_by(contrast,whatfix) |>
+norm <- res |> dplyr::group_by(contrast, whatfix) |>
     dplyr::summarize(meanpAUC_10 = mean(pAUC_10))
@@
 nrgg <- dds |> ggplot2::ggplot(ggplot2::aes(x = name, y = nrcontrasts )) + 
-  ggplot2::geom_bar(stat = "identity", fill="white", colour = "black") + 
+  ggplot2::geom_bar(stat = "identity", fill = "white", colour = "black") + 
   ggplot2::coord_cartesian(ylim = c(min(dds$nrcontrasts) - 200, max(dds$nrcontrasts) + 10)) +
@@
 knitr::kable(tibble::as_tibble(xM), caption = "Prolfqua functions that can be used to fit various models.",
-             format = "latex", booktabs=TRUE) |>
-    kableExtra::kable_styling(latex_options="scale_down") |> 
+             format = "latex", booktabs = TRUE) |>
+    kableExtra::kable_styling(latex_options = "scale_down") |> 
     kableExtra::add_footnote(label = "in development")
@@
 knitr::asis_output(stringr::str_replace(tmp, "HACKIT", "$^{a}$"))
+
 ```
@@
 # column with protein abundances
-atable$setWorkIntensity("total.intensity")
+atable$set_work_intensity("total.intensity")
 

### 2022-10-06 R/utilities.R

@@ -1,0 +1,8 @@
+tolong <- function(assay, rowData, colData ){
+    annot <- as_tibble(data.frame(colData))
+    hierarchy <- prolfqua::matrix_to_tibble( data.frame( rowData ) )
+    ll <- prolfqua::matrix_to_tibble(assay)
+    ll <- ll |> pivot_longer(cols = -all_of("row.names"), names_to = "sampleName", values_to = "intensity")
+    hl <- inner_join(hierarchy, ll)
+    ahl <- inner_join(annot, hl)
+}

### 2022-10-12 R/utilities.R
adding benchmark based on cptac dataset

@@ -6,3 +6,19 @@ tolong <- function(assay, rowData, colData ){
     hl <- inner_join(hierarchy, ll)
     ahl <- inner_join(annot, hl)
 }
+
+
+#' preprocess cptac dataset for benchmarking
+#' @export
+cptac_bench_preprocess <- function(data, idcol = "protein_Id") {
+    tmp <- data |>
+        ungroup() |>
+        mutate(species  = case_when(
+            grepl("YEAST", !!sym(idcol)) ~ "YEAST",
+            grepl("UPS", !!sym(idcol)) ~ "UPS",
+            TRUE ~ "OTHER"
+        ))
+    res <- tmp |> dplyr::filter(!.data$species == "OTHER")
+    res <- res |> mutate(TP = (.data$species == "UPS"))
+    return(list(data = res , table = table(tmp$species)))
+}

### 2022-10-12 inst/MyArticle/paper.Rmd
adding benchmark based on cptac dataset
update vignettes

@@
 ## load peptide level data
-d <- prolfqua::old2new(prolfqua::prolfqua_data('data_ionstar')$filtered())
-d <- prolfqua::LFQData$new(d$data, d$config)
+d <- prolfqua::prolfqua_data('data_ionstar')$filtered()
+d <- prolfqua::LFQData$new(d$data, prolfqua::old2new(d$config))
 ## create R6 obejct
@@
 ttt <- sapply(allBenchmarks, function(x){x$complete(FALSE)})
+names(allBenchmarks)
+
+for(i in seq_along(allBenchmarks)) {
+  allBenchmarks[[i]]$pAUC()
+}
+
 res <- purrr::map_df(allBenchmarks, function(x){x$pAUC()})
@@
 # column with protein abundances
-atable$set_work_intensity("total.intensity")
+atable$set_response("total.intensity")
 

### 2022-10-20 inst/MyArticle/paper.Rmd
rename tidy table to tidy data

@@
 library(BiocStyle)
-options(width=70)
+options(width = 70)
 ```
@@
 
-We store all the data needed for analysis in a tidy table, i.e., every column is a variable, every row is an observation, and every cell is a single value [@tidydata2014].
-Using an R6 [@R6cite] configuration object (Figure \@ref(fig:LFQData)), we specify what variable is in which column, making it easy to integrate new inputs in `r BiocStyle::Githubpkg("fgcz/prolfqua")` if provided as a tidy table.
+We store all the data needed for analysis in a data frame as tidy data, i.e., every column is a variable, every row is an observation, and every cell is a single value [@tidydata2014].
+Using an R6 [@R6cite] configuration object (Figure \@ref(fig:LFQData)), we specify what variable is in which column, making it easy to integrate new inputs in `r BiocStyle::Githubpkg("fgcz/prolfqua")` if provided as a tidy data.
 For example, to visualize tidy Spectronaut[@bruderer2015extending], DiaNN[@demichev2020dia], or Skyline[@maclean2010skyline] outputs, or data in `r BiocStyle::Biocpkg("MSstats")`[@Choi2014] format, only a few lines of code are needed to update the `r BiocStyle::Githubpkg("fgcz/prolfqua")` `AnalysisTableConfiguration` configuration.
@@
 We show an example code for creating an MSFragger[@yu2020fast] configuration in the Appendix.
-For popular software like MaxQuant[@Cox2008MaxQuant], or MSFragger, which stores the same variable, e.g., intensity, in multiple columns, one for each sample, we implemented methods that transform the data into tidy tables.
+For popular software like MaxQuant[@Cox2008MaxQuant], or MSFragger, which stores the same variable, e.g., intensity, in multiple columns, one for each sample, we implemented methods that transform the data into tidy data.
 Relying on the tidy data table enables us to interface with many data manipulation, visualization, and modeling methods, implemented in base R [@RCoreTeam2021] and the tidyverse [@tidyverse2019], easily.
@@
 
-### Determining linear parameter combinations for treatment comparison
+The package `r BiocStyle::Githubpkg("fgcz/prolfqua")` provides functions to determine the vector of $parameter$ weights $c$, from a linear model and a contrast specification string.
 
-The package `r BiocStyle::Githubpkg("fgcz/prolfqua")` provides a function to determine the vector of $parameter$ weights $c$, from a linear models and a contrast specification string.
+The linear model below explain the observed protein abundances using the explanatory variables `condition_` and `batch_` and the interaction among them `condition_:batch_`,
 
-The linear model below explain the observed protein abundances using the explanatory variables `factor_1` and `factor_2` and the interaction among them `factor_1:factor_2`,
-
-```{r}
-Intensity ~ factor_1 + factor_2 + factor_1:factor_2
+```{r, echo = TRUE}
+data_Yeast2Factor <- prolfqua::prolfqua_data("data_Yeast2Factor")
+mlm <- prolfqua::strategy_lm(transformedIntensity ~ condition_ + batch_ + condition_:batch_)
+mm <- prolfqua::build_model(data_Yeast2Factor$data, mlm, subject = "protein_Id")
 ```
 
-then the contrasts among `group_1` and `group_2` defined by `factor_1` can be specified using the string below.
+then the contrasts among the groups `Ethanol` and `Glucose` defined by the factor `condition`, and similarly for the difference of the groups defined by `batch_`, can be specified as:
 
-```{r eval=FALSE, echo=TRUE}
-c("contrast_name" = "factor_1group_1 - factor_1group_2")
+```{r contrast1,  echo=TRUE}
+d1 <- c("diff_condition" = "condition_Ethanol - condition_Glucose",
+  "diff_batch" = "batch_p2370 - batch_p2691")
 ```
 
-Furthermore, contrasts for subgroups can be specified using the code below,
+Furthermore, the difference among `Ethanol` and `Glucose` within the batches `p2370` and `p2691` can be specified as follows:
 
-```{r eval=FALSE, echo=TRUE}
-c("contrast_name" =
-    "`factor_1group_1:factor_2group_a` - `factor_1group_1:factor_2group_b`")
+```{r contrast2, eval=TRUE, echo=TRUE}
+d2 <- c("difference_within_p2370" =
+    "`condition_Glucose:batch_p2370` - `condition_Ethanol:batch_p2691`",
+  "difference_within_p2691" = 
+   "`condition_Glucose:batch_p2691` - `condition_Ethanol:batch_p2691`")
+```
+
+Finally, we can examine if the differences among `conditions_` observed in both batches significantly differ by specifying the contrast:
+
+```{r contrast3, eval=TRUE, echo=TRUE}
+d3 <- c("difference_of_difference" = "difference_within_p2370 - difference_within_p2691")
+```
+
+The code to compute the array of weights $c$ used to multiply the model coefficient $\beta$, to obtain the effect size  estimates, is shown below:
+
+```{r contrast4, echo=TRUE}
+Contrasts <- c(d1, d2, d3)
+linfct <- prolfqua::linfct_from_model(mm$modelDF$linear_model[[1]],
+                                      as_list = FALSE)
+mC <- prolfqua::linfct_matrix_contrasts(linfct, Contrasts)
+```
+
+while Table \@ref(tab:contrcoefficient) shows the weight $c$ of each model parameter in each contrast.
+
+```{r contrcoefficient}
+knitr::kable(mC, caption = "Weights of each model parameter.",format = "latex")
 
@@
 
-where `factor_x` is the name of the explanatory variable, and `group_x`  are group labels.
-
-The following code shows an example where we specify two contrast: the first to compare Cells of type CMP/MEP with cells of type HSC, and the second to compare therapy NO with therapy NU for the celltype CMP/MEP [@meier2021reduced].
-Finally, we compute the array of weights defining the contrast, used to multiply the model coefficient $\beta$.
-
-```{r echo=TRUE}
-Contrasts <- 
-    c("CMP/MEPvsHSC" = "`CelltypeCMP/MEP` - `CelltypeHSC`",
-   "NOvsHU" = "`class_therapyc.NO:CelltypeCMP/MEP` - `class_therapyp.HU:CelltypeCMP/MEP`")
-
-m <- prolfqua::prolfqua_data('data_basicModel_p1807')
-linfct <- prolfqua::linfct_from_model(m,as_list = FALSE)
-
-prolfqua::linfct_matrix_contrasts(linfct, Contrasts )
-```
+This example is taken from the `r BiocStyle::Githubpkg("fgcz/prolfqua")` package vignette file.
+Given the coefficients, $c$ contrasts for models obtained with `r BiocStyle::Biocpkg('proDA')`, or `r BiocStyle::Biocpkg('msqrob2')` can be determined, enabling us to integrate these packages transparently.
 
@@
 The code snippets in this section demonstrate how a differential expression analysis workflow can be implemented using the `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package.
-To speed up the computation of these examples, we use a random subset of the Ionstar dataset containing $163$ proteins and $1258$ peptides.
+To speed up the computation of these examples, we use subset of the Ionstar dataset generated by randomly selecting $400$ proteins.
 Peptide abundances are $\log_2$ transformed and robust z-score scaled using the method `robscale`.
@@
 ```{r prepro,  echo = TRUE , fig.cap="(ref:scaling)", out.width = '90%', fig.width = 8, fig.height = 8}
-## load peptide level data
-d <- prolfqua::prolfqua_data('data_ionstar')$filtered()
-d <- prolfqua::LFQData$new(d$data, prolfqua::old2new(d$config))
-## create R6 obejct
-lfqd <- prolfqua::LFQData$new(d$data, d$config)
+R.version.string; packageVersion("prolfqua")
 
+# read MQ peptide.txt and annotation table
+startdata <- prolfqua::tidyMQ_Peptides(system.file(
+  "samples/maxquant_txt/tiny2.zip", package = "prolfqua"))
+annot <- readxl::read_xlsx(system.file(
+  "samples/maxquant_txt/annotation_Ionstar2018_PXD003881.xlsx", package = "prolfqua"))
+startdata <- dplyr::inner_join(annot, startdata, by = "raw.file")
+# create proflqua configuration
+config <- prolfqua::create_config_MQ_peptide()
+config$table$factors[['dilution.']] = "sample"
+## create R6 object
+lfqd <- prolfqua::LFQData$new(startdata, config, setup = TRUE) 
+lfqd$hierarchy_counts()
+## remove observation with 0 intensity and filter for 2 peptides per protein
+lfqd$remove_small_intensities()
+lfqd$filter_proteins_by_peptide_count()
+lfqd$hierarchy_counts()
 ##  transform intensities
-t <- lfqd$get_Transformer()
-
-lfqd <-  t$log2()$robscale()$lfq
-lfqd$rename_response("peptide_abundance")
-
-## infer protein intensities from peptide intensity
+lfqd <- lfqd$get_Transformer()$log2()$robscale()$lfq
+lfqd$rename_response("log_peptide_abundance")
 agr <- lfqd$get_Aggregator()
 lfqp <- agr$medpolish()
-lfqp$rename_response("protein_abundance")
-
+lfqp$rename_response("log_protein_abundance")
 ## plot panels A-D
@@
   ggplot2::labs(tag = "A") + ggplot2::theme(legend.position = "none")
-panelB <- agr$plot()$plots[[1]] + ggplot2::labs(tag = "B")
+panelB <- agr$plot()$plots[[54]] + ggplot2::labs(tag = "B")
 panelC <- lfqp$get_Stats()$violin() + ggplot2::labs(tag = "C")
 pl <- lfqp$get_Plotter()
-panelD <- pl$boxplots()$boxplot[[1]] + ggplot2::labs(tag = "D")
+panelD <- pl$boxplots()$boxplot[[54]] + ggplot2::labs(tag = "D")
 ggpubr::ggarrange(panelA, panelB, panelC, panelD)
+
 ```
@@
 
-## compute contrasts from linear model, moderate t-statistics and p-values
-contr <- prolfqua::Contrasts$new(models, contrasts) |> 
-  prolfqua::ContrastsModerated$new()
-
-# compute contrasts using imputation imputation and moderate
-conI <- prolfqua::ContrastsSimpleImpute$new( lfqp, contrasts) |> 
-  prolfqua::ContrastsModerated$new()
-
+## compute contrasts from linear model
+contr <- prolfqua::Contrasts$new(models, contrasts) 
+# compute contrasts using imputation imputation
+conI <- prolfqua::ContrastsSimpleImpute$new( lfqp, contrasts)
 # merge contrasts, to obtain differences for all proteins
 contrasts <- prolfqua::merge_contrasts_results(prefer = contr, add = conI)
-
+# moderate t-statistics and p-values
+prolfqua::ContrastsModerated$undebug("get_contrasts")
+contrasts <-  contrasts$merged |> prolfqua::ContrastsModerated$new() 
 ## visualize results in panels
-pl <- contrasts$merged$get_Plotter() 
+pl <- contrasts$get_Plotter() 
 panelA <- pl$histogram()$p.value + ggplot2::labs(tag = "A")                                                        
@@
 ttt <- sapply(allBenchmarks, function(x){x$complete(FALSE)})
-names(allBenchmarks)
-
-for(i in seq_along(allBenchmarks)) {
+for (i in seq_along(allBenchmarks)) {
   allBenchmarks[[i]]$pAUC()
 }
-
 res <- purrr::map_df(allBenchmarks, function(x){x$pAUC()})
@@
 
-`r BiocStyle::Githubpkg("fgcz/prolfqua")` is an easy-to-use and feature-rich R package to analyze quantitative mass spectrometric data with simple or complex experimental designs.
+`r BiocStyle::Githubpkg("fgcz/prolfqua")` is an feature-rich, object oriented R package to analyze quantitative mass spectrometric data with simple or complex experimental designs.
 While other R packages for differential expression analysis of proteins typically only implement one modeling approach, `r BiocStyle::Githubpkg("fgcz/prolfqua")` supports various types of models (see Figure 2 and Table 3).
@@
 The modular design of `r BiocStyle::Githubpkg("fgcz/prolfqua")` allows for adding new features, e.g., generalized linear models to model the presence or absence information of a protein, or robust linear models, in the future.
-The developed framework is flexible enough to integrate other modeling methods, e.g., the probabilistic dropout model [@bioxrvproDA2020] or accurate variance estimation [@zhu2020deqms].
+Furthermore, the developed framework is flexible enough to integrate other modeling methods, e.g., the probabilistic dropout model [@bioxrvproDA2020] or accurate variance estimation [@zhu2020deqms].
 Hence, `r BiocStyle::Githubpkg("fgcz/prolfqua")` enables you to call various methods and quickly select the best differential expression analysis algorithm for your problem.
@@
 ```
+
+```{r roctest}
+ttmsqrob <- allBenchmarks$benchmark_msqrob$data()
+ddmsqrob <- pROC::roc(ttmsqrob$TP, ttmsqrob$scaled.p.value, partial.auc = c(1, 0.9))
+ttprolfqua <- allBenchmarks$benchmark_merged$data()
+ddprolfqua <- pROC::roc(ttprolfqua$TP, ttprolfqua$scaled.p.value, partial.auc = c(1, 0.9))
+ttproda <- allBenchmarks$benchmark_proDA$data()
+ddproda <- pROC::roc(ttproda$TP, ttproda$scaled.p.value, partial.auc = c(1, 0.9))
+
+msqrob_vs_prolfqua <- pROC::roc.test(ddmsqrob,ddprolfqua)
+msqrob_vs_proda <- pROC::roc.test(ddmsqrob,ddproda)
+prolfqua_vs_proda <- pROC::roc.test(ddprolfqua,ddproda)
+
+```

### 2022-10-20 inst/MyArticle/showThatFilesAreOK.R

@@ -1,0 +1,11 @@
+library(curl)
+
+download.file("https://gitlab.bfabric.org/wolski/prolfquadata/-/raw/master/inst/quantdata/MAXQuant_IonStar2018_PXD003881.zip?inline=false", "MAXQuant_IonStar2018_PXD003881.zip")
+tmp <- unzip("MAXQuant_IonStar2018_PXD003881.zip", list = RUE)
+pg <- read.table( unz("MAXQuant_IonStar2018_PXD003881.zip", "proteinGroups.txt"), sep = "\t", header=TRUE)
+iddiff <- diff(pg$id)  # same as : pg$id[2:nrow(pg)] - pg$id[1:(nrow(pg)-1)]
+iddiff |> table()
+
+
+download.file("https://gitlab.bfabric.org/wolski/prolfquadata/-/raw/master/inst/quantdata/MSFragger_IonStar2018_PXD003881.zip?inline=false", "MSFragger_IonStar2018_PXD003881.zip")
+tmp <- unzip("MSFragger_IonStar2018_PXD003881.zip", list = TRUE)

### 2022-10-20 inst/MyArticle/analysingDataStartingfromProteinGroups.Rmd
keep for the future.

---
title: "Modelling IonstarData with Interactions"
author: "Witold E. Wolski"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
vignette: |
  %\VignetteIndexEntry{Modelling with Interactions}  
  %\VignetteEncoding{UTF-8}   
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
bibliography:
  - prolfqua.bib
---

TODO(wew,jg): Can you please add a oneliner of the purpose of the vignette?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
a <- c(a = 3, b = 4.5, c = 6, d = 7.5, e = 9)
f1 <- list(l1 = 3, l2 = 4.5)
f2 <- list(l1 = 0, l2 = 3)

a <- f1$l1 + f2$l1
b <- f1$l2 + f2$l1
c <- f1$l1 + f2$l2
d <- f1$l2 + f2$l2
c(a, b, c, d)
```

First, we load the data and do the configuration.

```{r LoadDataAndConfigure}
datadir <- file.path(find.package("prolfqua") , "samples/maxquant_txt")
inputMQfile <- file.path(datadir, "tiny2.zip")
inputAnnotation <- file.path(datadir, "annotation_Ionstar2018_PXD003881.xlsx")
startdata <- prolfqua::tidyMQ_ProteinGroups(inputMQfile)
```

Read the sample annotation. The sample annotation must contain the `raw.file` name and the explanatory variables of your experiment, e.g. treatment, time point, genetic background, or other factors which you would like to check for confounding.

```{r readAnnotation}
annotation <- readxl::read_xlsx(inputAnnotation)
head(annotation)
annotation <- annotation |> dplyr::filter(sample != "e")
annotation <- annotation |> 
  dplyr::mutate(f1 = dplyr::case_when(sample %in% c("a","c") ~ "l1", TRUE ~ "l2"),
    f2 = dplyr::case_when(sample %in% c("a","b") ~ "l1", TRUE ~ "l2")) |>
  dplyr::arrange(sample)
```

Merge the annotation with quantitative data using `inner_join` joining by
`raw.file`.

```{r addAnnotationToData}
startdata <- dplyr::inner_join(annotation, startdata, by = "raw.file")
```

We remove all proteins identified only by a single peptide.

```{r filterForAtLeastTwoPeptides}
startdata <- dplyr::filter(startdata, nr.peptides > 1)
```

Then you need to _tell_ `prolfqua` which columns in the data frame contain what
information. You do it using the `AnalysisTableAnnotation` class.

```{r setupConfigs}
atable <- prolfqua::AnalysisTableAnnotation$new()
```

The `AnalysisTableAnnotation` has the following fields that need to be populated:
- fileName
- hierarchy
- factors
- workingIntensity
, and we will discuss it in more detail below.

The `fileName` is the column with the raw file names, however for labelled TMT experiments, it can be used to hold the name of the TMT channel.

```{r specifyRawFile}
atable$fileName = "raw.file"
```

The `hierarchy` field describes the structure of the MS data e.g, 
- protein 
- peptides
- modified peptides
- precursor
In case of the MaxQuant `proteinGroups.txt` file we have data on protein level.

```{r specifyProteinID}
atable$hierarchy[["protein_Id"]] <- c("proteinID")
```

In addition, we need to describe the `factors` of the analysis, i.e., the column
containing the explanatory variables.

```{r specifyFactors}
atable$factors[["f1."]] = "f1"
atable$factors[["f2."]] = "f2"

```

We also need to specify the column containing the protein abundances.

```{r specifyIntensity}
atable$set_response("mq.protein.intensity")
```

Finally, we create the `AnalysisConfiguration`, which needs the `AnalysisTableAnnotation` we just created and the `AnalysisParameters`.

```{r createAnalysisConfig}
config <- prolfqua::AnalysisConfiguration$new(atable)
adata <- prolfqua::setup_analysis(startdata, config)
```

Create the `LFQData` class instance and remove zeros from data. Of note, the MaxQuant software encodes missing values with zero.

```{r removeSmallIntensities}
lfqdata <- prolfqua::LFQData$new(adata, config)
lfqdata$remove_small_intensities()
lfqdata$factors()
```

Below, the code snippet graphs the heatmap of the protein abundances and a hierarchical clustering.

```{r plotHeatmap}
hm <- lfqdata$get_Plotter()$heatmap()
hm
```

```{r}
tr <- lfqdata$get_Transformer()
lfqTrans <- tr$log2()$lfq
lfqTrans$get_Plotter()$intensity_distribution_density()
lfqTrans$response()
lfqTrans$rename_response("abundance")
```

# Model Fitting

defines the contrasts

```{r specifyModel}
formula_Batches <- prolfqua::strategy_lm("abundance ~ f1. * f2. ")

# specify model definition
modelName  <- "Model"

# TODO(wew): why debug?
DEBUG <- TRUE

Contrasts <- c("f1.l1vsf1.l2" = "f1.l1 - f1.l2",
               "f2.l1vsf2.l2" = "f2.l1 - f2.l2",
               "f1l1vsf1l2_gv_f2.l1" = "`f1.l1:f2.l1` - `f1.l2:f2.l1`",
               "f1l1vsf1l2_gv_f2.l2" = "`f1.l1:f2.l2` - `f1.l2:f2.l2`",
               "Interaction" = "`f1l1vsf1l2_gv_f2.l1` - `f1l1vsf1l2_gv_f2.l2`"
               )
```

The following line fits the model.

```{r buildModel}
mod <- prolfqua::build_model(lfqTrans, formula_Batches)
```

```{r anovaPvaluePlots, fig.cap="p-value distributions for ANOVA analysis."}
mod$anova_histogram(what = "FDR.Pr..F.")
```

## ANOVA

Examine proteins with a significant interaction between the two factors treatment and batch.

```{r anovaAnalysis}
ANOVA <- mod$get_anova()
ANOVA |>
  dplyr::filter(factor == "f1.:f2.") |>
  dplyr::arrange(FDR.Pr..F.) |>
  head(5)

ANOVA$factor |> unique()
protIntSig <- ANOVA |> dplyr::filter(factor == "f1.") |>
  dplyr::filter(FDR.Pr..F. < 0.25)

protInt <-  lfqTrans$get_copy()
protInt$data <- protInt$data[protInt$data$protein_Id %in%
                               protIntSig$protein_Id, ]
```

```{r fig.with=15, fig.height=15, fig.cap="Proteins with FDR < 0.5 for condition batch interaction in ANOVA."}
ggpubr::ggarrange(plotlist = protInt$get_Plotter()$boxplots()$boxplot)
```

# Compute contrasts

```{r computeModeratedContrasts}
contr <- prolfqua::ContrastsModerated$new(prolfqua::Contrasts$new(mod, Contrasts))
#contr$get_contrasts_sides()
contrdf <- contr$get_contrasts()
```

The code snippets graph the volcano and ma plot.
```{r graphVolcano}
plotter <- contr$get_Plotter()
plotter$volcano()
plotter$ma_plot()
```

## Annalyse contrasts with missing data imputation

```{r}
lfqTrans$config$table$factorDepth <- 2
# ContrastsSimpleImpute$debug("get_contrasts")
contrSimple <- prolfqua::ContrastsSimpleImpute$new(lfqdata = lfqTrans,
                                                   Contrasts)
contrdfSimple <- contrSimple$get_contrasts()
# na.omit(contrdfSimple)
pl <- contrSimple$get_Plotter()
pl$histogram_diff()
pl$volcano()
```

## Merge nonimputed and imputed data.

```{r mergeNonImpyted}
dim(contr$get_contrasts())
dim(contrSimple$get_contrasts())

mergedContrasts <- prolfqua::merge_contrasts_results(prefer = contr, add = contrSimple)$merged
cM <- mergedContrasts$get_Plotter()
```

```{r plotNergeNonImpyted}
plot <- cM$volcano()
plot$FDR
```

The `prolfqua` package is described in [@Wolski2022.06.07.494524].

# Session Info

```{r}
sessionInfo()
```

# References

### 2022-10-20 inst/MyArticle/paper.Rmd
Edits in method section.

@@
 
-The linear model below explain the observed protein abundances using the explanatory variables `condition_` and `batch_` and the interaction among them `condition_:batch_`,
+The linear model below explain the observed protein abundances using the explanatory variables `condition_` and `batch_` and the interaction term  `condition_:batch_`:
 
@@
 data_Yeast2Factor <- prolfqua::prolfqua_data("data_Yeast2Factor")
-mlm <- prolfqua::strategy_lm(transformedIntensity ~ condition_ + batch_ + condition_:batch_)
+mlm <- prolfqua::strategy_lm(
+  transformedIntensity ~ condition_ + batch_ + condition_:batch_
+  )
 mm <- prolfqua::build_model(data_Yeast2Factor$data, mlm, subject = "protein_Id")
@@
 
-then the contrasts among the groups `Ethanol` and `Glucose` defined by the factor `condition`, and similarly for the difference of the groups defined by `batch_`, can be specified as:
+The treatment difference among the groups `Ethanol` and `Glucose` defined by the factor `condition`, and similarly among the groups defined by `batch_`, can be specified.
 
@@
 
-Furthermore, the difference among `Ethanol` and `Glucose` within the batches `p2370` and `p2691` can be specified as follows:
+Furthermore, the treatment differences among `Ethanol` and `Glucose` within the batches `p2370` and `p2691` can be specified.
 
@@
 
-Finally, we can examine if the differences among `conditions_` observed in both batches significantly differ by specifying the contrast:
+Finally, we can examine if the treatment differences in both batches are the same or are different.
 
@@
 
-The code to compute the array of weights $c$ used to multiply the model coefficient $\beta$, to obtain the effect size  estimates, is shown below:
+The code to compute the array of weights $c$ used to multiply the model coefficient $\beta$, to obtain the effect size estimates, and the variance of the estimate, is shown below:
 
@@
 
-while Table \@ref(tab:contrcoefficient) shows the weight $c$ of each model parameter in each contrast.
+Table \@ref(tab:contrcoefficient) shows weights $c$ for each of the five contrasts we specified.
 
@@
 This example is taken from the `r BiocStyle::Githubpkg("fgcz/prolfqua")` package vignette file.
-Given the coefficients, $c$ contrasts for models obtained with `r BiocStyle::Biocpkg('proDA')`, or `r BiocStyle::Biocpkg('msqrob2')` can be determined, enabling us to integrate these packages transparently.
+Given the weights $c$ contrasts from model parameter estimates obtained with `r BiocStyle::Biocpkg('proDA')`, or `r BiocStyle::Biocpkg('msqrob2')` can be determined, by this enabling us to call these packages using the same application programming interface (API).
 
@@
 pl <- contrasts$get_Plotter() 
-panelA <- pl$histogram()$p.value + ggplot2::labs(tag = "A")                                                        
-panelB <- pl$volcano()$FDR +
+histpval <- pl$histogram()$p.value + ggplot2::labs(tag = "A")               maplot <- pl$ma_plot(rank = TRUE) + ggplot2::labs(tag = "C")
+volcano <- pl$volcano()$FDR +
    ggplot2::theme(legend.position = "bottom") +
@@
 
-gridExtra::grid.arrange(panelA, panelB, ncol = 2)
+gridExtra::grid.arrange(histpval, maplot, volcano, layout_matrix = rbind(c(1,3),c(2,3)))
 ```

### 2022-10-21 inst/MyArticle/paper.Rmd
benchmarks

@@
 For quantification, intact peptide ions [@yu2020fast; @Cox2008MaxQuant] or products of peptide ion fragmentation [@rost2014openswath; @demichev2020dia] are counted and aggregated to obtain peptide abundances.
-Finally, the identified and quantified peptides are assigned to proteins based on protein sequence information.
+Finally, the identified and quantified peptides are assigned to proteins based on protein sequence information [@nesvizhskii2005interpretation].
 
@@
 Several data analysis packages exist to model mass spectrometry protein quantification experiments, e.g., `r BiocStyle::Biocpkg("limma")` [@Ritchie2015], `r BiocStyle::Biocpkg("MSstats")` [@Choi2014], `r BiocStyle::Biocpkg("PECA")` [@Suomi2017bEnhanced],  `r BiocStyle::Biocpkg("msqrob2")` [@Goeminne2016] or `r BiocStyle::Biocpkg("proDA")` [@bioxrvproDA2020], to mention some, all implemented in R.
-Each of them has some unique features, for example, `r BiocStyle::Biocpkg("MSstats")` infers the model and generates the model formula from the sample annotation, allowing users with limited statistical knowledge to perform differential expression analysis (DEA).
+Each of them has some unique features, for example, `r BiocStyle::Biocpkg("MSstats")` determines the statistical model from the structure of the sample annotation, allowing users with limited statistical knowledge to perform differential expression analysis (DEA).
 At the same time, `r BiocStyle::Biocpkg("limma")` allows to specify a design matrix using a linear model formula and implements the empirical Bayes variance shrinkage method.
 The package `r BiocStyle::Biocpkg("PECA")`, performs a roll-up of peptide level differences and peptide level $p$-value estimates, obtained from `r BiocStyle::Biocpkg("limma")` or `r BiocStyle::Biocpkg("PECA")`, to protein level estimates.
-Furthermore, `r BiocStyle::Biocpkg("msqrob2")` implements two models: robust linear models fitted to protein intensities and robust ridge regression fitted to peptide intensities, and combines them with empirical Bayes variance shrinkage.
+Furthermore, `r BiocStyle::Biocpkg("msqrob2")` combines two models: robust linear models fitted to protein abundances and a generalized linear model fitted to presence absence information of peptides, into Hurdle model.
 The `r BiocStyle::Biocpkg("proDA")` package implements a linear probabilistic dropout model to account for missing data without imputation.
-`r BiocStyle::Biocpkg("MSstats")` handles missing data by feature filtering and using imputation.
-Other means of modelling missing observations are the Hurdle model discussed by @goeminne2020msqrob, while the R package `r BiocStyle::Biocpkg("proDA")` models missingness using probabilistic dropout models [@bioxrvproDA2020].
 
-When analyzing parallel-group designs using a single explanatory variable, and contrasting groups, we can use all R packages; but, we can use only some of them if we want to model factorial designs or repeated measurements.
+Of note are the various approaches to handling missing observations, which are inherent in mass spectrometric bottom-up experiments.
+For instance, `r BiocStyle::Biocpkg("MSstats")` handles missing data by feature filtering and imputation.
+Other means of modeling missing observations are the Hurdle model discussed by @goeminne2020msqrob, while the R package `r BiocStyle::Biocpkg("proDA")` models missingness using probabilistic dropout models [@bioxrvproDA2020].
+
+We can use all the R packages discussed when analyzing parallel-group designs using a single explanatory variable and contrasting groups.
+However, we can use only some of them if we want to model factorial designs or repeated measurements.
 Table \@ref(tab:tableOverview) gives an overview of the models and features supported by these packages.
-We see that packages such as `r BiocStyle::Biocpkg("limma")` and `r BiocStyle::Biocpkg("proDA")` allow us to fit a comprehensive variety of models and test various hypotheses; however, in-depth knowledge of design matrix specification using the R formula interface is required [@law2020guide].
+We see that, for instance, `r BiocStyle::Biocpkg("limma")` and `r BiocStyle::Biocpkg("proDA")` allow us to fit a comprehensive variety of models and test various hypotheses; however, good knowledge of the design matrix specification using the R formula interface is required [@law2020guide].
 
@@
   MSstats = c("pd" = "Y", "rm" = "Y", "mem" = "Y"),
-  ROPECA = c("pd" = "Y", "rm" = "Y", fd = "Y", int = "Y", eb = "Y"),
+  PECA = c("pd" = "Y", "rm" = "Y", eb = "Y"),
   limma  = c("pd" = "Y", "rm" = "Y", fd = "Y", int = "Y",eb = "Y"),
@@
 When developing the R package `r BiocStyle::Githubpkg("fgcz/prolfqua")` we were inspired by the R package `r BiocStyle::CRANpkg("caret")` [@caret2008], which enables us to call various machine learning (ML) methods, and makes selecting the best ML algorithm for your problem easy.
-We aimed for a similar R package for the differential expression analysis of proteomics data.
-However, after examining the R packages for modeling proteomics quantification data, we observed that model specification, input, and output formats differ widely.
-At the same time, they have in common that: they fit linear models either to peptide or protein intensities, determine differences among groups, and afterward apply empirical Bayes variance shrinkage.
-Therefore, the revised goal was to provide a modular object-oriented (OO) design, with R linear models as a core, where we can add features such as $p$-value aggregation, e.g., `r BiocStyle::Biocpkg("PECA")`, variance shrinkage, or modeling of missing observations.
+We aimed for a similar R package for the differential expression analysis of quantitative proteomics data.
+However, the existing packages differ widely regarding supported designs, model specifications, and output formats.
+At the same time, they have in common: they fit linear models to either peptide or protein intensities, determine differences among groups, and afterward apply empirical Bayes variance shrinkage.
+Therefore, the revised goal was to provide a modular object-oriented (OO) design, with R linear models as a core, and add features such as $p$-value aggregation, variance shrinkage, or modeling of missing observations.
 
@@
 For example, we implemented strategies to estimate protein intensities from peptide intensities: top N [@Grossmann2010], Tukey's median polish  [@tukey1977exploratory], robust linear models [@goeminne2020msqrob].
-Furthermore, peptide or protein abundances can be scaled and transformed, using robust scaling, _quantile_ normalization or `r BiocStyle::Biocpkg("vsn")` to remove systematic differences among samples and heteroscedasticity.
-In this respect, `r BiocStyle::Githubpkg("fgcz/prolfqua")` can be compared with other R packages such as `r BiocStyle::Biocpkg("DEP")` [@DEP2018] or `r BiocStyle::Biocpkg("POMA")` [@POMA2021] which support the entire differential expression analysis pipeline.
+Furthermore, peptide or protein abundances can be scaled and transformed using robust scaling, _quantile_ normalization or `r BiocStyle::Biocpkg("vsn")` to remove systematic differences among samples and heteroscedasticity.
+In this respect, `r BiocStyle::Githubpkg("fgcz/prolfqua")` can be compared with R packages such as `r BiocStyle::Biocpkg("DEP")` [@DEP2018] or `r BiocStyle::Biocpkg("POMA")` [@POMA2021] which support the entire differential expression analysis pipeline.
 
-We also implemented functionality and use the Ionstar [@shen2018ionstar] dataset to benchmark the modeling methods implemented within `r BiocStyle::Githubpkg("fgcz/prolfqua")` and to compare our results with those of `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`.
+We also implemented functionality and use the Ionstar [@shen2018ionstar] and CPTAC (available from `r BiocStyle::Biocpkg("msdata")`) dataset to benchmark the modeling methods implemented within `r BiocStyle::Githubpkg("fgcz/prolfqua")` and to compare our results with those of `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`.
 Since group sizes are relatively small, typically with four or five subjects per group, the high power of the tests is a relevant criterion to assess the modeling method.
@@
 (ref:LFQData) Class Diagram of classes representing the proteomics data.
-The `LFQData` class encapsulates the quantitative proteomics data stored in a tidy table.
-An instance of the AnalysisTableConfiguration class specifies a mapping of columns in the tidy table.
+The `LFQData` class encapsulates the quantitative proteomics data stored in a table of tidy data.
+An instance of the AnalysisTableConfiguration class specifies a mapping of table columns to sample names, peptide or protein identifiers, explanatory variables, and response variables.
 The `LFQDataPlotter` class and other classes decorate the `LFQData` class with additional functionality.
@@
 R's formula interface for linear models is flexible, widely used, and well documented [@faraway2016extending].
-We use the formula interface to specify the models, making it easy to reproduce an analysis performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` in other statistical methods programming languages.
-In addition, we implement features specific to high throughput experiments, such as the experimental Bayes variance and $p$-value moderation, which utilizes the parallel structure of the protein measurements and the analysis [@Ritchie2015].
+We use the formula interface to specify the models, making it easy to reproduce an analysis performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` in other statistical programming languages.
+In addition, we implement features specific to high throughput experiments, such as the empirical Bayes variance and $p$-value moderation, which utilizes the parallel structure of the protein measurements and the analysis [@Ritchie2015].
 We also compute probabilities of differential protein regulation based on peptide level models [@Suomi2017bEnhanced].
@@
 
-(ref:ContrastUML)  The UML (unified modeling language) diagram of modeling and contrast related classes.
-Different strategies, e.g., _lm_, _lmer_, and _glm_ (Table \@ref(tab:prolfquaModels)), can fit various types of models.
+(ref:ContrastUML)  Unified modeling language (UML) diagram of modeling and contrast related classes.
+Different strategies, e.g., _lm_, _lmer_, and _glm_ (Table \@ref(tab:prolfquaModels)), reference methods to fit models, and compute contrasts.
 The model builder method fits the statistical model given the data and a strategy.
-The obtained model can then be used for the analysis of variance or to estimate contrasts.
-All classes estimating contrasts implement the _Contrasts_ interface.
-Results of external tools, e.g., SAINTexpress, can be adapted by implementing the Contrasts Interface.
+The models are used to analyze variance (ANOVA) or to estimate contrasts.
+All classes estimating contrasts implement the _ContrastsInterface_.
+Results of external tools, e.g., SAINTexpress, or proDA are adapted to implement the Contrasts Interface.
 
@@
 Protein abundances are inferred from the peptide intensities using Tukey's median polish.
-Finally, we fitted the fixed effect linear models, the dropout model `r BiocStyle::Biocpkg("proDA")` to protein abundances, the mixed effect linear model, and the `r BiocStyle::Biocpkg("PECA")` model to peptide intensities.
+Finally, we fitted the fixed effect linear models, the dropout model `r BiocStyle::Biocpkg("proDA")` to protein abundances, the mixed effect linear model, the `r BiocStyle::Biocpkg("PECA")` model, and hurdle model `r BiocStyle::Biocpkg("msqrob2")` to peptide intensities.
 
@@
 
-The linear model below explain the observed protein abundances using the explanatory variables `condition_` and `batch_` and the interaction term  `condition_:batch_`:
+The linear model explains the observed protein abundances using the explanatory variables `condition_` and `batch_` and the interaction term  `condition_:batch_`.
+The treatment difference among the groups `Ethanol` and `Glucose` defined by the factor `condition`, and similarly among the groups defined by `batch_`, can be specified.
+Furthermore, the treatment differences among `Ethanol` and `Glucose` within the batches `p2370` and `p2691` can be specified.
+Finally, we can examine if the treatment differences in both batches are the same or are different.
+The code to compute the array of weights $c$ used to multiply the model coefficient $\beta$, to obtain the effect size estimates, and the variance of the estimate, is shown below:
 
-```{r, echo = TRUE}
-data_Yeast2Factor <- prolfqua::prolfqua_data("data_Yeast2Factor")
+```{r contrast2, eval=TRUE, echo=TRUE}
 mlm <- prolfqua::strategy_lm(
@@
   )
-mm <- prolfqua::build_model(data_Yeast2Factor$data, mlm, subject = "protein_Id")
-```
+Yeast2Factor <- prolfqua::prolfqua_data("data_Yeast2Factor")
+mm <- prolfqua::build_model(Yeast2Factor$data, mlm, subject_Id = "protein_Id")
 
-The treatment difference among the groups `Ethanol` and `Glucose` defined by the factor `condition`, and similarly among the groups defined by `batch_`, can be specified.
+Contrasts <- c(
+  "diff_condition" = "condition_Ethanol - condition_Glucose",
+  "diff_batch" = "batch_p2370 - batch_p2691",
+  "difference_within_p2370" = "`condition_Glucose:batch_p2370` - `condition_Ethanol:batch_p2691`",
+  "difference_within_p2691" = "`condition_Glucose:batch_p2691` - `condition_Ethanol:batch_p2691`",
+  "difference_of_difference" = "difference_within_p2370 - difference_within_p2691")
 
-```{r contrast1,  echo=TRUE}
-d1 <- c("diff_condition" = "condition_Ethanol - condition_Glucose",
-  "diff_batch" = "batch_p2370 - batch_p2691")
-```
-
-Furthermore, the treatment differences among `Ethanol` and `Glucose` within the batches `p2370` and `p2691` can be specified.
-
-```{r contrast2, eval=TRUE, echo=TRUE}
-d2 <- c("difference_within_p2370" =
-    "`condition_Glucose:batch_p2370` - `condition_Ethanol:batch_p2691`",
-  "difference_within_p2691" = 
-   "`condition_Glucose:batch_p2691` - `condition_Ethanol:batch_p2691`")
-```
-
-Finally, we can examine if the treatment differences in both batches are the same or are different.
-
-```{r contrast3, eval=TRUE, echo=TRUE}
-d3 <- c("difference_of_difference" = "difference_within_p2370 - difference_within_p2691")
-```
-
-The code to compute the array of weights $c$ used to multiply the model coefficient $\beta$, to obtain the effect size estimates, and the variance of the estimate, is shown below:
-
-```{r contrast4, echo=TRUE}
-Contrasts <- c(d1, d2, d3)
 linfct <- prolfqua::linfct_from_model(mm$modelDF$linear_model[[1]],
@@
 mC <- prolfqua::linfct_matrix_contrasts(linfct, Contrasts)
+
 ```
@@
 ```{r contrcoefficient}
-knitr::kable(mC, caption = "Weights of each model parameter.",format = "latex")
+knitr::kable(mC, caption = "Weights $c$ of each model parameter (columns) to compute contrast (rows).",format = "latex")
 
@@
 Missing observations lead to different group sizes, which results in unbalanced designs.
-Linear and mixed effect models can handle unbalanced designs, as long as one observation per group is available, they will produce unbiased estimates, and no imputation is needed.
+Linear and mixed effect models can handle unbalanced designs.
+As long as at least one observation in a group is available, and sufficient observations to estimate the variance are available, they will produce unbiased estimates.
+Therefore, no imputation is needed.
 
-If there is no observation in a group, we assume that the protein is unobserved because the protein abundance is below the limit of detection (LOD) of the MS instrument.
-In this case, we will impute the mean using the protein abundance at the detection limit $A_{LOD}$.
-We estimate the abundance at the detection limit using the abundances of the proteins observed only in one sample of a group of samples.
-Then, we compute the median of the distribution and use it as the group mean if a protein is absent in a treatment group.
+However, if there is no observation in a group, the model fit fails.
+For example, suppose a protein is unobserved in all the samples of a group.
+In that case, a plausible explanation is that the protein abundance is below the limit of detection (LOD) of the MS instrument.
+In such a case, we will substitute the group mean, using the expected protein abundance $A$ at the LOD $A_{LOD}$.
+To estimate $A_{LOD}$ we are using the abundances from those groups where the protein was unobserved except for one sample. Typically there are many such cases, and hence we take the median.
 
-When computing differences $\Delta$ among two groups $a$ and $b$, we will use the group mean $\bar{a}$ or $\bar{b}$ estimated from the data, or if no observations are present in a group, we use $A_{LOD}$, e.g., If there are no observations in group $b$ then :
+When computing differences $\Delta$ among two groups $a$ and $b$, we will use either the group mean $\bar{a}$ or $\bar{b}$ estimated from the data.
+However, if for instance no observations are present in group $b$, we will use $\bar{b} = A_{LOD}$.
+Furthermore, if $\bar{a} < A_{LOD}$ we also set $\bar{a} = A_{LOD}$.
 
@@
 
-To estimate the variance, we assume that the variance of the protein is constant in all the groups and use the pooled variance based on all data:
+We use the pooled variance in all groups to estimate the protein variance, assuming they are the same. The pooled variance $s_p^2$ is given by:
 
@@
 Where, $n_g$ is the number of groups and $n$ is the number of observations.
-If variance can not be estimated for a protein, we will use the median pooled variance of all the proteins.
+If variance can not be estimated for a protein, because there are to few observations in other groups, we use the median pooled variance of all other proteins in the dataset.
 
-This methods are implemented in the class `ContrastSimpleImpute` (see Figure \@ref(fig:ContrastUML)).
+This method is implemented in the class `ContrastSimpleImpute` (see Figure \@ref(fig:ContrastUML)).
 
@@
 
-# read MQ peptide.txt and annotation table
+## read MQ peptide.txt and annotation table
 startdata <- prolfqua::tidyMQ_Peptides(system.file(
@@
 startdata <- dplyr::inner_join(annot, startdata, by = "raw.file")
-# create proflqua configuration
+## create MaxQuant configuration
 config <- prolfqua::create_config_MQ_peptide()
+## specify explanatory variable
 config$table$factors[['dilution.']] = "sample"
 ## create R6 object
-lfqd <- prolfqua::LFQData$new(startdata, config, setup = TRUE) 
-lfqd$hierarchy_counts()
+lfqpep <- prolfqua::LFQData$new(startdata, config, setup = TRUE) 
 ## remove observation with 0 intensity and filter for 2 peptides per protein
-lfqd$remove_small_intensities()
-lfqd$filter_proteins_by_peptide_count()
-lfqd$hierarchy_counts()
+lfqpep$remove_small_intensities()$filter_proteins_by_peptide_count()
 ##  transform intensities
-lfqd <- lfqd$get_Transformer()$log2()$robscale()$lfq
-lfqd$rename_response("log_peptide_abundance")
-agr <- lfqd$get_Aggregator()
-lfqp <- agr$medpolish()
-lfqp$rename_response("log_protein_abundance")
+lfqpep <- lfqpep$get_Transformer()$log2()$robscale()$lfq
+lfqpep$rename_response("log_peptide_abundance")
+agr <- lfqpep$get_Aggregator()
+lfqpro <- agr$medpolish()
+lfqpro$rename_response("log_protein_abundance")
 ## plot panels A-D
-pl <- lfqd$get_Plotter()
+pl <- lfqpep$get_Plotter()
 panelA <- pl$intensity_distribution_density() +
@@
 
-(ref:exampleContrasts) Panel A - Histogram showing the distribution of p-values for $163$ proteins.
-Panel B -  Volcano plot showing $-\log_{10}$ transformed $FDR$ as function of the difference between groups for $163$ proteins.
+(ref:exampleContrasts) Panel A - Histogram showing the distribution of p-values for $332$ proteins.
+Panel B -  Volcano plot showing $-\log_{10}$ transformed $FDR$ as function of the difference between groups for $332$ proteins.
+With black dots we show effect size and FDR estimates obtained from the linear model, while in green we show those obtained using imputation.
 
@@
 ## compute contrasts from linear model
-contr <- prolfqua::Contrasts$new(models, contrasts) 
+contr <- prolfqua::Contrasts$new(models, contrasts, modelName = "lm")
 # compute contrasts using imputation imputation
-conI <- prolfqua::ContrastsSimpleImpute$new( lfqp, contrasts)
+conI <- prolfqua::ContrastsSimpleImpute$new( lfqp, contrasts, modelName = "imp")
 # merge contrasts, to obtain differences for all proteins
@@
 pl <- contrasts$get_Plotter() 
-histpval <- pl$histogram()$p.value + ggplot2::labs(tag = "A")               maplot <- pl$ma_plot(rank = TRUE) + ggplot2::labs(tag = "C")
+histpval <- pl$histogram()$p.value + ggplot2::labs(tag = "A")               
+maplot <- pl$ma_plot(rank = TRUE,legend = FALSE) + ggplot2::labs(tag = "C")
 volcano <- pl$volcano()$FDR +
@@
 
-Using the tidy table to model the data ensures interoperability with other proteomics-related packages that manage their data with tidy-tables, e.g., `r BiocStyle::CRANpkg("protti")` [@quast2022protti].
-The use of R6 classes, which encapsulate the configuration and the data, allows for writing very concise code (no function arguments needed).
-Auto-completion support in the Rstudio editor makes it easy for novices to explore `r BiocStyle::Githubpkg("fgcz/prolfqua")`s functionality (Figure \@ref(fig:tabCompletion)).
+Using the table of tidy data ensures interoperability with other proteomics-related packages that manage their data with tidy-tables, e.g., `r BiocStyle::CRANpkg("protti")` [@quast2022protti].
 To simplify integration of `r BiocStyle::Githubpkg("fgcz/prolfqua")` with Bioconductor based workflows there is a method to convert the `LFQData` class into a `r BiocStyle::Biocpkg("SummarizedExperiment")`.
+The use of R6 classes, which encapsulate the configuration and the data, allows for writing very concise code, where functions can have only very few arguments.
+Auto-completion support for R6 classes in the Rstudio editor makes it easy for novices to explore `r BiocStyle::Githubpkg("fgcz/prolfqua")`s functionality (see Supplement Figure S[X]).
 
-To ease the usage barriers of the R package to users not being familiar with statistics and R programming, we integrated some workflows into our data management platform [b-fabric](https://fgcz-bfabric.uzh.ch/) [@bfabric].
+To ease the usage barriers of the R package to users not being familiar with statistics and R programming, we integrated some workflows into our data management platform [B-Fabric](https://fgcz-bfabric.uzh.ch/) [@bfabric,].
 This integration enables our users to select the input and basic settings in a graphical user interface.
-Then, as an output, the user receives a report including input files, the R markdown file, and R scripts necessary to replicate the analysis using their in-house R installation.
-The b-fabric system runs a computing infrastructure controlled by a local resource management system that supports cloud-bursting [@VMMAD].
-In this way, [b-fabric](https://fgcz-bfabric.uzh.ch/) helps scientists to meet requirements from funding agencies, journals, and academic institutions to publish data according to the FAIR (Findable, Accessible, Interoperable and Reusable) [@FAIR] data principles.
+The user receives a report including input files, the R markdown file, and R scripts necessary to replicate the analysis using their in-house R installation.
+The B-Fabric system runs a computing infrastructure controlled by a local resource management system that supports cloud-bursting [@VMMAD].
+In this way, [B-Fabric](https://fgcz-bfabric.uzh.ch/) helps scientists to meet requirements from funding agencies, journals, and academic institutions to publish data according to the FAIR (Findable, Accessible, Interoperable and Reusable) [@FAIR] data principles.
 
@@
 
-gridExtra::grid.arrange(p1_pAUC, pB, pC, layout_matrix=hlay)
+gridExtra::grid.arrange(p1_pAUC, pB, pC, layout_matrix = hlay)
 
@@
              "Input File - name of MaxQuant file used as input."),
-             format = "latex", booktabs=TRUE) |>
-    kableExtra::kable_styling(latex_options="scale_down") |> 
+             format = "latex", booktabs = TRUE) |>
+    kableExtra::kable_styling(latex_options = "scale_down") |> 
     kableExtra::add_footnote(label = "prolfqua_merged - Merges results of the prot_med_lm_moderated and prot_imputation modeling pipeline, preferring those of prot_med_lm_moderated if available.")
@@
 knitr::asis_output(stringr::str_replace(tmp, "HACKIT", "$^{a}$"))
+
+```
+
+```{r roctest, include = FALSE}
+ttmsqrob <- allBenchmarks$benchmark_msqrob$data()
+ddmsqrob <- pROC::roc(ttmsqrob$TP, ttmsqrob$scaled.p.value, partial.auc = c(1, 0.9))
+ttprolfqua <- allBenchmarks$benchmark_merged$data()
+ddprolfqua <- pROC::roc(ttprolfqua$TP, ttprolfqua$scaled.p.value, partial.auc = c(1, 0.9))
+ttproda <- allBenchmarks$benchmark_proDA$data()
+ddproda <- pROC::roc(ttproda$TP, ttproda$scaled.p.value, partial.auc = c(1, 0.9))
+
+msqrob_vs_prolfqua <- pROC::roc.test(ddmsqrob,ddprolfqua)
+msqrob_vs_proda <- pROC::roc.test(ddmsqrob,ddproda)
+prolfqua_vs_proda <- pROC::roc.test(ddprolfqua,ddproda)
 
@@
 
-The authors thank the technology platform fund (TPF) of the University of Zurich and all FGCZ proteomics colleagues for fruitful discussions.
+The authors thank the technology platform fund (TPF) of the University of Zurich and all colleagues of the FGCZ proteomics team for fruitful discussions.
 
@@
 
-## Benchmark Code
-
-All the code needed to replicate the benchmark results and this document is available at:
-
-https://github.com/wolski/prolfquabenchmark.
-
-- [Benchmarking of methods implemented in prolfqua](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkingIonstarData.html)
-- [Benchmarking of MSstats](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_MSStats.html)
-- [Benchmarking of proDA](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_proDA_medpolish.html)
-
-## Creating a prolfqua configuration {.unlisted .unnumbered}
-
-The following code demonstrates how we use `r BiocStyle::Githubpkg("fgcz/prolfqua")` to analyze protein intensities reported in the MSFragger `combined_protein.tsv` file.
-First, we create a tidy table containing the protein abundances by reading the `combined_protein.tsv` file using  `tidy_MSFragger_combined_protein.` Then, we read the sample annotation from the file `annotation.xlsx` file.
-Next, we create an `AnalysisTableAnnotation` R6 object.
-Bottom-up proteomics data is hierarchical, i.e., a protein has peptides, peptides might be modified, etc.
-Therefore, the `AnalysisTableAnnotation` has a `hierarchy` field storing a list with an entry for each hierarchy level.
-Since `combined_portein.tsv` only holds protein level data, the hierarchy list has one element, and we use it to specify which column contains the protein identifiers.
-We also need to define which column contains the protein abundances we want to use for the data analysis.
-Finally, we have to specify which columns contain the explanatory variables of the analysis.
-The `AnalysisTableAnnotation` has the field `factors,` a list with as many entries as explanatory variables.
-Here we include two explanatory variables, the dilution, specified in the column 'sample', and 'run' stored in the column 'run_ID', representing the order of the measurement.
-
-```{r echo=TRUE}
-datadir <- file.path(find.package("prolfquadata") , "quantdata")
-inputFragfile <-  file.path(datadir, "MSFragger_IonStar2018_PXD003881.zip")
-inputAnnotation <- file.path(datadir, "annotation_Ionstar2018_PXD003881.xlsx")
-# read input annotation
-annotation <- readxl::read_xlsx(inputAnnotation)
-
-protein <- tibble::as_tibble(
-    read.csv(unz(inputFragfile,"IonstarWithMSFragger/combined_protein.tsv"),
-             header = TRUE, sep = "\t", stringsAsFactors = FALSE))
-
-# read combined_protein.tsv 
-protein <- prolfqua::tidy_FragPipe_combined_protein_deprec(protein)
-# remove proteins identified by a single peptide
-protein <- protein |> dplyr::filter(unique.stripped.peptides > 1)
-
-# annotate the data
-merged <- dplyr::inner_join(annotation, protein)
-atable <- prolfqua::AnalysisTableAnnotation$new()
-atable$fileName = "raw.file"
-# specify column containing protein identifiers
-atable$hierarchy[["protein_Id"]] = "protein"
-
-# column with protein abundances
-atable$set_response("total.intensity")
-
-# the factors of the analysis
-atable$factors[["dilution."]] = "sample"
-atable$factors[["run"]] = "run_ID"
-
-config <- prolfqua::AnalysisConfiguration$new(atable)
-
-adata <- prolfqua::setup_analysis(merged, config)
-lfqdata <- prolfqua::LFQData$new(adata, config)
-# remove small intensities
-lfqdata$remove_small_intensities()
-
-```
-
 ## Session information {.unlisted .unnumbered}
@@
 sessionInfo()
-```
-
-## Miscellaneous {.unlisted .unnumbered}
-
-(ref:tabCompletion) The screenshot displays the command-line completion (tab completion) of RStudio on the `prolfqua::LFQData` R6 object. In the example, it shows the getter methods of the object.
-
-```{r tabCompletion, echo=FALSE, fig.cap="(ref:tabCompletion)", out.width = '66%'}
-knitr::include_graphics("codeSnippet1TabCompletion.png")
 ```
@@
 ```
-
-```{r roctest}
-ttmsqrob <- allBenchmarks$benchmark_msqrob$data()
-ddmsqrob <- pROC::roc(ttmsqrob$TP, ttmsqrob$scaled.p.value, partial.auc = c(1, 0.9))
-ttprolfqua <- allBenchmarks$benchmark_merged$data()
-ddprolfqua <- pROC::roc(ttprolfqua$TP, ttprolfqua$scaled.p.value, partial.auc = c(1, 0.9))
-ttproda <- allBenchmarks$benchmark_proDA$data()
-ddproda <- pROC::roc(ttproda$TP, ttproda$scaled.p.value, partial.auc = c(1, 0.9))
-
-msqrob_vs_prolfqua <- pROC::roc.test(ddmsqrob,ddprolfqua)
-msqrob_vs_proda <- pROC::roc.test(ddmsqrob,ddproda)
-prolfqua_vs_proda <- pROC::roc.test(ddprolfqua,ddproda)
-
-```

### 2022-10-21 inst/MyArticle/paper.Rmd
fix rename lfqp -> lfqpro

@@
 panelB <- agr$plot()$plots[[54]] + ggplot2::labs(tag = "B")
-panelC <- lfqp$get_Stats()$violin() + ggplot2::labs(tag = "C")
-pl <- lfqp$get_Plotter()
+panelC <- lfqpro$get_Stats()$violin() + ggplot2::labs(tag = "C")
+pl <- lfqpro$get_Plotter()
 panelD <- pl$boxplots()$boxplot[[54]] + ggplot2::labs(tag = "D")
@@
 ## fit model
-lmmodel <- paste(lfqp$response()," ~ dilution.")
+lmmodel <- paste(lfqpro$response()," ~ dilution.")
 modelFunction <- prolfqua::strategy_lm( lmmodel, model_name = "lm")
-models <- prolfqua::build_model(lfqp, modelFunction)
+models <- prolfqua::build_model(lfqpro, modelFunction)
 
@@
 # compute contrasts using imputation imputation
-conI <- prolfqua::ContrastsSimpleImpute$new( lfqp, contrasts, modelName = "imp")
+conI <- prolfqua::ContrastsSimpleImpute$new( lfqpro, contrasts, modelName = "imp")
 # merge contrasts, to obtain differences for all proteins

### 2022-10-21 inst/MyArticle/paper.Rmd
code examples fixed.

@@
       keep_tex: true
-      citation_package: natbib
 date: "`r format(Sys.time())`"
-bibliography: bibliography.bib
+bibliography: prolfqua.bib
 vignette: >
@@
 
-\begin{quote}
-To paraphrase provocatively, `machine learning is statistics minus any checking of models and assumptions'.
-
-\begin{flushright}
--- Brian D. Ripley, useR! 2004, Vienna.
-\end{flushright}
-\end{quote}
-
 Proteins carry out most crucial functions and give structure to living cells.
@@
 Each of them has some unique features, for example, `r BiocStyle::Biocpkg("MSstats")` determines the statistical model from the structure of the sample annotation, allowing users with limited statistical knowledge to perform differential expression analysis (DEA).
-At the same time, `r BiocStyle::Biocpkg("limma")` allows to specify a design matrix using a linear model formula and implements the empirical Bayes variance shrinkage method.
+At the same time, `r BiocStyle::Biocpkg("limma")` allows the specification of a design matrix using a linear model formula and implements the empirical Bayes variance shrinkage method.
 The package `r BiocStyle::Biocpkg("PECA")`, performs a roll-up of peptide level differences and peptide level $p$-value estimates, obtained from `r BiocStyle::Biocpkg("limma")` or `r BiocStyle::Biocpkg("PECA")`, to protein level estimates.
-Furthermore, `r BiocStyle::Biocpkg("msqrob2")` combines two models: robust linear models fitted to protein abundances and a generalized linear model fitted to presence absence information of peptides, into Hurdle model.
+Furthermore, `r BiocStyle::Biocpkg("msqrob2")` combines two models: robust linear models fitted to protein abundances and a quasibinomial generalized linear model fitted to peptide counts, into Hurdle model.
 The `r BiocStyle::Biocpkg("proDA")` package implements a linear probabilistic dropout model to account for missing data without imputation.
@@
 For example, to visualize tidy Spectronaut[@bruderer2015extending], DiaNN[@demichev2020dia], or Skyline[@maclean2010skyline] outputs, or data in `r BiocStyle::Biocpkg("MSstats")`[@Choi2014] format, only a few lines of code are needed to update the `r BiocStyle::Githubpkg("fgcz/prolfqua")` `AnalysisTableConfiguration` configuration.
-The configuration encapsulates the differences in column names among the various input formats, and enables the use of the methods without additional arguments.
+The configuration encapsulates the differences in column names among the various input formats and enables the use of the methods without additional arguments.
 We show an example code for creating an MSFragger[@yu2020fast] configuration in the Appendix.
@@
 We use R6 classes to structure the functionality of the package (see Figure \@ref(fig:LFQData) and Figure \@ref(fig:ContrastUML)).
-R6 classes are well supported by command-line completion features in RStudio [@RStudio2022], and help to implement argument free functions (Figure \@ref(fig:tabCompletion).
+R6 classes are well supported by command-line completion features in RStudio [@RStudio2022], and help to implement argument free functions (Supplement Figure SX).
 
@@
 
-R's formula interface for linear models is flexible, widely used, and well documented [@faraway2016extending].
+R's formula interface for linear models is flexible, widely used, and well documented [@faraway2016extending,@law2020guide].
 We use the formula interface to specify the models, making it easy to reproduce an analysis performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` in other statistical programming languages.
@@
 coli_.
-We annotated these dilutions from smallest to largest with the letters _a_, _b_, _c_, _d_, _e_.
+We annotated these dilutions from smallest to largest with the letters _a_, _b_, _c_, _d_ and _e_.
 By comparing the various dilutions, we obtain different effect sizes, e.g., when comparing dilution _e_ (9%) against dilution _d_ (7.5%), the expected difference is $1.2$ for E.
 coli proteins and $1$ for human proteins.
-There are four technical replicates for each dilution, and hence 20 raw files in total.
+There are four technical replicates for each dilution, hence 20 raw files in total.
 We processed the raw data using MaxQuant [@Cox2008MaxQuant] version Version 1.6.10.43, with MaxQuant default settings.
@@
 
-By plotting the $TPR$ versus the $FPR$ we obtain the receiver operator characteristic curve (ROC curve).
+In order to compute the confusion matrices based on the $p$-value we first need to rescale it (see Equation \@ref(eq:scalepvalue)).
+
+By plotting the $TPR$ versus the $FPR$ we obtain the receiver operator characteristic curve (ROC curve) [@pROC_2011].
 The area under the curve ($AUC$) or partial areas under the curve ($pAUC$), at various values of the $FPR$, are measures of performance derived from the ROC curve.
-By using these measures we can compare the performances of the statistics produced by the various methods examined.
-
-In order to compute the confusion matrices based on the $p$-value we first need to rescale it (see Equation \@ref(eq:scalepvalue)).
-Thus, the $p$-value is a function of the $t$-statistics and the degrees of freedom.
+Using these measures, we can compare the performances of the statistics or the various models examined and test if these differences are statistically significant.
 
@@
   "diff_batch" = "batch_p2370 - batch_p2691",
-  "difference_within_p2370" = "`condition_Glucose:batch_p2370` - `condition_Ethanol:batch_p2691`",
-  "difference_within_p2691" = "`condition_Glucose:batch_p2691` - `condition_Ethanol:batch_p2691`",
-  "difference_of_difference" = "difference_within_p2370 - difference_within_p2691")
+  "diff_within_p2370" = "`condition_Glucose:batch_p2370` - `condition_Ethanol:batch_p2691`",
+  "diff_within_p2691" = "`condition_Glucose:batch_p2691` - `condition_Ethanol:batch_p2691`",
+  "diff_of_difference" = "diff_within_p2370 - diff_within_p2691")
 
@@
 ```{r contrcoefficient}
-knitr::kable(mC, caption = "Weights $c$ of each model parameter (columns) to compute contrast (rows).",format = "latex")
+knitr::kable(mC, caption = "Weights $c$ for each of the contrasts (rows) which will be applied to the model parameter (columns).",format = "latex")
 
@@
 The code snippets in this section demonstrate how a differential expression analysis workflow can be implemented using the `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package.
-To speed up the computation of these examples, we use subset of the Ionstar dataset generated by randomly selecting $400$ proteins.
-Peptide abundances are $\log_2$ transformed and robust z-score scaled using the method `robscale`.
-Using the `LFQDataPlotter` class, we show the distribution of the normalized peptide abundances in Figure \@ref(fig:prepro) Panel A.
+To speed up the computation of these examples, we use a subset of the Ionstar dataset generated by randomly selecting $400$ proteins.
+First, we remove all proteins with a single peptide and all observations for which MaxQuant reports zero intensities which leaves $332$ proteins.
+Next, peptide abundances are $\log_2$ transformed and robust z-score scaled using the method `robscale`.
+Then, using the `LFQDataPlotter` class, we show the distribution of the normalized peptide abundances in Figure \@ref(fig:prepro) Panel A.
 Afterwards, protein intensities are estimated from peptide intensities using Tukey's median polish.
@@
 For each sample a line with a different colour is shown.
-Panel B - Peptide intensities for protein _5NTD_ are shown using a line of different colour, and the protein intensity estimate is shown using a fat black line, Panel C - distribution of standard deviations of all proteins in each dilution group ($a$, $b$, $c$, $d$, $e$) and overall (all), Panel D - Distribution of protein intensities of Protein _5NTD_ in each dilution group.
+Panel B - Peptide intensities for protein HFQ_ECOLI are shown using a line of different colour, and the protein intensity estimate is shown using a fat black line, Panel C - distribution of standard deviations of all proteins in each dilution group ($a$, $b$, $c$, $d$, $e$) and overall (all), Panel D - Distribution of protein intensities of Protein HFQ_ECOLI in each dilution group.
 
@@
 Afterward, we apply $t$-statistic moderation using the `ContrastModerated` class.
-Finally, the `addContrastResults` function merges both sets of contrast estimates, preferring the one obtained from the linear model if both are available.
+Finally, the `merge_contrasts_results` function merges both sets of contrast estimates, preferring the one obtained from the linear model if both are available.
 Then we create the plots shown in Figure \@ref(fig:exampleContrasts).
-Panel A shows the distribution of the $p$-values, while Panel B shows the volcano plot for each comparison.
+Panel A shows the distribution of the $p$-values, Panel B shows the volcano plot for each comparison, and Panel C shows a Bland-Altman plot which shows the difference between groups as a function of the rank of the protein abundance.
 
 (ref:exampleContrasts) Panel A - Histogram showing the distribution of p-values for $332$ proteins.
-Panel B -  Volcano plot showing $-\log_{10}$ transformed $FDR$ as function of the difference between groups for $332$ proteins.
-With black dots we show effect size and FDR estimates obtained from the linear model, while in green we show those obtained using imputation.
+Panel B -  Volcano plot showing $-\log_{10}$ transformed $FDR$ as a function of the difference between groups for $332$ proteins.
+With black dots, we show effect size and FDR estimates obtained from the linear model, while in green, we plot those obtained using imputation.
+Finally, panel C shows the difference between groups as a function of the rank of the abundance of the proteins.
 
@@
 Furthermore, this glass box approach should make it easy to reimplement an analysis performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` using base R or other programming languages by reading the analysis script.
-However, acknowledging the R formula interface’s complexity to non-statisticians and the popularity of `r BiocStyle::Biocpkg("MSstats")`, we provide the functionality to suggest a model formula from the sample annotation provided in tabular form similarly to `r BiocStyle::Biocpkg("MSstats")`.
 
-Using the table of tidy data ensures interoperability with other proteomics-related packages that manage their data with tidy-tables, e.g., `r BiocStyle::CRANpkg("protti")` [@quast2022protti].
+Using the data frame of tidy data ensures interoperability with other proteomics-related packages that manage their data with tidy-tables, e.g., `r BiocStyle::CRANpkg("protti")` [@quast2022protti].
 To simplify integration of `r BiocStyle::Githubpkg("fgcz/prolfqua")` with Bioconductor based workflows there is a method to convert the `LFQData` class into a `r BiocStyle::Biocpkg("SummarizedExperiment")`.
-The use of R6 classes, which encapsulate the configuration and the data, allows for writing very concise code, where functions can have only very few arguments.
+The use of R6 classes, which encapsulate the configuration and the data, allows for writing very concise code, where functions can have few arguments.
 Auto-completion support for R6 classes in the Rstudio editor makes it easy for novices to explore `r BiocStyle::Githubpkg("fgcz/prolfqua")`s functionality (see Supplement Figure S[X]).
 
-To ease the usage barriers of the R package to users not being familiar with statistics and R programming, we integrated some workflows into our data management platform [B-Fabric](https://fgcz-bfabric.uzh.ch/) [@bfabric,].
-This integration enables our users to select the input and basic settings in a graphical user interface.
+To ease the usage barriers of the R package to users not familiar with statistics and R programming, we developed an application based on the prolfqua package into our data management platform [B-Fabric](https://fgcz-bfabric.uzh.ch/) [@bfabric,@bfabricjib2022].
+The B-Fabric system runs a computing infrastructure controlled by a local resource management system that supports cloud-bursting [@VMMAD].
+This integration enables users to select the input data and basic settings in a graphical user interface.
+We determine the model formula and which comparisons to compute from the sample annotation provided in tabular form, similarly to SAINTexpress and `r BiocStyle::Biocpkg("MSstats")`.
+However, only parallel group designs with and without repeated measurements are analyzed.
 The user receives a report including input files, the R markdown file, and R scripts necessary to replicate the analysis using their in-house R installation.
-The B-Fabric system runs a computing infrastructure controlled by a local resource management system that supports cloud-bursting [@VMMAD].
-In this way, [B-Fabric](https://fgcz-bfabric.uzh.ch/) helps scientists to meet requirements from funding agencies, journals, and academic institutions to publish data according to the FAIR (Findable, Accessible, Interoperable and Reusable) [@FAIR] data principles.
+In this way, [B-Fabric](https://fgcz-bfabric.uzh.ch/) helps scientists to meet requirements from funding agencies, journals, and academic institutions to publish data according to the FAIR (Findable, Accessible, Interoperable, and Reusable) [@FAIR] data principles.
+We are working on creating a stand-alone shiny application with the described functionality.
 
@@
 We use those scores, i.e.
-the $pAUC$ at $10%$ $FDR$ and the $FDP$, to examine how well the methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` model quantitative mass spectrometric high throughput data and compare them with results produced by `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`.
+the $pAUC$ at $10%$ $FDR$ and the $FDP$, to examine how well the methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` model quantitative mass spectrometric high throughput data and compare them with results produced by `r BiocStyle::Biocpkg("MSstats")`, `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Biocpkg("msqrob2")`.
 Table \@ref(tab:describeBenchmarkedModels) summarizes all modelling methods we evaluated.
@@
 However, some methods can not estimate all of them.
-The set of proteins with effect size estimates might differ for each method, biasing direct comparison of scores such as $pAUC$.
+The set of proteins with effect size estimates might differ for each method, biasing direct comparison of scores such as the $pAUC$.
 For instance, we observe that `r BiocStyle::Biocpkg("MSstats")` estimates $16058$ group differences while the mixed effect models estimates the fewest with $15940$.
@@
 Interestingly, the $pAUC$ increases when using $p$-values instead of the $t$-statistics for linear models, while it decreases for mixed effect models.
-The reason is an erroneous denominator degrees of freedom estimation for many proteins in the case of the mixed effect models.
+The reason is an erroneous denominator degree of freedom estimation for many proteins, a known problem in the case of mixed effect models.
 
 We also benchmark if the $FDR$ obtained from a model is an unbiased estimate of the false discovery proportion $FDP$.
-Figure \@ref(fig:FDRfdp) Panel C draws on the horizontal axis the $FDR$ determined from the model, and on the vertical axis, the $FDP$ obtained from the confusion matrix.
-Most lines are below the diagonal, indicating that the $FDR$ estimates are modestly conservative for this benchmark dataset.
+Figure \@ref(fig:FDRfdp) Panel C shows the $FDP$, obtained from the confusion matrix, as a function of the $FDR$ determined from the model.
+Most lines are below the diagonal, indicating that the $FDR$ estimates are modestly conservative for this particular benchmark dataset.
 In the case of `r BiocStyle::Biocpkg("MSstats")`, we observe a high proportion of false discoveries for small $FDR$ values.
-In the case of `r BiocStyle::Biocpkg("PECA")`, the $FDR$ estimates, obtained by Benjamini-Hochberg correcting the regulation probabilities, strongly overestimate the $FDP$.
+In the case of the ROPECA method, the $FDR$ estimates, obtained by applying the Benjamini-Hochberg correction to the Beta distribution-based regulation probabilities, strongly overestimate the $FDP$.
 
@@
 ddb <- dd |> dplyr::filter(contrast == "all")
+ddb$FDP <- NULL
+ddb <- ddb |> dplyr::rename(FDR = scorecol, FDP = FDP_)
 
-pC <- ddb |> ggplot2::ggplot(ggplot2::aes(y = FDP_,  x  = scorecol )) + 
+pC <- ddb |> ggplot2::ggplot(ggplot2::aes(y = FDP,  x  = FDR )) + 
   ggplot2::geom_line(ggplot2::aes(color = model_name, linetype = model_name)) +
@@
   ggplot2::theme(legend.position = "bottom") +
-  ggplot2::labs(tag = "C")
-
+  ggplot2::labs(tag = "C") + ggplot2::xlim(c(0,0.5)) + ggplot2::ylim(c(0,0.5))
 hlay = rbind(c(2,1,1),
@@
     pic = c("runSaint, ContrastsSaintExpress", "protein interaction scoring (SAINTExpress)"),
-    prodA = c( "strategy_proDA*, Contrasts_proDA*", "adapter to the probabilistic dropout model implemented in proDA"), check.names =  FALSE)
+    prodA = c( "strategy_proDA*, ContrastsProDA", "adapter to the probabilistic dropout model implemented in proDA"), check.names =  FALSE)
 xM <- data.frame(t(xM))
@@
 
-Using a benchmark dataset whose ground truth is known (see Methods), we assessed the performance of different modeling approaches implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` (Tables \@ref(tab:prolfquaModels) and \@ref(tab:describeBenchmarkedModels)), `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`.
-Table \@ref(tab:describeBenchmarkedModels) summarizes which methods we have evaluated, which MaxQuant results we used, and if the models are fitted to peptide or protein intensities.
+Using a benchmark dataset whose ground truth is known (see Methods), we assessed the performance of different modeling approaches implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` (Tables \@ref(tab:prolfquaModels) and \@ref(tab:describeBenchmarkedModels)), `r BiocStyle::Biocpkg("MSstats")`, `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Biocpkg("msqrob2")`.
+Table \@ref(tab:describeBenchmarkedModels) summarizes which methods we have evaluated, which MaxQuant files we used as input, and if the models are fitted to peptide or protein intensities.
 We make the R-markdown files to replicate the benchmarking available at `r BiocStyle::Githubpkg("wolski/prolfquabenchmark")`.
@@
     msstats = c("MSstats", "preprocess with default parameters",  "", "evidence.txt"),
+    msqrob = c("msqrob2", "hurdle msqrob model", "protein and peptide", "peptide.txt"),
     proDA = c("proDA", "probabilistic dropout model",  "protein", "peptide.txt"),
@@
 
-```{r roctest, include = FALSE}
+```{r roctest, include = FALSE, eval = FALSE}
 ttmsqrob <- allBenchmarks$benchmark_msqrob$data()
@@
 Since only technical replicates are available for each dilution, essential sources of variation present in every experiment, such as the biochemical and biological, are not measured.
-Therefore, the dataset captures only the variance from the chromatography, electro-spray, and mass spectrometric measurement method.
-Thus, while we can extrapolate some of the results obtained to more realistic datasets, we need to be careful not to over-interpret the results.
+Therefore, the IonStar dataset (see Methods) captures only the variance from the chromatography, electro-spray, and mass spectrometric measurements.
+Furthermore, this benchmark with a parallel group design does not allow for benchmarking models with interactions.
+Unfortunately,  other benchmark datasets have similar constraints [@edwards2015cptac, @o2018proteome].
+At the same time, we could not analyze the benchmark dataset [@wessels2012comprehensive] with a full factorial design with an up-to-date version of MaxQuant.
+Thus, while we can extrapolate some of the results obtained to more realistic datasets, we need to be careful not to over-interpret.
 Specifically, the observed variances will be higher in more realistic experiments, and the power will be lower for the same sample sizes.
@@
 We can conclude that if we want to sort the proteins according to the likelihood of being differentially regulated to perform gene set enrichment analysis [@subramanian2005gene], the $t$-statistic is better suited than the fold-change estimate.
-Modeling the degrees of freedom when computing the $p$-values might improve the inference.
-However, this improvement is minuscule (see Figure \@ref(fig:FDRfdp) panel B).
-There is no such improvement for the mixed effect model, most likely because the degrees of freedom are erroneously estimated for many models.
-Furthermore, for the fixed effect linear model, the empirical Bayes variance shrinkage, as suggested by @Smyth2004linear, consistently improves the ranking of proteins compared with the unmoderated estimates (not shown) and fails to do so for mixed effect models.
+Modeling the degrees of freedom when computing the $p$-values improves the inference (see Figure \@ref(fig:FDRfdp) panel B).
+There is no such improvement for the mixed effect model, because the degrees of freedom are erroneously estimated.
+Furthermore, for the fixed effect linear model, the empirical Bayes variance shrinkage, as suggested by @Smyth2004linear, consistently improves the ranking of proteins compared with the unmoderated estimates (not shown).
+However, since also for this method, a correct degree of freedom estimate is required, it does not work for mixed effect models.
 
@@
 Furthermore, by using the Beta distribution to model the number of peptides observed, we can further improve the $pAUC$ scores (see Figure \@ref(fig:FDRfdp) Panel A center).
-However, the properties of Beta-based probabilities are not well understood; for instance, the p-values are not uniformly distributed under the null hypothesis (not shown).
+However, the properties of Beta-based probabilities are not well understood; for instance, the p-values are not uniformly distributed under the null hypothesis.
 Furthermore, the $FDR$ estimates obtained when correcting for multiple testing with the Benjamini-Hochberg method are biased and overestimate the false discovery proportion (see Figure \@ref(fig:FDRfdp) Panel C).
@@
 
-The probabilistic dropout analysis implemented in the `r BiocStyle::Biocpkg("proDA")` produces inferences comparable to those of other methods (Figure \@ref(fig:FDRfdp) Panel A).
-Because of the robustness of the dropout model to missing observations we obtain difference estimates for all proteins and contrasts (Figure \@ref(fig:FDRfdp) Panel B).
-Moreover, the estimated differences are unbiased and show high diagnostic accuracy (Figure \@ref(fig:FDRfdp) Panel A).
+The probabilistic dropout analysis implemented in  `r BiocStyle::Biocpkg("proDA")`, as well as the hurdle model implemented in `r BiocStyle::Biocpkg("msqrob2")`, produce inferences comparable to those of other methods (Figure \@ref(fig:FDRfdp) Panel A).
+Because of the robustness of the dropout and hurdle model to missing observations we obtain difference estimates for all proteins and contrasts (Figure \@ref(fig:FDRfdp) Panel B).
+Moreover, the estimated differences of the probabilistic dropout model are unbiased and show high diagnostic accuracy (Figure \@ref(fig:FDRfdp) Panel A).
 Furthermore, the performance of the scaled $p$-values or the $t$-statistics is comparable with that of the linear model with variance moderation (`prot_med_lm_moderated`).

### 2022-10-21 inst/MyArticle/paper.Rmd
reviewed text. Still references to supplement are missing.

@@
 # Introduction
+
+\begin{quote}
+To paraphrase provocatively, `machine learning is statistics minus any checking of models and assumptions'.
+\begin{flushright}
+-- Brian D. Ripley, useR! 2004, Vienna.
+\end{flushright}
+\end{quote}
 
@@
 
-The probabilistic dropout analysis implemented in  `r BiocStyle::Biocpkg("proDA")`, as well as the hurdle model implemented in `r BiocStyle::Biocpkg("msqrob2")`, produce inferences comparable to those of other methods (Figure \@ref(fig:FDRfdp) Panel A).
-Because of the robustness of the dropout and hurdle model to missing observations we obtain difference estimates for all proteins and contrasts (Figure \@ref(fig:FDRfdp) Panel B).
-Moreover, the estimated differences of the probabilistic dropout model are unbiased and show high diagnostic accuracy (Figure \@ref(fig:FDRfdp) Panel A).
-Furthermore, the performance of the scaled $p$-values or the $t$-statistics is comparable with that of the linear model with variance moderation (`prot_med_lm_moderated`).
-Therefore, we are planning to integrate the `r BiocStyle::Biocpkg("proDA")` package as an additional modeling option into `r BiocStyle::Githubpkg("fgcz/prolfqua")` (see Figure \@ref(fig:ContrastUML)).
-
-The R-package `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Biocpkg("prolfqua")` model the missing data directly, while MSstats imputes the data using an accelerated failure model.
-Despite imputation, `r BiocStyle::Biocpkg("MSstats")` does not estimate group differences for more proteins and does not achieve a higher $pAUC$ score than `r BiocStyle::Githubpkg("fgcz/prolfqua")`.
+The R-package `r BiocStyle::Biocpkg("proDA")`, `r BiocStyle::Biocpkg("msqrob2")` and `r BiocStyle::Biocpkg("prolfqua")` (see Methods) model the missing data directly, while MSstats imputes the data using an accelerated failure model.
+Despite imputation, `r BiocStyle::Biocpkg("MSstats")` estimates group differences for less proteins and does not achieve a higher $pAUC_{10}$.
 Furthermore, Figure \@ref(fig:FDRfdp) Panel C shows that when using `r BiocStyle::Biocpkg("MSstats")` the proportion of false discoveries might be very high even when filtering for a low $FDR$.
 
-We focused our benchmark on comparing the statistical modeling methods using three different scores, while we fixed the pre-processing steps.
+We obtained difference estimates for all proteins, as shown in Figure \@ref(fig:FDRfdp) Panel A when using the probabilistic dropout model (`r BiocStyle::Biocpkg("proDA")`), the hurdle model (`r BiocStyle::Biocpkg("msqrob2")`), and when combining the results of the linear model with contrast estimates in the presence of missing data (prolfqua_merged).
+Furthermore, the performance of the scaled $p$-values or the $t$-statistics is comparable with that of the other methods (Figure \@ref(fig:FDRfdp) Panel B).
+All three models perform similarly well when benchmarked against the IonStar/MaxQuant, IonStar/FragPipe, or CPTAC dataset (supplement).
+We tested if the $pAUC_{10}$ differences for the three methods were statistically significant, and none of them were at a significance level of $0.4$ (see supplement).
+Furthermore, for the IonStar/MaxQuant dataset, msqrob2 performed best, proDA performed best for the CPTAC/MaxQuant data, and prolfqua_merged for the IonStar/FragPipe dataset.
+Also, the FDR estimates (Figure \@ref(fig:FDRfdp) Panel C) are comparable for all three methods.
+
+Since, the API of `r BiocStyle::Biocpkg("proDA")` and the produced outputs are similar to those of prolfqua, we are planning to integrate the `r BiocStyle::Biocpkg("proDA")` package as an additional modeling option into `r BiocStyle::Githubpkg("fgcz/prolfqua")` (see Figure \@ref(fig:ContrastUML)), to make it easy for application developers to use both.
+
+We focused our benchmark on comparing the statistical modeling methods using three different scores while we fixed the pre-processing steps.
 However, there are other equally or even more important parameters of a protein quantification pipeline [@frohlich2022benchmarking].
@@
 For instance, the original `r BiocStyle::Biocpkg("proDA")` publication uses MaxLFQ [@cox2014accurate] protein estimates.
-However, when using MaxLFQ intensities reported by MaxQuant, the $pAUC_{10}$ is significantly lower ($pAUC_{10}(\textrm{t-statistics}) = 66%$) compared with results obtained when protein abundances are estimated from peptide abundances using Tukey's median polish ($pAUC_{10}(\textrm{t-statistics}) = 73%$).
-Last but not least, the software [@Cox2008MaxQuant; @yu2020fast], identifying proteins and generating the quantification values can also significantly contribute to the performance of the entire pipeline, altering the number of identified proteins and the sensitivity and specificity of the differential expression analysis.
+However, when using MaxLFQ intensities reported by MaxQuant, the $pAUC_{10}$ is significantly lower ($pAUC_{10}(\textrm{t-statistics}) = 66%$) compared with results obtained when protein abundances are estimated from peptide abundances using Tukey's median polish ($pAUC_{10}(\textrm{t-statistics}) = 75%$).
+Last but not least, the software [@Cox2008MaxQuant; @yu2020fast], identifying proteins and generating the quantification values, significantly contributes to the entire pipeline's performance, altering the number of identified proteins and the sensitivity and specificity of the differential expression analysis.
+In the supplement we show the benchmarking results for `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Githubpkg("fgcz/prolfqua")` when the IonStar dataset is prepossessed with the MSFragger/IonQuant [@yu2020fast] quantification pipeline (v15).
+Although the number of proteins identified with two peptides is slightly lower than with MaxQuant (3929 vs.
+4046), the $pAUC_{10}$ is significantly higher ($81%$ vs.
+$72%$).
 
@@
 
-`r BiocStyle::Githubpkg("fgcz/prolfqua")` is an feature-rich, object oriented R package to analyze quantitative mass spectrometric data with simple or complex experimental designs.
+`r BiocStyle::Githubpkg("fgcz/prolfqua")` is a feature-rich, object-oriented R package to analyze quantitative mass spectrometric data with simple or complex experimental designs.
 While other R packages for differential expression analysis of proteins typically only implement one modeling approach, `r BiocStyle::Githubpkg("fgcz/prolfqua")` supports various types of models (see Figure 2 and Table 3).
-The contrast specification is explicit and consistent for all models.
-The modular design of `r BiocStyle::Githubpkg("fgcz/prolfqua")` allows for adding new features, e.g., generalized linear models to model the presence or absence information of a protein, or robust linear models, in the future.
+Furthermore, the contrast specification is explicit and consistent for all models and allows for testing interactions.
+Furthermore, the modular design of `r BiocStyle::Githubpkg("fgcz/prolfqua")` enables adding new features, e.g., generalized linear models to model the presence or absence information of a protein, or robust linear models, in the future.
 Furthermore, the developed framework is flexible enough to integrate other modeling methods, e.g., the probabilistic dropout model [@bioxrvproDA2020] or accurate variance estimation [@zhu2020deqms].
-Hence, `r BiocStyle::Githubpkg("fgcz/prolfqua")` enables you to call various methods and quickly select the best differential expression analysis algorithm for your problem.
-It also can generate conclusive reports and can be used to benchmark MS software and statistical methods.
+Hence, `r BiocStyle::Githubpkg("fgcz/prolfqua")` enables the implementation of applications where the user can select an alternative normalization method, protein abundance estimation method, or differential expression analysis algorithm.
 Furthermore, with minimal adaptations, this R package can analyze different quantitative proteomics data (e.g., labeling-based TMT-, PRM- DIA-data).
@@
 It is relevant that an analysis pipeline shows good performance in all these categories.
-Leveraging these computer experiments, we can provide the following advice: i) estimate protein abundances from peptide abundances using a robust or nonparametric regression method; ii) use linear models because they show good performance in all categories; iii) if the measurements are correlated, as for technical replicates, mixed effect models might work if the sample sizes are large; if not, aggregate the replicates and fit a linear model instead; iv) if you use fixed-effect linear models, apply variance moderation to improve the $t$-statistics and $p$-value estimates; v) If you want to sort your protein lists to perform gene set enrichment analysis, use the $t$-statistic instead of the difference; vi) do not impute missing observation but statistically model missingness to estimate parameters, i.e., group differences.
-Finally, the differential expression analysis results obtained with prolfqua are comparable to or better than those of other differential expression analysis tools.
+Leveraging these computer experiments, we can provide the following advice: i) estimate protein abundances from peptide abundances using a robust or nonparametric regression method; ii) use linear models because they show good performance in all categories; iii) if the measurements are correlated, as for technical replicates, mixed effect models might work if the sample sizes are large; if not, aggregate the replicates and fit a linear model instead; iv) if you use fixed-effect linear models, apply variance moderation to improve the $t$-statistics and $p$-value estimates; v) If you want to sort your protein lists to perform gene set enrichment analysis, use the $t$-statistic instead of the difference; and last but not least vi) do not impute missing observation but statistically model missingness to estimate parameters, i.e., group differences.
+Finally, the differential expression analysis results obtained with prolfqua are comparable to those of other differential expression analysis tools.
 

### 2022-10-21 inst/MyArticle/paper.Rmd
changes to conclusion.

@@
 We use R6 classes to structure the functionality of the package (see Figure \@ref(fig:LFQData) and Figure \@ref(fig:ContrastUML)).
-R6 classes are well supported by command-line completion features in RStudio [@RStudio2022], and help to implement argument free functions (Supplement Figure SX).
+R6 classes are well supported by command-line completion features in RStudio [@RStudio2022], and help to implement argument free functions (see Supplementary Figure SX).
 
@@
 The use of R6 classes, which encapsulate the configuration and the data, allows for writing very concise code, where functions can have few arguments.
-Auto-completion support for R6 classes in the Rstudio editor makes it easy for novices to explore `r BiocStyle::Githubpkg("fgcz/prolfqua")`s functionality (see Supplement Figure S[X]).
+Auto-completion support for R6 classes in the Rstudio editor makes it easy for novices to explore `r BiocStyle::Githubpkg("fgcz/prolfqua")`s functionality (see Supplementary Figure S[X]).
 
@@
 Furthermore, the performance of the scaled $p$-values or the $t$-statistics is comparable with that of the other methods (Figure \@ref(fig:FDRfdp) Panel B).
-All three models perform similarly well when benchmarked against the IonStar/MaxQuant, IonStar/FragPipe, or CPTAC dataset (supplement).
-We tested if the $pAUC_{10}$ differences for the three methods were statistically significant, and none of them were at a significance level of $0.4$ (see supplement).
+All three models perform similarly well when benchmarked against the IonStar/MaxQuant, IonStar/FragPipe, or CPTAC dataset (see Supplement, Benchmark CPTAC).
+We tested if the $pAUC_{10}$ differences for the three methods were statistically significant, and none of them were at a significance level of $0.4$ (see Supplement, Testing ROC curves).
 Furthermore, for the IonStar/MaxQuant dataset, msqrob2 performed best, proDA performed best for the CPTAC/MaxQuant data, and prolfqua_merged for the IonStar/FragPipe dataset.

### 2022-10-21 inst/MyArticle/paper.Rmd
using figure1 figure2 figure3

@@
       keep_tex: true
+      #citation_package: natbib
 date: "`r format(Sys.time())`"
@@
 
-```{r contrast2, eval=TRUE, echo=TRUE}
+```{r contrast2, eval = TRUE, echo = TRUE}
 mlm <- prolfqua::strategy_lm(
@@
 Next, peptide abundances are $\log_2$ transformed and robust z-score scaled using the method `robscale`.
-Then, using the `LFQDataPlotter` class, we show the distribution of the normalized peptide abundances in Figure \@ref(fig:prepro) Panel A.
+Then, using the `LFQDataPlotter` class, we show the distribution of the normalized peptide abundances in Figure \@ref(fig:figure1) Panel A.
 Afterwards, protein intensities are estimated from peptide intensities using Tukey's median polish.
-Figure \@ref(fig:prepro) Panel B shows the peptide intensities and the estimated protein intensities.
+Figure \@ref(fig:figure1) Panel B shows the peptide intensities and the estimated protein intensities.
 Next, we compute the standard deviation of all the proteins in each group and display their distribution using violin plots (Panel C).
@@
 
-```{r prepro,  echo = TRUE , fig.cap="(ref:scaling)", out.width = '90%', fig.width = 8, fig.height = 8}
+```{r figure1,  echo = TRUE , fig.cap="(ref:scaling)", out.width = '90%', fig.width = 8, fig.height = 8}
 R.version.string; packageVersion("prolfqua")
@@
 Finally, the `merge_contrasts_results` function merges both sets of contrast estimates, preferring the one obtained from the linear model if both are available.
-Then we create the plots shown in Figure \@ref(fig:exampleContrasts).
+Then we create the plots shown in Figure \@ref(fig:figure2).
 Panel A shows the distribution of the $p$-values, Panel B shows the volcano plot for each comparison, and Panel C shows a Bland-Altman plot which shows the difference between groups as a function of the rank of the protein abundance.
 
-(ref:exampleContrasts) Panel A - Histogram showing the distribution of p-values for $332$ proteins.
+(ref:figure2) Panel A - Histogram showing the distribution of p-values for $332$ proteins.
 Panel B -  Volcano plot showing $-\log_{10}$ transformed $FDR$ as a function of the difference between groups for $332$ proteins.
@@
 
-```{r exampleContrasts, fig.cap="(ref:exampleContrasts)", echo=TRUE, out.width = '90%', fig.width = 8, fig.height = 5}
+```{r figure2, fig.cap="(ref:figure2)", echo=TRUE, out.width = '90%', fig.width = 8, fig.height = 5}
 # specify differences among groups
@@
 
-When comparing differential expression analysis performance, a relevant parameter is the number of proteins for which a method estimated differences (see Figure \@ref(fig:FDRfdp) Panel A), which indicates how robust the procedure works in the presence of missing observations.
+When comparing differential expression analysis performance, a relevant parameter is the number of proteins for which a method estimated differences (see Figure \@ref(fig:figure3) Panel A), which indicates how robust the procedure works in the presence of missing observations.
 For each protein, we tried to determine four differences ($\Delta = (1.20, 1.25, 1.30, 1.50)$).
@@
 
-Figure \@ref(fig:FDRfdp) Panel B shows how various estimates obtained from the models, i.e., the difference between groups, $t$-statistics, and scaled $p$-values allow identifying true positives (TP) given a false positive rate (FPR) of $10%$, by displaying the partial area under the ROC.
+Figure \@ref(fig:figure3) Panel B shows how various estimates obtained from the models, i.e., the difference between groups, $t$-statistics, and scaled $p$-values allow identifying true positives (TP) given a false positive rate (FPR) of $10%$, by displaying the partial area under the ROC.
 Ordering proteins by the $t$-statistic or $p$-value leads to a higher $pAUC_{10}$ than when ordering by the estimated difference among groups.
@@
 
-There are only minor differences in the $pAUC_{10}$ between the $t$-statistics or the scaled $p$-value  (see Figure \@ref(fig:FDRfdp) Panel A).
+There are only minor differences in the $pAUC_{10}$ between the $t$-statistics or the scaled $p$-value  (see Figure \@ref(fig:figure3) Panel A).
 Interestingly, the $pAUC$ increases when using $p$-values instead of the $t$-statistics for linear models, while it decreases for mixed effect models.
@@
 We also benchmark if the $FDR$ obtained from a model is an unbiased estimate of the false discovery proportion $FDP$.
-Figure \@ref(fig:FDRfdp) Panel C shows the $FDP$, obtained from the confusion matrix, as a function of the $FDR$ determined from the model.
+Figure \@ref(fig:figure3) Panel C shows the $FDP$, obtained from the confusion matrix, as a function of the $FDR$ determined from the model.
 Most lines are below the diagonal, indicating that the $FDR$ estimates are modestly conservative for this particular benchmark dataset.
@@
 
-(ref:FDRfdp) Panel A - Number of estimated contrasts for each modeling method (higher is better).
+(ref:figure3) Panel A - Number of estimated contrasts for each modeling method (higher is better).
 Panel B - Partial area under the ROC curve at $10\%$ FPR ($pAUC_{10}$) for all contrasts and three different statistics: the difference among groups (diff - Panel B left), the scaled $p$-value (sign(diff) $\cdot$ p.value) (scaled.p.value - Panel B center) and the $t$-statistics (statistic - Panel B right), where a higher pAUC is better.
@@
 
-```{r FDRfdp, fig.cap = "(ref:FDRfdp)", out.width = '90%', fig.width=12, fig.height=10}
+```{r figure3, fig.cap = "(ref:figure3)", out.width = '90%', fig.width=12, fig.height=10}
 dd <- purrr::map_df(allBenchmarks, function(x){res <- x$smc$summary; res$name <- x$model_name;res})
@@
 We can conclude that if we want to sort the proteins according to the likelihood of being differentially regulated to perform gene set enrichment analysis [@subramanian2005gene], the $t$-statistic is better suited than the fold-change estimate.
-Modeling the degrees of freedom when computing the $p$-values improves the inference (see Figure \@ref(fig:FDRfdp) panel B).
+Modeling the degrees of freedom when computing the $p$-values improves the inference (see Figure \@ref(fig:figure3) panel B).
 There is no such improvement for the mixed effect model, because the degrees of freedom are erroneously estimated.
@@
 
-Computing the statistics at the peptide level, e.g., the $t$-statistics or $p$-value, then summarizing these statistics using their median produces the highest $AUC$ scores among all the tested models (see Figure \@ref(fig:FDRfdp) Panel A left).
-Furthermore, by using the Beta distribution to model the number of peptides observed, we can further improve the $pAUC$ scores (see Figure \@ref(fig:FDRfdp) Panel A center).
+Computing the statistics at the peptide level, e.g., the $t$-statistics or $p$-value, then summarizing these statistics using their median produces the highest $AUC$ scores among all the tested models (see Figure \@ref(fig:figure3) Panel A left).
+Furthermore, by using the Beta distribution to model the number of peptides observed, we can further improve the $pAUC$ scores (see Figure \@ref(fig:figure3) Panel A center).
 However, the properties of Beta-based probabilities are not well understood; for instance, the p-values are not uniformly distributed under the null hypothesis.
-Furthermore, the $FDR$ estimates obtained when correcting for multiple testing with the Benjamini-Hochberg method are biased and overestimate the false discovery proportion (see Figure \@ref(fig:FDRfdp) Panel C).
+Furthermore, the $FDR$ estimates obtained when correcting for multiple testing with the Benjamini-Hochberg method are biased and overestimate the false discovery proportion (see Figure \@ref(fig:figure3) Panel C).
 Therefore, we can not recommend this method if an unbiased estimate of $FDR$ is essential, which is frequently the case.
-In addition, peptides are stronger affected by missing values, reducing the number of contrasts we could estimate for the dataset using this method (see Figure \@ref(fig:FDRfdp) Panel C).
+In addition, peptides are stronger affected by missing values, reducing the number of contrasts we could estimate for the dataset using this method (see Figure \@ref(fig:figure3) Panel C).
 
@@
 Despite imputation, `r BiocStyle::Biocpkg("MSstats")` estimates group differences for less proteins and does not achieve a higher $pAUC_{10}$.
-Furthermore, Figure \@ref(fig:FDRfdp) Panel C shows that when using `r BiocStyle::Biocpkg("MSstats")` the proportion of false discoveries might be very high even when filtering for a low $FDR$.
+Furthermore, Figure \@ref(fig:figure3) Panel C shows that when using `r BiocStyle::Biocpkg("MSstats")` the proportion of false discoveries might be very high even when filtering for a low $FDR$.
 
-We obtained difference estimates for all proteins, as shown in Figure \@ref(fig:FDRfdp) Panel A when using the probabilistic dropout model (`r BiocStyle::Biocpkg("proDA")`), the hurdle model (`r BiocStyle::Biocpkg("msqrob2")`), and when combining the results of the linear model with contrast estimates in the presence of missing data (prolfqua_merged).
-Furthermore, the performance of the scaled $p$-values or the $t$-statistics is comparable with that of the other methods (Figure \@ref(fig:FDRfdp) Panel B).
+We obtained difference estimates for all proteins, as shown in Figure \@ref(fig:figure3) Panel A when using the probabilistic dropout model (`r BiocStyle::Biocpkg("proDA")`), the hurdle model (`r BiocStyle::Biocpkg("msqrob2")`), and when combining the results of the linear model with contrast estimates in the presence of missing data (prolfqua_merged).
+Furthermore, the performance of the scaled $p$-values or the $t$-statistics is comparable with that of the other methods (Figure \@ref(fig:figure3) Panel B).
 All three models perform similarly well when benchmarked against the IonStar/MaxQuant, IonStar/FragPipe, or CPTAC dataset (see Supplement, Benchmark CPTAC).
@@
 Furthermore, for the IonStar/MaxQuant dataset, msqrob2 performed best, proDA performed best for the CPTAC/MaxQuant data, and prolfqua_merged for the IonStar/FragPipe dataset.
-Also, the FDR estimates (Figure \@ref(fig:FDRfdp) Panel C) are comparable for all three methods.
+Also, the FDR estimates (Figure \@ref(fig:figure3) Panel C) are comparable for all three methods.
 

### 2022-10-21 inst/MyArticle/paper.Rmd

--- 2022-10-21 inst/MyArticle/paper.Rmd
@@
       keep_tex: true
-      #citation_package: natbib
+      citation_package: natbib
 date: "`r format(Sys.time())`"

### 2022-10-21 inst/MyArticle/paper.Rmd
cod snippet cosmetics

@@
 mC <- prolfqua::linfct_matrix_contrasts(linfct, Contrasts)
-
+## see Table 3 for content of mC
 ```
@@
 startdata <- dplyr::inner_join(annot, startdata, by = "raw.file")
+
 ## create MaxQuant configuration
 config <- prolfqua::create_config_MQ_peptide()
+
 ## specify explanatory variable
 config$table$factors[['dilution.']] = "sample"
+
 ## create R6 object
 lfqpep <- prolfqua::LFQData$new(startdata, config, setup = TRUE)
+
 ## remove observation with 0 intensity and filter for 2 peptides per protein
 lfqpep$remove_small_intensities()$filter_proteins_by_peptide_count()
+
 ## transform intensities
@@
 lfqpro$rename_response("log_protein_abundance")
+
 ## plot panels A-D
@@
 ggpubr::ggarrange(panelA, panelB, panelC, panelD)
-
 ```
@@
 ```{r figure2, fig.cap="(ref:figure2)", echo=TRUE, out.width = '90%', fig.width = 8, fig.height = 5}
-# specify differences among groups
+## specify differences among groups
 contrasts <- c(
@@
 contr <- prolfqua::Contrasts$new(models, contrasts, modelName = "lm")
-# compute contrasts using imputation imputation
+
+## compute contrasts using imputation imputation
 conI <- prolfqua::ContrastsSimpleImpute$new( lfqpro, contrasts, modelName = "imp")
-# merge contrasts, to obtain differences for all proteins
+
+## merge contrasts, to obtain differences for all proteins
 contrasts <- prolfqua::merge_contrasts_results(prefer = contr, add = conI)
-# moderate t-statistics and p-values
+
+## moderate t-statistics and p-values
 prolfqua::ContrastsModerated$undebug("get_contrasts")
 contrasts <-  contrasts$merged |> prolfqua::ContrastsModerated$new()
+
 ## visualize results in panels

### 2022-10-22 inst/MyArticle/paper.Rmd
Once more review paper and fix bibliography

--- 2022-10-21 inst/MyArticle/paper.Rmd
@@
 Furthermore, the package implements benchmark functionality that can help to compare data acquisition, data preprocessing, or data modeling methods using a gold standard dataset.
-The application programmer interface of `r BiocStyle::Githubpkg("fgcz/prolfqua")`  strives to be clear,  predictable, discoverable, and consistent to make proteomics data analysis easy and exciting.
+The application programmer interface of `r BiocStyle::Githubpkg("fgcz/prolfqua")`  strives to be clear,  predictable, discoverable, and consistent to make proteomics data analysis application development easy and exciting.
 Finally, the `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package is available on GitHub https://github.com/fgcz/prolfqua, distributed under the MIT license.
@@
 
-By plotting the $TPR$ versus the $FPR$ we obtain the receiver operator characteristic curve (ROC curve) [@pROC_2011].
+By plotting the $TPR$ versus the $FPR$ we obtain the receiver operator characteristic curve (ROC curve) [@pROC2011].
 The area under the curve ($AUC$) or partial areas under the curve ($pAUC$), at various values of the $FPR$, are measures of performance derived from the ROC curve.
@@
 colnames(xM) <- c("prolfqua functions", "model")
-knitr::kable(tibble::as_tibble(xM), caption = "Prolfqua functions that can be used to fit various models.",
+knitr::kable(tibble::as_tibble(xM), caption = "prolfqua functions that can be used to fit various models.",
              format = "latex", booktabs = TRUE) |>
@@
 
-Since only technical replicates are available for each dilution, essential sources of variation present in every experiment, such as the biochemical and biological, are not measured.
-Therefore, the IonStar dataset (see Methods) captures only the variance from the chromatography, electro-spray, and mass spectrometric measurements.
-Furthermore, this benchmark with a parallel group design does not allow for benchmarking models with interactions.
-Unfortunately,  other benchmark datasets have similar constraints [@edwards2015cptac, @o2018proteome].
-At the same time, we could not analyze the benchmark dataset [@wessels2012comprehensive] with a full factorial design with an up-to-date version of MaxQuant.
-Thus, while we can extrapolate some of the results obtained to more realistic datasets, we need to be careful not to over-interpret.
-Specifically, the observed variances will be higher in more realistic experiments, and the power will be lower for the same sample sizes.
-Furthermore, the proportion of missing observations in real-life datasets might also be higher.
+The IonStar dataset (see Methods) captures only the variance from the chromatography, electro-spray, and mass spectrometric measurements since only technical replicates are available for each dilution.
+Therefore essential sources of variation typically present in other experiments, such as the biochemical and biological, are not measured.
+Furthermore, this dataset with a parallel group design does not allow for benchmarking models with interactions.
+However, other benchmark datasets have similar constraints [@edwards2015cptac, @o2018proteome].
+Furthermore, we could not analyze the benchmark dataset [@wessels2012comprehensive] with a complete factorial design with an up-to-date version of MaxQuant.
+Thus, while we can extrapolate some of the results obtained to more realistic datasets, we need to be careful not to over-interpret our findings.
+Specifically, the observed variances will be higher in datasets with biological replicates, and the power will be lower for the same number of samples.
+Furthermore, the proportion of missing observations in real-life datasets might be higher or distributed differently in groups.
 
@@
 
-The R-package `r BiocStyle::Biocpkg("proDA")`, `r BiocStyle::Biocpkg("msqrob2")` and `r BiocStyle::Biocpkg("prolfqua")` (see Methods) model the missing data directly, while MSstats imputes the data using an accelerated failure model.
+The R-package `r BiocStyle::Biocpkg("proDA")`, `r BiocStyle::Biocpkg("msqrob2")` and `r BiocStyle::Biocpkg("prolfqua")` do not impute missing data but integrate them into the statistical model, while MSstats filters and imputes the data using an accelerated failure model.
 Despite imputation, `r BiocStyle::Biocpkg("MSstats")` estimates group differences for less proteins and does not achieve a higher $pAUC_{10}$.
@@
 
-We obtained difference estimates for all proteins, as shown in Figure \@ref(fig:figure3) Panel A when using the probabilistic dropout model (`r BiocStyle::Biocpkg("proDA")`), the hurdle model (`r BiocStyle::Biocpkg("msqrob2")`), and when combining the results of the linear model with contrast estimates in the presence of missing data (prolfqua_merged).
-Furthermore, the performance of the scaled $p$-values or the $t$-statistics is comparable with that of the other methods (Figure \@ref(fig:figure3) Panel B).
-All three models perform similarly well when benchmarked against the IonStar/MaxQuant, IonStar/FragPipe, or CPTAC dataset (see Supplement, Benchmark CPTAC).
-We tested if the $pAUC_{10}$ differences for the three methods were statistically significant, and none of them were at a significance level of $0.4$ (see Supplement, Testing ROC curves).
-Furthermore, for the IonStar/MaxQuant dataset, msqrob2 performed best, proDA performed best for the CPTAC/MaxQuant data, and prolfqua_merged for the IonStar/FragPipe dataset.
+We obtained difference estimates for all proteins, as shown in Figure \@ref(fig:figure3) Panel A when using: a) the probabilistic dropout model (`r BiocStyle::Biocpkg("proDA")`), b) the hurdle model (`r BiocStyle::Biocpkg("msqrob2")`), and c) when combining the results of the linear model with contrast estimates in the presence of missing data (prolfqua_merged).
+Furthermore, the performance of the scaled $p$-values or the $t$-statistics is comparable among these three methods (Figure \@ref(fig:figure3) Panel B).
+We tested if there is no difference between the $pAUC_{10}$ for all three methods and did not reject this hypothesis at a significance level of $0.4$ (see Supplement, Testing ROC curves).
+Furthermore, all three models perform similarly well when examined using the IonStar/MaxQuant, IonStar/FragPipe, or CPTAC dataset (see Supplement, Benchmark CPTAC, and Benchmark FragPipe).
+While, for the IonStar/MaxQuant dataset, msqrob2 performed best, proDA performed best for the CPTAC/MaxQuant data, and prolfqua_merged for the IonStar/FragPipe dataset.
 Also, the FDR estimates (Figure \@ref(fig:figure3) Panel C) are comparable for all three methods.
 
-Since, the API of `r BiocStyle::Biocpkg("proDA")` and the produced outputs are similar to those of prolfqua, we are planning to integrate the `r BiocStyle::Biocpkg("proDA")` package as an additional modeling option into `r BiocStyle::Githubpkg("fgcz/prolfqua")` (see Figure \@ref(fig:ContrastUML)), to make it easy for application developers to use both.
+Since, the API of `r BiocStyle::Biocpkg("proDA")` and the produced outputs are similar to those of prolfqua, we are integrating the `r BiocStyle::Biocpkg("proDA")` package as an additional modeling option into `r BiocStyle::Githubpkg("fgcz/prolfqua")` (see Figure \@ref(fig:ContrastUML)).
 
 We focused our benchmark on comparing the statistical modeling methods using three different scores while we fixed the pre-processing steps.
-However, there are other equally or even more important parameters of a protein quantification pipeline [@frohlich2022benchmarking].
+However, some of these steps are of utmost significance when performing a differential expression analysis [@frohlich2022benchmarking].
 One of them is the normalization of the intensities within the samples to remove systematic differences [@pursiheimo2015optimization].
-The method used to infer protein intensities from peptide intensities is an additional important factor [@Grossmann2010].
+The method used to infer proteins from peptide identifications [@nesvizhskii2005interpretation] and protein abundances from peptide abundances is an additional important factor [@Grossmann2010].
 For instance, the original `r BiocStyle::Biocpkg("proDA")` publication uses MaxLFQ [@cox2014accurate] protein estimates.
-However, when using MaxLFQ intensities reported by MaxQuant, the $pAUC_{10}$ is significantly lower ($pAUC_{10}(\textrm{t-statistics}) = 66%$) compared with results obtained when protein abundances are estimated from peptide abundances using Tukey's median polish ($pAUC_{10}(\textrm{t-statistics}) = 75%$).
-Last but not least, the software [@Cox2008MaxQuant; @yu2020fast], identifying proteins and generating the quantification values, significantly contributes to the entire pipeline's performance, altering the number of identified proteins and the sensitivity and specificity of the differential expression analysis.
-In the supplement we show the benchmarking results for `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Githubpkg("fgcz/prolfqua")` when the IonStar dataset is prepossessed with the MSFragger/IonQuant [@yu2020fast] quantification pipeline (v15).
+However, when using MaxLFQ intensities reported by MaxQuant, the $pAUC_{10}$ is lower ($pAUC_{10}(\textrm{t-statistics}) = 66%$) compared with results obtained when protein abundances are estimated from peptides abundances using Tukey's median polish ($pAUC_{10}(\textrm{t-statistics}) = 72%$).
+Last but not least, the software [@Cox2008MaxQuant; @yu2020fast], used to identify and quantify proteins, significantly contributes to the entire pipeline's performance, altering the number of identified proteins and the sensitivity and specificity of the differential expression analysis.
+In the supplement we show the benchmarking results for `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Githubpkg("fgcz/prolfqua")` when the IonStar dataset is prepossessed with the FragPipe v15 [@yu2020fast] quantification pipeline.
 Although the number of proteins identified with two peptides is slightly lower than with MaxQuant (3929 vs.
-4046), the $pAUC_{10}$ is significantly higher ($81%$ vs.
+4046), the pAUC_{10} is significantly higher ($81%$ vs.
 $72%$).
@@
 `r BiocStyle::Githubpkg("fgcz/prolfqua")` is a feature-rich, object-oriented R package to analyze quantitative mass spectrometric data with simple or complex experimental designs.
-While other R packages for differential expression analysis of proteins typically only implement one modeling approach, `r BiocStyle::Githubpkg("fgcz/prolfqua")` supports various types of models (see Figure 2 and Table 3).
+While other R packages for differential expression analysis of proteins typically only implement one modeling approach, `r BiocStyle::Githubpkg("fgcz/prolfqua")` supports various types of models (see Figure \@ref(fig:ContrastUML) and Table \@ref(tab:prolfquaModels)).
 Furthermore, the contrast specification is explicit and consistent for all models and allows for testing interactions.
-Furthermore, the modular design of `r BiocStyle::Githubpkg("fgcz/prolfqua")` enables adding new features, e.g., generalized linear models to model the presence or absence information of a protein, or robust linear models, in the future.
-Furthermore, the developed framework is flexible enough to integrate other modeling methods, e.g., the probabilistic dropout model [@bioxrvproDA2020] or accurate variance estimation [@zhu2020deqms].
+The modular design of `r BiocStyle::Githubpkg("fgcz/prolfqua")` enables adding new features, e.g., generalized linear models to model the presence or absence information of a protein, or robust linear models, in the future.
+Furthermore, the developed framework can integrate other modeling methods, e.g., the probabilistic dropout model [@bioxrvproDA2020] or accurate variance estimation [@zhu2020deqms].
 Hence, `r BiocStyle::Githubpkg("fgcz/prolfqua")` enables the implementation of applications where the user can select an alternative normalization method, protein abundance estimation method, or differential expression analysis algorithm.
-Furthermore, with minimal adaptations, this R package can analyze different quantitative proteomics data (e.g., labeling-based TMT-, PRM- DIA-data).
+Furthermore, this R package can analyze other types of quantitative proteomics data,e.g., labeling-based TMT-, PRM- DIA-data.
 

### 2022-10-22 inst/MyArticle/prolfqua_supplement.Rmd
add supplement.

@@
 author: 
-  - Witold E.
-Wolski^[Functional Genomics Center Zurich (FGCZ), ETH Zurich / University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland] ^[Swiss Institute of Bioinformatics (SIB), Quartier Sorge - Batiment Amphipole, 1015 Lausanne, Switzerland]
-  - Paolo Nanni$^*$
-  - Jonas Grossmann$^*$$^\dag$
-  - Maria d'Errico$^*$$^\dag$
-  - Ralph Schlapbach$^*$
-  - Christian Panse$^*$$^\dag$
+  - name: Witold E. Wolski
+    affiliation:
+      - Functional Genomics Center Zurich (FGCZ), ETH Zurich / University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland
+      - Swiss Institute of Bioinformatics (SIB), Quartier Sorge - Batiment Amphipole, 1015 Lausanne, Switzerland
+    email: wew@fgcz.ethz.ethz.ch
+  - name: Paolo Nanni
+    affiliation: Functional Genomics Center Zurich (FGCZ), ETH Zurich / University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland
+  - name: Jonas Grossmann
+    affiliation:
+      - Functional Genomics Center Zurich (FGCZ), ETH Zurich / University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland
+      - Swiss Institute of Bioinformatics (SIB), Quartier Sorge - Batiment Amphipole, 1015 Lausanne, Switzerland
+  - name: Maria d'Errico
+    affiliation:
+      - Functional Genomics Center Zurich (FGCZ), ETH Zurich / University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland
+      - Swiss Institute of Bioinformatics (SIB), Quartier Sorge - Batiment Amphipole, 1015 Lausanne, Switzerland
+  - name: Ralph Schlapbach
+    affiliation: Functional Genomics Center Zurich (FGCZ), ETH Zurich / University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland
+  - name: Christian Panse
+    affiliation:
+      - Functional Genomics Center Zurich (FGCZ), ETH Zurich / University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland
+      - Swiss Institute of Bioinformatics (SIB), Quartier Sorge - Batiment Amphipole, 1015 Lausanne, Switzerland
+package: prolfqua
 output:
-  pdf_document:
-    keep_tex: yes
-    latex_engine: pdflatex
+  BiocStyle::pdf_document
 abstract: |
-  This document contains the supplements for the prolfqua manuscript available through \url{https://doi.org/10.1101/2022.06.07.494524 }.
-urlcolor: blue
-toc: true
-header-includes:
-  - \usepackage{pdfpages}
+  This document contains the supplements for the prolfqua manuscript available
+  through \url{https://doi.org/10.1101/2022.06.07.494524}.
+vignette: |
+  %\VignetteIndexEntry{Vignette Title}
+  %\VignetteEncoding{UTF-8}
+  %\VignetteEngine{knitr::rmarkdown}
+editor_options: 
+  chunk_output_type: console
+#bibliography: prolfqua.bib
 ---
+
 \renewcommand{\thepage}{S--\arabic{page}}
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-knitr::opts_chunk$set(error = TRUE)
+knitr::opts_chunk$set(echo = FALSE, error = FALSE, warning = FALSE, message = FALSE)
+```
+
+# prolfqua Installation
+
+To install the `r BiocStyle::Githubpkg("fgcz/prolfqua")` package ([ACS_JPR](https://github.com/fgcz/prolfqua/releases/tag/v0.4.3) release), including all R package dependencies, execute in the R console:
+
+```{r installation, echo=TRUE, eval=FALSE}
+remotes::install_github("https://github.com/fgcz/prolfqua/releases/tag/v0.4.3",
+    dependencies = TRUE)
+```
+
+# Benchmarking msqrob2, proDA and prolfqua using the CPTAC dataset
+
+The Rmarkdown file with R code to execute the benchmarks is  [Benchmarking of msqrob2, prolfqua and proDA using the CPTAC dataset MaxQuant output](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_cptac.html)
+
+# Benchmark Vignettes
+
+The R-markdown files with all the R code to run the benchmark for `r BiocStyle::Githubpkg("prolfqua")`, `r BiocStyle::Biocpkg("msqrob2")`, `BiocStyle::Biocpkg("MSStats")` and `BiocStyle::Biocpkg("proDA")` are available in the R package `r BiocStyle::Githubpkg('wolski/prolfquabenchmark')`.
+
+To build the vignettes with the benchmarks execute the following commands in the R console:
+
+```{r installbenchmark, echo=TRUE, eval=FALSE}
+install.packages('remotes')
+remotes::install_gitlab("wolski/prolfquadata", host="gitlab.bfabric.org") # installs the data
+remotes::install_github("wolski/prolfquabenchmark", build_vignettes = TRUE, dependencies = TRUE)
+```
+
+Pre-build version of the vignettes are available:
+
+- [Benchmarking of methods implemented in prolfqua using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkingIonstarData.html) 
+- [Benchmarking of MSstats using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_MSStats.html) 
+- [Benchmarking of proDA  using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_proDA_medpolish.html) 
+- [Benchmarking of msqrob2  using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkMSqRob2.html)  
+- [Benchmarking of proDA and proflqua using IonStar dataset MSFragger IonQuant combined_protein.txt](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkMSFraggerProteinIonStar.html)
+
+# Creating a prolfqua configuration
+
+The following code demonstrates how we use `r BiocStyle::Githubpkg("fgcz/prolfqua")` to analyze protein intensities reported in the MSFragger `combined_protein.tsv` file.
+First, we create a tidy table containing the protein abundances by reading the `combined_protein.tsv` file using  `tidy_MSFragger_combined_protein.` Then, we read the sample annotation from the file `annotation.xlsx` file.
+Next, we create an `AnalysisTableAnnotation` R6 object.
+Bottom-up proteomics data is hierarchical, i.e., a protein has peptides, peptides might be modified, etc.
+Therefore, the `AnalysisTableAnnotation` has a `hierarchy` field storing a list with an entry for each hierarchy level.
+Since `combined_portein.tsv` only holds protein level data, the hierarchy list has one element, and we use it to specify which column contains the protein identifiers.
+We also need to define which column contains the protein abundances we want to use for the data analysis.
+Finally, we have to specify which columns contain the explanatory variables of the analysis.
+The `AnalysisTableAnnotation` has the field `factors,` a list with as many entries as explanatory variables.
+Here we include two explanatory variables, the dilution, specified in the column 'sample', and 'run' stored in the column 'run_ID', representing the order of the measurement.
+
+```{r echo=TRUE}
+datadir <- file.path(find.package("prolfquadata") , "quantdata")
+inputFragfile <-  file.path(datadir, "MSFragger_IonStar2018_PXD003881.zip")
+inputAnnotation <- file.path(datadir, "annotation_Ionstar2018_PXD003881.xlsx")
+# read input annotation
+annotation <- readxl::read_xlsx(inputAnnotation)
+
+protein <- tibble::as_tibble(
+    read.csv(unz(inputFragfile,"IonstarWithMSFragger/combined_protein.tsv"),
+             header = TRUE, sep = "\t", stringsAsFactors = FALSE))
+
+# read combined_protein.tsv 
+protein <- prolfqua::tidy_FragPipe_combined_protein_deprec(protein)
+# remove proteins identified by a single peptide
+protein <- protein |> dplyr::filter(unique.stripped.peptides > 1)
+
+# annotate the data
+merged <- dplyr::inner_join(annotation, protein)
+atable <- prolfqua::AnalysisTableAnnotation$new()
+atable$fileName = "raw.file"
+# specify column containing protein identifiers
+atable$hierarchy[["protein_Id"]] = "protein"
+
+# column with protein abundances
+atable$set_response("total.intensity")
+
+# the factors of the analysis
+atable$factors[["dilution."]] = "sample"
+atable$factors[["run"]] = "run_ID"
+
+config <- prolfqua::AnalysisConfiguration$new(atable)
+
+adata <- prolfqua::setup_analysis(merged, config)
+lfqdata <- prolfqua::LFQData$new(adata, config)
+# show number of proteins in the dataset
+lfqdata$hierarchy_counts()
 
@@
 
-# This
-# And that
-# And whatsoever
+# Miscellaneous
 
-# R Session Information
+(ref:tabCompletion) The screenshot displays the command-line completion (tab completion) of RStudio on the `prolfqua::LFQData` R6 object. In the example, it shows the getter methods of the object.
 
-```{r sessioninfo, echo=FALSE}
+```{r tabCompletion, echo=FALSE, fig.cap="(ref:tabCompletion)", out.width = '66%'}
+knitr::include_graphics("graphics/codeSnippet1TabCompletion.png")
+```
+
+```{r proLFQuaSticker, echo=TRUE, out.height="5cm", eval=TRUE, fig.cap="Sticker maintainer: Witold E.
+Wolski; License: Creative Commons Attribution CC-BY.
+Feel free to share and adapt, but don't forget to credit the author."}
+file.path("graphics/hexStickerProlfqua.png") |>
+  knitr::include_graphics()
+```
+
+# Session info {.unnumbered}
+
+```{r sessionInfo, echo=FALSE}
 sessionInfo()

### 2022-10-22 inst/MyArticle/paper.Rmd
updated supplement.

@@
 The area under the curve ($AUC$) or partial areas under the curve ($pAUC$), at various values of the $FPR$, are measures of performance derived from the ROC curve.
-Using these measures, we can compare the performances of the statistics or the various models examined and test if these differences are statistically significant.
+Using these measures, we can compare the performances of the statistics for a single model or the various models and test if the differences are statistically significant, using a test to compare ROC curves.
 
@@
 knitr::asis_output(stringr::str_replace(tmp, "HACKIT", "$^{a}$"))
+saveRDS(allBenchmarks, "allBenchmarks.RDS")
 
@@
 ```{r sessioninfo, echo=FALSE}
-sessionInfo()
+pander::pander(sessionInfo())
 ```

### 2022-10-22 inst/MyArticle/prolfqua_supplement.Rmd
updated supplement.

@@
 ```{r installation, echo=TRUE, eval=FALSE}
-remotes::install_github("https://github.com/fgcz/prolfqua/releases/tag/v0.4.3",
-    dependencies = TRUE)
+remotes::install_github("https://github.com/fgcz/prolfqua/releases/tag/v0.4.5",
+    dependencies = TRUE, build_vignettes=TRUE)
+```
+
+The R-markdown files with all the R code to run the benchmark for `r BiocStyle::Githubpkg("prolfqua")`, `r BiocStyle::Biocpkg("msqrob2")`, `BiocStyle::Biocpkg("MSStats")` and `BiocStyle::Biocpkg("proDA")` are available in the R package `r BiocStyle::Githubpkg('wolski/prolfquabenchmark')`.
+
+```{r installation2, echo=TRUE, eval=FALSE}
+remotes::install_gitlab("wolski/prolfquadata", host="gitlab.bfabric.org")
+remotes::install_github("https://github.com/wolski/prolfquabenchmark",
+    dependencies = TRUE, build_vignettes=TRUE)
+```
+
+# Comparing partial areas under the curve at 0.1 FDR for Ionstar/MaxQuant dataset
+
+Here we test if we should reject the null hypothesis that the area under the ROC curve at 0.1 FDR does not differ for 
+msqrob2, proDA, and prolfqua when computed based on scaled p.values. To this task we use the function roc.test from the R package `r BiocStyle::CRANpkg("pROC")`.
+
+```{r roctest, include = FALSE, eval = TRUE}
+allBenchmarks <- readRDS("allBenchmarks.RDS")
+ttmsqrob <- allBenchmarks$benchmark_msqrob$data()
+ddmsqrob <- pROC::roc(ttmsqrob$TP, ttmsqrob$scaled.p.value, partial.auc = c(1, 0.9))
+
+ttprolfqua <- allBenchmarks$benchmark_merged$data()
+ddprolfqua <- pROC::roc(ttprolfqua$TP, ttprolfqua$scaled.p.value, partial.auc = c(1, 0.9))
+ttproda <- allBenchmarks$benchmark_proDA$data()
+ddproda <- pROC::roc(ttproda$TP, ttproda$scaled.p.value, partial.auc = c(1, 0.9))
+
+tmp <- c(msqrob2_vs_prolfqua = pROC::roc.test(ddmsqrob,ddprolfqua, progress = "none")$p.value,
+         msqrob2_vs_proda = pROC::roc.test(ddmsqrob,ddproda, progress = "none")$p.value,
+         prolfqua_vs_proda = pROC::roc.test(ddprolfqua,ddproda, progress = "none")$p.value)
+
+```
+
+The Table \@ref(tab:shwoROCtestREsultsIonStarMQ) shows the p-values of the bootstrap test for two ROC curves.
+We observe that for this benchmark data, there are no significant differences among the $AUC_{10}$ for the three packages.
+
+```{r shwoROCtestREsultsIonStarMQ}
+cap = "p-values for pairwise comparsions of partial AUC at 0.1 for msqrob2, proDA and prolfqua for IonStar MaxQuant data."
+knitr::kable(tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)), format = "latex", caption = cap)
+
 ```
@@
 
-The Rmarkdown file with R code to execute the benchmarks is  [Benchmarking of msqrob2, prolfqua and proDA using the CPTAC dataset MaxQuant output](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_cptac.html)
+The Rmarkdown file which created these results, with more details about the CPTAC / MaxQuant dataset, can be found here:
+["Benchmarking of msqrob2, prolfqua and proDA using the CPTAC dataset MaxQuant output"](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_cptac.html)
+
+```{r}
+getpath <- function(filN){
+    f1 <- paste0("../../inst/Benchresults/",filN)
+    if (f1 == "") {
+        f1 <- system.file("Benchresults",filN, package = "prolfquabenchmark")
+    }
+    message(f1)
+    return(f1)
+}
+CPTACBenchmarks <- readRDS(getpath("CPTAC_Benchmark.RDS"))
+mrob <- CPTACBenchmarks$benchmark_msqrob$pAUC_summaries()$ftable$content
+mrob$package <- "msqrob2"
+proda <- CPTACBenchmarks$benchmark_proDA$pAUC_summaries()$ftable$content
+proda$package <- "proDA"
+prolfqua <- CPTACBenchmarks$benchmark_merged$pAUC_summaries()$ftable$content
+prolfqua$package <- "prolfqua"
+
+all <- dplyr::bind_rows(list(mrob, proda, prolfqua))
+all <- all |> filter(contrast == "b_vs_a")
+all$what[all$what == "statistic"] <- "t_statistic"
+all$what[all$what == "t"] <- "t_statistic"
+all$what[all$what == "scaled.pval"] <- "scaled.p.value"
+all$what[all$what == "logFC"] <- "diff"
+
+```
+
+Figure \@ref(fig:plotCPTACbarplot) shows the partial area under the curve for `r BiocStyle::Biocpkg("msqrob2")`, `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Githubpkg("fgcz/prolfqua")`.
+
+```{r plotCPTACbarplot, fig.cap= "Barplot showing the partial area under the ROC at 0.1 FDR."}
+all |> ggplot(aes(x = what, y = pAUC_10, group = package, color = NULL, fill = package)) +
+  geom_bar(stat = "identity",  position = position_dodge())
+
+```
+
+```{r roctestCPTAC, include = TRUE, message=FALSE, error=FALSE}
+
+ttmsqrob <- CPTACBenchmarks$benchmark_msqrob$data()
+ddmsqrob <- pROC::roc(ttmsqrob$TP, ttmsqrob$scaled.pval, partial.auc = c(1, 0.9))
+ttprolfqua <- CPTACBenchmarks$benchmark_merged$data()
+ddprolfqua <- pROC::roc(ttprolfqua$TP, ttprolfqua$scaled.p.value, partial.auc = c(1, 0.9))
+ttproda <- CPTACBenchmarks$benchmark_proDA$data()
+ddproda <- pROC::roc(ttproda$TP, ttproda$scaled.pval, partial.auc = c(1, 0.9))
+
+tmp <- c(msqrob2_vs_prolfqua = pROC::roc.test(ddmsqrob,ddprolfqua, progress = "none")$p.value,
+         msqrob2_vs_proda = pROC::roc.test(ddmsqrob,ddproda, progress = "none")$p.value,
+         prolfqua_vs_proda = pROC::roc.test(ddprolfqua,ddproda, progress = "none")$p.value)
+
+```
+
+Table \@ref(tab:shwoROCtestREsultsCPTAC) shows the results of the Bootstrap test for two ROC curves. 
+We observe that for this benchmark data, there are no significant differences among the $AUC_{10}$ for the three packages.
+
+```{r shwoROCtestREsultsCPTAC}
+cap = "p-values for pairwise comparsions of partial AUC for msqrob2, proDA and prolfqua CPTAC/MaxQuant data."
+knitr::kable(tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)), format = "latex", caption = cap)
+
+```
+
+# Benchmark IonQuant FragPipe v15 combined_protein.tsv
+
+The Rmarkdown file which created these results, with more details about the IonQuant FragPipe v15 dataset, can be found here:
+["Benchmarking proDA and prolfqua using Ionstar Dataset and FragPipe v15 compbined_protein output"](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_cptac.html)
+
+Direct comparison with `msqrob2` is impossible since, to fit the dropout model, the peptide intensities are required, while here, we are starting the analysis from the `combined_proteins.tsv` file.
+
+```{r FragPipe}
+FragPipeBenchmarks <- readRDS(getpath("FragPipev15_comb_prot.RDS"))
+proda <- FragPipeBenchmarks$benchmark_proDA$pAUC_summaries()$ftable$content
+proda$package <- "proDA"
+prolfqua <- FragPipeBenchmarks$benchmark_prolfqua$pAUC_summaries()$ftable$content
+prolfqua$package <- "prolfqua"
+
+tmp <- dplyr::bind_rows(proda, prolfqua)
+tmp$what |> unique()
+tmp$what[tmp$what == "statistic"] <- "t_statistic"
+tmp$what[tmp$what == "scaled.pval"] <- "scaled.p.value"
+tmp <- tmp |> filter(contrast == "all")
+```
+
+Figure \@ref(fig:plotCPTACbarplot) shows the partial area under the curve for `r BiocStyle::Biocpkg("msqrob2")`, `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Githubpkg("fgcz/prolfqua")`.
+
+```{r plotFragPipebarplot, fig.cap= "Barplot showing the partial area under the ROC at 0.1 FDR."}
+tmp |> ggplot2::ggplot(ggplot2::aes(x = what, y = pAUC_10, group = package, color = NULL, fill = package)) +
+  ggplot2::geom_bar(stat = "identity",  position = ggplot2::position_dodge()) +
+  ggplot2::facet_wrap(~ contrast) +
+  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust = 1))
+
+```
+
+```{r roctestFRAGPIPE, include = TRUE, message=FALSE, error=FALSE}
+
+ttprolfqua <- FragPipeBenchmarks$benchmark_prolfqua$data()
+ddprolfqua <- pROC::roc(ttprolfqua$TP, ttprolfqua$scaled.p.value, partial.auc = c(1, 0.9))
+ttproda <- FragPipeBenchmarks$benchmark_proDA$data()
+ddproda <- pROC::roc(ttproda$TP, ttproda$scaled.pval, partial.auc = c(1, 0.9))
+
+tmp <- c(prolfqua_vs_proda = pROC::roc.test(ddprolfqua,ddproda, progress = "none")$p.value)
+
+```
+
+Table \@ref(tab:shwoROCtestREsultsFRAGPIPE) shows the results of the Bootstrap test for the difference of two ROC curves.
+We observe that for this benchmark data, there are no significant differences among the $AUC_{10}$ for the two packages.
+
+```{r shwoROCtestREsultsFRAGPIPE}
+cap = "p-values for pairwise comparsions of partial AUC for msqrob2, proDA and prolfqua CPTAC/MaxQuant data."
+knitr::kable(tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)), format = "latex", caption = cap)
+
+```
 
 # Benchmark Vignettes
-
-The R-markdown files with all the R code to run the benchmark for `r BiocStyle::Githubpkg("prolfqua")`, `r BiocStyle::Biocpkg("msqrob2")`, `BiocStyle::Biocpkg("MSStats")` and `BiocStyle::Biocpkg("proDA")` are available in the R package `r BiocStyle::Githubpkg('wolski/prolfquabenchmark')`.
-
-To build the vignettes with the benchmarks execute the following commands in the R console:
-
-```{r installbenchmark, echo=TRUE, eval=FALSE}
-install.packages('remotes')
-remotes::install_gitlab("wolski/prolfquadata", host="gitlab.bfabric.org") # installs the data
-remotes::install_github("wolski/prolfquabenchmark", build_vignettes = TRUE, dependencies = TRUE)
-```
 
@@
 ```{r sessionInfo, echo=FALSE}
-sessionInfo()
+pander::pander(sessionInfo())
 ```

### 2022-10-22 inst/MyArticle/prolfqua_supplement.Rmd
cleanup

@@
 tmp <- dplyr::bind_rows(proda, prolfqua)
-tmp$what |> unique()
 tmp$what[tmp$what == "statistic"] <- "t_statistic"

### 2022-10-24 inst/MyArticle/paper.Rmd
fix ref

--- 2022-10-21 inst/MyArticle/paper.Rmd
@@
 
-R's formula interface for linear models is flexible, widely used, and well documented [@faraway2016extending,@law2020guide].
+R's formula interface for linear models is flexible, widely used, and well documented [@faraway2016extending; @law2020guide].
 We use the formula interface to specify the models, making it easy to reproduce an analysis performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` in other statistical programming languages.
@@
 
-To ease the usage barriers of the R package to users not familiar with statistics and R programming, we developed an application based on the prolfqua package into our data management platform [B-Fabric](https://fgcz-bfabric.uzh.ch/) [@bfabric,@bfabricjib2022].
+To ease the usage barriers of the R package to users not familiar with statistics and R programming, we developed an application based on the prolfqua package into our data management platform [B-Fabric](https://fgcz-bfabric.uzh.ch/) [@bfabric; @bfabricjib2022].
 The B-Fabric system runs a computing infrastructure controlled by a local resource management system that supports cloud-bursting [@VMMAD].

### 2022-10-25 inst/MyArticle/paper.Rmd

@@
 There is a large variety of signal processing software and analysis tools.
-Nevertheless, there is a need for a modular, easy-to-use application programming interface in R that transparently supports a variety of well principled statistical procedures to make applying them to proteomics data, comparing and understanding their differences easy.
+Nevertheless, there is a need for a modular, easy-to-use application programming interface in R that transparently supports a variety of well principled statistical procedures to make applying them to proteomics data, comparing and understanding their differences easily.
 The `r BiocStyle::Githubpkg("fgcz/prolfqua")` package integrates essential steps of the mass spectrometry-based differential expression analysis workflow: quality control, data normalization, protein aggregation, statistical modeling, hypothesis testing, and sample size estimation.
@@
 Several data analysis packages exist to model mass spectrometry protein quantification experiments, e.g., `r BiocStyle::Biocpkg("limma")` [@Ritchie2015], `r BiocStyle::Biocpkg("MSstats")` [@Choi2014], `r BiocStyle::Biocpkg("PECA")` [@Suomi2017bEnhanced],  `r BiocStyle::Biocpkg("msqrob2")` [@Goeminne2016] or `r BiocStyle::Biocpkg("proDA")` [@bioxrvproDA2020], to mention some, all implemented in R.
-Each of them has some unique features, for example, `r BiocStyle::Biocpkg("MSstats")` determines the statistical model from the structure of the sample annotation, allowing users with limited statistical knowledge to perform differential expression analysis (DEA).
+Each of them has some unique features, for example, `r BiocStyle::Biocpkg("MSstats")` determines the statistical model from the structure of the sample annotation, which allows users with limited statistical knowledge to perform differential expression analysis (DEA).
 At the same time, `r BiocStyle::Biocpkg("limma")` allows the specification of a design matrix using a linear model formula and implements the empirical Bayes variance shrinkage method.
-The package `r BiocStyle::Biocpkg("PECA")`, performs a roll-up of peptide level differences and peptide level $p$-value estimates, obtained from `r BiocStyle::Biocpkg("limma")` or `r BiocStyle::Biocpkg("PECA")`, to protein level estimates.
+The package `r BiocStyle::Biocpkg("PECA")` performs a roll-up of peptide level differences and peptide level $p$-value estimates obtained from `r BiocStyle::Biocpkg("limma")` or `r BiocStyle::Biocpkg("PECA")`, to protein level estimates.
 Furthermore, `r BiocStyle::Biocpkg("msqrob2")` combines two models: robust linear models fitted to protein abundances and a quasibinomial generalized linear model fitted to peptide counts, into Hurdle model.
@@
 
-Of note are the various approaches to handling missing observations, which are inherent in mass spectrometric bottom-up experiments.
+Of note are the various approaches to handle missing observations, which are inherent in mass spectrometric bottom-up experiments.
 For instance, `r BiocStyle::Biocpkg("MSstats")` handles missing data by feature filtering and imputation.
-Other means of modeling missing observations are the Hurdle model discussed by @goeminne2020msqrob, while the R package `r BiocStyle::Biocpkg("proDA")` models missingness using probabilistic dropout models [@bioxrvproDA2020].
+Other means of modeling missing observations are the Hurdle models discussed by @goeminne2020msqrob, while the R package `r BiocStyle::Biocpkg("proDA")` models missingness using probabilistic dropout models [@bioxrvproDA2020].
 
@@
 
-When developing the R package `r BiocStyle::Githubpkg("fgcz/prolfqua")` we were inspired by the R package `r BiocStyle::CRANpkg("caret")` [@caret2008], which enables us to call various machine learning (ML) methods, and makes selecting the best ML algorithm for your problem easy.
+When developing the R package `r BiocStyle::Githubpkg("fgcz/prolfqua")` we were inspired by the R package `r BiocStyle::CRANpkg("caret")` [@caret2008] which enables us to call various Machine Learning (ML) methods, and makes selecting the best ML algorithm for your problem easy.
 We aimed for a similar R package for the differential expression analysis of quantitative proteomics data.
 However, the existing packages differ widely regarding supported designs, model specifications, and output formats.
-At the same time, they have in common: they fit linear models to either peptide or protein intensities, determine differences among groups, and afterward apply empirical Bayes variance shrinkage.
+At the same time, they have the following features in common: fitting linear models to either peptide or protein intensities, determining differences among groups, and afterward applying empirical Bayes variance shrinkage.
 Therefore, the revised goal was to provide a modular object-oriented (OO) design, with R linear models as a core, and add features such as $p$-value aggregation, variance shrinkage, or modeling of missing observations.
 
-Furthermore, the functionality of `r BiocStyle::Githubpkg("fgcz/prolfqua")` also includes methods specific to proteomics data.
+Furthermore, `r BiocStyle::Githubpkg("fgcz/prolfqua")` also includes methods specific to proteomics data.
 For example, we implemented strategies to estimate protein intensities from peptide intensities: top N [@Grossmann2010], Tukey's median polish  [@tukey1977exploratory], robust linear models [@goeminne2020msqrob].
-Furthermore, peptide or protein abundances can be scaled and transformed using robust scaling, _quantile_ normalization or `r BiocStyle::Biocpkg("vsn")` to remove systematic differences among samples and heteroscedasticity.
+Peptide or protein abundances can then be scaled and transformed using robust scaling, _quantile_ normalization or `r BiocStyle::Biocpkg("vsn")` to remove systematic differences among samples and heteroscedasticity.
 In this respect, `r BiocStyle::Githubpkg("fgcz/prolfqua")` can be compared with R packages such as `r BiocStyle::Biocpkg("DEP")` [@DEP2018] or `r BiocStyle::Biocpkg("POMA")` [@POMA2021] which support the entire differential expression analysis pipeline.
 
-We also implemented functionality and use the Ionstar [@shen2018ionstar] and CPTAC (available from `r BiocStyle::Biocpkg("msdata")`) dataset to benchmark the modeling methods implemented within `r BiocStyle::Githubpkg("fgcz/prolfqua")` and to compare our results with those of `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`.
+We use the Ionstar [@shen2018ionstar] and CPTAC (available from `r BiocStyle::Biocpkg("msdata")`) dataset to benchmark the modeling methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` and to compare our results with those of `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`.
 Since group sizes are relatively small, typically with four or five subjects per group, the high power of the tests is a relevant criterion to assess the modeling method.
@@
 We store all the data needed for analysis in a data frame as tidy data, i.e., every column is a variable, every row is an observation, and every cell is a single value [@tidydata2014].
-Using an R6 [@R6cite] configuration object (Figure \@ref(fig:LFQData)), we specify what variable is in which column, making it easy to integrate new inputs in `r BiocStyle::Githubpkg("fgcz/prolfqua")` if provided as a tidy data.
+Using an R6 [@R6cite] configuration object (Figure \@ref(fig:LFQData)), we specify what variable is in which column making it easy to integrate new inputs in `r BiocStyle::Githubpkg("fgcz/prolfqua")` if provided as a tidy data.
 For example, to visualize tidy Spectronaut[@bruderer2015extending], DiaNN[@demichev2020dia], or Skyline[@maclean2010skyline] outputs, or data in `r BiocStyle::Biocpkg("MSstats")`[@Choi2014] format, only a few lines of code are needed to update the `r BiocStyle::Githubpkg("fgcz/prolfqua")` `AnalysisTableConfiguration` configuration.
-The configuration encapsulates the differences in column names among the various input formats and enables the use of the methods without additional arguments.
-We show an example code for creating an MSFragger[@yu2020fast] configuration in the Appendix.
-For popular software like MaxQuant[@Cox2008MaxQuant], or MSFragger, which stores the same variable, e.g., intensity, in multiple columns, one for each sample, we implemented methods that transform the data into tidy data.
+The configuration encapsulates the differences among the various input formats in column names and enables the use of the methods without additional arguments.
+An example code for creating an MSFragger[@yu2020fast] configuration can be found in the Appendix.
+We implemented methods that transform the data into tidy data for popular software like MaxQuant[@Cox2008MaxQuant], or MSFragger, which stores the same variable, e.g., intensity, in multiple columns, one for each sample.
 Relying on the tidy data table enables us to interface with many data manipulation, visualization, and modeling methods, implemented in base R [@RCoreTeam2021] and the tidyverse [@tidyverse2019], easily.

### 2022-10-25 inst/MyArticle/prolfqua_supplement.Rmd
updates to supplement

@@
 
-```{r roctest, include = FALSE, eval = TRUE}
+```{r roctest, include = TRUE, eval = TRUE}
 allBenchmarks <- readRDS("allBenchmarks.RDS")
@@
 
+# ROPECA does not yield p-values
+
+We show here a simulation experiment.
+For all the peptides (nrPep = 10000) $H0$ is true, that is there are no difference among groups.
+We know that in such a case the p.values distribution should be uniform.
+We assign these 10000 peptides to 800 proteins.
+Afterwards we apply the ropeca methods as discussed in the methods section of the manuscript and in the original publication [].
+
+```{r , fig.cap = "Left panel - distibution of peptide level p.values.
+Center panel - number of proteins as a function of the number of peptides per protein.
+Right panel - distribution of protein level p-values(?) obtained with the ropeca method."}
+set.seed(10)
+nrPep <- 10000
+nrProtein <- 800
+p.value <- runif(nrPep)
+estimate <- rnorm(nrPep)
+avgAbd <- runif(nrPep)
+protein_Id <- sample(1:800, size = nrPep,
+   replace = TRUE, prob = dexp(seq(0,5,length = 800)))
+
+testdata <- data.frame(contrast = "contrast1",
+   protein_Id = protein_Id,
+   estimate = estimate,
+   pseudo_estimate = estimate,
+   p.value = p.value,
+   avgAbd = avgAbd )
+
+xx30 <- summary_ROPECA_median_p.scaled(testdata,
+                                     subject_Id = "protein_Id",
+                                     estimate = "estimate",
+                                     p.value = "p.value",
+                                     max.n = 30)
+
+par(mfrow = c(1,3))
+hist(testdata$p.value, breaks = 20, xlab = "p-value", main = "")
+plot(table(table(protein_Id)), xlab = "nr of peptides/protein", ylab = "# nr of proteins", main = "")
+hist(xx30$beta.based.significance, breaks = 20, xlab = "p-value(?) obtained with Ropeca.", main = "")
+
+```
+
 # Miscellaneous

### 2022-10-25 inst/MyArticle/paper.Rmd
adressing @JG suggestions

--- 2022-10-24 inst/MyArticle/paper.Rmd
@@
 
-R's formula interface for linear models is flexible, widely used, and well documented [@faraway2016extending; @law2020guide].
+R's formula interface for linear models is flexible, widely used, and well documented [@faraway2016extending,@law2020guide].
 We use the formula interface to specify the models, making it easy to reproduce an analysis performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` in other statistical programming languages.
@@
 
-Given a linear model $y = \beta X$, we can compute the difference $\beta_{c}$ between two groups by a linear combination $c$ of linear model parameters $\beta$, where $c$ is a column vector with as many elements as there are coefficients $\beta$ in the linear model.
+Given a linear model $y = \beta X$, we can compute the difference $\beta_{c}$ between two groups by a linear combination of weights $c$ and model parameters $\beta$, where $c$ is a column vector with as many elements as there are coefficients $\beta$ in the linear model.
 If $c$ has $0$ for one or more of its rows, then the corresponding coefficient in $\beta$ is not involved in determining the contrast [@ph525xseries].
@@
 
-The linear model explains the observed protein abundances using the explanatory variables `condition_` and `batch_` and the interaction term  `condition_:batch_`.
-The treatment difference among the groups `Ethanol` and `Glucose` defined by the factor `condition`, and similarly among the groups defined by `batch_`, can be specified.
-Furthermore, the treatment differences among `Ethanol` and `Glucose` within the batches `p2370` and `p2691` can be specified.
-Finally, we can examine if the treatment differences in both batches are the same or are different.
-The code to compute the array of weights $c$ used to multiply the model coefficient $\beta$, to obtain the effect size estimates, and the variance of the estimate, is shown below:
+The following code illustrates how to specify a model with two factors and an interaction term and define the differences we want to estimate.
+The example dataset was acquired in two batches (p2370, p2691).
+In each, we measured Yeast samples grown under two different conditions (Glucose and Ethanol).
+This example is taken from the `r BiocStyle::Githubpkg("fgcz/prolfqua")` package vignette file.
+To estimate treatment differences, we first specify the linear model that explains the observed protein abundances (transformedIntensities) using the explanatory variables `condition_` and `batch_` and the interaction term  `condition_:batch_`.
+Then, after fitting the model, we can estimate the treatment difference among the groups `Ethanol` and `Glucose` defined by the factor `condition` and similarly among the groups defined by `batch_`.
+Furthermore, we can examine differences between `Ethanol` and `Glucose` within each batch.
+We also can assess if these differences are the same or are different in both batches, i.e., if there is an interaction between condition and batch.
+We also show the code to compute the array of weights $c$ used to multiply the model coefficient $\beta$.
 
 ```{r contrast2, eval = TRUE, echo = TRUE}
+
+Yeast2Factor <- prolfqua::prolfqua_data("data_Yeast2Factor")
+Yeast2Factor$data |> dplyr::select(condition_, batch_) |> dplyr::distinct()
 mlm <- prolfqua::strategy_lm(
@@
   )
-Yeast2Factor <- prolfqua::prolfqua_data("data_Yeast2Factor")
 mm <- prolfqua::build_model(Yeast2Factor$data, mlm, subject_Id = "protein_Id")
@@
 mC <- prolfqua::linfct_matrix_contrasts(linfct, Contrasts)
-## see Table 3 for content of mC
 ```
@@
 knitr::kable(mC, caption = "Weights $c$ for each of the contrasts (rows) which will be applied to the model parameter (columns).",format = "latex")
-
 ```
 
-This example is taken from the `r BiocStyle::Githubpkg("fgcz/prolfqua")` package vignette file.
 Given the weights $c$ contrasts from model parameter estimates obtained with `r BiocStyle::Biocpkg('proDA')`, or `r BiocStyle::Biocpkg('msqrob2')` can be determined, by this enabling us to call these packages using the same application programming interface (API).
@@
 In such a case, we will substitute the group mean, using the expected protein abundance $A$ at the LOD $A_{LOD}$.
-To estimate $A_{LOD}$ we are using the abundances from those groups where the protein was unobserved except for one sample. Typically there are many such cases, and hence we take the median.
+To estimate $A_{LOD}$ we are using the protein abundances of those groups where the protein was observed in only a single sample. Typically there are many such cases, and hence we take the median.
 
@@
 However, if for instance no observations are present in group $b$, we will use $\bar{b} = A_{LOD}$.
-Furthermore, if $\bar{a} < A_{LOD}$ we also set $\bar{a} = A_{LOD}$.
+Furthermore, if $\bar{a} < A_{LOD}$ we also set $\bar{a} = A_{LOD}$, or more formally:
 
@@
 $$
-
 We use the pooled variance in all groups to estimate the protein variance, assuming they are the same. The pooled variance $s_p^2$ is given by:
@@
 startdata <- dplyr::inner_join(annot, startdata, by = "raw.file")
-
 ## create MaxQuant configuration
 config <- prolfqua::create_config_MQ_peptide()
-
 ## specify explanatory variable
 config$table$factors[['dilution.']] = "sample"
-
 ## create R6 object
 lfqpep <- prolfqua::LFQData$new(startdata, config, setup = TRUE) 
-
 ## remove observation with 0 intensity and filter for 2 peptides per protein
 lfqpep$remove_small_intensities()$filter_proteins_by_peptide_count()
-
 ##  transform intensities
@@
 lfqpro$rename_response("log_protein_abundance")
-
 ## plot panels A-D
@@
 ggpubr::ggarrange(panelA, panelB, panelC, panelD)
+
 ```
@@
 ```{r figure2, fig.cap="(ref:figure2)", echo=TRUE, out.width = '90%', fig.width = 8, fig.height = 5}
-## specify differences among groups
+# specify differences among groups
 contrasts <- c(
@@
 contr <- prolfqua::Contrasts$new(models, contrasts, modelName = "lm")
-
-## compute contrasts using imputation imputation
+# compute contrasts using imputation imputation
 conI <- prolfqua::ContrastsSimpleImpute$new( lfqpro, contrasts, modelName = "imp")
-
-## merge contrasts, to obtain differences for all proteins
+# merge contrasts, to obtain differences for all proteins
 contrasts <- prolfqua::merge_contrasts_results(prefer = contr, add = conI)
-
-## moderate t-statistics and p-values
+# moderate t-statistics and p-values
 prolfqua::ContrastsModerated$undebug("get_contrasts")
 contrasts <-  contrasts$merged |> prolfqua::ContrastsModerated$new() 
-
 ## visualize results in panels
@@
 R linear model and linear mixed effect models allow modeling parallel designs, repeated measurements, factorial designs, and many more.
-Models in `r BiocStyle::Githubpkg("fgcz/prolfqua")` are specified using R’s linear model and mixed model formula interface.
+Models in `r BiocStyle::Githubpkg("fgcz/prolfqua")` are specified using R's linear model and mixed model formula interface.
 Therefore, knowledge of the R regression model infrastructure [@faraway2016extending; @MASS2002] is advantageous when using our package.
 Furthermore, this glass box approach should make it easy to reimplement an analysis performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` using base R or other programming languages by reading the analysis script.
+However, in the package documentation, we showcase how a user, without this knowledge, can analyze experiments with a parallel group design and a factorial design.
 
@@
 
-To ease the usage barriers of the R package to users not familiar with statistics and R programming, we developed an application based on the prolfqua package into our data management platform [B-Fabric](https://fgcz-bfabric.uzh.ch/) [@bfabric; @bfabricjib2022].
+To further lower the usage barriers of the R package to users not familiar with statistics and R programming, we developed an application based on the prolfqua package into our data management platform [B-Fabric](https://fgcz-bfabric.uzh.ch/) [@bfabric,@bfabricjib2022].
 The B-Fabric system runs a computing infrastructure controlled by a local resource management system that supports cloud-bursting [@VMMAD].
-This integration enables users to select the input data and basic settings in a graphical user interface.
-We determine the model formula and which comparisons to compute from the sample annotation provided in tabular form, similarly to SAINTexpress and `r BiocStyle::Biocpkg("MSstats")`.
-However, only parallel group designs with and without repeated measurements are analyzed.
+This integration enables users to select the input data and basic settings in a graphical user interface (GUI).
+
+In this GUI application, we determine the model formula and which comparisons to compute from the sample annotation provided in tabular form, similarly to SAINTexpress and `r BiocStyle::Biocpkg("MSstats")`.
+However, this application can analyze only parallel group designs with and without repeated measurements or factorial designs without interactions.
 The user receives a report including input files, the R markdown file, and R scripts necessary to replicate the analysis using their in-house R installation.
-In this way, [B-Fabric](https://fgcz-bfabric.uzh.ch/) helps scientists to meet requirements from funding agencies, journals, and academic institutions to publish data according to the FAIR (Findable, Accessible, Interoperable, and Reusable) [@FAIR] data principles.
-We are working on creating a stand-alone shiny application with the described functionality.
+In this way, `r BiocStyle::Githubpkg(`fgcz/prolfqua`)` and [B-Fabric](https://fgcz-bfabric.uzh.ch/) and helps scientists to meet requirements from funding agencies, journals, and academic institutions to publish data according to the FAIR (Findable, Accessible, Interoperable, and Reusable) [@FAIR] data principles.
+We are working on creating a shiny stand-alone application with the described functionality and making it available soon.
 

### 2022-10-25 inst/MyArticle/paper.Rmd

--- 2022-10-25 inst/MyArticle/paper.Rmd
@@
 
-When developing the R package `r BiocStyle::Githubpkg("fgcz/prolfqua")` we were inspired by the R package `r BiocStyle::CRANpkg("caret")` [@caret2008] which enables us to call various Machine Learning (ML) methods, and makes selecting the best ML algorithm for your problem easy.
+When developing the R package `r BiocStyle::Githubpkg("fgcz/prolfqua")` we were inspired by the R package `r BiocStyle::CRANpkg("caret")` [@caret2008] which enables us to call various machine learning (ML) methods, and makes selecting the best ML algorithm for your problem easy.
 We aimed for a similar R package for the differential expression analysis of quantitative proteomics data.
@@
 
-To ease the usage barriers of the R package to users not familiar with statistics and R programming, we developed an application based on the prolfqua package into our data management platform [B-Fabric](https://fgcz-bfabric.uzh.ch/) [@bfabric; @bfabricjib2022].
+To ease the usage barriers of the R package to users not familiar with statistics and R programming, we developed an application based on the prolfqua package into our data management platform [B-Fabric](https://fgcz-bfabric.uzh.ch/) [@bfabric,@bfabricjib2022].
 The B-Fabric system runs a computing infrastructure controlled by a local resource management system that supports cloud-bursting [@VMMAD].
-This integration enables users to select the input data and basic settings in a graphical user interface.
-We determine the model formula and which comparisons to compute from the sample annotation provided in tabular form, similarly to SAINTexpress and `r BiocStyle::Biocpkg("MSstats")`.
-However, only parallel group designs with and without repeated measurements are analyzed.
+This integration enables users to select the input data and basic settings in a graphical user interface (GUI).
+
+In this GUI application, we determine the model formula and which comparisons to compute from the sample annotation provided in tabular form, similarly to SAINTexpress and `r BiocStyle::Biocpkg("MSstats")`.
+However, this application can analyze only parallel group designs with and without repeated measurements or factorial designs without interactions.
 The user receives a report including input files, the R markdown file, and R scripts necessary to replicate the analysis using their in-house R installation.
-In this way, [B-Fabric](https://fgcz-bfabric.uzh.ch/) helps scientists to meet requirements from funding agencies, journals, and academic institutions to publish data according to the FAIR (Findable, Accessible, Interoperable, and Reusable) [@FAIR] data principles.
-We are working on creating a stand-alone shiny application with the described functionality.
+In this way, `r BiocStyle::Githubpkg(`fgcz/prolfqua`)` and [B-Fabric](https://fgcz-bfabric.uzh.ch/) and helps scientists to meet requirements from funding agencies, journals, and academic institutions to publish data according to the FAIR (Findable, Accessible, Interoperable, and Reusable) [@FAIR] data principles.
+We are working on creating a shiny stand-alone application with the described functionality and making it available soon.
 

### 2022-10-25 inst/MyArticle/prolfqua_supplement.Rmd
Make prolfqua supplement work
- Rmd cosmetics
- show benchmark code snippet
- refactor `getpath` helper function

@@
 
-The R-markdown files with all the R code to run the benchmark for `r BiocStyle::Githubpkg("prolfqua")`, `r BiocStyle::Biocpkg("msqrob2")`, `BiocStyle::Biocpkg("MSStats")` and `BiocStyle::Biocpkg("proDA")` are available in the R package `r BiocStyle::Githubpkg('wolski/prolfquabenchmark')`.
+The R-markdown files with all the R code to run the benchmark for `r BiocStyle::Githubpkg("prolfqua")`, `r BiocStyle::Biocpkg("msqrob2")`, `r BiocStyle::Biocpkg("MSStats")` and `r `BiocStyle::Biocpkg("proDA")` are available through the R package `r BiocStyle::Githubpkg('wolski/prolfquabenchmark')`.
 
@@
 
-Here we test if we should reject the null hypothesis that the area under the ROC curve at 0.1 FDR does not differ for 
-msqrob2, proDA, and prolfqua when computed based on scaled p.values. To this task we use the function roc.test from the R package `r BiocStyle::CRANpkg("pROC")`.
+In the code snippet below, we we test if we should reject the null hypothesis that the area under the ROC curve at 0.1 FDR does not differ for  msqrob2, proDA, and prolfqua when computed based on scaled p.values.
+To this task we use the function roc.test from the R package `r BiocStyle::CRANpkg("pROC")`.
 
-```{r roctest, include = TRUE, eval = TRUE}
-allBenchmarks <- readRDS("allBenchmarks.RDS")
+```{r roctest, include = TRUE, eval = TRUE, echo=TRUE}
+## read pre-processed benchmarks from prolfquabenchmark install
+allBenchmarks <- system.file(package = "prolfquabenchmark") |>
+  file.path('MyArticle') |> 
+  file.path('allBenchmarks.RDS') |> 
+  readRDS("allBenchmarks.RDS")
+
 ttmsqrob <- allBenchmarks$benchmark_msqrob$data()
-ddmsqrob <- pROC::roc(ttmsqrob$TP, ttmsqrob$scaled.p.value, partial.auc = c(1, 0.9))
+ddmsqrob <- pROC::roc(ttmsqrob$TP,
+                      ttmsqrob$scaled.p.value,
+                      partial.auc = c(1, 0.9))
 
 ttprolfqua <- allBenchmarks$benchmark_merged$data()
-ddprolfqua <- pROC::roc(ttprolfqua$TP, ttprolfqua$scaled.p.value, partial.auc = c(1, 0.9))
+ddprolfqua <- pROC::roc(ttprolfqua$TP,
+                        ttprolfqua$scaled.p.value,
+                        partial.auc = c(1, 0.9))
+
 ttproda <- allBenchmarks$benchmark_proDA$data()
-ddproda <- pROC::roc(ttproda$TP, ttproda$scaled.p.value, partial.auc = c(1, 0.9))
+ddproda <- pROC::roc(ttproda$TP, ttproda$scaled.p.value,
+                     partial.auc = c(1, 0.9))
 
-tmp <- c(msqrob2_vs_prolfqua = pROC::roc.test(ddmsqrob,ddprolfqua, progress = "none")$p.value,
-         msqrob2_vs_proda = pROC::roc.test(ddmsqrob,ddproda, progress = "none")$p.value,
-         prolfqua_vs_proda = pROC::roc.test(ddprolfqua,ddproda, progress = "none")$p.value)
-
+## compose table1
+table1 <- c(msqrob2_vs_prolfqua = pROC::roc.test(ddmsqrob,ddprolfqua,
+                                              progress = "none")$p.value,
+         msqrob2_vs_proda = pROC::roc.test(ddmsqrob,ddproda,
+                                           progress = "none")$p.value,
+         prolfqua_vs_proda = pROC::roc.test(ddprolfqua,ddproda,
+                                            progress = "none")$p.value)
 ```
@@
 cap = "p-values for pairwise comparsions of partial AUC at 0.1 for msqrob2, proDA and prolfqua for IonStar MaxQuant data."
-knitr::kable(tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)), format = "latex", caption = cap)
-
+knitr::kable(dplyr::tibble(pROC_test = names(table1), p.value = round(table1, digits = 2)), format = "latex", caption = cap)
 ```
@@
 ```{r}
-getpath <- function(filN){
-    f1 <- paste0("../../inst/Benchresults/",filN)
-    if (f1 == "") {
-        f1 <- system.file("Benchresults",filN, package = "prolfquabenchmark")
-    }
-    message(f1)
-    return(f1)
+getpath <- function(f){
+    fp <- system.file("Benchresults", package = "prolfquabenchmark", f)
+    stopifnot(file.exists(fp))
+    fp
 }
@@
 all$what[all$what == "logFC"] <- "diff"
-
 ```
@@
   geom_bar(stat = "identity",  position = position_dodge()) 
-
 ```
@@
 ```{r roctestCPTAC, include = TRUE, message=FALSE, error=FALSE}
+ttmsqrob <- CPTACBenchmarks$benchmark_msqrob$data()
+ddmsqrob <- pROC::roc(ttmsqrob$TP, ttmsqrob$scaled.pval,
+                      partial.auc = c(1, 0.9))
+ttprolfqua <- CPTACBenchmarks$benchmark_merged$data()
+ddprolfqua <- pROC::roc(ttprolfqua$TP, ttprolfqua$scaled.p.value,
+                        partial.auc = c(1, 0.9))
+ttproda <- CPTACBenchmarks$benchmark_proDA$data()
+ddproda <- pROC::roc(ttproda$TP, ttproda$scaled.pval,
+                     partial.auc = c(1, 0.9))
 
-ttmsqrob <- CPTACBenchmarks$benchmark_msqrob$data()
-ddmsqrob <- pROC::roc(ttmsqrob$TP, ttmsqrob$scaled.pval, partial.auc = c(1, 0.9))
-ttprolfqua <- CPTACBenchmarks$benchmark_merged$data()
-ddprolfqua <- pROC::roc(ttprolfqua$TP, ttprolfqua$scaled.p.value, partial.auc = c(1, 0.9))
-ttproda <- CPTACBenchmarks$benchmark_proDA$data()
-ddproda <- pROC::roc(ttproda$TP, ttproda$scaled.pval, partial.auc = c(1, 0.9))
-
-tmp <- c(msqrob2_vs_prolfqua = pROC::roc.test(ddmsqrob,ddprolfqua, progress = "none")$p.value,
-         msqrob2_vs_proda = pROC::roc.test(ddmsqrob,ddproda, progress = "none")$p.value,
-         prolfqua_vs_proda = pROC::roc.test(ddprolfqua,ddproda, progress = "none")$p.value)
+tmp <- c(msqrob2_vs_prolfqua = pROC::roc.test(ddmsqrob,ddprolfqua,
+                                              progress = "none")$p.value,
+         msqrob2_vs_proda = pROC::roc.test(ddmsqrob,ddproda,
+                                           progress = "none")$p.value,
+         prolfqua_vs_proda = pROC::roc.test(ddprolfqua,ddproda,
+                                            progress = "none")$p.value)
 
@@
 cap = "p-values for pairwise comparsions of partial AUC for msqrob2, proDA and prolfqua CPTAC/MaxQuant data."
-knitr::kable(tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)), format = "latex", caption = cap)
+knitr::kable(dplyr::tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)), format = "latex", caption = cap)
 

### 2022-10-25 inst/MyArticle/prolfqua_supplement.Rmd
Fix typo

@@
 
-In the code snippet below, we we test if we should reject the null hypothesis that the area under the ROC curve at 0.1 FDR does not differ for  msqrob2, proDA, and prolfqua when computed based on scaled p.values.
+In the code snippet below, we test if we should reject the null hypothesis that the area under the ROC curve at 0.1 FDR does not differ for  msqrob2, proDA, and prolfqua when computed based on scaled p.values.
 To this task we use the function roc.test from the R package `r BiocStyle::CRANpkg("pROC")`.

### 2022-10-26 inst/MyArticle/prolfqua_supplement.Rmd
explicit dplyr::filter etc. close #24

--- 2022-10-25 inst/MyArticle/prolfqua_supplement.Rmd
@@
 
+The sourcode used to produce this Document can be found in the  [prolfqua_supplement.Rmd](https://github.com/wolski/prolfquabenchmark/blob/main/inst/MyArticle/prolfqua_supplement.Rmd) file.
+
 # prolfqua Installation
@@
 
-Here we test if we should reject the null hypothesis that the area under the ROC curve at 0.1 FDR does not differ for 
-msqrob2, proDA, and prolfqua when computed based on scaled p.values. To this task we use the function roc.test from the R package `r BiocStyle::CRANpkg("pROC")`.
+We test if we should reject the null hypothesis that the area under the ROC curve at 0.1 FDR, computed based on scaled p.values, does not differ for msqrob2, proDA, and prolfqua (Figure 5 in the article).
+To this task we use the function `roc.test` from the R package `r BiocStyle::CRANpkg("pROC")`.
 
@@
 
-The Table \@ref(tab:shwoROCtestREsultsIonStarMQ) shows the p-values of the bootstrap test for two ROC curves.
-We observe that for this benchmark data, there are no significant differences among the $AUC_{10}$ for the three packages.
+The Table \@ref(tab:shwoROCtestREsultsIonStarMQ) shows the p-values of the pairwise bootstrap test for partial areas under the ROC curves.
+We will not reject the null hypothesis that there are no difference among the ROC curves.
 
@@
 cap = "p-values for pairwise comparsions of partial AUC at 0.1 for msqrob2, proDA and prolfqua for IonStar MaxQuant data."
-knitr::kable(tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)), format = "latex", caption = cap)
+knitr::kable(tibble::tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)), format = "latex", caption = cap)
 
@@
 
-The Rmarkdown file which created these results, with more details about the CPTAC / MaxQuant dataset, can be found here:
+We present here the results of benchmarking the msqrob2, proDA and prolfqua packages using the CPTAC dataset.
+The Rmarkdown file which generated these results, with more details about the CPTAC / MaxQuant dataset, can be found here:
 ["Benchmarking of msqrob2, prolfqua and proDA using the CPTAC dataset MaxQuant output"](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_cptac.html)
@@
 all <- dplyr::bind_rows(list(mrob, proda, prolfqua))
-all <- all |> filter(contrast == "b_vs_a")
+all <- all |> 
+  dplyr::filter(contrast == "b_vs_a")
 all$what[all$what == "statistic"] <- "t_statistic"
@@
 
-Figure \@ref(fig:plotCPTACbarplot) shows the partial area under the curve for `r BiocStyle::Biocpkg("msqrob2")`, `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Githubpkg("fgcz/prolfqua")`.
+Figure \@ref(fig:plotCPTACbarplot) shows the partial area under the curve for `r BiocStyle::Biocpkg("msqrob2")`, `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Githubpkg("fgcz/prolfqua")`, when using the difference among the groups, the scaled p-values or the t-statistics to compute the ROC curves.
 
 ```{r plotCPTACbarplot, fig.cap= "Barplot showing the partial area under the ROC at 0.1 FDR."}
-all |> ggplot(aes(x = what, y = pAUC_10, group = package, color = NULL, fill = package)) +
-  geom_bar(stat = "identity",  position = position_dodge())
+all |> ggplot2::ggplot(ggplot2::aes(x = what, y = pAUC_10, group = package, color = NULL, fill = package)) +
+  ggplot2::geom_bar(stat = "identity",  position = ggplot2::position_dodge())
 
@@
 Table \@ref(tab:shwoROCtestREsultsCPTAC) shows the results of the Bootstrap test for two ROC curves. 
-We observe that for this benchmark data, there are no significant differences among the $AUC_{10}$ for the three packages.
+We observe that for this benchmark data, we can not reject the null hypothesis that there are no significant differences among the $pAUC_{10}$ (ROC based on scaled p-value) for the three packages.
 
@@
 cap = "p-values for pairwise comparsions of partial AUC for msqrob2, proDA and prolfqua CPTAC/MaxQuant data."
-knitr::kable(tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)), format = "latex", caption = cap)
+knitr::kable(tibble::tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)), format = "latex", caption = cap)
 
@@
 
-# Benchmark IonQuant FragPipe v15 combined_protein.tsv
+# Benchmark based on IonStar FragPipe v15 combined_protein.tsv
 
-The Rmarkdown file which created these results, with more details about the IonQuant FragPipe v15 dataset, can be found here:
+The Rmarkdown file which created these results, with more details about the IonStar FragPipe v15 dataset, can be found here:
 ["Benchmarking proDA and prolfqua using Ionstar Dataset and FragPipe v15 compbined_protein output"](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_cptac.html)
 
-Direct comparison with `msqrob2` is impossible since, to fit the dropout model, the peptide intensities are required, while here, we are starting the analysis from the `combined_proteins.tsv` file.
+Direct comparison with `msqrob2` is impossible since to fit the hurdle model, the peptide intensities are required, while here, we are starting the analysis from the `combined_proteins.tsv` file.
+Starting from protein level intensities has the advantage that the effect size estimates obtained from the model agree with the input data, a feature frequently requested by the biologist.
 
@@
 tmp$what[tmp$what == "scaled.pval"] <- "scaled.p.value"
-tmp <- tmp |> filter(contrast == "all")
+tmp <- tmp |> 
+  dplyr::filter(contrast == "all")
 ```
 
-Figure \@ref(fig:plotCPTACbarplot) shows the partial area under the curve for `r BiocStyle::Biocpkg("msqrob2")`, `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Githubpkg("fgcz/prolfqua")`.
+Figure \@ref(fig:plotCPTACbarplot) shows the partial area under the curve for `r BiocStyle::Biocpkg("msqrob2")`, `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Githubpkg("fgcz/prolfqua")`, computed using the effect size estimate among groups (diff), the scaled p-value (scaled.p.value) and the t-statistics (t_statistics),
 
@@
 Table \@ref(tab:shwoROCtestREsultsFRAGPIPE) shows the results of the Bootstrap test for the difference of two ROC curves.
-We observe that for this benchmark data, there are no significant differences among the $AUC_{10}$ for the two packages.
+We observe that for this benchmark data, there are no significant differences among the $pAUC_{10}$ (for scaled p-value) for the two packages.
 
 ```{r shwoROCtestREsultsFRAGPIPE}
-cap = "p-values for pairwise comparsions of partial AUC for msqrob2, proDA and prolfqua CPTAC/MaxQuant data."
-knitr::kable(tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)), format = "latex", caption = cap)
+cap = "p-values for pairwise comparsions of partial AUC for proDA and prolfqua CPTAC/MaxQuant data."
+knitr::kable(tibble::tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)), format = "latex", caption = cap)
 
@@
 - [Benchmarking of proDA and proflqua using IonStar dataset MSFragger IonQuant combined_protein.txt](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkMSFraggerProteinIonStar.html)
+
+# Estimating $A_{LOD}$
+
+The main text describes how we estimate the abundance at the detection limit $A_{LOD}$. 
+Figure \@ref(fig:missignessHistogram) shows the distribution of the average abundance within a group for all proteins.
+The red density shows the distribution for those proteins without a missing value, and the olive, turquoise, and lilac density show for proteins with 1,2 and 3 missing values, respectively.
+We see that the proteins with three missing values have, on average lower abundances than those with no missing values.
+We are using the median of the Lila density (one observation in the group), as the $A_{LOD}$ estimate.
+
+```{r missignessHistogram, fig.cap= "Distribution of the average abundance within group a and c for all proteins.
+Each density are color coded depending the number of missing observations per protein in a group.", echo=FALSE}
+
+datadir <- file.path(find.package("prolfquadata") , "quantdata")
+inputMQfile <-  file.path(datadir,
+                          "MAXQuant_IonStar2018_PXD003881.zip")
+inputAnnotation <- file.path(datadir, "annotation_Ionstar2018_PXD003881.xlsx")
+mqdata <- list()
+
+mqdata$data <- prolfqua::tidyMQ_Peptides(inputMQfile)
+mqdata$config <- prolfqua::create_config_MQ_peptide()
+annotation <- readxl::read_xlsx(inputAnnotation)
+res <- prolfqua::add_annotation(
+  mqdata$data,
+  annotation,
+  fileName = "raw.file"
+)
+mqdata$config$table$factors[["dilution."]] = "sample"
+mqdata$config$table$factors[["run_Id"]] = "run_ID"
+mqdata$config$table$factorDepth <- 1
+mqdata$data <- prolfqua::setup_analysis(res, mqdata$config)
+lfqdata <- prolfqua::LFQData$new(mqdata$data, mqdata$config)
+lfqdata$data <- lfqdata$data |> 
+  dplyr::filter(dilution. == "a" | dilution. =="c")
+lfqdata$remove_small_intensities()
+bb <- lfqdata$get_Plotter()
+bb$missigness_histogram()
+```
+
+# The probabilities produced by ROPECA are not p-values
+
+In the main text discuss that the properties of the Beta distribution-based probabilities, computed from peptide level p-values, using the ROPECA method are not well understood. 
+We are running here a simulation experiment, which shows that if we start with a dataset where the null hypothesis is valid for all the peptides, i.e., the p-values are uniformly distributed, the produced Beta distribution-based probabilities are not uniformely distributed.
+
+For all the peptides (`nrPep = 10000`) $H0$ is true.
+We know that if $H0$ is true, the distribution of the $p-values$ will be uniform.
+(Figure \@ref(fig:figROPECA) left Panel).
+Therefore, we can omit the step of generating the data from $H0$ and computing the p-values but simulate the p-values directly by sampling from the uniform distribution (`runif(nrPep)`).
+
+By sampling 800 protein id's 10000 times (with replacement), we assign the peptides to proteins ('sample(1:800, size = nrPep,
+   replace = TRUE, ...)').
+To make the data more realistic, that is that some proteins have a lot of peptides, and many have just a few peptides; the density of sampling the protein Id is exponential (`dexp(seq(0,5,length = 800)`).
+Figure \@ref(fig:figROPECA) center panel shows the number of proteins as a function of the number of peptides per protein.
+Afterward, we apply the ROPECA methods as discussed in the methods section of the manuscript.
+Figure \@ref(fig:figROPECA) shows the distribution of the Beta distribution-based probabilities, which is not uniform.
+Hence, they do not have the same interpretation as p-values, i.e., the probability of falsely rejecting the $H0$ if $H0$ is true.
+
+```{r figROPECA , fig.cap = "Left panel - distibution of peptide level p.values.
+Center panel - number of proteins as a function of the number of peptides per protein.
+Right panel - distribution of protein level p-values(?) obtained with the ropeca method."}
+set.seed(10)
+nrPep <- 10000
+nrProtein <- 800
+p.value <- runif(nrPep)
+estimate <- rnorm(nrPep)
+avgAbd <- runif(nrPep)
+protein_Id <- sample(1:800, size = nrPep,
+   replace = TRUE, prob = dexp(seq(0,5,length = 800)))
+
+testdata <- data.frame(contrast = "contrast1",
+   protein_Id = protein_Id,
+   estimate = estimate,
+   pseudo_estimate = estimate,
+   p.value = p.value,
+   avgAbd = avgAbd )
+
+xx30 <- prolfqua::summary_ROPECA_median_p.scaled(testdata,
+                                     subject_Id = "protein_Id",
+                                     estimate = "estimate",
+                                     p.value = "p.value",
+                                     max.n = 30)
+
+par(mfrow = c(1,3))
+hist(testdata$p.value, breaks = 20, xlab = "p-value", main = "")
+plot(table(table(protein_Id)), xlab = "nr of peptides/protein", ylab = "# nr of proteins", main = "")
+hist(xx30$beta.based.significance, breaks = 20, xlab = "p-value(?) obtained with Ropeca.", main = "")
+
+```
 
@@
 # remove proteins identified by a single peptide
-protein <- protein |> dplyr::filter(unique.stripped.peptides > 1)
+protein <- protein |> 
+  dplyr::filter(unique.stripped.peptides > 1)
 
@@
 
-# ROPECA does not yield p-values
-
-We show here a simulation experiment.
-For all the peptides (nrPep = 10000) $H0$ is true, that is there are no difference among groups.
-We know that in such a case the p.values distribution should be uniform.
-We assign these 10000 peptides to 800 proteins.
-Afterwards we apply the ropeca methods as discussed in the methods section of the manuscript and in the original publication [].
-
-```{r , fig.cap = "Left panel - distibution of peptide level p.values.
-Center panel - number of proteins as a function of the number of peptides per protein.
-Right panel - distribution of protein level p-values(?) obtained with the ropeca method."}
-set.seed(10)
-nrPep <- 10000
-nrProtein <- 800
-p.value <- runif(nrPep)
-estimate <- rnorm(nrPep)
-avgAbd <- runif(nrPep)
-protein_Id <- sample(1:800, size = nrPep,
-   replace = TRUE, prob = dexp(seq(0,5,length = 800)))
-
-testdata <- data.frame(contrast = "contrast1",
-   protein_Id = protein_Id,
-   estimate = estimate,
-   pseudo_estimate = estimate,
-   p.value = p.value,
-   avgAbd = avgAbd )
-
-xx30 <- summary_ROPECA_median_p.scaled(testdata,
-                                     subject_Id = "protein_Id",
-                                     estimate = "estimate",
-                                     p.value = "p.value",
-                                     max.n = 30)
-
-par(mfrow = c(1,3))
-hist(testdata$p.value, breaks = 20, xlab = "p-value", main = "")
-plot(table(table(protein_Id)), xlab = "nr of peptides/protein", ylab = "# nr of proteins", main = "")
-hist(xx30$beta.based.significance, breaks = 20, xlab = "p-value(?) obtained with Ropeca.", main = "")
-
-```
-
 # Miscellaneous

### 2022-10-26 inst/MyArticle/paper.Rmd

@@
 gridExtra::grid.arrange(histpval, maplot, volcano, layout_matrix = rbind(c(1,3),c(2,3)))
+
 ```
@@
 The user receives a report including input files, the R markdown file, and R scripts necessary to replicate the analysis using their in-house R installation.
-In this way, `r BiocStyle::Githubpkg(`fgcz/prolfqua`)` and [B-Fabric](https://fgcz-bfabric.uzh.ch/) and helps scientists to meet requirements from funding agencies, journals, and academic institutions to publish data according to the FAIR (Findable, Accessible, Interoperable, and Reusable) [@FAIR] data principles.
+In this way, `r BiocStyle::Githubpkg("fgcz/prolfqua")` and [B-Fabric](https://fgcz-bfabric.uzh.ch/) and helps scientists to meet requirements from funding agencies, journals, and academic institutions to publish data according to the FAIR (Findable, Accessible, Interoperable, and Reusable) [@FAIR] data principles.
 We are working on creating a shiny stand-alone application with the described functionality and making it available soon.

### 2022-10-26 inst/MyArticle/paper.Rmd
Small changes in Methods and Modelling sections

--- 2022-10-25 inst/MyArticle/paper.Rmd
@@
 
-When developing the R package `r BiocStyle::Githubpkg("fgcz/prolfqua")` we were inspired by the R package `r BiocStyle::CRANpkg("caret")` [@caret2008] which enables us to call various machine learning (ML) methods, and makes selecting the best ML algorithm for your problem easy.
+When developing the R package `r BiocStyle::Githubpkg("fgcz/prolfqua")` we were inspired by the R package `r BiocStyle::CRANpkg("caret")` [@caret2008] which enables us to call various Machine Learning (ML) methods, and makes selecting the best ML algorithm for your problem easy.
 We aimed for a similar R package for the differential expression analysis of quantitative proteomics data.
@@
 
-To evaluate the performance of differential expression analysis, we use the IonStar benchmark dataset[@shen2018ionstar], available from the Proteomics Identifications Database (PRIDE) identifier PXD003881.
+To evaluate the performance of differential expression analysis, we use the IonStar benchmark dataset[@shen2018ionstar] available from the Proteomics Identifications Database (PRIDE) identifier PXD003881.
 $DH5\alpha$ _E.
@@
 There are four technical replicates for each dilution, hence 20 raw files in total.
-We processed the raw data using MaxQuant [@Cox2008MaxQuant] version Version 1.6.10.43, with MaxQuant default settings.
+We processed the raw data using MaxQuant [@Cox2008MaxQuant] version 1.6.10.43 with default settings.
 The text files generated by MaxQuant are available in the _prolfquadata_ R package [@prolfquadata].
@@
 
-The peptide abundances (from the MaxQuant _peptide.txt_ file) were $log_2$ transformed and subsequently scaled, where median and the mean average deviation (mad) were obtained from the human proteins only.
+The peptide abundances (from the MaxQuant _peptide.txt_ file) were $log_2$ transformed and subsequently scaled, where median and the mean average deviation were obtained from the human proteins only.
 We removed _one hit wonders_, i.e., proteins with a single peptide assignment.
-Protein abundances are inferred from the peptide intensities using Tukey's median polish.
+Protein abundances were inferred from the peptide intensities using Tukey's median polish.
 Finally, we fitted the fixed effect linear models, the dropout model `r BiocStyle::Biocpkg("proDA")` to protein abundances, the mixed effect linear model, the `r BiocStyle::Biocpkg("PECA")` model, and hurdle model `r BiocStyle::Biocpkg("msqrob2")` to peptide intensities.
@@
 We know that for _H.
-sapiens_ proteins, the difference $\beta$ between two dilutions should be $\beta = 0$, while for _E.
-coli_ proteins, we know that the difference between dilutions should be $\beta \ne 0$.
+sapiens_ proteins the difference $\beta$ between two dilutions should be $\beta = 0$, while for _E.
+coli_ proteins we know that the difference between dilutions should be $\beta \ne 0$.
 
@@
 For each statistic and each value of the statistics we then compute a confusion matrix (see Table \@ref(tab:knitrConfusionMatrix)).
-From the confusion matrix we obtain measures such as true positive rate ($TPR$), false positive rate ($FPR$), or false discovery proportion ($FDP$) which are given by:
+From the confusion matrix we compute measures such as true positive rate ($TPR$), false positive rate ($FPR$), or false discovery proportion ($FDP$) using the following code snippet.
 
@@
 
-with
+where TPR, FPR and FDP are defined as follow:
 
@@
 @valikangas2016systematic discuss and benchmark various methods of peptide or protein intensity normalization, such as variance stabilizing normalization [@huber2002variance] or quantile normalization [@bolstad2003comparison].
-In this work, we use a robust version of the z-score, where instead of the mean $\bar{x}$ we use the median, and instead of the standard deviation $\tilde{S}$ we use the median absolute deviation (mad):
+In this work, we use a robust version of the z-score, where instead of the mean we use the median $\tilde{x}$, and instead of the standard deviation we use the median absolute deviation $\tilde{S}$:
 
@@
 
-Because we need to estimate the differences among groups on the original scale, we must multiply the $z$-score by the average variance of all the $n$ samples in the experiment.
+Because we need to estimate the differences among groups on the original scale, we must multiply the $z$-score by the average standard deviation of all the $n$ samples in the experiment.
 
 \begin{eqnarray} 
-z' = z \cdot \frac{1}{n}\sum_{i=1}^n S_i
+z' = z \cdot \frac{1}{n}\sum_{i=1}^n \tilde{S}_i
 \end{eqnarray}
@@
 For the Ionstar dataset, we used the intensities of _H.
-sapiens_ proteins, whose concentrations do not change, to determine $\bar{x}$ and $S$ and then applied it to all the intensities (including _E.
+sapiens_ proteins, whose concentrations do not change, to determine $\tilde{x}$ and $\tilde{S}$ and then applied it to all the intensities (including _E.
 coli_) in the sample.
@@
 
-Given a linear model $y = \beta X$, we can compute the difference $\beta_{c}$ between two groups by a linear combination of weights $c$ and model parameters $\beta$, where $c$ is a column vector with as many elements as there are coefficients $\beta$ in the linear model.
+Given a linear model $y = \beta X$, we can compute the difference $\beta_{c}$ between two groups by a linear combination $c$ of parameters $\beta$, where $c$ is a column vector with as many elements as there are coefficients $\beta$ in the linear model.
 If $c$ has $0$ for one or more of its rows, then the corresponding coefficient in $\beta$ is not involved in determining the contrast [@ph525xseries].
 
-The difference estimate $\beta_c$, is given by the dot product:
+The estimate of the difference $\beta_c$ is given by the dot product:
 
@@
 
-The following code illustrates how to specify a model with two factors and an interaction term and define the differences we want to estimate.
-The example dataset was acquired in two batches (p2370, p2691).
-In each, we measured Yeast samples grown under two different conditions (Glucose and Ethanol).
-This example is taken from the `r BiocStyle::Githubpkg("fgcz/prolfqua")` package vignette file.
-To estimate treatment differences, we first specify the linear model that explains the observed protein abundances (transformedIntensities) using the explanatory variables `condition_` and `batch_` and the interaction term  `condition_:batch_`.
-Then, after fitting the model, we can estimate the treatment difference among the groups `Ethanol` and `Glucose` defined by the factor `condition` and similarly among the groups defined by `batch_`.
-Furthermore, we can examine differences between `Ethanol` and `Glucose` within each batch.
-We also can assess if these differences are the same or are different in both batches, i.e., if there is an interaction between condition and batch.
-We also show the code to compute the array of weights $c$ used to multiply the model coefficient $\beta$.
+The linear model explains the observed protein abundances using the explanatory variables `condition_` and `batch_` and the interaction term  `condition_:batch_`.
+The treatment difference among the groups `Ethanol` and `Glucose` defined by the factor `condition`, and similarly among the groups defined by `batch_`, can be specified.
+The treatment differences among `Ethanol` and `Glucose` within the batches `p2370` and `p2691` can be also specified.
+Finally, we can examine if the treatment differences in both batches are the same or are different.
+The code used to obtain the effect size estimates and the variance of the estimate by computing the array of weights $c$ is shown below:
 
 ```{r contrast2, eval = TRUE, echo = TRUE}
-
-Yeast2Factor <- prolfqua::prolfqua_data("data_Yeast2Factor")
-Yeast2Factor$data |> dplyr::select(condition_, batch_) |> dplyr::distinct()
 mlm <- prolfqua::strategy_lm(
@@
   )
+Yeast2Factor <- prolfqua::prolfqua_data("data_Yeast2Factor")
 mm <- prolfqua::build_model(Yeast2Factor$data, mlm, subject_Id = "protein_Id")
@@
 knitr::kable(mC, caption = "Weights $c$ for each of the contrasts (rows) which will be applied to the model parameter (columns).",format = "latex")
+
 ```
 
-Given the weights $c$ contrasts from model parameter estimates obtained with `r BiocStyle::Biocpkg('proDA')`, or `r BiocStyle::Biocpkg('msqrob2')` can be determined, by this enabling us to call these packages using the same application programming interface (API).
+This example is taken from the `r BiocStyle::Githubpkg("fgcz/prolfqua")` package vignette file.
+Given the weights $c$, contrasts from model parameter estimates obtained with `r BiocStyle::Biocpkg('proDA')`, or `r BiocStyle::Biocpkg('msqrob2')` can be determined, which enables us to call these packages using the same API.
 
@@
 In such a case, we will substitute the group mean, using the expected protein abundance $A$ at the LOD $A_{LOD}$.
-To estimate $A_{LOD}$ we are using the protein abundances of those groups where the protein was observed in only a single sample. Typically there are many such cases, and hence we take the median.
+To estimate $A_{LOD}$ we are using the abundances from those groups where the protein was unobserved except for one sample. Typically there are many such cases, and hence we take the median.
 
@@
 However, if for instance no observations are present in group $b$, we will use $\bar{b} = A_{LOD}$.
-Furthermore, if $\bar{a} < A_{LOD}$ we also set $\bar{a} = A_{LOD}$, or more formally:
+Furthermore, if $\bar{a} < A_{LOD}$ we also set $\bar{a} = A_{LOD}$.
 
@@
 $$
+
 We use the pooled variance in all groups to estimate the protein variance, assuming they are the same. The pooled variance $s_p^2$ is given by:
@@
 R linear model and linear mixed effect models allow modeling parallel designs, repeated measurements, factorial designs, and many more. 
-Models in `r BiocStyle::Githubpkg("fgcz/prolfqua")` are specified using R's linear model and mixed model formula interface.
+Models in `r BiocStyle::Githubpkg("fgcz/prolfqua")` are specified using R’s linear model and mixed model formula interface.
 Therefore, knowledge of the R regression model infrastructure [@faraway2016extending; @MASS2002] is advantageous when using our package.
 Furthermore, this glass box approach should make it easy to reimplement an analysis performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` using base R or other programming languages by reading the analysis script.
-However, in the package documentation, we showcase how a user, without this knowledge, can analyze experiments with a parallel group design and a factorial design.
 
@@
 
-To ease the usage barriers of the R package to users not familiar with statistics and R programming, we developed an application based on the prolfqua package into our data management platform [B-Fabric](https://fgcz-bfabric.uzh.ch/) [@bfabric,@bfabricjib2022].
+To ease the usage barriers of the R package to users not familiar with statistics and R programming, we developed an application based on the prolfqua package into our data management platform [B-Fabric](https://fgcz-bfabric.uzh.ch/) [@bfabric; @bfabricjib2022].
 The B-Fabric system runs a computing infrastructure controlled by a local resource management system that supports cloud-bursting [@VMMAD].
-This integration enables users to select the input data and basic settings in a graphical user interface (GUI).
-
-In this GUI application, we determine the model formula and which comparisons to compute from the sample annotation provided in tabular form, similarly to SAINTexpress and `r BiocStyle::Biocpkg("MSstats")`.
-However, this application can analyze only parallel group designs with and without repeated measurements or factorial designs without interactions.
+This integration enables users to select the input data and basic settings in a graphical user interface.
+We determine the model formula and which comparisons to compute from the sample annotation provided in tabular form, similarly to SAINTexpress and `r BiocStyle::Biocpkg("MSstats")`.
+However, only parallel group designs with and without repeated measurements are analyzed.
 The user receives a report including input files, the R markdown file, and R scripts necessary to replicate the analysis using their in-house R installation.
-In this way, `r BiocStyle::Githubpkg(`fgcz/prolfqua`)` and [B-Fabric](https://fgcz-bfabric.uzh.ch/) and helps scientists to meet requirements from funding agencies, journals, and academic institutions to publish data according to the FAIR (Findable, Accessible, Interoperable, and Reusable) [@FAIR] data principles.
-We are working on creating a shiny stand-alone application with the described functionality and making it available soon.
+In this way, [B-Fabric](https://fgcz-bfabric.uzh.ch/) helps scientists to meet requirements from funding agencies, journals, and academic institutions to publish data according to the FAIR (Findable, Accessible, Interoperable, and Reusable) [@FAIR] data principles.
+We are working on creating a stand-alone shiny application with the described functionality.
 

### 2022-10-26 inst/MyArticle/paper.Rmd
added changes from MdE.

--- 2022-10-26 inst/MyArticle/paper.Rmd
@@
 
-The peptide abundances (from the MaxQuant _peptide.txt_ file) were $log_2$ transformed and subsequently scaled, where median and the mean average deviation (mad) were obtained from the human proteins only.
+The peptide abundances (from the MaxQuant _peptide.txt_ file) were $log_2$ transformed and subsequently scaled, where median and the mean absolute deviation were obtained from the human proteins only.
 We removed _one hit wonders_, i.e., proteins with a single peptide assignment.
@@
 @valikangas2016systematic discuss and benchmark various methods of peptide or protein intensity normalization, such as variance stabilizing normalization [@huber2002variance] or quantile normalization [@bolstad2003comparison].
-In this work, we use a robust version of the z-score, where instead of the mean $\bar{x}$ we use the median, and instead of the standard deviation $\tilde{S}$ we use the median absolute deviation (mad):
+In this work, we use a robust version of the z-score, where instead of the mean we use the median $\tilde{x}$ , and instead of the standard deviation we use the median absolute deviation $\tilde{S}$:
 
@@
 
-Given a linear model $y = \beta X$, we can compute the difference $\beta_{c}$ between two groups by a linear combination of weights $c$ and model parameters $\beta$, where $c$ is a column vector with as many elements as there are coefficients $\beta$ in the linear model.
+Given a linear model $y = \beta X$, we can compute the difference $\beta_{c}$ between two groups by the dot product of weights $c$ and model parameters $\beta$, where $c$ is a column vector with as many elements as there are coefficients $\beta$ in the linear model.
 If $c$ has $0$ for one or more of its rows, then the corresponding coefficient in $\beta$ is not involved in determining the contrast [@ph525xseries].
@@
 
-Given the weights $c$ contrasts from model parameter estimates obtained with `r BiocStyle::Biocpkg('proDA')`, or `r BiocStyle::Biocpkg('msqrob2')` can be determined, by this enabling us to call these packages using the same application programming interface (API).
+Given the weights $c$ contrasts from model parameters obtained with `r BiocStyle::Biocpkg('proDA')`, or `r BiocStyle::Biocpkg('msqrob2')` can be determined, which will enable us to implement adapters to `proDA` or `msqrob2`.
 
@@
 In such a case, we will substitute the group mean, using the expected protein abundance $A$ at the LOD $A_{LOD}$.
-To estimate $A_{LOD}$ we are using the protein abundances of those groups where the protein was observed in only a single sample. Typically there are many such cases, and hence we take the median.
+To estimate $A_{LOD}$ we are using the protein abundances of those groups where the protein was observed in only a single sample (see Supplement section Estimating $A_{LOD}$).
+Typically there are many such cases, and hence we take the median.
 

### 2022-10-27 inst/MyArticle/paper.Rmd
Add changes to Modelling, Results and Conclusion

@@
 
-However, if there is no observation in a group, the model fit fails.
+However, if there is no observation in a group the model fit fails.
 For example, suppose a protein is unobserved in all the samples of a group.
 In that case, a plausible explanation is that the protein abundance is below the limit of detection (LOD) of the MS instrument.
-In such a case, we will substitute the group mean, using the expected protein abundance $A$ at the LOD $A_{LOD}$.
+In such a case, we will substitute the group mean using the expected protein abundance $A$ at the LOD $A_{LOD}$.
 To estimate $A_{LOD}$ we are using the protein abundances of those groups where the protein was observed in only a single sample (see Supplement section Estimating $A_{LOD}$).
@@
 
-Where, $n_g$ is the number of groups and $n$ is the number of observations.
-If variance can not be estimated for a protein, because there are to few observations in other groups, we use the median pooled variance of all other proteins in the dataset.
+Where $n_g$ is the number of groups and $n$ is the number of observations.
+If variance can not be estimated for a protein, because there are too few observations in other groups, we use the median pooled variance of all other proteins in the dataset.
 
@@
 From the linear and the mixed effect models, we can obtain the residual standard deviation $\sigma$, and degrees of freedom $df$.
-@Smyth2004linear discuss how, to use the $\sigma$ and $df$ of all models to estimate a prior $\sigma$ and prior $df$, and posterior $\tilde \sigma$.
+@Smyth2004linear discuss how, to use the $\sigma$ and $df$ of all models to estimate the corresponding priors and posterior $\tilde \sigma$.
 These can be used to moderate the $t$-statistics by:
@@
 
-To summarize peptide level models to protein models, we apply the method suggested by @Suomi2017bEnhanced that use the median scaled $p$-value of the peptide models and the cumulative distribution function of the Beta distribution (CDF) to determine a regulation probability of the protein.
+To summarize peptide level models to protein models, we apply the method suggested by @Suomi2017bEnhanced that uses the median scaled $p$-value of the peptide models and the cumulative distribution function of the Beta distribution (CDF) to determine a regulation probability of the protein.
 
-To obtain the $\tilde{p}$ of a protein, we first rescaled the peptide $p$-values by taking the sign of the fold-change $\hat \beta$ into account, i.e.:
+To obtain the $\tilde{p}$ of a protein we first rescaled the peptide $p$-values by taking the sign of the fold-change $\hat \beta$ into account, i.e.:
 
@@
 
-Afterwards, the median scaled $p$-value $\tilde{p}_s$ is determined and using the transformation below, transformed back onto the original scale:
+Afterwards, the median scaled $p$-value $\tilde{p}_s$ is determined and, using the transformation below, transformed back onto the original scale:
 
@@
 
-Because we use the median, with the i-th order statistic $i = \frac{n}{2} + 0.5$, we parametrize the CDF of the Beta distribution with $\gamma = i = \frac{n}{2} + 0.5$ and $\delta = n - i + 1 = n - (\frac{n}{2} + 0.5) + 1 = \frac{n}{2} + 0.5 = \gamma$.
+Because we use the median with the i-th order statistic $i = \frac{n}{2} + 0.5$, we parametrize the CDF of the Beta distribution with $\gamma = i = \frac{n}{2} + 0.5$ and $\delta = n - i + 1 = n - (\frac{n}{2} + 0.5) + 1 = \frac{n}{2} + 0.5 = \gamma$.
 We implemented this method in the class `ContrastROPECA` (Figure \@ref(fig:ContrastUML)).
@@
 Then we create the plots shown in Figure \@ref(fig:figure2).
-Panel A shows the distribution of the $p$-values, Panel B shows the volcano plot for each comparison, and Panel C shows a Bland-Altman plot which shows the difference between groups as a function of the rank of the protein abundance.
+Panel A shows the distribution of the $p$-values, Panel B shows the volcano plot for each comparison, and Panel C shows a Bland-Altman plot reporting the difference between groups as a function of the rank of the protein abundance.
 
@@
 Using the data frame of tidy data ensures interoperability with other proteomics-related packages that manage their data with tidy-tables, e.g., `r BiocStyle::CRANpkg("protti")` [@quast2022protti].
-To simplify integration of `r BiocStyle::Githubpkg("fgcz/prolfqua")` with Bioconductor based workflows there is a method to convert the `LFQData` class into a `r BiocStyle::Biocpkg("SummarizedExperiment")`.
-The use of R6 classes, which encapsulate the configuration and the data, allows for writing very concise code, where functions can have few arguments.
-Auto-completion support for R6 classes in the Rstudio editor makes it easy for novices to explore `r BiocStyle::Githubpkg("fgcz/prolfqua")`s functionality (see Supplementary Figure S[X]).
-
+To simplify the integration of `r BiocStyle::Githubpkg("fgcz/prolfqua")` with Bioconductor based workflows there is a method that converts the `LFQData` class into a `r BiocStyle::Biocpkg("SummarizedExperiment")`.
+The use of R6 classes, which encapsulate the configuration and the data, allows for writing very concise code where functions can have few arguments.
+Auto-completion support for R6 classes in the Rstudio editor makes it easy for novices to explore `r BiocStyle::Githubpkg("fgcz/prolfqua")'s functionality (see Supplementary Figure S[X]).
 To ease the usage barriers of the R package to users not familiar with statistics and R programming, we developed an application based on the prolfqua package into our data management platform [B-Fabric](https://fgcz-bfabric.uzh.ch/) [@bfabric,@bfabricjib2022].
@@
 The user receives a report including input files, the R markdown file, and R scripts necessary to replicate the analysis using their in-house R installation.
-In this way, `r BiocStyle::Githubpkg("fgcz/prolfqua")` and [B-Fabric](https://fgcz-bfabric.uzh.ch/) and helps scientists to meet requirements from funding agencies, journals, and academic institutions to publish data according to the FAIR (Findable, Accessible, Interoperable, and Reusable) [@FAIR] data principles.
+In this way, `r BiocStyle::Githubpkg("fgcz/prolfqua")` and [B-Fabric](https://fgcz-bfabric.uzh.ch/) and helps scientists to meet requirements from funding agencies, journals, and academic institutions while publishing their data according to the FAIR (Findable, Accessible, Interoperable, and Reusable) [@FAIR] data principles.
 We are working on creating a shiny stand-alone application with the described functionality and making it available soon.
@@
 We use those scores, i.e.
-the $pAUC$ at $10%$ $FDR$ and the $FDP$, to examine how well the methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` model quantitative mass spectrometric high throughput data and compare them with results produced by `r BiocStyle::Biocpkg("MSstats")`, `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Biocpkg("msqrob2")`.
-Table \@ref(tab:describeBenchmarkedModels) summarizes all modelling methods we evaluated.
+the $pAUC$ at $10%$ $FDR$ and the $FDP$, to examine how well the methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` model quantitative mass spectrometric high throughput data and compare them with the results produced by `r BiocStyle::Biocpkg("MSstats")`, `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Biocpkg("msqrob2")`.
+Table \@ref(tab:describeBenchmarkedModels) summarizes all the modelling methods we evaluated.
 
@@
 
-When comparing differential expression analysis performance, a relevant parameter is the number of proteins for which a method estimated differences (see Figure \@ref(fig:figure3) Panel A), which indicates how robust the procedure works in the presence of missing observations.
+When comparing differential expression analysis performance, a relevant parameter is the number of proteins for which a method estimates differences (see Figure \@ref(fig:figure3) Panel A), which indicates how robust the procedure works in the presence of missing observations.
 For each protein, we tried to determine four differences ($\Delta = (1.20, 1.25, 1.30, 1.50)$).
@@
 
-Figure \@ref(fig:figure3) Panel B shows how various estimates obtained from the models, i.e., the difference between groups, $t$-statistics, and scaled $p$-values allow identifying true positives (TP) given a false positive rate (FPR) of $10%$, by displaying the partial area under the ROC.
+Figure \@ref(fig:figure3) Panel B shows how various estimates obtained from the models, i.e., the difference between groups, $t$-statistics and scaled $p$-values, allow identifying true positives (TP) given a false positive rate (FPR) of $10%$ by displaying the partial area under the ROC.
 Ordering proteins by the $t$-statistic or $p$-value leads to a higher $pAUC_{10}$ than when ordering by the estimated difference among groups.
@@
 In that case, the linear models fitted to protein intensities, calculated using Tukey's median polish, perform better than those directly modeling peptide intensities, e.g., `ropeca` or `prot_mixed_effect_moderated` (see Table \@ref(tab:describeBenchmarkedModels) Abundance column).
-We speculate that outliers at the peptide level do not affect the protein estimates when using Tukey's median polish method, a non-parametric method to infer protein abundances.
-This hypothesis could be examined, by including other forms of protein intensity inferences implemented in `prolfqua`, e.g.
-`top-N` or `rlm`, into the benchmark.
+We speculate that outliers at the peptide level do not affect the protein estimates when using Tukey's median polish method.
+This hypothesis is examined in the benchmark, by including other forms of protein intensity inferences implemented in `prolfqua`, e.g.
+`top-N` or `rlm`.
 
@@
 Figure \@ref(fig:figure3) Panel C shows the $FDP$, obtained from the confusion matrix, as a function of the $FDR$ determined from the model.
-Most lines are below the diagonal, indicating that the $FDR$ estimates are modestly conservative for this particular benchmark dataset.
+Most lines are below the diagonal, which indicates that the $FDR$ estimates are modestly conservative for this particular benchmark dataset.
 In the case of `r BiocStyle::Biocpkg("MSstats")`, we observe a high proportion of false discoveries for small $FDR$ values.
@@
 The IonStar dataset (see Methods) captures only the variance from the chromatography, electro-spray, and mass spectrometric measurements since only technical replicates are available for each dilution.
-Therefore essential sources of variation typically present in other experiments, such as the biochemical and biological, are not measured.
+Therefore essential sources of variation typically present in other experiments, such as the biochemical and biological ones, are not measured.
 Furthermore, this dataset with a parallel group design does not allow for benchmarking models with interactions.
@@
 Furthermore, for the fixed effect linear model, the empirical Bayes variance shrinkage, as suggested by @Smyth2004linear, consistently improves the ranking of proteins compared with the unmoderated estimates (not shown).
-However, since also for this method, a correct degree of freedom estimate is required, it does not work for mixed effect models.
+However, since also for this method a correct degree of freedom estimate is required, it does not work for mixed effect models.
 
@@
 However, the properties of Beta-based probabilities are not well understood; for instance, the p-values are not uniformly distributed under the null hypothesis.
-Furthermore, the $FDR$ estimates obtained when correcting for multiple testing with the Benjamini-Hochberg method are biased and overestimate the false discovery proportion (see Figure \@ref(fig:figure3) Panel C).
+Furthermore, the $FDR$ estimates obtained when correcting for multiple testing with the Benjamini-Hochberg method are biased and overestimate the FDP (see Figure \@ref(fig:figure3) Panel C).
 Therefore, we can not recommend this method if an unbiased estimate of $FDR$ is essential, which is frequently the case.
@@
 
-Since, the API of `r BiocStyle::Biocpkg("proDA")` and the produced outputs are similar to those of prolfqua, we are integrating the `r BiocStyle::Biocpkg("proDA")` package as an additional modeling option into `r BiocStyle::Githubpkg("fgcz/prolfqua")` (see Figure \@ref(fig:ContrastUML)).
+Since the API of `r BiocStyle::Biocpkg("proDA")` and the produced outputs are similar to those of prolfqua, we are integrating the `r BiocStyle::Biocpkg("proDA")` package as an additional modeling option into `r BiocStyle::Githubpkg("fgcz/prolfqua")` (see Figure \@ref(fig:ContrastUML)).
 
@@
 However, when using MaxLFQ intensities reported by MaxQuant, the $pAUC_{10}$ is lower ($pAUC_{10}(\textrm{t-statistics}) = 66%$) compared with results obtained when protein abundances are estimated from peptides abundances using Tukey's median polish ($pAUC_{10}(\textrm{t-statistics}) = 72%$).
-Last but not least, the software [@Cox2008MaxQuant; @yu2020fast], used to identify and quantify proteins, significantly contributes to the entire pipeline's performance, altering the number of identified proteins and the sensitivity and specificity of the differential expression analysis.
-In the supplement we show the benchmarking results for `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Githubpkg("fgcz/prolfqua")` when the IonStar dataset is prepossessed with the FragPipe v15 [@yu2020fast] quantification pipeline.
+Last but not least, the software [@Cox2008MaxQuant; @yu2020fast] used to identify and quantify proteins significantly contributes to the entire pipeline's performance altering the number of identified proteins and the sensitivity and specificity of the differential expression analysis.
+In the Supplement we show the benchmarking results for `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Githubpkg("fgcz/prolfqua")` when the IonStar dataset is prepossessed with the FragPipe v15 [@yu2020fast] quantification pipeline.
 Although the number of proteins identified with two peptides is slightly lower than with MaxQuant (3929 vs.
@@
 Furthermore, the developed framework can integrate other modeling methods, e.g., the probabilistic dropout model [@bioxrvproDA2020] or accurate variance estimation [@zhu2020deqms].
-Hence, `r BiocStyle::Githubpkg("fgcz/prolfqua")` enables the implementation of applications where the user can select an alternative normalization method, protein abundance estimation method, or differential expression analysis algorithm.
+Hence, `r BiocStyle::Githubpkg("fgcz/prolfqua")` enables the implementation of applications where the user can select an alternative normalization method, protein abundance estimation methods, or differential expression analysis algorithms.
 Furthermore, this R package can analyze other types of quantitative proteomics data,e.g., labeling-based TMT-, PRM- DIA-data.
@@
 It is relevant that an analysis pipeline shows good performance in all these categories.
-Leveraging these computer experiments, we can provide the following advice: i) estimate protein abundances from peptide abundances using a robust or nonparametric regression method; ii) use linear models because they show good performance in all categories; iii) if the measurements are correlated, as for technical replicates, mixed effect models might work if the sample sizes are large; if not, aggregate the replicates and fit a linear model instead; iv) if you use fixed-effect linear models, apply variance moderation to improve the $t$-statistics and $p$-value estimates; v) If you want to sort your protein lists to perform gene set enrichment analysis, use the $t$-statistic instead of the difference; and last but not least vi) do not impute missing observation but statistically model missingness to estimate parameters, i.e., group differences.
+Leveraging these computational experiments, we can provide the following advice: i) estimate protein abundances from peptide abundances using a robust or nonparametric regression method; ii) use linear models because they show good performance in all categories; iii) if the measurements are correlated, as for technical replicates, mixed effect models might work if the sample sizes are large; if not, aggregate the replicates and fit a linear model instead; iv) if you use fixed-effect linear models, apply variance moderation to improve the $t$-statistics and $p$-value estimates; v) If you want to sort your protein lists to perform gene set enrichment analysis, use the $t$-statistic instead of the difference; and last but not least vi) do not impute missing observation but statistically model missingness to estimate parameters, i.e., group differences.
 Finally, the differential expression analysis results obtained with prolfqua are comparable to those of other differential expression analysis tools.

### 2022-10-27 inst/MyArticle/paper.Rmd
new Biocpkg function

@@
 ```{r setup, include=FALSE}
+
+Biocpkg <- function(pkg, label = NULL){
+  url <- file.path("https://bioconductor.org/packages", pkg)
+  BiocStyle:::labelled_link(pkg , label , url)
+}
+
 ## See ACS guidelines [https://publish.acs.org/publish/author_guidelines?coden=ancham#manuscript_text_components]
@@
 
-Several data analysis packages exist to model mass spectrometry protein quantification experiments, e.g., `r BiocStyle::Biocpkg("limma")` [@Ritchie2015], `r BiocStyle::Biocpkg("MSstats")` [@Choi2014], `r BiocStyle::Biocpkg("PECA")` [@Suomi2017bEnhanced],  `r BiocStyle::Biocpkg("msqrob2")` [@Goeminne2016] or `r BiocStyle::Biocpkg("proDA")` [@bioxrvproDA2020], to mention some, all implemented in R.
-Each of them has some unique features, for example, `r BiocStyle::Biocpkg("MSstats")` determines the statistical model from the structure of the sample annotation, which allows users with limited statistical knowledge to perform differential expression analysis (DEA).
-At the same time, `r BiocStyle::Biocpkg("limma")` allows the specification of a design matrix using a linear model formula and implements the empirical Bayes variance shrinkage method.
-The package `r BiocStyle::Biocpkg("PECA")` performs a roll-up of peptide level differences and peptide level $p$-value estimates obtained from `r BiocStyle::Biocpkg("limma")` or `r BiocStyle::Biocpkg("PECA")`, to protein level estimates.
-Furthermore, `r BiocStyle::Biocpkg("msqrob2")` combines two models: robust linear models fitted to protein abundances and a quasibinomial generalized linear model fitted to peptide counts, into Hurdle model.
-The `r BiocStyle::Biocpkg("proDA")` package implements a linear probabilistic dropout model to account for missing data without imputation.
+Several data analysis packages exist to model mass spectrometry protein quantification experiments, e.g., `r Biocpkg("limma")` [@Ritchie2015], `r Biocpkg("MSstats")` [@Choi2014], `r Biocpkg("PECA")` [@Suomi2017bEnhanced],  `r Biocpkg("msqrob2")` [@Goeminne2016] or `r Biocpkg("proDA")` [@bioxrvproDA2020], to mention some, all implemented in R.
+Each of them has some unique features, for example, `r Biocpkg("MSstats")` determines the statistical model from the structure of the sample annotation, which allows users with limited statistical knowledge to perform differential expression analysis (DEA).
+At the same time, `r Biocpkg("limma")` allows the specification of a design matrix using a linear model formula and implements the empirical Bayes variance shrinkage method.
+The package `r Biocpkg("PECA")` performs a roll-up of peptide level differences and peptide level $p$-value estimates obtained from `r Biocpkg("limma")` or `r Biocpkg("PECA")`, to protein level estimates.
+Furthermore, `r Biocpkg("msqrob2")` combines two models: robust linear models fitted to protein abundances and a quasibinomial generalized linear model fitted to peptide counts, into Hurdle model.
+The `r Biocpkg("proDA")` package implements a linear probabilistic dropout model to account for missing data without imputation.
 
 Of note are the various approaches to handle missing observations, which are inherent in mass spectrometric bottom-up experiments.
-For instance, `r BiocStyle::Biocpkg("MSstats")` handles missing data by feature filtering and imputation.
-Other means of modeling missing observations are the Hurdle models discussed by @goeminne2020msqrob, while the R package `r BiocStyle::Biocpkg("proDA")` models missingness using probabilistic dropout models [@bioxrvproDA2020].
+For instance, `r Biocpkg("MSstats")` handles missing data by feature filtering and imputation.
+Other means of modeling missing observations are the Hurdle models discussed by @goeminne2020msqrob, while the R package `r Biocpkg("proDA")` models missingness using probabilistic dropout models [@bioxrvproDA2020].
 
@@
 Table \@ref(tab:tableOverview) gives an overview of the models and features supported by these packages.
-We see that, for instance, `r BiocStyle::Biocpkg("limma")` and `r BiocStyle::Biocpkg("proDA")` allow us to fit a comprehensive variety of models and test various hypotheses; however, good knowledge of the design matrix specification using the R formula interface is required [@law2020guide].
+We see that, for instance, `r Biocpkg("limma")` and `r Biocpkg("proDA")` allow us to fit a comprehensive variety of models and test various hypotheses; however, good knowledge of the design matrix specification using the R formula interface is required [@law2020guide].
 
@@
 For example, we implemented strategies to estimate protein intensities from peptide intensities: top N [@Grossmann2010], Tukey's median polish  [@tukey1977exploratory], robust linear models [@goeminne2020msqrob].
-Peptide or protein abundances can then be scaled and transformed using robust scaling, _quantile_ normalization or `r BiocStyle::Biocpkg("vsn")` to remove systematic differences among samples and heteroscedasticity.
-In this respect, `r BiocStyle::Githubpkg("fgcz/prolfqua")` can be compared with R packages such as `r BiocStyle::Biocpkg("DEP")` [@DEP2018] or `r BiocStyle::Biocpkg("POMA")` [@POMA2021] which support the entire differential expression analysis pipeline.
+Peptide or protein abundances can then be scaled and transformed using robust scaling, _quantile_ normalization or `r Biocpkg("vsn")` to remove systematic differences among samples and heteroscedasticity.
+In this respect, `r BiocStyle::Githubpkg("fgcz/prolfqua")` can be compared with R packages such as `r Biocpkg("DEP")` [@DEP2018] or `r Biocpkg("POMA")` [@POMA2021] which support the entire differential expression analysis pipeline.
 
-We use the Ionstar [@shen2018ionstar] and CPTAC (available from `r BiocStyle::Biocpkg("msdata")`) dataset to benchmark the modeling methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` and to compare our results with those of `r BiocStyle::Biocpkg("MSstats")` and `r BiocStyle::Biocpkg("proDA")`.
+We use the Ionstar [@shen2018ionstar] and CPTAC (available from `r Biocpkg("msdata")`) dataset to benchmark the modeling methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` and to compare our results with those of `r Biocpkg("MSstats")` and `r Biocpkg("proDA")`.
 Since group sizes are relatively small, typically with four or five subjects per group, the high power of the tests is a relevant criterion to assess the modeling method.
@@
 Using an R6 [@R6cite] configuration object (Figure \@ref(fig:LFQData)), we specify what variable is in which column making it easy to integrate new inputs in `r BiocStyle::Githubpkg("fgcz/prolfqua")` if provided as a tidy data.
-For example, to visualize tidy Spectronaut[@bruderer2015extending], DiaNN[@demichev2020dia], or Skyline[@maclean2010skyline] outputs, or data in `r BiocStyle::Biocpkg("MSstats")`[@Choi2014] format, only a few lines of code are needed to update the `r BiocStyle::Githubpkg("fgcz/prolfqua")` `AnalysisTableConfiguration` configuration.
+For example, to visualize tidy Spectronaut[@bruderer2015extending], DiaNN[@demichev2020dia], or Skyline[@maclean2010skyline] outputs, or data in `r Biocpkg("MSstats")`[@Choi2014] format, only a few lines of code are needed to update the `r BiocStyle::Githubpkg("fgcz/prolfqua")` `AnalysisTableConfiguration` configuration.
 The configuration encapsulates the differences among the various input formats in column names and enables the use of the methods without additional arguments.
@@
 Several implementations allow to perform differential expression analysis given linear or mixed effect models (`Contrasts`), variance shrinkage (`ContrastsModerated`), or to impute contrasts in cases when observations are missing for an entire group of samples (`ContrastsSimpleImpute`).
-Further implementations of the interface encapsulate and integrate differential expression analysis results of external tools such as `r BiocStyle::Biocpkg("proDA")` or of SAINTexpress [@teo2014saintexpress] used to analyze data from protein interaction studies.
+Further implementations of the interface encapsulate and integrate differential expression analysis results of external tools such as `r Biocpkg("proDA")` or of SAINTexpress [@teo2014saintexpress] used to analyze data from protein interaction studies.
 
@@
 Protein abundances are inferred from the peptide intensities using Tukey's median polish.
-Finally, we fitted the fixed effect linear models, the dropout model `r BiocStyle::Biocpkg("proDA")` to protein abundances, the mixed effect linear model, the `r BiocStyle::Biocpkg("PECA")` model, and hurdle model `r BiocStyle::Biocpkg("msqrob2")` to peptide intensities.
+Finally, we fitted the fixed effect linear models, the dropout model `r Biocpkg("proDA")` to protein abundances, the mixed effect linear model, the `r Biocpkg("PECA")` model, and hurdle model `r Biocpkg("msqrob2")` to peptide intensities.
 
@@
 The package `r BiocStyle::Githubpkg("fgcz/prolfqua")` provides functions to determine the vector of $parameter$ weights $c$, from a linear model and a contrast specification string.
-
-The following code illustrates how to specify a model with two factors and an interaction term and define the differences we want to estimate.
-The example dataset was acquired in two batches (p2370, p2691).
-In each, we measured Yeast samples grown under two different conditions (Glucose and Ethanol).
-This example is taken from the `r BiocStyle::Githubpkg("fgcz/prolfqua")` package vignette file.
-To estimate treatment differences, we first specify the linear model that explains the observed protein abundances (transformedIntensities) using the explanatory variables `condition_` and `batch_` and the interaction term  `condition_:batch_`.
-Then, after fitting the model, we can estimate the treatment difference among the groups `Ethanol` and `Glucose` defined by the factor `condition` and similarly among the groups defined by `batch_`.
-Furthermore, we can examine differences between `Ethanol` and `Glucose` within each batch.
-We also can assess if these differences are the same or are different in both batches, i.e., if there is an interaction between condition and batch.
-We also show the code to compute the array of weights $c$ used to multiply the model coefficient $\beta$.
-
-```{r contrast2, eval = TRUE, echo = TRUE}
-
-Yeast2Factor <- prolfqua::prolfqua_data("data_Yeast2Factor")
-Yeast2Factor$data |> dplyr::select(condition_, batch_) |> dplyr::distinct()
-mlm <- prolfqua::strategy_lm(
-  transformedIntensity ~ condition_ + batch_ + condition_:batch_
-  )
-mm <- prolfqua::build_model(Yeast2Factor$data, mlm, subject_Id = "protein_Id")
-
-Contrasts <- c(
-"diff_condition" = "condition_Ethanol - condition_Glucose",
-"diff_batch" = "batch_p2370 - batch_p2691",
-"diff_within_p2370" = "`condition_Glucose:batch_p2370` - `condition_Ethanol:batch_p2691`",
-"diff_within_p2691" = "`condition_Glucose:batch_p2691` - `condition_Ethanol:batch_p2691`",
-"diff_of_difference" = "diff_within_p2370 - diff_within_p2691")
-
-linfct <- prolfqua::linfct_from_model(mm$modelDF$linear_model[[1]],
-                                      as_list = FALSE)
-mC <- prolfqua::linfct_matrix_contrasts(linfct, Contrasts)
-## see Table 3 for content of mC
-```
-
-Table \@ref(tab:contrcoefficient) shows weights $c$ for each of the five contrasts we specified.
-
-```{r contrcoefficient}
-knitr::kable(mC, caption = "Weights $c$ for each of the contrasts (rows) which will be applied to the model parameter (columns).",format = "latex")
-```
-
-Given the weights $c$ contrasts from model parameters obtained with `r BiocStyle::Biocpkg('proDA')`, or `r BiocStyle::Biocpkg('msqrob2')` can be determined, which will enable us to implement adapters to `proDA` or `msqrob2`.
+In the Supplement we provided an example, how to specify contrasts for a dataset with two explanatory variable with an interaction term (Section "Specifying Contrasts for Models with two Factors and Interaction Term").
 
@@
 If variance can not be estimated for a protein, because there are too few observations in other groups, we use the median pooled variance of all other proteins in the dataset.
-
 This method is implemented in the class `ContrastSimpleImpute` (see Figure \@ref(fig:ContrastUML)).
@@
 Using the data frame of tidy data ensures interoperability with other proteomics-related packages that manage their data with tidy-tables, e.g., `r BiocStyle::CRANpkg("protti")` [@quast2022protti].
-To simplify the integration of `r BiocStyle::Githubpkg("fgcz/prolfqua")` with Bioconductor based workflows there is a method that converts the `LFQData` class into a `r BiocStyle::Biocpkg("SummarizedExperiment")`.
+To simplify the integration of `r BiocStyle::Githubpkg("fgcz/prolfqua")` with Bioconductor based workflows there is a method that converts the `LFQData` class into a `r Biocpkg("SummarizedExperiment")`.
 The use of R6 classes, which encapsulate the configuration and the data, allows for writing very concise code where functions can have few arguments.
-Auto-completion support for R6 classes in the Rstudio editor makes it easy for novices to explore `r BiocStyle::Githubpkg("fgcz/prolfqua")'s functionality (see Supplementary Figure S[X]).
+Auto-completion support for R6 classes in the Rstudio editor makes it easy for novices to explore `r BiocStyle::Githubpkg("fgcz/prolfqua")`'s functionality (see Supplementary Figure S[X]).
 To ease the usage barriers of the R package to users not familiar with statistics and R programming, we developed an application based on the prolfqua package into our data management platform [B-Fabric](https://fgcz-bfabric.uzh.ch/) [@bfabric,@bfabricjib2022].
@@
 
-In this GUI application, we determine the model formula and which comparisons to compute from the sample annotation provided in tabular form, similarly to SAINTexpress and `r BiocStyle::Biocpkg("MSstats")`.
+In this GUI application, we determine the model formula and which comparisons to compute from the sample annotation provided in tabular form, similarly to SAINTexpress and `r Biocpkg("MSstats")`.
 However, this application can analyze only parallel group designs with and without repeated measurements or factorial designs without interactions.
 The user receives a report including input files, the R markdown file, and R scripts necessary to replicate the analysis using their in-house R installation.
-In this way, `r BiocStyle::Githubpkg("fgcz/prolfqua")` and [B-Fabric](https://fgcz-bfabric.uzh.ch/) and helps scientists to meet requirements from funding agencies, journals, and academic institutions while publishing their data according to the FAIR (Findable, Accessible, Interoperable, and Reusable) [@FAIR] data principles.
+In this way, `r BiocStyle::Githubpkg("fgcz/prolfqua")` and [B-Fabric](https://fgcz-bfabric.uzh.ch/) and helps scientists to meet requirements from funding agencies, journals, and academic institutions while publishing their data according to the FAIR  [@FAIR] data principles.
 We are working on creating a shiny stand-alone application with the described functionality and making it available soon.
@@
 We use those scores, i.e.
-the $pAUC$ at $10%$ $FDR$ and the $FDP$, to examine how well the methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` model quantitative mass spectrometric high throughput data and compare them with the results produced by `r BiocStyle::Biocpkg("MSstats")`, `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Biocpkg("msqrob2")`.
+the $pAUC$ at $10%$ $FDR$ and the $FDP$, to examine how well the methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` model quantitative mass spectrometric high throughput data and compare them with the results produced by `r Biocpkg("MSstats")`, `r Biocpkg("proDA")` and `r Biocpkg("msqrob2")`.
 Table \@ref(tab:describeBenchmarkedModels) summarizes all the modelling methods we evaluated.
@@
 
-p1_pAUC <- ggplot2::ggplot(resAllB, ggplot2::aes(x = Name, y = pAUC_10)) +
+pB_pAUC <- ggplot2::ggplot(resAllB, ggplot2::aes(x = Name, y = pAUC_10)) +
   ggplot2::geom_bar(stat = "identity") +
@@
   ggplot2::xlab("") +
-  ggplot2::labs(tag = "A")
+  ggplot2::labs(tag = "B")
 
@@
 The set of proteins with effect size estimates might differ for each method, biasing direct comparison of scores such as the $pAUC$.
-For instance, we observe that `r BiocStyle::Biocpkg("MSstats")` estimates $16058$ group differences while the mixed effect models estimates the fewest with $15940$.
+For instance, we observe that `r Biocpkg("MSstats")` estimates $16058$ group differences while the mixed effect models estimates the fewest with $15940$.
 Hence, to conclude that one method shows a better performance, it does not suffice if the $pAUC$ is greater, but the number of proteins with differential expression results needs to be equal or larger.
@@
 Suppose an accurate estimate of the difference among groups is essential.
-In that case, the linear models fitted to protein intensities, calculated using Tukey's median polish, perform better than those directly modeling peptide intensities, e.g., `ropeca` or `prot_mixed_effect_moderated` (see Table \@ref(tab:describeBenchmarkedModels) Abundance column).
+In that case, the linear models fitted to protein intensities, calculated using Tukey's median polish, perform better than those when directly modeling peptide intensities, e.g., `ropeca` or `prot_mixed_effect_moderated` (see Table \@ref(tab:describeBenchmarkedModels) Abundance column).
 We speculate that outliers at the peptide level do not affect the protein estimates when using Tukey's median polish method.
-This hypothesis is examined in the benchmark, by including other forms of protein intensity inferences implemented in `prolfqua`, e.g.
-`top-N` or `rlm`.
-
-There are only minor differences in the $pAUC_{10}$ between the $t$-statistics or the scaled $p$-value  (see Figure \@ref(fig:figure3) Panel A).
-Interestingly, the $pAUC$ increases when using $p$-values instead of the $t$-statistics for linear models, while it decreases for mixed effect models.
-The reason is an erroneous denominator degree of freedom estimation for many proteins, a known problem in the case of mixed effect models.
 
@@
 Most lines are below the diagonal, which indicates that the $FDR$ estimates are modestly conservative for this particular benchmark dataset.
-In the case of `r BiocStyle::Biocpkg("MSstats")`, we observe a high proportion of false discoveries for small $FDR$ values.
+In the case of `r Biocpkg("MSstats")`, we observe a high proportion of false discoveries for small $FDR$ values.
 In the case of the ROPECA method, the $FDR$ estimates, obtained by applying the Benjamini-Hochberg correction to the Beta distribution-based regulation probabilities, strongly overestimate the $FDP$.
@@
             vjust = 0, hjust = -0.2, angle = -90) #+ 
-pB <- nrgg + ggplot2::labs(tag = "B")
+pA_barplot <- nrgg + ggplot2::labs(tag = "A")
 
@@
 
-pC <- ddb |> ggplot2::ggplot(ggplot2::aes(y = FDP,  x  = FDR )) + 
+pC_FDR <- ddb |> ggplot2::ggplot(ggplot2::aes(y = FDP,  x  = FDR )) + 
   ggplot2::geom_line(ggplot2::aes(color = model_name, linetype = model_name)) +
@@
 
-gridExtra::grid.arrange(p1_pAUC, pB, pC, layout_matrix = hlay)
+gridExtra::grid.arrange(pB_pAUC, pA_barplot, pC_FDR, layout_matrix = hlay)
 
@@
 
-Using a benchmark dataset whose ground truth is known (see Methods), we assessed the performance of different modeling approaches implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` (Tables \@ref(tab:prolfquaModels) and \@ref(tab:describeBenchmarkedModels)), `r BiocStyle::Biocpkg("MSstats")`, `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Biocpkg("msqrob2")`.
+Using a benchmark dataset whose ground truth is known (see Methods), we assessed the performance of different modeling approaches implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` (Tables \@ref(tab:prolfquaModels) and \@ref(tab:describeBenchmarkedModels)), `r Biocpkg("MSstats")`, `r Biocpkg("proDA")` and `r Biocpkg("msqrob2")`.
 Table \@ref(tab:describeBenchmarkedModels) summarizes which methods we have evaluated, which MaxQuant files we used as input, and if the models are fitted to peptide or protein intensities.
@@
     kableExtra::kable_styling(latex_options = "scale_down") |> 
-    kableExtra::add_footnote(label = "prolfqua_merged - Merges results of the prot_med_lm_moderated and prot_imputation modeling pipeline, preferring those of prot_med_lm_moderated if available.")
+    kableExtra::add_footnote(label = "'prolfqua_merged' - Augments estimates which are missing in 'prot_med_lm_mod' with those from 'prot_impute'.")
 
@@
 
-The IonStar dataset (see Methods) captures only the variance from the chromatography, electro-spray, and mass spectrometric measurements since only technical replicates are available for each dilution.
+The IonStar/MaxQuant dataset (see Methods) captures only the variance from the chromatography, electro-spray, and mass spectrometric measurements since only technical replicates are available for each dilution.
 Therefore essential sources of variation typically present in other experiments, such as the biochemical and biological ones, are not measured.
@@
 We can conclude that if we want to sort the proteins according to the likelihood of being differentially regulated to perform gene set enrichment analysis [@subramanian2005gene], the $t$-statistic is better suited than the fold-change estimate.
-Modeling the degrees of freedom when computing the $p$-values improves the inference (see Figure \@ref(fig:figure3) panel B).
-There is no such improvement for the mixed effect model, because the degrees of freedom are erroneously estimated.
+When computing the $p$-values from the t-statistics we are incorporating the degrees of freedom, which improve the inferences (see Figure \@ref(fig:figure3) panel B center versus panel B left).
+There is no such improvement for the mixed effect model.
+The reason is an erroneous denominator degree of freedom estimation for many proteins, a known problem in the case of mixed effect models.
 Furthermore, for the fixed effect linear model, the empirical Bayes variance shrinkage, as suggested by @Smyth2004linear, consistently improves the ranking of proteins compared with the unmoderated estimates (not shown).
@@
 
-Computing the statistics at the peptide level, e.g., the $t$-statistics or $p$-value, then summarizing these statistics using their median produces the highest $AUC$ scores among all the tested models (see Figure \@ref(fig:figure3) Panel A left).
-Furthermore, by using the Beta distribution to model the number of peptides observed, we can further improve the $pAUC$ scores (see Figure \@ref(fig:figure3) Panel A center).
-However, the properties of Beta-based probabilities are not well understood; for instance, the p-values are not uniformly distributed under the null hypothesis.
+Computing the statistics at the peptide level, e.g., the $t$-statistics, then summarizing it for each protein using their median, produces the highest $pAUC_{10}$ scores among all the tested models (see Figure \@ref(fig:figure3) Panel B `prolfqua_ropeca`).
+Furthermore, by using the Beta distribution to model the number of peptides observed, we can further improve the $pAUC$ scores (see Figure \@ref(fig:figure3) Panel B center).
+However, the properties of Beta-based probabilities are not well understood; for instance, it is not uniformly distributed under the null hypothesis (see Supplementary Section "The probabilities produced by ROPECA are not p- values").
 Furthermore, the $FDR$ estimates obtained when correcting for multiple testing with the Benjamini-Hochberg method are biased and overestimate the FDP (see Figure \@ref(fig:figure3) Panel C).
@@
 
-The R-package `r BiocStyle::Biocpkg("proDA")`, `r BiocStyle::Biocpkg("msqrob2")` and `r BiocStyle::Biocpkg("prolfqua")` do not impute missing data but integrate them into the statistical model, while MSstats filters and imputes the data using an accelerated failure model.
-Despite imputation, `r BiocStyle::Biocpkg("MSstats")` estimates group differences for less proteins and does not achieve a higher $pAUC_{10}$.
-Furthermore, Figure \@ref(fig:figure3) Panel C shows that when using `r BiocStyle::Biocpkg("MSstats")` the proportion of false discoveries might be very high even when filtering for a low $FDR$.
+The R-package `r Biocpkg("proDA")`, `r Biocpkg("msqrob2")` and `r Biocpkg("prolfqua")` do not impute missing data but integrate them into the statistical model, while `r Biocpkg("MSstats")` filters and imputes the data using an accelerated failure model.
+Despite imputation, `r Biocpkg("MSstats")` estimates group differences for less proteins, and does not achieve a higher $pAUC_{10}$.
+Furthermore, Figure \@ref(fig:figure3) Panel C shows that when using `r Biocpkg("MSstats")` the proportion of false discoveries might be very high for a low $FDR$ because of false positives.
+Of note, MSStats uses the _evidence.txt_ file, while all the other methods use _peptide.txt_ files, as input (see Table \@ref(tab:describeBenchmarkedModels))), a possible confounding factor to consider, but we can not control for.
 
-We obtained difference estimates for all proteins, as shown in Figure \@ref(fig:figure3) Panel A when using: a) the probabilistic dropout model (`r BiocStyle::Biocpkg("proDA")`), b) the hurdle model (`r BiocStyle::Biocpkg("msqrob2")`), and c) when combining the results of the linear model with contrast estimates in the presence of missing data (prolfqua_merged).
-Furthermore, the performance of the scaled $p$-values or the $t$-statistics is comparable among these three methods (Figure \@ref(fig:figure3) Panel B).
-We tested if there is no difference between the $pAUC_{10}$ for all three methods and did not reject this hypothesis at a significance level of $0.4$ (see Supplement, Testing ROC curves).
-Furthermore, all three models perform similarly well when examined using the IonStar/MaxQuant, IonStar/FragPipe, or CPTAC dataset (see Supplement, Benchmark CPTAC, and Benchmark FragPipe).
-While, for the IonStar/MaxQuant dataset, msqrob2 performed best, proDA performed best for the CPTAC/MaxQuant data, and prolfqua_merged for the IonStar/FragPipe dataset.
+We obtained difference and FDR estimates for all proteins, as shown in Figure \@ref(fig:figure3) Panel A when using: a) the probabilistic dropout model (*proDA*), b) the hurdle model (*msqrob2*), and c) `prolfqua_merge` (Table \@ref{tab:describeBenchmarkedModels}).
+We observe that the performance of the scaled $p$-values or the $t$-statistics is comparable among these three methods (Figure \@ref(fig:figure3) Panel B).
+We tested if there is a significant difference between the $pAUC_{10}$ for all three methods, but did not reject the null hypothesis that there is no such difference (Supplementary Section "Comparing pAUC for IonStar/MaxQuant dataset").
 Also, the FDR estimates (Figure \@ref(fig:figure3) Panel C) are comparable for all three methods.
+Furthermore, all three models perform similarly well when examined using CPTAC/MaxQuant dataset (see Supplementary section "Benchmarking msqrob2, proDA and prolfqua using the CPTAC dataset").
+In addition we also examined the performance of the *proDA* and `prolfqua_merged` model using the IonStar dataset but processed with the `FragPipe v15` software (see Supplement).
+While, for the IonStar/MaxQuant dataset, *msqrob2* performed best, *proDA* performed best for the CPTAC/MaxQuant data, and `prolfqua_merged` for the IonStar/FragPipe dataset.
 
-Since the API of `r BiocStyle::Biocpkg("proDA")` and the produced outputs are similar to those of prolfqua, we are integrating the `r BiocStyle::Biocpkg("proDA")` package as an additional modeling option into `r BiocStyle::Githubpkg("fgcz/prolfqua")` (see Figure \@ref(fig:ContrastUML)).
-
-We focused our benchmark on comparing the statistical modeling methods using three different scores while we fixed the pre-processing steps.
-However, some of these steps are of utmost significance when performing a differential expression analysis [@frohlich2022benchmarking].
+We focused our benchmark on comparing the statistical modeling methods while we fixed the pre-processing steps.
+However, some of these steps are of utmost significance when performing differential expression analysis [@frohlich2022benchmarking].
 One of them is the normalization of the intensities within the samples to remove systematic differences [@pursiheimo2015optimization].
 The method used to infer proteins from peptide identifications [@nesvizhskii2005interpretation] and protein abundances from peptide abundances is an additional important factor [@Grossmann2010].
-For instance, the original `r BiocStyle::Biocpkg("proDA")` publication uses MaxLFQ [@cox2014accurate] protein estimates.
+For instance, the original `r Biocpkg("proDA")` publication uses MaxLFQ [@cox2014accurate] protein estimates.
 However, when using MaxLFQ intensities reported by MaxQuant, the $pAUC_{10}$ is lower ($pAUC_{10}(\textrm{t-statistics}) = 66%$) compared with results obtained when protein abundances are estimated from peptides abundances using Tukey's median polish ($pAUC_{10}(\textrm{t-statistics}) = 72%$).
 Last but not least, the software [@Cox2008MaxQuant; @yu2020fast] used to identify and quantify proteins significantly contributes to the entire pipeline's performance altering the number of identified proteins and the sensitivity and specificity of the differential expression analysis.
-In the Supplement we show the benchmarking results for `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Githubpkg("fgcz/prolfqua")` when the IonStar dataset is prepossessed with the FragPipe v15 [@yu2020fast] quantification pipeline.
-Although the number of proteins identified with two peptides is slightly lower than with MaxQuant (3929 vs.
-4046), the pAUC_{10} is significantly higher ($81%$ vs.
+In the Supplement we show the benchmarking results for *proDA* and *prolfqua* when the IonStar dataset is prepossessed with the FragPipe v15 [@yu2020fast] quantification pipeline.
+Although the number of proteins identified with two peptides is slightly lower with FragPipe than with MaxQuant (3929 vs.
+4046), the $pAUC_{10}$ is significantly higher ($81%$ vs.
 $72%$).
@@
 \toprule
-Abbreviations & Explaination\\
+Abbreviations & Explanation\\
 \midrule
+API    &application programming interface          \\
 AUC    &Area Under the Curve                               \\
@@
 ESI-MS &Electro-Spray-Ionization Mass Spectrometry         \\
+FAIR   &Findable, Accessible, Interoperable, and Reusable  \\
 $FDP$  &False Discovery Proportion                         \\
@@
 LOD    &Limit Of Detection                                 \\
-MAR    & Missing At Random                                 \\ 
-MCAR   & Missing Completely At Random                      \\
+MAR    &Missing At Random                                 \\ 
+MCAR   &Missing Completely At Random                      \\
+ML     & machine learning                                 \\
 MS     &mass spectrometry                                  \\
 OO     &Object-Oriented                                    \\
+pAUC   & partial Are Under the Curve                       \\
+ROC    & Receiver operator curve                           \\
 UML    & Unified Modeling Language                        \\

### 2022-10-27 inst/MyArticle/prolfqua_supplement.Rmd
explicit dplyr::filter etc. close #20
new Biocpkg function

@@
 \renewcommand{\thepage}{S--\arabic{page}}
-\renewcommand{\tablename}{Supp. Table}
-\renewcommand{\figurename}{Supp. Figure}
+\renewcommand{\tablename}{Supplementary Table}
+\renewcommand{\figurename}{Supplementary Figure}
 
@@
 ```{r setup, include=FALSE}
+Biocpkg <- function(pkg, label = NULL){
+  url <- file.path("https://bioconductor.org/packages", pkg)
+  BiocStyle:::labelled_link(pkg , label , url)
+}
+
 knitr::opts_chunk$set(echo = FALSE, error = FALSE, warning = FALSE, message = FALSE)
@@
 
-# Comparing partial areas under the curve at 0.1 FDR for Ionstar/MaxQuant dataset
+# Comparing pAUC at 0.1 FDR for Ionstar/MaxQuant dataset
 
@@
 
-Figure \@ref(fig:plotCPTACbarplot) shows the partial area under the curve for `r BiocStyle::Biocpkg("msqrob2")`, `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Githubpkg("fgcz/prolfqua")`, when using the difference among the groups, the scaled p-values or the t-statistics to compute the ROC curves.
+Supplementary Figure \@ref(fig:plotCPTACbarplot) shows the partial area under the curve for `r BiocStyle::Biocpkg("msqrob2")`, `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Githubpkg("fgcz/prolfqua")`, when using the difference among the groups, the scaled p-values or the t-statistics to compute the ROC curves.
 
@@
 
-Figure \@ref(fig:plotCPTACbarplot) shows the partial area under the curve for `r BiocStyle::Biocpkg("msqrob2")`, `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Githubpkg("fgcz/prolfqua")`, computed using the effect size estimate among groups (diff), the scaled p-value (scaled.p.value) and the t-statistics (t_statistics),
+Supplementary Figure \@ref(fig:plotCPTACbarplot) shows the partial area under the curve for `r BiocStyle::Biocpkg("msqrob2")`, `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Githubpkg("fgcz/prolfqua")`, computed using the effect size estimate among groups (diff), the scaled p-value (scaled.p.value) and the t-statistics (t_statistics),
 
@@
 The main text describes how we estimate the abundance at the detection limit $A_{LOD}$. 
-Figure \@ref(fig:missignessHistogram) shows the distribution of the average abundance within a group for all proteins.
+Supplementary Figure \@ref(fig:missignessHistogram) shows the distribution of the average abundance within a group for all proteins.
 The red density shows the distribution for those proteins without a missing value, and the olive, turquoise, and lilac density show for proteins with 1,2 and 3 missing values, respectively.
@@
 We know that if $H0$ is true, the distribution of the $p-values$ will be uniform.
-(Figure \@ref(fig:figROPECA) left Panel).
+(Supplementary Figure \@ref(fig:figROPECA) left Panel).
 Therefore, we can omit the step of generating the data from $H0$ and computing the p-values but simulate the p-values directly by sampling from the uniform distribution (`runif(nrPep)`).
@@
 To make the data more realistic, that is that some proteins have a lot of peptides, and many have just a few peptides; the density of sampling the protein Id is exponential (`dexp(seq(0,5,length = 800)`).
-Figure \@ref(fig:figROPECA) center panel shows the number of proteins as a function of the number of peptides per protein.
+Supplementary Figure \@ref(fig:figROPECA) center panel shows the number of proteins as a function of the number of peptides per protein.
 Afterward, we apply the ROPECA methods as discussed in the methods section of the manuscript.
-Figure \@ref(fig:figROPECA) shows the distribution of the Beta distribution-based probabilities, which is not uniform.
+Supplementary Figure \@ref(fig:figROPECA) shows the distribution of the Beta distribution-based probabilities, which is not uniform.
 Hence, they do not have the same interpretation as p-values, i.e., the probability of falsely rejecting the $H0$ if $H0$ is true.
 
-```{r figROPECA , fig.cap = "Left panel - distibution of peptide level p.values.
+```{r figROPECA, fig.cap = "Left panel - distibution of peptide level p.values.
 Center panel - number of proteins as a function of the number of peptides per protein.
@@
 ```
+
+# Specifying Contrasts for Models with two Factors and Interaction Term
+
+The following code illustrates how to specify a model with two factors and an interaction term and define the differences we want to estimate.
+The example dataset was acquired in two batches (p2370, p2691).
+In each, we measured Yeast samples grown under two different conditions (Glucose and Ethanol).
+This example is taken from the `r BiocStyle::Githubpkg("fgcz/prolfqua")` package vignette file.
+To estimate treatment differences, we first specify the linear model that explains the observed protein abundances (transformedIntensities) using the explanatory variables `condition_` and `batch_` and the interaction term  `condition_:batch_`.
+Then, after fitting the model, we can estimate the treatment difference among the groups `Ethanol` and `Glucose` defined by the factor `condition` and similarly among the groups defined by `batch_`.
+Furthermore, we can examine differences between `Ethanol` and `Glucose` within each batch.
+We also can assess if these differences are the same or are different in both batches, i.e., if there is an interaction between condition and batch.
+We also show the code to compute the array of weights $c$ used to multiply the model coefficient $\beta$.
+
+```{r contrast2, eval = TRUE, echo = TRUE}
+
+Yeast2Factor <- prolfqua::prolfqua_data("data_Yeast2Factor")
+Yeast2Factor$data |> dplyr::select(condition_, batch_) |> dplyr::distinct()
+mlm <- prolfqua::strategy_lm(
+  transformedIntensity ~ condition_ + batch_ + condition_:batch_
+  )
+mm <- prolfqua::build_model(Yeast2Factor$data, mlm, subject_Id = "protein_Id")
+
+Contrasts <- c(
+"diff_condition" = "condition_Ethanol - condition_Glucose",
+"diff_batch" = "batch_p2370 - batch_p2691",
+"diff_within_p2370" = 
+  "`condition_Glucose:batch_p2370` - `condition_Ethanol:batch_p2691`",
+"diff_within_p2691" = 
+  "`condition_Glucose:batch_p2691` - `condition_Ethanol:batch_p2691`",
+"diff_of_difference" = "diff_within_p2370 - diff_within_p2691")
+
+linfct <- prolfqua::linfct_from_model(mm$modelDF$linear_model[[1]],
+                                      as_list = FALSE)
+mC <- prolfqua::linfct_matrix_contrasts(linfct, Contrasts)
+## see Table 3 for content of mC
+```
+
+Table \@ref(tab:contrcoefficient) shows weights $c$ for each of the five contrasts we specified.
+
+```{r contrcoefficient}
+knitr::kable(mC, caption = "Weights $c$ for each of the contrasts (rows) which will be applied to the model parameter (columns).",format = "latex")
+```
+
+Given the weights $c$ contrasts from model parameters obtained with `r Biocpkg('proDA')`, or `r Biocpkg('msqrob2')` can be determined, which will enables us to implement adapters to `proDA` or `msqrob2`.
+
+For more details see `r BiocStyle::Githubpkg("fgcz/prolfqua")` vignette: ["Modelling dataset with two Factors"](https://fgcz.github.io/prolfqua/articles/Modelling2Factors.html).
 

### 2022-10-28 inst/MyArticle/paper.Rmd
Cleanup in paper.Rmd

@@
 At the same time, they have the following features in common: fitting linear models to either peptide or protein intensities, determining differences among groups, and afterward applying empirical Bayes variance shrinkage.
-Therefore, the revised goal was to provide a modular object-oriented (OO) design, with R linear models as a core, and add features such as $p$-value aggregation, variance shrinkage, or modeling of missing observations.
+Therefore, the revised goal was to provide a modular object-oriented design, with R linear models as a core, and add features such as $p$-value aggregation, variance shrinkage, or modeling of missing observations.
 
@@
 
-## Dataset for benchmarking
+```{r prolfquaModels}
+xM <- data.frame( 
+    lm = c("strategy_lm, Contrasts", "linear modelling of peptide or protein abundances and group difference estimation"),
+    lmer = c("strategy_lmer, Contrasts", "mixed effect modelling of peptide or protein abundances and group differences estimation" ),
+    gmi = c("ContrastsSimpleImpute" , "group difference estimation when no observations in one of the groups" ),
+    ROPECA = c("ContrastsROPECA", "estimating group differences for proteins by summarizing peptide differences"),
+    limma = c( "ContrastsModerated", "empirical Bayes variance shrinkage for group difference estimates (limma)"),
+    pic = c("runSaint, ContrastsSaintExpress", "protein interaction scoring (SAINTExpress)"),
+    prodA = c( "strategy_proDA*, ContrastsProDA", "adapter to the probabilistic dropout model implemented in proDA"), check.names =  FALSE)
+xM <- data.frame(t(xM))
+
+colnames(xM) <- c("prolfqua functions", "model")
+knitr::kable(tibble::as_tibble(xM), caption = "prolfqua functions that can be used to fit various models.",
+             format = "latex", booktabs = TRUE) |>
+    kableExtra::kable_styling(latex_options = "scale_down") |> 
+    kableExtra::add_footnote(label = "in development")
+
+```
+
+## Datasets for benchmarking
+
+### IonStar/MaxQuant
 
@@
 To compare the performance of the various methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` we use only the contrasts resulting in minor differences $\Delta = (1.20, 1.25, 1.30, 1.50)$, because for bigger differences all models perform similarly.
+
+### IonStar/FragPipe
+
+### CPTAC/MaxQuant
 
@@
 The package `r BiocStyle::Githubpkg("fgcz/prolfqua")` provides functions to determine the vector of $parameter$ weights $c$, from a linear model and a contrast specification string.
-In the Supplement we provided an example, how to specify contrasts for a dataset with two explanatory variable with an interaction term (Section "Specifying Contrasts for Models with two Factors and Interaction Term").
+In the Supplement we provided an example, how to specify contrasts for a dataset with two explanatory variable with an interaction term (Supplementary Section "Specifying Contrasts for Models with two Factors and Interaction Term").
 
@@
 
-## plot panels A-D
+## plot Figure 3 panels A-D
 pl <- lfqpep$get_Plotter()
@@
 
-## visualize results in panels
+## plot Figure 4 panels A - C
 pl <- contrasts$get_Plotter() 
@@
 For each protein, we tried to determine four differences ($\Delta = (1.20, 1.25, 1.30, 1.50)$).
-Given 4046 proteins with at least two peptides, there are in total $16184$ possible differences.
+Given 4178 proteins with at least two peptides, there are in total $16712$ possible differences.
 However, some methods can not estimate all of them.
 The set of proteins with effect size estimates might differ for each method, biasing direct comparison of scores such as the $pAUC$.
-For instance, we observe that `r Biocpkg("MSstats")` estimates $16058$ group differences while the mixed effect models estimates the fewest with $15940$.
+For instance, we observe that `r Biocpkg("MSstats")` estimates $16038$ group differences while the mixed effect models estimates the fewest with $15756$.
 Hence, to conclude that one method shows a better performance, it does not suffice if the $pAUC$ is greater, but the number of proteins with differential expression results needs to be equal or larger.
 
-Figure \@ref(fig:figure3) Panel B shows how various estimates obtained from the models, i.e., the difference between groups, $t$-statistics and scaled $p$-values, allow identifying true positives (TP) given a false positive rate (FPR) of $10%$ by displaying the partial area under the ROC.
+Figure \@ref(fig:figure3) Panel B shows how various estimates obtained from the models, i.e., the difference between groups, $t$-statistics and scaled $p$-values, allow identifying true positives (TP) given a false positive rate (FPR) of $10%$ by displaying the partial area under the ROC ($pAUC_{10}$).
 Ordering proteins by the $t$-statistic or $p$-value leads to a higher $pAUC_{10}$ than when ordering by the estimated difference among groups.
 Suppose an accurate estimate of the difference among groups is essential.
-In that case, the linear models fitted to protein intensities, calculated using Tukey's median polish, perform better than those when directly modeling peptide intensities, e.g., `ropeca` or `prot_mixed_effect_moderated` (see Table \@ref(tab:describeBenchmarkedModels) Abundance column).
+In that case, the linear models fitted to protein intensities, calculated using Tukey's median polish, perform better than those when directly modeling peptide intensities, i.e., `ropeca` (see Table \@ref(tab:describeBenchmarkedModels) Abundance column).
 We speculate that outliers at the peptide level do not affect the protein estimates when using Tukey's median polish method.
@@
 (ref:figure3) Panel A - Number of estimated contrasts for each modeling method (higher is better).
-Panel B - Partial area under the ROC curve at $10\%$ FPR ($pAUC_{10}$) for all contrasts and three different statistics: the difference among groups (diff - Panel B left), the scaled $p$-value (sign(diff) $\cdot$ p.value) (scaled.p.value - Panel B center) and the $t$-statistics (statistic - Panel B right), where a higher pAUC is better.
+Panel B - Partial area under the ROC curve at $10\%$ FPR ($pAUC_{10}$) for all contrasts and three different statistics: the difference among groups (diff - Panel B left), the scaled $p$-value (sign(diff) $\cdot$ p.value) (scaled.p.value - Panel B center) and the $t$-statistics (statistic - Panel B right), where a higher $pAUC_{10}$ is better.
 The red line indicates the average area under the curve of all methods.
@@
 
-```{r figure3, fig.cap = "(ref:figure3)", out.width = '90%', fig.width=12, fig.height=10}
+```{r figure3, fig.cap = "(ref:figure3)", out.width = '90%', fig.width=10, fig.height=8}
 dd <- purrr::map_df(allBenchmarks, function(x){res <- x$smc$summary; res$name <- x$model_name;res})
@@
 
-```{r prolfquaModels}
-xM <- data.frame( 
-    lm = c("strategy_lm, Contrasts", "linear modelling of peptide or protein abundances and group difference estimation"),
-    lmer = c("strategy_lmer, Contrasts", "mixed effect modelling of peptide or protein abundances and group differences estimation" ),
-    gmi = c("ContrastsSimpleImpute" , "group difference estimation when no observations in one of the groups" ),
-    ROPECA = c("ContrastsROPECA", "estimating group differences for proteins by summarizing peptide differences"),
-    limma = c( "ContrastsModerated", "empirical Bayes variance shrinkage for group difference estimates (limma)"),
-    pic = c("runSaint, ContrastsSaintExpress", "protein interaction scoring (SAINTExpress)"),
-    prodA = c( "strategy_proDA*, ContrastsProDA", "adapter to the probabilistic dropout model implemented in proDA"), check.names =  FALSE)
-xM <- data.frame(t(xM))
-
-colnames(xM) <- c("prolfqua functions", "model")
-knitr::kable(tibble::as_tibble(xM), caption = "prolfqua functions that can be used to fit various models.",
-             format = "latex", booktabs = TRUE) |>
-    kableExtra::kable_styling(latex_options = "scale_down") |> 
-    kableExtra::add_footnote(label = "in development")
-
-```
-
-Using a benchmark dataset whose ground truth is known (see Methods), we assessed the performance of different modeling approaches implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` (Tables \@ref(tab:prolfquaModels) and \@ref(tab:describeBenchmarkedModels)), `r Biocpkg("MSstats")`, `r Biocpkg("proDA")` and `r Biocpkg("msqrob2")`.
+Using a benchmark dataset whose ground truth is known (see Methods), we assessed the performance of different modeling approaches implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")`, `r Biocpkg("MSstats")`, `r Biocpkg("proDA")` and `r Biocpkg("msqrob2")`.
 Table \@ref(tab:describeBenchmarkedModels) summarizes which methods we have evaluated, which MaxQuant files we used as input, and if the models are fitted to peptide or protein intensities.
@@
 saveRDS(allBenchmarks, "allBenchmarks.RDS")
-
-```
-
-```{r roctest, include = FALSE, eval = FALSE}
-ttmsqrob <- allBenchmarks$benchmark_msqrob$data()
-ddmsqrob <- pROC::roc(ttmsqrob$TP, ttmsqrob$scaled.p.value, partial.auc = c(1, 0.9))
-ttprolfqua <- allBenchmarks$benchmark_merged$data()
-ddprolfqua <- pROC::roc(ttprolfqua$TP, ttprolfqua$scaled.p.value, partial.auc = c(1, 0.9))
-ttproda <- allBenchmarks$benchmark_proDA$data()
-ddproda <- pROC::roc(ttproda$TP, ttproda$scaled.p.value, partial.auc = c(1, 0.9))
-
-msqrob_vs_prolfqua <- pROC::roc.test(ddmsqrob,ddprolfqua)
-msqrob_vs_proda <- pROC::roc.test(ddmsqrob,ddproda)
-prolfqua_vs_proda <- pROC::roc.test(ddprolfqua,ddproda)
 
@@
 Hence, `r BiocStyle::Githubpkg("fgcz/prolfqua")` enables the implementation of applications where the user can select an alternative normalization method, protein abundance estimation methods, or differential expression analysis algorithms.
-Furthermore, this R package can analyze other types of quantitative proteomics data,e.g., labeling-based TMT-, PRM- DIA-data.
+Furthermore, this R package can analyze other types of quantitative proteomics data,e.g., label-free DIA or labeling-based TMT data.
 
@@
 \midrule
-API    &application programming interface          \\
-AUC    &Area Under the Curve                               \\
-CDF    &Cumulative Distribution Function                   \\
-ESI-MS &Electro-Spray-Ionization Mass Spectrometry         \\
-FAIR   &Findable, Accessible, Interoperable, and Reusable  \\
-$FDP$  &False Discovery Proportion                         \\
-$FDR$  &False Discovery Rate                               \\
-LC     &Liquid Chromatography                              \\
-LC-MS  &Liquid Chromatography followed by Mass Spectrometry\\
-LOD    &Limit Of Detection                                 \\
-MAR    &Missing At Random                                 \\ 
-MCAR   &Missing Completely At Random                      \\
+API    & application programming interface          \\
+AUC    & Area Under the Curve                               \\
+CDF    & Cumulative Distribution Function                   \\
+DIA    & data independent acquisition                        \\
+ESI-MS & Electro-Spray-Ionization Mass Spectrometry         \\
+FAIR   & Findable, Accessible, Interoperable, and Reusable  \\
+$FDP$  & False Discovery Proportion                         \\
+$FDR$  & False Discovery Rate                               \\
+LC     & Liquid Chromatography                              \\
+LC-MS  & Liquid Chromatography followed by Mass Spectrometry\\
+LOD    & Limit Of Detection                                 \\
+MAR    & Missing At Random                                 \\ 
+MCAR   & Missing Completely At Random                      \\
 ML     & machine learning                                 \\
-MS     &mass spectrometry                                  \\
-OO     &Object-Oriented                                    \\
+MS     & mass spectrometry                                  \\
 pAUC   & partial Are Under the Curve                       \\
 ROC    & Receiver operator curve                           \\
+TMT    & Tandem Mass Tag                                   \\
 UML    & Unified Modeling Language                        \\

### 2022-10-28 inst/MyArticle/prolfqua_supplement.Rmd
Cleanup Supplement

@@
 
-# Comparing pAUC at 0.1 FDR for Ionstar/MaxQuant dataset
+# Comparing $pAUC_{0.1}$ for Ionstar/MaxQuant dataset
 
-We test if we should reject the null hypothesis that the area under the ROC curve at 0.1 FDR, computed based on scaled p.values, does not differ for msqrob2, proDA, and prolfqua (Figure 5 in the article).
-To this task we use the function `roc.test` from the R package `r BiocStyle::CRANpkg("pROC")`.
+We test if we should reject the null hypothesis that the area under the ROC curve at $0.1$ FDR, computed based on scaled p.values, does not differ for *msqrob2*, *proDA*, and `prolfqua_merged` (Figure 5 in the Article).
+To this task we use the function `roc.test`, that performs a bootstrap method test for a difference in the (partial) AUC of the ROC curves, from the R package `r BiocStyle::CRANpkg("pROC")`, which
 

### 2022-10-28 inst/MyArticle/prolfqua_supplement.Rmd
created benchmark for MSstats.tsv ouptut from Fragpipe v15

@@
 Supplementary Figure \@ref(fig:plotCPTACbarplot) shows the partial area under the curve for `r BiocStyle::Biocpkg("msqrob2")`, `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Githubpkg("fgcz/prolfqua")`, when using the difference among the groups, the scaled p-values or the t-statistics to compute the ROC curves.
+For this dataset, the differences among groups show a higher performance than the t-statistic and scaled p-value.
+The *proDA* package performs slightly better then *prolfqua*  or *msqrob* although the differences are between the $pAUC_{10}$ for the scaled p-value score are not statistically significant.
+Table \@ref(tab:shwoROCtestREsultsCPTAC) shows the results of the Bootstrap test for two ROC curves.
+We observe that for this benchmark data, we can not reject the null hypothesis that there are no significant differences among the $pAUC_{10}$ (ROC based on scaled p-value) for the three packages.
+
+(ref:plotCPTACbarplot) Barplot showing the partial area under the ROC at 0.1 FDR. diff - differences among groups (logFC), scaled.p.value - $sign(diff) \cdot \textrm{p-value}$.
 
@@
 
-Direct comparison with `msqrob2` is impossible since to fit the hurdle model, the peptide intensities are required, while here, we are starting the analysis from the `combined_proteins.tsv` file.
+Direct comparison with `msqrob2` is impossible since to fit the hurdle model, the peptide intensities are required, while here, we are starting the analysis from the `combined_proteins.tsv` file containing protein intensities.
 Starting from protein level intensities has the advantage that the effect size estimates obtained from the model agree with the input data, a feature frequently requested by the biologist.
@@
 We see that the proteins with three missing values have, on average lower abundances than those with no missing values.
-We are using the median of the Lila density (one observation in the group), as the $A_{LOD}$ estimate.
+We are using the median of the proteins with a single observation per group; the lila density, as the $A_{LOD}$ estimate.
 

### 2022-10-29 inst/MyArticle/paper.Rmd
text updates

@@
 Several data analysis packages exist to model mass spectrometry protein quantification experiments, e.g., `r Biocpkg("limma")` [@Ritchie2015], `r Biocpkg("MSstats")` [@Choi2014], `r Biocpkg("PECA")` [@Suomi2017bEnhanced],  `r Biocpkg("msqrob2")` [@Goeminne2016] or `r Biocpkg("proDA")` [@bioxrvproDA2020], to mention some, all implemented in R.
-Each of them has some unique features, for example, `r Biocpkg("MSstats")` determines the statistical model from the structure of the sample annotation, which allows users with limited statistical knowledge to perform differential expression analysis (DEA).
-At the same time, `r Biocpkg("limma")` allows the specification of a design matrix using a linear model formula and implements the empirical Bayes variance shrinkage method.
-The package `r Biocpkg("PECA")` performs a roll-up of peptide level differences and peptide level $p$-value estimates obtained from `r Biocpkg("limma")` or `r Biocpkg("PECA")`, to protein level estimates.
-Furthermore, `r Biocpkg("msqrob2")` combines two models: robust linear models fitted to protein abundances and a quasibinomial generalized linear model fitted to peptide counts, into Hurdle model.
-The `r Biocpkg("proDA")` package implements a linear probabilistic dropout model to account for missing data without imputation.
+Each of them has some unique features, for example, *MSstats* determines the statistical model from the structure of the sample annotation, which allows users with limited statistical knowledge to perform differential expression analysis (DEA).
+At the same time, *limma* allows the specification of a design matrix using a linear model formula and implements the empirical Bayes variance shrinkage method.
+The package *PECA* performs a roll-up of peptide level differences and peptide level $p$-value estimates obtained from *limma* or *PECA*, to protein level estimates.
+Furthermore, *msqrob2* combines two models: robust linear models fitted to protein abundances and a quasibinomial generalized linear model fitted to peptide counts, into Hurdle model.
+The *proDA* package implements a linear probabilistic dropout model to account for missing data without imputation.
 
 Of note are the various approaches to handle missing observations, which are inherent in mass spectrometric bottom-up experiments.
-For instance, `r Biocpkg("MSstats")` handles missing data by feature filtering and imputation.
-Other means of modeling missing observations are the Hurdle models discussed by @goeminne2020msqrob, while the R package `r Biocpkg("proDA")` models missingness using probabilistic dropout models [@bioxrvproDA2020].
+For instance, *MSstats* handles missing data by feature filtering and imputation.
+Other means of modeling missing observations are the Hurdle models discussed by @goeminne2020msqrob, while the R package *proDA* models missingness using probabilistic dropout models [@bioxrvproDA2020].
 
@@
 Table \@ref(tab:tableOverview) gives an overview of the models and features supported by these packages.
-We see that, for instance, `r Biocpkg("limma")` and `r Biocpkg("proDA")` allow us to fit a comprehensive variety of models and test various hypotheses; however, good knowledge of the design matrix specification using the R formula interface is required [@law2020guide].
+We see that, for instance, *limma* and *proDA* allow us to fit a comprehensive variety of models and test various hypotheses; however, good knowledge of the design matrix specification using the R formula interface is required [@law2020guide].
 
@@
 
-We use the Ionstar [@shen2018ionstar] and CPTAC (available from `r Biocpkg("msdata")`) dataset to benchmark the modeling methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` and to compare our results with those of `r Biocpkg("MSstats")` and `r Biocpkg("proDA")`.
 Since group sizes are relatively small, typically with four or five subjects per group, the high power of the tests is a relevant criterion to assess the modeling method.
@@
 We will use the partial area under the receiver operator curve (ROC) to assess the power of the tests and compare the FDR with the false discovery proportion ($FDP$).
+We use the IonStar [@shen2018ionstar] and CPTAC[@edwards2015cptac]  datasets, processed with *MaxQuant* and *FragPipe* to benchmark the modeling methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` and to compare our results with those of `r Biocpkg("MSstats")`, `r Biocpkg("msqrob2")` and `r Biocpkg("proDA")`.
+
+Although other benchmark datasets exist[@wessels2012comprehensive;@o2018proteome], the IonStar dataset has the advantage that the expected differences of the spike in proteins among groups are slight compared to other benchmark datasets, making DEA more difficult and enabling us to see performance differences among the methods.
+For the same reason of making the DEA more challenging and enhancing the differences between the statistical modeling approaches, we focus on the MaxQuant outputs since they are noisier than those produced by FragPipe.
 
@@
 We use R6 classes to structure the functionality of the package (see Figure \@ref(fig:LFQData) and Figure \@ref(fig:ContrastUML)).
-R6 classes are well supported by command-line completion features in RStudio [@RStudio2022], and help to implement argument free functions (see Supplementary Figure SX).
+R6 classes are well supported by command-line completion features (see Supplementary Figure SX) in RStudio [@RStudio2022], and help to implement argument free functions.
 
@@
 Several implementations allow to perform differential expression analysis given linear or mixed effect models (`Contrasts`), variance shrinkage (`ContrastsModerated`), or to impute contrasts in cases when observations are missing for an entire group of samples (`ContrastsSimpleImpute`).
-Further implementations of the interface encapsulate and integrate differential expression analysis results of external tools such as `r Biocpkg("proDA")` or of SAINTexpress [@teo2014saintexpress] used to analyze data from protein interaction studies.
+Further implementations of the interface encapsulate and integrate differential expression analysis results of external tools such as `r Biocpkg("proDA")` or of *SAINTexpress* [@teo2014saintexpress] used to analyze data from protein interaction studies.
 
@@
 All classes estimating contrasts implement the _ContrastsInterface_.
-Results of external tools, e.g., SAINTexpress, or proDA are adapted to implement the Contrasts Interface.
+Results of external tools, e.g., *SAINTexpress*, or *proDA* are adapted to implement the Contrasts Interface.
 
@@
 
-### IonStar/MaxQuant
+### IonStar
 
@@
 There are four technical replicates for each dilution, hence 20 raw files in total.
-We processed the raw data using MaxQuant [@Cox2008MaxQuant] version Version 1.6.10.43, with MaxQuant default settings.
-The text files generated by MaxQuant are available in the _prolfquadata_ R package [@prolfquadata].
 To compare the performance of the various methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` we use only the contrasts resulting in minor differences $\Delta = (1.20, 1.25, 1.30, 1.50)$, because for bigger differences all models perform similarly.
+
+However, other  have similar constraints. Furthermore, we could not analyze the benchmark dataset [@wessels2012comprehensive] with a complete factorial design with an up-to-date version of MaxQuant.
+
+### IonStar/MaxQuant
+
+We processed the raw data of the IonStar dataset using MaxQuant [@Cox2008MaxQuant] version Version 1.6.10.43, with MaxQuant default settings for orbitrap data.
+The text files generated by MaxQuant are available in the *prolfquadata* R package [@prolfquadata].
+MaxQuant produces various output files which can be used for DEA.
+We are using the quantification results reported in the 'peptide.txt' file for DEA.
+However, *MSstats* is using the 'evidence.txt' file for the DEA.
 
@@
 
+We processed the raw data of the IonStar dataset using the FragPipe [@yu2020fast] Version 14, with default workflow for label free quantification with match between runs.
+The text files generated by FragPipe are available in the *prolfquadata* R package [@prolfquadata].
+Similarily to MaxQuant, the FragPipe produces various outputs which can e used for DEA.
+We used the total protein intensities reported in the `combined_protein.tsv` file as input for the DEA and call this dataset _IonStar/FragPipe/combined_protein.tsv_ Alternatively, we benchmarked the DEA using the MSstats.tsv file as input and call this dataset _IonStar/FragPipe/combined_protein.tsv_.
+
 ### CPTAC/MaxQuant
+
+We used the CPTAC dataset, available in the R package *msdata*, and described in [@Goeminne2016].
+In brief, the Sigma Universal Protein Standard mixture 1 (UPS1) containing $48$ different human proteins was spiked in a protein background of 60 ng/$\mu L$ Saccharomyces cerevisiae strain BY4741.
+Two different spike-in concentrations were used, 6A ($0.25$ fmol UPS1 proteins/$\mu L$) and 6B ($0.74$ fmol UPS1 proteins/$\mu L$).
+Three replicates are available for each concentrations.
+The data were searched with *MaxQuant* version `1.5.2.8.`.
 
@@
 
-The peptide abundances (from the MaxQuant _peptide.txt_ file) were $log_2$ transformed and subsequently scaled, where median and the mean absolute deviation were obtained from the human proteins only.
+The peptide abundances (from the *MaxQuant* _peptide.txt_ file) were $log_2$ transformed and subsequently scaled, where median and the mean absolute deviation were obtained from the human proteins only.
 We removed _one hit wonders_, i.e., proteins with a single peptide assignment.
@@
 To speed up the computation of these examples, we use a subset of the Ionstar dataset generated by randomly selecting $400$ proteins.
-First, we remove all proteins with a single peptide and all observations for which MaxQuant reports zero intensities which leaves $332$ proteins.
+First, we remove all proteins with a single peptide and all observations for which *MaxQuant* reports zero intensities which leaves $332$ proteins.
 Next, peptide abundances are $\log_2$ transformed and robust z-score scaled using the method `robscale`.
@@
 Auto-completion support for R6 classes in the Rstudio editor makes it easy for novices to explore `r BiocStyle::Githubpkg("fgcz/prolfqua")`'s functionality (see Supplementary Figure S[X]).
-To ease the usage barriers of the R package to users not familiar with statistics and R programming, we developed an application based on the prolfqua package into our data management platform [B-Fabric](https://fgcz-bfabric.uzh.ch/) [@bfabric,@bfabricjib2022].
+
+To ease the usage barriers of the R package to users not familiar with statistics and R programming, we developed an application based on the prolfqua package into our data management platform [B-Fabric](https://fgcz-bfabric.uzh.ch/) [@bfabric;@bfabricjib2022].
 The B-Fabric system runs a computing infrastructure controlled by a local resource management system that supports cloud-bursting [@VMMAD].
@@
 
-In this GUI application, we determine the model formula and which comparisons to compute from the sample annotation provided in tabular form, similarly to SAINTexpress and `r Biocpkg("MSstats")`.
+In this GUI application, we determine the model formula, and which comparisons to compute from the sample annotation provided in tabular form, similarly to *SAINTexpress* or *MSstats*.
 However, this application can analyze only parallel group designs with and without repeated measurements or factorial designs without interactions.
@@
 We use those scores, i.e.
-the $pAUC$ at $10%$ $FDR$ and the $FDP$, to examine how well the methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` model quantitative mass spectrometric high throughput data and compare them with the results produced by `r Biocpkg("MSstats")`, `r Biocpkg("proDA")` and `r Biocpkg("msqrob2")`.
+the $pAUC$ at $10%$ $FDR$ and the $FDP$, to examine how well the methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` model quantitative mass spectrometric high throughput data and compare them with the results produced by *MSstats*, *proDA* and *msqrob2*.
 Table \@ref(tab:describeBenchmarkedModels) summarizes all the modelling methods we evaluated.
@@
 The set of proteins with effect size estimates might differ for each method, biasing direct comparison of scores such as the $pAUC$.
-For instance, we observe that `r Biocpkg("MSstats")` estimates $16038$ group differences while the mixed effect models estimates the fewest with $15756$.
+For instance, we observe that *MSstats* estimates $16038$ group differences while the mixed effect models estimates the fewest with $15756$.
 Hence, to conclude that one method shows a better performance, it does not suffice if the $pAUC$ is greater, but the number of proteins with differential expression results needs to be equal or larger.
@@
 Most lines are below the diagonal, which indicates that the $FDR$ estimates are modestly conservative for this particular benchmark dataset.
-In the case of `r Biocpkg("MSstats")`, we observe a high proportion of false discoveries for small $FDR$ values.
+In the case of *MSstats*, we observe a high proportion of false discoveries for small $FDR$ values.
 In the case of the ROPECA method, the $FDR$ estimates, obtained by applying the Benjamini-Hochberg correction to the Beta distribution-based regulation probabilities, strongly overestimate the $FDP$.
@@
 
-Using a benchmark dataset whose ground truth is known (see Methods), we assessed the performance of different modeling approaches implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")`, `r Biocpkg("MSstats")`, `r Biocpkg("proDA")` and `r Biocpkg("msqrob2")`.
-Table \@ref(tab:describeBenchmarkedModels) summarizes which methods we have evaluated, which MaxQuant files we used as input, and if the models are fitted to peptide or protein intensities.
+Using a benchmark dataset whose ground truth is known (see Methods), we assessed the performance of different modeling approaches implemented in *prolfqua*, *MSstats*, *proDA* and *msqrob2*.
+Table \@ref(tab:describeBenchmarkedModels) summarizes which methods we have evaluated, which *MaxQuant* files we used as input, and if the models are fitted to peptide or protein intensities.
 We make the R-markdown files to replicate the benchmarking available at `r BiocStyle::Githubpkg("wolski/prolfquabenchmark")`.
@@
 Furthermore, this dataset with a parallel group design does not allow for benchmarking models with interactions.
-However, other benchmark datasets have similar constraints [@edwards2015cptac, @o2018proteome].
-Furthermore, we could not analyze the benchmark dataset [@wessels2012comprehensive] with a complete factorial design with an up-to-date version of MaxQuant.
 Thus, while we can extrapolate some of the results obtained to more realistic datasets, we need to be careful not to over-interpret our findings.
@@
 
-The R-package `r Biocpkg("proDA")`, `r Biocpkg("msqrob2")` and `r Biocpkg("prolfqua")` do not impute missing data but integrate them into the statistical model, while `r Biocpkg("MSstats")` filters and imputes the data using an accelerated failure model.
-Despite imputation, `r Biocpkg("MSstats")` estimates group differences for less proteins, and does not achieve a higher $pAUC_{10}$.
-Furthermore, Figure \@ref(fig:figure3) Panel C shows that when using `r Biocpkg("MSstats")` the proportion of false discoveries might be very high for a low $FDR$ because of false positives.
-Of note, MSStats uses the _evidence.txt_ file, while all the other methods use _peptide.txt_ files, as input (see Table \@ref(tab:describeBenchmarkedModels))), a possible confounding factor to consider, but we can not control for.
+The R-package *proDA*, *msqrob2* and *prolfqua* do not impute missing data but integrate them into the statistical model, while *MSstats* filters and imputes the data using an accelerated failure model.
+Despite imputation, *MSstats* estimates group differences for less proteins, and does not achieve a higher $pAUC_{10}$.
+Furthermore, Figure \@ref(fig:figure3) Panel C shows that when using *MSstats* the proportion of false discoveries might be very high for a low $FDR$ because of false positives.
+Of note, *MSStats* uses the _evidence.txt_ file, while all the other methods use _peptide.txt_ files, as input (see Table \@ref(tab:describeBenchmarkedModels))).
+Furthermore, *MSstats* uses equalized medians normalization, while all the other methods use robust scaling (see Methods).
+These are possible confounding factors to consider.
+While *prolfqua*, as well as *proDA*, is highly modular, and to a lesser extent *msqrob2*, enabling us to use the same data preprocessing and normalization, *MSstats* is a monolithic software, making it unfeasible to change the preprocessing or normalization.
+Therefore, we could not control these confounders.
 
@@
 Furthermore, all three models perform similarly well when examined using CPTAC/MaxQuant dataset (see Supplementary section "Benchmarking msqrob2, proDA and prolfqua using the CPTAC dataset").
-In addition we also examined the performance of the *proDA* and `prolfqua_merged` model using the IonStar dataset but processed with the `FragPipe v15` software (see Supplement).
+In addition we also examined the performance of the *proDA* and `prolfqua_merged` model using the IonStar dataset but processed with the `FragPipe v14` software.
 While, for the IonStar/MaxQuant dataset, *msqrob2* performed best, *proDA* performed best for the CPTAC/MaxQuant data, and `prolfqua_merged` for the IonStar/FragPipe dataset.
@@
 Last but not least, the software [@Cox2008MaxQuant; @yu2020fast] used to identify and quantify proteins significantly contributes to the entire pipeline's performance altering the number of identified proteins and the sensitivity and specificity of the differential expression analysis.
-In the Supplement we show the benchmarking results for *proDA* and *prolfqua* when the IonStar dataset is prepossessed with the FragPipe v15 [@yu2020fast] quantification pipeline.
+In the Supplement we show the benchmarking results for *proDA* and *prolfqua* when the IonStar dataset is prepossessed with the FragPipe v14 [@yu2020fast] quantification pipeline.
 Although the number of proteins identified with two peptides is slightly lower with FragPipe than with MaxQuant (3929 vs.
@@
 
-`r BiocStyle::Githubpkg("fgcz/prolfqua")` is a feature-rich, object-oriented R package to analyze quantitative mass spectrometric data with simple or complex experimental designs.
+`r BiocStyle::Githubpkg("fgcz/prolfqua")` is a feature-rich, object-oriented, and modular R package to analyze quantitative mass spectrometric data with simple or complex experimental designs.
 While other R packages for differential expression analysis of proteins typically only implement one modeling approach, `r BiocStyle::Githubpkg("fgcz/prolfqua")` supports various types of models (see Figure \@ref(fig:ContrastUML) and Table \@ref(tab:prolfquaModels)).

### 2022-10-29 inst/MyArticle/prolfqua_supplement.Rmd
text updates

@@
 
-# prolfqua Installation
+# How to install *prolfqua* and *prolfquabenchmark*
 
-To install the `r BiocStyle::Githubpkg("fgcz/prolfqua")` package ([ACS_JPR](https://github.com/fgcz/prolfqua/releases/tag/v0.4.3) release), including all R package dependencies, execute in the R console:
+To install the *prolfqua* package ([ACS_JPR](https://github.com/fgcz/prolfqua/releases/tag/v0.4.3) release), including all R package dependencies, execute in the R console:
 
@@
 
-The R-markdown files with all the R code to run the benchmark for `r BiocStyle::Githubpkg("prolfqua")`, `r BiocStyle::Biocpkg("msqrob2")`, `BiocStyle::Biocpkg("MSStats")` and `BiocStyle::Biocpkg("proDA")` are available in the R package `r BiocStyle::Githubpkg('wolski/prolfquabenchmark')`.
+The R-markdown files with all the R code to run the benchmark for *prolfqua*, *msqrob2*, *MSStats* and *proDA* are available in the R package `r BiocStyle::Githubpkg('wolski/prolfquabenchmark')`.
 
@@
 
-# Comparing $pAUC_{0.1}$ for Ionstar/MaxQuant dataset
+# DEA benchmark : IonStar/MaxQuant/peptide.txt - Significance test
 
-We test if we should reject the null hypothesis that the area under the ROC curve at $0.1$ FDR, computed based on scaled p.values, does not differ for *msqrob2*, *proDA*, and `prolfqua_merged` (Figure 5 in the Article).
-To this task we use the function `roc.test`, that performs a bootstrap method test for a difference in the (partial) AUC of the ROC curves, from the R package `r BiocStyle::CRANpkg("pROC")`, which
+Table \@ref(tab:ISMQpeptideSum) summarizes $pAUC$ of the DEA benchmark when using the IonStar/MaxQuant/peptide.txt data. Figure 5 in the Article, shows a barplot showing the $pAUC_{10}$.
+
+We test if we should reject the null hypothesis that the area under the ROC curve at $0.1$ FDR ($pAUC_{10}$), computed using the scaled p-values (see scaled.p.value's in Table \@ref(tab:ISMQpeptideSum) ), do not differ for *msqrob2*, *proDA*, and `prolfqua_merged`.
+To this task we use the function `roc.test`, that performs a bootstrap method test for a difference in the pAUC of the ROC curves, from the R package `r BiocStyle::CRANpkg("pROC")`.
+The Table \@ref(tab:shwoROCtestREsultsIonStarMQ) shows the p-values of the pairwise bootstrap test for partial areas under the ROC curves.
 
@@
 ddprolfqua <- pROC::roc(ttprolfqua$TP, ttprolfqua$scaled.p.value, partial.auc = c(1, 0.9))
+
 ttproda <- allBenchmarks$benchmark_proDA$data()
@@
 
-The Table \@ref(tab:shwoROCtestREsultsIonStarMQ) shows the p-values of the pairwise bootstrap test for partial areas under the ROC curves.
-We will not reject the null hypothesis that there are no difference among the ROC curves.
+```{r ISMQpeptideSum}
+mrob <- allBenchmarks$benchmark_msqrob$pAUC_summaries()$ftable$content
+mrob$package <- "msqrob2"
+proda <- allBenchmarks$benchmark_proDA$pAUC_summaries()$ftable$content
+proda$package <- "proDA"
+prolfqua <- allBenchmarks$benchmark_merged$pAUC_summaries()$ftable$content
+prolfqua$package <- "prolfqua"
+msstats <- allBenchmarks$benchmark_mssstats$pAUC_summaries()$ftable$content
+msstats$package <- "MSstats"
+
+ISMpeptide <- dplyr::bind_rows(mrob,proda,prolfqua, msstats) |> dplyr::filter(contrast == "all")
+ISMpeptide$contrast <- NULL
+ISMpeptide <-dplyr::relocate(ISMpeptide , package, .before="AUC") |> dplyr::arrange(what)
+ISMpeptide$what[ISMpeptide$what=="statistic"] <- "t_statistic"
+ISMpeptide |> 
+  knitr::kable(digits = 1, caption = "Results of the DEA beanchmark for IonStar/MaxQuant/peptide.txt")
+```
 
@@
 
-# Benchmarking msqrob2, proDA and prolfqua using the CPTAC dataset
+## Benchmark Vignettes
 
-We present here the results of benchmarking the msqrob2, proDA and prolfqua packages using the CPTAC dataset.
-The Rmarkdown file which generated these results, with more details about the CPTAC / MaxQuant dataset, can be found here:
-["Benchmarking of msqrob2, prolfqua and proDA using the CPTAC dataset MaxQuant output"](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_cptac.html)
+Pre-build version of the vignettes, we created to run the DEA benchmark, are available:
+
+- [Benchmarking of methods implemented in prolfqua using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkingIonstarData.html) 
+- [Benchmarking of MSstats using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_MSStats.html) 
+- [Benchmarking of proDA  using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_proDA_medpolish.html) 
+- [Benchmarking of msqrob2  using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkMSqRob2.html)
+
+# DEA benchmark : CPTAC/MaxQuant/peptide.txt
+
+We applied DEA using *msqrob2*, *proDA* and *prolfqua*  to the CPTAC/MaxQuant dataset and benchmarked the results. We are using the `peptide.txt` file containing the peptide intensities.
+
+The Rmarkdown file which generated these results, can be found here:
+["DEA benchmark IonStar/FragPipeV14/combined_protein.tsv"](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_cptac.html)
 
@@
 
+all$contrast <- NULL
+all |> dplyr::relocate(package, .before = "AUC") |> dplyr::arrange(what) |> 
+  knitr::kable(digits = 1, caption = "Results of the DEA beanchmark for CPTAC/MaxQuant/peptide.txt")
+
 ```
 
-Supplementary Figure \@ref(fig:plotCPTACbarplot) shows the partial area under the curve for `r BiocStyle::Biocpkg("msqrob2")`, `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Githubpkg("fgcz/prolfqua")`, when using the difference among the groups, the scaled p-values or the t-statistics to compute the ROC curves.
+Supplementary Figure \@ref(fig:plotCPTACbarplot) shows the partial area under the curve for *msqrob2*, *proDA* and *prolfqua*, when using the difference among the groups, the scaled p-values or the t-statistics to rank the proteins.
 For this dataset, the differences among groups show a higher performance than the t-statistic and scaled p-value.
@@
 
-(ref:plotCPTACbarplot) Barplot showing the partial area under the ROC at 0.1 FDR. diff - differences among groups (logFC), scaled.p.value - $sign(diff) \cdot \textrm{p-value}$.
+(ref:plotCPTACbarplot) Barplot showing the partial area under the ROC at 0.1 FDR for diff - differences among groups, scaled.p.value - $sign(diff) \cdot \textrm{p-value}$ and t-statistics.
+These results where obtained when using CPTAC/MaxQuant/peptide.txt as input.
 
-```{r plotCPTACbarplot, fig.cap= "Barplot showing the partial area under the ROC at 0.1 FDR."}
-all |> ggplot2::ggplot(ggplot2::aes(x = what, y = pAUC_10, group = package, color = NULL, fill = package)) +
-  ggplot2::geom_bar(stat = "identity",  position = ggplot2::position_dodge())
+```{r plotCPTACbarplot, fig.cap= "(ref:plotCPTACbarplot)"}
+
+ggplot2::ggplot(all, ggplot2::aes(x = package, y = pAUC_10)) +
+  ggplot2::geom_bar(stat = "identity") +
+  ggplot2::facet_wrap(~what)  + 
+  ggplot2::theme_minimal() + 
+  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = -90, vjust = 0.5)) +
+  ggplot2::xlab("")
 
@@
 
-# Benchmark based on IonStar FragPipe v15 combined_protein.tsv
+# DEA benchmark : IonStar/FragPipeV14/combined_protein.tsv
 
-The Rmarkdown file which created these results, with more details about the IonStar FragPipe v15 dataset, can be found here:
-["Benchmarking proDA and prolfqua using Ionstar Dataset and FragPipe v15 compbined_protein output"](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_cptac.html)
+The Rmarkdown file which created these results, with more details about the IonStar FragPipe v14 dataset, and the data processing, can be found here:
+["DEA benchmark IonStar/FragPipeV14/combined_protein.tsv"](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkFragPipeProteinIonStar.html)
 
-Direct comparison with `msqrob2` is impossible since to fit the hurdle model, the peptide intensities are required, while here, we are starting the analysis from the `combined_proteins.tsv` file containing protein intensities.
-Starting from protein level intensities has the advantage that the effect size estimates obtained from the model agree with the input data, a feature frequently requested by the biologist.
+After filtering for two peptides per protein the dataset comprises of $3836$ proteins.
+Since we are using the combined_protein.tsv as input, comparison with neither *msqrob2* nor *MSstats* is possible.
+To fit the hurdle (*msqrob*) model, peptide intensities are required, while *MSstats* requires the `MSstats.tsv` file, containing precursor level abundances.
+However, starting from protein level intensities has the advantage that the effect size estimates obtained from the model agree with protein abundances, a feature frequently requested by the biologist.
 
 ```{r FragPipe}
-FragPipeBenchmarks <- readRDS(getpath("FragPipev15_comb_prot.RDS"))
+FragPipeBenchmarks <- readRDS(getpath("FragPipev14_comb_prot.RDS"))
 proda <- FragPipeBenchmarks$benchmark_proDA$pAUC_summaries()$ftable$content
@@
 
-tmp <- dplyr::bind_rows(proda, prolfqua)
-tmp$what[tmp$what == "statistic"] <- "t_statistic"
-tmp$what[tmp$what == "scaled.pval"] <- "scaled.p.value"
-tmp <- tmp |> 
+ISFPprot <- dplyr::bind_rows(proda, prolfqua)
+ISFPprot$what[ISFPprot$what == "statistic"] <- "t_statistic"
+ISFPprot$what[ISFPprot$what == "scaled.pval"] <- "scaled.p.value"
+ISFPprot <- ISFPprot |> 
   dplyr::filter(contrast == "all")
+ISFPprot$contrast <- NULL
+ISFPprot |> dplyr::relocate(package, .before = "AUC") |> dplyr::arrange(what) |>
+  knitr::kable(digits = 1, caption = "Results of the DEA beanchmark for IonStar/FragPipeV14/MSstats.tsv")
+
 ```
 
-Supplementary Figure \@ref(fig:plotCPTACbarplot) shows the partial area under the curve for `r BiocStyle::Biocpkg("msqrob2")`, `r BiocStyle::Biocpkg("proDA")` and `r BiocStyle::Githubpkg("fgcz/prolfqua")`, computed using the effect size estimate among groups (diff), the scaled p-value (scaled.p.value) and the t-statistics (t_statistics),
+Supplementary Figure \@ref(fig:plotFragPipebarplot) shows the partial area under the curve for  *proDA* and *prolfqua*, computed by ranking the proteins using the difference among groups (diff), the scaled p-value (scaled.p.value) and the t-statistics (t_statistics),
 
-```{r plotFragPipebarplot, fig.cap= "Barplot showing the partial area under the ROC at 0.1 FDR."}
-tmp |> ggplot2::ggplot(ggplot2::aes(x = what, y = pAUC_10, group = package, color = NULL, fill = package)) +
-  ggplot2::geom_bar(stat = "identity",  position = ggplot2::position_dodge()) +
-  ggplot2::facet_wrap(~ contrast) +
-  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust = 1))
+(ref:plotFragPipebarplot) Barplot showing the partial area under the ROC at 0.1 FDR, for diff - difference among groups, scaled.p.value and t_statistic.
+These results where obtained when using FragPipeV14 combined_protein.txt as input.
+
+```{r plotFragPipebarplot, fig.cap= "(ref:plotFragPipebarplot)"}
+
+ggplot2::ggplot(ISFPprot, ggplot2::aes(x = package, y = pAUC_10)) +
+  ggplot2::geom_bar(stat = "identity") +
+  ggplot2::facet_wrap(~what)  + 
+  ggplot2::theme_minimal() + 
+  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = -90, vjust = 0.5)) +
+  ggplot2::xlab("")
 
@@
 
-# Benchmark Vignettes
+# DEA benchmark : IonStar/FragPipeV14/MSstats.tsv
 
-Pre-build version of the vignettes are available:
+The Rmarkdown file which created these results, with more details about the analysis, can be found here:
+["DEA benchmark IonStar/FragPipeV14/MSstats.tsv"](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkFragPipeMSStats.Rmd)
 
-- [Benchmarking of methods implemented in prolfqua using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkingIonstarData.html) 
-- [Benchmarking of MSstats using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_MSStats.html) 
-- [Benchmarking of proDA  using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_proDA_medpolish.html) 
-- [Benchmarking of msqrob2  using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkMSqRob2.html)  
-- [Benchmarking of proDA and proflqua using IonStar dataset MSFragger IonQuant combined_protein.txt](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkMSFraggerProteinIonStar.html)
+Since we are using the MSstats.tsv file containing precursor level abundances as input, we can benchmark the DEA performed of *msqrob2*, *MSstats*, *proflqua*, and *proda*.
+The dataset has 3899 proteins after filtering for two peptides per protein.
+Since we make four comparisons between groups, 15596 is the maximum possible number of effect size estimates.
+Supplementary Figure \@ref(fig:nrcomp) shows the number of comparisons obtained.
+Supplementary Figure \@ref(fig:plotFragPipebarplot) shows the partial area under the curve for  *proDA* and *prolfqua*, computed by ranking the proteins using the difference among groups (diff), the scaled p-value (scaled.p.value) and the t-statistics (t_statistics),
+
+```{r laodResutls, message=FALSE}
+allB <- readRDS(getpath("FragPipev14_comb_MSStats.RDS"))
+```
+
+```{r nrcomp, fig.cap= "Number of comparisons for each method for IonStar/FragPipeV14/MSstats.tsv data."}
+a1 <- allB$benchmark_proDA$smc$summary
+names(a1)[2] <- "protein_Id"
+a1$name <- allB$benchmark_proDA$model_name
+
+a2 <- allB$benchmark_prolfqua$smc$summary
+names(a2)[2] <- "protein_Id"
+a2$name <- allB$benchmark_prolfqua$model_name
+
+a3 <- allB$benchmark_msqrob$smc$summary
+names(a3)[2] <- "protein_Id"
+a3$name <- allB$benchmark_msqrob$model_name
+a4 <- allB$benchmark_msstats$smc$summary
+names(a4)[2] <- "protein_Id"
+a4$name <- allB$benchmark_msstats$model_name
+dd <- dplyr::bind_rows(list(a1,a2,a3,a4))
+dd <- dd |> dplyr::mutate(nrcontrasts = protein_Id * (4 - nr_missing))
+dds <- dd |> dplyr::group_by(name) |> dplyr::summarize(nrcontrasts = sum(nrcontrasts))
+dds$percent <- dds$nrcontrasts/max(dds$nrcontrasts) * 100
+
+nrgg <- dds |> ggplot2::ggplot(ggplot2::aes(x = name, y = nrcontrasts )) + 
+  ggplot2::geom_bar(stat = "identity", fill = "white", colour = "black") + 
+  ggplot2::coord_cartesian(ylim = c(min(dds$nrcontrasts) - 200, max(dds$nrcontrasts) + 10)) +
+  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = -90, vjust = 0.5)) +
+  ggplot2::geom_text(
+    ggplot2::aes(label = paste0(round(nrcontrasts, digits = 1),
+                                paste0("  (",round(percent, digits = 1),"%)"))),
+    vjust = 0, hjust = -0.2, angle = -90) #+ 
+nrgg
+```
+
+(ref:isFPbarplot)
+
+```{r isFPbarplot, fig.cap="(ref:isFPbarplot)"}
+proda <- allB$benchmark_proDA$pAUC_summaries()$ftable$content
+proda$package <- "proDA"
+prolfqua <- allB$benchmark_prolfqua$pAUC_summaries()$ftable$content
+prolfqua$package <- "prolfqua"
+msqrob2 <- allB$benchmark_msqrob$pAUC_summaries()$ftable$content
+msqrob2$package <- "msqrob2"
+tmp <- dplyr::bind_rows(proda, prolfqua, msqrob2)
+bmsstats <- allB$benchmark_msstats$pAUC_summaries()$ftable$content
+bmsstats$package <- "MSstats"
+bmsstats$contrast <- bmsstats$Label
+bmsstats$Label <- NULL
+ISFPmsstats <- dplyr::bind_rows(list(proda, prolfqua, msqrob2, bmsstats))
+
+ISFPmsstats$what[ISFPmsstats$what == "statistic"] <- "t_statistic"
+ISFPmsstats$what[ISFPmsstats$what == "scaled.pval"] <- "scaled.p.value"
+ISFPmsstats$what[ISFPmsstats$what == "scaled.pvalue"] <- "scaled.p.value"
+
+ISFPmsstats$what[ISFPmsstats$what == "logFC"] <- "diff"
+ISFPmsstats$what[ISFPmsstats$what == "log2FC"] <- "diff"
+ISFPmsstats$what[ISFPmsstats$what == "t"] <- "t_statistic"
+ISFPmsstats$what[ISFPmsstats$what == "Tvalue"] <- "t_statistic"
+ISFPmsstats$what |> unique()
+
+ISFPmsstats <- ISFPmsstats |> dplyr::filter(contrast == "all")
+ISFPmsstats$contrast <- NULL
+ISFPmsstats <- dplyr::relocate(ISFPmsstats, package, .before="AUC") |> dplyr::arrange(ISFPmsstats)
+
+ISFPmsstats |> 
+  knitr::kable(digits = 1, caption = "Results of the DEA beanchmark for IonStar/FragPipeV14/MSstats.tsv")
+
+ggplot2::ggplot(ISFPmsstats, ggplot2::aes(x = package, y = pAUC_10)) +
+  ggplot2::geom_bar(stat = "identity") +
+  ggplot2::facet_wrap(~what)  + 
+  ggplot2::theme_minimal() + 
+  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = -90, vjust = 0.5)) +
+  ggplot2::xlab("")
+
+```
+
+# Comparing DEA results for MaxQuant and FragPipe
+
+In the Article, we discussed other factors influencing the performance of the DEA analysis.
+One of these factors is the quantification software.
+Here we compare the benchmarking results obtained with MaxQuant and FragPipe.
+We see (Supplementary Figure \@ref(fig:compareMaxQuantFP)) that the choice of the signal processing software changes the partial area under the curve ($pAUC_{10}$) much stronger (by up to $8%$) than the DEA analysis method (by about 4%).
+Furthermore, we observe that the DEA benchmark results differ depending on if _MSstats.tsv_ or _combined_protein.tsv_ is used as input.
+
+(ref:compareMaxQuantFP) Barplot, showing the DEA results ($pAUC_{10}$ - y axis) for four DEA methods and two signal processing methods (FragPipeV14/MSstats.tsv in black, FragPipeV14/combined_protein.txt dark gray and MaxQuant/peptide.txt in gray).
+
+```{r compareMaxQuantFP, fig.cap="(ref:compareMaxQuantFP)"}
+ISFPmsstats$software <- "FragPipe_precursor"
+ISMpeptide$software <- "MaxQuant"
+ISFPprot$software <- "FragPipe_protein"
+
+tmp <- dplyr::bind_rows(list(ISFPmsstats, ISMpeptide,ISFPprot))
+
+ggplot2::ggplot(tmp, aes(x=package, y = pAUC_10, fill = software))+
+  ggplot2::geom_bar(stat = "identity",position = "dodge") +
+  scale_fill_grey() +
+  ggplot2::facet_wrap(~what)  + 
+  ggplot2::theme_minimal() + 
+  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = -90, vjust = 0.5)) +
+  ggplot2::xlab("")
+
+```
 
@@
 In each, we measured Yeast samples grown under two different conditions (Glucose and Ethanol).
-This example is taken from the `r BiocStyle::Githubpkg("fgcz/prolfqua")` package vignette file.
+This example is taken from the *prolfqua* package vignette file.
 To estimate treatment differences, we first specify the linear model that explains the observed protein abundances (transformedIntensities) using the explanatory variables `condition_` and `batch_` and the interaction term  `condition_:batch_`.
@@
 
-Given the weights $c$ contrasts from model parameters obtained with `r Biocpkg('proDA')`, or `r Biocpkg('msqrob2')` can be determined, which will enables us to implement adapters to `proDA` or `msqrob2`.
+Given the weights $c$ contrasts from model parameters obtained with *proDA*, or *msqrob2* can be estimated, which will enables us to implement adapters to *proDA* or *msqrob2*.
 
-For more details see `r BiocStyle::Githubpkg("fgcz/prolfqua")` vignette: ["Modelling dataset with two Factors"](https://fgcz.github.io/prolfqua/articles/Modelling2Factors.html).
+For more details see *prolfqua* vignette: ["Modelling dataset with two Factors"](https://fgcz.github.io/prolfqua/articles/Modelling2Factors.html).
 
@@
 
-The following code demonstrates how we use `r BiocStyle::Githubpkg("fgcz/prolfqua")` to analyze protein intensities reported in the MSFragger `combined_protein.tsv` file.
+The following code demonstrates how we use *prolfqua* to analyze protein intensities reported in the MSFragger `combined_protein.tsv` file.
 First, we create a tidy table containing the protein abundances by reading the `combined_protein.tsv` file using  `tidy_MSFragger_combined_protein.` Then, we read the sample annotation from the file `annotation.xlsx` file.

### 2022-10-30 inst/MyArticle/paper.Rmd
edits

@@
 When developing the R package `r BiocStyle::Githubpkg("fgcz/prolfqua")` we were inspired by the R package `r BiocStyle::CRANpkg("caret")` [@caret2008] which enables us to call various machine learning (ML) methods, and makes selecting the best ML algorithm for your problem easy.
-We aimed for a similar R package for the differential expression analysis of quantitative proteomics data.
+We aimed for a similar R package for the DEA of quantitative proteomics data.
 However, the existing packages differ widely regarding supported designs, model specifications, and output formats.
@@
 Peptide or protein abundances can then be scaled and transformed using robust scaling, _quantile_ normalization or `r Biocpkg("vsn")` to remove systematic differences among samples and heteroscedasticity.
-In this respect, `r BiocStyle::Githubpkg("fgcz/prolfqua")` can be compared with R packages such as `r Biocpkg("DEP")` [@DEP2018] or `r Biocpkg("POMA")` [@POMA2021] which support the entire differential expression analysis pipeline.
+In this respect, *prolfqua* can be compared with R packages such as `r Biocpkg("DEP")` [@DEP2018] or `r Biocpkg("POMA")` [@POMA2021] which support the entire
+DEA pipeline.
 
@@
 We will use the partial area under the receiver operator curve (ROC) to assess the power of the tests and compare the FDR with the false discovery proportion ($FDP$).
-We use the IonStar [@shen2018ionstar] and CPTAC[@edwards2015cptac]  datasets, processed with *MaxQuant* and *FragPipe* to benchmark the modeling methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` and to compare our results with those of `r Biocpkg("MSstats")`, `r Biocpkg("msqrob2")` and `r Biocpkg("proDA")`.
+We use the *IonStar*[@shen2018ionstar] and *CPTAC*[@edwards2015cptac]  datasets, processed with *MaxQuant* and *FragPipe* to benchmark the modeling methods implemented in *prolfqua* and to compare our results with those of *MSstats*, *msqrob2* and *proDA*.
 
-Although other benchmark datasets exist[@wessels2012comprehensive;@o2018proteome], the IonStar dataset has the advantage that the expected differences of the spike in proteins among groups are slight compared to other benchmark datasets, making DEA more difficult and enabling us to see performance differences among the methods.
-For the same reason of making the DEA more challenging and enhancing the differences between the statistical modeling approaches, we focus on the MaxQuant outputs since they are noisier than those produced by FragPipe.
+Although other benchmark datasets exist[@wessels2012comprehensive;@o2018proteome], the *IonStar* dataset has the advantage that the expected differences among groups of the spike in proteins are slight compared to other benchmark datasets, making DEA more difficult and enabling us to see performance differences.
+For the same reason, of making the DEA more challenging and enhancing the differences between the statistical modeling approaches, we focus on the *MaxQuant* outputs since they are noisier than those produced by *FragPipe*.
 
@@
 We store all the data needed for analysis in a data frame as tidy data, i.e., every column is a variable, every row is an observation, and every cell is a single value [@tidydata2014].
-Using an R6 [@R6cite] configuration object (Figure \@ref(fig:LFQData)), we specify what variable is in which column making it easy to integrate new inputs in `r BiocStyle::Githubpkg("fgcz/prolfqua")` if provided as a tidy data.
-For example, to visualize tidy Spectronaut[@bruderer2015extending], DiaNN[@demichev2020dia], or Skyline[@maclean2010skyline] outputs, or data in `r Biocpkg("MSstats")`[@Choi2014] format, only a few lines of code are needed to update the `r BiocStyle::Githubpkg("fgcz/prolfqua")` `AnalysisTableConfiguration` configuration.
+Using an R6 [@R6cite] configuration object (Figure \@ref(fig:LFQData)), we specify what variable is in which column making it easy to integrate new inputs in *prolfqua* if provided as a tidy data.
+For example, to visualize tidy Spectronaut[@bruderer2015extending], DiaNN[@demichev2020dia], or Skyline[@maclean2010skyline] outputs, or data in *MSstats*[@Choi2014] format, only a few lines of code are needed to update the *prolfqua* `AnalysisTableConfiguration` configuration.
 The configuration encapsulates the differences among the various input formats in column names and enables the use of the methods without additional arguments.
-An example code for creating an MSFragger[@yu2020fast] configuration can be found in the Appendix.
-We implemented methods that transform the data into tidy data for popular software like MaxQuant[@Cox2008MaxQuant], or MSFragger, which stores the same variable, e.g., intensity, in multiple columns, one for each sample.
+An example code for creating an *FragPipe*[@yu2020fast] configuration can be found in the Supplementary Section: "Creating a prolfqua configuration".
+We implemented methods that transform the data into tidy data for popular software like *MaxQuant*[@Cox2008MaxQuant], or *FragPipe*, which stores the same variable, e.g., intensity, in multiple columns, one for each sample.
 Relying on the tidy data table enables us to interface with many data manipulation, visualization, and modeling methods, implemented in base R [@RCoreTeam2021] and the tidyverse [@tidyverse2019], easily.
 We use R6 classes to structure the functionality of the package (see Figure \@ref(fig:LFQData) and Figure \@ref(fig:ContrastUML)).
-R6 classes are well supported by command-line completion features (see Supplementary Figure SX) in RStudio [@RStudio2022], and help to implement argument free functions.
+R6 classes are well supported by command-line completion features (see Supplementary Figure 8) in *RStudio* [@RStudio2022], and help to implement argument free functions.
 
@@
 R's formula interface for linear models is flexible, widely used, and well documented [@faraway2016extending; @law2020guide].
-We use the formula interface to specify the models, making it easy to reproduce an analysis performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` in other statistical programming languages.
+We use the formula interface to specify the models, making it easy to reproduce an analysis performed with *prolfqua* in other statistical programming languages.
 In addition, we implement features specific to high throughput experiments, such as the empirical Bayes variance and $p$-value moderation, which utilizes the parallel structure of the protein measurements and the analysis [@Ritchie2015].
 We also compute probabilities of differential protein regulation based on peptide level models [@Suomi2017bEnhanced].
-We used R6 classes to encapsulate the statistical modelling functionality in `r BiocStyle::Githubpkg("fgcz/prolfqua")` (see Figure \@ref(fig:ContrastUML)).
+We used R6 classes to encapsulate the statistical modelling functionality in *prolfqua* (see Figure \@ref(fig:ContrastUML)).
 We specify a contrast interface (`ContrastsInterface`).
-Several implementations allow to perform differential expression analysis given linear or mixed effect models (`Contrasts`), variance shrinkage (`ContrastsModerated`), or to impute contrasts in cases when observations are missing for an entire group of samples (`ContrastsSimpleImpute`).
-Further implementations of the interface encapsulate and integrate differential expression analysis results of external tools such as `r Biocpkg("proDA")` or of *SAINTexpress* [@teo2014saintexpress] used to analyze data from protein interaction studies.
+Several implementations allow to perform DEA given linear or mixed effect models (`Contrasts`), variance shrinkage (`ContrastsModerated`), or to impute contrasts in cases when observations are missing for an entire group of samples (`ContrastsSimpleImpute`).
+Further implementations of the interface encapsulate and integrate DEA results of external tools such as *proDA* or of *SAINTexpress* [@teo2014saintexpress] used to analyze data from protein interaction studies.
 
@@
     pic = c("runSaint, ContrastsSaintExpress", "protein interaction scoring (SAINTExpress)"),
-    prodA = c( "strategy_proDA*, ContrastsProDA", "adapter to the probabilistic dropout model implemented in proDA"), check.names =  FALSE)
+    proDA = c( "strategy_proDA*, ContrastsProDA", "adapter to the probabilistic dropout model implemented in proDA"), check.names =  FALSE)
 xM <- data.frame(t(xM))
@@
 
-To evaluate the performance of differential expression analysis, we use the IonStar benchmark dataset[@shen2018ionstar], available from the Proteomics Identifications Database (PRIDE) identifier PXD003881.
+To evaluate the performance of DEA, we use the *IonStar* benchmark dataset[@shen2018ionstar], available from the Proteomics Identifications Database (PRIDE) identifier PXD003881.
 $DH5\alpha$ _E.
@@
 There are four technical replicates for each dilution, hence 20 raw files in total.
-To compare the performance of the various methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` we use only the contrasts resulting in minor differences $\Delta = (1.20, 1.25, 1.30, 1.50)$, because for bigger differences all models perform similarly.
-
-However, other  have similar constraints. Furthermore, we could not analyze the benchmark dataset [@wessels2012comprehensive] with a complete factorial design with an up-to-date version of MaxQuant.
+To compare the performance of the various methods implemented in *prolfqua* we use only the contrasts resulting in minor differences $\Delta = (1.20, 1.25, 1.30, 1.50)$, because for bigger differences all models perform similarly.
 
@@
 
-We processed the raw data of the IonStar dataset using MaxQuant [@Cox2008MaxQuant] version Version 1.6.10.43, with MaxQuant default settings for orbitrap data.
-The text files generated by MaxQuant are available in the *prolfquadata* R package [@prolfquadata].
-MaxQuant produces various output files which can be used for DEA.
+We processed the raw data of the *IonStar* dataset using *MaxQuant* [@Cox2008MaxQuant] version Version 1.6.10.43, with *MaxQuant* default settings for orbitrap data.
+The text files generated by *MaxQuant* are available in the *prolfquadata* R package [@prolfquadata].
+*MaxQuant* produces various output files which can be used for DEA.
 We are using the quantification results reported in the 'peptide.txt' file for DEA.
@@
 
-We processed the raw data of the IonStar dataset using the FragPipe [@yu2020fast] Version 14, with default workflow for label free quantification with match between runs.
-The text files generated by FragPipe are available in the *prolfquadata* R package [@prolfquadata].
-Similarily to MaxQuant, the FragPipe produces various outputs which can e used for DEA.
-We used the total protein intensities reported in the `combined_protein.tsv` file as input for the DEA and call this dataset _IonStar/FragPipe/combined_protein.tsv_ Alternatively, we benchmarked the DEA using the MSstats.tsv file as input and call this dataset _IonStar/FragPipe/combined_protein.tsv_.
+We processed the raw data of the *IonStar* dataset using the *FragPipe* [@yu2020fast] Version 14, with the default workflow for label free quantification with match between runs.
+The text files generated by *FragPipe* are available in the *prolfquadata* R package [@prolfquadata].
+Similarily to *MaxQuant*, the *FragPipe* software produces various outputs which can be used for DEA.
+We used the total protein intensities reported in the 'combined_protein.tsv' file as input for the DEA and call this dataset IonStar/FragPipe/combined_protein.tsv.
+Alternatively, we benchmarked the DEA using the 'MSstats.tsv' file as input and call this dataset IonStar/FragPipe/combined_protein.tsv.
 
@@
 Protein abundances are inferred from the peptide intensities using Tukey's median polish.
-Finally, we fitted the fixed effect linear models, the dropout model `r Biocpkg("proDA")` to protein abundances, the mixed effect linear model, the `r Biocpkg("PECA")` model, and hurdle model `r Biocpkg("msqrob2")` to peptide intensities.
+Finally, we fitted the fixed effect linear models, the dropout model *proDA* to protein abundances, the mixed effect linear model, the *ROPECA* model, and hurdle model implemented in *msqrob2* to peptide intensities.
 
@@
 
-The IonStar dataset contains _H.
+The *IonStar* dataset contains _H.
 sapiens_ proteins with constant concentrations and _E.
@@
 
-The package `r BiocStyle::Githubpkg("fgcz/prolfqua")` provides functions to determine the vector of $parameter$ weights $c$, from a linear model and a contrast specification string.
+The package *prolfqua* provides functions to determine the vector of $parameter$ weights $c$, from a linear model and a contrast specification string.
 In the Supplement we provided an example, how to specify contrasts for a dataset with two explanatory variable with an interaction term (Supplementary Section "Specifying Contrasts for Models with two Factors and Interaction Term").
@@
 In such a case, we will substitute the group mean using the expected protein abundance $A$ at the LOD $A_{LOD}$.
-To estimate $A_{LOD}$ we are using the protein abundances of those groups where the protein was observed in only a single sample (see Supplement section Estimating $A_{LOD}$).
+To estimate $A_{LOD}$ we are using the protein abundances of those groups where the protein was observed in only a single sample (see Supplementary Section "Estimating $A_{LOD}$").
 Typically there are many such cases, and hence we take the median.
@@
 
-The code snippets in this section demonstrate how a differential expression analysis workflow can be implemented using the `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package.
+The code snippets in this section demonstrate how a DEA workflow can be implemented using the *prolfqua* R package.
 To speed up the computation of these examples, we use a subset of the Ionstar dataset generated by randomly selecting $400$ proteins.
@@
 R linear model and linear mixed effect models allow modeling parallel designs, repeated measurements, factorial designs, and many more.
-Models in `r BiocStyle::Githubpkg("fgcz/prolfqua")` are specified using R's linear model and mixed model formula interface.
+Models in *prolfqua* are specified using R's linear model and mixed model formula interface.
 Therefore, knowledge of the R regression model infrastructure [@faraway2016extending; @MASS2002] is advantageous when using our package.
-Furthermore, this glass box approach should make it easy to reimplement an analysis performed with `r BiocStyle::Githubpkg("fgcz/prolfqua")` using base R or other programming languages by reading the analysis script.
+Furthermore, this glass box approach should make it easy to reimplement an analysis performed with *prolfqua* using base R or other programming languages by reading the analysis script.
 However, in the package documentation, we showcase how a user, without this knowledge, can analyze experiments with a parallel group design and a factorial design.
@@
 Using the data frame of tidy data ensures interoperability with other proteomics-related packages that manage their data with tidy-tables, e.g., `r BiocStyle::CRANpkg("protti")` [@quast2022protti].
-To simplify the integration of `r BiocStyle::Githubpkg("fgcz/prolfqua")` with Bioconductor based workflows there is a method that converts the `LFQData` class into a `r Biocpkg("SummarizedExperiment")`.
+To simplify the integration of *prolfqua* with Bioconductor based workflows there is a method that converts the `LFQData` class into a `r Biocpkg("SummarizedExperiment")`.
 The use of R6 classes, which encapsulate the configuration and the data, allows for writing very concise code where functions can have few arguments.
-Auto-completion support for R6 classes in the Rstudio editor makes it easy for novices to explore `r BiocStyle::Githubpkg("fgcz/prolfqua")`'s functionality (see Supplementary Figure S[X]).
+Auto-completion support for R6 classes in the *RStudio* editor makes it easy for novices to explore *prolfqua*'s functionality (see Supplementary Figure 8).
 
-To ease the usage barriers of the R package to users not familiar with statistics and R programming, we developed an application based on the prolfqua package into our data management platform [B-Fabric](https://fgcz-bfabric.uzh.ch/) [@bfabric;@bfabricjib2022].
+To ease the usage barriers of the R package to users not familiar with statistics and R programming, we developed an application based on the *prolfqua* package into our data management platform [B-Fabric](https://fgcz-bfabric.uzh.ch/) [@bfabric;@bfabricjib2022].
 The B-Fabric system runs a computing infrastructure controlled by a local resource management system that supports cloud-bursting [@VMMAD].
@@
 The user receives a report including input files, the R markdown file, and R scripts necessary to replicate the analysis using their in-house R installation.
-In this way, `r BiocStyle::Githubpkg("fgcz/prolfqua")` and [B-Fabric](https://fgcz-bfabric.uzh.ch/) and helps scientists to meet requirements from funding agencies, journals, and academic institutions while publishing their data according to the FAIR  [@FAIR] data principles.
+In this way, *prolfqua* and [B-Fabric](https://fgcz-bfabric.uzh.ch/) and help scientists to meet requirements from funding agencies, journals, and academic institutions while publishing their data according to the FAIR  [@FAIR] data principles.
 We are working on creating a shiny stand-alone application with the described functionality and making it available soon.
@@
 
-The Benchmark functionality of `r BiocStyle::Githubpkg("fgcz/prolfqua")` includes receiver operator curves (ROC) and computes partial areas under those curves ($pAUC$) and other scores, e.g., the false discovery proportion $FDP$.
+The Benchmark functionality of *prolfqua* includes receiver operator curves (ROC) and computes partial areas under those curves ($pAUC$) and other scores, e.g., the false discovery proportion $FDP$.
 We use those scores, i.e.
-the $pAUC$ at $10%$ $FDR$ and the $FDP$, to examine how well the methods implemented in `r BiocStyle::Githubpkg("fgcz/prolfqua")` model quantitative mass spectrometric high throughput data and compare them with the results produced by *MSstats*, *proDA* and *msqrob2*.
+the $pAUC$ at $10%$ $FDR$ and the $FDP$, to examine how well the methods implemented in *prolfqua* model quantitative mass spectrometric high throughput data and compare them with the results produced by *MSstats*, *proDA* and *msqrob2*.
 Table \@ref(tab:describeBenchmarkedModels) summarizes all the modelling methods we evaluated.
@@
 
-When comparing differential expression analysis performance, a relevant parameter is the number of proteins for which a method estimates differences (see Figure \@ref(fig:figure3) Panel A), which indicates how robust the procedure works in the presence of missing observations.
+When comparing DEA performance, a relevant parameter is the number of proteins for which a method estimates differences (see Figure \@ref(fig:figure3) Panel A), which indicates how robust the procedure works in the presence of missing observations.
 For each protein, we tried to determine four differences ($\Delta = (1.20, 1.25, 1.30, 1.50)$).
@@
 We observe that the performance of the scaled $p$-values or the $t$-statistics is comparable among these three methods (Figure \@ref(fig:figure3) Panel B).
-We tested if there is a significant difference between the $pAUC_{10}$ for all three methods, but did not reject the null hypothesis that there is no such difference (Supplementary Section "Comparing pAUC for IonStar/MaxQuant dataset").
+We tested if there is a significant difference between the $pAUC_{10}$ for all three methods, but did not reject the null hypothesis that there is no such difference (Supplementary Section "DEA benchmark : IonStar/MaxQuant/peptide.txt - Significance test").
 Also, the FDR estimates (Figure \@ref(fig:figure3) Panel C) are comparable for all three methods.
-Furthermore, all three models perform similarly well when examined using CPTAC/MaxQuant dataset (see Supplementary section "Benchmarking msqrob2, proDA and prolfqua using the CPTAC dataset").
-In addition we also examined the performance of the *proDA* and `prolfqua_merged` model using the IonStar dataset but processed with the `FragPipe v14` software.
-While, for the IonStar/MaxQuant dataset, *msqrob2* performed best, *proDA* performed best for the CPTAC/MaxQuant data, and `prolfqua_merged` for the IonStar/FragPipe dataset.
+
+Furthermore, all three models perform similarly when examined using a different benchmark dataset CPTAC/MaxQuant (see Methods).
+For this dataset, *proDA* performed slightly but not significantly better than *prolfqua* and *msqrob* (and Supplementary Section "DEA benchmark : CPTAC/MaxQuant/peptide.txt").
+
+In addition, we examined the DEA performance when using protein intensities reported by quantification software *FragPipe* for the *IonStar* dataset as input.
+Using protein abundances as input significantly simplifies the analysis and interpretation and might benefit from optimization implemented in the signal processing software.
+However, we can only easily fit the *proDA* and `prolfqua_merged` models to protein intensities.
+In this DEA benchmark, *prolfqua* performed slightly, but not significantly better than *proDA*  (Supplementary Section "DEA benchmark : IonStar/FragPipeV14/combined_protein.tsv").
+
+Finally, we also compared the DEA performances when starting the analysis from the precursor abundances reported in the 'MSstats.tsv' file, generated by *FragPipe* v14, from the *IonStar* dataset.
+Since *MSstats*, *msqrob2*, *proDA*, and *prolfqua* all read `MStstats.tsv` files, we could eliminate a confounding factor, i.e., different input abundances (Supplementary Section "DEA benchmark: IonStar/FragPipeV14/MSstats.tsv").
+In this DEA benchmark, *msqrob2* and *prolfqua* perform best but not significantly better than *proDA* or *MSstats*.
+Furthermore, by processing the *IonStar* dataset with both *MaxQuant* and *FragPipe*, we can compare their performances (Supplementary Section "Comparing DEA results for *MaxQuant* and *FragPipe*").
 
@@
 For instance, the original `r Biocpkg("proDA")` publication uses MaxLFQ [@cox2014accurate] protein estimates.
-However, when using MaxLFQ intensities reported by MaxQuant, the $pAUC_{10}$ is lower ($pAUC_{10}(\textrm{t-statistics}) = 66%$) compared with results obtained when protein abundances are estimated from peptides abundances using Tukey's median polish ($pAUC_{10}(\textrm{t-statistics}) = 72%$).
+However, when using MaxLFQ intensities reported by *MaxQuant*, the $pAUC_{10}$ is lower ($pAUC_{10}(\textrm{t-statistics}) = 66%$) compared with results obtained when protein abundances are estimated from peptides abundances using Tukey's median polish ($pAUC_{10}(\textrm{t-statistics}) = 72%$).
 Last but not least, the software [@Cox2008MaxQuant; @yu2020fast] used to identify and quantify proteins significantly contributes to the entire pipeline's performance altering the number of identified proteins and the sensitivity and specificity of the differential expression analysis.
-In the Supplement we show the benchmarking results for *proDA* and *prolfqua* when the IonStar dataset is prepossessed with the FragPipe v14 [@yu2020fast] quantification pipeline.
-Although the number of proteins identified with two peptides is slightly lower with FragPipe than with MaxQuant (3929 vs.
-4046), the $pAUC_{10}$ is significantly higher ($81%$ vs.
-$72%$).
+In the Supplementary Section "Comparing DEA results for *MaxQuant* and *FragPipe* we compare DEA benchmarking results for *MaxQuant* and *FragPipe*.
+While the number of proteins identified with two peptides is practically the same, the DEA benchmark performance differs significantly by $\approx 10%$ for the $pAUC_{10}$ score.
 
@@
 `r BiocStyle::Githubpkg("fgcz/prolfqua")` is a feature-rich, object-oriented, and modular R package to analyze quantitative mass spectrometric data with simple or complex experimental designs.
-While other R packages for differential expression analysis of proteins typically only implement one modeling approach, `r BiocStyle::Githubpkg("fgcz/prolfqua")` supports various types of models (see Figure \@ref(fig:ContrastUML) and Table \@ref(tab:prolfquaModels)).
+While other R packages for differential expression analysis of proteins typically only implement one modeling approach, *prolfqua* supports various types of models (see Figure \@ref(fig:ContrastUML) and Table \@ref(tab:prolfquaModels)).
 Furthermore, the contrast specification is explicit and consistent for all models and allows for testing interactions.
-The modular design of `r BiocStyle::Githubpkg("fgcz/prolfqua")` enables adding new features, e.g., generalized linear models to model the presence or absence information of a protein, or robust linear models, in the future.
+The modular design of *prolfqua* enables adding new features, e.g., generalized linear models to model the presence or absence information of a protein, or robust linear models, in the future.
 Furthermore, the developed framework can integrate other modeling methods, e.g., the probabilistic dropout model [@bioxrvproDA2020] or accurate variance estimation [@zhu2020deqms].
-Hence, `r BiocStyle::Githubpkg("fgcz/prolfqua")` enables the implementation of applications where the user can select an alternative normalization method, protein abundance estimation methods, or differential expression analysis algorithms.
+Hence, *prolfqua* enables the implementation of applications where the user can select an alternative normalization method, protein abundance estimation methods, or DEA algorithms.
 Furthermore, this R package can analyze other types of quantitative proteomics data,e.g., label-free DIA or labeling-based TMT data.
 
-When comparing statistical modeling methods for the differential expression analysis, we assessed performance measures such as the number of estimated contrasts, the $pAUC$, and if the $FDR$ is an unbiased estimate of the $FDP$.
+When comparing statistical modeling methods for the DEA, we assessed performance measures such as the number of estimated contrasts, the $pAUC$, and if the $FDR$ is an unbiased estimate of the $FDP$.
 It is relevant that an analysis pipeline shows good performance in all these categories.
 Leveraging these computational experiments, we can provide the following advice: i) estimate protein abundances from peptide abundances using a robust or nonparametric regression method; ii) use linear models because they show good performance in all categories; iii) if the measurements are correlated, as for technical replicates, mixed effect models might work if the sample sizes are large; if not, aggregate the replicates and fit a linear model instead; iv) if you use fixed-effect linear models, apply variance moderation to improve the $t$-statistics and $p$-value estimates; v) If you want to sort your protein lists to perform gene set enrichment analysis, use the $t$-statistic instead of the difference; and last but not least vi) do not impute missing observation but statistically model missingness to estimate parameters, i.e., group differences.
-Finally, the differential expression analysis results obtained with prolfqua are comparable to those of other differential expression analysis tools.
+Finally, the DEA results obtained with *prolfqua* are comparable to those of other DEA tools.
 

### 2022-10-30 inst/MyArticle/prolfqua_supplement.Rmd
edits

@@
 ISMpeptide <-dplyr::relocate(ISMpeptide , package, .before="AUC") |> dplyr::arrange(what)
-ISMpeptide$what[ISMpeptide$what=="statistic"] <- "t_statistic"
+ISMpeptide$what[ISMpeptide$what == "statistic"] <- "t_statistic"
 ISMpeptide |> 
-  knitr::kable(digits = 1, caption = "Results of the DEA beanchmark for IonStar/MaxQuant/peptide.txt")
+  knitr::kable(
+    digits = 1, 
+    caption = "Results of the DEA beanchmark for IonStar/MaxQuant/peptide.txt",
+    format = "latex")
 ```
@@
 cap = "p-values for pairwise comparsions of partial AUC at 0.1 for msqrob2, proDA and prolfqua for IonStar MaxQuant data."
-knitr::kable(tibble::tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)), format = "latex", caption = cap)
+knitr::kable(tibble::tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)),
+             format = "latex", caption = cap)
 
@@
 all |> dplyr::relocate(package, .before = "AUC") |> dplyr::arrange(what) |> 
-  knitr::kable(digits = 1, caption = "Results of the DEA beanchmark for CPTAC/MaxQuant/peptide.txt")
+  knitr::kable(digits = 1,
+               caption = "Results of the DEA beanchmark for CPTAC/MaxQuant/peptide.txt", 
+               format = "latex")
 
@@
 cap = "p-values for pairwise comparsions of partial AUC for msqrob2, proDA and prolfqua CPTAC/MaxQuant data."
-knitr::kable(tibble::tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)), format = "latex", caption = cap)
+knitr::kable(tibble::tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)),
+             format = "latex",
+             caption = cap)
 
@@
 ISFPprot |> dplyr::relocate(package, .before = "AUC") |> dplyr::arrange(what) |>
-  knitr::kable(digits = 1, caption = "Results of the DEA beanchmark for IonStar/FragPipeV14/MSstats.tsv")
+  knitr::kable(digits = 1,
+               caption = "Results of the DEA beanchmark for IonStar/FragPipeV14/MSstats.tsv",
+               format = "latex")
 
@@
 
-(ref:plotFragPipebarplot) Barplot showing the partial area under the ROC at 0.1 FDR, for diff - difference among groups, scaled.p.value and t_statistic.
-These results where obtained when using FragPipeV14 combined_protein.txt as input.
+(ref:plotFragPipebarplot) Barplot showing the partial area under the ROC at $0.1$ FDR ($pAUC_{10}$), for diff - difference among groups, scaled.p.value and t_statistic.
+These results where obtained when using IonStar/FragPipeV14/combined_protein.txt as input.
 
@@
 cap = "p-values for pairwise comparsions of partial AUC for proDA and prolfqua CPTAC/MaxQuant data."
-knitr::kable(tibble::tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)), format = "latex", caption = cap)
+knitr::kable(tibble::tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)),
+             format = "latex",
+             caption = cap)
 
@@
 
-(ref:isFPbarplot)
+(ref:isFPbarplot) Barplot showing the partial area under the ROC at 0.1 FDR ($pAUC_{10}$), for diff - difference among groups, scaled.p.value and t_statistic.
+These results where obtained when using IonStar/FragPipeV14/MSstats.tsv as input.
 
@@
 ISFPmsstats$what[ISFPmsstats$what == "Tvalue"] <- "t_statistic"
-ISFPmsstats$what |> unique()
 
@@
 ISFPmsstats |> 
-  knitr::kable(digits = 1, caption = "Results of the DEA beanchmark for IonStar/FragPipeV14/MSstats.tsv")
+  knitr::kable(
+    digits = 1,
+    caption = "Results of the DEA beanchmark for IonStar/FragPipeV14/MSstats.tsv",
+    format = "latex")
 
@@
 
-# Comparing DEA results for MaxQuant and FragPipe
+# Comparing DEA results for *MaxQuant* and *FragPipe*
 
@@
 
-ggplot2::ggplot(tmp, aes(x=package, y = pAUC_10, fill = software))+
+ggplot2::ggplot(tmp, ggplot2::aes(x=package, y = pAUC_10, fill = software))+
   ggplot2::geom_bar(stat = "identity",position = "dodge") +
-  scale_fill_grey() +
+  ggplot2::scale_fill_grey() +
   ggplot2::facet_wrap(~what)  + 
@@
 ```{r contrcoefficient}
-knitr::kable(mC, caption = "Weights $c$ for each of the contrasts (rows) which will be applied to the model parameter (columns).",format = "latex")
+cap = "Weights $c$ for each of the contrasts (rows) which will be applied to the model parameter (columns)."
+knitr::kable(mC, caption = cap,format = "latex")
 ```

### 2022-10-30 inst/MyArticle/paper.Rmd
added test for difference between MaxQuant and FragPipe

@@
     Mass spectrometry is widely used for quantitative proteomics studies, relative protein quantification, and differential expression analysis of proteins.
-There is a large variety of signal processing software and analysis tools.
+There is a large variety of quantification software and analysis tools.
 Nevertheless, there is a need for a modular, easy-to-use application programming interface in R that transparently supports a variety of well principled statistical procedures to make applying them to proteomics data, comparing and understanding their differences easily.
@@
 The package *prolfqua* provides functions to determine the vector of $parameter$ weights $c$, from a linear model and a contrast specification string.
-In the Supplement we provided an example, how to specify contrasts for a dataset with two explanatory variable with an interaction term (Supplementary Section "Specifying Contrasts for Models with two Factors and Interaction Term").
+In the Supplement we provided an example, how to specify contrasts for a dataset with two explanatory variables and an interaction term (Supplementary Section "Specifying Contrasts for Models with two Factors and Interaction Term").
 
@@
 In addition, we examined the DEA performance when using protein intensities reported by quantification software *FragPipe* for the *IonStar* dataset as input.
-Using protein abundances as input significantly simplifies the analysis and interpretation and might benefit from optimization implemented in the signal processing software.
+Using protein abundances as input significantly simplifies the analysis and interpretation and might benefit from optimization implemented in the quantification software.
 However, we can only easily fit the *proDA* and `prolfqua_merged` models to protein intensities.
@@
 While the number of proteins identified with two peptides is practically the same, the DEA benchmark performance differs significantly by $\approx 10%$ for the $pAUC_{10}$ score.
+This difference is more significant than the differences due to the choice of the modeling method.
 
@@
 
-The authors thank the technology platform fund (TPF) of the University of Zurich and all colleagues of the FGCZ proteomics team for fruitful discussions.
+The authors thank the technology platform fund of the University of Zurich and all colleagues of the FGCZ proteomics team for fruitful discussions.
 
@@
 CDF    & Cumulative Distribution Function                   \\
+DEA & Differential Expression Analysis \\
 DIA    & data independent acquisition                        \\
@@
 $FDR$  & False Discovery Rate                               \\
+$FP$ & False Postive \\
+$FPR$ & False Positive Rage    \\
 LC     & Liquid Chromatography                              \\
@@
 MS     & mass spectrometry                                  \\
-pAUC   & partial Are Under the Curve                       \\
+pAUC   & partial Area Under the Curve                       \\
+$pAUC_{10}$ & partial Area Unter the Curve for an FPR of 0 to 10% \\
 ROC    & Receiver operator curve                           \\
+TP & True Positive \\
 TMT    & Tandem Mass Tag                                   \\

### 2022-10-30 inst/MyArticle/prolfqua_supplement.Rmd
added test for difference between MaxQuant and FragPipe

@@
 ```{r shwoROCtestREsultsIonStarMQ}
-cap = "p-values for pairwise comparsions of partial AUC at 0.1 for msqrob2, proDA and prolfqua for IonStar MaxQuant data."
+cap = "p-values for pairwise comparsions of $pAUC_{10}$ for msqrob2, proDA and prolfqua for IonStar MaxQuant data."
 knitr::kable(tibble::tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)),
@@
 ```{r shwoROCtestREsultsCPTAC}
-cap = "p-values for pairwise comparsions of partial AUC for msqrob2, proDA and prolfqua CPTAC/MaxQuant data."
+cap = "p-values for pairwise comparsions of $pAUC_{10}$ for *msqrob2*, *proDA* and *prolfqua* CPTAC/MaxQuant/peptide.txt data."
 knitr::kable(tibble::tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)),
@@
 ```{r shwoROCtestREsultsFRAGPIPE}
-cap = "p-values for pairwise comparsions of partial AUC for proDA and prolfqua CPTAC/MaxQuant data."
+cap = "p-values for pairwise comparsions of $pAUC_{10}$ for proDA and prolfqua CPTAC/MaxQuant data."
 knitr::kable(tibble::tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)),
@@
 Here we compare the benchmarking results obtained with MaxQuant and FragPipe.
-We see (Supplementary Figure \@ref(fig:compareMaxQuantFP)) that the choice of the signal processing software changes the partial area under the curve ($pAUC_{10}$) much stronger (by up to $8%$) than the DEA analysis method (by about 4%).
-Furthermore, we observe that the DEA benchmark results differ depending on if _MSstats.tsv_ or _combined_protein.tsv_ is used as input.
+We see (Supplementary Figure \@ref(fig:compareMaxQuantFP)) that the choice of the quantification software changes the partial area under the curve ($pAUC_{10}$) much stronger (by up to $8%$) than the DEA analysis method (by about 4%).
+Furthermore, we observe that the DEA benchmark results differ depending on if _MSstats.tsv_ or _combined_protein.tsv_ is used as input (see Figure \@ref(fig:compareMaxQuantFP) Left Panel, methods *proDA* and *prolfqua*).
+Supplementary Table \@ref(tab:shwoROCtestMQandFP) shows the p-values when we test if the $pAUC_{10}$, summarizing the DEA benchmark results, significantly differ between the *MaxQuant* and *FragPipe*.
+At significance level of $1%$ we will reject the null hypothesis that there is no difference between *FragPipe* and *MaxQuant*.
 
-(ref:compareMaxQuantFP) Barplot, showing the DEA results ($pAUC_{10}$ - y axis) for four DEA methods and two signal processing methods (FragPipeV14/MSstats.tsv in black, FragPipeV14/combined_protein.txt dark gray and MaxQuant/peptide.txt in gray).
+(ref:compareMaxQuantFP) Barplot, showing the DEA results ($pAUC_{10}$ - y axis) for four DEA methods and two quantification methods (FragPipeV14/MSstats.tsv in black, FragPipeV14/combined_protein.txt dark gray and MaxQuant/peptide.txt in gray).
 
@@
   ggplot2::xlab("")
+
+```
+
+```{r roctestFPMaxQuant, include = TRUE, eval = TRUE}
+allBenchmarks <- readRDS("allBenchmarks.RDS")
+allB <- readRDS(getpath("FragPipev14_comb_MSStats.RDS"))
+
+ttprolfquaMQ <- allBenchmarks$benchmark_merged$data()
+ddprolfquaMQp.val <- pROC::roc(ttprolfquaMQ$TP, ttprolfquaMQ$scaled.p.value, partial.auc = c(1, 0.9))
+ddprolfquaMQdiff <- pROC::roc(ttprolfquaMQ$TP, ttprolfquaMQ$diff, partial.auc = c(1, 0.9))
+
+ttprolfquaFP <- allB$benchmark_prolfqua$data()
+ddprolfquaFPp.val <- pROC::roc(ttprolfquaFP$TP, ttprolfquaFP$scaled.p.value, partial.auc = c(1, 0.9))
+ddprolfquaFPdiff <- pROC::roc(ttprolfquaFP$TP, ttprolfquaFP$diff, partial.auc = c(1, 0.9))
+
+tmp <- c(MQ_vs_FP_scaled.p.value = pROC::roc.test(ddprolfquaMQp.val,ddprolfquaFPp.val, progress = "none")$p.value,
+         MQ_vs_FP_diff = pROC::roc.test(ddprolfquaMQdiff, ddprolfquaFPdiff, progress = "none")$p.value)
+tmp
+```
+
+```{r shwoROCtestMQandFP}
+cap = "p-values for pairwise comparsions of $pAUC_{10}$ for MaxQuant and FragPipe based DEA analysis with prolfqua.
+We compared the pAUC for the difference estimates (diff) and for the scaled.p.values."
+knitr::kable(tibble::tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)),
+             format = "latex", caption = cap)
 

### 2022-10-31 inst/MyArticle/paper.Rmd
Changes related to the manuscript resubmission ACS JPR.

@@
 There is a large variety of quantification software and analysis tools.
-Nevertheless, there is a need for a modular, easy-to-use application programming interface in R that transparently supports a variety of well principled statistical procedures to make applying them to proteomics data, comparing and understanding their differences easily.
+Nevertheless, there is a need for a modular, easy-to-use application programming interface in *R* that transparently supports a variety of well principled statistical procedures to make applying them to proteomics data, comparing and understanding their differences easily.
 The `r BiocStyle::Githubpkg("fgcz/prolfqua")` package integrates essential steps of the mass spectrometry-based differential expression analysis workflow: quality control, data normalization, protein aggregation, statistical modeling, hypothesis testing, and sample size estimation.
@@
 The application programmer interface of `r BiocStyle::Githubpkg("fgcz/prolfqua")`  strives to be clear,  predictable, discoverable, and consistent to make proteomics data analysis application development easy and exciting.
-Finally, the `r BiocStyle::Githubpkg("fgcz/prolfqua")` R package is available on GitHub https://github.com/fgcz/prolfqua, distributed under the MIT license.
-It runs on all platforms supported by the [R free software environment for statistical computing and graphics](https://www.r-project.org/).
+Finally, the `r BiocStyle::Githubpkg("fgcz/prolfqua")` *R* package is available on GitHub https://github.com/fgcz/prolfqua, distributed under the MIT license.
+It runs on all platforms supported by the [*R* free software environment for statistical computing and graphics](https://www.r-project.org/).
 
@@
 
-\begin{quote}
-To paraphrase provocatively, `machine learning is statistics minus any checking of models and assumptions'.
-\begin{flushright}
--- Brian D. Ripley, useR! 2004, Vienna.
-\end{flushright}
-\end{quote}
-
 Proteins carry out most crucial functions and give structure to living cells.
@@
 
-Several data analysis packages exist to model mass spectrometry protein quantification experiments, e.g., `r Biocpkg("limma")` [@Ritchie2015], `r Biocpkg("MSstats")` [@Choi2014], `r Biocpkg("PECA")` [@Suomi2017bEnhanced],  `r Biocpkg("msqrob2")` [@Goeminne2016] or `r Biocpkg("proDA")` [@bioxrvproDA2020], to mention some, all implemented in R.
+Several data analysis packages exist to model mass spectrometry protein quantification experiments, e.g., `r Biocpkg("limma")` [@Ritchie2015], `r Biocpkg("MSstats")` [@Choi2014], `r Biocpkg("PECA")` [@Suomi2017bEnhanced],  `r Biocpkg("msqrob2")` [@Goeminne2016] or `r Biocpkg("proDA")` [@bioxrvproDA2020], to mention some, all implemented in *R* [@RCoreTeam2021].
 Each of them has some unique features, for example, *MSstats* determines the statistical model from the structure of the sample annotation, which allows users with limited statistical knowledge to perform differential expression analysis (DEA).
@@
 For instance, *MSstats* handles missing data by feature filtering and imputation.
-Other means of modeling missing observations are the Hurdle models discussed by @goeminne2020msqrob, while the R package *proDA* models missingness using probabilistic dropout models [@bioxrvproDA2020].
+Other means of modeling missing observations are the Hurdle models discussed by @goeminne2020msqrob, while the *R* package *proDA* models missingness using probabilistic dropout models [@bioxrvproDA2020].
 
-We can use all the R packages discussed when analyzing parallel-group designs using a single explanatory variable and contrasting groups.
+We can use all the *R* packages discussed when analyzing parallel-group designs using a single explanatory variable and contrasting groups.
 However, we can use only some of them if we want to model factorial designs or repeated measurements.
 Table \@ref(tab:tableOverview) gives an overview of the models and features supported by these packages.
-We see that, for instance, *limma* and *proDA* allow us to fit a comprehensive variety of models and test various hypotheses; however, good knowledge of the design matrix specification using the R formula interface is required [@law2020guide].
+We see that, for instance, *limma* and *proDA* allow us to fit a comprehensive variety of models and test various hypotheses; however, good knowledge of the design matrix specification using the *R* formula interface is required [@law2020guide].
 
@@
 
-When developing the R package `r BiocStyle::Githubpkg("fgcz/prolfqua")` we were inspired by the R package `r BiocStyle::CRANpkg("caret")` [@caret2008] which enables us to call various machine learning (ML) methods, and makes selecting the best ML algorithm for your problem easy.
-We aimed for a similar R package for the DEA of quantitative proteomics data.
+When developing the *R* package `r BiocStyle::Githubpkg("fgcz/prolfqua")` we were inspired by the *R* package `r BiocStyle::CRANpkg("caret")` [@caret2008] which enables us to call various machine learning (ML) methods, and makes selecting the best ML algorithm for your problem easy.
+We aimed for a similar *R* package for the DEA of quantitative proteomics data.
 However, the existing packages differ widely regarding supported designs, model specifications, and output formats.
 At the same time, they have the following features in common: fitting linear models to either peptide or protein intensities, determining differences among groups, and afterward applying empirical Bayes variance shrinkage.
-Therefore, the revised goal was to provide a modular object-oriented design, with R linear models as a core, and add features such as $p$-value aggregation, variance shrinkage, or modeling of missing observations.
+Therefore, the revised goal was to provide a modular object-oriented design, with *R* linear models as a core, and add features such as $p$-value aggregation, variance shrinkage, or modeling of missing observations.
 
@@
 Peptide or protein abundances can then be scaled and transformed using robust scaling, _quantile_ normalization or `r Biocpkg("vsn")` to remove systematic differences among samples and heteroscedasticity.
-In this respect, *prolfqua* can be compared with R packages such as `r Biocpkg("DEP")` [@DEP2018] or `r Biocpkg("POMA")` [@POMA2021] which support the entire
+In this respect, *prolfqua* can be compared with *R* packages such as `r Biocpkg("DEP")` [@DEP2018] or `r Biocpkg("POMA")` [@POMA2021] which support the entire
 DEA pipeline.
@@
 We use the *IonStar*[@shen2018ionstar] and *CPTAC*[@edwards2015cptac]  datasets, processed with *MaxQuant* and *FragPipe* to benchmark the modeling methods implemented in *prolfqua* and to compare our results with those of *MSstats*, *msqrob2* and *proDA*.
-
-Although other benchmark datasets exist[@wessels2012comprehensive;@o2018proteome], the *IonStar* dataset has the advantage that the expected differences among groups of the spike in proteins are slight compared to other benchmark datasets, making DEA more difficult and enabling us to see performance differences.
-For the same reason, of making the DEA more challenging and enhancing the differences between the statistical modeling approaches, we focus on the *MaxQuant* outputs since they are noisier than those produced by *FragPipe*.
+Although other benchmark datasets exist[@wessels2012comprehensive;@o2018proteome], the *IonStar* dataset has the advantage that the expected differences, for the spike in proteins, among groups are small compared to other benchmark datasets, making DEA more difficult and enabling us to see performance differences among the modeling methods.
 
@@
 The configuration encapsulates the differences among the various input formats in column names and enables the use of the methods without additional arguments.
-An example code for creating an *FragPipe*[@yu2020fast] configuration can be found in the Supplementary Section: "Creating a prolfqua configuration".
+An example code for creating an *FragPipe*[@yu2020fast] configuration can be found in the Supplementary Section "Creating a prolfqua configuration".
 We implemented methods that transform the data into tidy data for popular software like *MaxQuant*[@Cox2008MaxQuant], or *FragPipe*, which stores the same variable, e.g., intensity, in multiple columns, one for each sample.
-Relying on the tidy data table enables us to interface with many data manipulation, visualization, and modeling methods, implemented in base R [@RCoreTeam2021] and the tidyverse [@tidyverse2019], easily.
+Relying on the tidy data table enables us to interface with many data manipulation, visualization, and modeling methods, implemented in base *R* [@RCoreTeam2021] and the tidyverse [@tidyverse2019], easily.
 We use R6 classes to structure the functionality of the package (see Figure \@ref(fig:LFQData) and Figure \@ref(fig:ContrastUML)).
@@
 The `LFQDataPlotter` class and other classes decorate the `LFQData` class with additional functionality.
-Similarly, the `LFQDataStats` and `LFQDataSummary` reference the `LFQData` class and group methods for variance and sample size estimation or summarizing peptide and protein counts.
+For instance, the `LFQDataStats` and `LFQDataSummary` reference the `LFQData` class and group methods for variance and sample size estimation or summarizing peptide and protein counts.
+Furthermore, the `LFQDataTransformer` and `LFQDataAggregator` classes group functions for data normalization and for estimating protein from peptide intensities.
 
@@
 
-R's formula interface for linear models is flexible, widely used, and well documented [@faraway2016extending; @law2020guide].
+*R*'s formula interface for linear models is flexible, widely used, and well documented [@faraway2016extending; @law2020guide].
 We use the formula interface to specify the models, making it easy to reproduce an analysis performed with *prolfqua* in other statistical programming languages.
@@
 We annotated these dilutions from smallest to largest with the letters _a_, _b_, _c_, _d_ and _e_.
-By comparing the various dilutions, we obtain different effect sizes, e.g., when comparing dilution _e_ (9%) against dilution _d_ (7.5%), the expected difference is $1.2$ for E.
+By comparing the various dilutions, we obtain different effect sizes, e.g., when comparing dilution _e_ ($9\%$) against dilution _d_ ($7.5\%$), the expected difference is $1.2$ for E.
 coli proteins and $1$ for human proteins.
@@
 We processed the raw data of the *IonStar* dataset using *MaxQuant* [@Cox2008MaxQuant] version Version 1.6.10.43, with *MaxQuant* default settings for orbitrap data.
-The text files generated by *MaxQuant* are available in the *prolfquadata* R package [@prolfquadata].
+The text files generated by *MaxQuant* are available in the *prolfquadata* *R* package [@prolfquadata].
 *MaxQuant* produces various output files which can be used for DEA.
@@
 We processed the raw data of the *IonStar* dataset using the *FragPipe* [@yu2020fast] Version 14, with the default workflow for label free quantification with match between runs.
-The text files generated by *FragPipe* are available in the *prolfquadata* R package [@prolfquadata].
+The text files generated by *FragPipe* are available in the *prolfquadata* *R* package [@prolfquadata].
 Similarily to *MaxQuant*, the *FragPipe* software produces various outputs which can be used for DEA.
@@
 
-We used the CPTAC dataset, available in the R package *msdata*, and described in [@Goeminne2016].
+We used the CPTAC dataset, available in the *R* package *msdata*, and described in [@Goeminne2016].
 In brief, the Sigma Universal Protein Standard mixture 1 (UPS1) containing $48$ different human proteins was spiked in a protein background of 60 ng/$\mu L$ Saccharomyces cerevisiae strain BY4741.
@@
 The degrees of freedom for the contrast are equal to the residual degrees of freedom of the linear model.
-For estimating contrasts from mixed effects models we used the function `contest` implemented in the R package `r BiocStyle::CRANpkg("lmerTest")`  and used the Satterthwaite [@Kuznetsova2017lmerTest] method to estimate the denominator degrees of freedom.
+For estimating contrasts from mixed effects models we used the function `contest` implemented in the *R* package `r BiocStyle::CRANpkg("lmerTest")`  and used the Satterthwaite [@Kuznetsova2017lmerTest] method to estimate the denominator degrees of freedom.
 These methods are available in the class `Contrast` (see Figure \@ref(fig:ContrastUML))
@@
 
-# Results and Discussion.
+# Results and Discussion
 
@@
 
-The code snippets in this section demonstrate how a DEA workflow can be implemented using the *prolfqua* R package.
+The code snippets in this section demonstrate how a DEA workflow can be implemented using the *prolfqua* *R* package.
 To speed up the computation of these examples, we use a subset of the Ionstar dataset generated by randomly selecting $400$ proteins.
@@
 
-(ref:figure2) Panel A - Histogram showing the distribution of p-values for $332$ proteins.
+(ref:figure2) Panel A - Histogram showing the distribution of p-values for $332$ proteins for contrasts "e_vs_d" and "d_vs_c".
 Panel B -  Volcano plot showing $-\log_{10}$ transformed $FDR$ as a function of the difference between groups for $332$ proteins.
 With black dots, we show effect size and FDR estimates obtained from the linear model, while in green, we plot those obtained using imputation.
-Finally, panel C shows the difference between groups as a function of the rank of the abundance of the proteins.
+Finally, panel C shows the difference between groups, as a function of the rank of the abundance of the proteins.
 
@@
 contrasts <- c(
-  "dilution_(9/7.5)_1.2" =   "dilution.e - dilution.d",
-  "dilution_(7.5/6)_1.25" =   "dilution.d - dilution.c"
+  "e_vs_d" =   "dilution.e - dilution.d",
+  "d_vs_c" =   "dilution.d - dilution.c"
 )
@@
 
-R linear model and linear mixed effect models allow modeling parallel designs, repeated measurements, factorial designs, and many more.
-Models in *prolfqua* are specified using R's linear model and mixed model formula interface.
-Therefore, knowledge of the R regression model infrastructure [@faraway2016extending; @MASS2002] is advantageous when using our package.
-Furthermore, this glass box approach should make it easy to reimplement an analysis performed with *prolfqua* using base R or other programming languages by reading the analysis script.
+*R* linear model and linear mixed effect models allow modeling parallel designs, repeated measurements, factorial designs, and many more.
+Models in *prolfqua* are specified using *R*'s linear model and mixed model formula interface.
+Therefore, knowledge of the *R* regression model infrastructure [@faraway2016extending; @MASS2002] is advantageous when using our package.
+Furthermore, this glass box approach should make it easy to reimplement an analysis performed with *prolfqua* using base *R* or other programming languages by reading the analysis script.
 However, in the package documentation, we showcase how a user, without this knowledge, can analyze experiments with a parallel group design and a factorial design.
@@
 
-To ease the usage barriers of the R package to users not familiar with statistics and R programming, we developed an application based on the *prolfqua* package into our data management platform [B-Fabric](https://fgcz-bfabric.uzh.ch/) [@bfabric;@bfabricjib2022].
+To ease the usage barriers of the *R* package to users not familiar with statistics and *R* programming, we developed an application based on the *prolfqua* package into our data management platform [B-Fabric](https://fgcz-bfabric.uzh.ch/) [@bfabric;@bfabricjib2022].
 The B-Fabric system runs a computing infrastructure controlled by a local resource management system that supports cloud-bursting [@VMMAD].
@@
 However, this application can analyze only parallel group designs with and without repeated measurements or factorial designs without interactions.
-The user receives a report including input files, the R markdown file, and R scripts necessary to replicate the analysis using their in-house R installation.
-In this way, *prolfqua* and [B-Fabric](https://fgcz-bfabric.uzh.ch/) and help scientists to meet requirements from funding agencies, journals, and academic institutions while publishing their data according to the FAIR  [@FAIR] data principles.
+The user receives a report including input files, the *R* markdown file, and *R* scripts necessary to replicate the analysis using their in-house *R* installation.
+In this way, *prolfqua* and [B-Fabric](https://fgcz-bfabric.uzh.ch/) help scientists to meet requirements from funding agencies, journals, and academic institutions while publishing their data according to the FAIR  [@FAIR] data principles.
 We are working on creating a shiny stand-alone application with the described functionality and making it available soon.
@@
 We use those scores, i.e.
-the $pAUC$ at $10%$ $FDR$ and the $FDP$, to examine how well the methods implemented in *prolfqua* model quantitative mass spectrometric high throughput data and compare them with the results produced by *MSstats*, *proDA* and *msqrob2*.
+the $pAUC$ at $10\%$ $FDR$ and the $FDP$, to examine how well the methods implemented in *prolfqua* model quantitative mass spectrometric high throughput data and compare them with the results produced by *MSstats*, *proDA* and *msqrob2*.
 Table \@ref(tab:describeBenchmarkedModels) summarizes all the modelling methods we evaluated.
@@
 
-Figure \@ref(fig:figure3) Panel B shows how various estimates obtained from the models, i.e., the difference between groups, $t$-statistics and scaled $p$-values, allow identifying true positives (TP) given a false positive rate (FPR) of $10%$ by displaying the partial area under the ROC ($pAUC_{10}$).
+Figure \@ref(fig:figure3) Panel B shows how various estimates obtained from the models, i.e., the difference between groups, $t$-statistics and scaled $p$-values, allow identifying true positives (TP) given a false positive rate (FPR) of $10\%$ by displaying the partial area under the ROC ($pAUC_{10}$).
 Ordering proteins by the $t$-statistic or $p$-value leads to a higher $pAUC_{10}$ than when ordering by the estimated difference among groups.
@@
 Table \@ref(tab:describeBenchmarkedModels) summarizes which methods we have evaluated, which *MaxQuant* files we used as input, and if the models are fitted to peptide or protein intensities.
-We make the R-markdown files to replicate the benchmarking available at `r BiocStyle::Githubpkg("wolski/prolfquabenchmark")`.
+We make the *R*-markdown files to replicate the benchmarking available at `r BiocStyle::Githubpkg("wolski/prolfquabenchmark")`.
 
@@
 
-The R-package *proDA*, *msqrob2* and *prolfqua* do not impute missing data but integrate them into the statistical model, while *MSstats* filters and imputes the data using an accelerated failure model.
+The *R*-package *proDA*, *msqrob2* and *prolfqua* do not impute missing data but integrate them into the statistical model, while *MSstats* filters and imputes the data using an accelerated failure model.
 Despite imputation, *MSstats* estimates group differences for less proteins, and does not achieve a higher $pAUC_{10}$.
@@
 
-We obtained difference and FDR estimates for all proteins, as shown in Figure \@ref(fig:figure3) Panel A when using: a) the probabilistic dropout model (*proDA*), b) the hurdle model (*msqrob2*), and c) `prolfqua_merge` (Table \@ref{tab:describeBenchmarkedModels}).
+We obtained difference and FDR estimates for all proteins, as shown in Figure \@ref(fig:figure3) Panel A when using: a) the probabilistic dropout model (*proDA*), b) the hurdle model (*msqrob2*), and c) 'prolfqua_merge' (Table \@ref(tab:describeBenchmarkedModels).
 We observe that the performance of the scaled $p$-values or the $t$-statistics is comparable among these three methods (Figure \@ref(fig:figure3) Panel B).
@@
 Using protein abundances as input significantly simplifies the analysis and interpretation and might benefit from optimization implemented in the quantification software.
-However, we can only easily fit the *proDA* and `prolfqua_merged` models to protein intensities.
+However, we can only easily fit the *proDA* and 'prolfqua_merged' (see Table \@ref(tab:describeBenchmarkedModels)) models to protein intensities.
 In this DEA benchmark, *prolfqua* performed slightly, but not significantly better than *proDA* (Supplementary Section "DEA benchmark : IonStar/FragPipeV14/combined_protein.tsv").
@@
 Since *MSstats*, *msqrob2*, *proDA*, and *prolfqua* all read `MStstats.tsv` files, we could eliminate a confounding factor, i.e., different input abundances (Supplementary Section "DEA benchmark: IonStar/FragPipeV14/MSstats.tsv").
-In this DEA benchmark, *msqrob2* and *prolfqua* perform best but not significantly better than *proDA* or *MSstats*.
+In this DEA benchmark, *msqrob2* and *prolfqua* ('prolfqua_merged') perform best but not significantly better than *proDA* or *MSstats*.
 Furthermore, by processing the *IonStar* dataset with both *MaxQuant* and *FragPipe*, we can compare their performances (Supplementary Section "Comparing DEA results for *MaxQuant* and *FragPipe*").
@@
 The method used to infer proteins from peptide identifications [@nesvizhskii2005interpretation] and protein abundances from peptide abundances is an additional important factor [@Grossmann2010].
-For instance, the original `r Biocpkg("proDA")` publication uses MaxLFQ [@cox2014accurate] protein estimates.
-However, when using MaxLFQ intensities reported by *MaxQuant*, the $pAUC_{10}$ is lower ($pAUC_{10}(\textrm{t-statistics}) = 66%$) compared with results obtained when protein abundances are estimated from peptides abundances using Tukey's median polish ($pAUC_{10}(\textrm{t-statistics}) = 72%$).
+For instance, the original *proDA* publication uses MaxLFQ [@cox2014accurate] protein estimates.
+However, when using MaxLFQ intensities reported by *MaxQuant*, the $pAUC_{10}$ is lower ($pAUC_{10}(\textrm{t-statistics}) = 66\%$) compared with results obtained when protein abundances are estimated from peptide abundances using Tukey's median polish ($pAUC_{10}(\textrm{t-statistics}) = 72\%$).
 Last but not least, the software [@Cox2008MaxQuant; @yu2020fast] used to identify and quantify proteins significantly contributes to the entire pipeline's performance altering the number of identified proteins and the sensitivity and specificity of the differential expression analysis.
-In the Supplementary Section "Comparing DEA results for *MaxQuant* and *FragPipe* we compare DEA benchmarking results for *MaxQuant* and *FragPipe*.
-While the number of proteins identified with two peptides is practically the same, the DEA benchmark performance differs significantly by $\approx 10%$ for the $pAUC_{10}$ score.
+In the Supplementary Section "Comparing DEA results for *MaxQuant* and *FragPipe*" we compare DEA benchmarking results for the quantification software.
+While the number of proteins identified with two peptides is practically the same, the DEA benchmark performance differs significantly by $\approx 10\%$ for the $pAUC_{10}$ score.
 This difference is more significant than the differences due to the choice of the modeling method.
@@
 
-`r BiocStyle::Githubpkg("fgcz/prolfqua")` is a feature-rich, object-oriented, and modular R package to analyze quantitative mass spectrometric data with simple or complex experimental designs.
-While other R packages for differential expression analysis of proteins typically only implement one modeling approach, *prolfqua* supports various types of models (see Figure \@ref(fig:ContrastUML) and Table \@ref(tab:prolfquaModels)).
+`r BiocStyle::Githubpkg("fgcz/prolfqua")` is a feature-rich, object-oriented, and modular *R* package to analyze quantitative mass spectrometric data with simple or complex experimental designs.
+While other *R* packages for differential expression analysis of proteins typically only implement one modeling approach, *prolfqua* supports various types of models (see Figure \@ref(fig:ContrastUML) and Table \@ref(tab:prolfquaModels)).
 Furthermore, the contrast specification is explicit and consistent for all models and allows for testing interactions.
@@
 Hence, *prolfqua* enables the implementation of applications where the user can select an alternative normalization method, protein abundance estimation methods, or DEA algorithms.
-Furthermore, this R package can analyze other types of quantitative proteomics data,e.g., label-free DIA or labeling-based TMT data.
+Furthermore, this *R* package can analyze other types of quantitative proteomics data, e.g., label-free DIA or labeling-based TMT data.
 
@@
 pAUC   & partial Area Under the Curve                       \\
-$pAUC_{10}$ & partial Area Unter the Curve for an FPR of 0 to 10% \\
-ROC    & Receiver operator curve                           \\
+$pAUC_{10}$ & partial Area Under the Curve for an FPR of 0 to $10\%$ \\
+ROC    & Receiver Operator Curve                           \\
 TP & True Positive \\

### 2022-10-31 inst/MyArticle/prolfqua_supplement.Rmd
Changes related to the manuscript resubmission ACS JPR.

@@
 
-The sourcode used to produce this Document can be found in the  [prolfqua_supplement.Rmd](https://github.com/wolski/prolfquabenchmark/blob/main/inst/MyArticle/prolfqua_supplement.Rmd) file.
+The *R* markdown file, with all the *R* code used to produce this pdf-document can be found here :   [prolfqua_supplement.Rmd](https://github.com/wolski/prolfquabenchmark/blob/main/inst/MyArticle/prolfqua_supplement.Rmd).
+To replicated the analysis you need to install the *prolfquabenchmark* *R* package.
 
@@
 
-To install the *prolfqua* package ([ACS_JPR](https://github.com/fgcz/prolfqua/releases/tag/v0.4.3) release), including all R package dependencies, execute in the R console:
+To install the *prolfqua* package ([ACS_JPR](https://github.com/fgcz/prolfqua/releases/tag/v0.4.3) release), including all *R* package dependencies, execute in the *R* console:
 
 ```{r installation, echo=TRUE, eval=FALSE}
-remotes::install_github("https://github.com/fgcz/prolfqua/releases/tag/v0.4.5",
+remotes::install_github("https://github.com/fgcz/prolfqua/releases/tag/v1.0.0",
     dependencies = TRUE, build_vignettes=TRUE)
@@
 
-The R-markdown files with all the R code to run the benchmark for *prolfqua*, *msqrob2*, *MSStats* and *proDA* are available in the R package `r BiocStyle::Githubpkg('wolski/prolfquabenchmark')`.
+The R-markdown files with all the *R* code to run the benchmark for *prolfqua*, *msqrob2*, *MSStats* and *proDA* are available in the R package `r BiocStyle::Githubpkg('wolski/prolfquabenchmark')`.
 
@@
 remotes::install_gitlab("wolski/prolfquadata", host="gitlab.bfabric.org")
-remotes::install_github("https://github.com/wolski/prolfquabenchmark",
+remotes::install_github("https://github.com/wolski/prolfquabenchmark/releases/tag/v0.2.0",
     dependencies = TRUE, build_vignettes=TRUE)
@@
 We test if we should reject the null hypothesis that the area under the ROC curve at $0.1$ FDR ($pAUC_{10}$), computed using the scaled p-values (see scaled.p.value's in Table \@ref(tab:ISMQpeptideSum) ), do not differ for *msqrob2*, *proDA*, and `prolfqua_merged`.
-To this task we use the function `roc.test`, that performs a bootstrap method test for a difference in the pAUC of the ROC curves, from the R package `r BiocStyle::CRANpkg("pROC")`.
+To this task we use the function `roc.test`, that performs a bootstrap method test for a difference in the pAUC of the ROC curves, from the *R* package `r BiocStyle::CRANpkg("pROC")`.
 The Table \@ref(tab:shwoROCtestREsultsIonStarMQ) shows the p-values of the pairwise bootstrap test for partial areas under the ROC curves.
@@
 ```{r shwoROCtestREsultsIonStarMQ}
-cap = "p-values for pairwise comparsions of $pAUC_{10}$ for msqrob2, proDA and prolfqua for IonStar MaxQuant data."
+cap = "p-values for pairwise comparsions of $pAUC_{10}$, for *msqrob2*, *proDA* and *prolfqua* for IonStar MaxQuant data."
 knitr::kable(tibble::tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)),
@@
 
-We applied DEA using *msqrob2*, *proDA* and *prolfqua*  to the CPTAC/MaxQuant dataset and benchmarked the results. We are using the `peptide.txt` file containing the peptide intensities.
+We applied DEA using *msqrob2*, *proDA* and *prolfqua* ('prolfqua_merged')  to the CPTAC/MaxQuant dataset and benchmarked the results.
+We are using the 'peptide.txt' file containing the peptide intensities.
 
@@
 ```{r shwoROCtestREsultsCPTAC}
-cap = "p-values for pairwise comparsions of $pAUC_{10}$ for *msqrob2*, *proDA* and *prolfqua* CPTAC/MaxQuant/peptide.txt data."
+cap = "p-values for pairwise comparsions of $pAUC_{10}$, for *msqrob2*, *proDA* and *prolfqua* CPTAC/MaxQuant/peptide.txt data."
 knitr::kable(tibble::tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)),
@@
 After filtering for two peptides per protein the dataset comprises of $3836$ proteins.
-Since we are using the combined_protein.tsv as input, comparison with neither *msqrob2* nor *MSstats* is possible.
-To fit the hurdle (*msqrob*) model, peptide intensities are required, while *MSstats* requires the `MSstats.tsv` file, containing precursor level abundances.
+Since we are using the 'combined_protein.tsv' as input, comparison with neither *msqrob2* nor *MSstats* is possible.
+To fit the hurdle (*msqrob*) model, peptide intensities are required, while *MSstats* requires the 'MSstats.tsv' file, containing precursor level abundances.
 However, starting from protein level intensities has the advantage that the effect size estimates obtained from the model agree with protein abundances, a feature frequently requested by the biologist.
@@
 Table \@ref(tab:shwoROCtestREsultsFRAGPIPE) shows the results of the Bootstrap test for the difference of two ROC curves.
-We observe that for this benchmark data, there are no significant differences among the $pAUC_{10}$ (for scaled p-value) for the two packages.
+We observe that for this benchmark, there is no significant differences (at significance level of $0.1$) among the $pAUC_{10}$, where the scaled p-value was use to rank the proteins, for the two packages.
 
 ```{r shwoROCtestREsultsFRAGPIPE}
-cap = "p-values for pairwise comparsions of $pAUC_{10}$ for proDA and prolfqua CPTAC/MaxQuant data."
+cap = "p-values for pairwise comparsions of $pAUC_{10}$, for *proDA* and *prolfqua* CPTAC/MaxQuant data."
 knitr::kable(tibble::tibble(pROC_test = names(tmp), p.value = round(tmp, digits = 2)),
@@
 The Rmarkdown file which created these results, with more details about the analysis, can be found here:
-["DEA benchmark IonStar/FragPipeV14/MSstats.tsv"](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkFragPipeMSStats.Rmd)
+["DEA benchmark IonStar/FragPipeV14/MSstats.tsv"](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkFragPipeMSStats.html)
 
-Since we are using the MSstats.tsv file containing precursor level abundances as input, we can benchmark the DEA performed of *msqrob2*, *MSstats*, *proflqua*, and *proda*.
+Since we are using the 'MSstats.tsv' file containing precursor level abundances as input, we can benchmark the DEA performed of *msqrob2*, *MSstats*, *proflqua*, and *proDA*.
 The dataset has 3899 proteins after filtering for two peptides per protein.
@@
 Here we compare the benchmarking results obtained with MaxQuant and FragPipe.
-We see (Supplementary Figure \@ref(fig:compareMaxQuantFP)) that the choice of the quantification software changes the partial area under the curve ($pAUC_{10}$) much stronger (by up to $8%$) than the DEA analysis method (by about 4%).
-Furthermore, we observe that the DEA benchmark results differ depending on if _MSstats.tsv_ or _combined_protein.tsv_ is used as input (see Figure \@ref(fig:compareMaxQuantFP) Left Panel, methods *proDA* and *prolfqua*).
+We see (Supplementary Figure \@ref(fig:compareMaxQuantFP)) that the choice of the quantification software changes the partial area under the curve ($pAUC_{10}$) much stronger (by up to $8\%$) than the DEA analysis method (by about $4\%$).
+Furthermore, we observe that the DEA benchmark results differ depending on if 'MSstats.tsv' or 'combined_protein.tsv' is used as input (see Figure \@ref(fig:compareMaxQuantFP) Left Panel, methods *proDA* and *prolfqua*).
 Supplementary Table \@ref(tab:shwoROCtestMQandFP) shows the p-values when we test if the $pAUC_{10}$, summarizing the DEA benchmark results, significantly differ between the *MaxQuant* and *FragPipe*.
-At significance level of $1%$ we will reject the null hypothesis that there is no difference between *FragPipe* and *MaxQuant*.
+At significance level of $0.01$ we will reject the null hypothesis that there is no difference between *FragPipe* and *MaxQuant*.
 
@@
          MQ_vs_FP_diff = pROC::roc.test(ddprolfquaMQdiff, ddprolfquaFPdiff, progress = "none")$p.value)
-tmp
 ```
@@
 In each, we measured Yeast samples grown under two different conditions (Glucose and Ethanol).
-This example is taken from the *prolfqua* package vignette file.
+
 To estimate treatment differences, we first specify the linear model that explains the observed protein abundances (transformedIntensities) using the explanatory variables `condition_` and `batch_` and the interaction term  `condition_:batch_`.
@@
 
-The following code demonstrates how we use *prolfqua* to analyze protein intensities reported in the MSFragger `combined_protein.tsv` file.
-First, we create a tidy table containing the protein abundances by reading the `combined_protein.tsv` file using  `tidy_MSFragger_combined_protein.` Then, we read the sample annotation from the file `annotation.xlsx` file.
+The following code demonstrates how we use *prolfqua* to analyze protein intensities reported in the *FragPipe* 'combined_protein.tsv' file.
+First, we create a tidy table containing the protein abundances by reading the `combined_protein.tsv` file using  `tidy_FragPipe_combined_protein.` Then, we read the sample annotation from the file `annotation.xlsx` file.
 Next, we create an `AnalysisTableAnnotation` R6 object.

### 2022-10-31 inst/MyArticle/createUploadFiles.R

@@ -1,0 +1,33 @@
+
+tex <- readLines("paper.tex")
+
+abstract <- tex[ grep("begin\\{abstract\\}", tex):(grep("end\\{abstract\\}", tex))]
+writeLines(abstract, "../../../prolfqua-achems/abstract.tex")
+
+introduction <- tex[ grep("hypertarget\\{introduction\\}", tex):(grep("hypertarget\\{methods\\}", tex) - 1)]
+writeLines(introduction, "../../../prolfqua-achems/introduction.tex")
+
+methods <- tex[ grep("hypertarget\\{methods\\}", tex):(grep("hypertarget\\{results-and-discussion\\}", tex) - 1)]
+writeLines(methods, "../../../prolfqua-achems/methods.tex")
+
+# create results.tex
+results <- tex[ grep("hypertarget\\{results-and-discussion\\}", tex):(grep("hypertarget\\{conclusion\\}", tex) - 1)]
+
+sB <- grep("begin\\{Shaded\\}", results)
+sE <- grep("end\\{Shaded\\}", results)
+
+results <- results[-(sB[1]:sE[2])]
+results <- append(results, "\\TODO{codesnippets/codesnippet2.pdf}", after = (sB[1] - 1))
+
+sB <- grep("begin\\{Shaded\\}", results)
+sE <- grep("end\\{Shaded\\}", results)
+results <- results[-(sB[1]:sE[1])]
+results <- append(results, "\\TODO{codesnippets/codesnippet3.pdf}", after = (sB[1] - 1))
+
+writeLines(results, "../../../prolfqua-achems/results.tex")
+
+conclusion <- tex[ grep("hypertarget\\{conclusion\\}", tex):(grep("hypertarget\\{acknowledgements\\}", tex) - 1)]
+writeLines(conclusion, "../../../prolfqua-achems/conclusion.tex")
+
+abbreviations <- tex[ grep("hypertarget\\{abbreviations\\}", tex):(grep("hypertarget\\{references\\}", tex) - 1)]
+writeLines(abbreviations, "../../../prolfqua-achems/abbreviations.tex")

### 2023-01-09 inst/MyArticle/paper.Rmd
adress comment regarding contrasts from tables by removing sentence and merging paragraphs.

@@
 This integration enables users to select the input data and basic settings in a graphical user interface (GUI).
-
-In this GUI application, we determine the model formula, and which comparisons to compute from the sample annotation provided in tabular form, similarly to *SAINTexpress* or *MSstats*.
-However, this application can analyze only parallel group designs with and without repeated measurements or factorial designs without interactions.
-The user receives a report including input files, the *R* markdown file, and *R* scripts necessary to replicate the analysis using their in-house *R* installation.
 In this way, *prolfqua* and [B-Fabric](https://fgcz-bfabric.uzh.ch/) help scientists to meet requirements from funding agencies, journals, and academic institutions while publishing their data according to the FAIR  [@FAIR] data principles.
@@
     proDA = c("proDA", "probabilistic dropout model",  "protein", "peptide.txt"),
-    imputation = c("prolfqua_impute", "ContrastsSimpleImpute, ContrastsModerated", "protein",  "peptide.txt"),
+    imputation = c("prolfqua_estimate", "ContrastsSimpleImpute, ContrastsModerated", "protein",  "peptide.txt"),
     lmmed = c("prolfqua_lm_mod", "strategy_lm, Contrasts, ContrastsModerated",  "protein", "peptide.txt"),
-    prot_merged = c("prolfqua_merged", "addContrastResults( prefer = prot_med_lm_moderated , add = prot_imputation)HACKIT ",  "protein", "peptide.txt"),
+    prot_merged = c("prolfqua_merged", "addContrastResults( prefer = proflqua_lm_mod , add = prolfqua_estimate)HACKIT ",  "protein", "peptide.txt"),
     mixed = c("prolfqua_mix_eff_mod", "strategy_lmer, Contrasts, ContrastsModerated", "peptide", "peptide.txt" ),

### 2023-01-09 inst/MyArticle/SIB.Rmd
rename ContrastSimpleImpute to ContrastMissing #28

@@
 conI <- 
-    prolfqua::ContrastsSimpleImpute$new(lfqp, contrasts) |> 
+    prolfqua::ContrastsMissing$new(lfqp, contrasts) |> 
     prolfqua::ContrastsModerated$new()
@@
     proDA = c("proDA", "probabilistic dropout model",  "protein", "peptide.txt"),
-    imputation = c("prot_imputation", "ContrastsSimpleImpute, ContrastsModerated", "protein",  "peptide.txt"),
+    imputation = c("prot_imputation", "ContrastsMissing, ContrastsModerated", "protein",  "peptide.txt"),
     lmmed = c("prot_med_lm_moderated", "strategy_lm, Contrasts, ContrastsModerated",  "protein", "peptide.txt"),

### 2023-01-09 inst/MyArticle/analysingDataStartingfromProteinGroups.Rmd
rename ContrastSimpleImpute to ContrastMissing #28

@@
 lfqTrans$config$table$factorDepth <- 2
-# ContrastsSimpleImpute$debug("get_contrasts")
-contrSimple <- prolfqua::ContrastsSimpleImpute$new(lfqdata = lfqTrans,
+# ContrastsMissing$debug("get_contrasts")
+contrSimple <- prolfqua::ContrastsMissing$new(lfqdata = lfqTrans,
                                                    Contrasts)

### 2023-01-09 inst/MyArticle/paper.Rmd
rename ContrastSimpleImpute to ContrastMissing #28

@@
 We specify a contrast interface (`ContrastsInterface`).
-Several implementations allow to perform DEA given linear or mixed effect models (`Contrasts`), variance shrinkage (`ContrastsModerated`), or to impute contrasts in cases when observations are missing for an entire group of samples (`ContrastsSimpleImpute`).
+Several implementations allow to perform DEA given linear or mixed effect models (`Contrasts`), variance shrinkage (`ContrastsModerated`), or to impute contrasts in cases when observations are missing for an entire group of samples (`ContrastsMissing`).
 Further implementations of the interface encapsulate and integrate DEA results of external tools such as *proDA* or of *SAINTexpress* [@teo2014saintexpress] used to analyze data from protein interaction studies.
@@
     lmer = c("strategy_lmer, Contrasts", "mixed effect modelling of peptide or protein abundances and group differences estimation" ),
-    gmi = c("ContrastsSimpleImpute" , "group difference estimation when no observations in one of the groups" ),
+    gmi = c("ContrastsMissing" , "group difference estimation when no observations in one of the groups" ),
     ROPECA = c("ContrastsROPECA", "estimating group differences for proteins by summarizing peptide differences"),
@@
 Afterward, the model is fitted to the data using the `build_model` function.
-Next, we estimate the contrasts either from the linear model using the `Contrasts` class or directly from the data using the `ContrastsSimpleImpute` class.
+Next, we estimate the contrasts either from the linear model using the `Contrasts` class or directly from the data using the `ContrastsMissing` class.
 Afterward, we apply $t$-statistic moderation using the `ContrastModerated` class.
@@
 
-## compute contrasts using imputation imputation
-conI <- prolfqua::ContrastsSimpleImpute$new( lfqpro, contrasts, modelName = "imp")
+## estimate contrasts when observations missing
+conI <- prolfqua::ContrastsMissing$new( lfqpro, contrasts, modelName = "mis")
 
@@
 
-allBenchmarks <- allBenchmarks[c("benchmark_imputation","benchmark_ProtModerated",  "benchmark_mixedModerated", "benchmark_ropeca","benchmark_merged","benchmark_mssstats","benchmark_proDA", "benchmark_msqrob")]
+allBenchmarks <- allBenchmarks[c("benchmark_missing","benchmark_ProtModerated",  "benchmark_mixedModerated", "benchmark_ropeca","benchmark_merged","benchmark_mssstats","benchmark_proDA", "benchmark_msqrob")]
 
@@
     proDA = c("proDA", "probabilistic dropout model",  "protein", "peptide.txt"),
-    imputation = c("prolfqua_estimate", "ContrastsSimpleImpute, ContrastsModerated", "protein",  "peptide.txt"),
+    imputation = c("prolfqua_missing", "ContrastsMissing, ContrastsModerated", "protein",  "peptide.txt"),
     lmmed = c("prolfqua_lm_mod", "strategy_lm, Contrasts, ContrastsModerated",  "protein", "peptide.txt"),
-    prot_merged = c("prolfqua_merged", "addContrastResults( prefer = proflqua_lm_mod , add = prolfqua_estimate)HACKIT ",  "protein", "peptide.txt"),
+    prot_merged = c("prolfqua_merged", "addContrastResults( prefer = proflqua_lm_mod , add = prolfqua_missing)HACKIT ",  "protein", "peptide.txt"),
     mixed = c("prolfqua_mix_eff_mod", "strategy_lmer, Contrasts, ContrastsModerated", "peptide", "peptide.txt" ),

### 2023-01-09 inst/MyArticle/paper.Rmd
updated table 1 close #29

@@
 ```{r tableOverview}
-caption = "Rows - R packages, Columns - types of models supported: pd - parallel design, rm - repeated measurements, fd - factorial design, int - interactions among factors, mem - mixed effect models, eb - empirical Bayes, md - missing data modelling.
-Y - yes, * - repeated measurements are modeled using a fixed effect."
+caption = "Rows - R packages, Columns - types of models supported: pd - parallel design, rm - repeated measurements, fd - factorial design, int - interactions among factors, mem - mixed effect models, eb - empirical Bayes, md - missing data modelling with no imputation needed.
+Y - yes"
 
 xx <- list(
-  MSstats = c("pd" = "Y", "rm" = "Y", "mem" = "Y"),
   PECA = c("pd" = "Y", "rm" = "Y", eb = "Y"),
   limma  = c("pd" = "Y", "rm" = "Y", fd = "Y", int = "Y",eb = "Y"),
+  MSstats = c("pd" = "Y", "rm" = "Y", fd = "Y", int = "Y", mem = "Y"),
+  proDA = c("pd" = "Y", "rm" = "Y", fd = "Y", int = "Y", eb = "Y", md = "Y"),
   msqrob2 = c("pd" = "Y", "rm" = "Y", fd = "Y", int = "Y", mem = "Y", eb = "Y", md = "Y"),
-  proDA = c("pd" = "Y", "rm" = "Y", fd = "Y", int = "Y", eb = "Y", md = "Y"),
   prolfqua = c("pd" = "Y", "rm" = "Y", fd = "Y", int = "Y", mem = "Y",eb = "Y", md = "Y"))

### 2023-01-12 inst/MyArticle/prolfqua_supplement.Rmd
moved away from BiocStyle

@@
 ---
-title: "Supporting Information:\n prolfqua - A Comprehensive R-package for Proteomics Differential Expression Analysis"
+title: | 
+    | Supporting Information
+    | prolfqua : A Comprehensive R-package for Proteomics Differential Expression Analysis
 author:
-  - name: Witold E. Wolski
-    affiliation:
-      - Functional Genomics Center Zurich (FGCZ), ETH Zurich / University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland
-      - Swiss Institute of Bioinformatics (SIB), Quartier Sorge - Batiment Amphipole, 1015 Lausanne, Switzerland
+  - Witold E. Wolski:
       email: wew@fgcz.ethz.ethz.ch
-  - name: Paolo Nanni
-    affiliation: Functional Genomics Center Zurich (FGCZ), ETH Zurich / University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland
-  - name: Jonas Grossmann
-    affiliation:
-      - Functional Genomics Center Zurich (FGCZ), ETH Zurich / University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland
-      - Swiss Institute of Bioinformatics (SIB), Quartier Sorge - Batiment Amphipole, 1015 Lausanne, Switzerland
-  - name: Maria d'Errico
-    affiliation:
-      - Functional Genomics Center Zurich (FGCZ), ETH Zurich / University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland
-      - Swiss Institute of Bioinformatics (SIB), Quartier Sorge - Batiment Amphipole, 1015 Lausanne, Switzerland
-  - name: Ralph Schlapbach
-    affiliation: Functional Genomics Center Zurich (FGCZ), ETH Zurich / University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland
-  - name: Christian Panse
-    affiliation:
-      - Functional Genomics Center Zurich (FGCZ), ETH Zurich / University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland
-      - Swiss Institute of Bioinformatics (SIB), Quartier Sorge - Batiment Amphipole, 1015 Lausanne, Switzerland
+      institute: [fgcz, sib]
+      correspondence: true
+  - Paolo Nanni:
+      institute: [fgcz]
+  - Jonas Grossmann:
+      institute: [fgcz, sib]
+  - Maria dErrico:
+      institute: [fgcz, sib]
+  - Ralph Schlapbach:
+      institute: [fgcz]
+  - Christian Panse:
+      institute: [fgcz, sib]
+institute:
+  - fgcz: Functional Genomics Center Zurich (FGCZ), ETH Zurich / University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland
+  - sib: Swiss Institute of Bioinformatics (SIB), Quartier Sorge - Batiment Amphipole, 1015 Lausanne, Switzerland
 package: prolfqua
 output:
-  BiocStyle::pdf_document
+  bookdown::pdf_document2:
+    toc: yes
+    pandoc_args:
+      - '--lua-filter=scholarly-metadata.lua'
+      - '--lua-filter=author-info-blocks.lua'
 abstract: |
@@
   chunk_output_type: console
-#bibliography: prolfqua.bib
+  bibliography: prolfqua.bib 
 ---

### 2023-01-13 inst/MyArticle/PosterBioconductor.Rmd
update files.

@@
   prolfqua::ContrastsModerated$new()
-conI <- prolfqua::ContrastsSimpleImpute$new( lfqp, contrasts) |> 
+conI <- prolfqua::ContrastsMissing$new( lfqp, contrasts) |> 
   prolfqua::ContrastsModerated$new()

### 2023-01-13 inst/MyArticle/paper.Rmd
update files.

@@
 However, the existing packages differ widely regarding supported designs, model specifications, and output formats.
-At the same time, they have the following features in common: fitting linear models to either peptide or protein intensities, determining differences among groups, and afterward applying empirical Bayes variance shrinkage.
+At the same time, they have the following features in common: fitting linear models to either peptide or protein abundances, determining differences among groups, and afterward applying empirical Bayes variance shrinkage.
 Therefore, the revised goal was to provide a modular object-oriented design, with *R* linear models as a core, and add features such as $p$-value aggregation, variance shrinkage, or modeling of missing observations.
@@
 
-### Contrast estimation in presence of missing data.
+### Contrast estimation in presence of missing data using LOD
 
@@
 
-The Benchmark functionality of *prolfqua* includes receiver operator curves (ROC) and computes partial areas under those curves ($pAUC$) and other scores, e.g., the false discovery proportion $FDP$.
-We use those scores, i.e.
-the $pAUC$ at $10\%$ $FDR$ and the $FDP$, to examine how well the methods implemented in *prolfqua* model quantitative mass spectrometric high throughput data and compare them with the results produced by *MSstats*, *proDA* and *msqrob2*.
-Table \@ref(tab:describeBenchmarkedModels) summarizes all the modelling methods we evaluated.
+Using a benchmark dataset whose ground truth is known (see Methods), we assessed the performance of different modeling approaches implemented in *prolfqua*, *MSstats*, *proDA* and *msqrob2*.
+Table \@ref(tab:describeBenchmarkedModels) summarizes which methods we have evaluated, which *MaxQuant* files we used as input, and if the models are fitted to peptide or protein intensities.
+We make the *R*-markdown files to replicate the benchmarking available at `r BiocStyle::Githubpkg("wolski/prolfquabenchmark")`.
+
+The IonStar/MaxQuant dataset (see Methods) captures only the variance from the chromatography, electro-spray, and mass spectrometric measurements since only technical replicates are available for each dilution.
+Therefore essential sources of variation typically present in other experiments, such as the biochemical and biological ones, are not measured.
+Furthermore, this dataset with a parallel group design does not allow for benchmarking models with interactions.
+Thus, while we can extrapolate some of the results obtained to more realistic datasets, we need to be careful not to over-interpret our findings.
+Specifically, the observed variances will be higher in datasets with biological replicates, and the power will be lower for the same number of samples.
+Furthermore, the proportion of missing observations in real-life datasets might be higher or distributed differently in groups.
 
@@
 
-When comparing DEA performance, a relevant parameter is the number of proteins for which a method estimates differences (see Figure \@ref(fig:figure3) Panel A), which indicates how robust the procedure works in the presence of missing observations.
-For each protein, we tried to determine four differences ($\Delta = (1.20, 1.25, 1.30, 1.50)$).
-Given 4178 proteins with at least two peptides, there are in total $16712$ possible differences.
-However, some methods can not estimate all of them.
-The set of proteins with effect size estimates might differ for each method, biasing direct comparison of scores such as the $pAUC$.
-For instance, we observe that *MSstats* estimates $16038$ group differences while the mixed effect models estimates the fewest with $15756$.
+When comparing DEA performance, a relevant parameter is the number of differences among conditions a method can estimate  (see Figure \@ref(fig:figure3) Panel A).
+For each protein, we tried to determine four differences ($\Delta = (1.20, 1.25, 1.30, 1.50)$) and therefore given $4178$ proteins with at least two peptides, there are, in total, $16712$ possible differences.
+Since *msqrob2*, *proDA*, *proflqua_missing*, and *prolfqua_merged* are directly modeling missing observation, they estimate all possible contrasts.
+However, some models fail to estimate differences when abundances are unobserved or rely on imputation.
+For instance, when using the mixed effect models, sensitive to missing data, we estimate the fewest number of contrasts with $15756$.
+
+The Benchmark functionality of *prolfqua* includes receiver operator curves (ROC) and computes partial areas under those curves ($pAUC$) and other scores, e.g., the false discovery proportion $FDP$.
+Since the set of effect size estimates will differ for some methods, e.g., $16712$ vs.
+$15756$ (see Figure \@ref(fig:figure3) Panel A), this introduces a bias when computing receiver operator curves and the $pAUC$.
 Hence, to conclude that one method shows a better performance, it does not suffice if the $pAUC$ is greater, but the number of proteins with differential expression results needs to be equal or larger.
+However, for *proDA*, *msqrob2*, and *prolfqua_merged*, we can compare the $pAUC$ to asses which method performs best.
 
@@
 Ordering proteins by the $t$-statistic or $p$-value leads to a higher $pAUC_{10}$ than when ordering by the estimated difference among groups.
+
+We can conclude that if we want to sort the proteins according to the likelihood of being differentially regulated to perform gene set enrichment analysis [@subramanian2005gene], the $t$-statistic is better suited than the fold-change estimate.
+When computing the $p$-values from the t-statistics we are incorporating the degrees of freedom, which improve the inferences (see Figure \@ref(fig:figure3) panel B center versus panel B left).
+There is no such improvement for the mixed effect model.
+The reason is an erroneous denominator degree of freedom estimation for many proteins, a known problem in the case of mixed effect models.
+Furthermore, for the fixed effect linear model, the empirical Bayes variance shrinkage, as suggested by @Smyth2004linear, consistently improves the ranking of proteins compared with the unmoderated estimates (not shown).
+However, since also for this method a correct degree of freedom estimate is required, it does not work for mixed effect models.
+
 Suppose an accurate estimate of the difference among groups is essential.
-In that case, the linear models fitted to protein intensities, calculated using Tukey's median polish, perform better than those when directly modeling peptide intensities, i.e., `ropeca` (see Table \@ref(tab:describeBenchmarkedModels) Abundance column).
-We speculate that outliers at the peptide level do not affect the protein estimates when using Tukey's median polish method.
+In that case, among the models fitted to protein intensities, calculated using Tukey's median polish, the *proDA* model performed best (see Figure \@ref(fig:figure3) panel B left).
+The dropout model more accurately models the posterior protein intensities, compared with *prolfqua_missing*, which uses a point estimate of the $LOD$.
+Furthermore, the *prolfqua_ropeca* model that first fits peptide level models and then summarizes differences performed worst.
+When using Tukey's median polish method, we speculate that the peptide-level outliers do not affect the protein estimates.
 
@@
 In the case of *MSstats*, we observe a high proportion of false discoveries for small $FDR$ values.
-In the case of the ROPECA method, the $FDR$ estimates, obtained by applying the Benjamini-Hochberg correction to the Beta distribution-based regulation probabilities, strongly overestimate the $FDP$.
+In the case of the *prolfqua_ropeca* method, the $FDR$ estimates, obtained by applying the Benjamini-Hochberg correction to the Beta distribution-based regulation probabilities, strongly overestimate the $FDP$.
+
+However, computing the t-statistics at the peptide level and then summarizing it for each protein using their median produces the highest $pAUC_{10}$ scores among all the tested models (see Figure \@ref(fig:figure3) Panel B *prolfqua_ropeca*).
+Furthermore, by using the Beta distribution to model the number of peptides observed, we can further improve the $pAUC$ scores (see Figure \@ref(fig:figure3) Panel B center).
+However, the properties of Beta-based probabilities need to be better understood; their distribution is not uniform under the null hypothesis (see Supplementary Section "The probabilities produced by ROPECA are not p-values").
+Therefore, the resulting $FDR$ estimates are biased (see Figure \@ref(fig:figure3) Panel C).
+Consequently, we can not recommend this method if an unbiased estimate of $FDR$ is essential, which is frequently the case.
+In addition, peptides are stronger affected by missing values, reducing the number of contrasts we could estimate for the dataset using this method (see Figure \@ref(fig:figure3) Panel C).
 
@@
 
-Using a benchmark dataset whose ground truth is known (see Methods), we assessed the performance of different modeling approaches implemented in *prolfqua*, *MSstats*, *proDA* and *msqrob2*.
-Table \@ref(tab:describeBenchmarkedModels) summarizes which methods we have evaluated, which *MaxQuant* files we used as input, and if the models are fitted to peptide or protein intensities.
-We make the *R*-markdown files to replicate the benchmarking available at `r BiocStyle::Githubpkg("wolski/prolfquabenchmark")`.
-
 ```{r describeBenchmarkedModels, results='asis' }
@@
     msstats = c("MSstats", "preprocess with default parameters",  "", "evidence.txt"),
-    msqrob = c("msqrob2", "hurdle msqrob model", "protein and peptide", "peptide.txt"),
+    msqrob = c("msqrob2::msqrobHurdle", "merge of msqrobHurdleIntensity and msqrobHurdleCount results", "protein and peptide", "peptide.txt"),
     proDA = c("proDA", "probabilistic dropout model",  "protein", "peptide.txt"),
@@
     mixed = c("prolfqua_mix_eff_mod", "strategy_lmer, Contrasts, ContrastsModerated", "peptide", "peptide.txt" ),
-    ropeca = c("prolfqua_ropeca", "strategy_lm, Contrasts, ContrastsModerated, ContrastsROPECA", "peptide", "peptide.txt" )
+    ROPECA = c("prolfqua_ropeca", "strategy_lm, Contrasts, ContrastsModerated, ContrastsROPECA", "peptide", "peptide.txt" )
 )
@@
 
-The IonStar/MaxQuant dataset (see Methods) captures only the variance from the chromatography, electro-spray, and mass spectrometric measurements since only technical replicates are available for each dilution.
-Therefore essential sources of variation typically present in other experiments, such as the biochemical and biological ones, are not measured.
-Furthermore, this dataset with a parallel group design does not allow for benchmarking models with interactions.
-Thus, while we can extrapolate some of the results obtained to more realistic datasets, we need to be careful not to over-interpret our findings.
-Specifically, the observed variances will be higher in datasets with biological replicates, and the power will be lower for the same number of samples.
-Furthermore, the proportion of missing observations in real-life datasets might be higher or distributed differently in groups.
+The *R*-package *proDA*, *msqrob2*, and *prolfqua* do not impute missing data but integrate them into the statistical model, while *MSstats* filters and imputes the data using an accelerated failure model.
+Despite imputation, *MSstats* estimates fewer group differences ($16038$), and does not achieve a higher $pAUC_{10}$  (see Figure \@ref(fig:figure3)).
+Furthermore, Figure \@ref(fig:figure3) Panel C shows that when using *MSstats*, the proportion of false discoveries might be very high for a low $FDR$ because of false positives.
+Hence, augmenting the linear model to handle missing observations, using the quasibinomial generalized linear model, the dropout model, or estimating missing differences using the LOD, simplifies the analysis pipeline since no imputation is needed and improves the quality of the estimates.
 
-We can conclude that if we want to sort the proteins according to the likelihood of being differentially regulated to perform gene set enrichment analysis [@subramanian2005gene], the $t$-statistic is better suited than the fold-change estimate.
-When computing the $p$-values from the t-statistics we are incorporating the degrees of freedom, which improve the inferences (see Figure \@ref(fig:figure3) panel B center versus panel B left).
-There is no such improvement for the mixed effect model.
-The reason is an erroneous denominator degree of freedom estimation for many proteins, a known problem in the case of mixed effect models.
-Furthermore, for the fixed effect linear model, the empirical Bayes variance shrinkage, as suggested by @Smyth2004linear, consistently improves the ranking of proteins compared with the unmoderated estimates (not shown).
-However, since also for this method a correct degree of freedom estimate is required, it does not work for mixed effect models.
-
-Computing the statistics at the peptide level, e.g., the $t$-statistics, then summarizing it for each protein using their median, produces the highest $pAUC_{10}$ scores among all the tested models (see Figure \@ref(fig:figure3) Panel B `prolfqua_ropeca`).
-Furthermore, by using the Beta distribution to model the number of peptides observed, we can further improve the $pAUC$ scores (see Figure \@ref(fig:figure3) Panel B center).
-However, the properties of Beta-based probabilities are not well understood; for instance, it is not uniformly distributed under the null hypothesis (see Supplementary Section "The probabilities produced by ROPECA are not p- values").
-Furthermore, the $FDR$ estimates obtained when correcting for multiple testing with the Benjamini-Hochberg method are biased and overestimate the FDP (see Figure \@ref(fig:figure3) Panel C).
-Therefore, we can not recommend this method if an unbiased estimate of $FDR$ is essential, which is frequently the case.
-In addition, peptides are stronger affected by missing values, reducing the number of contrasts we could estimate for the dataset using this method (see Figure \@ref(fig:figure3) Panel C).
-
-The *R*-package *proDA*, *msqrob2* and *prolfqua* do not impute missing data but integrate them into the statistical model, while *MSstats* filters and imputes the data using an accelerated failure model.
-Despite imputation, *MSstats* estimates group differences for less proteins, and does not achieve a higher $pAUC_{10}$.
-Furthermore, Figure \@ref(fig:figure3) Panel C shows that when using *MSstats* the proportion of false discoveries might be very high for a low $FDR$ because of false positives.
 Of note, *MSStats* uses the _evidence.txt_ file, while all the other methods use _peptide.txt_ files, as input (see Table \@ref(tab:describeBenchmarkedModels))).
@@
 These are possible confounding factors to consider.
-While *prolfqua*, as well as *proDA*, is highly modular, and to a lesser extent *msqrob2*, enabling us to use the same data preprocessing and normalization, *MSstats* is a monolithic software, making it unfeasible to change the preprocessing or normalization.
-Therefore, we could not control these confounders.
+While *prolfqua*, as well as *proDA*, is highly modular, and to a lesser extent *msqrob2*, enabling us to use the same data preprocessing and normalization, *MSstats* is monolithic, making it unfeasible to use a preprocessing or normalization method not available in *MSstats*.
 
-We obtained difference and FDR estimates for all proteins, as shown in Figure \@ref(fig:figure3) Panel A when using: a) the probabilistic dropout model (*proDA*), b) the hurdle model (*msqrob2*), and c) 'prolfqua_merge' (Table \@ref(tab:describeBenchmarkedModels).
+We obtained difference and FDR estimates for all proteins and comparisons, as shown in Figure \@ref(fig:figure3) Panel A when using: a) the probabilistic dropout model (*proDA*), b) the hurdle model (*msqrobHurdle*), and c) *prolfqua_merged*.
 We observe that the performance of the scaled $p$-values or the $t$-statistics is comparable among these three methods (Figure \@ref(fig:figure3) Panel B).
@@
 Furthermore, all three models perform similarly when examined using a different benchmark dataset CPTAC/MaxQuant (see Methods).
-For this dataset, *proDA* performed slightly but not significantly better than *prolfqua* and *msqrob* (and Supplementary Section "DEA benchmark : CPTAC/MaxQuant/peptide.txt").
+For this dataset, *proDA* performed slightly but not significantly better than *prolfqua* and *msqrobHurdle* (and Supplementary Section "DEA benchmark : CPTAC/MaxQuant/peptide.txt").
 
@@
 Using protein abundances as input significantly simplifies the analysis and interpretation and might benefit from optimization implemented in the quantification software.
-However, we can only easily fit the *proDA* and 'prolfqua_merged' (see Table \@ref(tab:describeBenchmarkedModels)) models to protein intensities.
-In this DEA benchmark, *prolfqua* performed slightly, but not significantly better than *proDA* (Supplementary Section "DEA benchmark : IonStar/FragPipeV14/combined_protein.tsv").
+However, we can only fit the *proDA* and *prolfqua_merged* (see Table \@ref(tab:describeBenchmarkedModels)) models to protein abundances, while *MSstats* and *msqrobHurdle* require psm or peptide level abundances In this DEA benchmark, *prolfqua* performed slightly, but not significantly better than *proDA* (Supplementary Section "DEA benchmark : IonStar/FragPipeV14/combined_protein.tsv").
 
@@
 Since *MSstats*, *msqrob2*, *proDA*, and *prolfqua* all read `MStstats.tsv` files, we could eliminate a confounding factor, i.e., different input abundances (Supplementary Section "DEA benchmark: IonStar/FragPipeV14/MSstats.tsv").
-In this DEA benchmark, *msqrob2* and *prolfqua* ('prolfqua_merged') perform best but not significantly better than *proDA* or *MSstats*.
+In this DEA benchmark, *msqrob2* and *prolfqua_merged* perform best but not significantly better than *proDA* or *MSstats*.
 Furthermore, by processing the *IonStar* dataset with both *MaxQuant* and *FragPipe*, we can compare their performances (Supplementary Section "Comparing DEA results for *MaxQuant* and *FragPipe*").
@@
 However, some of these steps are of utmost significance when performing differential expression analysis [@frohlich2022benchmarking].
-One of them is the normalization of the intensities within the samples to remove systematic differences [@pursiheimo2015optimization].
+One of them is the normalization of the abundances within the samples to remove systematic differences [@pursiheimo2015optimization].
 The method used to infer proteins from peptide identifications [@nesvizhskii2005interpretation] and protein abundances from peptide abundances is an additional important factor [@Grossmann2010].
 For instance, the original *proDA* publication uses MaxLFQ [@cox2014accurate] protein estimates.
-However, when using MaxLFQ intensities reported by *MaxQuant*, the $pAUC_{10}$ is lower ($pAUC_{10}(\textrm{t-statistics}) = 66\%$) compared with results obtained when protein abundances are estimated from peptide abundances using Tukey's median polish ($pAUC_{10}(\textrm{t-statistics}) = 72\%$).
+However, when using MaxLFQ abundances reported by *MaxQuant*, the $pAUC_{10}$ is lower ($pAUC_{10}(\textrm{t-statistics}) = 66\%$) compared with results obtained when protein abundances are estimated from peptide abundances using Tukey's median polish ($pAUC_{10}(\textrm{t-statistics}) = 72\%$).
 Last but not least, the software [@Cox2008MaxQuant; @yu2020fast] used to identify and quantify proteins significantly contributes to the entire pipeline's performance altering the number of identified proteins and the sensitivity and specificity of the differential expression analysis.
@@
 It is relevant that an analysis pipeline shows good performance in all these categories.
-Leveraging these computational experiments, we can provide the following advice: i) estimate protein abundances from peptide abundances using a robust or nonparametric regression method; ii) use linear models because they show good performance in all categories; iii) if the measurements are correlated, as for technical replicates, mixed effect models might work if the sample sizes are large; if not, aggregate the replicates and fit a linear model instead; iv) if you use fixed-effect linear models, apply variance moderation to improve the $t$-statistics and $p$-value estimates; v) If you want to sort your protein lists to perform gene set enrichment analysis, use the $t$-statistic instead of the difference; and last but not least vi) do not impute missing observation but statistically model missingness to estimate parameters, i.e., group differences.
-Finally, the DEA results obtained with *prolfqua* are comparable to those of other DEA tools.
+The examined models *prolfqua_merged*, *proDa*, and *msqrob2* showed good performance in all these categories.
+Leveraging these computational experiments, we can provide the following advice: i) estimate protein abundances from peptide abundances using a robust or nonparametric regression method; ii) fit linear models to protein abundances, iii) do not impute missing observation but statistically, model missingness to estimate parameters, i.e., group differences, iv) explicitly report the model used v) if the measurements are correlated, as for technical replicates, mixed effect models might work if the sample sizes are large; if not, aggregate the replicates and fit a linear model instead; iv) if you use fixed-effect linear models, apply variance moderation to improve the $t$-statistics and $p$-value estimates; v) If you want to sort your protein lists to perform gene set enrichment analysis, use the $t$-statistic instead of the difference; and last but not least.
 

### 2023-01-13 inst/MyArticle/paper.Rmd
caption

@@
 ```{r tableOverview}
-caption = "Rows - R packages, Columns - types of models supported: pd - parallel design, rm - repeated measurements, fd - factorial design, int - interactions among factors, mem - mixed effect models, eb - empirical Bayes, md - missing data modelling with no imputation needed.
+caption = "Rows - R packages, Columns - types of models supported: pd - parallel design, rm - repeated measurements, fd - factorial design, int - interactions among factors, mem - mixed effect models, eb - empirical Bayes, md - missing data modelling (no imputation needed).
 Y - yes"

### 2023-01-16 inst/MyArticle/paper.Rmd
cleanup figures

@@
 The configuration encapsulates the differences among the various input formats in column names and enables the use of the methods without additional arguments.
-An example code for creating an *FragPipe*[@yu2020fast] configuration can be found in the Supplementary Section "Creating a prolfqua configuration".
+An example code for creating an *FragPipe*[@yu2020fast] configuration can be found in the Supporting Material S3 Section "Creating a prolfqua configuration".
 We implemented methods that transform the data into tidy data for popular software like *MaxQuant*[@Cox2008MaxQuant], or *FragPipe*, which stores the same variable, e.g., intensity, in multiple columns, one for each sample.
@@
 We use R6 classes to structure the functionality of the package (see Figure \@ref(fig:LFQData) and Figure \@ref(fig:ContrastUML)).
-R6 classes are well supported by command-line completion features (see Supplementary Figure 8) in *RStudio* [@RStudio2022], and help to implement argument free functions.
+R6 classes are well supported by command-line completion features (see Supporting Figure 8) in *RStudio* [@RStudio2022], and help to implement argument free functions.
 
@@
 ```{r LFQData, echo=FALSE, fig.cap="(ref:LFQData)", out.width = '90%'}
-knitr::include_graphics("LFQData_UML_LARGE.png")
+knitr::include_graphics("graphics/LFQData_UML_LARGE.png")
 ```
@@
 
-knitr::include_graphics("ContrastClassesUML_LARGE.png")
+knitr::include_graphics("graphics/ContrastClassesUML_LARGE.png")
 ```
@@
 The package *prolfqua* provides functions to determine the vector of $parameter$ weights $c$, from a linear model and a contrast specification string.
-In the Supplement we provided an example, how to specify contrasts for a dataset with two explanatory variables and an interaction term (Supplementary Section "Specifying Contrasts for Models with two Factors and Interaction Term").
+In the Supporting Material S10 we provided an example, how to specify contrasts for a dataset with two explanatory variables and an interaction term.
 
@@
 In such a case, we will substitute the group mean using the expected protein abundance $A$ at the LOD $A_{LOD}$.
-To estimate $A_{LOD}$ we are using the protein abundances of those groups where the protein was observed in only a single sample (see Supplementary Section "Estimating $A_{LOD}$").
+To estimate $A_{LOD}$ we are using the protein abundances of those groups where the protein was observed in only a single sample (see Supporting Material S8 "Estimating $A_{LOD}$").
 Typically there are many such cases, and hence we take the median.
@@
 The use of R6 classes, which encapsulate the configuration and the data, allows for writing very concise code where functions can have few arguments.
-Auto-completion support for R6 classes in the *RStudio* editor makes it easy for novices to explore *prolfqua*'s functionality (see Supplementary Figure 8).
+Auto-completion support for R6 classes in the *RStudio* editor makes it easy for novices to explore *prolfqua*'s functionality (see Supporting Figure 8).
 
@@
 Furthermore, by using the Beta distribution to model the number of peptides observed, we can further improve the $pAUC$ scores (see Figure \@ref(fig:figure3) Panel B center).
-However, the properties of Beta-based probabilities need to be better understood; their distribution is not uniform under the null hypothesis (see Supplementary Section "The probabilities produced by ROPECA are not p-values").
+However, the properties of Beta-based probabilities need to be better understood; their distribution is not uniform under the null hypothesis (see Supporting Material S9 "The probabilities produced by ROPECA are not p-values").
 Therefore, the resulting $FDR$ estimates are biased (see Figure \@ref(fig:figure3) Panel C).
@@
 We observe that the performance of the scaled $p$-values or the $t$-statistics is comparable among these three methods (Figure \@ref(fig:figure3) Panel B).
-We tested if there is a significant difference between the $pAUC_{10}$ for all three methods, but did not reject the null hypothesis that there is no such difference (Supplementary Section "DEA benchmark : IonStar/MaxQuant/peptide.txt - Significance test").
+We tested if there is a significant difference between the $pAUC_{10}$ for all three methods, but did not reject the null hypothesis that there is no such difference (Supporting Material S3 "DEA benchmark : IonStar/MaxQuant/peptide.txt - Significance test").
 Also, the FDR estimates (Figure \@ref(fig:figure3) Panel C) are comparable for all three methods.
@@
 Furthermore, all three models perform similarly when examined using a different benchmark dataset CPTAC/MaxQuant (see Methods).
-For this dataset, *proDA* performed slightly but not significantly better than *prolfqua* and *msqrobHurdle* (and Supplementary Section "DEA benchmark : CPTAC/MaxQuant/peptide.txt").
+For this dataset, *proDA* performed slightly but not significantly better than *prolfqua* and *msqrobHurdle* (and Supporting Material S4 "DEA benchmark : CPTAC/MaxQuant/peptide.txt").
 
@@
 Using protein abundances as input significantly simplifies the analysis and interpretation and might benefit from optimization implemented in the quantification software.
-However, we can only fit the *proDA* and *prolfqua_merged* (see Table \@ref(tab:describeBenchmarkedModels)) models to protein abundances, while *MSstats* and *msqrobHurdle* require psm or peptide level abundances In this DEA benchmark, *prolfqua* performed slightly, but not significantly better than *proDA* (Supplementary Section "DEA benchmark : IonStar/FragPipeV14/combined_protein.tsv").
+However, we can only fit the *proDA* and *prolfqua_merged* (see Table \@ref(tab:describeBenchmarkedModels)) models to protein abundances, while *MSstats* and *msqrobHurdle* require psm or peptide level abundances In this DEA benchmark, *prolfqua* performed slightly, but not significantly better than *proDA* (Supporting Material S5 "DEA benchmark : IonStar/FragPipeV14/combined_protein.tsv").
 
 Finally, we also compared the DEA performances when starting the analysis from the precursor abundances reported in the 'MSstats.tsv' file, generated by *FragPipe* v14, from the *IonStar* dataset.
-Since *MSstats*, *msqrob2*, *proDA*, and *prolfqua* all read `MStstats.tsv` files, we could eliminate a confounding factor, i.e., different input abundances (Supplementary Section "DEA benchmark: IonStar/FragPipeV14/MSstats.tsv").
+Since *MSstats*, *msqrob2*, *proDA*, and *prolfqua* all read `MStstats.tsv` files, we could eliminate a confounding factor, i.e., different input abundances (Supporting Material S6 "DEA benchmark: IonStar/FragPipeV14/MSstats.tsv").
 In this DEA benchmark, *msqrob2* and *prolfqua_merged* perform best but not significantly better than *proDA* or *MSstats*.
-Furthermore, by processing the *IonStar* dataset with both *MaxQuant* and *FragPipe*, we can compare their performances (Supplementary Section "Comparing DEA results for *MaxQuant* and *FragPipe*").
+Furthermore, by processing the *IonStar* dataset with both *MaxQuant* and *FragPipe*, we can compare their performances (Supporting Material S7 "Comparing DEA results for *MaxQuant* and *FragPipe*").
 
@@
 Last but not least, the software [@Cox2008MaxQuant; @yu2020fast] used to identify and quantify proteins significantly contributes to the entire pipeline's performance altering the number of identified proteins and the sensitivity and specificity of the differential expression analysis.
-In the Supplementary Section "Comparing DEA results for *MaxQuant* and *FragPipe*" we compare DEA benchmarking results for the quantification software.
+In the Supporting Material S7 "Comparing DEA results for *MaxQuant* and *FragPipe*" we compare DEA benchmarking results for the quantification software.
 While the number of proteins identified with two peptides is practically the same, the DEA benchmark performance differs significantly by $\approx 10\%$ for the $pAUC_{10}$ score.
@@
 
+# Supporting Information
+
+- Supporting Material S1: How to install *prolfqua* and *prolfquabenchmark*
+- Supporting Material S2: Benchmark Vignettes
+- Supporting Material S3: DEA benchmark IonStar/MaxQuant/peptide.txt - Significance test
+- Supporting Material S4: DEA benchmark : CPTAC/MaxQuant/peptide.txt
+- Supporting Material S5: DEA benchmark : IonStar/FragPipeV14/combined_protein.tsv
+- Supporting Material S6: DEA benchmark : IonStar/FragPipeV14/MSstats.tsv
+- Supporting Material S7: Comparing DEA results for *MaxQuant* and *FragPipe*
+- Supporting Material S8: Estimating $A_{LOD}$
+- Supporting Material S9: The probabilities produced by ROPECA are not p-values
+- Supporting Material S10: Specifying Contrasts for Models with two Factors and Interaction Term
+- Supporting Material S11: Creating a prolfqua configuration
+
 # Appendix  {.unnumbered}
@@
 
-file.path("hexStickerProlfqua.png") |>
+file.path("graphics/hexStickerProlfqua.png") |>
   knitr::include_graphics()

### 2023-01-16 inst/MyArticle/prolfqua_supplement.Rmd
cleanup figures

@@
     toc: yes
+    number_sections: false
     pandoc_args:
@@
 
-# How to install *prolfqua* and *prolfquabenchmark*
+# Material S1: How to install *prolfqua* and *prolfquabenchmark*
 
@@
 
-# DEA benchmark : IonStar/MaxQuant/peptide.txt - Significance test
+# Material S2: Benchmark Vignettes (IonStar/MaxQuant)
+
+Pre-build version of the vignettes, created to run the DEA benchmark, are available:
+
+- [Benchmarking of methods implemented in prolfqua using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkingIonstarData.html) 
+- [Benchmarking of MSstats using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_MSStats.html) 
+- [Benchmarking of proDA  using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_proDA_medpolish.html) 
+- [Benchmarking of msqrob2  using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkMSqRob2.html)
+
+# Material S3: DEA benchmark IonStar/MaxQuant/peptide.txt - Significance test
 
@@
 
-## Benchmark Vignettes
-
-Pre-build version of the vignettes, we created to run the DEA benchmark, are available:
-
-- [Benchmarking of methods implemented in prolfqua using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkingIonstarData.html) 
-- [Benchmarking of MSstats using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_MSStats.html) 
-- [Benchmarking of proDA  using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/Benchmark_proDA_medpolish.html) 
-- [Benchmarking of msqrob2  using the IonStar dataset MaxQuant](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkMSqRob2.html)
-
-# DEA benchmark : CPTAC/MaxQuant/peptide.txt
+# Material S4: DEA benchmark : CPTAC/MaxQuant/peptide.txt
 
@@
 
-# DEA benchmark : IonStar/FragPipeV14/combined_protein.tsv
+# Material S5: DEA benchmark : IonStar/FragPipeV14/combined_protein.tsv
 
@@
 
-# DEA benchmark : IonStar/FragPipeV14/MSstats.tsv
+# Material S6: DEA benchmark : IonStar/FragPipeV14/MSstats.tsv
 
@@
 
-# Comparing DEA results for *MaxQuant* and *FragPipe*
+# Material S7: Comparing DEA results for *MaxQuant* and *FragPipe*
 
@@
 
-# Estimating $A_{LOD}$
+# Material S8: Estimating $A_{LOD}$
 
@@
 
-# The probabilities produced by ROPECA are not p-values
+# Material S9: The probabilities produced by ROPECA are not p-values
 
@@
 
-# Specifying Contrasts for Models with two Factors and Interaction Term
+# Material S10: Specifying Contrasts for Models with two Factors and Interaction Term
 
@@
 
-# Creating a prolfqua configuration
+# Material S11: Creating a prolfqua configuration
 
@@
 Feel free to share and adapt, but don't forget to credit the author."}
-file.path("graphics/hexStickerProlfqua.png") |>
+file.path("graphics/hexstrickerPRL2.png") |>
   knitr::include_graphics()

### 2023-01-19 inst/MyArticle/paper.Rmd
Added TOC to supplement. Further fixes related to renaming ContrastsSimpleImpute to ContrastsMissing

@@
 If variance can not be estimated for a protein, because there are too few observations in other groups, we use the median pooled variance of all other proteins in the dataset.
-This method is implemented in the class `ContrastSimpleImpute` (see Figure \@ref(fig:ContrastUML)).
+This method is implemented in the class `ContrastsMissing` (see Figure \@ref(fig:ContrastUML)).
 
@@
     msstats = c("MSstats", "preprocess with default parameters",  "", "evidence.txt"),
-    msqrob = c("msqrob2::msqrobHurdle", "merge of msqrobHurdleIntensity and msqrobHurdleCount results", "protein and peptide", "peptide.txt"),
+    msqrob = c("msqrob2", "merge of msqrobHurdleIntensity and msqrobHurdleCount (msqrobHurdle)", "protein and peptide", "peptide.txt"),
     proDA = c("proDA", "probabilistic dropout model",  "protein", "peptide.txt"),
@@
              caption = paste0("All benchmarked models. ",
-             "Description - prolfqua function names, ",
+             "Label - name of the method, ", 
+             "Description - functions used in the respective packages, ",
              "Abudances - indicates if model is fitted to peptide or protein abundances, ",
@@
 
-# References {-}
-
-<div id="refs"></div>
-
-\newpage
-
 # Supporting Information
@@
 
-# Appendix  {.unnumbered}
+\newpage
+
+# References {-}
+
+<div id="refs"></div>
 

### 2023-01-19 inst/MyArticle/prolfqua_supplement.Rmd
Added TOC to supplement. Further fixes related to renaming ContrastsSimpleImpute to ContrastsMissing

@@
 
+(ref:toc) The figure illustrates some possible data analysis procedures that can
+be achieved with only very few functions from prolfqua. As input
+different levels from a variety of quantitation tools are accepted and
+along with an annotation table configured for prolfqua using either
+pre-built import functions or made ready with a little reshaping. In
+prolfqua (depending on the input), one can choose roll-up
+procedures and perform basic steps of mass spectrometry-based data
+processing steps such as quality control, data normalization, protein
+aggregation, as well as statistical modeling, hypothesis testing, and
+sample size estimation. At all levels, one can always use
+functions to show diagnostic plots or even inspect individual proteins.
+Furthermore, prolfqua supports multiple models for differential
+expression analysis such as linear (mixed effect) models and allows to
+specify multiple contrasts.
+
+```{r toc, echo=FALSE, fig.cap="(ref:toc)"}
+knitr::include_graphics("graphics/TOCv2.pdf")
+
+```
+
+\newpage
+
 # Session info {.unnumbered}

### 2023-01-19 inst/MyArticle/createUploadFiles.R

@@ -29,5 +29,8 @@ writeLines(results, "../../../prolfqua-achems/results.tex")
 conclusion <- tex[ grep("hypertarget\\{conclusion\\}", tex):(grep("hypertarget\\{acknowledgements\\}", tex) - 1)]
 writeLines(conclusion, "../../../prolfqua-achems/conclusion.tex")
 
-abbreviations <- tex[ grep("hypertarget\\{abbreviations\\}", tex):(grep("hypertarget\\{references\\}", tex) - 1)]
+abbreviations <- tex[ grep("hypertarget\\{abbreviations\\}", tex):(grep("hypertarget\\{supporting-information\\}", tex) - 1)]
 writeLines(abbreviations, "../../../prolfqua-achems/abbreviations.tex")
+
+abbreviations <- tex[ grep("hypertarget\\{supporting-information\\}", tex):(grep("hypertarget\\{references\\}", tex) - 1)]
+writeLines(abbreviations, "../../../prolfqua-achems/supporting.tex")

### 2023-01-19 inst/MyArticle/paper.Rmd
improve end

@@
 When comparing statistical modeling methods for the DEA, we assessed performance measures such as the number of estimated contrasts, the $pAUC$, and if the $FDR$ is an unbiased estimate of the $FDP$.
-It is relevant that an analysis pipeline shows good performance in all these categories.
-The examined models *prolfqua_merged*, *proDa*, and *msqrob2* showed good performance in all these categories.
+An analysis pipeline must demonstrate strong performance in all of these categories.
+Among the examined models, *prolfqua_merged*, *proDa*, and *msqrob2* all exhibit good performance in all these categories.
 Leveraging these computational experiments, we can provide the following advice: i) estimate protein abundances from peptide abundances using a robust or nonparametric regression method; ii) fit linear models to protein abundances, iii) do not impute missing observation but statistically, model missingness to estimate parameters, i.e., group differences, iv) explicitly report the model used v) if the measurements are correlated, as for technical replicates, mixed effect models might work if the sample sizes are large; if not, aggregate the replicates and fit a linear model instead; iv) if you use fixed-effect linear models, apply variance moderation to improve the $t$-statistics and $p$-value estimates; v) If you want to sort your protein lists to perform gene set enrichment analysis, use the $t$-statistic instead of the difference; and last but not least.

### 2023-03-03 inst/MyArticle/prolfqua_supplement.Rmd
fix spelling in author name.

@@
     | Supporting Information
-    | prolfqua : A Comprehensive R-package for Proteomics Differential Expression Analysis
+    | prolfqua: A Comprehensive R-package for Proteomics Differential Expression Analysis
 author:
@@
       institute: [fgcz, sib]
-  - Maria dErrico:
+  - Maria d'Errico:
       institute: [fgcz, sib]

### 2023-04-03 inst/MyArticle/prolfqua_supplement.Rmd
fix email

@@
   - Witold E. Wolski:
-      email: wew@fgcz.ethz.ethz.ch
+      email: wew@fgcz.ethz.ch
       institute: [fgcz, sib]
@@
 
-# Material S1: How to install *prolfqua* and *prolfquabenchmark*
+# Material S1: How to Install *prolfqua* and *prolfquabenchmark*
 

### 2024-06-28 inst/MyArticle/PosterBioconductor.Rmd
update vignettes to work with latest version of prolfqua

@@
 ```{r LFQData, echo=FALSE, fig.cap="Class Diagram for LFQData and related objects. The LFQData_Plotter class uses the LFQData class to implement methods for plotting.", out.width = '90%'}
-knitr::include_graphics("LFQData_UML_V2.png")
+knitr::include_graphics("graphics/LFQData_UML_V2.png")
 ```
@@
 ```{r ContrastUML, echo=FALSE, fig.cap="(ref:ContrastUML)", out.width = '90%'}
-knitr::include_graphics("ContrastClassesUML.png")
+knitr::include_graphics("graphics/ContrastClassesUML.png")
 ```
@@
 lfqd <- prolfqua::LFQData$new(d$data, d$config)
+
 t <- lfqd$get_Transformer()
@@
 ```{r}
-allBenchmarks <- readRDS("../../inst/Benchresults/allBenchmarks.RDS")
-benchmark_msstats <- readRDS("../../inst/Benchresults/benchmark_msstats.RDS")
-msFragger <- readRDS(file = "../../inst/Benchresults/MSFragger_medpol_benchmark.RDS")
+allBenchmarks <- readRDS(file.path("../../inst/Benchresults/",format( Sys.Date(), "%Y%m%d"),"allBenchmarks.RDS"))
+benchmark_msstats <- readRDS(file.path("../../inst/Benchresults/",format( Sys.Date(), "%Y%m%d"),"benchmark_msstats.RDS"))
+msFragger <- readRDS(file = file.path("../../inst/Benchresults/",format( Sys.Date(), "%Y%m%d"),"MSFragger_medpol_benchmark.RDS"))
 

### 2024-06-28 inst/MyArticle/SIB.Rmd
update vignettes to work with latest version of prolfqua

@@
 getpath <- function(filN){
-    f1 <- paste0("../../inst/Benchresults/",filN)
-    if(f1 == ""){
-       f1 <- system.file("Benchresults",filN,
-                         package = "prolfquabenchmark")
+    f1 <- file.path("../../inst/Benchresults/",format( Sys.Date(), "%Y%m%d"),filN)
+    if (f1 == "") {
+        f1 <- system.file("Benchresults",format( Sys.Date(), "%Y%m%d"),filN, package = "prolfquabenchmark")
     }
+    message(f1)
     return(f1)

### 2024-06-28 inst/MyArticle/analysingDataStartingfromProteinGroups.Rmd
update vignettes to work with latest version of prolfqua

@@
 
-TODO(wew,jg): Can you please add a oneliner of the purpose of the vignette?
+TODO(wew,jg): Can you please add a one liner of the purpose of the vignette?
 
@@
 ```{r anovaPvaluePlots, fig.cap="p-value distributions for ANOVA analysis."}
-mod$anova_histogram(what = "FDR.Pr..F.")
+mod$anova_histogram(what = "FDR")
 ```
@@
   dplyr::filter(factor == "f1.:f2.") |>
-  dplyr::arrange(FDR.Pr..F.) |>
+  dplyr::arrange(FDR) |>
   head(5)
@@
 protIntSig <- ANOVA |> dplyr::filter(factor == "f1.") |>
-  dplyr::filter(FDR.Pr..F. < 0.25)
+  dplyr::filter(FDR < 0.25)
 

### 2024-06-28 inst/MyArticle/paper.Rmd
update vignettes to work with latest version of prolfqua

@@
 ## read MQ peptide.txt and annotation table
-startdata <- prolfqua::tidyMQ_Peptides(system.file(
+startdata <- prolfquapp::tidyMQ_Peptides(system.file(
   "samples/maxquant_txt/tiny2.zip", package = "prolfqua"))
@@
 getpath <- function(filN){
-    f1 <- paste0("../../inst/Benchresults/",filN)
+    f1 <- file.path("../../inst/Benchresults/",format( Sys.Date(), "%Y%m%d"),filN)
     if (f1 == "") {
-        f1 <- system.file("Benchresults",filN, package = "prolfquabenchmark")
+        f1 <- system.file("Benchresults",format( Sys.Date(), "%Y%m%d"),filN, package = "prolfquabenchmark")
     }

### 2024-06-28 inst/MyArticle/prolfqua_supplement.Rmd
update vignettes to work with latest version of prolfqua

@@
 
-mqdata$data <- prolfqua::tidyMQ_Peptides(inputMQfile)
+mqdata$data <- prolfquapp::tidyMQ_Peptides(inputMQfile)
 mqdata$config <- prolfqua::create_config_MQ_peptide()
 annotation <- readxl::read_xlsx(inputAnnotation)
-res <- prolfqua::add_annotation(
+res <- prolfquapp::add_annotation(
   mqdata$data,

### 2024-06-28 inst/triqler/Benchmark_Triqler.Rmd
update vignettes to work with latest version of prolfqua

---
title: "Looking into Triqler results"
output: html_document
date: "2023-04-18"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

## R Markdown

```{r}
library(readr)
library(tidyverse)
resTriq <- read_tsv("../inst/triqler/proteins.tsv")
resTriq$peptides <- NULL

resTriq <- as.data.frame(resTriq)
resTriq$DECOY <- ifelse(grepl("^DECOY_", resTriq$protein), TRUE, FALSE)
resTriq$protein <-  gsub("^DECOY_", "", resTriq$protein)
resTriq <- resTriq |> tidyr::separate(protein, into =  c("ID", "Organism"), sep = "_",remove = FALSE)
resTriq$Organism |> table()

resTriq <- resTriq |> mutate(color = case_when(Organism == "YEAST" ~ "green",
                                               Organism == "HUMAN" ~ "black",
                                               Organism == "ECOLI" ~ "green",
                                               TRUE ~ "pink"))

resTriq <- resTriq |> arrange(color)
plot(resTriq$log2_fold_change, -log10(resTriq$q_value), col = resTriq$color, pch = 20)
Honly <- resTriq |> filter(color == "black")
points(Honly$log2_fold_change, -log10(Honly$q_value), pch = 20)

abline(h = -log10(0.1))
abline(v = c(-1,1))

```

# Analyse data with prolfqua

```{r}
res <- read_tsv("../inst/triqler/LFQBench_PS.csv")
res$condition <- paste0("c",res$condition)
res <- res |> filter(!grepl("^DECOY_",res$proteins))
hist(res$searchScore)

# res <- res |> filter(searchScore > 9)

resAgr <- res |> group_by(run, condition, peptide, proteins ) |>
  summarize(intensity = sum(intensity, na.rm = TRUE), searchScore = max(searchScore, na.rm = TRUE), n = n())

resAgr <- ungroup(resAgr)
resAgr$identQ <- max(resAgr$searchScore) - resAgr$searchScore
resAgr <- resAgr |> filter(!grepl("^DECOY_",resAgr$proteins))

atable <- prolfqua::AnalysisTableAnnotation$new()
atable$hierarchy[["protein_Id"]] <- "proteins"
atable$hierarchy[["peptide_Id"]] <- "peptide"
atable$hierarchy
atable$hierarchy_keys()
atable$factors[["condition"]] <- "condition"
atable$ident_Score <- "searchScore"
atable$ident_qValue <- "identQ"
atable$set_response("intensity")
atable$fileName <- "run"
config <- prolfqua::AnalysisConfiguration$new(atable)

undebug(prolfqua::setup_analysis)
xx <- prolfqua::setup_analysis(resAgr, config)

library(prolfqua)
psmdata <- LFQData$new(xx, config)
psmdata$to_wide()$data |> dim()

psmdata$config$table$hierarchyDepth <- 1
psmdata$config$table$hierarchy_keys_depth()
psmdata$hierarchy_counts()
psmdata$remove_small_intensities()

tr <- psmdata$get_Transformer()$intensity_array(log)
ag <- tr$lfq$get_Aggregator()
ag$medpolish()
protd <- ag$lfq_agg
tr <- protd$get_Transformer()
tr$intensity_array(exp, force = TRUE)
protd <- tr$lfq
protd$is_transformed(FALSE)

stprot <- protd$get_Stats()
stprot$violin_median()

pl <- protd$get_Plotter()
hm <- pl$heatmap()

```

```{r}
print(hm)

```

```{r}

pl$intensity_distribution_violin()
tr <- protd$get_Transformer()
tr$intensity_matrix(vsn::justvsn)
lfqdata <- tr$lfq
lfqdata$get_Plotter()$intensity_distribution_violin()
hm <- lfqdata$get_Plotter()$heatmap()
```

```{r}
print(hm)
```

```{r}
st <- lfqdata$get_Stats()
st$violin()

```

## Prolfqua contrasts

```{r}
lfqdata$rename_response("intensity")
contr <- c("2vs1" = "conditionc2 - conditionc1")
lmstr <- strategy_lm("intensity ~ condition")
models <- build_model(lfqdata, lmstr)
sum(models$modelDF$isSingular,na.rm = TRUE)
sum(is.na(models$modelDF$isSingular))

contrlm <- Contrasts$new(models, contr)
contrmiss <- ContrastsMissing$new(lfqdata, contr) 
contrlmmod <- contrlm |> ContrastsModerated$new()
contrmissmod <- contrmiss |> ContrastsModerated$new()

contrmergmod <- prolfqua::merge_contrasts_results(contrlmmod, contrmissmod)$merged
ctrmergmod <- contrmergmod$get_contrasts()

```

```{r}

pl <- contrmergmod$get_Plotter()
pl$contrastDF$color <- ifelse(grepl("HUMAN",ctrmergmod$protein_Id), "black", "red")
pl$volcano(colour = "color")

```

```{r}

tripleprot_bench_preprocess <- function(data, idcol = "protein_Id", drop = "YEAST") {

  truepos = setdiff(c("YEAST", "ECOLI"), drop)
  print(truepos)

  tmp <- data |>
    ungroup() |>
    mutate(species  = case_when(
      grepl("HUMAN", !!sym(idcol)) ~ "HUMAN",
      grepl("ECOLI", !!sym(idcol)) ~ "ECOLI",
      grepl("YEAST", !!sym(idcol)) ~ "YEAST",
      TRUE ~ "OTHER"
    ))
  res <- tmp |> dplyr::filter(!.data$species == "OTHER"  & !.data$species == drop)
  res <- res |> mutate(TP = (.data$species == truepos))
  return(list(data = res , table = table(tmp$species)))
}

```

```{r moderated}
ttd <- tripleprot_bench_preprocess(ctrmerg,drop = "YEAST" )
dim(ttd$data)
benchmark_mod_prolfqua <- prolfqua::make_benchmark(
  ttd$data,
  model_description = "prolfqua_merged",
  model_name = "prolfqua_merged",
  FDRvsFDP = list(list(score = "FDR", desc = FALSE))
)
benchmark_mod_prolfqua$pAUC_summaries()$ftable$content
knitr::kable(benchmark_mod_prolfqua$pAUC_summaries()$ftable$content)
prolfqua::table_facade(benchmark_mod_prolfqua$smc$summary, "Nr of estimated contrasts")
benchmark_mod_prolfqua$plot_score_distribution()
benchmark_mod_prolfqua$plot_ROC(0.05)
benchmark_mod_prolfqua$plot_FDRvsFDP()

```

```{r}
ttd <- tripleprot_bench_preprocess(ctrmergmod,drop = "ECOLI" )
dim(ttd$data)
benchmark_mod_prolfqua <- prolfqua::make_benchmark(
  ttd$data,
  model_description = "prolfqua_merged",
  model_name = "prolfqua_merged",
  FDRvsFDP = list(list(score = "FDR", desc = FALSE)),
  benchmark = list(
    list(score = "diff", desc = FALSE),
    list(score = "statistic", desc = FALSE),
    list(score = "scaled.p.value", desc = FALSE)
  )
)

benchmark_mod_prolfqua$pAUC_summaries()$ftable$content
prolfqua::table_facade(benchmark_mod_prolfqua$smc$summary, "Nr of estimated contrasts")
benchmark_mod_prolfqua$plot_score_distribution()
benchmark_mod_prolfqua$plot_ROC(0.05)

benchmark_mod_prolfqua$plot_FDRvsFDP()
benchmark_mod_prolfqua$.get_confusion
prolfqua:::ms_bench_add_scores

dx <- benchmark_mod_prolfqua$get_confusion_FDRvsFDP()
head(dx)
plot(dx$scorecol, dx$FDP_)

```

```{r}
resTriqM <- resTriq |> filter(DECOY == FALSE)
nrow(resTriqM)

resTriqM <- resTriqM |> select(protein,log2_fold_change,q_value, color)
ctr <- ungroup(ctrmergmod) |> select(protein_Id,diff,FDR)

xx <- inner_join(resTriqM, ctr, by = c(protein = "protein_Id"))

plot( xx$log2_fold_change, -xx$diff, col = xx$color)
abline(0, 1, col = "red", lwd = 2)
abline(h = c(-1,1), v = c(-1,1))

plot(-log10(xx$FDR), -log10(xx$q_value), log="xy", col = xx$color )
abline(h = -log10(0.05), v = -log10(0.05))
```

```{r}

```

### 2024-06-28 inst/MyArticle/paper.Rmd
create benchpath with date

@@
 ## read MQ peptide.txt and annotation table
-startdata <- prolfquapp::tidyMQ_Peptides(system.file(
-  "samples/maxquant_txt/tiny2.zip", package = "prolfqua"))
-annot <- readxl::read_xlsx(system.file(
-  "samples/maxquant_txt/annotation_Ionstar2018_PXD003881.xlsx", package = "prolfqua"))
-startdata <- dplyr::inner_join(annot, startdata, by = "raw.file")
-
-## create MaxQuant configuration
-config <- prolfqua::create_config_MQ_peptide()
-
-## specify explanatory variable
-config$table$factors[['dilution.']] = "sample"
+startdata <- prolfqua::sim_lfq_data_peptide_config(Nprot = 100)
 
 ## create R6 object
-lfqpep <- prolfqua::LFQData$new(startdata, config, setup = TRUE)
+lfqpep <- prolfqua::LFQData$new(startdata$data, startdata$config, setup = FALSE)
 
@@
 contrasts <- c(
-  "e_vs_d" =   "dilution.e - dilution.d",
-  "d_vs_c" =   "dilution.d - dilution.c"
+  "a_vs_c" =   "group_A - group_Ctrl",
+  "b_vs_c" =   "group_B - group_Ctrl"
 )
 ## fit model
-lmmodel <- paste(lfqpro$response()," ~ dilution.")
+lmmodel <- paste(lfqpro$response()," ~ group_")
 modelFunction <- prolfqua::strategy_lm( lmmodel, model_name = "lm")

### 2025-01-08 R/utilities.R
getting cptac vignette working

@@ -12,13 +12,13 @@ tolong <- function(assay, rowData, colData ){
 #' @export
 cptac_bench_preprocess <- function(data, idcol = "protein_Id") {
     tmp <- data |>
-        ungroup() |>
-        mutate(species  = case_when(
+        dplyr::ungroup() |>
+        dplyr::mutate(species  = dplyr::case_when(
             grepl("YEAST", !!sym(idcol)) ~ "YEAST",
             grepl("UPS", !!sym(idcol)) ~ "UPS",
             TRUE ~ "OTHER"
         ))
     res <- tmp |> dplyr::filter(!.data$species == "OTHER")
-    res <- res |> mutate(TP = (.data$species == "UPS"))
+    res <- res |> dplyr::mutate(TP = (.data$species == "UPS"))
     return(list(data = res , table = table(tmp$species)))
 }
